---
title: "Vietnam Draft Lottery — Replication and Outcome-Weight Diagnostics"
subtitle: "Kappa estimators, Double ML, learner sensitivity, and exact weight checks"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body   { font-size: 15px; color: #2c2c2c; }
h1     { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2     { color: #2c7bb6; margin-top: 28px; }
h3     { color: #444; margin-top: 20px; }

.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
.result-box {
  background: #f3f8f3;
  border-left: 4px solid #3c8d40;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
.warn-box {
  background: #fff7e6;
  border-left: 4px solid #d89000;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  cache   = TRUE,
  autodep = TRUE,
  fig.width  = 12,
  fig.height = 6
)
options(knitr.kable.NA = "—")
```

---

# Setup

## Packages & Functions

```{r packages-and-source}
# Core
library(haven)
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
# DML smoother + balance plots
library(OutcomeWeights)
library(cobalt)
library(viridis)
library(gridExtra)
library(ggplot2)
# DoubleML learners
library(DoubleML)
library(mlr3verse)
lgr::get_logger("mlr3")$set_threshold("warn")

source(file.path("..", "functions_all.R"))   # estimators, analytical SEs, and weight diagnostics
```

## Data & Design

<div class="note-box">
**Setting.** Angrist (1990) SIPP 1984 subsample (N = 3,027 white men, born 1950–53). Instrument Z = `rsncode` (draft-eligibility indicator). Treatment D = `nvstat` (Vietnam-era veteran status). Outcome: log wages in two units (`lwage = log(kwage)` dollars; `lwage_cnt = log(100·kwage)` cents) — the units only enter the SUW reproduction table. Variable definitions follow the SUW Stata replication exactly: `age = age_5`, with three covariate specifications — **Linear** (`age`), **Cubic** (`age, age², age³`), and **Saturated** (`i.age` = age dummies).
</div>

```{r data}
sipp_path <- "sipp.dta"
stopifnot(file.exists(sipp_path))
sipp <- read_dta(sipp_path)

# Sample restriction: N = 3,027  (drop rsncode==999, missing educ / kwage)
sipp_clean <- sipp[sipp$rsncode != 999 &
                     !is.na(sipp$educ) &
                     !is.na(sipp$kwage), ]

# Age terms — EXACTLY as in suw_jbes_replicate.do: age = age_5
sipp_clean$age     <- sipp_clean$age_5
sipp_clean$age2    <- sipp_clean$age_5^2
sipp_clean$age3    <- sipp_clean$age_5^3
sipp_clean$age_fac <- factor(sipp_clean$age_5)

# Outcomes in two units (note: dollars vs. cents convention is FLIPPED vs. Card)
sipp_clean$lwage     <- log(sipp_clean$kwage)         # dollars
sipp_clean$lwage_cnt <- log(100 * sipp_clean$kwage)   # cents

# Core vectors
Y_cnt <- sipp_clean$lwage_cnt   # cents  (used only for the SUW reproduction layout)
Y_dol <- sipp_clean$lwage       # dollars (the working outcome everywhere else)
Z     <- as.integer(sipp_clean$rsncode)
D     <- as.integer(sipp_clean$nvstat)

stopifnot(nrow(sipp_clean) == 3027L)
```

```{r design-matrices}
# Kappa design matrices — WITH intercept (for logit / CBPS)
X_lin <- model.matrix(~ age,                data = sipp_clean)
X_cub <- model.matrix(~ age + age2 + age3,  data = sipp_clean)
X_sat <- model.matrix(~ age_fac,            data = sipp_clean)

# X_df for 2SLS (no intercept column)
X_lin_df <- data.frame(age = sipp_clean$age)
X_cub_df <- data.frame(age = sipp_clean$age, age2 = sipp_clean$age2, age3 = sipp_clean$age3)
X_sat_df <- data.frame(age_fac = sipp_clean$age_fac)

# DML design matrices — no intercept, attributes stripped for grf (make_X)
X_dml_cub      <- make_X(~ 0 + age + I(age^2) + I(age^3), data = sipp_clean)
X_dml_sat_full <- make_X(~ 0 + age_fac, data = sipp_clean)
X_dml_sat      <- X_dml_sat_full[, -1, drop = FALSE]   # drop one age dummy (reference)

stopifnot(nrow(X_dml_cub) == nrow(sipp_clean),
          nrow(X_dml_sat) == nrow(sipp_clean))
```

---

# 0. Empirical Design Audit

<div class="note-box">
The lottery design is conditionally random within birth cohort. The pooled
eligibility share is therefore a sample-composition statistic, not the basis
for the assignment claim. The relevant empirical checks are whether draft
eligibility shifts veteran status after age adjustment and whether fitted
eligibility probabilities remain away from the boundaries.
</div>

```{r design-audit, echo=FALSE}
first_stage_row <- function(label, formula, X) {
  fit <- lm(formula, data = data.frame(D = D, Z = Z,
                                       age = sipp_clean$age,
                                       age2 = sipp_clean$age2,
                                       age3 = sipp_clean$age3,
                                       age_fac = sipp_clean$age_fac))
  ct <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC1"))
  p <- logit_mle(Z, X)
  data.frame(
    Specification = label,
    `Adjusted first stage` = unname(ct["Z", "Estimate"]),
    `HC1 SE` = unname(ct["Z", "Std. Error"]),
    `Robust t²` = unname(ct["Z", "t value"]^2),
    `Min p̂(X)` = min(p),
    `Max p̂(X)` = max(p),
    `Boundary share` = mean(p < 0.05 | p > 0.95),
    check.names = FALSE
  )
}

design_audit <- rbind(
  first_stage_row("Linear age", D ~ Z + age, X_lin),
  first_stage_row("Cubic age", D ~ Z + age + age2 + age3, X_cub),
  first_stage_row("Saturated age", D ~ Z + age_fac, X_sat)
)

sample_audit <- data.frame(
  N = length(Z),
  `Pr(Z=1)` = mean(Z),
  `Pr(D=1)` = mean(D),
  `Pr(D=1 | Z=0)` = mean(D[Z == 0]),
  `Pr(D=1 | Z=1)` = mean(D[Z == 1]),
  `Raw uptake gap` = mean(D[Z == 1]) - mean(D[Z == 0]),
  check.names = FALSE
)

sample_audit %>%
  kbl(digits = 4, caption = "Sample and unconditional treatment uptake",
      align = rep("r", ncol(sample_audit))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13)

design_audit %>%
  kbl(digits = 4,
      caption = "Age-adjusted first-stage strength and estimated instrument overlap",
      align = c("l", rep("r", 6))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) %>%
  footnote(general = "Robust t² is the squared HC1 t-statistic for draft eligibility; it is a descriptive strength diagnostic, not a formal weak-IV-robust test. Boundary share uses 0.05 and 0.95.")
```

<div class="warn-box">
**Qualification.** Sample first-stage differences are not positive in every
age cell (ages 29 and 39 are negative). These noisy cell contrasts neither
establish nor refute individual-level monotonicity. The cubic raw-power basis
is also ill-conditioned; its estimates are retained because they are stable,
but the saturated specification is numerically cleaner.
</div>

---

# 1. Reproduction — SUW Table 2

<div class="note-box">
**Layout.** One treatment (veteran status), three age specifications, and two
outcome units (cents / dollars). Columns are specification × unit. The
normalized estimators (τ_cb_u, τ_ml_u, τ_ml_a10) are translation invariant by
construction, so their cents and dollars results match. The unnormalized
variants (τ_ml_a, τ_ml_t, τ_ml_a0) are not generally translation invariant:
they change in the linear and cubic specifications but coincide empirically in
the saturated-age specification.

**Notation.** `cb` and `ml` denote CBPS and logit-MLE instrument propensities;
`u` is Uysal normalization and `a10` separately normalizes the treated and
control kappa components. The `a`, `t`, and `a0` rows are the corresponding
SUW unnormalized denominator variants.
</div>

```{r build-columns, cache=TRUE}
# One full column = 2SLS + all six kappa estimators with analytical SEs.
run_viet_column <- function(Y, X_kappa, X_df) {
  list(
    tsls  = run_2sls(Y, D, Z, X_df, endog_name = "nvstat"),
    kappa = kappa_analytic_se_all(Y, Z, D, X_kappa)
  )
}

col_lin_cnt <- run_viet_column(Y_cnt, X_lin, X_lin_df)
col_lin_dol <- run_viet_column(Y_dol, X_lin, X_lin_df)
col_cub_cnt <- run_viet_column(Y_cnt, X_cub, X_cub_df)
col_cub_dol <- run_viet_column(Y_dol, X_cub, X_cub_df)
col_sat_cnt <- run_viet_column(Y_cnt, X_sat, X_sat_df)
col_sat_dol <- run_viet_column(Y_dol, X_sat, X_sat_df)
```

```{r table-builder, echo=FALSE}
cell_fmt <- function(est, se, digits = 3) {
  if (is.na(est)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0)
    return(sprintf("%.3f<br><span style='color:#888'>(NA)</span>", est))
  sprintf("%.3f<br><span style='color:#888'>(%.3f)</span>", est, se)
}

build_spec_table <- function(c_lin_cnt, c_lin_dol, c_cub_cnt, c_cub_dol, c_sat_cnt, c_sat_dol) {
  ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
  labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")

  rows <- list(data.frame(
    Estimator = "2SLS",
    `Linear / cents`      = cell_fmt(c_lin_cnt$tsls$coef, c_lin_cnt$tsls$se),
    `Linear / dollars`    = cell_fmt(c_lin_dol$tsls$coef, c_lin_dol$tsls$se),
    `Cubic / cents`       = cell_fmt(c_cub_cnt$tsls$coef, c_cub_cnt$tsls$se),
    `Cubic / dollars`     = cell_fmt(c_cub_dol$tsls$coef, c_cub_dol$tsls$se),
    `Saturated / cents`   = cell_fmt(c_sat_cnt$tsls$coef, c_sat_cnt$tsls$se),
    `Saturated / dollars` = cell_fmt(c_sat_dol$tsls$coef, c_sat_dol$tsls$se),
    check.names = FALSE, stringsAsFactors = FALSE
  ))

  for (i in seq_along(ests)) {
    e <- ests[i]
    rows[[length(rows) + 1]] <- data.frame(
      Estimator = labels[i],
      `Linear / cents`      = cell_fmt(c_lin_cnt$kappa$estimates[e], c_lin_cnt$kappa$se[e]),
      `Linear / dollars`    = cell_fmt(c_lin_dol$kappa$estimates[e], c_lin_dol$kappa$se[e]),
      `Cubic / cents`       = cell_fmt(c_cub_cnt$kappa$estimates[e], c_cub_cnt$kappa$se[e]),
      `Cubic / dollars`     = cell_fmt(c_cub_dol$kappa$estimates[e], c_cub_dol$kappa$se[e]),
      `Saturated / cents`   = cell_fmt(c_sat_cnt$kappa$estimates[e], c_sat_cnt$kappa$se[e]),
      `Saturated / dollars` = cell_fmt(c_sat_dol$kappa$estimates[e], c_sat_dol$kappa$se[e]),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }
  do.call(rbind, rows)
}
```

## Results

```{r table-reproduction}
build_spec_table(col_lin_cnt, col_lin_dol, col_cub_cnt, col_cub_dol, col_sat_cnt, col_sat_dol) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c","c","c"),
      caption = "SUW Table 2 — Treatment: Vietnam veteran status (nvstat). Effect of military service on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12.5) %>%
  add_header_above(c(" " = 1, "Linear age" = 2, "Cubic age" = 2, "Saturated age" = 2)) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",            1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell reports the estimate with its analytical or HC1-robust standard error below. Blue = normalized (translation invariant); orange = unnormalized. age = age_5, following the SUW replication.")
```

---

# 2. Double ML — PLR-IV & Wald-AIPW (honest forest, tuned vs. untuned)

<div class="note-box">
**Setting.** DML uses `dml_with_smoother()` with five-fold cross-fitting on
the **cubic-age** and **saturated-age** specifications. For each specification,
the same call jointly returns PLR-IV and Wald-AIPW. We estimate an untuned
honest forest (`tune.parameters = "none"`) and a tuned honest forest
(`tune.parameters = "all"`). The latter invokes GRF's automatic,
out-of-bag-loss-based tuning; it is distinct from the learner comparison in
Section 3. The linear-age specification remains in the kappa/2SLS reproduction
but is too low-dimensional to provide a meaningful honest-forest comparison.
`set.seed(42)` is reset before every fit so that the four calls begin from the
same random-number state, although tuning itself consumes additional random
draws.
</div>

```{r dml-fit-all, cache=TRUE, results='hide'}
# Cubic-age specification.
set.seed(42); dml_cub_untuned <- dml_with_smoother(
  Y_dol, D, X_dml_cub, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_cub_tuned <- dml_with_smoother(
  Y_dol, D, X_dml_cub, Z, n_cf_folds = 5, tune.parameters = "all")

# Saturated-age specification.
set.seed(42); dml_sat_untuned <- dml_with_smoother(
  Y_dol, D, X_dml_sat, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_sat_tuned <- dml_with_smoother(
  Y_dol, D, X_dml_sat, Z, n_cf_folds = 5, tune.parameters = "all")

res_cub_untuned <- summary(dml_cub_untuned)
res_cub_tuned   <- summary(dml_cub_tuned)
res_sat_untuned <- summary(dml_sat_untuned)
res_sat_tuned   <- summary(dml_sat_tuned)

extract_w <- function(dml_obj) {
  om <- get_outcome_weights(dml_obj)
  check_omega_rows(om)
  list(plriv    = as.vector(om$omega["PLR-IV",    ]),
       waldaipw = as.vector(om$omega["Wald-AIPW", ]))
}
w_cub_untuned <- extract_w(dml_cub_untuned)
w_cub_tuned   <- extract_w(dml_cub_tuned)
w_sat_untuned <- extract_w(dml_sat_untuned)
w_sat_tuned   <- extract_w(dml_sat_tuned)
```

## Estimates {.tabset}

```{r dml-cell-display, echo=FALSE}
show_dml_cell <- function(res, wlist, Y, cap) {
  ids <- c(
    check_weight_identity(wlist$plriv,    Y, get_estimate(res, "PLR-IV")),
    check_weight_identity(wlist$waldaipw, Y, get_estimate(res, "Wald-AIPW"))
  )
  df <- as.data.frame(res[c("PLR-IV", "Wald-AIPW"), , drop = FALSE])
  df$Estimator  <- rownames(df)
  df$`ω′Y reproduces τ̂` <- as.character(ids)
  df <- df[, c("Estimator", setdiff(names(df), c("Estimator","ω′Y reproduces τ̂")),
               "ω′Y reproduces τ̂")]
  df %>%
    kbl(caption = cap, digits = 4, row.names = FALSE,
        align = c("l", rep("r", ncol(df) - 2), "c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "Estimator-specific numerical reproduction uses the all.equal() criterion employed by Knaus, with R's default tolerance.")
}
```

### Cubic age

```{r dml-cub-untuned}
show_dml_cell(res_cub_untuned, w_cub_untuned, Y_dol,
              "Cubic age — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-cub-tuned}
show_dml_cell(res_cub_tuned, w_cub_tuned, Y_dol,
              "Cubic age — tuned honest forest (tune.parameters = \"all\")")
```

### Saturated age

```{r dml-sat-untuned}
show_dml_cell(res_sat_untuned, w_sat_untuned, Y_dol,
              "Saturated age — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-sat-tuned}
show_dml_cell(res_sat_tuned, w_sat_tuned, Y_dol,
              "Saturated age — tuned honest forest (tune.parameters = \"all\")")
```

<div class="note-box">
**How to read the comparison.** Within each age specification, compare the
untuned and tuned rows for changes in the point estimate, standard error, and
later weight geometry. The `ω′Y reproduces τ̂` column checks the package's
algebraic outcome-weight representation; it is not a tuning-quality or
normalization criterion.
</div>

## DML vs. Kappa — Point Estimates

```{r dml-summary-table}
# Kappa estimates (dollars) for both specs — reused later by the learner table.
kappa_cub <- kappa_analytic_se_all(Y_dol, Z, D, X_cub)
kappa_sat <- kappa_analytic_se_all(Y_dol, Z, D, X_sat)

cell <- function(coef, se) {
  if (is.na(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  sprintf("%.3f (%.3f)", coef, se)
}

dml_cell <- function(res, est) {
  coef_val <- get_estimate(res, est)
  se_val   <- if ("SE" %in% colnames(res)) as.numeric(res[est, "SE"])
              else if ("Std. Error" %in% colnames(res)) as.numeric(res[est, "Std. Error"])
              else NA_real_
  cell(coef_val, se_val)
}

dml_rows <- rbind(
  data.frame(Estimator = "PLR-IV (untuned forest)",
             `Cubic age`     = dml_cell(res_cub_untuned, "PLR-IV"),
             `Saturated age` = dml_cell(res_sat_untuned, "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (untuned forest)",
             `Cubic age`     = dml_cell(res_cub_untuned, "Wald-AIPW"),
             `Saturated age` = dml_cell(res_sat_untuned, "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "PLR-IV (tuned forest)",
             `Cubic age`     = dml_cell(res_cub_tuned, "PLR-IV"),
             `Saturated age` = dml_cell(res_sat_tuned, "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (tuned forest)",
             `Cubic age`     = dml_cell(res_cub_tuned, "Wald-AIPW"),
             `Saturated age` = dml_cell(res_sat_tuned, "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE)
)

ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")
kappa_rows <- do.call(rbind, lapply(seq_along(ests), function(i) {
  e <- ests[i]
  data.frame(
    Estimator = labels[i],
    `Cubic age`     = cell(kappa_cub$estimates[e], kappa_cub$se[e]),
    `Saturated age` = cell(kappa_sat$estimates[e], kappa_sat$se[e]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}))

rbind(dml_rows, kappa_rows) %>%
  kbl(align = c("l","c","c"),
      caption = "DML vs. Kappa — Point Estimates (SE), Y = log wages (dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Double ML — honest forest", 1, 4) %>%
  pack_rows("Kappa — normalized", 5, 7) %>%
  pack_rows("Kappa — unnormalized", 8, 10) %>%
  row_spec(1:4, background = "#eef6ff") %>%
  row_spec(8:10, background = "#fffbf0") %>%
  footnote(general = "Each cell reports estimate (SE). Blue = DML; orange = unnormalized kappa.")
```

---

# 3. Learner Comparison — PLR-IV and Wald-AIPW

<div class="note-box">
**Purpose.** Section 2 varies GRF's internal tuning while holding the honest-
forest learner family fixed. This section instead holds the five-fold
DoubleML procedure fixed and varies the nuisance learners: parametric
linear/logistic models, Ranger, and XGBoost. Each learner family is used for
both estimators. PLR-IV is implemented with `DoubleMLPLIV`; Wald-AIPW/LATE is
implemented with `DoubleMLIIVM`. All models use the same cubic-age and
saturated-age designs defined above. Linear and Ranger are fitted in this
notebook. XGBoost is imported from `vietnam_xgb_tuning_export.rds` and
`vietnam_xgb_pliv_tuning_export.rds`, which preserve both the fixed untuned
baseline and the `tune_on_folds = TRUE` estimator without rerunning expensive
nested tuning here.

**Terminology.** “Linear learner” refers to the nuisance-model class: OLS for
PLR-IV and OLS/logit for Wald-AIPW. It does **not** refer to the linear-age
feature specification in Section 1. The learner comparison uses either the
cubic-age feature matrix or the saturated-age feature matrix throughout.
</div>

```{r ext-setup}
make_dml_data <- function(X_dml) {
  df <- as.data.frame(X_dml)
  df$y <- as.numeric(Y_dol)
  df$d <- as.numeric(D)
  df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X_dml))
}

dat_cub <- make_dml_data(X_dml_cub)
dat_sat <- make_dml_data(X_dml_sat)
```

```{r ext-fit-helpers, echo=FALSE}
ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))
ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))
omega_vec <- function(om, estimator) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && estimator %in% rownames(w))
    return(as.vector(w[estimator, ]))
  as.vector(w[1, ])
}

# Fit one IIVM model. On failure, return NA rather than aborting the knit.
fit_iivm <- function(dml_data, learners, Y) {
  tryCatch({
    set.seed(42)
    obj <- DoubleMLIIVM$new(data = dml_data,
                            ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                            n_folds = 5, score = "LATE")
    invisible(capture.output(
      obj$fit(store_models = TRUE, store_predictions = TRUE)
    ))
    om  <- get_outcome_weights(object = obj, dml_data = dml_data)
    w   <- omega_vec(om, "Wald-AIPW")
    ok  <- check_doubleml_identity(om, Y, obj$coef)
    list(obj = obj, w = w, ok = ok)
  }, error = function(e) {
    message("fit_iivm failed: ", conditionMessage(e))
    list(obj = NULL, w = rep(NA_real_, length(Y)), ok = NA)
  })
}

# PLR-IV uses three regression nuisances: E[Y|X], E[Z|X], and E[D|X].
fit_pliv <- function(dml_data, learners, Y) {
  tryCatch({
    set.seed(42)
    obj <- DoubleMLPLIV$new(data = dml_data,
                            ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
                            n_folds = 5)
    invisible(capture.output(
      obj$fit(store_models = TRUE, store_predictions = TRUE)
    ))
    om <- get_outcome_weights(object = obj, dml_data = dml_data)
    w  <- omega_vec(om, "PLR-IV")
    ok <- check_weight_identity(w, Y, get_dml_coef(obj))
    list(obj = obj, w = w, ok = ok)
  }, error = function(e) {
    message("fit_pliv failed: ", conditionMessage(e))
    list(obj = NULL, w = rep(NA_real_, length(Y)), ok = NA)
  })
}

safe_coef <- function(f) {
  if (is.null(f$obj)) return(NA_real_)
  out <- NA_real_
  invisible(capture.output(out <- get_dml_coef(f$obj)))
  as.numeric(out)
}
safe_se <- function(f) {
  if (is.null(f$obj)) return(NA_real_)
  out <- NA_real_
  invisible(capture.output(out <- get_dml_se(f$obj)))
  as.numeric(out)
}

# Compact estimator-specific reproduction table for one learner family.
learner_pair_table <- function(pliv_cub, pliv_sat, iivm_cub, iivm_sat, cap) {
  data.frame(
    Estimator = c("PLR-IV", "PLR-IV",
                  "Wald-AIPW / LATE", "Wald-AIPW / LATE"),
    Specification = c("Cubic age", "Saturated age",
                      "Cubic age", "Saturated age"),
    Estimate = round(c(safe_coef(pliv_cub), safe_coef(pliv_sat),
                       safe_coef(iivm_cub), safe_coef(iivm_sat)), 4),
    SE = round(c(safe_se(pliv_cub), safe_se(pliv_sat),
                 safe_se(iivm_cub), safe_se(iivm_sat)), 4),
    `ω′Y reproduces τ̂` = as.character(c(
      pliv_cub$ok, pliv_sat$ok, iivm_cub$ok, iivm_sat$ok
    )),
    check.names = FALSE
  ) %>%
    kbl(caption = cap, align = c("l","l","r","r","c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "PLR-IV uses DoubleMLPLIV; Wald-AIPW/LATE uses DoubleMLIIVM. Each fitted estimator is checked separately using the all.equal() criterion employed by Knaus, with R's default tolerance.")
}
```

## Estimates {.tabset}

### OLS / Logistic nuisance learners

```{r ext-fit-lm, cache=TRUE, cache.extra='quiet-output-v2'}
lm_iivm_cub <- fit_iivm(dat_cub, ml_lm_iivm(), Y_dol)
lm_iivm_sat <- fit_iivm(dat_sat, ml_lm_iivm(), Y_dol)
lm_pliv_cub <- fit_pliv(dat_cub, ml_lm_pliv(), Y_dol)
lm_pliv_sat <- fit_pliv(dat_sat, ml_lm_pliv(), Y_dol)

learner_pair_table(lm_pliv_cub, lm_pliv_sat, lm_iivm_cub, lm_iivm_sat,
                   "Parametric learners — Linear PLR-IV and Linear/Logistic Wald-AIPW")
```

<div class="note-box">
**Cubic OLS PLR-IV.** The `FALSE` is a borderline numerical-conditioning
artifact, not evidence against the PLR-IV representation. The reported estimate
is 0.2387739352 versus \(\omega'Y=0.2387739389\): the relative difference
\(1.553\times10^{-8}\) lies just above R's default tolerance
\(1.490\times10^{-8}\). Raw cubic cross-products are ill-conditioned
(\(7.16\times10^{15}\) to \(7.63\times10^{15}\)); a QR-based reconstruction,
matching `lm`, reduces the gap to \(3.34\times10^{-13}\) and returns `TRUE`,
consistent with Knaus (2024, Table 5, C1a) for linear smoothers.
</div>

### Ranger

```{r ext-fit-rf, cache=TRUE, cache.extra='quiet-output-v2'}
rf_iivm_cub <- fit_iivm(dat_cub, ml_rf_iivm(), Y_dol)
rf_iivm_sat <- fit_iivm(dat_sat, ml_rf_iivm(), Y_dol)
rf_pliv_cub <- fit_pliv(dat_cub, ml_rf_pliv(), Y_dol)
rf_pliv_sat <- fit_pliv(dat_sat, ml_rf_pliv(), Y_dol)

learner_pair_table(rf_pliv_cub, rf_pliv_sat, rf_iivm_cub, rf_iivm_sat,
                   "Ranger learners — PLR-IV and Wald-AIPW/LATE")
```

### Imported XGBoost fits

```{r ext-xgb-wald-import, include=FALSE}
validate_xgb_export <- function(x,expected_class) {
  required <- c("estimate_comparison","nuisance_losses","weight_diagnostics",
                "omega_vectors","identity_checks","settings","sample_info")
  if (!all(required %in% names(x))) stop("XGBoost export has an incomplete schema.")
  if (!identical(x$settings$dml_class,expected_class)) stop("XGBoost export has the wrong DoubleML class.")
  if (!isTRUE(x$settings$tune_on_folds)) stop("XGBoost export is not tune_on_folds=TRUE.")
  if (!all(c("Cubic age","Saturated age") %in% names(x$omega_vectors)))
    stop("XGBoost export does not contain both Vietnam age specifications.")
  if (!identical(as.integer(x$sample_info$n), as.integer(length(Z))) ||
      !isTRUE(all.equal(as.numeric(x$sample_info$mean_Z), mean(Z), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(x$sample_info$mean_D), mean(D), tolerance = 1e-12)))
    stop("XGBoost export sample metadata do not match the current analysis data.")
  invisible(TRUE)
}
xgb_wald_path <- "vietnam_xgb_tuning_export.rds"
xgb_wald_available <- file.exists(xgb_wald_path)
if (!xgb_wald_available) stop("Missing vietnam_xgb_tuning_export.rds.")
xgb_wald_export <- readRDS(xgb_wald_path)
validate_xgb_export(xgb_wald_export,"DoubleMLIIVM")
```

```{r ext-xgb-pliv-import, include=FALSE}
xgb_pliv_path <- "vietnam_xgb_pliv_tuning_export.rds"
xgb_pliv_available <- file.exists(xgb_pliv_path)
if (!xgb_pliv_available) stop("Missing vietnam_xgb_pliv_tuning_export.rds.")
xgb_pliv_export <- readRDS(xgb_pliv_path)
validate_xgb_export(xgb_pliv_export,"DoubleMLPLIV")
```

```{r ext-xgb-export-helpers, echo=FALSE}
empty_xgb_bundle <- function() list(w=NULL,ok=NA,estimate=NA_real_,gap=NA_real_)
get_xgb_estimate_value <- function(export,cell,variant) {
  if (is.null(export)) return(NA_real_)
  df <- export$weight_diagnostics
  keep_variant <- if (variant=="untuned") {
    grepl("untuned$",df$Learner)
  } else {
    grepl(" tuned$",df$Learner)
  }
  hit <- df[df$Cell==cell & keep_variant,,drop=FALSE]
  if (!nrow(hit)) NA_real_ else as.numeric(hit$Estimate[1])
}
get_xgb_bundle <- function(export,cell,variant) {
  if (is.null(export)) return(empty_xgb_bundle())
  w <- tryCatch(export$omega_vectors[[cell]][[variant]],error=function(e)NULL)
  estimate <- get_xgb_estimate_value(export,cell,variant)
  gap <- if (is.null(w) || !is.finite(estimate)) NA_real_ else sum(w*Y_dol)-estimate
  ok <- if (is.finite(gap)) check_weight_identity(w,Y_dol,estimate) else NA
  list(w=w,ok=ok,estimate=estimate,gap=gap)
}
get_xgb_estimate_cell <- function(export,cell,variant) {
  if (is.null(export)) return("—")
  df <- export$estimate_comparison
  keep_variant <- if (variant=="untuned") grepl("untuned$",df$Learner) else grepl("tune_on_folds=TRUE",df$Learner,fixed=TRUE)
  hit <- df[df$Cell==cell & keep_variant,,drop=FALSE]
  if (!nrow(hit)) "—" else as.character(hit[["Estimate (SE)"]][1])
}
get_xgb_identity_cell <- function(export,cell,variant) {
  value <- tryCatch(export$identity_checks[[cell]][[variant]], error=function(e) NA)
  if (isTRUE(value)) "TRUE" else if (identical(value,FALSE)) "FALSE" else "—"
}

xgb_wald_obj <- if (exists("xgb_wald_export")) xgb_wald_export else NULL
xgb_pliv_obj <- if (exists("xgb_pliv_export")) xgb_pliv_export else NULL

xgb_iivm_cub_untuned <- get_xgb_bundle(xgb_wald_obj,"Cubic age","untuned")
xgb_iivm_cub_tuned   <- get_xgb_bundle(xgb_wald_obj,"Cubic age","tuned")
xgb_iivm_sat_untuned <- get_xgb_bundle(xgb_wald_obj,"Saturated age","untuned")
xgb_iivm_sat_tuned   <- get_xgb_bundle(xgb_wald_obj,"Saturated age","tuned")
xgb_pliv_cub_untuned <- get_xgb_bundle(xgb_pliv_obj,"Cubic age","untuned")
xgb_pliv_cub_tuned   <- get_xgb_bundle(xgb_pliv_obj,"Cubic age","tuned")
xgb_pliv_sat_untuned <- get_xgb_bundle(xgb_pliv_obj,"Saturated age","untuned")
xgb_pliv_sat_tuned   <- get_xgb_bundle(xgb_pliv_obj,"Saturated age","tuned")
```

```{r ext-xgb-comparison, echo=FALSE}
xgb_import_table <- do.call(rbind, lapply(list(
  list("PLR-IV", "Cubic age",     xgb_pliv_obj),
  list("PLR-IV", "Saturated age", xgb_pliv_obj),
  list("Wald-AIPW / LATE", "Cubic age",     xgb_wald_obj),
  list("Wald-AIPW / LATE", "Saturated age", xgb_wald_obj)
), function(x) {
  data.frame(
    Estimator=x[[1]],
    Specification=x[[2]],
    Untuned=get_xgb_estimate_cell(x[[3]],x[[2]],"untuned"),
    Tuned=get_xgb_estimate_cell(x[[3]],x[[2]],"tuned"),
    `ω′Y reproduces τ̂ (untuned / tuned)`=paste(
      get_xgb_identity_cell(x[[3]],x[[2]],"untuned"),
      get_xgb_identity_cell(x[[3]],x[[2]],"tuned"),
      sep=" / "),
    check.names=FALSE
  )
}))

xgb_import_table %>%
  kbl(caption="Imported XGBoost estimates — Untuned and fold-specific tuning",
      align=c("l","l","c","c","c")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE,font_size=13) %>%
  footnote(general="Each estimate cell reports estimate (SE). Each fitted estimator is checked separately using the all.equal() criterion employed by Knaus, with R's default tolerance. Reproduction is distinct from normalization.")
```

<div class="note-box">
**Interpretation.** None of the reconstructed XGBoost vectors passes the
estimator-specific `all.equal()` reproduction check. The XGBoost point
estimates remain in the learner comparison, but these vectors are excluded
from the outcome-weight diagnostics and Love plots. This representation check
is distinct from the normalization conditions studied by Knaus.
</div>

## Comparison Table

```{r ext-comparison}
ext_cell <- function(coef, se) {
  if (is.na(coef) || !is.finite(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  sprintf("%.3f (%.3f)", coef, se)
}

learner_row <- function(label, f_cub, f_sat) {
  data.frame(
    Estimator = label,
    `Cubic age`     = ext_cell(safe_coef(f_cub), safe_se(f_cub)),
    `Saturated age` = ext_cell(safe_coef(f_sat), safe_se(f_sat)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

kappa_ref_row <- function(label, key) {
  data.frame(
    Estimator = label,
    `Cubic age`     = ext_cell(kappa_cub$estimates[key], kappa_cub$se[key]),
    `Saturated age` = ext_cell(kappa_sat$estimates[key], kappa_sat$se[key]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

comparison_rows <- list(
  learner_row("PLR-IV (OLS nuisances)",                  lm_pliv_cub,  lm_pliv_sat),
  learner_row("Wald-AIPW / LATE (OLS/logit nuisances)", lm_iivm_cub, lm_iivm_sat),
  learner_row("PLR-IV (ranger)",                  rf_pliv_cub,  rf_pliv_sat),
  learner_row("Wald-AIPW / LATE (ranger)",       rf_iivm_cub, rf_iivm_sat),
  kappa_ref_row("τ_cb_u (kappa, ref)",  "tau_cb_u"),
  kappa_ref_row("τ_ml_u (kappa, ref)",  "tau_ml_u")
)
xgb_rows <- list(
  data.frame(Estimator="PLR-IV (XGBoost, untuned)",
    `Cubic age`=get_xgb_estimate_cell(xgb_pliv_obj,"Cubic age","untuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_pliv_obj,"Saturated age","untuned"),check.names=FALSE),
  data.frame(Estimator="PLR-IV (XGBoost, tune_on_folds=TRUE)",
    `Cubic age`=get_xgb_estimate_cell(xgb_pliv_obj,"Cubic age","tuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_pliv_obj,"Saturated age","tuned"),check.names=FALSE),
  data.frame(Estimator="Wald-AIPW / LATE (XGBoost, untuned)",
    `Cubic age`=get_xgb_estimate_cell(xgb_wald_obj,"Cubic age","untuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_wald_obj,"Saturated age","untuned"),check.names=FALSE),
  data.frame(Estimator="Wald-AIPW / LATE (XGBoost, tune_on_folds=TRUE)",
    `Cubic age`=get_xgb_estimate_cell(xgb_wald_obj,"Cubic age","tuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_wald_obj,"Saturated age","tuned"),check.names=FALSE)
)
comp_df <- do.call(rbind,c(comparison_rows[1:4],xgb_rows,comparison_rows[5:6]))

comp_df %>%
  kbl(align = c("l","c","c"),
      caption = "PLR-IV and Wald-AIPW Learner Comparison — DoubleML vs. Kappa (Y = log wages, dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("DoubleML learners", 1, 8) %>%
  pack_rows("Kappa reference",   9, 10) %>%
  row_spec(1:8, background = "#eef6ff") %>%
  footnote(general = "Each cell reports estimate (SE). Linear denotes the nuisance learner, not the linear-age feature specification. XGBoost rows import the stored nested-tuning fits; their point estimates are retained although their reconstructed weights fail the reproduction gate.")
```

---

# 4. Weight Diagnostics

<div class="note-box">
**Representation gate.** The estimator-specific checks in Sections 2–3
determine which vectors enter the outcome-weight analysis. A vector is included
if it is algebraic by construction or if `ω′Y reproduces τ̂` is `TRUE`.
Accordingly, the XGBoost vectors and cubic-age OLS PLR-IV vector remain in the
estimate tables but are excluded below. Reproduction, normalization, balance,
and concentration are distinct properties.
</div>

```{r diag-builders, echo=FALSE}
# kappa weight bundles for both specs.
kw_cub <- kappa_weights_bundle(Z, D, X_cub)
kw_sat <- kappa_weights_bundle(Z, D, X_sat)

exact_w <- function(fit) {
  if (isTRUE(fit$ok) && !is.null(fit$w)) fit$w else NULL
}
```

## Detailed Outcome-Weight Statistics {.tabset}

<div class="note-box">
**Knaus summary.** The tables report Overall / Treated / Control statistics
using Knaus's sign convention. `Neg. share` is an observation-count share and
is distinct from the absolute-mass sign-cancellation measure reported in the
extension below. To avoid repeating the tuning comparison, the honest-forest
entries use the tuned fits. All six algebraic kappa variants are included.
</div>

```{r bridge-helpers, echo=FALSE}
# Named list of exact or algebraic weight vectors per cell.
cell_weight_list <- function(w_tuned,
                             pliv_lm, iivm_lm, pliv_rf, iivm_rf, kw) {
  lst <- list(
    "PLR-IV (tuned forest)"      = w_tuned$plriv,
    "Wald-AIPW (tuned forest)"   = w_tuned$waldaipw,
    "PLR-IV (OLS nuisances)"      = exact_w(pliv_lm),
    "Wald-AIPW (OLS/logit nuisances)" = exact_w(iivm_lm),
    "PLR-IV (ranger)"             = exact_w(pliv_rf),
    "Wald-AIPW (ranger)"          = exact_w(iivm_rf),
    "τ_cb_u (kappa)"   = kw$tau_cb_u,
    "τ_ml_u (kappa)"   = kw$tau_ml_u,
    "τ_ml_a10 (kappa)" = kw$tau_ml_a10,
    "τ_ml_a (kappa, unnormalized)"  = kw$tau_ml_a,
    "τ_ml_t (kappa, untreated-unnormalized)" = kw$tau_ml_t,
    "τ_ml_a0 (kappa, treated-unnormalized)" = kw$tau_ml_a0
  )
  Filter(function(w) !is.null(w) && length(w) && !all(is.na(w)), lst)
}

bridge_cells <- list(
  "Cubic age" = list(w = cell_weight_list(
    w_cub_tuned,
    lm_pliv_cub, lm_iivm_cub, rf_pliv_cub, rf_iivm_cub, kw_cub)),
  "Saturated age" = list(w = cell_weight_list(
    w_sat_tuned,
    lm_pliv_sat, lm_iivm_sat, rf_pliv_sat, rf_iivm_sat, kw_sat))
)

bridge_table <- function(wlist, cap) {
  ests <- names(wlist)
  df   <- do.call(rbind, lapply(ests, function(nm) weight_bridge(wlist[[nm]], D, nm)))
  disp <- df[, c("Scope","Min","Max","Neg_share","Sum_top10","Sum_w",
                 "Sum_abs_w")]

  kk <- disp %>%
    kbl(caption = cap, row.names = FALSE,
        col.names = c("Scope","Min","Max","Neg. share","ΣTop10%","Σω","Σ|ω|"),
        align = c("l", rep("r", 6))) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 12) %>%
    add_header_above(c(" " = 1, "Knaus summary() statistics" = 6))
  for (i in seq_along(ests)) kk <- pack_rows(kk, ests[i], (i - 1) * 3 + 1, i * 3)
  kk %>% footnote(general = "Neg. share is Knaus's observation-count statistic in [0,1]. Control rows use the sign-flipped −ω convention. Only exact or algebraic representations enter this table. Knaus's 10^-4 epsilon is a display convention for small individual weights, not an equality or normalization tolerance.")
}
```

### Cubic age

```{r bridge-cub}
bridge_table(bridge_cells[["Cubic age"]]$w,
             "Detailed outcome-weight statistics — Cubic age")
```

### Saturated age

```{r bridge-sat}
bridge_table(bridge_cells[["Saturated age"]]$w,
             "Detailed outcome-weight statistics — Saturated age")
```

## ESS and Sign-Cancellation Extension {.tabset}

<div class="note-box">
**Thesis extension.** Modified ESS is shown separately for treated and control
observations, followed by the value relative to each observed arm size.
`Opposite-sign mass T/C` is the percentage of absolute mass with the opposite
orientation within each arm. It measures sign cancellation rather than the
number of negative observations. In an IV design, signs are also shaped by the
residualized instrument, so a high value is not by itself evidence of an
invalid estimator. Conventional Kish ESS is omitted because signed outcome
weights are not a nonnegative normalized weighted sample.
</div>

```{r extension-helpers, echo=FALSE}
extension_table <- function(wlist, cap) {
  fmt_pair <- function(a, b, digits) {
    if (any(!is.finite(c(a, b)))) return(NA_character_)
    paste0(formatC(a, format = "f", digits = digits), " / ",
           formatC(b, format = "f", digits = digits))
  }
  df <- do.call(rbind, lapply(names(wlist), function(nm) {
    x <- outcome_weight_diagnostics(wlist[[nm]], D, nm)
    data.frame(
      Estimator = nm,
      `ESS mod T/C` = fmt_pair(x$ESS_mod_T, x$ESS_mod_C, 0),
      `ESS mod/N T/C` = fmt_pair(
        x$ESS_mod_T / sum(D == 1), x$ESS_mod_C / sum(D == 0), 3
      ),
      `Opposite-sign mass T/C (%)` = fmt_pair(
        100 * x$Wrong_mass_T, 100 * x$Wrong_mass_C, 1
      ),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }))
  df %>%
    kbl(caption = cap, row.names = FALSE, align = c("l","c","c","c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 12) %>%
    footnote(general = "T/C denotes treated / sign-oriented control. Only exact or algebraic representations enter this extension.")
}
```

### Cubic age

```{r extension-cub}
extension_table(bridge_cells[["Cubic age"]]$w,
                "Modified ESS and sign cancellation — Cubic age")
```

### Saturated age

```{r extension-sat}
extension_table(bridge_cells[["Saturated age"]]$w,
                "Modified ESS and sign cancellation — Saturated age")
```

---

# 5. Love Plots

<div class="note-box">
**Balance scale.** Each panel reports absolute mean differences only (no KS
statistic). Continuous covariates are standardized by the pooled unadjusted
standard deviation. Starred binary covariates are left on their raw 0/1 scale.
Because the adjusted comparison uses sign-oriented outcome weights, a raw binary
difference is a signed weighted-moment difference and need not compare two
literal proportions when opposite-oriented weights remain.
</div>

```{r love-helpers, echo=FALSE}
# Return NULL unless an exact/algebraic weight representation is available.
maybe_love <- function(title, w, D, X, ok = TRUE) {
  if (!isTRUE(ok) || all(is.na(w))) NULL else make_love(title, w, D, X)
}
```

## DML vs. Kappa {.tabset}

### Cubic age (untuned/tuned forests and kappa)

```{r love-cub, fig.height=9, fig.width=13}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nforest untuned",    w_cub_untuned$plriv,    D, X_dml_cub),
  maybe_love("Wald-AIPW\nforest untuned", w_cub_untuned$waldaipw, D, X_dml_cub),
  maybe_love("PLR-IV\nforest tuned",      w_cub_tuned$plriv,      D, X_dml_cub),
  maybe_love("Wald-AIPW\nforest tuned",   w_cub_tuned$waldaipw,   D, X_dml_cub),
  maybe_love("τ_cb_u (CBPS)",    kw_cub$tau_cb_u,   D, X_dml_cub),
  maybe_love("τ_ml_u",           kw_cub$tau_ml_u,   D, X_dml_cub),
  maybe_love("τ_ml_a10",         kw_cub$tau_ml_a10, D, X_dml_cub)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — Cubic age: untuned/tuned honest forest vs. kappa")
```

### Saturated age (untuned/tuned forests and kappa)

```{r love-sat, fig.height=9, fig.width=13}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nforest untuned",    w_sat_untuned$plriv,    D, X_dml_sat),
  maybe_love("Wald-AIPW\nforest untuned", w_sat_untuned$waldaipw, D, X_dml_sat),
  maybe_love("PLR-IV\nforest tuned",      w_sat_tuned$plriv,      D, X_dml_sat),
  maybe_love("Wald-AIPW\nforest tuned",   w_sat_tuned$waldaipw,   D, X_dml_sat),
  maybe_love("τ_cb_u (CBPS)",    kw_sat$tau_cb_u,   D, X_dml_sat),
  maybe_love("τ_ml_u",           kw_sat$tau_ml_u,   D, X_dml_sat),
  maybe_love("τ_ml_a10",         kw_sat$tau_ml_a10, D, X_dml_sat)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — Saturated age: untuned/tuned honest forest vs. kappa")
```

## Learners vs. Kappa {.tabset}

<div class="note-box">
**Included weights.** These panels contain only exact or algebraic outcome-
weight representations: the eligible OLS/logit and Ranger fits together with
the two main normalized kappa references. The XGBoost vectors and cubic-age OLS
PLR-IV vector fail the estimator-specific reproduction gate and are therefore
not used for balance diagnostics.
</div>

### Cubic age (PLR-IV and Wald-AIPW by learner)

```{r ext-love-cub, fig.height=11, fig.width=14}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nOLS nuisances", lm_pliv_cub$w, D, X_dml_cub, lm_pliv_cub$ok),
  maybe_love("Wald-AIPW\nOLS/logit nuisances", lm_iivm_cub$w, D, X_dml_cub, lm_iivm_cub$ok),
  maybe_love("PLR-IV\nranger",       rf_pliv_cub$w,   D, X_dml_cub, rf_pliv_cub$ok),
  maybe_love("Wald-AIPW\nranger",    rf_iivm_cub$w,   D, X_dml_cub, rf_iivm_cub$ok),
  maybe_love("τ_cb_u\nkappa",        kw_cub$tau_cb_u, D, X_dml_cub),
  maybe_love("τ_ml_u\nkappa",        kw_cub$tau_ml_u, D, X_dml_cub)
))
love_layout_cub <- rbind(
  c(1, 1, 2, 2, 3, 3),
  c(NA, 4, 4, 5, 5, NA)
)
grid.arrange(grobs = grobs, layout_matrix = love_layout_cub,
             top = "Love Plots — PLR-IV and Wald-AIPW learners vs. kappa | Cubic age")
```

### Saturated age (PLR-IV and Wald-AIPW by learner)

```{r ext-love-sat, fig.height=17, fig.width=14}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nOLS nuisances", lm_pliv_sat$w, D, X_dml_sat, lm_pliv_sat$ok),
  maybe_love("Wald-AIPW\nOLS/logit nuisances", lm_iivm_sat$w, D, X_dml_sat, lm_iivm_sat$ok),
  maybe_love("PLR-IV\nranger",       rf_pliv_sat$w,   D, X_dml_sat, rf_pliv_sat$ok),
  maybe_love("Wald-AIPW\nranger",    rf_iivm_sat$w,   D, X_dml_sat, rf_iivm_sat$ok),
  maybe_love("τ_cb_u\nkappa",        kw_sat$tau_cb_u, D, X_dml_sat),
  maybe_love("τ_ml_u\nkappa",        kw_sat$tau_ml_u, D, X_dml_sat)
))
grid.arrange(grobs = grobs, ncol = 2,
             top = "Love Plots — PLR-IV and Wald-AIPW learners vs. kappa | Saturated age")
```
