---
title: "Vietnam Draft Lottery — Smoother-Matrix Diagnostics"
subtitle: "Affine smoothers (Condition 3) and treatment-prediction alignment (Condition 5)"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  autodep = TRUE
)
options(knitr.kable.NA = "—")
```

# Purpose

This notebook checks the sufficient smoother conditions in Knaus (2024)
directly on the fitted Vietnam nuisance functions. It does not replace the
outcome-weight reproduction gate.

For an outcome smoother \(\widehat m=S Y\), Condition 3 requires

\[
S\mathbf 1_N=\mathbf 1_N.
\]

Thus, every row of \(S\) must sum to one. For Wald-AIPW the condition is
checked separately for \(S_0^z\) and \(S_1^z\). Condition 3 is sufficient for
normalised outcome weights. Full normalisation additionally requires

\[
S D=\widehat D
\]

for PLR-IV (Condition 5a), or

\[
S_0^zD=\widehat D_0^z,\qquad
S_1^zD=\widehat D_1^z
\]

for Wald-AIPW (Condition 5b).

The main analysis below fits only the tuned honest forests. The optional
section reconstructs matrices for OLS/logit and `ranger`. It is disabled by
default because those models are already examined in the main Vietnam
notebook and the \(N\times N\) matrices are memory intensive.

```{r switches}
RUN_ALTERNATIVE_LEARNERS <- FALSE
```

# Setup

```{r packages}
library(haven)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(knitr)
library(kableExtra)

lgr::get_logger("mlr3")$set_threshold("warn")
source(file.path("..", "functions_all.R"))
```

```{r data}
sipp <- read_dta("sipp.dta")
sipp_clean <- sipp[
  sipp$rsncode != 999 &
    !is.na(sipp$educ) &
    !is.na(sipp$kwage),
]

sipp_clean$age <- sipp_clean$age_5
sipp_clean$age_fac <- factor(sipp_clean$age_5)

Y <- log(sipp_clean$kwage)
Z <- as.integer(sipp_clean$rsncode)
D <- as.integer(sipp_clean$nvstat)

X_cubic <- make_X(
  ~ 0 + age + I(age^2) + I(age^3),
  data = sipp_clean
)
X_saturated_full <- make_X(~ 0 + age_fac, data = sipp_clean)
X_saturated <- X_saturated_full[, -1, drop = FALSE]

stopifnot(
  length(Y) == 3027L,
  nrow(X_cubic) == length(Y),
  nrow(X_saturated) == length(Y)
)
```

# Diagnostic helpers

```{r helpers}
rep_matrix <- function(x, repetition = 1L) {
  dx <- dim(x)
  if (length(dx) == 3L) {
    return(x[repetition, , , drop = TRUE])
  }
  as.matrix(x)
}

rep_prediction <- function(x, repetition = 1L) {
  if (is.null(dim(x))) return(as.numeric(x))
  as.numeric(x[, repetition])
}

affine_row <- function(S, specification, estimator, smoother) {
  row_mass <- rowSums(S)
  target <- rep(1, nrow(S))
  data.frame(
    Specification = specification,
    Estimator = estimator,
    Smoother = smoother,
    `Minimum row sum` = min(row_mass),
    `Maximum row sum` = max(row_mass),
    `Maximum |row sum - 1|` = max(abs(row_mass - 1)),
    `Condition 3` = isTRUE(all.equal(row_mass, target)),
    check.names = FALSE
  )
}

alignment_row <- function(S, D, D_hat, specification, estimator, condition) {
  reconstructed <- as.numeric(S %*% D)
  difference <- reconstructed - as.numeric(D_hat)
  data.frame(
    Specification = specification,
    Estimator = estimator,
    Condition = condition,
    RMSE = sqrt(mean(difference^2)),
    `Maximum absolute difference` = max(abs(difference)),
    Satisfied = isTRUE(all.equal(reconstructed, as.numeric(D_hat))),
    check.names = FALSE
  )
}

show_table <- function(x, caption, digits = 8) {
  x |>
    kbl(
      caption = caption,
      row.names = FALSE,
      digits = digits
    ) |>
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE
    )
}
```

# Tuned honest forests

The `dml_with_smoother()` objects store the cross-fitted outcome smoother
matrices under `NuPa.hat$smoothers`. The cubic and saturated models are fitted
with the same five-fold design and tuning rule used in the Vietnam analysis.

```{r fit-honest-forests, cache=TRUE}
set.seed(42)
forest_cubic <- dml_with_smoother(
  Y, D, X_cubic, Z,
  n_cf_folds = 5,
  tune.parameters = "all"
)

set.seed(42)
forest_saturated <- dml_with_smoother(
  Y, D, X_saturated, Z,
  n_cf_folds = 5,
  tune.parameters = "all"
)

forest_fits <- list(
  "Cubic age" = forest_cubic,
  "Saturated age" = forest_saturated
)
```

## Condition 3: affine outcome smoothers

```{r forest-condition-3}
condition3_forest <- do.call(rbind, lapply(names(forest_fits), function(spec) {
  fit <- forest_fits[[spec]]
  sm <- fit$NuPa.hat$smoothers

  rbind(
    affine_row(
      rep_matrix(sm$S), spec,
      "PLR-IV", "S"
    ),
    affine_row(
      rep_matrix(sm$S.z0), spec,
      "Wald-AIPW", "S.z0"
    ),
    affine_row(
      rep_matrix(sm$S.z1), spec,
      "Wald-AIPW", "S.z1"
    )
  )
}))

show_table(
  condition3_forest,
  "Condition 3 — row sums of the tuned honest-forest outcome smoothers",
  digits = 15
)
```

## Conditions 5a and 5b: treatment-prediction alignment

```{r forest-condition-5}
condition5_forest <- do.call(rbind, lapply(names(forest_fits), function(spec) {
  fit <- forest_fits[[spec]]
  sm <- fit$NuPa.hat$smoothers
  pr <- fit$NuPa.hat$predictions

  rbind(
    alignment_row(
      rep_matrix(sm$S), D, rep_prediction(pr$D.hat),
      spec, "PLR-IV", "5a: S D = D.hat"
    ),
    alignment_row(
      rep_matrix(sm$S.z0), D, rep_prediction(pr$D.hat.z0),
      spec, "Wald-AIPW", "5b: S.z0 D = D.hat.z0"
    ),
    alignment_row(
      rep_matrix(sm$S.z1), D, rep_prediction(pr$D.hat.z1),
      spec, "Wald-AIPW", "5b: S.z1 D = D.hat.z1"
    )
  )
}))

show_table(
  condition5_forest,
  "Conditions 5a and 5b — treatment-prediction alignment",
  digits = 8
)
```

## Outcome-weight reproduction gate

Condition 3 is interpreted only after the extracted outcome weights reproduce
the estimator. This table keeps the representation check separate from the
smoother conditions.

```{r forest-reproduction}
forest_reproduction <- do.call(rbind, lapply(names(forest_fits), function(spec) {
  fit <- forest_fits[[spec]]
  omega <- get_outcome_weights(fit)$omega
  estimates <- summary(fit)

  data.frame(
    Specification = spec,
    Estimator = c("PLR-IV", "Wald-AIPW"),
    `omega'Y` = c(
      as.numeric(omega["PLR-IV", ] %*% Y),
      as.numeric(omega["Wald-AIPW", ] %*% Y)
    ),
    Estimate = c(
      get_estimate(estimates, "PLR-IV"),
      get_estimate(estimates, "Wald-AIPW")
    ),
    `Reproduces estimate` = c(
      isTRUE(all.equal(
        as.numeric(omega["PLR-IV", ] %*% Y),
        get_estimate(estimates, "PLR-IV")
      )),
      isTRUE(all.equal(
        as.numeric(omega["Wald-AIPW", ] %*% Y),
        get_estimate(estimates, "Wald-AIPW")
      ))
    ),
    check.names = FALSE
  )
}))

show_table(
  forest_reproduction,
  "Estimator-specific outcome-weight reproduction",
  digits = 12
)
```

# Optional OLS/logit and ranger audit

The development version of `OutcomeWeights` can reconstruct the cross-fitted
outcome smoother matrix from stored `DoubleML` fold models:

- OLS supplies a linear-smoother matrix. With an intercept its rows should sum
  to one, although the raw cubic cross-product reconstruction is numerically
  ill-conditioned.
- `ranger` reconstructs forest weights from terminal-node membership and
  in-bag counts. The learners must therefore be fitted with
  `keep.inbag = TRUE`.

The code below is disabled by default. Set
`RUN_ALTERNATIVE_LEARNERS <- TRUE` above to execute it.

```{r alternative-helpers, eval=RUN_ALTERNATIVE_LEARNERS}
make_dml_data <- function(X) {
  dat <- as.data.frame(X)
  dat$y <- Y
  dat$d <- D
  dat$z <- Z
  DoubleMLData$new(
    data = dat,
    y_col = "y",
    d_cols = "d",
    z_cols = "z",
    x_cols = colnames(X)
  )
}

fit_pliv_object <- function(dml_data, learner) {
  set.seed(42)
  object <- DoubleMLPLIV$new(
    data = dml_data,
    ml_l = learner(),
    ml_m = learner(),
    ml_r = learner(),
    n_folds = 5
  )
  object$fit(store_models = TRUE, store_predictions = TRUE)
  object
}

fit_iivm_object <- function(dml_data, outcome_learner,
                            instrument_learner, treatment_learner) {
  set.seed(42)
  object <- DoubleMLIIVM$new(
    data = dml_data,
    ml_g = outcome_learner(),
    ml_m = instrument_learner(),
    ml_r = treatment_learner(),
    n_folds = 5,
    score = "LATE"
  )
  object$fit(store_models = TRUE, store_predictions = TRUE)
  object
}

extract_doubleml_smoothers <- function(object, dml_data) {
  type <- class(object)[1]
  dat <- as.matrix(dml_data$data)
  test_ids <- object$smpls[[1]]$test_ids
  train_ids <- object$smpls[[1]]$train_ids
  x_cols <- dml_data$x_cols
  y_col <- dml_data$y_col

  smoother <- getFromNamespace("get_DoubleML_smoother", "OutcomeWeights")

  if (type == "DoubleMLPLIV") {
    fold_models <- object$models$ml_l$d[[1]]
    return(list(
      S = smoother(
        fold_models, test_ids, train_ids, dat,
        x_cols, y_col, NULL
      )
    ))
  }

  if (type == "DoubleMLIIVM") {
    z <- dat[, dml_data$z_cols]
    return(list(
      S.z0 = smoother(
        object$models$ml_g0$d[[1]], test_ids, train_ids, dat,
        x_cols, y_col, z == 0
      ),
      S.z1 = smoother(
        object$models$ml_g1$d[[1]], test_ids, train_ids, dat,
        x_cols, y_col, z == 1
      )
    ))
  }

  stop("Only DoubleMLPLIV and DoubleMLIIVM are supported here.")
}
```

```{r fit-alternative-learners, eval=RUN_ALTERNATIVE_LEARNERS, cache=TRUE}
dml_data <- list(
  "Cubic age" = make_dml_data(X_cubic),
  "Saturated age" = make_dml_data(X_saturated)
)

alternative_fits <- list()
for (spec in names(dml_data)) {
  dat <- dml_data[[spec]]

  alternative_fits[[paste(spec, "PLR-IV OLS", sep = " | ")]] <-
    list(
      specification = spec,
      estimator = "PLR-IV",
      learner = "OLS",
      data = dat,
      object = fit_pliv_object(dat, function() lrn("regr.lm"))
    )

  alternative_fits[[paste(spec, "Wald-AIPW OLS/logit", sep = " | ")]] <-
    list(
      specification = spec,
      estimator = "Wald-AIPW",
      learner = "OLS/logit",
      data = dat,
      object = fit_iivm_object(
        dat,
        function() lrn("regr.lm"),
        function() lrn("classif.log_reg"),
        function() lrn("classif.log_reg")
      )
    )

  alternative_fits[[paste(spec, "PLR-IV ranger", sep = " | ")]] <-
    list(
      specification = spec,
      estimator = "PLR-IV",
      learner = "ranger",
      data = dat,
      object = fit_pliv_object(
        dat,
        function() lrn("regr.ranger", keep.inbag = TRUE)
      )
    )

  alternative_fits[[paste(spec, "Wald-AIPW ranger", sep = " | ")]] <-
    list(
      specification = spec,
      estimator = "Wald-AIPW",
      learner = "ranger",
      data = dat,
      object = fit_iivm_object(
        dat,
        function() lrn("regr.ranger", keep.inbag = TRUE),
        function() lrn("classif.ranger", keep.inbag = TRUE),
        function() lrn("classif.ranger", keep.inbag = TRUE)
      )
    )
}
```

```{r alternative-results, eval=RUN_ALTERNATIVE_LEARNERS}
alternative_audit <- lapply(alternative_fits, function(entry) {
  object <- entry$object
  smoothers <- extract_doubleml_smoothers(object, entry$data)
  omega <- get_outcome_weights(object, dml_data = entry$data)$omega
  reproduces <- isTRUE(all.equal(
    as.numeric(omega %*% Y),
    as.numeric(object$coef)
  ))

  if (entry$estimator == "PLR-IV") {
    c3 <- affine_row(
      smoothers$S, entry$specification,
      paste(entry$estimator, entry$learner), "S"
    )
    c5 <- alignment_row(
      smoothers$S, D, as.numeric(object$predictions$ml_r),
      entry$specification,
      paste(entry$estimator, entry$learner),
      "5a: S D = D.hat"
    )
  } else {
    c3_z0 <- affine_row(
      smoothers$S.z0, entry$specification,
      paste(entry$estimator, entry$learner), "S.z0"
    )
    c3_z1 <- affine_row(
      smoothers$S.z1, entry$specification,
      paste(entry$estimator, entry$learner), "S.z1"
    )
    c3 <- rbind(c3_z0, c3_z1)
    c5 <- rbind(
      alignment_row(
        smoothers$S.z0, D, as.numeric(object$predictions$ml_r0),
        entry$specification,
        paste(entry$estimator, entry$learner),
        "5b: S.z0 D = D.hat.z0"
      ),
      alignment_row(
        smoothers$S.z1, D, as.numeric(object$predictions$ml_r1),
        entry$specification,
        paste(entry$estimator, entry$learner),
        "5b: S.z1 D = D.hat.z1"
      )
    )
  }

  c3_output <- data.frame(
    Specification = c3$Specification,
    Estimator = c3$Estimator,
    Smoother = c3$Smoother,
    `Maximum |row sum - 1|` = c3$`Maximum |row sum - 1|`,
    `Condition 3` = c3$`Condition 3`,
    `Reproduces estimate` = reproduces,
    check.names = FALSE
  )

  c5_output <- data.frame(
    Specification = c5$Specification,
    Estimator = c5$Estimator,
    Condition = c5$Condition,
    RMSE = c5$RMSE,
    `Maximum absolute difference` = c5$`Maximum absolute difference`,
    Satisfied = c5$Satisfied,
    `Reproduces estimate` = reproduces,
    check.names = FALSE
  )

  list(condition3 = c3_output, condition5 = c5_output)
})

alternative_condition3 <- do.call(
  rbind,
  lapply(alternative_audit, `[[`, "condition3")
)
alternative_condition5 <- do.call(
  rbind,
  lapply(alternative_audit, `[[`, "condition5")
)

show_table(
  alternative_condition3,
  "Optional affine-smoother audit — OLS/logit and ranger",
  digits = 15
)

show_table(
  alternative_condition5,
  "Optional treatment-prediction alignment — OLS/logit and ranger",
  digits = 8
)
```

# Why tuned XGBoost is not included

An XGBoost smoother matrix can be reconstructed only for a restricted class
of fitted boosters supported by `OutcomeWeights`. In particular, the current
method requires `alpha = 0`, `subsample = 1`, and `max_delta_step = 0`. The
stored Vietnam XGBoost exports contain estimates, reconstructed outcome-weight
vectors, and diagnostics, but not the fitted fold-specific booster objects
needed to recover the smoother matrices.

More importantly, all current XGBoost outcome-weight vectors fail the
estimator-reproduction gate. A row-sum calculation on another reconstructed
matrix would therefore not validate their normalisation or balance
diagnostics. For the thesis, XGBoost remains in the point-estimate and
translation-invariance comparisons but is excluded from the smoother and
outcome-weight diagnostics.

# Interpretation

The checks have a strict order:

1. Verify that \(\omega'Y\) reproduces the estimator.
2. If it does, use Condition 3 to explain zero total weight mass.
3. Use Conditions 5a or 5b to distinguish full from scale normalisation.

A passing affine row-sum check cannot rescue a failed outcome-weight
representation. Conversely, Condition 3 may pass while Condition 5 fails;
that is precisely the theoretical pattern associated with scale-normalised
weights.
