---
title: "Angrist (1990) — Translation Invariance: Complete-Pipeline Rerun"
subtitle: "Vietnam draft lottery: kappa, honest forest, DoubleML learners, and nested XGBoost"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size:15px; color:#2c2c2c; }
h1 { border-bottom:2px solid #2c7bb6; padding-bottom:6px; margin-top:36px; }
h2 { color:#2c7bb6; margin-top:28px; }
table { font-size:13px; }
pre:not(.sourceCode) { display:none; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE,autodep=TRUE,fig.width=12,fig.height=6,
                      cache.extra="translation-rerun-only-v3-canonical-kappa")
```

# Purpose

This notebook evaluates translation invariance by fitting the complete estimation
rerun to the original log-wage outcome and to

\[
Y_i^{\mathrm{shift}}=Y_i+\log(100),
\]

which is the cents recoding used in the SUW reproduction. The empirical check is
the direct difference between the two separately fitted point estimates. 
It extends the classical
kappa comparison to data-adaptive estimators by allowing outcome nuisances and
tuning choices to be re-estimated under the same cross-fitting protocol after
the translation. 


# Setup

```{r packages-and-source}
library(haven)
library(knitr)
library(kableExtra)
library(dplyr)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")
source(file.path("..", "functions_all.R"))

ti_vietnam_row <- function(name,tau_original,tau_shifted,k,digits=12) {
  difference <- as.numeric(tau_shifted) - as.numeric(tau_original)
  data.frame(
    Estimator = name,
    Shift_k = round(as.numeric(k), digits),
    tau_original = round(as.numeric(tau_original), digits),
    tau_shifted = round(as.numeric(tau_shifted), digits),
    rerun_difference = round(difference, digits),
    stringsAsFactors = FALSE,
    check.names = FALSE
  )
}

display_rerun_results <- function(x) {
  fixed <- function(value,digits) {
    formatC(as.numeric(value), format = "f", digits = digits)
  }
  change <- function(value) {
    value <- as.numeric(value)
    value[abs(value) < 0.5e-6] <- 0
    formatC(value,format="f",digits=6)
  }
  data.frame(
    Specification = x$Specification,
    Estimator = x$Estimator,
    Shift = fixed(x$Shift_k,4),
    `Original estimate` = fixed(x$tau_original,4),
    `Shifted estimate` = fixed(x$tau_shifted,4),
    `Rerun change` = change(x$rerun_difference),
    check.names = FALSE
  )
}
```

## Runtime switches

```{r switches}
RUN_KAPPA <- TRUE
RUN_DML <- TRUE
RUN_LM_RF <- TRUE
RUN_XGB_RERUN <- FALSE
RUN_XGB_WALD_AIPW <- TRUE
RUN_XGB_PLIV <- TRUE
USE_XGB_ORIGINAL_EXPORT <- TRUE

SEED <- 42
N_CF <- 5
TUNE_EVALS <- 15
INNER_FOLDS <- 3
```

**XGBoost rule.** The default `RUN_XGB_RERUN = FALSE` loads the verified
complete-rerun export and does not repeat nested tuning. Set it to `TRUE` only
for a deliberate rebuild. A rebuild imports the original-outcome estimates
from the two tuning exports and retunes the shifted outcome with the same seed,
five outer folds, three inner folds, fifteen random-search evaluations, and
`tune_on_folds = TRUE`.

## Data — the exact main-analysis sample

```{r data}
stopifnot(file.exists("sipp.dta"))
sipp <- read_dta("sipp.dta")
sipp <- sipp[sipp$rsncode!=999 & !is.na(sipp$educ) & !is.na(sipp$kwage),]
sipp$age <- sipp$age_5
sipp$age2 <- sipp$age_5^2
sipp$age3 <- sipp$age_5^3
sipp$age_fac <- factor(sipp$age_5)

Y_orig <- as.numeric(log(sipp$kwage))
k_shift <- log(100)
Y_shift <- Y_orig+k_shift
D <- as.integer(sipp$nvstat)
Z <- as.integer(sipp$rsncode)

X_kappa_cub <- model.matrix(~age+age2+age3,data=sipp)
X_kappa_sat <- model.matrix(~age_fac,data=sipp)
X_dml_cub <- make_X(~0+age+I(age^2)+I(age^3),data=sipp)
X_dml_sat_full <- make_X(~0+age_fac,data=sipp)
X_dml_sat <- X_dml_sat_full[,-1,drop=FALSE]

workflows <- list(
  "Cubic age"=list(label="Cubic age",Y_orig=Y_orig,Y_shift=Y_shift,k=k_shift,
    D=D,Z=Z,X_kappa=X_kappa_cub,X_dml=X_dml_cub),
  "Saturated age"=list(label="Saturated age",Y_orig=Y_orig,Y_shift=Y_shift,k=k_shift,
    D=D,Z=Z,X_kappa=X_kappa_sat,X_dml=X_dml_sat))

data.frame(
  Sample = "Main-analysis sample",
  N = nrow(sipp),
  Translation = "log wage + log(100)",
  Specifications = paste(names(workflows),collapse=", ")
) %>%
  kbl(caption="Rerun setup",row.names=FALSE) %>%
  kable_styling(bootstrap_options=c("striped","condensed"),full_width=FALSE)
```

# 1. Kappa estimators


```{r kappa-builders}
kappa_estimator_names <- c("tau_cb_u","tau_ml_u","tau_ml_a10",
  "tau_ml_a","tau_ml_t","tau_ml_a0")
build_kappa_rows <- function(wf) {
  est_o <- kappa_point_estimates(wf$Y_orig,wf$Z,wf$D,wf$X_kappa)
  est_s <- kappa_point_estimates(wf$Y_shift,wf$Z,wf$D,wf$X_kappa)
  do.call(rbind,lapply(kappa_estimator_names,function(est){
    row <- ti_vietnam_row(est,est_o[[est]],est_s[[est]],wf$k)
    row$Specification <- wf$label; row
  }))
}
```

```{r kappa-run, cache=TRUE}
if(RUN_KAPPA)kappa_ti_all<-do.call(rbind,lapply(workflows,build_kappa_rows))
```

```{r kappa-table}
if(RUN_KAPPA) {
  display_rerun_results(kappa_ti_all) %>%
    kbl(caption="Kappa estimators: original versus shifted log wage",row.names=FALSE) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
    footnote(general=" The visible changes are concentrated in the cubic unnormalized kappa variants.")
}
```

# 2. DML tuned forest

```{r dml-headline-builder}
build_dml_headline_rows <- function(wf) {
  set.seed(SEED)
  invisible(capture.output(
    fit_o <- dml_with_smoother(wf$Y_orig,wf$D,wf$X_dml,wf$Z,
      n_cf_folds=N_CF,tune.parameters="all")
  ))
  set.seed(SEED)
  invisible(capture.output(
    fit_s <- dml_with_smoother(wf$Y_shift,wf$D,wf$X_dml,wf$Z,
      n_cf_folds=N_CF,tune.parameters="all")
  ))
  invisible(capture.output(ro <- summary(fit_o)))
  invisible(capture.output(rs <- summary(fit_s)))
  do.call(rbind,lapply(c("PLR-IV","Wald-AIPW"),function(est){
    row <- ti_vietnam_row(est,get_estimate(ro,est),
                          get_estimate(rs,est),wf$k)
    row$Specification <- wf$label; row
  }))
}
```

```{r dml-headline-run, cache=TRUE}
if(RUN_DML)dml_headline_ti_all<-do.call(rbind,lapply(workflows,build_dml_headline_rows))
```

```{r dml-headline-table}
if(RUN_DML) {
  display_rerun_results(dml_headline_ti_all) %>%
    kbl(caption="Tuned honest forest: original versus shifted log wage",row.names=FALSE) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
    footnote(general="The signed rerun change is displayed to six decimal places. Values that round to zero at this precision are shown uniformly as 0.000000.")
}
```

# 3. OLS/logit and Ranger learners


```{r learner-setup}
make_dml_data <- function(Y,D,Z,X){df<-as.data.frame(X);df$y<-Y;df$d<-D;df$z<-Z
  DoubleMLData$new(data=df,y_col="y",d_cols="d",z_cols="z",x_cols=colnames(X))}
ml_lm_iivm<-function()list(g=lrn("regr.lm"),m=lrn("classif.log_reg"),r=lrn("classif.log_reg"))
ml_rf_iivm<-function()list(g=lrn("regr.ranger",keep.inbag=TRUE),m=lrn("classif.ranger",keep.inbag=TRUE),r=lrn("classif.ranger",keep.inbag=TRUE))
ml_lm_pliv<-function()list(l=lrn("regr.lm"),m=lrn("regr.lm"),r=lrn("regr.lm"))
ml_rf_pliv<-function()list(l=lrn("regr.ranger",keep.inbag=TRUE),m=lrn("regr.ranger",keep.inbag=TRUE),r=lrn("regr.ranger",keep.inbag=TRUE))
fit_iivm<-function(dat,L){set.seed(SEED);o<-DoubleMLIIVM$new(data=dat,ml_g=L$g,ml_m=L$m,ml_r=L$r,n_folds=N_CF,score="LATE")
  invisible(capture.output(o$fit(store_models=TRUE,store_predictions=TRUE)));o}
fit_pliv<-function(dat,L){set.seed(SEED);o<-DoubleMLPLIV$new(data=dat,ml_l=L$l,ml_m=L$m,ml_r=L$r,n_folds=N_CF)
  invisible(capture.output(o$fit(store_models=TRUE,store_predictions=TRUE)));o}
build_learner_row<-function(wf,label,Lfun,fitfun){
  fo<-fitfun(make_dml_data(wf$Y_orig,wf$D,wf$Z,wf$X_dml),Lfun())
  fs<-fitfun(make_dml_data(wf$Y_shift,wf$D,wf$Z,wf$X_dml),Lfun())
  row<-ti_vietnam_row(label,get_dml_coef(fo),get_dml_coef(fs),wf$k)
  row$Specification<-wf$label;row}
```

```{r learner-run, cache=TRUE}
if(RUN_LM_RF)learner_ti_all<-do.call(rbind,lapply(workflows,function(wf)rbind(
  build_learner_row(wf,"PLR-IV (OLS nuisances)",ml_lm_pliv,fit_pliv),
  build_learner_row(wf,"Wald-AIPW (OLS/logit nuisances)",ml_lm_iivm,fit_iivm),
  build_learner_row(wf,"PLR-IV (Ranger)",ml_rf_pliv,fit_pliv),
  build_learner_row(wf,"Wald-AIPW (Ranger)",ml_rf_iivm,fit_iivm))))
```

```{r learner-table}
if(RUN_LM_RF) {
  display_rerun_results(learner_ti_all) %>%
    kbl(caption="DoubleML learners — original versus shifted log wage",row.names=FALSE) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
    footnote(general="OLS/logit labels the nuisance learner class,  cubic and saturated identify the feature specification")
}
```

# 4. Tuned XGBoost complete-pipeline rerun

```{r xgb-setup}
XGB_RERUN_EXPORT <- "vietnam_translation_xgb_rerun_export.rds"

validate_xgb_rerun_export <- function(export, path, required_specs) {
  if (!is.list(export) || is.null(export$results) || is.null(export$settings))
    stop(sprintf("%s has an incomplete schema.", path))
  required_columns <- c("Estimator", "Shift_k", "tau_original", "tau_shifted",
                        "rerun_difference", "Specification")
  if (!all(required_columns %in% names(export$results)))
    stop(sprintf("%s is missing required result columns.", path))
  expected_settings <- c(seed=SEED,n_cf=N_CF,tune_evals=TUNE_EVALS,
                         inner_folds=INNER_FOLDS)
  if (!all(names(expected_settings) %in% names(export$settings)) ||
      !isTRUE(export$settings$tune_on_folds))
    stop(sprintf("%s is missing required tuning metadata.", path))
  observed_settings <- vapply(names(expected_settings),
    function(x) as.integer(export$settings[[x]]), integer(1))
  if (!identical(unname(observed_settings),unname(as.integer(expected_settings))))
    stop(sprintf("%s uses tuning settings that differ from this notebook.", path))
  required_estimators <- c("Wald-AIPW (XGBoost tuned)",
                           "PLR-IV (XGBoost tuned)")
  expected_keys <- as.vector(outer(required_specs,required_estimators,paste,sep="\r"))
  observed_keys <- paste(export$results$Specification,export$results$Estimator,sep="\r")
  if (anyDuplicated(observed_keys) || !setequal(observed_keys,expected_keys))
    stop(sprintf("%s does not contain exactly the required Vietnam XGBoost rows.", path))
  numeric_columns <- c("Shift_k","tau_original","tau_shifted","rerun_difference")
  if (any(!vapply(export$results[numeric_columns],
                  function(x) all(is.finite(as.numeric(x))),logical(1))))
    stop(sprintf("%s contains non-finite translation results.", path))
  export$results[,required_columns,drop=FALSE]
}

if(RUN_XGB_RERUN){
  if(!RUN_XGB_WALD_AIPW || !RUN_XGB_PLIV)
    stop("An official Vietnam XGBoost translation rebuild must include both Wald-AIPW and PLR-IV.")
  if(!USE_XGB_ORIGINAL_EXPORT)stop("USE_XGB_ORIGINAL_EXPORT must be TRUE.")
  req<-c("vietnam_xgb_tuning_export.rds","vietnam_xgb_pliv_tuning_export.rds")
  if(!all(file.exists(req)))stop("Run both Vietnam XGBoost tuning notebooks before the translation rerun.")
  xgb_wald_export<-readRDS(req[1]);xgb_pliv_export<-readRDS(req[2])
  for(ex in list(xgb_wald_export,xgb_pliv_export)){
    if(!isTRUE(ex$settings$tune_on_folds))stop("Export is not a tune_on_folds=TRUE fit.")
    if(!all(names(workflows)%in%ex$weight_diagnostics$Cell))stop("Export cells do not match the translation workflows.")
    if(!identical(as.integer(ex$settings$seed),as.integer(SEED))||
       !identical(as.integer(ex$settings$n_cf),as.integer(N_CF))||
       !identical(as.integer(ex$settings$tune_evals),as.integer(TUNE_EVALS))||
       !identical(as.integer(ex$settings$inner_folds),as.integer(INNER_FOLDS)))
      stop("Export tuning settings do not match the shifted-outcome rerun settings.")
  }
}

xgb_ps<-function()ps(nrounds=p_int(50,400),max_depth=p_int(2,5),eta=p_dbl(0.02,0.20),min_child_weight=p_int(1,8))
make_xgb_iivm<-function()list(g=lrn("regr.xgboost",alpha=0,subsample=1,max_delta_step=0,base_score=0),m=lrn("classif.xgboost",predict_type="prob"),r=lrn("classif.xgboost",predict_type="prob"))
make_xgb_pliv<-function(){b<-list(alpha=0,subsample=1,max_delta_step=0,base_score=0);list(l=do.call(lrn,c(list("regr.xgboost"),b)),m=do.call(lrn,c(list("regr.xgboost"),b)),r=do.call(lrn,c(list("regr.xgboost"),b)))}
tune_set_iivm<-function()list(ml_g=xgb_ps(),ml_m=xgb_ps(),ml_r=xgb_ps())
tune_set_pliv<-function()list(ml_l=xgb_ps(),ml_m=xgb_ps(),ml_r=xgb_ps())
tune_settings_iivm<-function()list(n_folds_tune=INNER_FOLDS,rsmp_tune=rsmp("cv",folds=INNER_FOLDS),measure=list(ml_g="regr.rmse",ml_m="classif.logloss",ml_r="classif.logloss"),terminator=trm("evals",n_evals=TUNE_EVALS),algorithm=tnr("random_search"))
tune_settings_pliv<-function()list(n_folds_tune=INNER_FOLDS,rsmp_tune=rsmp("cv",folds=INNER_FOLDS),measure=list(ml_l="regr.rmse",ml_m="regr.rmse",ml_r="regr.rmse"),terminator=trm("evals",n_evals=TUNE_EVALS),algorithm=tnr("random_search"))

get_original_xgb<-function(export,cell,learner_label){
  row<-export$weight_diagnostics%>%filter(.data$Cell==cell,.data$Learner==learner_label)
  if(nrow(row)!=1)stop(sprintf("Expected one diagnostic row for %s / %s.",cell,learner_label))
  tau<-as.numeric(row$Estimate[1])
  if(!is.finite(tau))stop(sprintf("Tuned estimate is unavailable for %s / %s.",cell,learner_label))
  tau}
fit_shifted_wald<-function(wf){dat<-make_dml_data(wf$Y_shift,wf$D,wf$Z,wf$X_dml);L<-make_xgb_iivm()
  o<-DoubleMLIIVM$new(data=dat,ml_g=L$g,ml_m=L$m,ml_r=L$r,n_folds=N_CF,score="LATE")
  invisible(capture.output(o$tune(param_set=tune_set_iivm(),tune_settings=tune_settings_iivm(),tune_on_folds=TRUE)))
  invisible(capture.output(o$fit(store_models=TRUE,store_predictions=TRUE)));get_dml_coef(o)}
fit_shifted_pliv<-function(wf){dat<-make_dml_data(wf$Y_shift,wf$D,wf$Z,wf$X_dml);L<-make_xgb_pliv()
  o<-DoubleMLPLIV$new(data=dat,ml_l=L$l,ml_m=L$m,ml_r=L$r,n_folds=N_CF)
  invisible(capture.output(o$tune(param_set=tune_set_pliv(),tune_settings=tune_settings_pliv(),tune_on_folds=TRUE)))
  invisible(capture.output(o$fit(store_models=TRUE,store_predictions=TRUE)));get_dml_coef(o)}
build_xgb_rows<-function(wf){out<-list()
  if(RUN_XGB_WALD_AIPW){orig<-get_original_xgb(xgb_wald_export,wf$label,"XGB tuned");set.seed(SEED);ts<-fit_shifted_wald(wf)
    r<-ti_vietnam_row("Wald-AIPW (XGBoost tuned)",orig,ts,wf$k);r$Specification<-wf$label;out$wald<-r}
  if(RUN_XGB_PLIV){orig<-get_original_xgb(xgb_pliv_export,wf$label,"PLR-IV XGB tuned");set.seed(SEED);ts<-fit_shifted_pliv(wf)
    r<-ti_vietnam_row("PLR-IV (XGBoost tuned)",orig,ts,wf$k);r$Specification<-wf$label;out$pliv<-r}
  do.call(rbind,out)}
```

```{r xgb-run, cache=FALSE}
if(RUN_XGB_RERUN){
  xgb_ti_all<-do.call(rbind,lapply(workflows,build_xgb_rows))
  rebuilt_xgb_export<-list(results=xgb_ti_all,settings=list(seed=SEED,n_cf=N_CF,
    tune_evals=TUNE_EVALS,inner_folds=INNER_FOLDS,tune_on_folds=TRUE,
    original_exports=c("vietnam_xgb_tuning_export.rds","vietnam_xgb_pliv_tuning_export.rds")))
  validate_xgb_rerun_export(rebuilt_xgb_export,XGB_RERUN_EXPORT,names(workflows))
  saveRDS(rebuilt_xgb_export,XGB_RERUN_EXPORT)
} else {
  if(!file.exists(XGB_RERUN_EXPORT))
    stop("Missing verified Vietnam XGBoost translation export. Restore it or set RUN_XGB_RERUN=TRUE to rebuild it.")
  xgb_ti_all<-validate_xgb_rerun_export(
    readRDS(XGB_RERUN_EXPORT),XGB_RERUN_EXPORT,names(workflows))
}
```

```{r xgb-table}
if(exists("xgb_ti_all")) {
  display_rerun_results(xgb_ti_all) %>%
    kbl(caption="Tuned XGBoost — original export versus shifted nested-tuning rerun",row.names=FALSE) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
    footnote(general=paste0(
      if(RUN_XGB_RERUN)
        "This render rebuilt the shifted-outcome nested-tuning results. " else
        "This render loaded the verified complete-rerun XGBoost export. ",
      "The signed rerun change is reported without a binary cutoff."
    ))
}
```



# 5. Save reusable rerun results

```{r save-complete-rerun, cache=FALSE, include=FALSE}
result_parts <- Filter(Negate(is.null), list(
  Kappa = if (exists("kappa_ti_all")) kappa_ti_all else NULL,
  `Tuned honest forest` = if (exists("dml_headline_ti_all")) dml_headline_ti_all else NULL,
  `OLS/logit and Ranger` = if (exists("learner_ti_all")) learner_ti_all else NULL,
  `Tuned XGBoost` = if (exists("xgb_ti_all")) xgb_ti_all else NULL
))
for (nm in names(result_parts)) result_parts[[nm]]$Family <- nm
translation_results <- bind_rows(result_parts)
if(RUN_XGB_RERUN){
  expected_families<-c("Kappa","Tuned honest forest","OLS/logit and Ranger","Tuned XGBoost")
  if(!setequal(unique(translation_results$Family),expected_families))
    stop("Refusing to replace the complete Vietnam translation export because a result family is missing.")
  saveRDS(list(
    results = translation_results,
    settings = list(seed=SEED,n_cf=N_CF,shift=log(100),
                    grf_tuning="all",xgboost_tune_on_folds=TRUE,
                    xgboost_tune_evals=TUNE_EVALS,xgboost_inner_folds=INNER_FOLDS)
  ), "vietnam_translation_rerun_results.rds")
}
```
