---
title: "The Vietnam Draft Lottery: XGBoost Tuning for PLR-IV"
subtitle: "Nested nuisance tuning within the cross-fitting design for PLR IV"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size:15px; color:#2c2c2c; }
h1 { border-bottom:2px solid #2c7bb6; padding-bottom:6px; margin-top:36px; }
h2 { color:#2c7bb6; margin-top:28px; }
table { font-size:13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE,cache=TRUE,autodep=TRUE,fig.width=11,fig.height=5.5)
```

```{r packages, include=FALSE}
library(haven)
library(knitr)
library(kableExtra)
library(dplyr)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)
source(file.path("..", "functions_all.R"))
lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

SEED <- 42
N_CF <- 5
TUNE_EVALS <- 15
INNER_FOLDS <- 3
EXPORT_WEIGHT_VECTORS <- TRUE
RERUN_TUNING <- FALSE
EXPORT_PATH <- "vietnam_xgb_pliv_tuning_export.rds"
```

**Purpose.** This notebook compares untuned and fold-specifically tuned
XGBoost nuisance learners for PLR-IV under the cubic- and saturated-age
specifications. Tuning is performed within each outer training fold. The
estimates, nuisance losses, tuning records, and outcome-weight reproduction
audit are exported to the main Vietnam notebook.

**Runtime.** A full rerun tunes three nuisance models using five outer folds,
three inner folds, and 15 random-search evaluations per nuisance and outer
fold. With `RERUN_TUNING = FALSE`, the notebook reads the existing export and
does not refit XGBoost.

# Setup

```{r data}
stopifnot(file.exists("sipp.dta"))
sipp <- read_dta("sipp.dta")
sipp <- sipp[sipp$rsncode != 999 & !is.na(sipp$educ) & !is.na(sipp$kwage), ]
sipp$age <- sipp$age_5
sipp$age_fac <- factor(sipp$age_5)
sipp$lwage <- log(sipp$kwage)

Y <- as.numeric(sipp$lwage)
D <- as.integer(sipp$nvstat)
Z <- as.integer(sipp$rsncode)
X_cub <- make_X(~0+age+I(age^2)+I(age^3),data=sipp)
X_sat_full <- make_X(~0+age_fac,data=sipp)
X_sat <- X_sat_full[,-1,drop=FALSE]
data.frame(
  Quantity=c("Analysis sample","Cubic-age covariates",
             "Saturated-age covariates"),
  Value=c(format(nrow(sipp),big.mark=","),ncol(X_cub),ncol(X_sat))
) %>%
  kbl(caption="Analysis dimensions") %>%
  kable_styling(bootstrap_options=c("striped","condensed"),full_width=FALSE)
```

```{r learners-and-tuning}
make_dml_data <- function(Y,D,Z,X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data=df,y_col="y",d_cols="d",z_cols="z",x_cols=colnames(X))
}

make_xgb_learners <- function() {
  base <- list(alpha=0,subsample=1,max_delta_step=0,base_score=0)
  list(l=do.call(lrn,c(list("regr.xgboost"),base)),
       m=do.call(lrn,c(list("regr.xgboost"),base)),
       r=do.call(lrn,c(list("regr.xgboost"),base)))
}
xgb_ps <- function() ps(nrounds=p_int(50,400),max_depth=p_int(2,5),
                         eta=p_dbl(0.02,0.20),min_child_weight=p_int(1,8))
tune_param_set <- function() list(ml_l=xgb_ps(),ml_m=xgb_ps(),ml_r=xgb_ps())
tune_settings <- function() list(
  n_folds_tune=INNER_FOLDS,rsmp_tune=rsmp("cv",folds=INNER_FOLDS),
  measure=list(ml_l="regr.rmse",ml_m="regr.rmse",ml_r="regr.rmse"),
  terminator=trm("evals",n_evals=TUNE_EVALS),algorithm=tnr("random_search"))

get_prediction_vector <- function(obj,name) {
  p <- tryCatch(obj$predictions[[name]],error=function(e)NULL)
  if(is.null(p))return(numeric(0)); as.numeric(drop(p))
}
rmse_loss <- function(truth,pred) {
  if(length(pred)!=length(truth)||!length(pred))return(NA_real_)
  keep <- is.finite(truth)&is.finite(pred)
  if(!any(keep))return(NA_real_)
  sqrt(mean((as.numeric(truth[keep])-as.numeric(pred[keep]))^2))
}
calc_nuisance_losses <- function(r,cell) {
  if(is.null(r))return(c(outcome_l_rmse=NA,instrument_m_rmse=NA,treatment_r_rmse=NA))
  c(outcome_l_rmse=rmse_loss(cell$Y,get_prediction_vector(r$obj,"ml_l")),
    instrument_m_rmse=rmse_loss(cell$Z,get_prediction_vector(r$obj,"ml_m")),
    treatment_r_rmse=rmse_loss(cell$D,get_prediction_vector(r$obj,"ml_r")))
}

fit_cell <- function(dat,learners,label,tune=FALSE) {
  tryCatch({
    set.seed(SEED)
    obj <- DoubleMLPLIV$new(data=dat,ml_l=learners$l,ml_m=learners$m,
                            ml_r=learners$r,n_folds=N_CF)
    if(tune)obj$tune(param_set=tune_param_set(),tune_settings=tune_settings(),
                     tune_on_folds=TRUE)
    obj$fit(store_models=TRUE,store_predictions=TRUE)
    list(label=label,tuned=tune,coef=get_dml_coef(obj),se=get_dml_se(obj),
         tuning_result=tryCatch(obj$tuning_result,error=function(e)NULL),
         params=tryCatch(obj$params,error=function(e)NULL),obj=obj)
  },error=function(e){message(sprintf("Fit failed for %s: %s",label,conditionMessage(e)));NULL})
}

cells <- list("Cubic age"=list(Y=Y,D=D,Z=Z,X=X_cub),
              "Saturated age"=list(Y=Y,D=D,Z=Z,X=X_sat))
```

# Fit: untuned versus tuned XGBoost

```{r fit-all, include=FALSE, cache.extra=list(RERUN_TUNING,SEED,N_CF,TUNE_EVALS,INNER_FOLDS)}
if (RERUN_TUNING) {
  results <- list()
  for(cn in names(cells)) {
    cell <- cells[[cn]]
    dat <- make_dml_data(cell$Y,cell$D,cell$Z,cell$X)
    results[[cn]] <- list(
      untuned=fit_cell(dat,make_xgb_learners(),
                       paste(cn,"pliv-xgb-untuned"),FALSE),
      tuned=fit_cell(dat,make_xgb_learners(),
                     paste(cn,"pliv-xgb-tuned"),TRUE))
  }
} else {
  if (!file.exists(EXPORT_PATH)) {
    stop("Verified PLR-IV XGBoost export not found. Set RERUN_TUNING=TRUE to rebuild it.")
  }
  stored_export <- readRDS(EXPORT_PATH)
}
```

## 1. Estimate comparison

```{r comparison-table}
fmt_est <- function(r)if(is.null(r))"—"else sprintf("%.3f (%.3f)",r$coef,r$se)
if (RERUN_TUNING) {
  rows <- do.call(rbind,lapply(names(results),function(cn){R<-results[[cn]]
    data.frame(Cell=cn,Learner=c("PLR-IV XGB untuned","PLR-IV XGB tuned (tune_on_folds=TRUE)"),
               `Estimate (SE)`=c(fmt_est(R$untuned),fmt_est(R$tuned)),check.names=FALSE)}))
} else {
  rows <- stored_export$estimate_comparison
}
rows %>% kbl(caption="Vietnam PLR-IV: XGBoost untuned vs. fold-specifically tuned",
             align=c("l","l","c")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
  collapse_rows(columns=1,valign="middle") %>%
  row_spec(which(rows$Learner=="PLR-IV XGB tuned (tune_on_folds=TRUE)"),background="#eef6ff") %>%
  footnote(general="Tuned nuisances select nrounds, max_depth, eta, and min_child_weight by three-fold random-search CV inside each outer training fold.")
```

## 2. Nuisance-prediction losses

```{r nuisance-loss-table}
if (RERUN_TUNING) {
  loss_rows <- do.call(rbind,lapply(names(results),function(cn){R<-results[[cn]];cell<-cells[[cn]]
    u<-calc_nuisance_losses(R$untuned,cell);t<-calc_nuisance_losses(R$tuned,cell)
    data.frame(Cell=cn,Learner=c("PLR-IV XGB untuned","PLR-IV XGB tuned"),
      `Outcome l RMSE`=c(u["outcome_l_rmse"],t["outcome_l_rmse"]),
      `Instrument m RMSE`=c(u["instrument_m_rmse"],t["instrument_m_rmse"]),
      `Treatment r RMSE`=c(u["treatment_r_rmse"],t["treatment_r_rmse"]),
      check.names=FALSE,row.names=NULL)}))
} else {
  loss_rows <- stored_export$nuisance_losses
}
loss_rows %>% kbl(digits=4,caption="PLR-IV nuisance losses: untuned vs. tuned (lower is better)") %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
  collapse_rows(columns=1,valign="middle") %>%
  row_spec(which(loss_rows$Learner=="PLR-IV XGB tuned"),background="#eef6ff")
```

## 3. Tuning protocol


```{r tuning-protocol}
data.frame(
  Component=c(
    "Outer cross-fitting","Inner validation","Search algorithm","Budget",
    "nrounds","max_depth","eta","min_child_weight",
    "Outcome nuisance","Instrument nuisance","Treatment nuisance"
  ),
  Specification=c(
    paste(N_CF,"folds"),paste(INNER_FOLDS,"folds"),"Random search",
    paste(TUNE_EVALS,"evaluations per nuisance and outer fold"),
    "50–400","2–5","0.02–0.20","1–8",
    "RMSE","RMSE","RMSE"
  )
) %>%
  kbl(caption="Nested XGBoost tuning protocol: PLR-IV") %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)
```

```{r hyperparameter-summary, include=FALSE}
extract_tuned_params <- function(r,cn){
  if(is.null(r)||!r$tuned)return(data.frame())
  src<-if(!is.null(r$params))r$params else r$tuning_result
  if(is.null(src))return(data.frame())
  u<-unlist(src,recursive=TRUE);u<-u[grepl("nrounds|max_depth|eta|min_child_weight",names(u))]
  if(!length(u))return(data.frame())
  hp<-data.frame(Cell=cn,Raw_name=names(u),Value=as.numeric(u))
  hp$Nuisance<-ifelse(grepl("ml_l",hp$Raw_name),"Outcome l",
               ifelse(grepl("ml_m",hp$Raw_name),"Instrument m",
               ifelse(grepl("ml_r",hp$Raw_name),"Treatment r","Unknown")))
  hp$Parameter<-sub(".*(nrounds|max_depth|eta|min_child_weight).*","\\1",hp$Raw_name);hp
}
if (RERUN_TUNING) {
  hp_parts<-lapply(names(results),function(cn)extract_tuned_params(results[[cn]]$tuned,cn))
  hp_parts<-Filter(function(x)nrow(x)>0,hp_parts)
  hp_raw<-if(length(hp_parts))do.call(rbind,hp_parts)else data.frame()
  if (!nrow(hp_raw)) stop("Fold-specific hyperparameters could not be extracted.")
  hp_summary<-hp_raw%>%filter(Nuisance!="Unknown")%>%group_by(Cell,Nuisance,Parameter)%>%
    summarise(Median_across_folds=median(Value),N_values_seen=n(),.groups="drop")
} else {
  hp_raw <- NULL
  hp_summary <- stored_export$hyperparameters
}
```

## 4. Outcome-weight audit and export

```{r weight-helpers}
omega_vec<-function(om){w<-om$omega;if(is.null(dim(w)))return(as.vector(w))
  if(!is.null(rownames(w))&&"PLR-IV"%in%rownames(w))return(as.vector(w["PLR-IV",]));as.vector(w[1,])}
extract_xgb_w<-function(r,cell){if(is.null(r)||is.null(r$obj))stop("fitted DoubleML object is NULL")
  omega_vec(get_outcome_weights(object=r$obj,dml_data=make_dml_data(cell$Y,cell$D,cell$Z,cell$X)))}
safe_extract_xgb_w<-function(r,cell)tryCatch(extract_xgb_w(r,cell),error=function(e)NULL)
one_diag_row<-function(cn,label,r,cell){
  w<-safe_extract_xgb_w(r,cell)
  if(is.null(r)||is.null(w))return(data.frame(
    Cell=cn,Learner=label,Estimate=if(is.null(r))NA else r$coef,
    `Overall sum`=NA,`Treated sum`=NA,`Control sum`=NA,
    `ω′Y reproduces τ̂`=NA,check.names=FALSE))
  data.frame(
    Cell=cn,Learner=label,Estimate=r$coef,
    `Overall sum`=sum(w),
    `Treated sum`=sum(w[cell$D==1]),
    `Control sum`=-sum(w[cell$D==0]),
    `ω′Y reproduces τ̂`=tryCatch(
      check_weight_identity(w,cell$Y,r$coef),error=function(e)NA),
    check.names=FALSE)}
```

```{r weight-table}
if (RERUN_TUNING) {
  xgb_pliv_weight_table<-do.call(rbind,lapply(names(cells),function(cn){R<-results[[cn]];cell<-cells[[cn]]
    rbind(one_diag_row(cn,"PLR-IV XGB untuned",R$untuned,cell),
          one_diag_row(cn,"PLR-IV XGB tuned",R$tuned,cell))}))
} else {
  xgb_pliv_weight_table <- stored_export$weight_diagnostics
}
if ("Identity ok" %in% names(xgb_pliv_weight_table))
  names(xgb_pliv_weight_table)[names(xgb_pliv_weight_table)=="Identity ok"] <- "ω′Y reproduces τ̂"

learner_variant<-function(label){
  if(grepl("untuned",label,fixed=TRUE))"untuned" else "tuned"
}
audit_vector<-function(cell_name,learner_label){
  variant<-learner_variant(learner_label)
  if(RERUN_TUNING){
    return(safe_extract_xgb_w(results[[cell_name]][[variant]],cells[[cell_name]]))
  }
  tryCatch(stored_export$omega_vectors[[cell_name]][[variant]],error=function(e)NULL)
}
xgb_pliv_weight_display<-do.call(rbind,lapply(seq_len(nrow(xgb_pliv_weight_table)),function(i){
  row<-xgb_pliv_weight_table[i,,drop=FALSE]
  cell_name<-as.character(row$Cell)
  learner_label<-as.character(row$Learner)
  w<-audit_vector(cell_name,learner_label)
  d<-cells[[cell_name]]$D
  data.frame(
    Cell=cell_name,
    Variant=if(learner_variant(learner_label)=="untuned")"Untuned XGBoost" else "Tuned XGBoost",
    `Overall sum`=if(is.null(w))NA_real_ else sum(w),
    `Treated sum`=if(is.null(w))NA_real_ else sum(w[d==1]),
    `Control sum`=if(is.null(w))NA_real_ else -sum(w[d==0]),
    `ω′Y reproduces τ̂`=if(is.null(w))NA else
      check_weight_identity(w,cells[[cell_name]]$Y,as.numeric(row$Estimate)),
    check.names=FALSE
  )
}))
xgb_pliv_weight_display%>%
  kbl(digits=4,caption="Vietnam PLR-IV — reconstructed XGBoost weight audit",
      align=c("l","l","r","r","r","c"))%>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=FALSE)%>%
  collapse_rows(columns=1,valign="middle")%>%
  row_spec(which(xgb_pliv_weight_display$Variant=="Tuned XGBoost"),background="#eef6ff")%>%
  footnote(general="The sums describe reconstructed outcome-weight vectors. Control weights are reported with reversed sign.Reproduction follows all.equal() check with R's default tolerance.")
```

```{r omega-and-identities}
if (RERUN_TUNING) {
  xgb_pliv_omega_vectors<-lapply(names(cells),function(cn)list(
    untuned=if(EXPORT_WEIGHT_VECTORS)safe_extract_xgb_w(results[[cn]]$untuned,cells[[cn]])else NULL,
    tuned=if(EXPORT_WEIGHT_VECTORS)safe_extract_xgb_w(results[[cn]]$tuned,cells[[cn]])else NULL))
  names(xgb_pliv_omega_vectors)<-names(cells)
  xgb_pliv_identity_checks<-lapply(names(cells),function(cn){cell<-cells[[cn]];R<-results[[cn]];W<-xgb_pliv_omega_vectors[[cn]]
    list(
      untuned=tryCatch(if(is.null(W$untuned)||is.null(R$untuned))NA else
        check_weight_identity(W$untuned,cell$Y,R$untuned$coef),error=function(e)NA),
      tuned=tryCatch(if(is.null(W$tuned)||is.null(R$tuned))NA else
        check_weight_identity(W$tuned,cell$Y,R$tuned$coef),error=function(e)NA))})
  names(xgb_pliv_identity_checks)<-names(cells)
} else {
  xgb_pliv_omega_vectors <- stored_export$omega_vectors
  xgb_pliv_identity_checks <- stored_export$identity_checks
}
```

```{r export, include=FALSE, cache=FALSE}
if (RERUN_TUNING) {
  xgb_pliv_export<-list(estimate_comparison=rows,nuisance_losses=loss_rows,
    hyperparameters=hp_summary,hyperparameters_raw=hp_raw,
    weight_diagnostics=xgb_pliv_weight_table,omega_vectors=xgb_pliv_omega_vectors,
    identity_checks=xgb_pliv_identity_checks,
    settings=list(seed=SEED,n_cf=N_CF,tune_evals=TUNE_EVALS,inner_folds=INNER_FOLDS,
      tune_on_folds=TRUE,learner="xgboost",dml_class="DoubleMLPLIV",
      score="partialling out",export_weight_vectors=EXPORT_WEIGHT_VECTORS,cells=names(cells)),
    sample_info=list(n=nrow(sipp),mean_Z=mean(Z),mean_D=mean(D),outcome="log(kwage)",
                     specifications=names(cells)))
  saveRDS(xgb_pliv_export,EXPORT_PATH)
}
```
