---
title: "Vietnam Forest Translation Audit"
subtitle: "Tuned honest forest and Ranger smoother matrices"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    code_folding: show
    df_print: kable
---

```{css, echo=FALSE}
body { font-size:15px; color:#2c2c2c; }
h1 { border-bottom:2px solid #2c7bb6; padding-bottom:6px; }
h2 { color:#2c7bb6; }
table { font-size:12px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  message=FALSE,
  warning=FALSE,
  cache=FALSE
)
options(knitr.kable.NA="—")
```

# Purpose

This internal notebook localizes the rerun translation changes for
the two forest implementations used in the Vietnam application. It compares
the original log-wage outcome with \(Y+\log(100)\) for the cubic-age and
saturated-age specifications.

Within every original/translated pair, \(D\), \(Z\), \(X\), observation order,
five-fold cross-fitting rule, initial random-number state, learner settings,
and nuisance-fit order are held fixed. The honest forest uses the tuned
headline specification (`tune.parameters = "all"`). Ranger uses the same
untuned learner settings as the main translation notebook. Only \(Y\) changes.
The audit stops if the realized cross-fitting folds differ.

For each outcome smoother, it checks whether the realized matrix changes and
verifies

\[
\widehat g^{\,k}-\widehat g-k\mathbf 1
=
(S^k-S)Y+k(S^k\mathbf 1-\mathbf 1).
\]

The first term isolates changes in the fitted smoother weights, the second
isolates departures of the translated smoother's row sums from one.

# Setup and fixed sample

```{r packages-and-data}
library(haven)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(dplyr)
library(knitr)
library(kableExtra)

lgr::get_logger("mlr3")$set_threshold("warn")
source(file.path("..","functions_all.R"))

if (!exists("get_DoubleML_smoother",
            envir=asNamespace("OutcomeWeights"),inherits=FALSE)) {
  stop("The installed OutcomeWeights version does not provide get_DoubleML_smoother().")
}

SEED<-42L
N_CF<-5L
SHIFT_K<-log(100)
WEIGHT_TOLERANCE<-1e-12
BLOCK_ROWS<-100L

sipp<-read_dta("sipp.dta")
sipp<-sipp[
  sipp$rsncode!=999 & !is.na(sipp$educ) & !is.na(sipp$kwage),
]
sipp$age<-sipp$age_5
sipp$age_fac<-factor(sipp$age_5)

Y<-as.numeric(log(sipp$kwage))
D<-as.integer(sipp$nvstat)
Z<-as.integer(sipp$rsncode)

X_cubic<-make_X(~0+age+I(age^2)+I(age^3),data=sipp)
X_saturated_full<-make_X(~0+age_fac,data=sipp)
X_saturated<-X_saturated_full[,-1,drop=FALSE]

specifications<-list(
  "Cubic age"=X_cubic,
  "Saturated age"=X_saturated
)

stopifnot(
  length(Y)==3027L,
  all(vapply(specifications,nrow,integer(1))==length(Y))
)
```

# Audit functions

```{r helpers}
max_abs<-function(x) max(abs(as.numeric(x)))
rmse<-function(x) sqrt(mean(as.numeric(x)^2))

make_dml_data<-function(outcome,X){
  dat<-as.data.frame(X)
  dat$y<-outcome
  dat$d<-D
  dat$z<-Z
  DoubleMLData$new(
    data=dat,
    y_col="y",
    d_cols="d",
    z_cols="z",
    x_cols=colnames(X)
  )
}

fit_honest_forest<-function(outcome,X){
  set.seed(SEED)
  invisible(capture.output(
    fit<-dml_with_smoother(
      outcome,D,X,Z,
      n_cf_folds=N_CF,
      tune.parameters="all"
    )
  ))
  fit
}

fit_ranger<-function(outcome,X,estimator){
  dml_data<-make_dml_data(outcome,X)

  if (estimator=="PLR-IV") {
    learners<-list(
      l=lrn("regr.ranger",keep.inbag=TRUE),
      m=lrn("regr.ranger",keep.inbag=TRUE),
      r=lrn("regr.ranger",keep.inbag=TRUE)
    )
    set.seed(SEED)
    object<-DoubleMLPLIV$new(
      data=dml_data,
      ml_l=learners$l,
      ml_m=learners$m,
      ml_r=learners$r,
      n_folds=N_CF
    )
  } else if (estimator=="Wald-AIPW") {
    learners<-list(
      g=lrn("regr.ranger",keep.inbag=TRUE),
      m=lrn("classif.ranger",keep.inbag=TRUE),
      r=lrn("classif.ranger",keep.inbag=TRUE)
    )
    set.seed(SEED)
    object<-DoubleMLIIVM$new(
      data=dml_data,
      ml_g=learners$g,
      ml_m=learners$m,
      ml_r=learners$r,
      n_folds=N_CF,
      score="LATE"
    )
  } else {
    stop("Unsupported estimator.")
  }

  invisible(capture.output(
    object$fit(store_models=TRUE,store_predictions=TRUE)
  ))
  list(object=object,data=dml_data)
}

extract_ranger_smoothers<-function(fit){
  object<-fit$object
  dml_data<-fit$data
  smoother<-getFromNamespace("get_DoubleML_smoother","OutcomeWeights")
  dat<-as.matrix(dml_data$data)
  test_ids<-object$smpls[[1]]$test_ids
  train_ids<-object$smpls[[1]]$train_ids

  if (inherits(object,"DoubleMLPLIV")) {
    return(list(
      S=smoother(
        object$models$ml_l$d[[1]],
        test_ids,train_ids,dat,dml_data$x_cols,dml_data$y_col,NULL
      )
    ))
  }

  z<-dat[,dml_data$z_cols]
  list(
    S.z0=smoother(
      object$models$ml_g0$d[[1]],
      test_ids,train_ids,dat,dml_data$x_cols,dml_data$y_col,z==0
    ),
    S.z1=smoother(
      object$models$ml_g1$d[[1]],
      test_ids,train_ids,dat,dml_data$x_cols,dml_data$y_col,z==1
    )
  )
}

smoother_dimensions<-function(S){
  dimensions<-dim(S)
  if (length(dimensions)==2L && dimensions[1L]==dimensions[2L]) {
    return(c(dimensions[1L],dimensions[2L]))
  }
  if (length(dimensions)==3L && dimensions[1L]==1L &&
      dimensions[2L]==dimensions[3L]) {
    return(c(dimensions[2L],dimensions[3L]))
  }
  stop("Expected an N x N matrix or a 1 x N x N smoother array.")
}

smoother_block<-function(S,rows,N){
  if (length(dim(S))==2L) return(S[rows,,drop=FALSE])
  matrix(S[1L,rows,,drop=FALSE],nrow=length(rows),ncol=N)
}

audit_smoother<-function(
    S_original,S_shifted,pred_original,pred_shifted,
    specification,learner,estimator,smoother_name){

  if (!identical(smoother_dimensions(S_original),
                 smoother_dimensions(S_shifted))) {
    stop("Original and translated smoother dimensions differ.")
  }

  N<-smoother_dimensions(S_original)[1L]
  observed<-as.numeric(pred_shifted)-as.numeric(pred_original)-SHIFT_K
  smoother_component<-numeric(N)
  normalization_component<-numeric(N)
  maximum_change<-0
  maximum_row_sum_error<-0
  changed_rows<-logical(N)

  for (start in seq.int(1L,N,by=BLOCK_ROWS)) {
    rows<-start:min(start+BLOCK_ROWS-1L,N)
    S_o<-smoother_block(S_original,rows,N)
    S_k<-smoother_block(S_shifted,rows,N)
    difference<-S_k-S_o
    absolute_difference<-abs(difference)

    maximum_change<-max(maximum_change,absolute_difference)
    changed_rows[rows]<-apply(absolute_difference,1L,max)>WEIGHT_TOLERANCE
    maximum_row_sum_error<-max(
      maximum_row_sum_error,
      abs(rowSums(S_o)-1),
      abs(rowSums(S_k)-1)
    )
    smoother_component[rows]<-as.vector(difference%*%Y)
    normalization_component[rows]<-SHIFT_K*(rowSums(S_k)-1)
  }

  closure<-observed-smoother_component-normalization_component
  data.frame(
    Specification=specification,
    Learner=learner,
    Estimator=estimator,
    Smoother=smoother_name,
    `S identical`=identical(S_original,S_shifted),
    `Maximum |S^k-S|`=maximum_change,
    `Share rows changed`=mean(changed_rows),
    `Maximum row-sum error`=maximum_row_sum_error,
    `Prediction residual RMSE`=rmse(observed),
    `Smoother component RMSE`=rmse(smoother_component),
    `Normalization component maximum`=max_abs(normalization_component),
    `Closure maximum`=max_abs(closure),
    check.names=FALSE
  )
}

estimate_row<-function(
    specification,learner,estimator,tau_original,tau_shifted,
    folds_identical,nonoutcome_change){
  data.frame(
    Specification=specification,
    Learner=learner,
    Estimator=estimator,
    `Original estimate`=tau_original,
    `Shifted estimate`=tau_shifted,
    `Rerun change`=tau_shifted-tau_original,
    `Folds identical`=folds_identical,
    `Maximum non-outcome nuisance change`=nonoutcome_change,
    check.names=FALSE
  )
}
```

# Paired fits

```{r run-audit}
estimate_parts<-list()
smoother_parts<-list()
result_index<-1L

for (specification in names(specifications)) {
  X<-specifications[[specification]]

  honest_original<-fit_honest_forest(Y,X)
  honest_shifted<-fit_honest_forest(Y+SHIFT_K,X)
  honest_folds_identical<-identical(
    honest_original$NuPa.hat$cf_mat,
    honest_shifted$NuPa.hat$cf_mat
  )
  if (!honest_folds_identical) {
    stop(sprintf("Honest-forest folds differ for %s.",specification))
  }

  honest_summary_original<-summary(honest_original)
  honest_summary_shifted<-summary(honest_shifted)
  p_o<-honest_original$NuPa.hat$predictions
  p_k<-honest_shifted$NuPa.hat$predictions
  s_o<-honest_original$NuPa.hat$smoothers
  s_k<-honest_shifted$NuPa.hat$smoothers
  honest_nonoutcome_change<-max_abs(c(
    p_k$D.hat-p_o$D.hat,
    p_k$D.hat.z0-p_o$D.hat.z0,
    p_k$D.hat.z1-p_o$D.hat.z1,
    p_k$Z.hat-p_o$Z.hat
  ))

  for (estimator in c("PLR-IV","Wald-AIPW")) {
    estimate_parts[[result_index]]<-estimate_row(
      specification,"Tuned honest forest",estimator,
      get_estimate(honest_summary_original,estimator),
      get_estimate(honest_summary_shifted,estimator),
      honest_folds_identical,honest_nonoutcome_change
    )
    result_index<-result_index+1L
  }

  smoother_parts[[length(smoother_parts)+1L]]<-audit_smoother(
    s_o$S,s_k$S,p_o$Y.hat,p_k$Y.hat,
    specification,"Tuned honest forest","PLR-IV","S"
  )
  smoother_parts[[length(smoother_parts)+1L]]<-audit_smoother(
    s_o$S.z0,s_k$S.z0,p_o$Y.hat.z0,p_k$Y.hat.z0,
    specification,"Tuned honest forest","Wald-AIPW","S.z0"
  )
  smoother_parts[[length(smoother_parts)+1L]]<-audit_smoother(
    s_o$S.z1,s_k$S.z1,p_o$Y.hat.z1,p_k$Y.hat.z1,
    specification,"Tuned honest forest","Wald-AIPW","S.z1"
  )

  rm(honest_original,honest_shifted,honest_summary_original,
     honest_summary_shifted,p_o,p_k,s_o,s_k)
  invisible(gc())

  for (estimator in c("PLR-IV","Wald-AIPW")) {
    ranger_original<-fit_ranger(Y,X,estimator)
    ranger_shifted<-fit_ranger(Y+SHIFT_K,X,estimator)
    ranger_folds_identical<-identical(
      ranger_original$object$smpls,
      ranger_shifted$object$smpls
    )
    if (!ranger_folds_identical) {
      stop(sprintf("Ranger folds differ for %s / %s.",
                   specification,estimator))
    }

    ranger_s_o<-extract_ranger_smoothers(ranger_original)
    ranger_s_k<-extract_ranger_smoothers(ranger_shifted)

    if (estimator=="PLR-IV") {
      predictions_o<-list(
        S=as.numeric(drop(ranger_original$object$predictions$ml_l))
      )
      predictions_k<-list(
        S=as.numeric(drop(ranger_shifted$object$predictions$ml_l))
      )
      nonoutcome_change<-max_abs(c(
        ranger_shifted$object$predictions$ml_m-
          ranger_original$object$predictions$ml_m,
        ranger_shifted$object$predictions$ml_r-
          ranger_original$object$predictions$ml_r
      ))
      smoother_names<-"S"
    } else {
      predictions_o<-list(
        S.z0=as.numeric(drop(ranger_original$object$predictions$ml_g0)),
        S.z1=as.numeric(drop(ranger_original$object$predictions$ml_g1))
      )
      predictions_k<-list(
        S.z0=as.numeric(drop(ranger_shifted$object$predictions$ml_g0)),
        S.z1=as.numeric(drop(ranger_shifted$object$predictions$ml_g1))
      )
      nonoutcome_change<-max_abs(c(
        ranger_shifted$object$predictions$ml_m-
          ranger_original$object$predictions$ml_m,
        ranger_shifted$object$predictions$ml_r0-
          ranger_original$object$predictions$ml_r0,
        ranger_shifted$object$predictions$ml_r1-
          ranger_original$object$predictions$ml_r1
      ))
      smoother_names<-c("S.z0","S.z1")
    }

    estimate_parts[[result_index]]<-estimate_row(
      specification,"Ranger",estimator,
      get_dml_coef(ranger_original$object),
      get_dml_coef(ranger_shifted$object),
      ranger_folds_identical,nonoutcome_change
    )
    result_index<-result_index+1L

    for (smoother_name in smoother_names) {
      smoother_parts[[length(smoother_parts)+1L]]<-audit_smoother(
        ranger_s_o[[smoother_name]],ranger_s_k[[smoother_name]],
        predictions_o[[smoother_name]],predictions_k[[smoother_name]],
        specification,"Ranger",estimator,smoother_name
      )
    }

    rm(ranger_original,ranger_shifted,ranger_s_o,ranger_s_k,
       predictions_o,predictions_k)
    invisible(gc())
  }
}

estimate_results<-bind_rows(estimate_parts)
smoother_results<-bind_rows(smoother_parts)
```

# Results

The first table reports estimate changes continuously and verifies the paired
design.

```{r estimate-table, echo=FALSE}
estimate_display<-estimate_results
estimate_display[c("Original estimate","Shifted estimate")]<-
  lapply(estimate_display[c("Original estimate","Shifted estimate")],
         function(x) formatC(x,format="f",digits=4))
estimate_display$`Rerun change`<-
  formatC(estimate_display$`Rerun change`,format="e",digits=3)
estimate_display$`Maximum non-outcome nuisance change`<-
  formatC(estimate_display$`Maximum non-outcome nuisance change`,
          format="e",digits=2)

estimate_display %>%
  kbl(caption="Estimate-level paired translation results",row.names=FALSE) %>%
  kable_styling(
    bootstrap_options=c("striped","hover","condensed"),
    full_width=TRUE
  ) %>%
  collapse_rows(columns=c(1,2),valign="middle")
```

The second table distinguishes a change in the smoother itself from a failure
of its rows to sum to one. 

```{r smoother-table, echo=FALSE}
closure_max<-max(as.numeric(smoother_results$`Closure maximum`))

display_columns<-c(
  "Specification","Learner","Estimator","Smoother",
  "Maximum |S^k-S|","Prediction residual RMSE",
  "Smoother component RMSE","Normalization component maximum"
)
smoother_display<-smoother_results[,display_columns,drop=FALSE]

diagnostic_columns<-setdiff(
  names(smoother_display),
  c("Specification","Learner","Estimator","Smoother")
)
smoother_display[diagnostic_columns]<-lapply(
  smoother_display[diagnostic_columns],
  function(x) formatC(x,format="e",digits=2)
)

smoother_display %>%
  kbl(caption="Outcome-smoother change and decomposition",row.names=FALSE) %>%
  kable_styling(
    bootstrap_options=c("striped","hover","condensed"),
    full_width=TRUE,
    font_size=10
  ) %>%
  collapse_rows(columns=c(1,2,3),valign="middle") %>%
  footnote(general=paste0(
    "Each observed prediction change decomposes into a smoother-weight term and a ",
    "normalization term. The decomposition closure ",
    "|observed - smoother - normalization| stayed below ",
    formatC(closure_max,format="e",digits=2)," across all cells."
  ))
```


