---
title: "Card (1995) — Replication and Outcome-Weight Diagnostics"
subtitle: "Kappa estimators, Double ML, learner sensitivity, and exact weight checks"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body   { font-size: 15px; color: #2c2c2c; }
h1     { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2     { color: #2c7bb6; margin-top: 28px; }
h3     { color: #444; margin-top: 20px; }

.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  cache   = TRUE,
  autodep = TRUE,
  fig.width  = 12,
  fig.height = 6
)
```

---

# Setup

## Packages & Functions

```{r packages-and-source, include=FALSE, cache=FALSE}
# Core
library(haven)
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(dplyr)
# DML smoother + balance plots
library(OutcomeWeights)
library(cobalt)
library(viridis)
library(gridExtra)
library(ggplot2)
# DoubleML learners
library(DoubleML)
library(mlr3verse)

source(file.path("..", "functions_all.R"))
```

## Data & Design

<div class="note-box">
**Setting.** Card (1995) NLSYM subsample (N = 3,010). Instrument Z = `nearc4` (grew up near a 4-year college). Two binary treatments: **somecol** = 1{educ > 12} and **educ16** = 1{educ ≥ 16}. Outcome: log wages, reported in two units (dollars / cents) only because SUW Table 3 is laid out that way. Two covariate specs: **Card** (exper, expersq, black, south, smsa, smsa66, reg661–668) and **Kitagawa** (black, south, smsa, smsa66, south66).
</div>

```{r data, include=FALSE}
card       <- read_dta("card.dta")

# Sample restriction: N = 3,010
card_clean <- card[!is.na(card$wage) & !is.na(card$educ), ]

# Outcomes in two units: lwage_cnt = lwage + log(100)
card_clean$lwage_cnt <- log(card_clean$wage)         # cents
card_clean$lwage     <- log(card_clean$wage / 100)   # dollars

# Binary treatments
card_clean$somecol <- as.integer(card_clean$educ > 12)
card_clean$educ16  <- as.integer(card_clean$educ >= 16)

# Instrument
card_clean$Z_inst <- as.integer(card_clean$nearc4)

# expersq if missing
if (!"expersq" %in% names(card_clean)) {
  card_clean$expersq <- card_clean$exper^2
}

```

```{r design-matrices, include=FALSE}
# Kappa design matrices — WITH intercept (for logit / CBPS)
X_card_kappa <- model.matrix(
  ~ exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card_clean
)
X_card_df <- as.data.frame(X_card_kappa[, -1, drop = FALSE])

X_kit_kappa <- model.matrix(
  ~ black + south + smsa + smsa66 + south66,
  data = card_clean
)
X_kit_df <- as.data.frame(X_kit_kappa[, -1, drop = FALSE])

# DML design matrices — no intercept, attributes stripped for grf (make_X)
X_dml_card <- make_X(
  ~ 0 + exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card_clean
)
X_dml_kit <- make_X(
  ~ 0 + black + south + smsa + smsa66 + south66,
  data = card_clean
)

# Core vectors
Y_cnt <- card_clean$lwage_cnt   # cents  (used only for the SUW Table 3 layout)
Y_dol <- card_clean$lwage       # dollars (the working outcome everywhere else)
Z     <- card_clean$Z_inst
D1    <- card_clean$somecol
D2    <- card_clean$educ16

```

# 0. Empirical Design Audit

<div class="note-box">
The compact design audit reports treatment uptake, conditional first-stage
strength, and fitted instrument-propensity support. Covariate balance is
examined later using estimator-specific outcome weights.
</div>

```{r design-audit, echo=FALSE}
first_stage_row <- function(label, D, X_df, X_kappa) {
  dat <- data.frame(D = D, Z = Z, X_df, check.names = FALSE)
  fit <- lm(D ~ ., data = dat)
  ct <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC1"))
  p <- logit_mle(Z, X_kappa)
  data.frame(
    Cell = label,
    `Adjusted first stage` = unname(ct["Z", "Estimate"]),
    `HC1 SE` = unname(ct["Z", "Std. Error"]),
    `Robust t²` = unname(ct["Z", "t value"]^2),
    `Min p̂(X)` = min(p),
    `Max p̂(X)` = max(p),
    `Boundary share` = mean(p < 0.05 | p > 0.95),
    check.names = FALSE
  )
}

sample_audit <- rbind(
  data.frame(
    Treatment = "somecol", N = length(D1), `Pr(Z=1)` = mean(Z),
    `Pr(D=1)` = mean(D1), `Pr(D=1 | Z=0)` = mean(D1[Z == 0]),
    `Pr(D=1 | Z=1)` = mean(D1[Z == 1]),
    `Raw uptake gap` = mean(D1[Z == 1]) - mean(D1[Z == 0]),
    check.names = FALSE
  ),
  data.frame(
    Treatment = "educ16", N = length(D2), `Pr(Z=1)` = mean(Z),
    `Pr(D=1)` = mean(D2), `Pr(D=1 | Z=0)` = mean(D2[Z == 0]),
    `Pr(D=1 | Z=1)` = mean(D2[Z == 1]),
    `Raw uptake gap` = mean(D2[Z == 1]) - mean(D2[Z == 0]),
    check.names = FALSE
  )
)

design_audit <- rbind(
  first_stage_row("Card / somecol", D1, X_card_df, X_card_kappa),
  first_stage_row("Card / educ16", D2, X_card_df, X_card_kappa),
  first_stage_row("Kitagawa / somecol", D1, X_kit_df, X_kit_kappa),
  first_stage_row("Kitagawa / educ16", D2, X_kit_df, X_kit_kappa)
)

sample_audit %>%
  kbl(digits = 4, caption = "Sample and unconditional treatment uptake") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13)

design_audit %>%
  kbl(digits = 4,
      caption = "Conditional first-stage strength and estimated instrument overlap") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) %>%
  footnote(general = "Robust t² is the squared HC1 t-statistic for college proximity. It is a descriptive strength diagnostic, not a formal weak-IV-robust test. Boundary share uses 0.05 and 0.95.")
```

---

# 1. Reproduction — SUW Table 3

<div class="note-box">
**Layout.** SUW Table 3 has 8 columns: 2 treatments × 2 specifications × 2 outcome units, split here into two tabs by treatment. Within each tab the four columns are spec × unit. Normalized estimators (τ_cb_u, τ_ml_u, τ_ml_a10) are translation invariant — cents and dollars columns match; unnormalized ones (τ_ml_a, τ_ml_t, τ_ml_a0) do not.
</div>

```{r build-columns, cache=TRUE}
# One full column = 2SLS + all six kappa estimators with analytical SEs.
run_card_column <- function(Y, D, X_kappa, X_df) {
  list(
    tsls  = run_2sls(Y, D, Z, X_df, endog_name = "D"),
    kappa = kappa_analytic_se_all(Y, Z, D, X_kappa)
  )
}

# Treatment 1: somecol (cols 1-4)
col1 <- run_card_column(Y_cnt, D1, X_card_kappa, X_card_df)  # cents,   Card
col2 <- run_card_column(Y_dol, D1, X_card_kappa, X_card_df)  # dollars, Card
col3 <- run_card_column(Y_cnt, D1, X_kit_kappa,  X_kit_df)   # cents,   Kitagawa
col4 <- run_card_column(Y_dol, D1, X_kit_kappa,  X_kit_df)   # dollars, Kitagawa

# Treatment 2: educ16 (cols 5-8)
col5 <- run_card_column(Y_cnt, D2, X_card_kappa, X_card_df)  # cents,   Card
col6 <- run_card_column(Y_dol, D2, X_card_kappa, X_card_df)  # dollars, Card
col7 <- run_card_column(Y_cnt, D2, X_kit_kappa,  X_kit_df)   # cents,   Kitagawa
col8 <- run_card_column(Y_dol, D2, X_kit_kappa,  X_kit_df)   # dollars, Kitagawa

```

```{r table-builder, echo=FALSE}
cell_fmt <- function(est, se, digits = 3) {
  if (is.na(est)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0)
    return(sprintf("%.3f<br><span style='color:#888'>(NA)</span>", est))
  pval  <- 2 * pnorm(-abs(est / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s<br><span style='color:#888'>(%.3f)</span>", est, stars, se)
}

build_treatment_table <- function(cA_cnt, cB_dol, cC_cnt, cD_dol) {
  ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
  labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")

  rows <- list(data.frame(
    Estimator = "2SLS",
    `Card / cents`       = cell_fmt(cA_cnt$tsls$coef, cA_cnt$tsls$se),
    `Card / dollars`     = cell_fmt(cB_dol$tsls$coef, cB_dol$tsls$se),
    `Kitagawa / cents`   = cell_fmt(cC_cnt$tsls$coef, cC_cnt$tsls$se),
    `Kitagawa / dollars` = cell_fmt(cD_dol$tsls$coef, cD_dol$tsls$se),
    check.names = FALSE, stringsAsFactors = FALSE
  ))

  for (i in seq_along(ests)) {
    e <- ests[i]
    rows[[length(rows) + 1]] <- data.frame(
      Estimator = labels[i],
      `Card / cents`       = cell_fmt(cA_cnt$kappa$estimates[e], cA_cnt$kappa$se[e]),
      `Card / dollars`     = cell_fmt(cB_dol$kappa$estimates[e], cB_dol$kappa$se[e]),
      `Kitagawa / cents`   = cell_fmt(cC_cnt$kappa$estimates[e], cC_cnt$kappa$se[e]),
      `Kitagawa / dollars` = cell_fmt(cD_dol$kappa$estimates[e], cD_dol$kappa$se[e]),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }
  do.call(rbind, rows)
}
```

## Results {.tabset}

### Treatment: Some College (educ > 12)

```{r table-somecol}
build_treatment_table(col1, col2, col3, col4) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c"),
      caption = "SUW Table 3 — Treatment: somecol (educ > 12). Effect of some college on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",           1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant). Orange = unnormalized. Card spec: exper, expersq, black, south, smsa, smsa66, reg661–668. Kitagawa spec: black, south, smsa, smsa66, south66. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

### Treatment: College Completion (educ ≥ 16)

```{r table-educ16}
build_treatment_table(col5, col6, col7, col8) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c"),
      caption = "SUW Table 3 — Treatment: educ16 (educ ≥ 16). Effect of college completion on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",           1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant). Orange = unnormalized. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 2. Double ML — PLR-IV & Wald-AIPW (honest forest, tuned vs. untuned)

<div class="note-box">
Each of the four treatment-by-covariate cells is fitted with an untuned honest
forest (`tune.parameters = "none"`) and a tuned honest forest
(`tune.parameters = "all"`), using five-fold cross-fitting. The tuned fits are
used in the downstream learner and eligible-weight comparisons. The untuned
fits are retained only for the direct tuning comparison.
</div>

```{r dml-fit-all, include=FALSE}
# Four treatment-by-covariate cells.
set.seed(42); dml_card_d1_untuned <- dml_with_smoother(
  Y_dol, D1, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_card_d1_tuned <- dml_with_smoother(
  Y_dol, D1, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "all")

set.seed(42); dml_card_d2_untuned <- dml_with_smoother(
  Y_dol, D2, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_card_d2_tuned <- dml_with_smoother(
  Y_dol, D2, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "all")

set.seed(42); dml_kit_d1_untuned <- dml_with_smoother(
  Y_dol, D1, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_kit_d1_tuned <- dml_with_smoother(
  Y_dol, D1, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "all")

set.seed(42); dml_kit_d2_untuned <- dml_with_smoother(
  Y_dol, D2, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_kit_d2_tuned <- dml_with_smoother(
  Y_dol, D2, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "all")

res_card_d1_untuned <- summary(dml_card_d1_untuned)
res_card_d1_tuned   <- summary(dml_card_d1_tuned)
res_card_d2_untuned <- summary(dml_card_d2_untuned)
res_card_d2_tuned   <- summary(dml_card_d2_tuned)
res_kit_d1_untuned  <- summary(dml_kit_d1_untuned)
res_kit_d1_tuned    <- summary(dml_kit_d1_tuned)
res_kit_d2_untuned  <- summary(dml_kit_d2_untuned)
res_kit_d2_tuned    <- summary(dml_kit_d2_tuned)

# Extract PLR-IV and Wald-AIPW outcome weights for every fit.
extract_w <- function(dml_obj) {
  om <- get_outcome_weights(dml_obj)
  check_omega_rows(om)
  list(plriv    = as.vector(om$omega["PLR-IV",    ]),
       waldaipw = as.vector(om$omega["Wald-AIPW", ]))
}
w_card_d1_untuned <- extract_w(dml_card_d1_untuned)
w_card_d1_tuned   <- extract_w(dml_card_d1_tuned)
w_card_d2_untuned <- extract_w(dml_card_d2_untuned)
w_card_d2_tuned   <- extract_w(dml_card_d2_tuned)
w_kit_d1_untuned  <- extract_w(dml_kit_d1_untuned)
w_kit_d1_tuned    <- extract_w(dml_kit_d1_tuned)
w_kit_d2_untuned  <- extract_w(dml_kit_d2_untuned)
w_kit_d2_tuned    <- extract_w(dml_kit_d2_tuned)

# Headline aliases use tuned forests.
dml_card_d1 <- dml_card_d1_tuned
dml_card_d2 <- dml_card_d2_tuned
dml_kit_d1  <- dml_kit_d1_tuned
dml_kit_d2  <- dml_kit_d2_tuned
res_card_d1 <- res_card_d1_tuned
res_card_d2 <- res_card_d2_tuned
res_kit_d1  <- res_kit_d1_tuned
res_kit_d2  <- res_kit_d2_tuned
w_card_d1   <- w_card_d1_tuned
w_card_d2   <- w_card_d2_tuned
w_kit_d1    <- w_kit_d1_tuned
w_kit_d2    <- w_kit_d2_tuned
```

## Estimates {.tabset}

```{r dml-cell-display, echo=FALSE}
show_dml_cell <- function(res, wlist, Y, cap) {
  ids <- c(
    check_weight_identity(wlist$plriv,    Y, get_estimate(res, "PLR-IV")),
    check_weight_identity(wlist$waldaipw, Y, get_estimate(res, "Wald-AIPW"))
  )
  df <- as.data.frame(res[c("PLR-IV", "Wald-AIPW"), , drop = FALSE])
  df$Estimator  <- rownames(df)
  df$`ω′Y reproduces τ̂` <- as.character(ids)
  df <- df[, c("Estimator",
               setdiff(names(df), c("Estimator", "ω′Y reproduces τ̂")),
               "ω′Y reproduces τ̂")]
  df %>%
    kbl(caption = cap, digits = 4, row.names = FALSE,
        align = c("l", rep("r", ncol(df) - 2), "c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "Numerical reproduction follows Knaus's all.equal() check with R's default tolerance.")
}
```

### Card — somecol

```{r dml-card-d1-untuned}
show_dml_cell(res_card_d1_untuned, w_card_d1_untuned, Y_dol,
              "Card spec, somecol — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-card-d1-tuned}
show_dml_cell(res_card_d1_tuned, w_card_d1_tuned, Y_dol,
              "Card spec, somecol — tuned honest forest (tune.parameters = \"all\")")
```

### Card — educ16

```{r dml-card-d2-untuned}
show_dml_cell(res_card_d2_untuned, w_card_d2_untuned, Y_dol,
              "Card spec, educ16 — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-card-d2-tuned}
show_dml_cell(res_card_d2_tuned, w_card_d2_tuned, Y_dol,
              "Card spec, educ16 — tuned honest forest (tune.parameters = \"all\")")
```

### Kitagawa — somecol

```{r dml-kit-d1-untuned}
show_dml_cell(res_kit_d1_untuned, w_kit_d1_untuned, Y_dol,
              "Kitagawa spec, somecol — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-kit-d1-tuned}
show_dml_cell(res_kit_d1_tuned, w_kit_d1_tuned, Y_dol,
              "Kitagawa spec, somecol — tuned honest forest (tune.parameters = \"all\")")
```

### Kitagawa — educ16

```{r dml-kit-d2-untuned}
show_dml_cell(res_kit_d2_untuned, w_kit_d2_untuned, Y_dol,
              "Kitagawa spec, educ16 — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-kit-d2-tuned}
show_dml_cell(res_kit_d2_tuned, w_kit_d2_tuned, Y_dol,
              "Kitagawa spec, educ16 — tuned honest forest (tune.parameters = \"all\")")
```

<div class="note-box">
Within each cell, the untuned and tuned fits can be compared in their point
estimates and standard errors. Numerical reproduction is a separate
representation gate; it is not a tuning-quality or normalization criterion.
</div>

## DML vs. Kappa — Point Estimates

```{r dml-summary-table}
# Kappa estimates for all four cells.
kappa_card_d1 <- kappa_analytic_se_all(Y_dol, Z, D1, X_card_kappa)
kappa_card_d2 <- kappa_analytic_se_all(Y_dol, Z, D2, X_card_kappa)
kappa_kit_d1  <- kappa_analytic_se_all(Y_dol, Z, D1, X_kit_kappa)
kappa_kit_d2  <- kappa_analytic_se_all(Y_dol, Z, D2, X_kit_kappa)

cell <- function(coef, se) {
  if (is.na(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

dml_cell <- function(res, est) {
  coef_val <- get_estimate(res, est)
  se_val   <- if ("SE" %in% colnames(res)) as.numeric(res[est, "SE"])
              else if ("Std. Error" %in% colnames(res)) as.numeric(res[est, "Std. Error"])
              else NA_real_
  cell(coef_val, se_val)
}

dml_rows <- rbind(
  data.frame(Estimator = "PLR-IV (DML; tuned forest)",
             `Card / somecol`     = dml_cell(res_card_d1, "PLR-IV"),
             `Card / educ16`      = dml_cell(res_card_d2, "PLR-IV"),
             `Kitagawa / somecol` = dml_cell(res_kit_d1,  "PLR-IV"),
             `Kitagawa / educ16`  = dml_cell(res_kit_d2,  "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (DML; tuned forest)",
             `Card / somecol`     = dml_cell(res_card_d1, "Wald-AIPW"),
             `Card / educ16`      = dml_cell(res_card_d2, "Wald-AIPW"),
             `Kitagawa / somecol` = dml_cell(res_kit_d1,  "Wald-AIPW"),
             `Kitagawa / educ16`  = dml_cell(res_kit_d2,  "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE)
)

ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")
kappa_rows <- do.call(rbind, lapply(seq_along(ests), function(i) {
  e <- ests[i]
  data.frame(
    Estimator = labels[i],
    `Card / somecol`     = cell(kappa_card_d1$estimates[e], kappa_card_d1$se[e]),
    `Card / educ16`      = cell(kappa_card_d2$estimates[e], kappa_card_d2$se[e]),
    `Kitagawa / somecol` = cell(kappa_kit_d1$estimates[e],  kappa_kit_d1$se[e]),
    `Kitagawa / educ16`  = cell(kappa_kit_d2$estimates[e],  kappa_kit_d2$se[e]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}))

rbind(dml_rows, kappa_rows) %>%
  kbl(align = c("l","c","c","c","c"),
      caption = "DML vs. Kappa — Point Estimates (SE), Y = log wages (dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  add_header_above(c(" " = 1, "Card covariates" = 2, "Kitagawa covariates" = 2)) %>%
  pack_rows("Double ML", 1, 2) %>%
  pack_rows("Kappa — normalized", 3, 5) %>%
  pack_rows("Kappa — unnormalized", 6, 8) %>%
  row_spec(1:2, background = "#eef6ff") %>%
  row_spec(6:8, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE). Blue = DML. Orange = unnormalized kappa. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 3. Learner Comparison — PLR-IV and Wald-AIPW

<div class="note-box">
This section holds the five-fold DoubleML procedure and the four Card design
cells fixed while varying the nuisance learners. Linear and Ranger models are
fitted here. Fold-specifically tuned XGBoost results are imported from the two
verified tuning exports and are not retuned in this notebook. The separate
XGBoost sensitivity notebook compares tuned and untuned specifications.
</div>

```{r ext-setup, include=FALSE}
# DoubleMLData objects — one per cell. d and z must be numeric 0/1.
make_dml_data <- function(D, X_dml) {
  df <- as.data.frame(X_dml)
  df$y <- as.numeric(Y_dol)
  df$d <- as.numeric(D)
  df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X_dml))
}

dat_card_d1 <- make_dml_data(D1, X_dml_card)
dat_card_d2 <- make_dml_data(D2, X_dml_card)
dat_kit_d1  <- make_dml_data(D1, X_dml_kit)
dat_kit_d2  <- make_dml_data(D2, X_dml_kit)
```

```{r ext-fit-helpers, echo=FALSE}
ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))
ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))

omega_vec <- function(om, estimator) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && estimator %in% rownames(w)) {
    return(as.vector(w[estimator, ]))
  }
  as.vector(w[1, ])
}

fit_iivm <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLIIVM$new(data = dml_data,
                          ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5, score = "LATE")
  invisible(capture.output(obj$fit(store_models = TRUE, store_predictions = TRUE)))
  om  <- get_outcome_weights(object = obj, dml_data = dml_data)
  w   <- omega_vec(om, "Wald-AIPW")
  ok  <- check_weight_identity(w, Y, get_dml_coef(obj))
  list(obj = obj, w = w, ok = ok)
}

# Fit one PLIV model with the same fold count and seed.
fit_pliv <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLPLIV$new(data = dml_data,
                          ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5)
  invisible(capture.output(obj$fit(store_models = TRUE, store_predictions = TRUE)))
  om <- get_outcome_weights(object = obj, dml_data = dml_data)
  w  <- omega_vec(om, "PLR-IV")
  coef <- NA_real_
  invisible(capture.output(coef <- get_dml_coef(obj)))
  ok <- check_weight_identity(w, Y, coef)
  list(obj = obj, w = w, ok = ok)
}

safe_coef <- function(f) {
  if (is.null(f$obj)) return(NA_real_)
  out <- NA_real_
  invisible(capture.output(out <- get_dml_coef(f$obj)))
  as.numeric(out)
}

safe_se <- function(f) {
  if (is.null(f$obj)) return(NA_real_)
  out <- NA_real_
  invisible(capture.output(out <- get_dml_se(f$obj)))
  as.numeric(out)
}

learner_estimate_table <- function(pliv, iivm, cap) {
  data.frame(
    Estimator = c("PLR-IV", "Wald-AIPW"),
    Estimate  = round(c(safe_coef(pliv), safe_coef(iivm)), 4),
    SE        = round(c(safe_se(pliv), safe_se(iivm)), 4),
    `ω′Y reproduces τ̂` = c(pliv$ok, iivm$ok),
    check.names = FALSE
  ) %>%
    kbl(caption = cap, align = c("l","r","r","c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "Numerical reproduction follows Knaus's all.equal() check with R's default tolerance.")
}
```

## Estimates {.tabset}

### Card — somecol

```{r ext-fit-card-d1, cache=TRUE}
lm_pliv_card_d1 <- fit_pliv(dat_card_d1, ml_lm_pliv(), Y_dol)
lm_iivm_card_d1 <- fit_iivm(dat_card_d1, ml_lm_iivm(), Y_dol)
rf_pliv_card_d1 <- fit_pliv(dat_card_d1, ml_rf_pliv(), Y_dol)
rf_iivm_card_d1 <- fit_iivm(dat_card_d1, ml_rf_iivm(), Y_dol)
learner_estimate_table(lm_pliv_card_d1, lm_iivm_card_d1,
                       "Card / somecol — Linear + Logistic")
learner_estimate_table(rf_pliv_card_d1, rf_iivm_card_d1,
                       "Card / somecol — Ranger")
```

### Card — educ16

```{r ext-fit-card-d2, cache=TRUE}
lm_pliv_card_d2 <- fit_pliv(dat_card_d2, ml_lm_pliv(), Y_dol)
lm_iivm_card_d2 <- fit_iivm(dat_card_d2, ml_lm_iivm(), Y_dol)
rf_pliv_card_d2 <- fit_pliv(dat_card_d2, ml_rf_pliv(), Y_dol)
rf_iivm_card_d2 <- fit_iivm(dat_card_d2, ml_rf_iivm(), Y_dol)
learner_estimate_table(lm_pliv_card_d2, lm_iivm_card_d2,
                       "Card / educ16 — Linear + Logistic")
learner_estimate_table(rf_pliv_card_d2, rf_iivm_card_d2,
                       "Card / educ16 — Ranger")
```

### Kitagawa — somecol

```{r ext-fit-kit-d1, cache=TRUE}
lm_pliv_kit_d1 <- fit_pliv(dat_kit_d1, ml_lm_pliv(), Y_dol)
lm_iivm_kit_d1 <- fit_iivm(dat_kit_d1, ml_lm_iivm(), Y_dol)
rf_pliv_kit_d1 <- fit_pliv(dat_kit_d1, ml_rf_pliv(), Y_dol)
rf_iivm_kit_d1 <- fit_iivm(dat_kit_d1, ml_rf_iivm(), Y_dol)
learner_estimate_table(lm_pliv_kit_d1, lm_iivm_kit_d1,
                       "Kitagawa / somecol — Linear + Logistic")
learner_estimate_table(rf_pliv_kit_d1, rf_iivm_kit_d1,
                       "Kitagawa / somecol — Ranger")
```

### Kitagawa — educ16

```{r ext-fit-kit-d2, cache=TRUE}
lm_pliv_kit_d2 <- fit_pliv(dat_kit_d2, ml_lm_pliv(), Y_dol)
lm_iivm_kit_d2 <- fit_iivm(dat_kit_d2, ml_lm_iivm(), Y_dol)
rf_pliv_kit_d2 <- fit_pliv(dat_kit_d2, ml_rf_pliv(), Y_dol)
rf_iivm_kit_d2 <- fit_iivm(dat_kit_d2, ml_rf_iivm(), Y_dol)
learner_estimate_table(lm_pliv_kit_d2, lm_iivm_kit_d2,
                       "Kitagawa / educ16 — Linear + Logistic")
learner_estimate_table(rf_pliv_kit_d2, rf_iivm_kit_d2,
                       "Kitagawa / educ16 — Ranger")
```

### XGBoost

```{r ext-xgb-import, include=FALSE}
validate_xgb_export <- function(x, expected_class) {
  required <- c("estimate_comparison", "nuisance_losses",
                "weight_diagnostics", "omega_vectors", "identity_checks",
                "settings", "sample_info")
  if (!all(required %in% names(x))) {
    stop("XGBoost export has an incomplete schema.")
  }
  if (!identical(x$settings$dml_class, expected_class)) {
    stop("XGBoost export has the wrong DoubleML class.")
  }
  if (!isTRUE(x$settings$tune_on_folds)) {
    stop("XGBoost export is not tune_on_folds=TRUE.")
  }
  required_cells <- c(
    "Card / some college", "Card / college completion",
    "Kitagawa / some college", "Kitagawa / college completion"
  )
  if (!all(required_cells %in% names(x$omega_vectors))) {
    stop("XGBoost export does not contain all four Card design cells.")
  }
  info <- x$sample_info
  if (!identical(as.integer(info$n), as.integer(length(Z))) ||
      !isTRUE(all.equal(as.numeric(info$mean_Z), mean(Z), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(info$somecol_mean_D), mean(D1), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(info$educ16_mean_D), mean(D2), tolerance = 1e-12))) {
    stop("XGBoost export metadata do not match the current Card data.")
  }
  invisible(TRUE)
}

xgb_wald_path <- "card_xgb_tuning_export.rds"
xgb_pliv_path <- "card_xgb_pliv_tuning_export.rds"
if (!file.exists(xgb_wald_path) || !file.exists(xgb_pliv_path)) {
  stop("Run both Card XGBoost tuning notebooks before card_test.Rmd.")
}
wald_xgb <- readRDS(xgb_wald_path)
pliv_xgb <- readRDS(xgb_pliv_path)
validate_xgb_export(wald_xgb, "DoubleMLIIVM")
validate_xgb_export(pliv_xgb, "DoubleMLPLIV")
```

```{r ext-xgb-export-helpers, echo=FALSE}
get_xgb_estimate_cell <- function(export, cell_name, variant) {
  tab <- export$estimate_comparison
  keep <- if (variant == "untuned") {
    grepl("untuned$", tab$Learner)
  } else {
    grepl("tune_on_folds=TRUE", tab$Learner, fixed = TRUE)
  }
  hit <- tab$Cell == cell_name & keep
  if (sum(hit) != 1L) return("—")
  as.character(tab[["Estimate (SE)"]][hit][1])
}

get_xgb_bundle <- function(export, cell_name, variant) {
  w <- tryCatch(export$omega_vectors[[cell_name]][[variant]],
                error = function(e) NULL)
  s <- tryCatch(export$fit_summaries[[cell_name]][[variant]],
                error = function(e) NULL)
  if (is.null(s)) {
    wd <- export$weight_diagnostics
    keep <- if (variant == "untuned") {
      grepl("untuned$", wd$Learner)
    } else {
      grepl(" tuned$", wd$Learner)
    }
    hit <- wd$Cell == cell_name & keep
    estimate <- if (sum(hit) == 1L) as.numeric(wd$Estimate[hit]) else NA_real_
    txt <- get_xgb_estimate_cell(export, cell_name, variant)
    nums <- regmatches(txt, gregexpr("-?[0-9]+\\.?[0-9]*", txt))[[1]]
    se <- if (length(nums) >= 2) as.numeric(nums[2]) else NA_real_
  } else {
    estimate <- as.numeric(s$coef)
    se <- as.numeric(s$se)
  }
  gap <- if (is.null(w) || !is.finite(estimate)) NA_real_ else
    sum(w * Y_dol) - estimate
  list(
    w = w, coef = estimate, se = se, gap = gap,
    ok = if (is.finite(gap)) check_weight_identity(w, Y_dol, estimate) else NA
  )
}

xgb_cells <- c(
  card_d1 = "Card / some college",
  card_d2 = "Card / college completion",
  kit_d1 = "Kitagawa / some college",
  kit_d2 = "Kitagawa / college completion"
)

xgb_card_d1 <- get_xgb_bundle(wald_xgb, xgb_cells["card_d1"], "tuned")
xgb_card_d2 <- get_xgb_bundle(wald_xgb, xgb_cells["card_d2"], "tuned")
xgb_kit_d1  <- get_xgb_bundle(wald_xgb, xgb_cells["kit_d1"], "tuned")
xgb_kit_d2  <- get_xgb_bundle(wald_xgb, xgb_cells["kit_d2"], "tuned")

xgb_pliv_card_d1 <- get_xgb_bundle(pliv_xgb, xgb_cells["card_d1"], "tuned")
xgb_pliv_card_d2 <- get_xgb_bundle(pliv_xgb, xgb_cells["card_d2"], "tuned")
xgb_pliv_kit_d1  <- get_xgb_bundle(pliv_xgb, xgb_cells["kit_d1"], "tuned")
xgb_pliv_kit_d2  <- get_xgb_bundle(pliv_xgb, xgb_cells["kit_d2"], "tuned")
```

```{r ext-xgb-comparison, echo=FALSE}
xgb_import_table <- do.call(rbind, lapply(list(
  list("PLR-IV", xgb_cells["card_d1"], pliv_xgb),
  list("PLR-IV", xgb_cells["card_d2"], pliv_xgb),
  list("PLR-IV", xgb_cells["kit_d1"], pliv_xgb),
  list("PLR-IV", xgb_cells["kit_d2"], pliv_xgb),
  list("Wald-AIPW", xgb_cells["card_d1"], wald_xgb),
  list("Wald-AIPW", xgb_cells["card_d2"], wald_xgb),
  list("Wald-AIPW", xgb_cells["kit_d1"], wald_xgb),
  list("Wald-AIPW", xgb_cells["kit_d2"], wald_xgb)
), function(x) {
  t <- get_xgb_bundle(x[[3]], x[[2]], "tuned")
  data.frame(
    Estimator = x[[1]], Cell = unname(x[[2]]),
    `Tuned estimate (SE)` = get_xgb_estimate_cell(x[[3]], x[[2]], "tuned"),
    `ω′Y reproduces τ̂` = ifelse(isTRUE(t$ok), "TRUE", "FALSE"),
    `Absolute gap` = sprintf("%.2e", abs(t$gap)),
    check.names = FALSE
  )
}))

xgb_import_table %>%
  kbl(caption = "Imported fold-specifically tuned XGBoost estimates") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 12) %>%
  footnote(general = "The separate XGBoost sensitivity notebook reports the untuned comparison. Numerical reproduction follows Knaus's all.equal() check with R's default tolerance and is distinct from tuning quality.")
```

## Comparison Table

```{r ext-comparison}
ext_cell <- function(coef, se) {
  if (is.na(coef) || !is.finite(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

learner_row <- function(label, f_card_d1, f_card_d2, f_kit_d1, f_kit_d2) {
  data.frame(
    Estimator = label,
    `Card / somecol`     = ext_cell(safe_coef(f_card_d1), safe_se(f_card_d1)),
    `Card / educ16`      = ext_cell(safe_coef(f_card_d2), safe_se(f_card_d2)),
    `Kitagawa / somecol` = ext_cell(safe_coef(f_kit_d1),  safe_se(f_kit_d1)),
    `Kitagawa / educ16`  = ext_cell(safe_coef(f_kit_d2),  safe_se(f_kit_d2)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

kappa_ref_row <- function(label, key) {
  data.frame(
    Estimator = label,
    `Card / somecol`     = ext_cell(kappa_card_d1$estimates[key], kappa_card_d1$se[key]),
    `Card / educ16`      = ext_cell(kappa_card_d2$estimates[key], kappa_card_d2$se[key]),
    `Kitagawa / somecol` = ext_cell(kappa_kit_d1$estimates[key],  kappa_kit_d1$se[key]),
    `Kitagawa / educ16`  = ext_cell(kappa_kit_d2$estimates[key],  kappa_kit_d2$se[key]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

xgb_row <- function(label, a, b, c, d) {
  data.frame(
    Estimator = label,
    `Card / somecol` = ext_cell(a$coef, a$se),
    `Card / educ16` = ext_cell(b$coef, b$se),
    `Kitagawa / somecol` = ext_cell(c$coef, c$se),
    `Kitagawa / educ16` = ext_cell(d$coef, d$se),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

rbind(
  learner_row("PLR-IV (OLS)", lm_pliv_card_d1, lm_pliv_card_d2, lm_pliv_kit_d1, lm_pliv_kit_d2),
  learner_row("PLR-IV (ranger)", rf_pliv_card_d1, rf_pliv_card_d2, rf_pliv_kit_d1, rf_pliv_kit_d2),
  xgb_row("PLR-IV (XGBoost tuned)", xgb_pliv_card_d1, xgb_pliv_card_d2, xgb_pliv_kit_d1, xgb_pliv_kit_d2),
  learner_row("Wald-AIPW (linear+logit)", lm_iivm_card_d1, lm_iivm_card_d2, lm_iivm_kit_d1, lm_iivm_kit_d2),
  learner_row("Wald-AIPW (ranger)", rf_iivm_card_d1, rf_iivm_card_d2, rf_iivm_kit_d1, rf_iivm_kit_d2),
  xgb_row("Wald-AIPW (XGBoost tuned)", xgb_card_d1, xgb_card_d2, xgb_kit_d1, xgb_kit_d2),
  kappa_ref_row("τ_cb_u (kappa, ref)",  "tau_cb_u"),
  kappa_ref_row("τ_ml_u (kappa, ref)",  "tau_ml_u")
) %>%
  kbl(align = c("l","c","c","c","c"),
      caption = "Learner Comparison — DoubleML vs. Kappa (Y = log wages, dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  add_header_above(c(" " = 1, "Card covariates" = 2, "Kitagawa covariates" = 2)) %>%
  pack_rows("PLR-IV", 1, 3) %>%
  pack_rows("Wald-AIPW", 4, 6) %>%
  pack_rows("Kappa reference", 7, 8) %>%
  row_spec(1:6, background = "#eef6ff") %>%
  footnote(general = "Each cell reports estimate (SE). XGBoost rows import the stored nested-tuning fits. Their reconstructed outcome weights are assessed separately.")
```

---

# 4. Weight Diagnostics

```{r diag-builders, echo=FALSE}
kw_card_d1 <- kappa_weights_bundle(Z, D1, X_card_kappa)
kw_card_d2 <- kappa_weights_bundle(Z, D2, X_card_kappa)
kw_kit_d1  <- kappa_weights_bundle(Z, D1, X_kit_kappa)
kw_kit_d2  <- kappa_weights_bundle(Z, D2, X_kit_kappa)

candidate <- function(w, estimate = NA_real_, ok = TRUE, source = "verified") {
  list(w = w, estimate = as.numeric(estimate), ok = isTRUE(ok), source = source)
}

cell_candidates <- function(w_forest, res_forest,
                            lm_pliv, lm_iivm, rf_pliv, rf_iivm,
                            xgb_pliv, xgb_iivm, kw) {
  pliv_forest_est <- get_estimate(res_forest, "PLR-IV")
  wald_forest_est <- get_estimate(res_forest, "Wald-AIPW")
  list(
    "PLR-IV, tuned forest" = candidate(
      w_forest$plriv, pliv_forest_est,
      check_weight_identity(w_forest$plriv, Y_dol, pliv_forest_est)
    ),
    "Wald-AIPW, tuned forest" = candidate(
      w_forest$waldaipw, wald_forest_est,
      check_weight_identity(w_forest$waldaipw, Y_dol, wald_forest_est)
    ),
    "PLR-IV, OLS" = candidate(
      lm_pliv$w, safe_coef(lm_pliv), lm_pliv$ok
    ),
    "Wald-AIPW, OLS/logit" = candidate(
      lm_iivm$w, safe_coef(lm_iivm), lm_iivm$ok
    ),
    "PLR-IV, ranger" = candidate(
      rf_pliv$w, safe_coef(rf_pliv), rf_pliv$ok
    ),
    "Wald-AIPW, ranger" = candidate(
      rf_iivm$w, safe_coef(rf_iivm), rf_iivm$ok
    ),
    "PLR-IV, tuned XGBoost" = candidate(
      xgb_pliv$w, xgb_pliv$coef, xgb_pliv$ok, "reconstructed"
    ),
    "Wald-AIPW, tuned XGBoost" = candidate(
      xgb_iivm$w, xgb_iivm$coef, xgb_iivm$ok, "reconstructed"
    ),
    "τ_cb_u (kappa, CBPS)" = candidate(kw$tau_cb_u, source = "algebraic"),
    "τ_ml_u (kappa, MLE)" = candidate(kw$tau_ml_u, source = "algebraic"),
    "τ_ml_a10 (kappa, MLE)" = candidate(kw$tau_ml_a10, source = "algebraic"),
    "τ_ml_a (kappa, MLE)" = candidate(kw$tau_ml_a, source = "algebraic"),
    "τ_ml_t (kappa, MLE)" = candidate(kw$tau_ml_t, source = "algebraic"),
    "τ_ml_a0 (kappa, MLE)" = candidate(kw$tau_ml_a0, source = "algebraic")
  )
}

candidate_cells <- list(
  "Card / some college" = list(
    D = D1,
    candidates = cell_candidates(
      w_card_d1_tuned, res_card_d1_tuned,
      lm_pliv_card_d1, lm_iivm_card_d1, rf_pliv_card_d1, rf_iivm_card_d1,
      xgb_pliv_card_d1, xgb_card_d1, kw_card_d1
    )
  ),
  "Card / college completion" = list(
    D = D2,
    candidates = cell_candidates(
      w_card_d2_tuned, res_card_d2_tuned,
      lm_pliv_card_d2, lm_iivm_card_d2, rf_pliv_card_d2, rf_iivm_card_d2,
      xgb_pliv_card_d2, xgb_card_d2, kw_card_d2
    )
  ),
  "Kitagawa / some college" = list(
    D = D1,
    candidates = cell_candidates(
      w_kit_d1_tuned, res_kit_d1_tuned,
      lm_pliv_kit_d1, lm_iivm_kit_d1, rf_pliv_kit_d1, rf_iivm_kit_d1,
      xgb_pliv_kit_d1, xgb_kit_d1, kw_kit_d1
    )
  ),
  "Kitagawa / college completion" = list(
    D = D2,
    candidates = cell_candidates(
      w_kit_d2_tuned, res_kit_d2_tuned,
      lm_pliv_kit_d2, lm_iivm_kit_d2, rf_pliv_kit_d2, rf_iivm_kit_d2,
      xgb_pliv_kit_d2, xgb_kit_d2, kw_kit_d2
    )
  )
)

eligible_weight_list <- function(cell) {
  eligible <- Filter(function(x) isTRUE(x$ok), cell$candidates)
  setNames(lapply(eligible, function(x) x$w), names(eligible))
}

diagnostic_cells <- lapply(candidate_cells, function(cell) {
  list(D = cell$D, w = eligible_weight_list(cell))
})
```

## Detailed Outcome-Weight Statistics {.tabset}

<div class="note-box">
**Knaus summary.** The tables report Overall / Treated / Control statistics.
`Neg. share` is an observation-count share and is distinct from the
absolute-mass sign-cancellation measure reported in the extension below. To
avoid repeating the tuning comparison, the honest-forest entries use the
tuned fits. Only algebraic or numerically verified outcome-weight
representations are included.
</div>

```{r detailed-weight-helpers, echo=FALSE}
detailed_weight_table <- function(cell, caption) {
  ests <- names(cell$w)
  df <- do.call(rbind, lapply(ests, function(nm) {
    weight_bridge(cell$w[[nm]], cell$D, nm)
  }))
  disp <- df[, c(
    "Scope", "Min", "Max", "Neg_share", "Sum_top10", "Sum_w", "Sum_abs_w"
  )]

  out <- disp %>%
    kbl(
      caption = caption, row.names = FALSE,
      col.names = c("Scope", "Min", "Max", "Neg. share", "ΣTop10%", "Σω", "Σ|ω|"),
      align = c("l", rep("r", 6))
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = TRUE, font_size = 12
    ) %>%
    add_header_above(c(" " = 1, "Knaus summary() statistics" = 6))

  for (i in seq_along(ests)) {
    out <- pack_rows(out, ests[i], (i - 1) * 3 + 1, i * 3)
  }

  out %>%
    footnote(general = paste(
      "Control rows use the sign-flipped −ω convention.",
      "Only exact or algebraic representations enter this table.",
      "Knaus's 10^-4 epsilon is a display convention for small individual weights,",
      "not an equality or normalization tolerance."
    ))
}
```

### Card — some college

```{r detailed-weight-card-d1}
detailed_weight_table(
  diagnostic_cells[["Card / some college"]],
  "Detailed outcome-weight statistics — Card, some college"
)
```

### Card — college completion

```{r detailed-weight-card-d2}
detailed_weight_table(
  diagnostic_cells[["Card / college completion"]],
  "Detailed outcome-weight statistics — Card, college completion"
)
```

### Kitagawa — some college

```{r detailed-weight-kit-d1}
detailed_weight_table(
  diagnostic_cells[["Kitagawa / some college"]],
  "Detailed outcome-weight statistics — Kitagawa, some college"
)
```

### Kitagawa — college completion

```{r detailed-weight-kit-d2}
detailed_weight_table(
  diagnostic_cells[["Kitagawa / college completion"]],
  "Detailed outcome-weight statistics — Kitagawa, college completion"
)
```

## ESS and Sign-Cancellation Extension {.tabset}

<div class="note-box">
**Thesis extension.** Modified ESS is shown separately for treated and control
observations, followed by the value relative to each observed arm size.
`Opposite-sign mass T/C` is the percentage of absolute mass with the opposite
orientation within each arm. It measures sign cancellation rather than the
number of negative observations. In an IV design, signs are also shaped by the
residualized instrument, so a high value is not by itself evidence of an
invalid estimator.
</div>

```{r weight-extension-helpers, echo=FALSE}
weight_extension_table <- function(cell, caption) {
  fmt_pair <- function(a, b, digits) {
    if (any(!is.finite(c(a, b)))) return(NA_character_)
    paste0(
      formatC(a, format = "f", digits = digits), " / ",
      formatC(b, format = "f", digits = digits)
    )
  }

  df <- do.call(rbind, lapply(names(cell$w), function(nm) {
    x <- outcome_weight_diagnostics(cell$w[[nm]], cell$D, nm)
    data.frame(
      Estimator = nm,
      `ESS mod T/C` = fmt_pair(x$ESS_mod_T, x$ESS_mod_C, 0),
      `ESS mod/N T/C` = fmt_pair(
        x$ESS_mod_T / sum(cell$D == 1),
        x$ESS_mod_C / sum(cell$D == 0), 3
      ),
      `Opposite-sign mass T/C (%)` = fmt_pair(
        100 * x$Wrong_mass_T, 100 * x$Wrong_mass_C, 1
      ),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }))

  df %>%
    kbl(caption = caption, row.names = FALSE, align = c("l", "c", "c", "c")) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE, font_size = 12
    ) %>%
    footnote(general = paste(
      "T/C denotes treated / sign-oriented control.",
      "Only exact or algebraic representations enter this extension."
    ))
}
```

### Card — some college

```{r weight-extension-card-d1}
weight_extension_table(
  diagnostic_cells[["Card / some college"]],
  "Modified ESS and sign cancellation — Card, some college"
)
```

### Card — college completion

```{r weight-extension-card-d2}
weight_extension_table(
  diagnostic_cells[["Card / college completion"]],
  "Modified ESS and sign cancellation — Card, college completion"
)
```

### Kitagawa — some college

```{r weight-extension-kit-d1}
weight_extension_table(
  diagnostic_cells[["Kitagawa / some college"]],
  "Modified ESS and sign cancellation — Kitagawa, some college"
)
```

### Kitagawa — college completion

```{r weight-extension-kit-d2}
weight_extension_table(
  diagnostic_cells[["Kitagawa / college completion"]],
  "Modified ESS and sign cancellation — Kitagawa, college completion"
)
```

---

# 5. Love Plots

<div class="note-box">
Panels report absolute mean differences. Continuous covariates are standardized;
binary covariates remain on the raw 0/1 scale.
</div>

```{r love-helper, echo=FALSE}
love_from_candidates <- function(cell_name, estimator_names, D, X) {
  candidates <- candidate_cells[[cell_name]]$candidates
  Filter(Negate(is.null), lapply(estimator_names, function(estimator) {
    x <- candidates[[estimator]]
    if (is.null(x) || !isTRUE(x$ok)) return(NULL)
    make_love(estimator, x$w, D, X)
  }))
}

headline_love_estimators <- c(
  "PLR-IV, tuned forest", "Wald-AIPW, tuned forest",
  "τ_cb_u (kappa, CBPS)", "τ_ml_u (kappa, MLE)",
  "τ_ml_a10 (kappa, MLE)", "τ_ml_a (kappa, MLE)",
  "τ_ml_t (kappa, MLE)", "τ_ml_a0 (kappa, MLE)"
)

learner_love_estimators <- c(
  "PLR-IV, OLS", "Wald-AIPW, OLS/logit",
  "PLR-IV, ranger", "Wald-AIPW, ranger",
  "τ_cb_u (kappa, CBPS)", "τ_ml_u (kappa, MLE)"
)
```

## DML vs. Kappa {.tabset}

### Card — somecol

```{r love-card-d1, fig.height=9, fig.width=13}
grid.arrange(
  grobs = love_from_candidates(
    "Card / some college", headline_love_estimators, D1, X_dml_card
  ),
  nrow = 2, top = "Love Plots — Card spec, somecol: eligible weights"
)
```

### Card — college completion

```{r love-card-d2, fig.height=9, fig.width=13}
grid.arrange(
  grobs = love_from_candidates(
    "Card / college completion", headline_love_estimators, D2, X_dml_card
  ),
  nrow = 2, top = "Love Plots — Card spec, college completion: eligible weights"
)
```

### Kitagawa — somecol

```{r love-kit-d1, fig.height=9, fig.width=13}
grid.arrange(
  grobs = love_from_candidates(
    "Kitagawa / some college", headline_love_estimators, D1, X_dml_kit
  ),
  nrow = 2, top = "Love Plots — Kitagawa spec, somecol: eligible weights"
)
```

### Kitagawa — college completion

```{r love-kit-d2, fig.height=9, fig.width=13}
grid.arrange(
  grobs = love_from_candidates(
    "Kitagawa / college completion", headline_love_estimators, D2, X_dml_kit
  ),
  nrow = 2, top = "Love Plots — Kitagawa spec, college completion: eligible weights"
)
```

## Learners vs. Kappa {.tabset}

### Card — somecol (learners + kappa)

```{r ext-love-card-d1, fig.height=7.5, fig.width=14}
grid.arrange(
  grobs = love_from_candidates(
    "Card / some college", learner_love_estimators, D1, X_dml_card
  ),
  nrow = 2, top = "Love Plots — eligible learners vs. kappa | Card, somecol"
)
```

### Card — college completion (learners + kappa)

```{r ext-love-card-d2, fig.height=7.5, fig.width=14}
grid.arrange(
  grobs = love_from_candidates(
    "Card / college completion", learner_love_estimators, D2, X_dml_card
  ),
  nrow = 2, top = "Love Plots — eligible learners vs. kappa | Card, college completion"
)
```

### Kitagawa — somecol (learners + kappa)

```{r ext-love-kit-d1, fig.height=7.5, fig.width=14}
grid.arrange(
  grobs = love_from_candidates(
    "Kitagawa / some college", learner_love_estimators, D1, X_dml_kit
  ),
  nrow = 2, top = "Love Plots — eligible learners vs. kappa | Kitagawa, somecol"
)
```

### Kitagawa — college completion (learners + kappa)

```{r ext-love-kit-d2, fig.height=7.5, fig.width=14}
grid.arrange(
  grobs = love_from_candidates(
    "Kitagawa / college completion", learner_love_estimators, D2, X_dml_kit
  ),
  nrow = 2, top = "Love Plots — eligible learners vs. kappa | Kitagawa, college completion"
)
```

## Focused Thesis Figures

```{r focused-balance-figures, include=FALSE, cache=FALSE}
source("card_balance_figures.R")
create_card_balance_figures(
  candidate_cells = candidate_cells,
  D1 = D1,
  D2 = D2,
  X_dml_card = X_dml_card,
  X_dml_kit = X_dml_kit,
  output_dir = "../pics"
)
```
