---
title: "Card (1995) — Smoother Conditions for DML Outcome Weights"
subtitle: "Conditions 3, 5a, and 5b across conditioning sets and treatment margins"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size:15px; color:#2c2c2c; }
h1 { border-bottom:2px solid #2c7bb6; padding-bottom:6px; margin-top:36px; }
h2 { color:#2c7bb6; margin-top:28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size:13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  message=FALSE,
  warning=FALSE,
  cache=TRUE,
  autodep=TRUE,
  cache.extra="card-smoother-condition-audit-v1"
)
options(knitr.kable.NA="—")
```

# Purpose

This notebook checks the smoother conditions in Knaus (2024) for the Card
DML implementations whose outcome-weight representations are eligible for
the main diagnostics. It covers PLR-IV and Wald-AIPW with tuned honest
forests, linear or linear/logit nuisance models, and `ranger`. The checks are
repeated for the Card and Kitagawa conditioning sets and for both treatment
margins.

For an outcome smoother \(\widehat m=S Y\), Condition 3 requires

\[
S\mathbf 1_N=\mathbf 1_N.
\]

Full normalization additionally requires

\[
S D=\widehat D
\]

for PLR-IV (Condition 5a), or

\[
S_0^zD=\widehat D_0^z,\qquad
S_1^zD=\widehat D_1^z
\]

for Wald-AIPW (Condition 5b). The two Wald-AIPW branches are reported
separately.

<div class="note-box">
The estimator and learner definitions match `card_test.Rmd`: five-fold
cross-fitting, seed 42, fold-specific honest-forest tuning, linear nuisances
for PLR-IV, linear outcome plus logistic binary nuisances for parametric
Wald-AIPW, and `ranger` for the alternative forest implementation.
</div>

<div class="warn-box">
The smoother matrices contain \(N^2\) elements. Each model is therefore
fitted, audited, and released before the next model is processed. The four
design cells make this notebook memory- and computation-intensive. XGBoost
is not included because its reconstructed CARD weight vectors do not pass the
estimator-reproduction gate.
</div>

# Setup

```{r packages, cache=FALSE}
library(haven)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(dplyr)
library(knitr)
library(kableExtra)

lgr::get_logger("mlr3")$set_threshold("warn")
source(file.path("..","functions_all.R"))

if (!exists(
  "get_DoubleML_smoother",
  envir=asNamespace("OutcomeWeights"),
  inherits=FALSE
)) {
  stop("The installed OutcomeWeights version does not provide get_DoubleML_smoother().")
}
```

```{r data}
card<-read_dta("card.dta")
card<-card[!is.na(card$wage)&!is.na(card$educ),]

if (!"expersq"%in%names(card)) {
  card$expersq<-card$exper^2
}

Y<-as.numeric(log(card$wage/100))
Z<-as.integer(card$nearc4)
D_somecol<-as.integer(card$educ>12)
D_educ16<-as.integer(card$educ>=16)

X_card<-make_X(
  ~0+exper+expersq+black+south+smsa+smsa66+
    reg661+reg662+reg663+reg664+reg665+reg666+reg667+reg668,
  data=card
)
X_kitagawa<-make_X(
  ~0+black+south+smsa+smsa66+south66,
  data=card
)

design_cells<-list(
  "Card / some college"=list(X=X_card,D=D_somecol),
  "Card / college completion"=list(X=X_card,D=D_educ16),
  "Kitagawa / some college"=list(X=X_kitagawa,D=D_somecol),
  "Kitagawa / college completion"=list(X=X_kitagawa,D=D_educ16)
)

stopifnot(
  length(Y)==3010L,
  all(vapply(design_cells,function(x)nrow(x$X),integer(1))==length(Y)),
  all(vapply(design_cells,function(x)length(x$D),integer(1))==length(Y))
)
```

# Diagnostic functions

```{r helpers}
make_dml_data<-function(X,D){
  dat<-as.data.frame(X)
  dat$y<-Y
  dat$d<-D
  dat$z<-Z
  DoubleMLData$new(
    data=dat,
    y_col="y",
    d_cols="d",
    z_cols="z",
    x_cols=colnames(X)
  )
}

fit_pliv<-function(dml_data,learner){
  set.seed(42)
  object<-DoubleMLPLIV$new(
    data=dml_data,
    ml_l=learner(),
    ml_m=learner(),
    ml_r=learner(),
    n_folds=5
  )
  invisible(capture.output(
    object$fit(store_models=TRUE,store_predictions=TRUE)
  ))
  object
}

fit_wald<-function(dml_data,outcome_learner,
                     instrument_learner,treatment_learner){
  set.seed(42)
  object<-DoubleMLIIVM$new(
    data=dml_data,
    ml_g=outcome_learner(),
    ml_m=instrument_learner(),
    ml_r=treatment_learner(),
    n_folds=5,
    score="LATE"
  )
  invisible(capture.output(
    object$fit(store_models=TRUE,store_predictions=TRUE)
  ))
  object
}

fit_honest_forest<-function(X,D){
  set.seed(42)
  dml_with_smoother(
    Y,D,X,Z,
    n_cf_folds=5,
    tune.parameters="all"
  )
}

prediction_vector<-function(x){
  as.numeric(drop(x))
}

first_rep_smoother<-function(x){
  if (length(dim(x))==3L) return(as.matrix(x[1,,]))
  as.matrix(x)
}

extract_doubleml_smoothers<-function(object,dml_data){
  smoother<-getFromNamespace("get_DoubleML_smoother","OutcomeWeights")
  dat<-as.matrix(dml_data$data)
  test_ids<-object$smpls[[1]]$test_ids
  train_ids<-object$smpls[[1]]$train_ids
  x_cols<-dml_data$x_cols
  y_col<-dml_data$y_col

  if (inherits(object,"DoubleMLPLIV")) {
    return(list(
      S=smoother(
        object$models$ml_l$d[[1]],
        test_ids,train_ids,dat,x_cols,y_col,NULL
      )
    ))
  }

  if (inherits(object,"DoubleMLIIVM")) {
    z<-dat[,dml_data$z_cols]
    return(list(
      S.z0=smoother(
        object$models$ml_g0$d[[1]],
        test_ids,train_ids,dat,x_cols,y_col,z==0
      ),
      S.z1=smoother(
        object$models$ml_g1$d[[1]],
        test_ids,train_ids,dat,x_cols,y_col,z==1
      )
    ))
  }

  stop("Unsupported DoubleML class.")
}

extract_weight_vector<-function(object,dml_data=NULL,estimator){
  omega<-if (inherits(object,"dml_with_smoother")) {
    get_outcome_weights(object)$omega
  } else {
    get_outcome_weights(object=object,dml_data=dml_data)$omega
  }
  if (is.null(dim(omega))) return(as.numeric(omega))
  if (!is.null(rownames(omega))&&estimator%in%rownames(omega)) {
    return(as.numeric(omega[estimator,]))
  }
  as.numeric(omega[1,])
}

honest_estimate<-function(object,estimator){
  result<-summary(object)
  get_estimate(result,estimator)
}

condition3_row<-function(S,cell,estimator,learner,smoother,reproduces){
  row_sums<-rowSums(S)
  target<-rep(1,nrow(S))
  data.frame(
    Cell=cell,
    Estimator=estimator,
    Learner=learner,
    Smoother=smoother,
    `Minimum row sum`=min(row_sums),
    `Maximum row sum`=max(row_sums),
    `Maximum |row sum - 1|`=max(abs(row_sums-1)),
    `Condition 3`=isTRUE(all.equal(row_sums,target)),
    `Reproduces estimate`=reproduces,
    check.names=FALSE
  )
}

condition5_row<-function(S,D,D_hat,cell,estimator,learner,
                           condition,reproduces){
  reconstructed<-as.numeric(S%*%D)
  difference<-reconstructed-as.numeric(D_hat)
  data.frame(
    Cell=cell,
    Estimator=estimator,
    Learner=learner,
    Condition=condition,
    RMSE=sqrt(mean(difference^2)),
    `Maximum absolute difference`=max(abs(difference)),
    Satisfied=isTRUE(all.equal(reconstructed,as.numeric(D_hat))),
    `Reproduces estimate`=reproduces,
    check.names=FALSE
  )
}

classify_weights<-function(total,treated,control,reproduces){
  if (!isTRUE(reproduces)) {
    return("Not interpreted: reproduction failed")
  }
  total_zero<-isTRUE(all.equal(as.numeric(total),0))
  treated_one<-isTRUE(all.equal(as.numeric(treated),1))
  control_one<-isTRUE(all.equal(as.numeric(control),1))

  if (total_zero&&treated_one&&control_one) return("Fully normalized")
  if (total_zero) return("Scale-normalized")
  if (treated_one&&!control_one) return("Untreated-unnormalized")
  if (!treated_one&&control_one) return("Treated-unnormalized")
  "Fully unnormalized"
}

weight_row<-function(w,D,cell,estimator,learner,reproduces){
  total<-sum(w)
  treated<-sum(w[D==1])
  control<--sum(w[D==0])
  data.frame(
    Cell=cell,
    Estimator=estimator,
    Learner=learner,
    `Overall sum`=total,
    `Treated sum`=treated,
    `Control sum`=control,
    `Realized class`=classify_weights(
      total,treated,control,reproduces
    ),
    `Reproduces estimate`=reproduces,
    check.names=FALSE
  )
}
```

# Sequential audit

```{r audit-functions}
audit_honest_forest<-function(cell,X,D){
  object<-fit_honest_forest(X,D)
  smoothers<-list(
    S=first_rep_smoother(object$NuPa.hat$smoothers$S),
    S.z0=first_rep_smoother(object$NuPa.hat$smoothers$S.z0),
    S.z1=first_rep_smoother(object$NuPa.hat$smoothers$S.z1)
  )

  w_pliv<-extract_weight_vector(object,estimator="PLR-IV")
  w_wald<-extract_weight_vector(object,estimator="Wald-AIPW")
  est_pliv<-honest_estimate(object,"PLR-IV")
  est_wald<-honest_estimate(object,"Wald-AIPW")
  rep_pliv<-check_weight_identity(w_pliv,Y,est_pliv)
  rep_wald<-check_weight_identity(w_wald,Y,est_wald)

  condition3<-rbind(
    condition3_row(
      smoothers$S,cell,"PLR-IV","Tuned honest forest","S",rep_pliv
    ),
    condition3_row(
      smoothers$S.z0,cell,"Wald-AIPW","Tuned honest forest","S.z0",rep_wald
    ),
    condition3_row(
      smoothers$S.z1,cell,"Wald-AIPW","Tuned honest forest","S.z1",rep_wald
    )
  )

  condition5<-rbind(
    condition5_row(
      smoothers$S,D,
      prediction_vector(object$NuPa.hat$predictions$D.hat),
      cell,"PLR-IV","Tuned honest forest",
      "5a: S D = D.hat",rep_pliv
    ),
    condition5_row(
      smoothers$S.z0,D,
      prediction_vector(object$NuPa.hat$predictions$D.hat.z0),
      cell,"Wald-AIPW","Tuned honest forest",
      "5b: S.z0 D = D.hat.z0",rep_wald
    ),
    condition5_row(
      smoothers$S.z1,D,
      prediction_vector(object$NuPa.hat$predictions$D.hat.z1),
      cell,"Wald-AIPW","Tuned honest forest",
      "5b: S.z1 D = D.hat.z1",rep_wald
    )
  )

  weights<-rbind(
    weight_row(
      w_pliv,D,cell,"PLR-IV","Tuned honest forest",rep_pliv
    ),
    weight_row(
      w_wald,D,cell,"Wald-AIPW","Tuned honest forest",rep_wald
    )
  )

  list(condition3=condition3,condition5=condition5,weights=weights)
}

audit_alternative<-function(cell,X,D,estimator,learner){
  dml_data<-make_dml_data(X,D)

  if (estimator=="PLR-IV"&&learner=="Linear") {
    object<-fit_pliv(dml_data,function()lrn("regr.lm"))
  } else if (estimator=="PLR-IV"&&learner=="ranger") {
    object<-fit_pliv(
      dml_data,
      function()lrn("regr.ranger",keep.inbag=TRUE)
    )
  } else if (estimator=="Wald-AIPW"&&learner=="Linear/logit") {
    object<-fit_wald(
      dml_data,
      function()lrn("regr.lm"),
      function()lrn("classif.log_reg"),
      function()lrn("classif.log_reg")
    )
  } else if (estimator=="Wald-AIPW"&&learner=="ranger") {
    object<-fit_wald(
      dml_data,
      function()lrn("regr.ranger",keep.inbag=TRUE),
      function()lrn("classif.ranger",keep.inbag=TRUE),
      function()lrn("classif.ranger",keep.inbag=TRUE)
    )
  } else {
    stop("Unsupported estimator and learner combination.")
  }

  smoothers<-extract_doubleml_smoothers(object,dml_data)
  w<-extract_weight_vector(object,dml_data,estimator)
  estimate<-get_dml_coef(object)
  reproduces<-check_weight_identity(w,Y,estimate)

  if (estimator=="PLR-IV") {
    condition3<-condition3_row(
      smoothers$S,cell,estimator,learner,"S",reproduces
    )
    condition5<-condition5_row(
      smoothers$S,D,prediction_vector(object$predictions$ml_r),
      cell,estimator,learner,"5a: S D = D.hat",reproduces
    )
  } else {
    condition3<-rbind(
      condition3_row(
        smoothers$S.z0,cell,estimator,learner,"S.z0",reproduces
      ),
      condition3_row(
        smoothers$S.z1,cell,estimator,learner,"S.z1",reproduces
      )
    )
    condition5<-rbind(
      condition5_row(
        smoothers$S.z0,D,
        prediction_vector(object$predictions$ml_r0),
        cell,estimator,learner,
        "5b: S.z0 D = D.hat.z0",reproduces
      ),
      condition5_row(
        smoothers$S.z1,D,
        prediction_vector(object$predictions$ml_r1),
        cell,estimator,learner,
        "5b: S.z1 D = D.hat.z1",reproduces
      )
    )
  }

  weights<-weight_row(
    w,D,cell,estimator,learner,reproduces
  )

  list(condition3=condition3,condition5=condition5,weights=weights)
}
```

```{r run-audit, cache=TRUE}
alternative_grid<-rbind(
  data.frame(Estimator="PLR-IV",Learner="Linear"),
  data.frame(Estimator="PLR-IV",Learner="ranger"),
  data.frame(Estimator="Wald-AIPW",Learner="Linear/logit"),
  data.frame(Estimator="Wald-AIPW",Learner="ranger")
)

condition3_parts<-list()
condition5_parts<-list()
weight_parts<-list()
result_index<-1L

for (cell in names(design_cells)) {
  X<-design_cells[[cell]]$X
  D<-design_cells[[cell]]$D

  result<-audit_honest_forest(cell,X,D)
  condition3_parts[[result_index]]<-result$condition3
  condition5_parts[[result_index]]<-result$condition5
  weight_parts[[result_index]]<-result$weights
  rm(result)
  invisible(gc())
  result_index<-result_index+1L

  for (i in seq_len(nrow(alternative_grid))) {
    result<-audit_alternative(
      cell,X,D,
      alternative_grid$Estimator[i],
      alternative_grid$Learner[i]
    )
    condition3_parts[[result_index]]<-result$condition3
    condition5_parts[[result_index]]<-result$condition5
    weight_parts[[result_index]]<-result$weights
    rm(result)
    invisible(gc())
    result_index<-result_index+1L
  }
}

condition3_results<-do.call(rbind,condition3_parts)
condition5_results<-do.call(rbind,condition5_parts)
weight_results<-do.call(rbind,weight_parts)

saveRDS(
  list(
    condition3=condition3_results,
    condition5=condition5_results,
    weights=weight_results,
    settings=list(
      seed=42L,
      n_folds=5L,
      honest_forest_tuning="all",
      reproduction="all.equal default tolerance"
    )
  ),
  "card_smoother_condition_results.rds"
)
```

# Results

```{r table-helper, echo=FALSE}
cell_table<-function(results,cell,caption,digits,collapse_columns){
  results%>%
    filter(.data$Cell==cell)%>%
    select(-Cell)%>%
    kbl(
      caption=caption,
      digits=digits,
      row.names=FALSE
    )%>%
    kable_styling(
      bootstrap_options=c("striped","hover","condensed"),
      full_width=TRUE
    )%>%
    collapse_rows(columns=collapse_columns,valign="middle")
}
```

## Condition 3: affine outcome smoothers {.tabset}

### Card — some college

```{r condition3-card-somecol}
cell_table(
  condition3_results,"Card / some college",
  "Condition 3 — Card, some college",12,c(1,2)
)
```

### Card — college completion

```{r condition3-card-educ16}
cell_table(
  condition3_results,"Card / college completion",
  "Condition 3 — Card, college completion",12,c(1,2)
)
```

### Kitagawa — some college

```{r condition3-kitagawa-somecol}
cell_table(
  condition3_results,"Kitagawa / some college",
  "Condition 3 — Kitagawa, some college",12,c(1,2)
)
```

### Kitagawa — college completion

```{r condition3-kitagawa-educ16}
cell_table(
  condition3_results,"Kitagawa / college completion",
  "Condition 3 — Kitagawa, college completion",12,c(1,2)
)
```

## Conditions 5a and 5b: treatment-prediction alignment {.tabset}

For Wald-AIPW, Condition 5b is satisfied only if both instrument-specific
rows are marked `TRUE`.

### Card — some college

```{r condition5-card-somecol}
cell_table(
  condition5_results,"Card / some college",
  "Conditions 5a and 5b — Card, some college",10,c(1,2)
)
```

### Card — college completion

```{r condition5-card-educ16}
cell_table(
  condition5_results,"Card / college completion",
  "Conditions 5a and 5b — Card, college completion",10,c(1,2)
)
```

### Kitagawa — some college

```{r condition5-kitagawa-somecol}
cell_table(
  condition5_results,"Kitagawa / some college",
  "Conditions 5a and 5b — Kitagawa, some college",10,c(1,2)
)
```

### Kitagawa — college completion

```{r condition5-kitagawa-educ16}
cell_table(
  condition5_results,"Kitagawa / college completion",
  "Conditions 5a and 5b — Kitagawa, college completion",10,c(1,2)
)
```

## Realized outcome-weight classification {.tabset}

### Card — some college

```{r weights-card-somecol}
cell_table(
  weight_results,"Card / some college",
  "Realized normalization class — Card, some college",8,c(1)
)
```

### Card — college completion

```{r weights-card-educ16}
cell_table(
  weight_results,"Card / college completion",
  "Realized normalization class — Card, college completion",8,c(1)
)
```

### Kitagawa — some college

```{r weights-kitagawa-somecol}
cell_table(
  weight_results,"Kitagawa / some college",
  "Realized normalization class — Kitagawa, some college",8,c(1)
)
```

### Kitagawa — college completion

```{r weights-kitagawa-educ16}
cell_table(
  weight_results,"Kitagawa / college completion",
  "Realized normalization class — Kitagawa, college completion",8,c(1)
)
```

<div class="note-box">
The smoother conditions and realized weight masses answer different
questions. Condition 3 and Conditions 5a/5b explain why normalization arises
from the fitted nuisance implementation. The final table classifies only
vectors that reproduce their estimator. A passing smoother condition does not
override a failed outcome-weight reproduction check.
</div>

# Session information

```{r session-info, cache=FALSE}
sessionInfo()
```
