---
title: "Card (1995) — XGBoost Tuning Sensitivity"
subtitle: "Predictive losses, estimates, and numerical outcome-weight reproduction"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size: 12.5px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  cache = FALSE, fig.width = 13, fig.height = 6
)

library(haven)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(viridis)

source(file.path("..", "functions_all.R"))
```

<div class="note-box">
This notebook compares the untuned and fold-specifically tuned XGBoost
specifications stored by the two tuning notebooks. It reads the verified
exports and does not refit or retune any model.
</div>

```{r data-and-exports, include=FALSE}
card <- read_dta("card.dta")
card <- card[!is.na(card$wage) & !is.na(card$educ), ]
if (nrow(card) != 3010L) stop("Expected the Card analysis sample to contain 3,010 observations.")

card$somecol <- as.integer(card$educ >= 13)
card$educ16  <- as.integer(card$educ >= 16)
Y <- log(card$wage / 100)

wald_export <- readRDS("card_xgb_tuning_export.rds")
pliv_export <- readRDS("card_xgb_pliv_tuning_export.rds")

required <- c(
  "estimate_comparison", "fit_summaries", "nuisance_losses",
  "weight_diagnostics", "omega_vectors", "settings", "sample_info"
)
if (!all(required %in% names(wald_export)) ||
    !all(required %in% names(pliv_export))) {
  stop("At least one XGBoost export has an incomplete schema.")
}
if (!isTRUE(wald_export$settings$tune_on_folds) ||
    !isTRUE(pliv_export$settings$tune_on_folds)) {
  stop("The tuned exports must use tune_on_folds=TRUE.")
}

cells <- c(
  "Card / some college",
  "Card / college completion",
  "Kitagawa / some college",
  "Kitagawa / college completion"
)
if (!all(cells %in% names(wald_export$omega_vectors)) ||
    !all(cells %in% names(pliv_export$omega_vectors))) {
  stop("The exports do not contain all four Card design cells.")
}

extract_comparison <- function(export, estimator) {
  do.call(rbind, lapply(cells, function(cell) {
    do.call(rbind, lapply(c("untuned", "tuned"), function(variant) {
      fit <- export$fit_summaries[[cell]][[variant]]
      w <- as.numeric(export$omega_vectors[[cell]][[variant]])
      estimate <- as.numeric(fit$coef)
      gap <- abs(sum(w * Y) - estimate)
      data.frame(
        Estimator = estimator,
        Cell = cell,
        Variant = ifelse(variant == "untuned", "Untuned", "Tuned"),
        Estimate = estimate,
        SE = as.numeric(fit$se),
        Reproduction_gap = gap,
        Reproduces = check_weight_identity(w, Y, estimate),
        stringsAsFactors = FALSE
      )
    }))
  }))
}

comparison <- rbind(
  extract_comparison(pliv_export, "PLR-IV"),
  extract_comparison(wald_export, "Wald-AIPW")
)
comparison$Variant <- factor(comparison$Variant, levels = c("Untuned", "Tuned"))
comparison$Cell <- factor(comparison$Cell, levels = cells)

loss_reduction <- function(export, estimator, columns, labels) {
  do.call(rbind, lapply(cells, function(cell) {
    z <- export$nuisance_losses[export$nuisance_losses$Cell == cell, ]
    untuned <- z[grepl("untuned", z$Learner, fixed = TRUE), ]
    tuned <- z[
      grepl("tuned", z$Learner, fixed = TRUE) &
        !grepl("untuned", z$Learner, fixed = TRUE),
    ]
    if (nrow(untuned) != 1L || nrow(tuned) != 1L) {
      stop("Could not identify one tuned and one untuned loss row.")
    }
    data.frame(
      Estimator = estimator,
      Cell = cell,
      Nuisance = labels,
      Reduction_pct = 100 * (
        unlist(untuned[1, columns], use.names = FALSE) -
          unlist(tuned[1, columns], use.names = FALSE)
      ) / unlist(untuned[1, columns], use.names = FALSE),
      stringsAsFactors = FALSE
    )
  }))
}

loss_plot_data <- rbind(
  loss_reduction(
    pliv_export, "PLR-IV",
    c("Outcome l RMSE", "Instrument m RMSE", "Treatment r RMSE"),
    c("Outcome", "Instrument", "Treatment")
  ),
  loss_reduction(
    wald_export, "Wald-AIPW",
    c("Outcome RMSE", "Instrument log-loss", "Treatment log-loss"),
    c("Outcome", "Instrument", "Treatment")
  )
)
loss_plot_data$Cell <- factor(loss_plot_data$Cell, levels = cells)
```

# 1. Nuisance-prediction losses

```{r loss-reduction-plot, fig.height=6.5}
ggplot(
  loss_plot_data,
  aes(x = Reduction_pct, y = Nuisance, color = Estimator)
) +
  geom_vline(xintercept = 0, color = "grey65", linewidth = 0.4) +
  geom_point(size = 2.8) +
  facet_wrap(~ Cell, ncol = 2) +
  scale_color_viridis_d(option = "D", end = 0.82) +
  labs(
    x = "Reduction in recorded validation loss after tuning (%)",
    y = NULL,
    color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

Positive values mean that the tuned specification has a lower recorded
validation loss. The loss functions differ across the two estimators and are
therefore compared only within estimator and nuisance.

# 2. Point estimates

```{r estimate-plot, fig.height=7.5}
ggplot(comparison, aes(x = Variant, y = Estimate, group = 1)) +
  geom_line(color = "#777777", linewidth = 0.55) +
  geom_point(aes(color = Variant), size = 2.8) +
  geom_text(
    aes(label = sprintf("%.3f", Estimate)),
    vjust = -0.75, size = 3.1, show.legend = FALSE
  ) +
  facet_wrap(~ Estimator + Cell, scales = "free_y", ncol = 4) +
  scale_color_viridis_d(option = "D", end = 0.82) +
  scale_y_continuous(expand = expansion(mult = c(0.10, 0.18))) +
  labs(x = NULL, y = "Estimate", color = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 25, hjust = 1),
    strip.text = element_text(size = 9)
  )
```

```{r estimate-table}
comparison %>%
  transmute(
    Estimator, Cell, Variant,
    Estimate = round(Estimate, 4),
    SE = round(SE, 4)
  ) %>%
  kbl(
    caption = "XGBoost estimates: untuned and fold-specifically tuned",
    align = c("l", "l", "l", "r", "r")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE
  ) %>%
  collapse_rows(columns = c(1, 2), valign = "middle")
```

The panels use separate vertical scales because the untuned Wald-AIPW
estimates and standard errors differ substantially across cells. The table
reports the standard errors; the plot is limited to the point-estimate
comparison.

# 3. Numerical outcome-weight reproduction

```{r reproduction-plot, fig.height=5.5}
ggplot(
  comparison,
  aes(x = Reproduction_gap, y = Cell, color = Variant)
) +
  geom_point(size = 2.8, position = position_dodge(width = 0.45)) +
  facet_wrap(~ Estimator, ncol = 1, scales = "free_y") +
  scale_x_log10() +
  scale_color_viridis_d(option = "D", end = 0.82) +
  labs(
    x = "Absolute reproduction gap |sum(omega_i Y_i) - tau_hat|",
    y = NULL,
    color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

This is a numerical diagnostic for the reconstructed outcome weights. It is
not a tuning criterion.

```{r reproduction-table}
comparison %>%
  transmute(
    Estimator, Cell, Variant,
    `|ω′Y−τ̂|` = Reproduction_gap,
    `ω′Y reproduces τ̂` = ifelse(Reproduces, "TRUE", "FALSE")
  ) %>%
  kbl(
    digits = 10,
    format.args = list(scientific = TRUE),
    caption = "XGBoost outcome-weight representation audit",
    align = c("l", "l", "l", "r", "c")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE
  ) %>%
  collapse_rows(columns = c(1, 2), valign = "middle") %>%
  footnote(
    general = paste(
      "Each fitted estimator is assessed separately using Knaus's all.equal()",
      "criterion with R's default tolerance. A FALSE vector is excluded from",
      "normalization, concentration, and balance diagnostics."
    )
  )
```

# 4. Reading the comparison

The recorded nuisance-prediction losses are lower after tuning in every
reported cell and nuisance. The reductions are larger under the Card
specification than under the Kitagawa specification.

For PLR-IV, tuning lowers the point estimate in all four cells. For Wald-AIPW,
the untuned estimates vary much more across cells, while the tuned estimates
lie between 0.238 and 0.730. These comparisons describe the effect of the
stated tuning protocol. They do not establish that the tuned causal estimates
are closer to the underlying treatment effects.

None of the reconstructed XGBoost vectors passes the estimator-reproduction
gate. The notebook therefore does not report normalization, concentration, or
balance diagnostics for those vectors. This numerical failure does not remove
the XGBoost point estimates from the learner-sensitivity or translation
analysis.
