---
title: "Angrist & Evans (1998) — Smoother Conditions for DML Outcome Weights"
subtitle: "Conditions 3, 5a, and 5b for work status and maternal log income"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size:15px; color:#2c2c2c; }
h1 { border-bottom:2px solid #2c7bb6; padding-bottom:6px; margin-top:36px; }
h2 { color:#2c7bb6; margin-top:28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size:13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE,
  message=FALSE,
  warning=FALSE,
  cache=TRUE,
  autodep=TRUE,
  cache.extra="child-smoother-condition-audit-v1"
)
options(knitr.kable.NA="—")
```

# Purpose

This notebook checks the smoother conditions in Knaus (2024) for the Child
DML implementations whose outcome-weight representations are eligible for
the main diagnostics. The audit is performed separately for work status and
maternal log income because the two outcomes use distinct fixed samples.

The learner definitions match `child_test.Rmd`: PLR-IV and Wald-AIPW are fitted
with tuned honest forests, linear or linear/logit nuisance models, and
`ranger`, using the same six covariates, five-fold cross-fitting, and seed 42.

For an outcome smoother \(\widehat m=S Y\), Condition 3 requires

\[
S\mathbf 1_N=\mathbf 1_N.
\]

Full normalization additionally requires

\[
S D=\widehat D
\]

for PLR-IV (Condition 5a), or

\[
S_0^zD=\widehat D_0^z,\qquad
S_1^zD=\widehat D_1^z
\]

for Wald-AIPW (Condition 5b). The two Wald-AIPW branches are reported
separately.

<div class="warn-box">
The smoother matrices contain \(N^2\) elements. Each model is fitted, audited,
and released before the next model is processed. XGBoost is not included
because its reconstructed Child outcome-weight vectors do not pass the
estimator-reproduction gate used for the main weight diagnostics.
</div>

# Setup

```{r packages, cache=FALSE}
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(dplyr)
library(knitr)
library(kableExtra)

lgr::get_logger("mlr3")$set_threshold("warn")
source(file.path("..","functions_all.R"))

if (!exists(
  "get_DoubleML_smoother",
  envir=asNamespace("OutcomeWeights"),
  inherits=FALSE
)) {
  stop("The installed OutcomeWeights version does not provide get_DoubleML_smoother().")
}
```

```{r data}
obj<-readRDS("ae98_sub.rds")
if (!all(c("ae98_labor_sub","ae98_income_sub")%in%names(obj))) {
  stop("ae98_sub.rds must contain ae98_labor_sub and ae98_income_sub.")
}

labor<-obj$ae98_labor_sub%>%
  mutate(
    workedm=as.integer(workedm),
    morekids=as.integer(morekids),
    samesex=as.integer(samesex)
  )
income<-obj$ae98_income_sub%>%
  mutate(
    morekids=as.integer(morekids),
    samesex=as.integer(samesex),
    lincomem=log(incomem)
  )

CHILD_COVARIATES<-c(
  "agem","agefstm","sexk","blackm","hispm","othracem"
)
child_formula<-reformulate(CHILD_COVARIATES,intercept=FALSE)

analysis_cells<-list(
  "Work status"=list(
    Y=as.numeric(labor$workedm),
    D=as.integer(labor$morekids),
    Z=as.integer(labor$samesex),
    X=make_X(child_formula,data=labor)
  ),
  "Maternal log income"=list(
    Y=as.numeric(income$lincomem),
    D=as.integer(income$morekids),
    Z=as.integer(income$samesex),
    X=make_X(child_formula,data=income)
  )
)

stopifnot(
  identical(vapply(analysis_cells,function(x)length(x$Y),integer(1)),
            c(`Work status`=3000L,`Maternal log income`=3000L)),
  all(vapply(analysis_cells,function(x)nrow(x$X),integer(1))==3000L),
  all(vapply(analysis_cells,function(x)ncol(x$X),integer(1))==6L),
  identical(colnames(analysis_cells[[1]]$X),CHILD_COVARIATES),
  identical(colnames(analysis_cells[[2]]$X),CHILD_COVARIATES)
)
```

# Diagnostic functions

```{r helpers}
make_dml_data<-function(cell){
  dat<-as.data.frame(cell$X)
  dat$y<-cell$Y
  dat$d<-cell$D
  dat$z<-cell$Z
  DoubleMLData$new(
    data=dat,
    y_col="y",
    d_cols="d",
    z_cols="z",
    x_cols=colnames(cell$X)
  )
}

fit_pliv<-function(dml_data,learner){
  set.seed(42)
  object<-DoubleMLPLIV$new(
    data=dml_data,
    ml_l=learner(),
    ml_m=learner(),
    ml_r=learner(),
    n_folds=5
  )
  invisible(capture.output(
    object$fit(store_models=TRUE,store_predictions=TRUE)
  ))
  object
}

fit_wald<-function(dml_data,outcome_learner,
                     instrument_learner,treatment_learner){
  set.seed(42)
  object<-DoubleMLIIVM$new(
    data=dml_data,
    ml_g=outcome_learner(),
    ml_m=instrument_learner(),
    ml_r=treatment_learner(),
    n_folds=5,
    score="LATE"
  )
  invisible(capture.output(
    object$fit(store_models=TRUE,store_predictions=TRUE)
  ))
  object
}

fit_honest_forest<-function(cell){
  set.seed(42)
  dml_with_smoother(
    cell$Y,cell$D,cell$X,cell$Z,
    n_cf_folds=5,
    tune.parameters="all"
  )
}

prediction_vector<-function(x){
  as.numeric(drop(x))
}

first_rep_smoother<-function(x){
  if (length(dim(x))==3L) return(as.matrix(x[1,,]))
  as.matrix(x)
}

extract_doubleml_smoothers<-function(object,dml_data){
  smoother<-getFromNamespace("get_DoubleML_smoother","OutcomeWeights")
  dat<-as.matrix(dml_data$data)
  test_ids<-object$smpls[[1]]$test_ids
  train_ids<-object$smpls[[1]]$train_ids
  x_cols<-dml_data$x_cols
  y_col<-dml_data$y_col

  if (inherits(object,"DoubleMLPLIV")) {
    return(list(
      S=smoother(
        object$models$ml_l$d[[1]],
        test_ids,train_ids,dat,x_cols,y_col,NULL
      )
    ))
  }

  if (inherits(object,"DoubleMLIIVM")) {
    z<-dat[,dml_data$z_cols]
    return(list(
      S.z0=smoother(
        object$models$ml_g0$d[[1]],
        test_ids,train_ids,dat,x_cols,y_col,z==0
      ),
      S.z1=smoother(
        object$models$ml_g1$d[[1]],
        test_ids,train_ids,dat,x_cols,y_col,z==1
      )
    ))
  }

  stop("Unsupported DoubleML class.")
}

extract_weight_vector<-function(object,dml_data=NULL,estimator){
  omega<-if (inherits(object,"dml_with_smoother")) {
    get_outcome_weights(object)$omega
  } else {
    get_outcome_weights(object=object,dml_data=dml_data)$omega
  }
  if (is.null(dim(omega))) return(as.numeric(omega))
  if (!is.null(rownames(omega))&&estimator%in%rownames(omega)) {
    return(as.numeric(omega[estimator,]))
  }
  as.numeric(omega[1,])
}

honest_estimate<-function(object,estimator){
  get_estimate(summary(object),estimator)
}

condition3_row<-function(S,outcome,estimator,learner,smoother,reproduces){
  row_sums<-rowSums(S)
  target<-rep(1,nrow(S))
  data.frame(
    Outcome=outcome,
    Estimator=estimator,
    Learner=learner,
    Smoother=smoother,
    `Minimum row sum`=min(row_sums),
    `Maximum row sum`=max(row_sums),
    `Maximum |row sum - 1|`=max(abs(row_sums-1)),
    `Condition 3`=isTRUE(all.equal(row_sums,target)),
    `Reproduces estimate`=reproduces,
    check.names=FALSE
  )
}

condition5_row<-function(S,D,D_hat,outcome,estimator,learner,
                           condition,reproduces){
  reconstructed<-as.numeric(S%*%D)
  difference<-reconstructed-as.numeric(D_hat)
  data.frame(
    Outcome=outcome,
    Estimator=estimator,
    Learner=learner,
    Condition=condition,
    RMSE=sqrt(mean(difference^2)),
    `Maximum absolute difference`=max(abs(difference)),
    Satisfied=isTRUE(all.equal(reconstructed,as.numeric(D_hat))),
    `Reproduces estimate`=reproduces,
    check.names=FALSE
  )
}

classify_weights<-function(total,treated,control,reproduces){
  if (!isTRUE(reproduces)) return("Not interpreted: reproduction failed")
  total_zero<-isTRUE(all.equal(as.numeric(total),0))
  treated_one<-isTRUE(all.equal(as.numeric(treated),1))
  control_one<-isTRUE(all.equal(as.numeric(control),1))

  if (total_zero&&treated_one&&control_one) return("Fully normalized")
  if (total_zero) return("Scale-normalized")
  if (treated_one&&!control_one) return("Untreated-unnormalized")
  if (!treated_one&&control_one) return("Treated-unnormalized")
  "Fully unnormalized"
}

weight_row<-function(w,D,outcome,estimator,learner,reproduces){
  total<-sum(w)
  treated<-sum(w[D==1])
  control<--sum(w[D==0])
  data.frame(
    Outcome=outcome,
    Estimator=estimator,
    Learner=learner,
    `Overall sum`=total,
    `Treated sum`=treated,
    `Control sum`=control,
    `Realized class`=classify_weights(total,treated,control,reproduces),
    `Reproduces estimate`=reproduces,
    check.names=FALSE
  )
}
```

# Sequential audit

```{r audit-functions}
audit_honest_forest<-function(outcome,cell){
  object<-fit_honest_forest(cell)
  smoothers<-list(
    S=first_rep_smoother(object$NuPa.hat$smoothers$S),
    S.z0=first_rep_smoother(object$NuPa.hat$smoothers$S.z0),
    S.z1=first_rep_smoother(object$NuPa.hat$smoothers$S.z1)
  )

  w_pliv<-extract_weight_vector(object,estimator="PLR-IV")
  w_wald<-extract_weight_vector(object,estimator="Wald-AIPW")
  rep_pliv<-check_weight_identity(
    w_pliv,cell$Y,honest_estimate(object,"PLR-IV")
  )
  rep_wald<-check_weight_identity(
    w_wald,cell$Y,honest_estimate(object,"Wald-AIPW")
  )

  condition3<-rbind(
    condition3_row(
      smoothers$S,outcome,"PLR-IV","Tuned honest forest","S",rep_pliv
    ),
    condition3_row(
      smoothers$S.z0,outcome,"Wald-AIPW","Tuned honest forest","S.z0",rep_wald
    ),
    condition3_row(
      smoothers$S.z1,outcome,"Wald-AIPW","Tuned honest forest","S.z1",rep_wald
    )
  )

  condition5<-rbind(
    condition5_row(
      smoothers$S,cell$D,
      prediction_vector(object$NuPa.hat$predictions$D.hat),
      outcome,"PLR-IV","Tuned honest forest","5a: S D = D.hat",rep_pliv
    ),
    condition5_row(
      smoothers$S.z0,cell$D,
      prediction_vector(object$NuPa.hat$predictions$D.hat.z0),
      outcome,"Wald-AIPW","Tuned honest forest",
      "5b: S.z0 D = D.hat.z0",rep_wald
    ),
    condition5_row(
      smoothers$S.z1,cell$D,
      prediction_vector(object$NuPa.hat$predictions$D.hat.z1),
      outcome,"Wald-AIPW","Tuned honest forest",
      "5b: S.z1 D = D.hat.z1",rep_wald
    )
  )

  weights<-rbind(
    weight_row(w_pliv,cell$D,outcome,"PLR-IV","Tuned honest forest",rep_pliv),
    weight_row(w_wald,cell$D,outcome,"Wald-AIPW","Tuned honest forest",rep_wald)
  )

  list(condition3=condition3,condition5=condition5,weights=weights)
}

audit_alternative<-function(outcome,cell,estimator,learner){
  dml_data<-make_dml_data(cell)

  if (estimator=="PLR-IV"&&learner=="Linear") {
    object<-fit_pliv(dml_data,function()lrn("regr.lm"))
  } else if (estimator=="PLR-IV"&&learner=="ranger") {
    object<-fit_pliv(
      dml_data,function()lrn("regr.ranger",keep.inbag=TRUE)
    )
  } else if (estimator=="Wald-AIPW"&&learner=="Linear/logit") {
    object<-fit_wald(
      dml_data,
      function()lrn("regr.lm"),
      function()lrn("classif.log_reg"),
      function()lrn("classif.log_reg")
    )
  } else if (estimator=="Wald-AIPW"&&learner=="ranger") {
    object<-fit_wald(
      dml_data,
      function()lrn("regr.ranger",keep.inbag=TRUE),
      function()lrn("classif.ranger",keep.inbag=TRUE),
      function()lrn("classif.ranger",keep.inbag=TRUE)
    )
  } else {
    stop("Unsupported estimator and learner combination.")
  }

  smoothers<-extract_doubleml_smoothers(object,dml_data)
  w<-extract_weight_vector(object,dml_data,estimator)
  reproduces<-check_weight_identity(w,cell$Y,get_dml_coef(object))

  if (estimator=="PLR-IV") {
    condition3<-condition3_row(
      smoothers$S,outcome,estimator,learner,"S",reproduces
    )
    condition5<-condition5_row(
      smoothers$S,cell$D,prediction_vector(object$predictions$ml_r),
      outcome,estimator,learner,"5a: S D = D.hat",reproduces
    )
  } else {
    condition3<-rbind(
      condition3_row(
        smoothers$S.z0,outcome,estimator,learner,"S.z0",reproduces
      ),
      condition3_row(
        smoothers$S.z1,outcome,estimator,learner,"S.z1",reproduces
      )
    )
    condition5<-rbind(
      condition5_row(
        smoothers$S.z0,cell$D,
        prediction_vector(object$predictions$ml_r0),
        outcome,estimator,learner,"5b: S.z0 D = D.hat.z0",reproduces
      ),
      condition5_row(
        smoothers$S.z1,cell$D,
        prediction_vector(object$predictions$ml_r1),
        outcome,estimator,learner,"5b: S.z1 D = D.hat.z1",reproduces
      )
    )
  }

  weights<-weight_row(w,cell$D,outcome,estimator,learner,reproduces)
  list(condition3=condition3,condition5=condition5,weights=weights)
}
```

```{r run-audit, cache=TRUE}
alternative_grid<-rbind(
  data.frame(Estimator="PLR-IV",Learner="Linear"),
  data.frame(Estimator="PLR-IV",Learner="ranger"),
  data.frame(Estimator="Wald-AIPW",Learner="Linear/logit"),
  data.frame(Estimator="Wald-AIPW",Learner="ranger")
)

condition3_parts<-list()
condition5_parts<-list()
weight_parts<-list()
result_index<-1L

for (outcome in names(analysis_cells)) {
  cell<-analysis_cells[[outcome]]

  result<-audit_honest_forest(outcome,cell)
  condition3_parts[[result_index]]<-result$condition3
  condition5_parts[[result_index]]<-result$condition5
  weight_parts[[result_index]]<-result$weights
  rm(result)
  invisible(gc())
  result_index<-result_index+1L

  for (i in seq_len(nrow(alternative_grid))) {
    result<-audit_alternative(
      outcome,cell,
      alternative_grid$Estimator[i],
      alternative_grid$Learner[i]
    )
    condition3_parts[[result_index]]<-result$condition3
    condition5_parts[[result_index]]<-result$condition5
    weight_parts[[result_index]]<-result$weights
    rm(result)
    invisible(gc())
    result_index<-result_index+1L
  }
}

condition3_results<-do.call(rbind,condition3_parts)
condition5_results<-do.call(rbind,condition5_parts)
weight_results<-do.call(rbind,weight_parts)

saveRDS(
  list(
    condition3=condition3_results,
    condition5=condition5_results,
    weights=weight_results,
    settings=list(
      seed=42L,
      n_folds=5L,
      honest_forest_tuning="all",
      covariates=CHILD_COVARIATES,
      samples=c(`Work status`=nrow(labor),
                `Maternal log income`=nrow(income)),
      reproduction="all.equal default tolerance"
    )
  ),
  "ae98_smoother_condition_results.rds"
)
```

# Results

```{r table-helper, echo=FALSE}
outcome_table<-function(results,outcome,caption,digits,collapse_columns){
  results%>%
    filter(.data$Outcome==outcome)%>%
    select(-Outcome)%>%
    kbl(caption=caption,digits=digits,row.names=FALSE)%>%
    kable_styling(
      bootstrap_options=c("striped","hover","condensed"),
      full_width=TRUE
    )%>%
    collapse_rows(columns=collapse_columns,valign="middle")
}
```

## Condition 3: affine outcome smoothers {.tabset}

### Work status

```{r condition3-labor}
outcome_table(
  condition3_results,"Work status",
  "Condition 3 — work status",12,c(1,2)
)
```

### Maternal log income

```{r condition3-income}
outcome_table(
  condition3_results,"Maternal log income",
  "Condition 3 — maternal log income",12,c(1,2)
)
```

## Conditions 5a and 5b: treatment-prediction alignment {.tabset}

For Wald-AIPW, Condition 5b is satisfied only if both instrument-specific
rows are marked `TRUE`.

### Work status

```{r condition5-labor}
outcome_table(
  condition5_results,"Work status",
  "Conditions 5a and 5b — work status",10,c(1,2)
)
```

### Maternal log income

```{r condition5-income}
outcome_table(
  condition5_results,"Maternal log income",
  "Conditions 5a and 5b — maternal log income",10,c(1,2)
)
```

## Realized outcome-weight classification {.tabset}

### Work status

```{r weights-labor}
outcome_table(
  weight_results,"Work status",
  "Realized normalization class — work status",8,c(1)
)
```

### Maternal log income

```{r weights-income}
outcome_table(
  weight_results,"Maternal log income",
  "Realized normalization class — maternal log income",8,c(1)
)
```

<div class="note-box">
The smoother conditions and realized weight masses answer different
questions. Condition 3 and Conditions 5a/5b explain why normalization should
arise from the fitted nuisance implementation. The final table classifies only
vectors that reproduce their estimator. A passing smoother condition does not
override a failed outcome-weight reproduction check.
</div>

# Session information

```{r session-info, cache=FALSE}
sessionInfo()
```
