---
title: "Fertility and Mothers' Labor-Market Outcomes: XGBoost Tuning for Wald-AIPW"
subtitle: "Nested nuisance tuning within the cross-fitting design"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      cache = FALSE,
                      fig.width = 11, fig.height = 5.5)
```

```{r packages, include=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

source(file.path("..", "functions_all.R"))

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

SEED        <- 42
N_CF        <- 5
TUNE_EVALS  <- 15
INNER_FOLDS <- 3
RERUN_TUNING <- FALSE
EXPORT_PATH  <- "ae98_xgb_tuning_export.rds"
```

<div class="note-box">
**Purpose.** This notebook compares untuned and fold-specifically tuned
XGBoost nuisances for Wald-AIPW in the two child-application samples.
Hyperparameters are selected by nuisance-prediction loss within each outer
training fold. The learner uses the common six-covariate Child design:
`agem`, `agefstm`, `sexk`, `blackm`, `hispm`, and `othracem`.
`sex2nd` is omitted because, together with `sexk`, it exactly reconstructs
the `samesex` instrument. Giving both indicators to an unrestricted tree
learner would therefore remove the residual instrument variation required
for DML. This is the same design used by every estimator in
`child_test.Rmd`.

Ordinary renders read the verified export. Set `RERUN_TUNING = TRUE` only
when deliberately re-estimating the nested tuning procedure.
</div>

<div class="warn-box">
**Runtime.** A substantive rerun tunes the instrument, treatment, and
arm-specific outcome nuisances separately within every outer training fold
and replaces `ae98_xgb_tuning_export.rds`.
</div>

---

# Setup

```{r data, include=FALSE}
obj_data <- readRDS("ae98_sub.rds")
if (!all(c("ae98_labor_sub", "ae98_income_sub") %in% names(obj_data))) {
  stop("ae98_sub.rds must contain ae98_labor_sub and ae98_income_sub.")
}

ae98_labor_sub  <- obj_data$ae98_labor_sub
ae98_income_sub <- obj_data$ae98_income_sub

ae98_labor_sub <- ae98_labor_sub %>%
  mutate(workedm = as.integer(workedm),
         morekids = as.integer(morekids),
         samesex  = as.integer(samesex))

ae98_income_sub <- ae98_income_sub %>%
  mutate(morekids  = as.integer(morekids),
         samesex   = as.integer(samesex),
         lincomem  = log(incomem))

CHILD_COVARIATES <- c(
  "agem", "agefstm", "sexk", "blackm", "hispm", "othracem"
)
RHS_FORMULA <- reformulate(CHILD_COVARIATES, intercept = FALSE)
X_dml_labor  <- make_X(RHS_FORMULA, data = ae98_labor_sub)
X_dml_income <- make_X(RHS_FORMULA, data = ae98_income_sub)

D_labor  <- ae98_labor_sub$morekids;  Z_labor  <- ae98_labor_sub$samesex
D_income <- ae98_income_sub$morekids; Z_income <- ae98_income_sub$samesex

```

```{r analysis-dimensions, echo=FALSE}
data.frame(
  Outcome = c("Work Status", "Maternal Log Income"),
  N = c(nrow(ae98_labor_sub), nrow(ae98_income_sub)),
  Covariates = c(ncol(X_dml_labor), ncol(X_dml_income))
) %>%
  kbl(caption = "Analysis dimensions") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE)
```

```{r learners-and-tuning-setup, include=FALSE}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                    x_cols = colnames(X))
}

make_xgb_learners <- function() {
  list(
    g = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
    m = lrn("classif.xgboost", predict_type = "prob"),
    r = lrn("classif.xgboost", predict_type = "prob")
  )
}

xgb_ps <- function() {
  ps(nrounds = p_int(50, 400), max_depth = p_int(2, 5),
     eta = p_dbl(0.02, 0.20), min_child_weight = p_int(1, 8))
}

tune_param_set <- function() list(ml_g = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())

tune_settings <- function() {
  list(
    n_folds_tune = INNER_FOLDS,
    rsmp_tune    = rsmp("cv", folds = INNER_FOLDS),
    measure      = list(ml_g = "regr.rmse", ml_m = "classif.logloss", ml_r = "classif.logloss"),
    terminator   = trm("evals", n_evals = TUNE_EVALS),
    algorithm    = tnr("random_search")
  )
}

get_prediction_vector <- function(obj, learner_name) {
  pred <- tryCatch(obj$predictions[[learner_name]], error = function(e) NULL)
  if (is.null(pred)) return(numeric(0))
  as.numeric(drop(pred))
}

get_branch_prediction <- function(obj, base_name, Z) {
  pred0 <- get_prediction_vector(obj, paste0(base_name, "0"))
  pred1 <- get_prediction_vector(obj, paste0(base_name, "1"))
  if (length(pred0) != length(Z) || length(pred1) != length(Z)) return(numeric(0))
  ifelse(as.numeric(Z) == 1, pred1, pred0)
}

propensity_boundary_shares <- function(obj, Z, threshold = 0.02) {
  m_hat <- get_prediction_vector(obj, "ml_m")
  r_hat <- get_branch_prediction(obj, "ml_r", Z)
  boundary_share <- function(p_hat) {
    p_hat <- p_hat[is.finite(p_hat)]
    if (length(p_hat) == 0) return(NA_real_)
    mean(p_hat < threshold | p_hat > 1 - threshold)
  }
  c(
    instrument = boundary_share(m_hat),
    treatment = boundary_share(r_hat)
  )
}

rmse_loss <- function(truth, pred) {
  if (length(pred) != length(truth) || length(pred) == 0) return(NA_real_)
  truth <- as.numeric(truth); pred <- as.numeric(pred)
  keep <- is.finite(truth) & is.finite(pred)
  if (!any(keep)) return(NA_real_)
  sqrt(mean((truth[keep] - pred[keep])^2))
}

logloss_loss <- function(truth, pred, eps = 1e-6) {
  if (length(pred) != length(truth) || length(pred) == 0) return(NA_real_)
  truth <- as.numeric(truth); pred <- as.numeric(pred)
  keep <- is.finite(truth) & is.finite(pred)
  if (!any(keep)) return(NA_real_)
  truth <- truth[keep]; pred <- pmin(pmax(pred[keep], eps), 1 - eps)
  -mean(truth * log(pred) + (1 - truth) * log(1 - pred))
}

calc_nuisance_losses <- function(r, cell) {
  if (is.null(r)) return(c(outcome_rmse = NA_real_, instrument_logloss = NA_real_, treatment_logloss = NA_real_))
  g_hat <- get_branch_prediction(r$obj, "ml_g", cell$Z)
  m_hat <- get_prediction_vector(r$obj, "ml_m")
  r_hat <- get_branch_prediction(r$obj, "ml_r", cell$Z)
  c(outcome_rmse        = rmse_loss(cell$Y, g_hat),
    instrument_logloss   = logloss_loss(cell$Z, m_hat),
    treatment_logloss    = logloss_loss(cell$D, r_hat))
}

fit_cell <- function(dml_data, learners, label, Z, tune = FALSE, tune_on_folds = FALSE) {
  tryCatch({
    set.seed(SEED)
    obj <- DoubleMLIIVM$new(data = dml_data, ml_g = learners$g, ml_m = learners$m,
                             ml_r = learners$r, n_folds = N_CF, score = "LATE")
    if (tune) {
      obj$tune(param_set = tune_param_set(), tune_settings = tune_settings(),
               tune_on_folds = tune_on_folds)
    }
    obj$fit(store_models = TRUE, store_predictions = TRUE)

    boundary <- propensity_boundary_shares(obj, Z)
    list(label = label, tuned = tune,
         coef = get_dml_coef(obj), se = get_dml_se(obj),
         instrument_boundary_share = unname(boundary["instrument"]),
         treatment_boundary_share = unname(boundary["treatment"]),
         tuning_result = tryCatch(obj$tuning_result, error = function(e) NULL),
         params        = tryCatch(obj$params,        error = function(e) NULL),
         obj = obj)
  }, error = function(e) {
    message(sprintf("Fit failed for %s: %s", label, conditionMessage(e)))
    NULL
  })
}

cells <- list(
  "Labor (workedm)"   = list(Y = ae98_labor_sub$workedm,  D = D_labor,  Z = Z_labor,  X = X_dml_labor),
  "Income (lincomem)" = list(Y = ae98_income_sub$lincomem, D = D_income, Z = Z_income, X = X_dml_income)
)
```

---

# Fit: untuned versus tuned XGBoost

```{r fit-all, include=FALSE}
if (RERUN_TUNING) {
  results <- list()
  for (cell_name in names(cells)) {
    cell <- cells[[cell_name]]
    dml_data <- make_dml_data(cell$Y, cell$D, cell$Z, cell$X)

    results[[cell_name]] <- list(
      untuned = fit_cell(dml_data, make_xgb_learners(),
                         paste(cell_name, "xgb-untuned"),
                         Z = cell$Z, tune = FALSE),
      tuned   = fit_cell(dml_data, make_xgb_learners(),
                         paste(cell_name, "xgb-tuned"),
                         Z = cell$Z, tune = TRUE, tune_on_folds = TRUE)
    )
  }
} else {
  if (!file.exists(EXPORT_PATH)) {
    stop("Verified Wald-AIPW XGBoost export not found. Set RERUN_TUNING=TRUE to rebuild it.")
  }
  stored_export <- readRDS(EXPORT_PATH)
}
```

## 1. Estimate comparison

```{r comparison-table}
fmt_est <- function(r) if (is.null(r)) "\u2014" else sprintf("%.3f (%.3f)", r$coef, r$se)

# Cached fits created before the boundary diagnostics were split do not contain
# the two fields below. Recover them from the stored out-of-fold predictions so
# that updating the reporting code does not require refitting XGBoost.
add_boundary_shares <- function(r, Z) {
  if (is.null(r)) return(NULL)

  has_scalar <- function(x) {
    length(x) == 1L && is.numeric(x) && (is.finite(x) || is.na(x))
  }
  has_instrument <- has_scalar(r$instrument_boundary_share)
  has_treatment  <- has_scalar(r$treatment_boundary_share)

  if ((!has_instrument || !has_treatment) && !is.null(r$obj)) {
    boundary <- propensity_boundary_shares(r$obj, Z)
    if (!has_instrument) {
      r$instrument_boundary_share <- unname(boundary["instrument"])
    }
    if (!has_treatment) {
      r$treatment_boundary_share <- unname(boundary["treatment"])
    }
  }

  if (!has_scalar(r$instrument_boundary_share)) {
    r$instrument_boundary_share <- NA_real_
  }
  if (!has_scalar(r$treatment_boundary_share)) {
    r$treatment_boundary_share <- NA_real_
  }
  r
}

fmt_boundary <- function(r, component) {
  value <- if (is.null(r)) NA_real_ else r[[component]]
  if (length(value) != 1L || !is.numeric(value) || !is.finite(value)) {
    "\u2014"
  } else {
    sprintf("%.4f", value)
  }
}

if (RERUN_TUNING) {
  for (cell_name in names(results)) {
    results[[cell_name]]$untuned <- add_boundary_shares(
      results[[cell_name]]$untuned, cells[[cell_name]]$Z
    )
    results[[cell_name]]$tuned <- add_boundary_shares(
      results[[cell_name]]$tuned, cells[[cell_name]]$Z
    )
  }

  rows <- do.call(rbind, lapply(names(results), function(cell_name) {
    R <- results[[cell_name]]
    data.frame(Cell = cell_name,
               Learner = c("XGB untuned", "XGB tuned (tune_on_folds=TRUE)"),
               `Estimate (SE)` = c(fmt_est(R$untuned), fmt_est(R$tuned)),
               `Instrument boundary share` = c(
                 fmt_boundary(R$untuned, "instrument_boundary_share"),
                 fmt_boundary(R$tuned, "instrument_boundary_share")
               ),
               `Treatment boundary share` = c(
                 fmt_boundary(R$untuned, "treatment_boundary_share"),
                 fmt_boundary(R$tuned, "treatment_boundary_share")
               ),
               check.names = FALSE, stringsAsFactors = FALSE)
  }))
} else {
  rows <- stored_export$estimate_comparison
}

rows %>%
  mutate(Cell = recode(
    Cell,
    "Labor (workedm)" = "Work Status",
    "Income (lincomem)" = "Maternal Log Income"
  )) %>%
  kbl(align = c("l","l","c","r","r"),
      caption = "Child Wald-AIPW: XGBoost untuned vs. fold-specifically tuned") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(rows$Learner == "XGB tuned (tune_on_folds=TRUE)"), background = "#eef6ff") %>%
  footnote(general = paste(
    "Boundary shares are reported separately as the fractions of fitted",
    "instrument and treatment propensities below 0.02 or above 0.98.",
    "Tuning uses nuisance-prediction loss within each outer training fold."
  ))
```

## 2. Nuisance-prediction losses

```{r nuisance-loss-table}
if (RERUN_TUNING) {
  loss_rows <- do.call(rbind, lapply(names(results), function(cell_name) {
    R <- results[[cell_name]]; cell <- cells[[cell_name]]
    ul <- calc_nuisance_losses(R$untuned, cell)
    tl <- calc_nuisance_losses(R$tuned, cell)
    data.frame(Cell = cell_name, Learner = c("XGB untuned", "XGB tuned"),
               `Outcome RMSE` = c(ul["outcome_rmse"], tl["outcome_rmse"]),
               `Instrument log-loss` = c(ul["instrument_logloss"], tl["instrument_logloss"]),
               `Treatment log-loss` = c(ul["treatment_logloss"], tl["treatment_logloss"]),
               check.names = FALSE, row.names = NULL)
  }))
} else {
  loss_rows <- stored_export$nuisance_losses
}

loss_rows %>%
  mutate(Cell = recode(
    Cell,
    "Labor (workedm)" = "Work Status",
    "Income (lincomem)" = "Maternal Log Income"
  )) %>%
  kbl(digits = 4, caption = "Nuisance-prediction loss: untuned vs. tuned (lower is better)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(loss_rows$Learner == "XGB tuned"), background = "#eef6ff") %>%
  footnote(general = "Losses are descriptive out-of-fold nuisance-prediction diagnostics; lower values indicate better prediction for the reported nuisance.")
```

## 3. Tuning protocol

The tuning design is the relevant reproducibility information because
selection occurs separately within each outer training fold.

```{r tuning-protocol}
data.frame(
  Component = c(
    "Outer cross-fitting", "Inner validation", "Search algorithm", "Budget",
    "nrounds", "max_depth", "eta", "min_child_weight",
    "Outcome nuisance", "Instrument nuisance", "Treatment nuisance"
  ),
  Specification = c(
    paste(N_CF, "folds"), paste(INNER_FOLDS, "folds"), "Random search",
    paste(TUNE_EVALS, "evaluations per nuisance and outer fold"),
    "50–400", "2–5", "0.02–0.20", "1–8",
    "RMSE", "Log-loss", "Log-loss"
  )
) %>%
  kbl(caption = "Nested XGBoost tuning protocol: Wald-AIPW") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE)
```

```{r hyperparameter-summary, include=FALSE}
extract_tuned_params <- function(r, cell_name) {
  if (is.null(r) || !r$tuned) return(data.frame())
  src <- if (!is.null(r$params)) r$params else r$tuning_result
  if (is.null(src)) return(data.frame())

  u <- unlist(src, recursive = TRUE)
  u <- u[grepl("nrounds|max_depth|eta|min_child_weight", names(u))]
  if (length(u) == 0) return(data.frame())

  hp <- data.frame(Cell = cell_name, Raw_name = names(u), Value = as.numeric(u), stringsAsFactors = FALSE)
  hp$Nuisance <- ifelse(grepl("ml_g", hp$Raw_name), "Outcome g",
                  ifelse(grepl("ml_m", hp$Raw_name), "Instrument m",
                  ifelse(grepl("ml_r", hp$Raw_name), "Treatment r", "Unknown")))
  hp$Branch <- ifelse(grepl("ml_g0|ml_r0", hp$Raw_name), "Z=0",
                ifelse(grepl("ml_g1|ml_r1", hp$Raw_name), "Z=1", ""))
  hp$Parameter <- sub(".*(nrounds|max_depth|eta|min_child_weight).*", "\\1", hp$Raw_name)
  hp
}

if (RERUN_TUNING) {
  hp_raw <- do.call(rbind, lapply(names(results), function(cn) {
    extract_tuned_params(results[[cn]]$tuned, cn)
  }))
  if (!nrow(hp_raw)) stop("Fold-specific hyperparameters could not be extracted.")
  hp_raw <- hp_raw[hp_raw$Nuisance != "Unknown", ]
  hp_summary <- hp_raw %>%
    group_by(Cell, Nuisance, Branch, Parameter) %>%
    summarise(Median_across_folds = median(Value), N_folds_seen = n(), .groups = "drop") %>%
    arrange(Cell, Nuisance, Branch, Parameter)
} else {
  hp_raw <- NULL
  hp_summary <- stored_export$hyperparameters
}
```

## 4. Outcome-weight representation audit and export

<div class="note-box">
Reconstructed weights enter subsequent diagnostics only when
\(\omega'Y\) reproduces the fitted coefficient.
</div>

```{r xgb-tuned-weight-table}
omega_vec <- function(om) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && "Wald-AIPW" %in% rownames(w)) return(as.vector(w["Wald-AIPW", ]))
  as.vector(w[1, ])
}

extract_xgb_w <- function(r, cell) {
  if (is.null(r) || is.null(r$obj)) {
    stop("Cannot extract weights because the fitted DoubleML object is NULL.")
  }
  dml_data <- make_dml_data(cell$Y, cell$D, cell$Z, cell$X)
  om <- get_outcome_weights(object = r$obj, dml_data = dml_data)
  omega_vec(om)
}

safe_extract_xgb_w <- function(r, cell) {
  tryCatch(extract_xgb_w(r, cell), error = function(e) NULL)
}

one_xgb_diag_row <- function(cell_name, learner_name, r, cell, w) {
  if (is.null(r) || is.null(r$obj)) {
    return(data.frame(Cell = cell_name, Learner = learner_name,
                      Estimate = NA_real_,
                      `Overall sum` = NA_real_,
                      `Treated sum` = NA_real_,
                      `Control sum` = NA_real_,
                      `ω′Y reproduces τ̂` = NA,
                      check.names = FALSE))
  }

  if (is.null(w)) {
    return(data.frame(Cell = cell_name, Learner = learner_name,
                      Estimate = as.numeric(r$coef),
                      `Overall sum` = NA_real_,
                      `Treated sum` = NA_real_,
                      `Control sum` = NA_real_,
                      `ω′Y reproduces τ̂` = NA,
                      check.names = FALSE))
  }

  identity_ok <- tryCatch(
    check_weight_identity(w, cell$Y, r$coef),
    error = function(e) NA
  )

  data.frame(Cell = cell_name, Learner = learner_name, Estimate = as.numeric(r$coef),
             `Overall sum` = sum(w),
             `Treated sum` = sum(w[cell$D == 1]),
             `Control sum` = -sum(w[cell$D == 0]),
             `ω′Y reproduces τ̂` = identity_ok,
             check.names = FALSE)
}

if (RERUN_TUNING) {
  xgb_omega_vectors <- lapply(names(cells), function(cell_name) {
    list(
      untuned = safe_extract_xgb_w(results[[cell_name]]$untuned,
                                    cells[[cell_name]]),
      tuned = safe_extract_xgb_w(results[[cell_name]]$tuned,
                                  cells[[cell_name]])
    )
  })
  names(xgb_omega_vectors) <- names(cells)

  xgb_tuned_weight_table <- do.call(rbind, lapply(names(cells), function(cell_name) {
    cell <- cells[[cell_name]]; R <- results[[cell_name]]
    W <- xgb_omega_vectors[[cell_name]]
    rbind(
      one_xgb_diag_row(cell_name, "XGB untuned", R$untuned, cell, W$untuned),
      one_xgb_diag_row(cell_name, "XGB tuned", R$tuned, cell, W$tuned)
    )
  }))
} else {
  xgb_omega_vectors <- stored_export$omega_vectors
  xgb_tuned_weight_table <- stored_export$weight_diagnostics
}

xgb_wald_audit_display <- do.call(rbind, lapply(
  seq_len(nrow(xgb_tuned_weight_table)),
  function(i) {
    row <- xgb_tuned_weight_table[i, , drop = FALSE]
    cell_name <- as.character(row$Cell)
    variant <- if (grepl("untuned", row$Learner, fixed = TRUE)) {
      "untuned"
    } else {
      "tuned"
    }
    w <- tryCatch(xgb_omega_vectors[[cell_name]][[variant]],
                  error = function(e) NULL)
    d <- cells[[cell_name]]$D
    estimate <- as.numeric(row$Estimate)
    data.frame(
      Cell = recode(cell_name,
                    "Labor (workedm)" = "Work Status",
                    "Income (lincomem)" = "Maternal Log Income"),
      Variant = if (variant == "untuned") "Untuned XGBoost" else "Tuned XGBoost",
      Estimate = estimate,
      `Overall sum` = if (is.null(w)) NA_real_ else sum(w),
      `Treated sum` = if (is.null(w)) NA_real_ else sum(w[d == 1]),
      `Control sum` = if (is.null(w)) NA_real_ else -sum(w[d == 0]),
      `ω′Y reproduces τ̂` = if (is.null(w) || !is.finite(estimate)) NA else
        check_weight_identity(w, cells[[cell_name]]$Y, estimate),
      check.names = FALSE
    )
  }
))

xgb_wald_audit_display %>%
  kbl(digits = 4,
      caption = "Angrist–Evans Wald-AIPW: Reconstructed XGBoost weight audit") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12) %>%
  row_spec(which(xgb_wald_audit_display$Variant == "Tuned XGBoost"), background = "#eef6ff") %>%
  footnote(general = paste(
    "The sums describe reconstructed outcome-weight vectors; control weights are reported with reversed sign.",
    "A normalized contrast has overall, treated, and control sums of 0, 1, and 1.",
    "Reproduction is assessed from the unrounded vectors and estimates using all.equal() with R's default tolerance.",
    "Failed vectors are excluded from further outcome-weight diagnostics."
  ))
```

---

### Export for child_test.Rmd

<div class="note-box">
Because tuning is fold-specific, the export stores the fitted estimates,
numeric coefficients, weight vectors, identity checks, and diagnostic tables.
The median hyperparameter summary is descriptive and is not reused as a fixed
recipe.
</div>

```{r xgb-identity-checks, include=FALSE}
if (RERUN_TUNING) {
  xgb_identity_checks <- lapply(names(cells), function(cell_name) {
    cell <- cells[[cell_name]]
    list(
      untuned = tryCatch(
        if (is.null(xgb_omega_vectors[[cell_name]]$untuned) ||
            is.null(results[[cell_name]]$untuned)) NA else
          check_weight_identity(xgb_omega_vectors[[cell_name]]$untuned,
                                cell$Y, results[[cell_name]]$untuned$coef),
        error = function(e) NA
      ),
      tuned = tryCatch(
        if (is.null(xgb_omega_vectors[[cell_name]]$tuned) ||
            is.null(results[[cell_name]]$tuned)) NA else
          check_weight_identity(xgb_omega_vectors[[cell_name]]$tuned,
                                cell$Y, results[[cell_name]]$tuned$coef),
        error = function(e) NA
      )
    )
  })
  names(xgb_identity_checks) <- names(cells)
} else {
  xgb_identity_checks <- stored_export$identity_checks
}
```

```{r xgb-export, cache=FALSE, include=FALSE}
if (RERUN_TUNING) {
xgb_export <- list(
  estimate_comparison = rows,
  fit_summaries = lapply(results, function(x) list(
    untuned = if (is.null(x$untuned)) {
      list(coef = NA_real_, se = NA_real_)
    } else {
      list(coef = x$untuned$coef, se = x$untuned$se)
    },
    tuned = if (is.null(x$tuned)) {
      list(coef = NA_real_, se = NA_real_)
    } else {
      list(coef = x$tuned$coef, se = x$tuned$se)
    }
  )),
  coefficients = lapply(results, function(x) list(
    untuned = if (is.null(x$untuned)) NA_real_ else x$untuned$coef,
    tuned = if (is.null(x$tuned)) NA_real_ else x$tuned$coef
  )),
  nuisance_losses     = loss_rows,
  hyperparameters     = if (exists("hp_summary")) hp_summary else NULL,
  weight_diagnostics  = xgb_tuned_weight_table,
  omega_vectors       = xgb_omega_vectors,
  identity_checks     = xgb_identity_checks,
  settings = list(
    seed          = SEED,
    n_cf          = N_CF,
    tune_evals    = TUNE_EVALS,
    inner_folds   = INNER_FOLDS,
    tune_on_folds = TRUE,
    learner       = "xgboost",
    dml_class     = "DoubleMLIIVM",
    score         = "LATE",
    covariates    = colnames(X_dml_labor),
    untuned_base  = list(g = list(alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
                         m = "classif.xgboost defaults with predict_type = 'prob'",
                         r = "classif.xgboost defaults with predict_type = 'prob'"),
    extract_weights = TRUE
  ),
  sample_info = list(
    labor_n       = nrow(ae98_labor_sub),
    income_n      = nrow(ae98_income_sub),
    labor_mean_Z  = mean(Z_labor),
    labor_mean_D  = mean(D_labor),
    income_mean_Z = mean(Z_income),
    income_mean_D = mean(D_income)
  )
)

saveRDS(xgb_export, EXPORT_PATH)
}
```

<div class="note-box">
The export lets `child_test.Rmd` display the fitted estimates and apply the
outcome-weight representation gate without repeating nested tuning.
</div>
