---
title: "Angrist & Evans (1998) — Main Analysis"
subtitle: "Reproduction tables, DML estimates, learner comparisons, and weight diagnostics"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body   { font-size: 15px; color: #2c2c2c; }
h1     { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2     { color: #2c7bb6; margin-top: 28px; }
h3     { color: #444; margin-top: 20px; }

.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  cache   = TRUE,
  autodep = TRUE,
  fig.width  = 12,
  fig.height = 6
)
```

---

# Setup

## Packages & Functions

```{r packages-and-source, include=FALSE, cache=FALSE}
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(cobalt)           # love.plot(), used inside make_love()
library(viridis)          # colour scale, used inside make_love()
library(OutcomeWeights)   # dml_with_smoother(), get_outcome_weights()
library(DoubleML)         # DoubleMLIIVM, DoubleMLData
library(mlr3verse)        # lrn() learners for linear/logit and ranger

lgr::get_logger("mlr3")$set_threshold("warn")

source(file.path("..", "functions_all.R"))
```

## Data & Design

<div class="note-box">
**Setting.** Angrist and Evans (1998) childbearing application. The instrument
is `samesex`; the treatment is `morekids`. The fixed diagnostic samples contain
3,000 observations for the primary labor outcome and 3,000 observations for
the primary log-income outcome. They are created in
`ae98_make_balanced_subsample_submission.Rmd`.

**Common covariate design.** Every estimator uses
`agem`, `agefstm`, `sexk`, `blackm`, `hispm`, and `othracem`. The second-child
sex indicator, `sex2nd`, is not
included because, together with first-child sex, it deterministically reveals the instrument
and creates exact collinearity within the instrument-specific nuisance regressions.
The common six-variable design also makes
the kappa, linear, forest, Ranger, and XGBoost comparisons use the same
information set. Angrist and Evans report that removing the individual
child-sex controls changes their estimates by no more than about two percent. See
[Angrist and Evans (1998)](https://www.dpipe.tsukuba.ac.jp/~naito/teaching_web/public_economics_2013_web/Angrist_Evans_AER_1998.pdf).

**Outcome coding.** Section 1 reports the alternative outcome recodings used
in the reproduction exercise. In all subsequent DML, learner-comparison,
outcome-weight, and covariate-balance analyses, work status is the binary
indicator `workedm`, and maternal log income is
`lincomem = log(incomem)`, where `incomem` is annual maternal income measured
in dollars. Alternative income units are therefore used only in the
reproduction and translation-invariance exercises.
</div>

```{r data, include=FALSE}
obj <- readRDS("ae98_sub.rds")
if (!all(c("ae98_labor_sub", "ae98_income_sub") %in% names(obj))) {
  stop("ae98_sub.rds must contain ae98_labor_sub and ae98_income_sub. Run ae98_make_balanced_subsample_submission.Rmd first.")
}

ae98_labor_sub  <- obj$ae98_labor_sub
ae98_income_sub <- obj$ae98_income_sub

ae98_labor_sub <- ae98_labor_sub %>%
  mutate(
    workedm    = as.integer(workedm),
    morekids   = as.integer(morekids),
    samesex    = as.integer(samesex),
    workedm_21 = workedm + 1L,        # 2 if worked, 1 otherwise
    workedm_01 = 1L - workedm         # 1 if did not work, 0 otherwise
  )

ae98_income_sub <- ae98_income_sub %>%
  mutate(
    morekids      = as.integer(morekids),
    samesex       = as.integer(samesex),
    lincomem      = log(incomem),             # dollars   -> col (5)
    lincomem_cnt  = log(100   * incomem),     # cents     -> col (4)
    lincomem_ths  = log(incomem / 1000),       # $1000s    -> col (6)
    lincomem_hths = log(incomem / 100000)      # $100,000s -> col (7)
  )

```

```{r design-matrices, include=FALSE}
CHILD_COVARIATES <- c(
  "agem", "agefstm", "sexk", "blackm", "hispm", "othracem"
)
RHS_FORMULA     <- reformulate(CHILD_COVARIATES)
RHS_DML_FORMULA <- reformulate(CHILD_COVARIATES, intercept = FALSE)

X_labor_kappa <- model.matrix(RHS_FORMULA, data = ae98_labor_sub)
X_labor_df    <- as.data.frame(X_labor_kappa[, -1, drop = FALSE])

X_income_kappa <- model.matrix(RHS_FORMULA, data = ae98_income_sub)
X_income_df    <- as.data.frame(X_income_kappa[, -1, drop = FALSE])

Z_labor <- ae98_labor_sub$samesex
D_labor <- ae98_labor_sub$morekids

Z_income <- ae98_income_sub$samesex
D_income <- ae98_income_sub$morekids

X_dml_labor  <- make_X(RHS_DML_FORMULA, data = ae98_labor_sub)
X_dml_income <- make_X(RHS_DML_FORMULA, data = ae98_income_sub)

```

```{r analysis-dimensions, echo=FALSE}
data.frame(
  Outcome = c("Work Status", "Maternal Log Income"),
  N = c(nrow(ae98_labor_sub), nrow(ae98_income_sub)),
  `Pr(Z=1)` = c(mean(Z_labor), mean(Z_income)),
  `Pr(D=1)` = c(mean(D_labor), mean(D_income)),
  Covariates = c(ncol(X_dml_labor), ncol(X_dml_income)),
  check.names = FALSE
) %>%
  kbl(digits = 3, caption = "Child-application analysis dimensions") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE)
```

---

# 1. Reproduction — SUW Table 4 (split by outcome family)

The two tables reproduce the recoding structure of SUW Table 4 on the fixed
diagnostic subsamples. They are not intended to reproduce the published
full-sample point estimates.

```{r build-columns, cache=TRUE, include=FALSE}
run_child_column <- function(Y, D, Z, X_kappa, X_df) {
  list(
    tsls  = run_2sls(Y, D, Z, X_df, endog_name = "D"),
    kappa = kappa_analytic_se_all(Y, Z, D, X_kappa)
  )
}

labor_cols <- list(
  "1 if worked, 0 else"    = run_child_column(ae98_labor_sub$workedm,
                                               D_labor, Z_labor, X_labor_kappa, X_labor_df),
  "2 if worked, 1 else"    = run_child_column(ae98_labor_sub$workedm_21,
                                               D_labor, Z_labor, X_labor_kappa, X_labor_df),
  "1 if did not work, 0 else" = run_child_column(ae98_labor_sub$workedm_01,
                                               D_labor, Z_labor, X_labor_kappa, X_labor_df)
)

income_cols <- list(
  "Cents"        = run_child_column(ae98_income_sub$lincomem_cnt,
                                     D_income, Z_income, X_income_kappa, X_income_df),
  "Dollars"      = run_child_column(ae98_income_sub$lincomem,
                                     D_income, Z_income, X_income_kappa, X_income_df),
  "$1000s"       = run_child_column(ae98_income_sub$lincomem_ths,
                                     D_income, Z_income, X_income_kappa, X_income_df),
  "$100,000s"    = run_child_column(ae98_income_sub$lincomem_hths,
                                     D_income, Z_income, X_income_kappa, X_income_df)
)

```

```{r table-builder, echo=FALSE}
cell_fmt <- function(est, se, digits = 3) {
  if (is.na(est)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0)
    return(sprintf("%.3f<br><span style='color:#888'>(NA)</span>", est))
  pval  <- 2 * pnorm(-abs(est / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s<br><span style='color:#888'>(%.3f)</span>", est, stars, se)
}

# flip_col applies SUW's display convention for the reflected labor outcome.
build_child_table <- function(col_list, flip_col = NULL) {
  ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
  labels <- c("\u03c4_cb_u","\u03c4_ml_u","\u03c4_ml_a10","\u03c4_ml_a","\u03c4_ml_t","\u03c4_ml_a0")

  fmt_one <- function(cn, est, se) {
    if (!is.null(flip_col) && cn == flip_col) est <- -est   # display-only sign flip; SE untouched
    cell_fmt(est, se)
  }

  row_2sls <- c(Estimator = "2SLS",
                vapply(names(col_list), function(cn)
                  fmt_one(cn, col_list[[cn]]$tsls$coef, col_list[[cn]]$tsls$se),
                  character(1)))

  kappa_rows <- lapply(seq_along(ests), function(i) {
    e <- ests[i]
    c(Estimator = labels[i],
      vapply(names(col_list), function(cn)
        fmt_one(cn, col_list[[cn]]$kappa$estimates[e], col_list[[cn]]$kappa$se[e]),
        character(1)))
  })

  df <- as.data.frame(do.call(rbind, c(list(row_2sls), kappa_rows)),
                       stringsAsFactors = FALSE)
  names(df) <- c("Estimator", names(col_list))
  rownames(df) <- NULL
  df
}
```

## Results {.tabset}

### Work Status

```{r table-labor, echo=FALSE}
build_child_table(labor_cols, flip_col = "1 if did not work, 0 else") %>%
  kbl(escape = FALSE,
      align = c("l", rep("c", length(labor_cols))),
      caption = "AE98 reproduction — Work Status") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",            1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",   2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators", 5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = paste(
    sprintf("Diagnostic subsample, N = %d; SUW Table 4 uses N = 394,840.", nrow(ae98_labor_sub)),
    "Each cell reports estimate with SE below. Blue rows are normalized; orange rows are unnormalized.",
    "The 'did not work' column follows SUW's reflected-outcome display convention by reversing",
    "the estimate sign while leaving the SE unchanged. Stars: * p<0.10, ** p<0.05, *** p<0.01."
  ))
```

### Maternal Log Income

```{r table-income, echo=FALSE}
build_child_table(income_cols) %>%
  kbl(escape = FALSE,
      align = c("l", rep("c", length(income_cols))),
      caption = "AE98 reproduction — Maternal Log Income") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",            1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",   2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators", 5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = paste(
    sprintf("Diagnostic subsample, N = %d; SUW Table 4 uses N = 220,502.", nrow(ae98_income_sub)),
    "Each cell reports estimate with SE below. Blue rows are normalized; orange rows are unnormalized.",
    "Common covariates: agem, agefstm, sexk, blackm, hispm, othracem.",
    "Stars: * p<0.10, ** p<0.05, *** p<0.01."
  ))
```

---

# 2. Double ML — PLR-IV & Wald-AIPW (honest forest, tuned vs. untuned)

<div class="note-box">
Each outcome-specific sample is fitted with untuned
(`tune.parameters = "none"`) and tuned (`"all"`) honest forests. All four
fits use five cross-fitting folds and seed 42. The income outcome in every fit
is log annual maternal income measured in dollars.
</div>

```{r dml-fit-all, cache=TRUE, include=FALSE}
set.seed(42); dml_labor_untuned  <- dml_with_smoother(
  ae98_labor_sub$workedm, D_labor, X_dml_labor, Z_labor,
  n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_labor_tuned    <- dml_with_smoother(
  ae98_labor_sub$workedm, D_labor, X_dml_labor, Z_labor,
  n_cf_folds = 5, tune.parameters = "all")

set.seed(42); dml_income_untuned <- dml_with_smoother(
  ae98_income_sub$lincomem, D_income, X_dml_income, Z_income,
  n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_income_tuned   <- dml_with_smoother(
  ae98_income_sub$lincomem, D_income, X_dml_income, Z_income,
  n_cf_folds = 5, tune.parameters = "all")

res_labor_untuned  <- summary(dml_labor_untuned)
res_labor_tuned    <- summary(dml_labor_tuned)
res_income_untuned <- summary(dml_income_untuned)
res_income_tuned   <- summary(dml_income_tuned)

extract_w <- function(dml_obj) {
  om <- get_outcome_weights(dml_obj)
  check_omega_rows(om)
  list(plriv    = as.vector(om$omega["PLR-IV",    ]),
       waldaipw = as.vector(om$omega["Wald-AIPW", ]))
}
w_labor_untuned  <- extract_w(dml_labor_untuned)
w_labor_tuned    <- extract_w(dml_labor_tuned)
w_income_untuned <- extract_w(dml_income_untuned)
w_income_tuned   <- extract_w(dml_income_tuned)

```

```{r dml-cell-display, echo=FALSE}
summary_value <- function(res, row, candidates) {
  hit <- intersect(candidates, colnames(res))
  if (!length(hit)) return(NA_real_)
  as.numeric(res[row, hit[1]])
}

dml_estimate_table <- function(res, wlist, Y, caption) {
  estimators <- c("PLR-IV", "Wald-AIPW")
  weights <- list(wlist$plriv, wlist$waldaipw)
  out <- data.frame(
    Estimator = estimators,
    Estimate = unname(vapply(estimators, function(x) get_estimate(res, x), numeric(1))),
    `Standard Error` = unname(vapply(estimators, function(x) {
      summary_value(res, x, c("SE", "Std. Error", "std err", "Std.Error"))
    }, numeric(1))),
    `ω′Y reproduces τ̂` = vapply(seq_along(estimators), function(i) {
      check_weight_identity(weights[[i]], Y, get_estimate(res, estimators[i]))
    }, logical(1)),
    check.names = FALSE
  )
  rownames(out) <- NULL
  out %>%
    kbl(
      digits = 4, row.names = FALSE, caption = caption,
      align = c("l", "r", "r", "c")
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE, font_size = 13
    ) %>%
    footnote(general = paste(
      "Numerical reproduction follows the all.equal() check",
      "with R's default tolerance."
    ))
}
```

## Estimates {.tabset}

### Work Status

```{r dml-labor-display, echo=FALSE}
dml_estimate_table(
  res_labor_untuned, w_labor_untuned, ae98_labor_sub$workedm,
  'Work Status — untuned honest forest (tune.parameters = "none")'
)
dml_estimate_table(
  res_labor_tuned, w_labor_tuned, ae98_labor_sub$workedm,
  'Work Status — tuned honest forest (tune.parameters = "all")'
)
```

### Maternal Log Income

```{r dml-income-display, echo=FALSE}
dml_estimate_table(
  res_income_untuned, w_income_untuned, ae98_income_sub$lincomem,
  'Maternal Log Income — untuned honest forest (tune.parameters = "none")'
)
dml_estimate_table(
  res_income_tuned, w_income_tuned, ae98_income_sub$lincomem,
  'Maternal Log Income — tuned honest forest (tune.parameters = "all")'
)
```

## DML vs. Kappa — Point Estimates

```{r dml-kappa-comparison, echo=FALSE}
dml_comparison_cell <- function(coef, se) {
  if (!is.finite(coef)) return("\u2014")
  if (!is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  sprintf("%.3f (%.3f)", coef, se)
}

dml_kappa_ref_row <- function(label, key) {
  data.frame(
    Estimator = label,
    `Work Status` = dml_comparison_cell(
      labor_cols[["1 if worked, 0 else"]]$kappa$estimates[key],
      labor_cols[["1 if worked, 0 else"]]$kappa$se[key]
    ),
    `Maternal Log Income` = dml_comparison_cell(
      income_cols[["Dollars"]]$kappa$estimates[key],
      income_cols[["Dollars"]]$kappa$se[key]
    ),
    check.names = FALSE
  )
}

dml_kappa_rows <- rbind(
  data.frame(
    Estimator = "PLR-IV (DML; tuned honest forest)",
    `Work Status` = dml_comparison_cell(
      get_estimate(res_labor_tuned, "PLR-IV"),
      summary_value(res_labor_tuned, "PLR-IV",
                    c("SE", "Std. Error", "std err", "Std.Error"))
    ),
    `Maternal Log Income` = dml_comparison_cell(
      get_estimate(res_income_tuned, "PLR-IV"),
      summary_value(res_income_tuned, "PLR-IV",
                    c("SE", "Std. Error", "std err", "Std.Error"))
    ),
    check.names = FALSE
  ),
  data.frame(
    Estimator = "Wald-AIPW (DML; tuned honest forest)",
    `Work Status` = dml_comparison_cell(
      get_estimate(res_labor_tuned, "Wald-AIPW"),
      summary_value(res_labor_tuned, "Wald-AIPW",
                    c("SE", "Std. Error", "std err", "Std.Error"))
    ),
    `Maternal Log Income` = dml_comparison_cell(
      get_estimate(res_income_tuned, "Wald-AIPW"),
      summary_value(res_income_tuned, "Wald-AIPW",
                    c("SE", "Std. Error", "std err", "Std.Error"))
    ),
    check.names = FALSE
  ),
  dml_kappa_ref_row("τ_cb_u (kappa, ref)", "tau_cb_u"),
  dml_kappa_ref_row("τ_ml_u (kappa, ref)", "tau_ml_u")
)

dml_kappa_rows %>%
  kbl(
    align = c("l", "c", "c"),
    row.names = FALSE,
    caption = "Tuned honest-forest DML versus normalized kappa estimates"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE, font_size = 13
  ) %>%
  pack_rows("Double ML", 1, 2) %>%
  pack_rows("Kappa reference", 3, 4) %>%
  footnote(general = "Each cell reports estimate (SE). All estimators use the common six-covariate specification.")
```

---

# 3. Learner Comparison — PLR-IV and Wald-AIPW

<div class="note-box">
The comparison uses the same learner families as Vietnam and Card. Every
learner is fitted on the common six-covariate design introduced above, so
differences are not confounded by learner-specific covariate sets.
</div>

```{r ext-setup, include=FALSE}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X))
}

dat_labor      <- make_dml_data(ae98_labor_sub$workedm,   D_labor,  Z_labor,  X_dml_labor)
dat_income     <- make_dml_data(ae98_income_sub$lincomem, D_income, Z_income, X_dml_income)

```

```{r ext-fit-helpers, echo=FALSE}
ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))

ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))
omega_vec <- function(om, estimator) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && estimator %in% rownames(w)) {
    return(as.vector(w[estimator, ]))
  }
  as.vector(w[1, ])
}

fit_iivm <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLIIVM$new(data = dml_data,
                          ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5, score = "LATE")
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  om  <- get_outcome_weights(object = obj, dml_data = dml_data)
  w   <- omega_vec(om, "Wald-AIPW")
  ok  <- check_weight_identity(w, Y, get_dml_coef(obj))
  list(obj = obj, w = w, ok = ok)
}

fit_pliv <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLPLIV$new(data = dml_data,
                          ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5)
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  om  <- get_outcome_weights(object = obj, dml_data = dml_data)
  w   <- omega_vec(om, "PLR-IV")
  ok  <- check_weight_identity(w, Y, get_dml_coef(obj))
  list(obj = obj, w = w, ok = ok)
}

```

```{r ext-fit-lm, cache=TRUE, include=FALSE}
lm_labor       <- fit_iivm(dat_labor,  ml_lm_iivm(), ae98_labor_sub$workedm)
lm_income      <- fit_iivm(dat_income, ml_lm_iivm(), ae98_income_sub$lincomem)
pliv_lm_labor  <- fit_pliv(dat_labor,  ml_lm_pliv(), ae98_labor_sub$workedm)
pliv_lm_income <- fit_pliv(dat_income, ml_lm_pliv(), ae98_income_sub$lincomem)

```

```{r ext-fit-rf, cache=TRUE, include=FALSE}
rf_labor       <- fit_iivm(dat_labor,  ml_rf_iivm(), ae98_labor_sub$workedm)
rf_income      <- fit_iivm(dat_income, ml_rf_iivm(), ae98_income_sub$lincomem)
pliv_rf_labor  <- fit_pliv(dat_labor,  ml_rf_pliv(), ae98_labor_sub$workedm)
pliv_rf_income <- fit_pliv(dat_income, ml_rf_pliv(), ae98_income_sub$lincomem)

```

```{r ext-xgb-import, include=FALSE}
validate_child_xgb_export <- function(x, expected_class, path) {
  required <- c(
    "estimate_comparison", "nuisance_losses", "weight_diagnostics",
    "omega_vectors", "identity_checks", "settings", "sample_info"
  )
  if (!all(required %in% names(x))) {
    stop(path, " has an incomplete export schema. Rerun its tuning notebook.")
  }
  if (!identical(x$settings$dml_class, expected_class)) {
    stop(path, " contains the wrong DoubleML estimator class.")
  }
  if (!isTRUE(x$settings$tune_on_folds)) {
    stop(path, " was not generated with fold-specific tuning.")
  }
  if (!identical(x$settings$covariates, colnames(X_dml_labor))) {
    stop(path, " does not use the common six-covariate CHILD design.")
  }
  required_cells <- c("Labor (workedm)", "Income (lincomem)")
  if (!all(required_cells %in% names(x$omega_vectors))) {
    stop(path, " does not contain both CHILD outcome cells.")
  }

  info <- x$sample_info
  valid_sample <-
    !is.null(info) &&
    identical(as.integer(info$labor_n), as.integer(nrow(ae98_labor_sub))) &&
    identical(as.integer(info$income_n), as.integer(nrow(ae98_income_sub))) &&
    isTRUE(all.equal(as.numeric(info$labor_mean_Z), mean(Z_labor), tolerance = 1e-12)) &&
    isTRUE(all.equal(as.numeric(info$labor_mean_D), mean(D_labor), tolerance = 1e-12)) &&
    isTRUE(all.equal(as.numeric(info$income_mean_Z), mean(Z_income), tolerance = 1e-12)) &&
    isTRUE(all.equal(as.numeric(info$income_mean_D), mean(D_income), tolerance = 1e-12))
  if (!valid_sample) {
    stop(path, " does not match the current CHILD subsamples.")
  }
  invisible(TRUE)
}

xgb_export_path      <- "ae98_xgb_tuning_export.rds"
xgb_pliv_export_path <- "ae98_xgb_pliv_tuning_export.rds"
if (!file.exists(xgb_export_path) || !file.exists(xgb_pliv_export_path)) {
  stop("Run both CHILD XGBoost tuning notebooks before child_test.Rmd.")
}

xgb_export      <- readRDS(xgb_export_path)
xgb_pliv_export <- readRDS(xgb_pliv_export_path)
validate_child_xgb_export(xgb_export, "DoubleMLIIVM", xgb_export_path)
validate_child_xgb_export(xgb_pliv_export, "DoubleMLPLIV", xgb_pliv_export_path)
```

```{r learner-display-helpers, echo=FALSE}
ext_cell <- function(coef, se) {
  if (is.na(coef) || !is.finite(coef)) return("\u2014")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

learner_row <- function(label, fit_labor, fit_income) {
  data.frame(Estimator = label,
             `Work Status` = ext_cell(get_dml_coef(fit_labor$obj),
                                      get_dml_se(fit_labor$obj)),
             `Maternal Log Income` = ext_cell(get_dml_coef(fit_income$obj),
                                              get_dml_se(fit_income$obj)),
             check.names = FALSE, stringsAsFactors = FALSE)
}

kappa_ref_row <- function(label, key) {
  data.frame(Estimator = label,
             `Work Status` = ext_cell(
               labor_cols[["1 if worked, 0 else"]]$kappa$estimates[key],
               labor_cols[["1 if worked, 0 else"]]$kappa$se[key]
             ),
             `Maternal Log Income` = ext_cell(
               income_cols[["Dollars"]]$kappa$estimates[key],
               income_cols[["Dollars"]]$kappa$se[key]
             ),
             check.names = FALSE, stringsAsFactors = FALSE)
}

export_identity <- function(export, sample_label, tuned_label) {
  out <- tryCatch(
    export$identity_checks[[sample_label]][[tuned_label]],
    error = function(e) NA
  )
  if (is.null(out) || !length(out)) NA else as.logical(out[1])
}

extract_se_from_display <- function(x) {
  hit <- regmatches(x, regexec("\\(([-+0-9.eE]+)\\)", x))[[1]]
  if (length(hit) < 2L) NA_real_ else as.numeric(hit[2])
}

xgb_stats <- function(export, sample_label, estimator, variant = "tuned") {
  tab <- export$estimate_comparison
  keep <- tab$Cell == sample_label
  if (variant == "tuned") {
    keep <- keep & grepl("tune_on_folds=TRUE", tab$Learner, fixed = TRUE)
  } else {
    keep <- keep & grepl("untuned$", tab$Learner)
  }
  display <- if (sum(keep) == 1L) tab[["Estimate (SE)"]][keep][1] else "—"
  stored_summary <- tryCatch(
    export$fit_summaries[[sample_label]][[variant]],
    error = function(e) NULL
  )
  list(
    estimator = estimator,
    coef = if (is.null(stored_summary)) {
      tryCatch(as.numeric(export$coefficients[[sample_label]][[variant]]),
               error = function(e) NA_real_)
    } else {
      as.numeric(stored_summary$coef)
    },
    se = if (is.null(stored_summary)) {
      extract_se_from_display(display)
    } else {
      as.numeric(stored_summary$se)
    },
    ok = export_identity(export, sample_label, variant)
  )
}

```

```{r ext-comparison-build, include=FALSE, cache.extra="quiet-doubleml-summary-v2"}
xgb_pliv_labor_tuned <- xgb_stats(
  xgb_pliv_export, "Labor (workedm)", "PLR-IV"
)
xgb_pliv_income_tuned <- xgb_stats(
  xgb_pliv_export, "Income (lincomem)", "PLR-IV"
)
xgb_wald_labor_tuned <- xgb_stats(
  xgb_export, "Labor (workedm)", "Wald-AIPW"
)
xgb_wald_income_tuned <- xgb_stats(
  xgb_export, "Income (lincomem)", "Wald-AIPW"
)

xgb_comparison_row <- function(label, labor, income) {
  data.frame(
    Estimator = label,
    `Work Status` = ext_cell(labor$coef, labor$se),
    `Maternal Log Income` = ext_cell(income$coef, income$se),
    check.names = FALSE
  )
}

comp_df <- rbind(
  learner_row("PLR-IV (OLS)", pliv_lm_labor, pliv_lm_income),
  learner_row("PLR-IV (ranger)", pliv_rf_labor, pliv_rf_income),
  xgb_comparison_row(
    "PLR-IV (XGBoost tuned)",
    xgb_pliv_labor_tuned, xgb_pliv_income_tuned
  ),
  learner_row(
    "Wald-AIPW (linear+logit)",
    lm_labor, lm_income
  ),
  learner_row("Wald-AIPW (ranger)", rf_labor, rf_income),
  xgb_comparison_row(
    "Wald-AIPW (XGBoost tuned)",
    xgb_wald_labor_tuned, xgb_wald_income_tuned
  ),
  kappa_ref_row("τ_cb_u (kappa, ref)", "tau_cb_u"),
  kappa_ref_row("τ_ml_u (kappa, ref)", "tau_ml_u")
)
```

## Estimates {.tabset}

### Work Status

```{r learner-estimates-labor, echo=FALSE, cache.extra="quiet-doubleml-summary-v2"}
comp_df %>%
  select(Estimator, `Work Status`) %>%
  kbl(
    align = c("l", "c"), row.names = FALSE,
    caption = "Learner comparison — Work Status"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE, font_size = 13
  ) %>%
  pack_rows("PLR-IV", 1, 3) %>%
  pack_rows("Wald-AIPW", 4, 6) %>%
  pack_rows("Kappa reference", 7, 8) %>%
  row_spec(1:6, background = "#eef6ff") %>%
  footnote(general = paste(
    "Cells report estimate (SE). XGBoost rows use the stored",
    "fold-specifically tuned fits. All estimators use the common",
    "six-covariate specification."
  ))
```

### Maternal Log Income

```{r learner-estimates-income, echo=FALSE, cache.extra="quiet-doubleml-summary-v2"}
comp_df %>%
  select(Estimator, `Maternal Log Income`) %>%
  kbl(
    align = c("l", "c"), row.names = FALSE,
    caption = "Learner comparison — Maternal Log Income"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE, font_size = 13
  ) %>%
  pack_rows("PLR-IV", 1, 3) %>%
  pack_rows("Wald-AIPW", 4, 6) %>%
  pack_rows("Kappa reference", 7, 8) %>%
  row_spec(1:6, background = "#eef6ff") %>%
  footnote(general = paste(
    "Cells report estimate (SE). XGBoost rows use the stored",
    "fold-specifically tuned fits. All estimators use the common",
    "six-covariate specification."
  ))
```

---

# 4. Weight Diagnostics

<div class="note-box">
The same candidate registry is used for the summary statistics, ESS and
sign-cancellation results, and Love plots. Algebraic kappa weights and DML
vectors that reproduce their fitted estimates are eligible. The honest-forest
entries use the tuned fits; untuned fits remain in the tuning comparison.
</div>

```{r diag-builders, echo=FALSE}
kw_labor  <- kappa_weights_bundle(Z_labor,  D_labor,  X_labor_kappa)
kw_income <- kappa_weights_bundle(Z_income, D_income, X_income_kappa)

get_xgb_w  <- function(sample_label, tuned_label) {
  if (!exists("xgb_export")) return(NULL)
  out <- tryCatch(xgb_export$omega_vectors[[sample_label]][[tuned_label]], error = function(e) NULL)
  out
}
get_xgb_ok <- function(sample_label, tuned_label) {
  if (!exists("xgb_export")) return(NA)
  out <- tryCatch(xgb_export$identity_checks[[sample_label]][[tuned_label]], error = function(e) NA)
  if (is.null(out)) NA else out
}
get_xgb_pliv_w <- function(sample_label, tuned_label) {
  if (!exists("xgb_pliv_export")) return(NULL)
  out <- tryCatch(xgb_pliv_export$omega_vectors[[sample_label]][[tuned_label]], error = function(e) NULL)
  out
}
get_xgb_pliv_ok <- function(sample_label, tuned_label) {
  if (!exists("xgb_pliv_export")) return(NA)
  out <- tryCatch(xgb_pliv_export$identity_checks[[sample_label]][[tuned_label]], error = function(e) NA)
  if (is.null(out)) NA else out
}

w_xgb_labor_untuned   <- get_xgb_w("Labor (workedm)",   "untuned")
w_xgb_labor_tuned     <- get_xgb_w("Labor (workedm)",   "tuned")
w_xgb_income_untuned  <- get_xgb_w("Income (lincomem)", "untuned")
w_xgb_income_tuned    <- get_xgb_w("Income (lincomem)", "tuned")
w_xgb_pliv_labor_untuned   <- get_xgb_pliv_w("Labor (workedm)",   "untuned")
w_xgb_pliv_labor_tuned     <- get_xgb_pliv_w("Labor (workedm)",   "tuned")
w_xgb_pliv_income_untuned  <- get_xgb_pliv_w("Income (lincomem)", "untuned")
w_xgb_pliv_income_tuned    <- get_xgb_pliv_w("Income (lincomem)", "tuned")
ok_xgb_labor_untuned  <- get_xgb_ok("Labor (workedm)",   "untuned")
ok_xgb_labor_tuned    <- get_xgb_ok("Labor (workedm)",   "tuned")
ok_xgb_income_untuned <- get_xgb_ok("Income (lincomem)", "untuned")
ok_xgb_income_tuned   <- get_xgb_ok("Income (lincomem)", "tuned")
ok_xgb_pliv_labor_untuned  <- get_xgb_pliv_ok("Labor (workedm)",   "untuned")
ok_xgb_pliv_labor_tuned    <- get_xgb_pliv_ok("Labor (workedm)",   "tuned")
ok_xgb_pliv_income_untuned <- get_xgb_pliv_ok("Income (lincomem)", "untuned")
ok_xgb_pliv_income_tuned   <- get_xgb_pliv_ok("Income (lincomem)", "tuned")

# The same candidate registry drives every diagnostic table and Love plot.
# Algebraic kappa weights are eligible by construction. Reconstructed weights
# are eligible only when omega'Y reproduces the fitted coefficient under the
# canonical all.equal() rule in functions_all.R.
candidate <- function(w, estimate = NA_real_, ok = TRUE, source = "verified") {
  list(w = w, estimate = as.numeric(estimate), ok = isTRUE(ok), source = source)
}

export_coef <- function(export, sample_label, tuned_label) {
  if (is.null(export)) return(NA_real_)
  out <- tryCatch(export$coefficients[[sample_label]][[tuned_label]],
                  error = function(e) NULL)
  if (!is.null(out) && length(out)) return(as.numeric(out[1]))

  # Backward-compatible fallback for exports created before numeric
  # coefficients were stored separately.
  tab <- export$estimate_comparison
  if (is.null(tab) || !nrow(tab)) return(NA_real_)
  learner_label <- if (identical(export$settings$dml_class, "DoubleMLPLIV")) {
    if (tuned_label == "tuned") "PLR-IV XGB tuned (tune_on_folds=TRUE)" else "PLR-IV XGB untuned"
  } else {
    if (tuned_label == "tuned") "XGB tuned (tune_on_folds=TRUE)" else "XGB untuned"
  }
  row <- tab[tab$Cell == sample_label & tab$Learner == learner_label, , drop = FALSE]
  if (!nrow(row)) return(NA_real_)
  suppressWarnings(as.numeric(sub("^[[:space:]]*([-+0-9.eE]+).*", "\\1",
                                  row[["Estimate (SE)"]][1])))
}

build_candidates <- function(Y, w_untuned, res_untuned, w_tuned, res_tuned,
                             pliv_lm, iivm_lm, pliv_rf, iivm_rf,
                             xgb_pliv_export, xgb_wald_export,
                             xgb_pliv_w_untuned, xgb_pliv_w_tuned,
                             xgb_wald_w_untuned, xgb_wald_w_tuned,
                             xgb_pliv_ok_untuned, xgb_pliv_ok_tuned,
                             xgb_wald_ok_untuned, xgb_wald_ok_tuned,
                             sample_label, kw) {
  pliv_u <- get_estimate(res_untuned, "PLR-IV")
  wald_u <- get_estimate(res_untuned, "Wald-AIPW")
  pliv_t <- get_estimate(res_tuned, "PLR-IV")
  wald_t <- get_estimate(res_tuned, "Wald-AIPW")
  list(
    "PLR-IV, untuned forest" = candidate(
      w_untuned$plriv, pliv_u, check_weight_identity(w_untuned$plriv, Y, pliv_u)),
    "Wald-AIPW, untuned forest" = candidate(
      w_untuned$waldaipw, wald_u, check_weight_identity(w_untuned$waldaipw, Y, wald_u)),
    "PLR-IV, tuned forest" = candidate(
      w_tuned$plriv, pliv_t, check_weight_identity(w_tuned$plriv, Y, pliv_t)),
    "Wald-AIPW, tuned forest" = candidate(
      w_tuned$waldaipw, wald_t, check_weight_identity(w_tuned$waldaipw, Y, wald_t)),
    "PLR-IV, linear" = candidate(pliv_lm$w, get_dml_coef(pliv_lm$obj), pliv_lm$ok),
    "Wald-AIPW, linear/logit" = candidate(iivm_lm$w, get_dml_coef(iivm_lm$obj), iivm_lm$ok),
    "PLR-IV, ranger" = candidate(pliv_rf$w, get_dml_coef(pliv_rf$obj), pliv_rf$ok),
    "Wald-AIPW, ranger" = candidate(iivm_rf$w, get_dml_coef(iivm_rf$obj), iivm_rf$ok),
    "PLR-IV, untuned XGBoost" = candidate(
      xgb_pliv_w_untuned, export_coef(xgb_pliv_export, sample_label, "untuned"),
      xgb_pliv_ok_untuned, "reconstructed"),
    "PLR-IV, tuned XGBoost" = candidate(
      xgb_pliv_w_tuned, export_coef(xgb_pliv_export, sample_label, "tuned"),
      xgb_pliv_ok_tuned, "reconstructed"),
    "Wald-AIPW, untuned XGBoost" = candidate(
      xgb_wald_w_untuned, export_coef(xgb_wald_export, sample_label, "untuned"),
      xgb_wald_ok_untuned, "reconstructed"),
    "Wald-AIPW, tuned XGBoost" = candidate(
      xgb_wald_w_tuned, export_coef(xgb_wald_export, sample_label, "tuned"),
      xgb_wald_ok_tuned, "reconstructed"),
    "\u03c4_cb_u (kappa, CBPS)" = candidate(kw$tau_cb_u, source = "algebraic"),
    "\u03c4_ml_u (kappa, MLE)" = candidate(kw$tau_ml_u, source = "algebraic"),
    "\u03c4_ml_a10 (kappa, MLE)" = candidate(kw$tau_ml_a10, source = "algebraic"),
    "\u03c4_ml_a (kappa, MLE)" = candidate(kw$tau_ml_a, source = "algebraic"),
    "\u03c4_ml_t (kappa, MLE)" = candidate(kw$tau_ml_t, source = "algebraic"),
    "\u03c4_ml_a0 (kappa, MLE)" = candidate(kw$tau_ml_a0, source = "algebraic")
  )
}

xgb_wald_obj <- if (exists("xgb_export")) xgb_export else NULL
xgb_pliv_obj <- if (exists("xgb_pliv_export")) xgb_pliv_export else NULL

candidate_cells <- list(
  "Labor (workedm)" = list(
    Y = ae98_labor_sub$workedm, D = D_labor, X = X_dml_labor,
    candidates = build_candidates(
      ae98_labor_sub$workedm,
      w_labor_untuned, res_labor_untuned, w_labor_tuned, res_labor_tuned,
      pliv_lm_labor, lm_labor, pliv_rf_labor, rf_labor,
      xgb_pliv_obj, xgb_wald_obj,
      w_xgb_pliv_labor_untuned, w_xgb_pliv_labor_tuned,
      w_xgb_labor_untuned, w_xgb_labor_tuned,
      ok_xgb_pliv_labor_untuned, ok_xgb_pliv_labor_tuned,
      ok_xgb_labor_untuned, ok_xgb_labor_tuned,
      "Labor (workedm)", kw_labor)),
  "Income (lincomem)" = list(
    Y = ae98_income_sub$lincomem, D = D_income, X = X_dml_income,
    candidates = build_candidates(
      ae98_income_sub$lincomem,
      w_income_untuned, res_income_untuned, w_income_tuned, res_income_tuned,
      pliv_lm_income, lm_income, pliv_rf_income, rf_income,
      xgb_pliv_obj, xgb_wald_obj,
      w_xgb_pliv_income_untuned, w_xgb_pliv_income_tuned,
      w_xgb_income_untuned, w_xgb_income_tuned,
      ok_xgb_pliv_income_untuned, ok_xgb_pliv_income_tuned,
      ok_xgb_income_untuned, ok_xgb_income_tuned,
      "Income (lincomem)", kw_income))
)

diagnostic_estimators <- c(
  "PLR-IV, tuned forest", "Wald-AIPW, tuned forest",
  "PLR-IV, linear", "Wald-AIPW, linear/logit",
  "PLR-IV, ranger", "Wald-AIPW, ranger",
  "PLR-IV, tuned XGBoost", "Wald-AIPW, tuned XGBoost",
  "τ_cb_u (kappa, CBPS)", "τ_ml_u (kappa, MLE)",
  "τ_ml_a10 (kappa, MLE)", "τ_ml_a (kappa, MLE)",
  "τ_ml_t (kappa, MLE)", "τ_ml_a0 (kappa, MLE)"
)

eligible_weight_list <- function(cell) {
  selected <- cell$candidates[
    intersect(diagnostic_estimators, names(cell$candidates))
  ]
  eligible <- Filter(function(x) isTRUE(x$ok), selected)
  setNames(lapply(eligible, function(x) x$w), names(eligible))
}

diagnostic_cells <- lapply(candidate_cells, function(cell) {
  list(D = cell$D, w = eligible_weight_list(cell))
})
```

## Outcome-Weight Summary Statistics {.tabset}

<div class="note-box">
The tables report statistics for the overall vector and its treated and
sign-oriented control components. `Neg. share` is the fraction of observations
with a negative oriented weight and differs from the absolute-mass
sign-cancellation measure reported below. Only eligible representations are
included.
</div>

```{r detailed-weight-helpers, echo=FALSE}
detailed_weight_table <- function(cell, caption) {
  estimators <- names(cell$w)
  statistics <- do.call(rbind, lapply(estimators, function(estimator) {
    weight_bridge(cell$w[[estimator]], cell$D, estimator)
  }))
  display <- statistics[, c(
    "Scope", "Min", "Max", "Neg_share", "Sum_top10", "Sum_w", "Sum_abs_w"
  )]

  out <- display %>%
    kbl(
      caption = caption, row.names = FALSE,
      col.names = c("Scope", "Min", "Max", "Neg. share", "ΣTop10%", "Σω", "Σ|ω|"),
      align = c("l", rep("r", 6))
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = TRUE, font_size = 12
    ) %>%
    add_header_above(c(" " = 1, "Outcome-weight summary statistics" = 6))

  for (i in seq_along(estimators)) {
    out <- pack_rows(out, estimators[i], (i - 1) * 3 + 1, i * 3)
  }

  out %>%
    footnote(general = paste(
      "Control rows use the sign-flipped −ω convention.",
      "ΣTop10% is the sum of the largest ten percent of weights within the row.",
      "Only algebraic or numerically verified representations are included."
    ))
}
```

### Work Status

```{r detailed-weight-labor, echo=FALSE}
detailed_weight_table(
  diagnostic_cells[["Labor (workedm)"]],
  "Outcome-weight summary statistics — Work Status"
)
```

### Maternal Log Income

```{r detailed-weight-income, echo=FALSE}
detailed_weight_table(
  diagnostic_cells[["Income (lincomem)"]],
  "Outcome-weight summary statistics — Maternal Log Income"
)
```

## ESS and Sign-Cancellation Extension {.tabset}

<div class="note-box">
Modified ESS is reported separately for treated and control observations and
relative to the observed size of each treatment arm. Opposite-sign mass is the
percentage of absolute weight mass with the opposite orientation within the
respective arm. It measures sign cancellation rather than the number of
negative observations.
</div>

```{r weight-extension-helpers, echo=FALSE}
weight_extension_table <- function(cell, caption) {
  format_pair <- function(a, b, digits) {
    if (any(!is.finite(c(a, b)))) return(NA_character_)
    paste0(
      formatC(a, format = "f", digits = digits), " / ",
      formatC(b, format = "f", digits = digits)
    )
  }

  results <- do.call(rbind, lapply(names(cell$w), function(estimator) {
    x <- outcome_weight_diagnostics(cell$w[[estimator]], cell$D, estimator)
    data.frame(
      Estimator = estimator,
      `ESS mod T/C` = format_pair(x$ESS_mod_T, x$ESS_mod_C, 0),
      `ESS mod/N T/C` = format_pair(
        x$ESS_mod_T / sum(cell$D == 1),
        x$ESS_mod_C / sum(cell$D == 0), 3
      ),
      `Opposite-sign mass T/C (%)` = format_pair(
        100 * x$Wrong_mass_T, 100 * x$Wrong_mass_C, 1
      ),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }))

  results %>%
    kbl(caption = caption, row.names = FALSE, align = c("l", "c", "c", "c")) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE, font_size = 12
    ) %>%
    footnote(general = paste(
      "T/C denotes treated / sign-oriented control.",
      "Only algebraic or numerically verified representations are included."
    ))
}
```

### Work Status

```{r weight-extension-labor, echo=FALSE}
weight_extension_table(
  diagnostic_cells[["Labor (workedm)"]],
  "Modified ESS and sign cancellation — Work Status"
)
```

### Maternal Log Income

```{r weight-extension-income, echo=FALSE}
weight_extension_table(
  diagnostic_cells[["Income (lincomem)"]],
  "Modified ESS and sign cancellation — Maternal Log Income"
)
```

# 5. Love Plots

<div class="note-box">
Panels report absolute mean differences. Continuous covariates are standardized;
binary covariates remain on the raw 0/1 scale. Only candidates that pass the
representation gate are plotted.
</div>

```{r love-helpers, echo=FALSE}
love_from_candidates <- function(cell_name, estimator_names) {
  cell <- candidate_cells[[cell_name]]
  Filter(Negate(is.null), lapply(estimator_names, function(estimator) {
    x <- cell$candidates[[estimator]]
    if (is.null(x) || !isTRUE(x$ok)) return(NULL)
    make_love(estimator, x$w, cell$D, cell$X)
  }))
}

selected_love_estimators <- c(
  "PLR-IV, tuned forest", "Wald-AIPW, tuned forest",
  "\u03c4_cb_u (kappa, CBPS)", "\u03c4_ml_u (kappa, MLE)",
  "\u03c4_ml_a10 (kappa, MLE)", "\u03c4_ml_a (kappa, MLE)",
  "\u03c4_ml_t (kappa, MLE)", "\u03c4_ml_a0 (kappa, MLE)"
)

learner_love_estimators <- c(
  "PLR-IV, linear", "Wald-AIPW, linear/logit",
  "PLR-IV, ranger", "Wald-AIPW, ranger",
  "PLR-IV, tuned XGBoost", "Wald-AIPW, tuned XGBoost",
  "\u03c4_cb_u (kappa, CBPS)", "\u03c4_ml_u (kappa, MLE)"
)
```

## DML vs. Kappa {.tabset}

### Work Status

```{r love-labor, echo=FALSE, fig.height=9, fig.width=13}
grid.arrange(
  grobs = love_from_candidates("Labor (workedm)", selected_love_estimators),
  nrow = 2, top = "Love plots — Work Status: eligible DML and kappa weights"
)
```

### Maternal Log Income

```{r love-income, echo=FALSE, fig.height=9, fig.width=13}
grid.arrange(
  grobs = love_from_candidates("Income (lincomem)", selected_love_estimators),
  nrow = 2, top = "Love plots — Maternal Log Income: eligible DML and kappa weights"
)
```

## Learners vs. Kappa {.tabset}

### Work Status

```{r ext-love-labor, echo=FALSE, fig.height=7.5, fig.width=13}
grid.arrange(
  grobs = love_from_candidates("Labor (workedm)", learner_love_estimators),
  nrow = 2, top = "Love plots — Work Status: eligible learners vs. kappa"
)
```

### Maternal Log Income

```{r ext-love-income, echo=FALSE, fig.height=7.5, fig.width=13}
grid.arrange(
  grobs = love_from_candidates("Income (lincomem)", learner_love_estimators),
  nrow = 2, top = "Love plots — Maternal Log Income: eligible learners vs. kappa"
)
```

<div class="note-box">
Only exact outcome-weight representations enter the figures. XGBoost remains
an estimate-level learner comparison when its reconstructed vectors fail the
reproduction check.
</div>
