---
title: "Angrist & Evans (1998) — Maternal Log Income Across Four Units"
subtitle: "Complete-pipeline translation diagnostic with high-precision output"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
table { font-size: 12.5px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  cache = TRUE, autodep = TRUE,
  fig.width = 12, fig.height = 6,
  cache.extra = "child-income-four-units-v1"
)
```

# Purpose

This diagnostic repeats the maternal-log-income analysis under four equivalent
measurement units:
\[
\begin{aligned}
Y_i^{\mathrm{dollars}} &= \log(\mathrm{income}_i),\\
Y_i^{\mathrm{cents}} &= Y_i^{\mathrm{dollars}}+\log(100),\\
Y_i^{\mathrm{thousands}} &= Y_i^{\mathrm{dollars}}-\log(1000),\\
Y_i^{\mathrm{hundred\ thousands}} &=
Y_i^{\mathrm{dollars}}-\log(100000).
\end{aligned}
\]
Every complete-pipeline rerun uses the same sample, six covariates, seed, and
cross-fitting protocol. Nuisance functions and tuning choices are re-estimated
for each outcome coding. The estimation vectors are constructed as exact
additive shifts of the stored dollar-log vector. Separately evaluated expressions
such as `log(100 * income)` are used only to verify the economic unit
interpretation, not as alternative learner inputs.

<div class="note-box">
**Inspection settings.** Results are displayed with eight decimal places and
classified using the strict diagnostic threshold (10^{-8}). To reproduce the
main Method A convention later, change `DISPLAY_DIGITS` to `4` and
`RERUN_TOLERANCE` to `1e-4` in the setup chunk.
</div>

<div class="warn-box">
The XGBoost block is computationally intensive. It imports only the verified
tuned dollar estimates. Cents, thousands, and hundred-thousands are separately
retuned within each outer training fold. The resulting export is stored under
a new filename and does not overwrite the existing tuning or Method A exports.
</div>

# Setup

```{r packages-and-functions, include=FALSE, cache=FALSE}
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyr)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

source(file.path("..", "functions_all.R"))

SEED <- 42L
N_CF <- 5L
DISPLAY_DIGITS <- 8L
RERUN_TOLERANCE <- 1e-8

RUN_KAPPA <- TRUE
RUN_HONEST_FOREST <- TRUE
RUN_LINEAR_RANGER <- TRUE
RUN_XGBOOST <- TRUE
RUN_METHOD_A_AUDIT <- TRUE

RUN_XGB_PLIV <- TRUE
RUN_XGB_WALD_AIPW <- TRUE
TUNE_EVALS <- 15L
INNER_FOLDS <- 3L

fmt_fixed <- function(x, digits = DISPLAY_DIGITS) {
  x <- round(as.numeric(x), digits)
  x[abs(x) < 0.5 * 10^(-digits)] <- 0
  formatC(x, format = "f", digits = digits)
}

add_translation_diagnostics <- function(x) {
  dollar <- x |>
    filter(.data$Unit == "Dollars") |>
    select(Family, Estimator, Dollar_estimate = Estimate)

  x |>
    left_join(dollar, by = c("Family", "Estimator")) |>
    mutate(
      Change_from_dollars = .data$Estimate - .data$Dollar_estimate,
      Invariant = abs(.data$Change_from_dollars) <= RERUN_TOLERANCE
    )
}

wide_translation_table <- function(x, caption) {
  metrics <- x |>
    group_by(.data$Family, .data$Estimator) |>
    summarise(
      `Maximum absolute change` = max(abs(.data$Change_from_dollars)),
      `Invariant` = all(.data$Invariant),
      .groups = "drop"
    )

  out <- x |>
    select(Family, Estimator, Unit, Estimate) |>
    pivot_wider(names_from = .data$Unit, values_from = .data$Estimate) |>
    left_join(metrics, by = c("Family", "Estimator")) |>
    select(
      Estimator, Dollars, Cents, Thousands, `Hundred-thousands`,
      `Maximum absolute change`, Invariant
    )

  display <- out
  numeric_columns <- c(
    "Dollars", "Cents", "Thousands", "Hundred-thousands",
    "Maximum absolute change"
  )
  display[numeric_columns] <- lapply(display[numeric_columns], fmt_fixed)

  display |>
    kbl(
      caption = caption, row.names = FALSE,
      align = c("l", rep("r", 5), "c")
    ) |>
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = TRUE, font_size = 12.5
    ) |>
    row_spec(which(!out$Invariant), background = "#fffbf0") |>
    footnote(general = paste0(
      "All changes are relative to log income measured in dollars. ",
      "Orange rows exceed the absolute diagnostic threshold of ",
      format(RERUN_TOLERANCE, scientific = TRUE), "."
    ))
}
```

```{r data-and-design, include=FALSE, cache=FALSE}
obj <- readRDS("ae98_sub.rds")
if (!"ae98_income_sub" %in% names(obj)) {
  stop("ae98_sub.rds must contain ae98_income_sub.")
}

ae98_income_sub <- obj$ae98_income_sub |>
  mutate(
    morekids = as.integer(.data$morekids),
    samesex = as.integer(.data$samesex)
  )

if (any(!is.finite(ae98_income_sub$incomem)) ||
    any(ae98_income_sub$incomem <= 0)) {
  stop("The income sample must contain strictly positive finite income.")
}

CHILD_COVARIATES <- c(
  "agem", "agefstm", "sexk", "blackm", "hispm", "othracem"
)
RHS_FORMULA <- reformulate(CHILD_COVARIATES)
RHS_DML_FORMULA <- reformulate(CHILD_COVARIATES, intercept = FALSE)

Z_income <- ae98_income_sub$samesex
D_income <- ae98_income_sub$morekids
X_income_kappa <- model.matrix(RHS_FORMULA, data = ae98_income_sub)
X_income_dml <- make_X(RHS_DML_FORMULA, data = ae98_income_sub)

Y_dollars <- log(ae98_income_sub$incomem)
income_units <- list(
  Dollars = list(Y = Y_dollars, shift = 0),
  Cents = list(Y = Y_dollars + log(100), shift = log(100)),
  Thousands = list(Y = Y_dollars - log(1000), shift = -log(1000)),
  `Hundred-thousands` = list(
    Y = Y_dollars - log(100000), shift = -log(100000)
  )
)

direct_unit_codings <- list(
  Dollars = log(ae98_income_sub$incomem),
  Cents = log(100 * ae98_income_sub$incomem),
  Thousands = log(ae98_income_sub$incomem / 1000),
  `Hundred-thousands` = log(ae98_income_sub$incomem / 100000)
)

unit_coding_max_deviation <- vapply(
  names(income_units),
  function(nm) max(abs(income_units[[nm]]$Y - direct_unit_codings[[nm]])),
  numeric(1)
)
stopifnot(all(unit_coding_max_deviation < 1e-10))

unit_design <- data.frame(
  Unit = names(income_units),
  Multiplicative_scale = c("1", "100", "1/1,000", "1/100,000"),
  Additive_log_shift = vapply(income_units, `[[`, numeric(1), "shift"),
  check.names = FALSE
)
```

```{r unit-design-table, echo=FALSE}
unit_design |>
  mutate(`Additive log shift` = fmt_fixed(.data$Additive_log_shift)) |>
  select(Unit, `Income scale` = Multiplicative_scale,
         `Additive log shift`) |>
  kbl(caption = "Four equivalent maternal-income codings", row.names = FALSE) |>
  kable_styling(
    bootstrap_options = c("striped", "condensed"),
    full_width = FALSE
  )
```

# 1. Normalized and unnormalized kappa estimators

```{r kappa-four-units, include=FALSE, cache=TRUE, cache.extra=list("kappa-v1", RERUN_TOLERANCE, CHILD_COVARIATES)}
kappa_estimator_names <- c(
  "tau_cb_u", "tau_ml_u", "tau_ml_a10",
  "tau_ml_a", "tau_ml_t", "tau_ml_a0"
)

if (RUN_KAPPA) {
  kappa_four_units <- bind_rows(lapply(names(income_units), function(unit) {
    estimates <- kappa_point_estimates(
      income_units[[unit]]$Y, Z_income, D_income, X_income_kappa
    )
    data.frame(
      Family = "Kappa",
      Estimator = kappa_estimator_names,
      Unit = unit,
      Shift = income_units[[unit]]$shift,
      Estimate = unname(estimates[kappa_estimator_names]),
      stringsAsFactors = FALSE
    )
  })) |>
    add_translation_diagnostics()
}
```

```{r kappa-four-units-table, echo=FALSE}
if (RUN_KAPPA) {
  wide_translation_table(
    kappa_four_units,
    "Kappa estimators across four maternal-income units"
  )
}
```

# 2. Tuned honest-forest DML

```{r honest-forest-four-units, include=FALSE, cache=TRUE, cache.extra=list("honest-forest-v1", SEED, N_CF, CHILD_COVARIATES)}
if (RUN_HONEST_FOREST) {
  honest_forest_four_units <- bind_rows(lapply(names(income_units), function(unit) {
    set.seed(SEED)
    fit <- dml_with_smoother(
      income_units[[unit]]$Y, D_income, X_income_dml, Z_income,
      n_cf_folds = N_CF, tune.parameters = "all"
    )
    result <- summary(fit)
    estimators <- c("PLR-IV", "Wald-AIPW")
    data.frame(
      Family = "Tuned honest forest",
      Estimator = estimators,
      Unit = unit,
      Shift = income_units[[unit]]$shift,
      Estimate = vapply(estimators, function(est) get_estimate(result, est), numeric(1)),
      stringsAsFactors = FALSE
    )
  })) |>
    add_translation_diagnostics()
}
```

```{r honest-forest-four-units-table, echo=FALSE}
if (RUN_HONEST_FOREST) {
  wide_translation_table(
    honest_forest_four_units,
    "Tuned honest-forest DML across four maternal-income units"
  )
}
```

# 3. Linear/logit and Ranger learners

```{r learner-functions, include=FALSE, cache=FALSE}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y)
  df$d <- as.numeric(D)
  df$z <- as.numeric(Z)
  DoubleMLData$new(
    data = df, y_col = "y", d_cols = "d", z_cols = "z",
    x_cols = colnames(X)
  )
}

ml_lm_iivm <- function() list(
  g = lrn("regr.lm"),
  m = lrn("classif.log_reg"),
  r = lrn("classif.log_reg")
)
ml_rf_iivm <- function() list(
  g = lrn("regr.ranger", keep.inbag = TRUE),
  m = lrn("classif.ranger", keep.inbag = TRUE),
  r = lrn("classif.ranger", keep.inbag = TRUE)
)
ml_lm_pliv <- function() list(
  l = lrn("regr.lm"), m = lrn("regr.lm"), r = lrn("regr.lm")
)
ml_rf_pliv <- function() list(
  l = lrn("regr.ranger", keep.inbag = TRUE),
  m = lrn("regr.ranger", keep.inbag = TRUE),
  r = lrn("regr.ranger", keep.inbag = TRUE)
)

fit_iivm <- function(Y, learners) {
  set.seed(SEED)
  obj <- DoubleMLIIVM$new(
    data = make_dml_data(Y, D_income, Z_income, X_income_dml),
    ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
    n_folds = N_CF, score = "LATE"
  )
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  get_dml_coef(obj)
}

fit_pliv <- function(Y, learners) {
  set.seed(SEED)
  obj <- DoubleMLPLIV$new(
    data = make_dml_data(Y, D_income, Z_income, X_income_dml),
    ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
    n_folds = N_CF
  )
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  get_dml_coef(obj)
}
```

```{r learners-four-units, include=FALSE, cache=TRUE, cache.extra=list("linear-ranger-v1", SEED, N_CF, CHILD_COVARIATES)}
learner_specs <- list(
  list(label = "PLR-IV (linear)", fit = function(Y) fit_pliv(Y, ml_lm_pliv())),
  list(label = "Wald-AIPW (linear+logit)", fit = function(Y) fit_iivm(Y, ml_lm_iivm())),
  list(label = "PLR-IV (ranger)", fit = function(Y) fit_pliv(Y, ml_rf_pliv())),
  list(label = "Wald-AIPW (ranger)", fit = function(Y) fit_iivm(Y, ml_rf_iivm()))
)

if (RUN_LINEAR_RANGER) {
  linear_ranger_four_units <- bind_rows(lapply(names(income_units), function(unit) {
    bind_rows(lapply(learner_specs, function(spec) {
      data.frame(
        Family = "Linear and Ranger",
        Estimator = spec$label,
        Unit = unit,
        Shift = income_units[[unit]]$shift,
        Estimate = spec$fit(income_units[[unit]]$Y),
        stringsAsFactors = FALSE
      )
    }))
  })) |>
    add_translation_diagnostics()
}
```

```{r learners-four-units-table, echo=FALSE}
if (RUN_LINEAR_RANGER) {
  wide_translation_table(
    linear_ranger_four_units,
    "Linear/logit and Ranger DML across four maternal-income units"
  )
}
```

# 4. Tuned XGBoost

```{r xgboost-functions, include=FALSE, cache=FALSE}
xgb_wald_export <- readRDS("ae98_xgb_tuning_export.rds")
xgb_pliv_export <- readRDS("ae98_xgb_pliv_tuning_export.rds")

validate_xgb_export <- function(export, path, expected_class) {
  settings <- export$settings
  required <- c("seed", "n_cf", "tune_evals", "inner_folds", "tune_on_folds")
  if (is.null(settings) || !all(required %in% names(settings))) {
    stop(sprintf("%s is missing required tuning metadata.", path))
  }
  if (!isTRUE(settings$tune_on_folds) ||
      as.integer(settings$seed) != SEED ||
      as.integer(settings$n_cf) != N_CF ||
      as.integer(settings$tune_evals) != TUNE_EVALS ||
      as.integer(settings$inner_folds) != INNER_FOLDS ||
      !identical(settings$dml_class, expected_class) ||
      !identical(settings$covariates, CHILD_COVARIATES)) {
    stop(sprintf("%s does not match the current four-unit protocol.", path))
  }
  info <- export$sample_info
  if (is.null(info) ||
      as.integer(info$income_n) != nrow(ae98_income_sub) ||
      !isTRUE(all.equal(as.numeric(info$income_mean_Z), mean(Z_income), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(info$income_mean_D), mean(D_income), tolerance = 1e-12))) {
    stop(sprintf("%s does not match the current income sample.", path))
  }
  invisible(TRUE)
}

validate_xgb_export(
  xgb_wald_export, "ae98_xgb_tuning_export.rds", "DoubleMLIIVM"
)
validate_xgb_export(
  xgb_pliv_export, "ae98_xgb_pliv_tuning_export.rds", "DoubleMLPLIV"
)

get_tuned_dollar_estimate <- function(export, cell = "Income (lincomem)") {
  value <- export$coefficients[[cell]]$tuned
  if (length(value) != 1L || !is.finite(value)) {
    stop(sprintf("A unique tuned dollar estimate is unavailable for %s.", cell))
  }
  as.numeric(value)
}

make_xgb_learners_iivm <- function() list(
  g = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
  m = lrn("classif.xgboost", predict_type = "prob"),
  r = lrn("classif.xgboost", predict_type = "prob")
)
make_xgb_learners_pliv <- function() list(
  l = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
  m = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
  r = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0)
)
xgb_ps <- function() {
  ps(
    nrounds = p_int(50, 400), max_depth = p_int(2, 5),
    eta = p_dbl(0.02, 0.20), min_child_weight = p_int(1, 8)
  )
}
tune_param_set_iivm <- function() list(
  ml_g = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps()
)
tune_param_set_pliv <- function() list(
  ml_l = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps()
)
tune_settings_iivm <- function() list(
  n_folds_tune = INNER_FOLDS,
  rsmp_tune = rsmp("cv", folds = INNER_FOLDS),
  measure = list(
    ml_g = "regr.rmse", ml_m = "classif.logloss", ml_r = "classif.logloss"
  ),
  terminator = trm("evals", n_evals = TUNE_EVALS),
  algorithm = tnr("random_search")
)
tune_settings_pliv <- function() list(
  n_folds_tune = INNER_FOLDS,
  rsmp_tune = rsmp("cv", folds = INNER_FOLDS),
  measure = list(ml_l = "regr.rmse", ml_m = "regr.rmse", ml_r = "regr.rmse"),
  terminator = trm("evals", n_evals = TUNE_EVALS),
  algorithm = tnr("random_search")
)

fit_xgb_wald <- function(Y) {
  learners <- make_xgb_learners_iivm()
  obj <- DoubleMLIIVM$new(
    data = make_dml_data(Y, D_income, Z_income, X_income_dml),
    ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
    n_folds = N_CF, score = "LATE"
  )
  obj$tune(
    param_set = tune_param_set_iivm(),
    tune_settings = tune_settings_iivm(), tune_on_folds = TRUE
  )
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  get_dml_coef(obj)
}

fit_xgb_pliv <- function(Y) {
  learners <- make_xgb_learners_pliv()
  obj <- DoubleMLPLIV$new(
    data = make_dml_data(Y, D_income, Z_income, X_income_dml),
    ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
    n_folds = N_CF
  )
  obj$tune(
    param_set = tune_param_set_pliv(),
    tune_settings = tune_settings_pliv(), tune_on_folds = TRUE
  )
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  get_dml_coef(obj)
}
```

```{r xgboost-four-units, include=FALSE, cache=TRUE, cache.extra=list("xgboost-four-units-v1", SEED, N_CF, TUNE_EVALS, INNER_FOLDS, CHILD_COVARIATES, RUN_XGB_PLIV, RUN_XGB_WALD_AIPW)}
if (RUN_XGBOOST) {
  xgb_rows <- list()

  if (RUN_XGB_PLIV) {
    pliv_values <- c(
      Dollars = get_tuned_dollar_estimate(xgb_pliv_export),
      vapply(names(income_units)[-1], function(unit) {
        set.seed(SEED)
        fit_xgb_pliv(income_units[[unit]]$Y)
      }, numeric(1))
    )
    names(pliv_values) <- names(income_units)
    xgb_rows[["pliv"]] <- data.frame(
      Family = "Tuned XGBoost",
      Estimator = "PLR-IV (XGBoost tuned)",
      Unit = names(income_units),
      Shift = vapply(income_units, `[[`, numeric(1), "shift"),
      Estimate = unname(pliv_values),
      stringsAsFactors = FALSE
    )
  }

  if (RUN_XGB_WALD_AIPW) {
    wald_values <- c(
      Dollars = get_tuned_dollar_estimate(xgb_wald_export),
      vapply(names(income_units)[-1], function(unit) {
        set.seed(SEED)
        fit_xgb_wald(income_units[[unit]]$Y)
      }, numeric(1))
    )
    names(wald_values) <- names(income_units)
    xgb_rows[["wald"]] <- data.frame(
      Family = "Tuned XGBoost",
      Estimator = "Wald-AIPW (XGBoost tuned)",
      Unit = names(income_units),
      Shift = vapply(income_units, `[[`, numeric(1), "shift"),
      Estimate = unname(wald_values),
      stringsAsFactors = FALSE
    )
  }

  xgboost_four_units <- bind_rows(xgb_rows) |>
    add_translation_diagnostics()
}
```

```{r save-xgboost-four-units, include=FALSE, cache=FALSE}
if (RUN_XGBOOST && exists("xgboost_four_units")) {
  saveRDS(
    list(
      results = xgboost_four_units,
      settings = list(
        seed = SEED, n_cf = N_CF, tune_evals = TUNE_EVALS,
        inner_folds = INNER_FOLDS, tune_on_folds = TRUE,
        display_digits = DISPLAY_DIGITS,
        rerun_tolerance = RERUN_TOLERANCE,
        covariates = CHILD_COVARIATES,
        dollar_source = c(
          "ae98_xgb_pliv_tuning_export.rds",
          "ae98_xgb_tuning_export.rds"
        )
      )
    ),
    "ae98_income_four_unit_xgb_export.rds"
  )
}
```

```{r xgboost-four-units-table, echo=FALSE}
if (RUN_XGBOOST) {
  wide_translation_table(
    xgboost_four_units,
    "Tuned XGBoost across four maternal-income units"
  )
}
```

# 5. Dollar–cent reproduction audit

This audit compares the dollar and cent estimates produced above with the
stored Method A results. It verifies that the expanded four-unit workflow
reproduces the two codings already used in the thesis analysis.

```{r method-a-audit, include=FALSE, cache=FALSE}
available_parts <- Filter(Negate(is.null), list(
  if (exists("kappa_four_units")) kappa_four_units else NULL,
  if (exists("honest_forest_four_units")) honest_forest_four_units else NULL,
  if (exists("linear_ranger_four_units")) linear_ranger_four_units else NULL,
  if (exists("xgboost_four_units")) xgboost_four_units else NULL
))
four_unit_results <- bind_rows(available_parts)

method_a_path <- "ae98_translation_rerun_results.rds"
if (RUN_METHOD_A_AUDIT && file.exists(method_a_path)) {
  method_a <- readRDS(method_a_path)$results
  sample_column <- intersect(c("Specification", "Sample"), names(method_a))[1]
  if (is.na(sample_column)) {
    stop("Method A results contain neither Specification nor Sample.")
  }

  method_a_income <- method_a |>
    filter(.data[[sample_column]] %in% c(
      "Income (lincomem)", "Maternal Log Income"
    ))

  method_a_long <- bind_rows(
    method_a_income |>
      transmute(
        Family, Estimator, Unit = "Dollars",
        Method_A_estimate = .data$tau_original
      ),
    method_a_income |>
      transmute(
        Family, Estimator, Unit = "Cents",
        Method_A_estimate = .data$tau_shifted
      )
  )

  method_a_audit <- four_unit_results |>
    filter(.data$Unit %in% c("Dollars", "Cents")) |>
    select(Family, Estimator, Unit,
           Four_unit_estimate = Estimate) |>
    inner_join(
      method_a_long,
      by = c("Family", "Estimator", "Unit")
    ) |>
    mutate(
      Difference = .data$Four_unit_estimate - .data$Method_A_estimate,
      Reproduced = abs(.data$Difference) <= RERUN_TOLERANCE
    )
}
```

```{r method-a-audit-table, echo=FALSE}
if (RUN_METHOD_A_AUDIT && exists("method_a_audit")) {
  audit_display <- method_a_audit |>
    transmute(
      Family, Estimator, Unit,
      `Four-unit estimate` = fmt_fixed(.data$Four_unit_estimate),
      `Method A estimate` = fmt_fixed(.data$Method_A_estimate),
      Difference = fmt_fixed(.data$Difference),
      Reproduced
    )

  audit_display |>
    kbl(
      caption = "Dollar–cent reproduction audit against Method A",
      row.names = FALSE,
      align = c("l", "l", "l", "r", "r", "r", "c")
    ) |>
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = TRUE, font_size = 12
    ) |>
    collapse_rows(columns = 1, valign = "top") |>
    row_spec(which(!method_a_audit$Reproduced), background = "#fffbf0") |>
    footnote(general = paste0(
      "Reproduction requires an absolute difference no larger than ",
      format(RERUN_TOLERANCE, scientific = TRUE), "."
    ))
}
```

# 6. Save complete diagnostic results

```{r save-results, include=FALSE, cache=FALSE}
saveRDS(
  list(
    results = four_unit_results,
    method_a_audit = if (exists("method_a_audit")) method_a_audit else NULL,
    unit_design = unit_design,
    settings = list(
      seed = SEED, n_cf = N_CF,
      display_digits = DISPLAY_DIGITS,
      rerun_tolerance = RERUN_TOLERANCE,
      grf_tuning = "all",
      xgboost_tune_on_folds = TRUE,
      xgboost_tune_evals = TUNE_EVALS,
      xgboost_inner_folds = INNER_FOLDS,
      covariates = CHILD_COVARIATES
    )
  ),
  "ae98_income_four_unit_translation_results.rds"
)
```

<div class="note-box">
The complete results are saved to
`ae98_income_four_unit_translation_results.rds`. The separately fitted
XGBoost results are also stored in `ae98_income_four_unit_xgb_export.rds`.
</div>
