---
title: "Angrist & Evans (1998) — Translation Invariance: Complete-Pipeline Rerun"
subtitle: "Kappa, honest forest, DoubleML learners, and nested XGBoost"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
h3   { color: #444; margin-top: 20px; }
table { font-size: 13px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 12, fig.height = 6, autodep = TRUE,
  cache.extra = "translation-rerun-only-v5-six-covariates"
)
```

---

# Purpose

This notebook evaluates translation invariance by fitting each complete
estimation pipeline to the original outcome and to its additively shifted
counterpart. The two CHILD workflows are
\[
Y^{\mathrm{shift}}_{i,\mathrm{work}}
=Y_{i,\mathrm{work}}+1
\]
and
\[
Y^{\mathrm{shift}}_{i,\mathrm{income}}
=Y_{i,\mathrm{income}}+\log(100)
=\log(100\,\mathrm{income}_i).
\]
Original and shifted fits use the same samples, covariates, seed, and
cross-fitting protocol, while nuisance functions and tuning choices are
re-estimated.

<div class="note-box">
The reflected labor outcome, `1 - workedm`, is excluded because it combines a
reflection with a translation and is therefore not a pure additive shift.
</div>

An estimate is classified as invariant when its absolute rerun change does not
exceed \(10^{-4}\). This operational threshold is separate from the
`all.equal()` check used for fitted outcome-weight representations.

<div class="note-box">
**Common design.** As in `child_test.Rmd`, every kappa and DML rerun uses
`agem`, `agefstm`, `sexk`, `blackm`, `hispm`, and `othracem`. The
second-child sex indicator is omitted because its interaction with first-child
sex exactly determines `samesex`. Using the same six covariates in every
pipeline isolates translation behavior from learner-specific differences in
the conditioning set.
</div>

# Setup

```{r packages-and-source, include=FALSE, cache=FALSE}
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(cobalt)
library(viridis)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

source(file.path("..", "functions_all.R"))

RERUN_TOLERANCE <- 1e-4

ti_child_row <- function(name, tau_original, tau_shifted, k, digits = 12) {
  out <- translation_rerun_row(
    name, tau_original, tau_shifted, k,
    tolerance = RERUN_TOLERANCE, digits = digits
  )
  out$invariant <- abs(out$rerun_difference) <= RERUN_TOLERANCE
  out
}

display_rerun_results <- function(x) {
  fixed_four <- function(value) {
    value <- round(as.numeric(value), 4)
    value[abs(value) < 0.00005] <- 0
    formatC(value, format = "f", digits = 4)
  }
  data.frame(
    Specification = recode(
      x$Specification,
      "Labor (workedm)" = "Work Status",
      "Income (lincomem)" = "Maternal Log Income"
    ),
    Estimator = x$Estimator,
    Shift = fixed_four(x$Shift_k),
    `Original estimate` = fixed_four(x$tau_original),
    `Shifted estimate` = fixed_four(x$tau_shifted),
    `Rerun change` = fixed_four(x$rerun_difference),
    Invariant = x$invariant,
    check.names = FALSE
  )
}
```

<div class="warn-box">
The default `RUN_XGB_RERUN = FALSE` loads the verified complete-rerun XGBoost
export. Set it to `TRUE` only to import the original tuned estimates and repeat
nested tuning for the shifted outcomes.
</div>

```{r switches, include=FALSE}
RUN_KAPPA             <- TRUE
RUN_DML               <- TRUE
RUN_LM_RF             <- TRUE
RUN_XGB_RERUN         <- FALSE
RUN_XGB_LABOR_ONLY    <- FALSE
RUN_XGB_WALD_AIPW     <- TRUE
RUN_XGB_PLIV          <- TRUE
USE_XGB_ORIGINAL_EXPORT <- TRUE
```

```{r data, include=FALSE}
obj <- readRDS("ae98_sub.rds")
if (!all(c("ae98_labor_sub", "ae98_income_sub") %in% names(obj))) {
  stop("ae98_sub.rds must contain ae98_labor_sub and ae98_income_sub.")
}
ae98_labor_sub  <- obj$ae98_labor_sub
ae98_income_sub <- obj$ae98_income_sub

ae98_labor_sub <- ae98_labor_sub %>%
  mutate(workedm = as.integer(workedm), morekids = as.integer(morekids),
         samesex = as.integer(samesex))
ae98_income_sub <- ae98_income_sub %>%
  mutate(morekids = as.integer(morekids), samesex = as.integer(samesex),
         lincomem = log(incomem))

Z_labor <- ae98_labor_sub$samesex;  D_labor  <- ae98_labor_sub$morekids
Z_income <- ae98_income_sub$samesex; D_income <- ae98_income_sub$morekids

k_labor  <- 1
k_income <- log(100)

Y_labor_orig   <- ae98_labor_sub$workedm
Y_labor_shift  <- ae98_labor_sub$workedm + k_labor

Y_income_orig  <- ae98_income_sub$lincomem
Y_income_shift <- ae98_income_sub$lincomem + k_income

max_dev <- max(abs(Y_income_shift - log(100 * ae98_income_sub$incomem)))
stopifnot(max_dev < 1e-9)
```

```{r design-matrices, include=FALSE}
CHILD_COVARIATES <- c(
  "agem", "agefstm", "sexk", "blackm", "hispm", "othracem"
)
RHS_FORMULA     <- reformulate(CHILD_COVARIATES)
RHS_DML_FORMULA <- reformulate(CHILD_COVARIATES, intercept = FALSE)

X_labor_kappa  <- model.matrix(RHS_FORMULA, data = ae98_labor_sub)
X_income_kappa <- model.matrix(RHS_FORMULA, data = ae98_income_sub)

X_dml_labor  <- make_X(RHS_DML_FORMULA, data = ae98_labor_sub)
X_dml_income <- make_X(RHS_DML_FORMULA, data = ae98_income_sub)

workflows <- list(
  Labor = list(label = "Labor (workedm)", D = D_labor, Z = Z_labor,
               X_kappa = X_labor_kappa, X_dml = X_dml_labor,
               Y_orig = Y_labor_orig, Y_shift = Y_labor_shift, k = k_labor),
  Income = list(label = "Income (lincomem)", D = D_income, Z = Z_income,
               X_kappa = X_income_kappa, X_dml = X_dml_income,
               Y_orig = Y_income_orig, Y_shift = Y_income_shift, k = k_income)
)
```

---

# 1. Kappa estimators

Normalized and unnormalized kappa estimators provide the benchmark for the
complete-pipeline learner reruns.

```{r kappa-estimator-names, include=FALSE}
kappa_estimator_names <- c(
  "tau_cb_u", "tau_ml_u", "tau_ml_a10",
  "tau_ml_a", "tau_ml_t", "tau_ml_a0"
)

build_kappa_ti_rows <- function(wf) {
  est_o <- kappa_point_estimates(wf$Y_orig,  wf$Z, wf$D, wf$X_kappa)
  est_s <- kappa_point_estimates(wf$Y_shift, wf$Z, wf$D, wf$X_kappa)
  rows <- lapply(kappa_estimator_names, function(est_name) {
    row <- ti_child_row(name = est_name, tau_original = est_o[[est_name]],
                        tau_shifted = est_s[[est_name]], k = wf$k)
    row$Specification <- wf$label
    row
  })
  do.call(rbind, rows)
}
```

```{r kappa-ti-run, cache=TRUE, include=FALSE}
if (RUN_KAPPA) {
  kappa_ti_all <- do.call(rbind, lapply(workflows, build_kappa_ti_rows))
}
```

```{r kappa-ti-table}
if (RUN_KAPPA) {
  display_rerun_results(kappa_ti_all) %>%
    kbl(caption = "Kappa estimators — original versus shifted outcome",
        digits = 4, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!kappa_ti_all$invariant), background = "#fffbf0") %>%
    footnote(general = paste0(
      "Orange rows exceed the pre-specified absolute rerun threshold of ",
      format(RERUN_TOLERANCE, scientific = TRUE), "."
    ))
}
```

---

# 2. Headline DML: tuned honest forest

<div class="note-box">
The headline rerun uses tuned honest forests. Original and shifted fits use
the same seed, five cross-fitting folds, and tuning setting.
</div>

```{r dml-headline-builder, include=FALSE}
build_dml_headline_ti_rows <- function(wf, n_cf_folds = 5) {
  set.seed(42)
  fit_o <- dml_with_smoother(wf$Y_orig, wf$D, wf$X_dml, wf$Z,
                             n_cf_folds = n_cf_folds, tune.parameters = "all")
  set.seed(42)
  fit_s <- dml_with_smoother(wf$Y_shift, wf$D, wf$X_dml, wf$Z,
                             n_cf_folds = n_cf_folds, tune.parameters = "all")

  res_o <- summary(fit_o); res_s <- summary(fit_s)
  rows <- lapply(c("PLR-IV", "Wald-AIPW"), function(est) {
    row <- ti_child_row(name = est, tau_original = get_estimate(res_o, est),
                        tau_shifted = get_estimate(res_s, est), k = wf$k)
    row$Specification <- wf$label
    row
  })
  do.call(rbind, rows)
}
```

```{r dml-headline-run, cache=TRUE, include=FALSE}
if (RUN_DML) {
  dml_headline_ti_all <- do.call(rbind, lapply(workflows, build_dml_headline_ti_rows))
}
```

```{r dml-headline-table}
if (RUN_DML) {
  display_rerun_results(dml_headline_ti_all) %>%
    kbl(caption = "Tuned honest forest — original versus shifted outcome",
        digits = 4, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!dml_headline_ti_all$invariant), background = "#fffbf0") %>%
    footnote(general = "Orange rows exceed the pre-specified absolute rerun threshold of 10^-4.")
}
```

---

# 3. OLS/logit and Ranger learners

The learner comparison repeats the specifications in `child_test.Rmd`.
All learners use the same six covariates. XGBoost is handled separately
because the shifted outcome requires a new nested-tuning run.

```{r dml-learner-setup, include=FALSE}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z", x_cols = colnames(X))
}

ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))

ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))

fit_iivm <- function(dml_data, learners) {
  set.seed(42)
  o <- DoubleMLIIVM$new(data = dml_data, ml_g = learners$g, ml_m = learners$m,
                        ml_r = learners$r, n_folds = 5, score = "LATE")
  o$fit(store_models = TRUE, store_predictions = TRUE)
  o
}

fit_pliv <- function(dml_data, learners) {
  set.seed(42)
  o <- DoubleMLPLIV$new(data = dml_data, ml_l = learners$l, ml_m = learners$m,
                        ml_r = learners$r, n_folds = 5)
  o$fit(store_models = TRUE, store_predictions = TRUE)
  o
}

build_dml_learner_ti_row <- function(wf, label, learners_fn, X_dml, fit_fn) {
  dat_o <- make_dml_data(wf$Y_orig,  wf$D, wf$Z, X_dml)
  dat_s <- make_dml_data(wf$Y_shift, wf$D, wf$Z, X_dml)

  fit_o <- fit_fn(dat_o, learners_fn())
  fit_s <- fit_fn(dat_s, learners_fn())

  row <- ti_child_row(name = label, tau_original = get_dml_coef(fit_o),
                      tau_shifted = get_dml_coef(fit_s), k = wf$k)
  row$Specification <- wf$label
  row
}
```

```{r dml-learner-run, cache=TRUE, include=FALSE}
if (RUN_LM_RF) {
  dml_learner_ti_all <- do.call(rbind, lapply(workflows, function(wf) {
    rbind(
      build_dml_learner_ti_row(wf, "PLR-IV (linear)",          ml_lm_pliv,  wf$X_dml, fit_pliv),
      build_dml_learner_ti_row(wf, "Wald-AIPW (linear+logit)", ml_lm_iivm,  wf$X_dml, fit_iivm),
      build_dml_learner_ti_row(wf, "PLR-IV (ranger)",          ml_rf_pliv,  wf$X_dml,    fit_pliv),
      build_dml_learner_ti_row(wf, "Wald-AIPW (ranger)",       ml_rf_iivm,  wf$X_dml,    fit_iivm)
    )
  }))
}
```

```{r dml-learner-table}
if (RUN_LM_RF) {
  display_rerun_results(dml_learner_ti_all) %>%
    kbl(caption = "DoubleML learners — original versus shifted outcome",
        digits = 4, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!dml_learner_ti_all$invariant), background = "#fffbf0") %>%
    footnote(general = paste(
      "Every learner uses the same six covariates.",
      "Orange rows exceed the pre-specified absolute rerun threshold of 10^-4."
    ))
}
```

# 4. Tuned XGBoost complete-pipeline rerun

<div class="warn-box">
The shifted-outcome XGBoost fits repeat the complete nested-tuning protocol
with the same seed, samples, folds, and tuning budget.
</div>

```{r xgb-setup, include=FALSE}
SEED <- 42; N_CF <- 5; TUNE_EVALS <- 15; INNER_FOLDS <- 3
XGB_WALD_EXPORT <- "ae98_xgb_tuning_export.rds"
XGB_PLIV_EXPORT <- "ae98_xgb_pliv_tuning_export.rds"
XGB_RERUN_EXPORT <- "ae98_translation_xgb_rerun_export.rds"

if (RUN_XGB_RERUN && !USE_XGB_ORIGINAL_EXPORT) {
  stop("This notebook expects USE_XGB_ORIGINAL_EXPORT <- TRUE for the original XGBoost side.")
}
if (RUN_XGB_RERUN && RUN_XGB_LABOR_ONLY) {
  stop("An official XGBoost translation rebuild must include both Child outcomes; set RUN_XGB_LABOR_ONLY=FALSE.")
}
if (RUN_XGB_RERUN && (!RUN_XGB_WALD_AIPW || !RUN_XGB_PLIV)) {
  stop("An official Child XGBoost translation rebuild must include both Wald-AIPW and PLR-IV.")
}

xgb_wald_export <- if (RUN_XGB_RERUN && file.exists(XGB_WALD_EXPORT)) readRDS(XGB_WALD_EXPORT) else NULL
xgb_pliv_export <- if (RUN_XGB_RERUN && file.exists(XGB_PLIV_EXPORT)) readRDS(XGB_PLIV_EXPORT) else NULL

validate_xgb_export <- function(export, path, required_cells, required_covariates) {
  settings <- export$settings
  if (is.null(settings) || !isTRUE(settings$tune_on_folds)) {
    stop(sprintf("%s is not a tune_on_folds=TRUE export.", path))
  }
  expected <- c(seed = SEED, n_cf = N_CF, tune_evals = TUNE_EVALS,
                inner_folds = INNER_FOLDS)
  if (!all(names(expected) %in% names(settings))) {
    stop(sprintf("%s is missing required tuning metadata.", path))
  }
  observed <- vapply(names(expected), function(x) as.integer(settings[[x]]), integer(1))
  if (!identical(unname(observed), unname(as.integer(expected)))) {
    stop(sprintf("%s uses tuning settings that differ from this rerun.", path))
  }
  if (is.null(export$weight_diagnostics) ||
      !all(required_cells %in% export$weight_diagnostics$Cell)) {
    stop(sprintf("%s does not contain every required child sample.", path))
  }
  if (is.null(settings$covariates) ||
      !identical(settings$covariates, required_covariates)) {
    stop(sprintf("%s does not use the common six-covariate Child design.", path))
  }
  info <- export$sample_info
  if (is.null(info) ||
      !identical(as.integer(info$labor_n), as.integer(nrow(ae98_labor_sub))) ||
      !identical(as.integer(info$income_n), as.integer(nrow(ae98_income_sub))) ||
      !isTRUE(all.equal(as.numeric(info$labor_mean_Z), mean(Z_labor), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(info$labor_mean_D), mean(D_labor), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(info$income_mean_Z), mean(Z_income), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(info$income_mean_D), mean(D_income), tolerance = 1e-12))) {
    stop(sprintf("%s metadata do not match the current Child samples.", path))
  }
  invisible(TRUE)
}

required_xgb_cells <- vapply(workflows, function(wf) wf$label, character(1))
if (RUN_XGB_RERUN) {
  if (!file.exists(XGB_WALD_EXPORT) || !file.exists(XGB_PLIV_EXPORT)) {
    stop("Run or restore both Child XGBoost tuning exports before rebuilding the translation rerun.")
  }
  validate_xgb_export(
    xgb_wald_export, XGB_WALD_EXPORT,
    required_xgb_cells, CHILD_COVARIATES
  )
  validate_xgb_export(
    xgb_pliv_export, XGB_PLIV_EXPORT,
    required_xgb_cells, CHILD_COVARIATES
  )
}

validate_xgb_rerun_export <- function(export, path, required_specs) {
  if (!is.list(export) || is.null(export$results) || is.null(export$settings)) {
    stop(sprintf("%s has an incomplete schema.", path))
  }
  required_columns <- c("Estimator", "Shift_k", "tau_original", "tau_shifted",
                        "rerun_difference", "invariant", "Specification")
  if (!all(required_columns %in% names(export$results))) {
    stop(sprintf("%s is missing required result columns.", path))
  }
  expected_settings <- c(seed = SEED, n_cf = N_CF, tune_evals = TUNE_EVALS,
                         inner_folds = INNER_FOLDS)
  if (!all(names(expected_settings) %in% names(export$settings)) ||
      !isTRUE(export$settings$tune_on_folds)) {
    stop(sprintf("%s is missing required tuning metadata.", path))
  }
  observed_settings <- vapply(
    names(expected_settings), function(x) as.integer(export$settings[[x]]), integer(1)
  )
  if (!identical(unname(observed_settings), unname(as.integer(expected_settings)))) {
    stop(sprintf("%s uses tuning settings that differ from this notebook.", path))
  }
  if (is.null(export$settings$covariates) ||
      !identical(export$settings$covariates, CHILD_COVARIATES)) {
    stop(sprintf("%s does not use the common six-covariate Child design.", path))
  }
  required_estimators <- c("Wald-AIPW (XGBoost tuned)",
                           "PLR-IV (XGBoost tuned)")
  expected_keys <- as.vector(outer(required_specs, required_estimators, paste, sep = "\r"))
  observed_keys <- paste(export$results$Specification, export$results$Estimator, sep = "\r")
  if (anyDuplicated(observed_keys) || !setequal(observed_keys, expected_keys)) {
    stop(sprintf("%s does not contain exactly the required Child XGBoost rows.", path))
  }
  numeric_columns <- c("Shift_k", "tau_original", "tau_shifted", "rerun_difference")
  if (any(!vapply(export$results[numeric_columns],
                  function(x) all(is.finite(as.numeric(x))), logical(1)))) {
    stop(sprintf("%s contains non-finite translation results.", path))
  }
  export$results
}

make_xgb_learners_iivm <- function() {
  list(g = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
       m = lrn("classif.xgboost", predict_type = "prob"),
       r = lrn("classif.xgboost", predict_type = "prob"))
}
make_xgb_learners_pliv <- function() {
  list(l = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
       m = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
       r = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0))
}
xgb_ps <- function() {
  ps(nrounds = p_int(50, 400), max_depth = p_int(2, 5),
     eta = p_dbl(0.02, 0.20), min_child_weight = p_int(1, 8))
}
tune_param_set_iivm <- function() list(ml_g = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())
tune_param_set_pliv <- function() list(ml_l = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())
tune_settings_iivm <- function() {
  list(n_folds_tune = INNER_FOLDS, rsmp_tune = rsmp("cv", folds = INNER_FOLDS),
       measure = list(ml_g = "regr.rmse", ml_m = "classif.logloss", ml_r = "classif.logloss"),
       terminator = trm("evals", n_evals = TUNE_EVALS), algorithm = tnr("random_search"))
}
tune_settings_pliv <- function() {
  list(n_folds_tune = INNER_FOLDS, rsmp_tune = rsmp("cv", folds = INNER_FOLDS),
       measure = list(ml_l = "regr.rmse", ml_m = "regr.rmse", ml_r = "regr.rmse"),
       terminator = trm("evals", n_evals = TUNE_EVALS), algorithm = tnr("random_search"))
}

get_xgb_original <- function(export, cell, learner) {
  row <- export$weight_diagnostics |>
    filter(.data$Cell == cell, .data$Learner == learner)
  if (nrow(row) != 1) {
    stop(sprintf("Expected one export row for %s / %s, found %d.", cell, learner, nrow(row)))
  }
  as.numeric(row$Estimate[[1]])
}

fit_xgb_wald_shifted <- function(Y, D, Z, X) {
  dml_data <- make_dml_data(Y, D, Z, X)
  learners <- make_xgb_learners_iivm()
  o <- DoubleMLIIVM$new(data = dml_data, ml_g = learners$g, ml_m = learners$m,
                        ml_r = learners$r, n_folds = N_CF, score = "LATE")
  o$tune(param_set = tune_param_set_iivm(), tune_settings = tune_settings_iivm(),
         tune_on_folds = TRUE)
  o$fit(store_models = TRUE, store_predictions = TRUE)
  list(obj = o, tau = get_dml_coef(o))
}

fit_xgb_pliv_shifted <- function(Y, D, Z, X) {
  dml_data <- make_dml_data(Y, D, Z, X)
  learners <- make_xgb_learners_pliv()
  o <- DoubleMLPLIV$new(data = dml_data, ml_l = learners$l, ml_m = learners$m,
                        ml_r = learners$r, n_folds = N_CF)
  o$tune(param_set = tune_param_set_pliv(), tune_settings = tune_settings_pliv(),
         tune_on_folds = TRUE)
  o$fit(store_models = TRUE, store_predictions = TRUE)
  list(obj = o, tau = get_dml_coef(o))
}

build_xgb_rerun_rows <- function(wf) {
  rows <- list()

  if (RUN_XGB_WALD_AIPW) {
    original <- get_xgb_original(xgb_wald_export, wf$label, "XGB tuned")
    set.seed(SEED)
    shifted <- fit_xgb_wald_shifted(wf$Y_shift, wf$D, wf$Z, wf$X_dml)
    row <- ti_child_row(name = "Wald-AIPW (XGBoost tuned)",
                        tau_original = original,
                        tau_shifted = shifted$tau,
                        k = wf$k)
    row$Specification <- wf$label
    rows[["wald"]] <- row
  }

  if (RUN_XGB_PLIV) {
    original <- get_xgb_original(xgb_pliv_export, wf$label, "PLR-IV XGB tuned")
    set.seed(SEED)
    shifted <- fit_xgb_pliv_shifted(wf$Y_shift, wf$D, wf$Z, wf$X_dml)
    row <- ti_child_row(name = "PLR-IV (XGBoost tuned)",
                        tau_original = original,
                        tau_shifted = shifted$tau,
                        k = wf$k)
    row$Specification <- wf$label
    rows[["pliv"]] <- row
  }

  do.call(rbind, rows)
}

save_xgb_rerun_export <- function(xgb_ti_all) {
  export <- list(
    results = xgb_ti_all,
    settings = list(seed = SEED, n_cf = N_CF,
                    rerun_tolerance = RERUN_TOLERANCE,
                    tune_evals = TUNE_EVALS,
                    inner_folds = INNER_FOLDS, tune_on_folds = TRUE,
                    covariates = CHILD_COVARIATES,
                    original_exports = c(XGB_WALD_EXPORT, XGB_PLIV_EXPORT))
  )
  validate_xgb_rerun_export(export, XGB_RERUN_EXPORT, required_xgb_cells)
  saveRDS(export, XGB_RERUN_EXPORT)
}
```

```{r xgb-run, include=FALSE, cache=FALSE}
if (RUN_XGB_RERUN) {
  xgb_ti_all <- do.call(rbind, lapply(workflows, build_xgb_rerun_rows))
  save_xgb_rerun_export(xgb_ti_all)
} else {
  if (!file.exists(XGB_RERUN_EXPORT)) {
    stop("Missing verified Child XGBoost translation export. Restore it or set RUN_XGB_RERUN=TRUE to rebuild it.")
  }
  xgb_ti_all <- validate_xgb_rerun_export(
    readRDS(XGB_RERUN_EXPORT), XGB_RERUN_EXPORT, required_xgb_cells
  )
}
xgb_ti_all$invariant <- abs(xgb_ti_all$rerun_difference) <= RERUN_TOLERANCE
```

```{r xgb-table}
if (exists("xgb_ti_all")) {
  display_rerun_results(xgb_ti_all) %>%
    kbl(caption = "Tuned XGBoost — original export versus shifted nested-tuning rerun",
        digits = 4, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!xgb_ti_all$invariant), background = "#fffbf0") %>%
    footnote(general = paste(
      if (RUN_XGB_RERUN)
        "This render rebuilt the shifted-outcome nested-tuning results." else
        "This render loaded the verified complete-rerun XGBoost export.",
      "Orange rows exceed the pre-specified absolute rerun threshold of 10^-4."
    ))
}
```

<div class="note-box">
The XGBoost rerun is stored separately and does not overwrite the original
tuning exports.
</div>

# 5. Save reusable rerun results

```{r save-complete-rerun, cache=FALSE, include=FALSE}
result_parts <- Filter(Negate(is.null), list(
  Kappa = if (exists("kappa_ti_all")) kappa_ti_all else NULL,
  `Tuned honest forest` = if (exists("dml_headline_ti_all")) dml_headline_ti_all else NULL,
  `Linear and Ranger` = if (exists("dml_learner_ti_all")) dml_learner_ti_all else NULL,
  `Tuned XGBoost` = if (exists("xgb_ti_all")) xgb_ti_all else NULL
))
for (nm in names(result_parts)) result_parts[[nm]]$Family <- nm
translation_results <- bind_rows(result_parts)
if (RUN_XGB_RERUN) {
  expected_families <- c("Kappa", "Tuned honest forest", "Linear and Ranger", "Tuned XGBoost")
  if (!setequal(unique(translation_results$Family), expected_families)) {
    stop("Refusing to replace the complete Child translation export because a result family is missing.")
  }
  saveRDS(list(
    results = translation_results,
    settings = list(seed=42,n_cf=5,shifts=c(Labor=1,Income=log(100)),
                    rerun_tolerance=RERUN_TOLERANCE,
                    covariates=CHILD_COVARIATES,
                    grf_tuning="all",xgboost_tune_on_folds=TRUE,
                    xgboost_tune_evals=TUNE_EVALS,xgboost_inner_folds=INNER_FOLDS)
  ), "ae98_translation_rerun_results.rds")
}
```
