---
title: "Child Untuned Honest-Forest Translation Audit"
subtitle: "Internal paired check with a strict 10^-8 tolerance"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    code_folding: show
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1 { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; }
h2 { color: #2c7bb6; }
.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
}
.warn-box {
  background: #fff3e0;
  border-left: 4px solid #fb8c00;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE
)
```

# Purpose

This is a standalone internal audit of complete-pipeline translation invariance
for the **untuned honest-forest** implementations of PLR-IV and Wald-AIPW in the
Child application. It is intended to be run from the `child` R project by
clicking **Run All** or **Knit**.

The audit is a paired computational experiment, not a change designed to obtain
a particular result. Within each pair, the sample, (D), (Z), (X),
observation ordering, learner defaults, five-fold cross-fitting rule, and
initial random-number state are held fixed. Only the outcome changes from
(mathbf Y) to (mathbf Y+kmathbf 1). The forest structures and predictions
are allowed to change; detecting whether they change is the purpose of the
exercise.

<div class="note-box">
The original-outcome calls deliberately reproduce the untuned specification in
<code>child_test.Rmd</code>: seed 42, five cross-fitting folds, and
<code>tune.parameters = "none"</code>. No active result export is read,
overwritten, or replaced.
</div>

<div class="warn-box">
This notebook performs six honest-forest fits. It keeps at most two large fitted
objects in memory simultaneously and does not use a knitr cache.
</div>

# Setup and fixed Child samples

```{r packages-and-settings}
library(OutcomeWeights)
library(knitr)
library(kableExtra)

source(file.path("..", "functions_all.R"))

SEED <- 42L
N_CF <- 5L
RERUN_TOLERANCE <- 1e-8
SMOOTHER_TOLERANCE <- 1e-12
SMOOTHER_BLOCK_ROWS <- 100L
DISPLAY_DIGITS <- 12L
SAVE_RESULTS <- FALSE
```

```{r data-and-design}
if (!file.exists("ae98_sub.rds")) {
  stop("Missing ae98_sub.rds. Rebuild the fixed Child subsamples first.")
}

obj <- readRDS("ae98_sub.rds")
required_samples <- c("ae98_labor_sub", "ae98_income_sub")
if (!all(required_samples %in% names(obj))) {
  stop("ae98_sub.rds must contain ae98_labor_sub and ae98_income_sub.")
}

ae98_labor_sub <- obj$ae98_labor_sub
ae98_income_sub <- obj$ae98_income_sub

ae98_labor_sub$workedm <- as.integer(ae98_labor_sub$workedm)
ae98_labor_sub$morekids <- as.integer(ae98_labor_sub$morekids)
ae98_labor_sub$samesex <- as.integer(ae98_labor_sub$samesex)

ae98_income_sub$morekids <- as.integer(ae98_income_sub$morekids)
ae98_income_sub$samesex <- as.integer(ae98_income_sub$samesex)

if (any(!is.finite(ae98_income_sub$incomem)) ||
    any(ae98_income_sub$incomem <= 0)) {
  stop("The income sample must contain strictly positive finite income.")
}

CHILD_COVARIATES <- c(
  "agem", "agefstm", "sexk", "blackm", "hispm", "othracem"
)
RHS_DML_FORMULA <- reformulate(CHILD_COVARIATES, intercept = FALSE)

X_labor <- make_X(RHS_DML_FORMULA, data = ae98_labor_sub)
X_income <- make_X(RHS_DML_FORMULA, data = ae98_income_sub)

D_labor <- ae98_labor_sub$morekids
Z_labor <- ae98_labor_sub$samesex
D_income <- ae98_income_sub$morekids
Z_income <- ae98_income_sub$samesex

Y_labor <- ae98_labor_sub$workedm
Y_income_dollars <- log(ae98_income_sub$incomem)

income_units <- list(
  Cents = list(Y = Y_income_dollars + log(100), k = log(100)),
  Thousands = list(Y = Y_income_dollars - log(1000), k = -log(1000)),
  `Hundred-thousands` = list(
    Y = Y_income_dollars - log(100000),
    k = -log(100000)
  )
)

# These separately evaluated expressions verify the economic interpretation.
# They are not supplied to the forest learners.
direct_income_units <- list(
  Cents = log(100 * ae98_income_sub$incomem),
  Thousands = log(ae98_income_sub$incomem / 1000),
  `Hundred-thousands` = log(ae98_income_sub$incomem / 100000)
)
unit_coding_max_deviation <- vapply(
  names(income_units),
  function(unit) {
    max(abs(income_units[[unit]]$Y - direct_income_units[[unit]]))
  },
  numeric(1)
)
stopifnot(all(unit_coding_max_deviation < 1e-10))
```

# Audit functions

```{r audit-functions}
fit_untuned <- function(Y, D, X, Z) {
  # This call matches the untuned honest-forest calls in child_test.Rmd.
  set.seed(SEED)
  dml_with_smoother(
    Y, D, X, Z,
    n_cf_folds = N_CF,
    tune.parameters = "none"
  )
}

max_abs <- function(x) {
  x <- as.numeric(x)
  if (!length(x)) return(NA_real_)
  max(abs(x))
}

rmse <- function(x) {
  x <- as.numeric(x)
  if (!length(x)) return(NA_real_)
  sqrt(mean(x^2))
}

prediction_diagnostic <- function(pred_original, pred_shifted, expected_shift,
                                  outcome, unit, nuisance) {
  error <- as.numeric(pred_shifted) -
    as.numeric(pred_original) - expected_shift
  data.frame(
    Outcome = outcome,
    Unit = unit,
    Nuisance = nuisance,
    Expected_shift = expected_shift,
    Maximum_absolute_error = max_abs(error),
    RMSE = rmse(error),
    Median_absolute_error = median(abs(error)),
    stringsAsFactors = FALSE,
    check.names = FALSE
  )
}

smoother_diagnostic <- function(
    smoother_original, smoother_shifted,
    Y_original, prediction_original, prediction_shifted, k,
    outcome, unit, estimator, smoother_name,
    tolerance = SMOOTHER_TOLERANCE,
    block_rows = SMOOTHER_BLOCK_ROWS) {

  if (!identical(dim(smoother_original), dim(smoother_shifted))) {
    stop(sprintf(
      "Smoother dimensions differ for %s / %s / %s.",
      outcome, unit, smoother_name
    ))
  }

  smoother_dim <- dim(smoother_original)
  if (length(smoother_dim) != 3L || smoother_dim[1L] != 1L ||
      smoother_dim[2L] != smoother_dim[3L]) {
    stop(sprintf(
      paste0(
        "Expected a 1 x N x N smoother array for %s / %s / %s; ",
        "found dimensions %s."
      ),
      outcome, unit, smoother_name,
      paste(smoother_dim, collapse = " x ")
    ))
  }

  N <- smoother_dim[2L]
  if (length(Y_original) != N ||
      length(prediction_original) != N ||
      length(prediction_shifted) != N) {
    stop(sprintf(
      paste0(
        "Outcome, prediction, and smoother dimensions do not match for ",
        "%s / %s / %s."
      ),
      outcome, unit, smoother_name
    ))
  }

  # Store only N-length summaries. Difference matrices are formed one row block
  # at a time and discarded immediately.
  row_l1_change <- numeric(N)
  row_max_change <- numeric(N)
  smoother_component <- numeric(N)
  normalization_component <- numeric(N)
  reconstructed_original <- numeric(N)
  reconstructed_shifted <- numeric(N)

  maximum_element_change <- 0
  squared_element_change <- 0
  changed_elements <- 0
  maximum_row_sum_deviation_original <- 0
  maximum_row_sum_deviation_shifted <- 0

  block_starts <- seq.int(1L, N, by = block_rows)
  for (block_start in block_starts) {
    rows <- block_start:min(block_start + block_rows - 1L, N)
    n_block <- length(rows)

    S_original_block <- matrix(
      smoother_original[1L, rows, , drop = FALSE],
      nrow = n_block,
      ncol = N
    )
    S_shifted_block <- matrix(
      smoother_shifted[1L, rows, , drop = FALSE],
      nrow = n_block,
      ncol = N
    )
    difference_block <- S_shifted_block - S_original_block
    absolute_difference_block <- abs(difference_block)

    row_l1_change[rows] <- rowSums(absolute_difference_block)
    row_max_change[rows] <- apply(absolute_difference_block, 1L, max)
    maximum_element_change <- max(
      maximum_element_change,
      absolute_difference_block
    )
    squared_element_change <- squared_element_change +
      sum(difference_block^2)
    changed_elements <- changed_elements +
      sum(absolute_difference_block > tolerance)

    row_sums_original <- rowSums(S_original_block)
    row_sums_shifted <- rowSums(S_shifted_block)
    maximum_row_sum_deviation_original <- max(
      maximum_row_sum_deviation_original,
      abs(row_sums_original - 1)
    )
    maximum_row_sum_deviation_shifted <- max(
      maximum_row_sum_deviation_shifted,
      abs(row_sums_shifted - 1)
    )

    smoother_component[rows] <- as.vector(
      difference_block %*% Y_original
    )
    normalization_component[rows] <- k * (row_sums_shifted - 1)
    reconstructed_original[rows] <- as.vector(
      S_original_block %*% Y_original
    )
    reconstructed_shifted[rows] <- as.vector(
      S_shifted_block %*% (Y_original + k)
    )

    rm(
      S_original_block, S_shifted_block, difference_block,
      absolute_difference_block
    )
  }

  observed_residual <- as.numeric(prediction_shifted) -
    as.numeric(prediction_original) - k
  decomposition_residual <- observed_residual -
    smoother_component - normalization_component
  reconstruction_error_original <- as.numeric(prediction_original) -
    reconstructed_original
  reconstruction_error_shifted <- as.numeric(prediction_shifted) -
    reconstructed_shifted

  matrix_row <- data.frame(
    Outcome = outcome,
    Unit = unit,
    Estimator = estimator,
    Smoother = smoother_name,
    Exactly_identical = identical(smoother_original, smoother_shifted),
    Maximum_element_change = maximum_element_change,
    Element_RMSE = sqrt(squared_element_change / (N * N)),
    Median_row_L1_change = median(row_l1_change),
    Maximum_row_L1_change = max(row_l1_change),
    Share_rows_changed = mean(row_max_change > tolerance),
    Share_elements_changed = changed_elements / (N * N),
    Maximum_row_sum_deviation_original =
      maximum_row_sum_deviation_original,
    Maximum_row_sum_deviation_shifted =
      maximum_row_sum_deviation_shifted,
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  decomposition_row <- data.frame(
    Outcome = outcome,
    Unit = unit,
    Estimator = estimator,
    Smoother = smoother_name,
    Observed_residual_RMSE = rmse(observed_residual),
    Observed_residual_maximum = max_abs(observed_residual),
    Smoother_component_RMSE = rmse(smoother_component),
    Smoother_component_maximum = max_abs(smoother_component),
    Normalization_component_RMSE = rmse(normalization_component),
    Normalization_component_maximum = max_abs(normalization_component),
    Maximum_decomposition_residual = max_abs(decomposition_residual),
    Maximum_reconstruction_error_original =
      max_abs(reconstruction_error_original),
    Maximum_reconstruction_error_shifted =
      max_abs(reconstruction_error_shifted),
    stringsAsFactors = FALSE,
    check.names = FALSE
  )

  list(matrix = matrix_row, decomposition = decomposition_row)
}

compare_fits <- function(fit_original, fit_shifted, outcome, unit, k) {
  folds_identical <- identical(
    fit_original$NuPa.hat$cf_mat,
    fit_shifted$NuPa.hat$cf_mat
  )
  if (!folds_identical) {
    stop(sprintf(
      paste0(
        "Cross-fitting matrices differ for %s / %s; ",
        "the paired comparison is not interpreted."
      ),
      outcome, unit
    ))
  }

  summary_original <- summary(fit_original)
  summary_shifted <- summary(fit_shifted)
  estimators <- c("PLR-IV", "Wald-AIPW")

  estimate_rows <- do.call(rbind, lapply(estimators, function(estimator) {
    tau_original <- get_estimate(summary_original, estimator)
    tau_shifted <- get_estimate(summary_shifted, estimator)
    difference <- tau_shifted - tau_original
    data.frame(
      Outcome = outcome,
      Estimator = estimator,
      Unit = unit,
      Shift_k = k,
      tau_original = tau_original,
      tau_shifted = tau_shifted,
      rerun_difference = difference,
      absolute_difference = abs(difference),
      invariant_1e_8 = is.finite(difference) &&
        abs(difference) <= RERUN_TOLERANCE,
      folds_identical = folds_identical,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  }))

  p_o <- fit_original$NuPa.hat$predictions
  p_s <- fit_shifted$NuPa.hat$predictions
  s_o <- fit_original$NuPa.hat$smoothers
  s_s <- fit_shifted$NuPa.hat$smoothers

  prediction_rows <- do.call(rbind, list(
    # Outcome nuisances should shift by k under location equivariance.
    prediction_diagnostic(
      p_o$Y.hat, p_s$Y.hat, k,
      outcome, unit, "Y.hat (PLR-IV outcome)"
    ),
    prediction_diagnostic(
      p_o$Y.hat.z0, p_s$Y.hat.z0, k,
      outcome, unit, "Y.hat.z0 (Wald outcome)"
    ),
    prediction_diagnostic(
      p_o$Y.hat.z1, p_s$Y.hat.z1, k,
      outcome, unit, "Y.hat.z1 (Wald outcome)"
    ),
    # These nuisances do not use Y and should not shift.
    prediction_diagnostic(
      p_o$D.hat, p_s$D.hat, 0,
      outcome, unit, "D.hat (PLR-IV treatment)"
    ),
    prediction_diagnostic(
      p_o$D.hat.z0, p_s$D.hat.z0, 0,
      outcome, unit, "D.hat.z0 (Wald treatment)"
    ),
    prediction_diagnostic(
      p_o$D.hat.z1, p_s$D.hat.z1, 0,
      outcome, unit, "D.hat.z1 (Wald treatment)"
    ),
    prediction_diagnostic(
      p_o$Z.hat, p_s$Z.hat, 0,
      outcome, unit, "Z.hat (instrument)"
    )
  ))

  smoother_checks <- list(
    smoother_diagnostic(
      s_o$S, s_s$S,
      fit_original$data$Y, p_o$Y.hat, p_s$Y.hat, k,
      outcome, unit, "PLR-IV", "S"
    ),
    smoother_diagnostic(
      s_o$S.z0, s_s$S.z0,
      fit_original$data$Y, p_o$Y.hat.z0, p_s$Y.hat.z0, k,
      outcome, unit, "Wald-AIPW", "S.z0"
    ),
    smoother_diagnostic(
      s_o$S.z1, s_s$S.z1,
      fit_original$data$Y, p_o$Y.hat.z1, p_s$Y.hat.z1, k,
      outcome, unit, "Wald-AIPW", "S.z1"
    )
  )

  smoother_matrix_rows <- do.call(
    rbind,
    lapply(smoother_checks, `[[`, "matrix")
  )
  smoother_decomposition_rows <- do.call(
    rbind,
    lapply(smoother_checks, `[[`, "decomposition")
  )

  list(
    estimates = estimate_rows,
    predictions = prediction_rows,
    smoother_matrices = smoother_matrix_rows,
    smoother_decompositions = smoother_decomposition_rows
  )
}

fmt_fixed <- function(x, digits = DISPLAY_DIGITS) {
  x <- as.numeric(x)
  x[abs(x) < 0.5 * 10^(-digits)] <- 0
  formatC(x, format = "f", digits = digits)
}

fmt_scientific <- function(x, digits = 5L) {
  formatC(as.numeric(x), format = "e", digits = digits)
}
```

# Run the paired untuned-forest fits

```{r run-audit}
estimate_results <- list()
prediction_results <- list()
smoother_matrix_results <- list()
smoother_decomposition_results <- list()
result_index <- 1L

# Work status: binary 0/1 coding versus the exact +1 translation.
fit_labor_original <- fit_untuned(
  Y_labor, D_labor, X_labor, Z_labor
)
fit_labor_shifted <- fit_untuned(
  Y_labor + 1, D_labor, X_labor, Z_labor
)

labor_comparison <- compare_fits(
  fit_labor_original,
  fit_labor_shifted,
  outcome = "Work status",
  unit = "0/1 to 1/2",
  k = 1
)
estimate_results[[result_index]] <- labor_comparison$estimates
prediction_results[[result_index]] <- labor_comparison$predictions
smoother_matrix_results[[result_index]] <-
  labor_comparison$smoother_matrices
smoother_decomposition_results[[result_index]] <-
  labor_comparison$smoother_decompositions
result_index <- result_index + 1L

rm(fit_labor_original, fit_labor_shifted, labor_comparison)
invisible(gc())

# Maternal income: retain one dollar reference fit and compare the three exact
# additive unit transformations with that common reference.
fit_income_dollars <- fit_untuned(
  Y_income_dollars, D_income, X_income, Z_income
)

for (unit in names(income_units)) {
  fit_income_shifted <- fit_untuned(
    income_units[[unit]]$Y,
    D_income,
    X_income,
    Z_income
  )

  comparison <- compare_fits(
    fit_income_dollars,
    fit_income_shifted,
    outcome = "Maternal log income",
    unit = unit,
    k = income_units[[unit]]$k
  )
  estimate_results[[result_index]] <- comparison$estimates
  prediction_results[[result_index]] <- comparison$predictions
  smoother_matrix_results[[result_index]] <-
    comparison$smoother_matrices
  smoother_decomposition_results[[result_index]] <-
    comparison$smoother_decompositions
  result_index <- result_index + 1L

  rm(fit_income_shifted, comparison)
  invisible(gc())
}

rm(fit_income_dollars)
invisible(gc())

estimate_results <- do.call(rbind, estimate_results)
prediction_results <- do.call(rbind, prediction_results)
smoother_matrix_results <- do.call(rbind, smoother_matrix_results)
smoother_decomposition_results <- do.call(
  rbind,
  smoother_decomposition_results
)
```

# Estimate-level results

The (10^{-8}) classification below uses the unrounded rerun difference.
Formatting is applied only for display.

```{r estimate-table, echo=FALSE}
estimate_display <- estimate_results
numeric_columns <- c(
  "Shift_k", "tau_original", "tau_shifted",
  "rerun_difference", "absolute_difference"
)
estimate_display[numeric_columns] <- lapply(
  estimate_display[numeric_columns], fmt_fixed
)

estimate_display |>
  kbl(
    caption = "Untuned honest-forest complete-pipeline translation results",
    row.names = FALSE,
    align = c("l", "l", "l", rep("r", 5), "c", "c")
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    font_size = 12
  ) |>
  row_spec(
    which(!estimate_results$invariant_1e_8),
    background = "#fffbf0"
  ) |>
  footnote(general = paste0(
    "Invariant means that the absolute unrounded rerun difference is at most ",
    format(RERUN_TOLERANCE, scientific = TRUE),
    ". Every interpreted pair has an identical realized cross-fitting matrix."
  ))
```

# Nuisance-prediction diagnostics

For outcome nuisances, the expected prediction change is (k). For treatment
and instrument nuisances, which do not use (Y), the expected change is zero.

```{r prediction-table, echo=FALSE}
prediction_display <- prediction_results
prediction_numeric_columns <- c(
  "Expected_shift", "Maximum_absolute_error", "RMSE",
  "Median_absolute_error"
)
prediction_display[prediction_numeric_columns] <- lapply(
  prediction_display[prediction_numeric_columns], fmt_fixed
)

prediction_display |>
  kbl(
    caption = "Nuisance-prediction translation residuals",
    row.names = FALSE
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    font_size = 11
  )
```

# Realized smoother-matrix diagnostics

The first table compares the original and translated realized outcome smoother
matrices. `Exactly_identical` is a bit-for-bit comparison. The numerical change
statistics use the matrix difference \(S^k-S\). A row is classified as changed
when at least one of its weights differs by more than
`r format(SMOOTHER_TOLERANCE, scientific = TRUE)`.

```{r smoother-matrix-table, echo=FALSE}
smoother_matrix_display <- smoother_matrix_results
smoother_matrix_numeric_columns <- c(
  "Maximum_element_change", "Element_RMSE",
  "Median_row_L1_change", "Maximum_row_L1_change",
  "Share_rows_changed", "Share_elements_changed",
  "Maximum_row_sum_deviation_original",
  "Maximum_row_sum_deviation_shifted"
)
smoother_matrix_display[smoother_matrix_numeric_columns] <- lapply(
  smoother_matrix_display[smoother_matrix_numeric_columns],
  fmt_scientific
)

smoother_matrix_display |>
  kbl(
    caption = "Original versus translated realized smoother matrices",
    row.names = FALSE
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    font_size = 10
  ) |>
  footnote(general = paste0(
    "Rows and elements are classified as numerically changed at an absolute ",
    "weight tolerance of ",
    format(SMOOTHER_TOLERANCE, scientific = TRUE), "."
  ))
```

The second table verifies the observation-level identity

\[
\widehat{\mathbf g}^{\,k}-\widehat{\mathbf g}-k\mathbf 1
=
(S^k-S)\mathbf Y+k(S^k\mathbf 1-\mathbf 1).
\]

The first term on the right is the **smoother-change component**. The second is
the **normalization component**. The decomposition residual also incorporates
any discrepancy between the stored forest predictions and their reconstruction
from the returned smoother weights.

```{r smoother-decomposition-table, echo=FALSE}
smoother_decomposition_display <- smoother_decomposition_results
smoother_decomposition_numeric_columns <- setdiff(
  names(smoother_decomposition_display),
  c("Outcome", "Unit", "Estimator", "Smoother")
)
smoother_decomposition_display[smoother_decomposition_numeric_columns] <-
  lapply(
    smoother_decomposition_display[smoother_decomposition_numeric_columns],
    fmt_scientific
  )

smoother_decomposition_display |>
  kbl(
    caption = paste0(
      "Decomposition of the outcome-nuisance translation residual: ",
      "observed = smoother change + normalization"
    ),
    row.names = FALSE
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    font_size = 9
  )
```

# Optional internal export

No output file is written under the default setting. Change `SAVE_RESULTS` to
`TRUE` only if the internal diagnostic results should be retained.

```{r optional-save}
settings <- list(
  seed = SEED,
  n_cf = N_CF,
  tuning = "none",
  tolerance = RERUN_TOLERANCE,
  smoother_tolerance = SMOOTHER_TOLERANCE,
  smoother_block_rows = SMOOTHER_BLOCK_ROWS,
  display_digits = DISPLAY_DIGITS,
  covariates = CHILD_COVARIATES,
  labor_n = nrow(ae98_labor_sub),
  income_n = nrow(ae98_income_sub),
  input_file = "ae98_sub.rds",
  OutcomeWeights_version = as.character(packageVersion("OutcomeWeights")),
  grf_version = as.character(packageVersion("grf"))
)

if (SAVE_RESULTS) {
  saveRDS(
    list(
      estimates = estimate_results,
      prediction_diagnostics = prediction_results,
      smoother_matrix_diagnostics = smoother_matrix_results,
      smoother_decomposition_diagnostics =
        smoother_decomposition_results,
      unit_coding_max_deviation = unit_coding_max_deviation,
      settings = settings
    ),
    "child_untuned_forest_translation_audit_results.rds"
  )
  message("Saved child_untuned_forest_translation_audit_results.rds")
} else {
  message("SAVE_RESULTS is FALSE; no result file was written.")
}
```
