---
title: "Angrist (1990) — Descriptive Diagnostics"
subtitle: "Sample & Missingness · Z/D/Y Distributions · Compliance · First Stage / Reduced Form / Wald · Propensity Overlap · Unadjusted Balance"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
h3   { color: #444; margin-top: 20px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.key-box  { background:#fdeaea; border-left:4px solid #d7191c; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 5.5, dpi = 110)
```

```{r packages}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(AER)
library(sandwich)
library(lmtest)
library(scales)
library(cobalt)

source("functions_all.R")

# Consistent palette
col_z1   <- "#2c7bb6"   # Z = 1 (draft eligible)
col_z0   <- "#d7191c"   # Z = 0 (draft ineligible)
col_grey <- "#4d4d4d"
```

```{r data}
data_path <- "/Users/alexung/desktop/suw_jbes_replicate"
sipp_raw  <- read_dta(file.path(data_path, "sipp.dta"))   # full sample, for missingness

# Sample restriction (suw_jbes_replicate.do): drop rsncode==999, missing educ / kwage
sipp <- sipp_raw[sipp_raw$rsncode != 999 &
                   !is.na(sipp_raw$educ) &
                   !is.na(sipp_raw$kwage), ]

# Variable definitions — EXACTLY as in the Stata file: age = age_5 (age, NOT cohort)
sipp$age     <- sipp$age_5
sipp$age2    <- sipp$age_5^2
sipp$age3    <- sipp$age_5^3
sipp$age_fac <- factor(sipp$age_5)

sipp$Z      <- as.integer(sipp$rsncode)   # instrument: draft-eligibility indicator
sipp$nvstat <- as.integer(sipp$nvstat)    # treatment: Vietnam-era veteran status
sipp$lwage  <- log(sipp$kwage)            # outcome (log wage, dollars)

# Design matrices (with intercept) for the three age specifications
X_lin <- model.matrix(~ age,               sipp)
X_cub <- model.matrix(~ age + age2 + age3, sipp)
X_sat <- model.matrix(~ age_fac,           sipp)

# Propensity scores p(X) = Pr(Z = 1 | X) under each spec  (engine: logit_mle)
sipp$p_lin <- logit_mle(sipp$Z, X_lin)
sipp$p_cub <- logit_mle(sipp$Z, X_cub)
sipp$p_sat <- logit_mle(sipp$Z, X_sat)

cat(sprintf("N (analysis) = %d | Pr(Z=1) = %.3f | Pr(D=1) = %.3f | age range %d–%d\n",
            nrow(sipp), mean(sipp$Z), mean(sipp$nvstat),
            min(sipp$age_5), max(sipp$age_5)))
```

<div class="note-box">
**Scope.** Purely descriptive: it characterises the sample, the instrument, the treatment, the outcome, and covariate overlap *before* any estimator is applied. Weight-level diagnostics (ESS, weight concentration, love plots) live in the main analysis notebook. **Variable note:** following the SUW Stata replication, the only covariate is `age = age_5` (respondent age — *not* birth cohort), entered linearly, cubically, or as age dummies (saturated). Sections are self-contained, so more descriptives can be appended without disturbing what is here.
</div>

---

# 1. Sample & Missingness

<div class="note-box">
**What we look at.** Raw sample size, how many rows the SUW restriction drops (`rsncode == 999`, missing `educ` / `kwage`), the type of every variable, and where missing values sit.
</div>

```{r sample-size}
n_drop <- nrow(sipp_raw) - nrow(sipp)
data.frame(
  Stage = c("Raw file", "Drop rsncode==999 / missing educ / missing kwage", "Analysis sample"),
  N     = c(nrow(sipp_raw), n_drop, nrow(sipp)),
  check.names = FALSE
) %>%
  kbl(caption = "Sample construction", align = c("l","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE, font_size = 14) %>%
  row_spec(3, bold = TRUE)
```

```{r missingness}
miss_tab <- data.frame(
  Variable    = names(sipp_raw),
  Class       = vapply(sipp_raw, function(x) class(x)[1], character(1)),
  n_missing   = vapply(sipp_raw, function(x) sum(is.na(x)), integer(1)),
  pct_missing = round(100 * vapply(sipp_raw, function(x) mean(is.na(x)), numeric(1)), 2),
  n_unique    = vapply(sipp_raw, function(x) length(unique(x)), integer(1)),
  row.names   = NULL
)
miss_tab <- miss_tab[order(-miss_tab$n_missing), ]

miss_tab %>%
  kbl(caption = "Variable inventory and missingness (raw file, sorted by missing count)",
      col.names = c("Variable","Class","# missing","% missing","# unique"),
      align = c("l","l","r","r","r"), row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12.5) %>%
  row_spec(which(miss_tab$n_missing > 0), background = "#fdeaea") %>%
  scroll_box(height = "360px") %>%
  footnote(general = "Highlighted rows have at least one missing value. After the restriction the analysis sample is used everywhere below.")
```

---

# 2. Distributions of Z, D, and Y

<div class="note-box">
**Z** = draft eligibility (instrument), **D** = `nvstat` (Vietnam-era veteran status), **Y** = log wage. Z and D are binary (shown as proportions); Y is continuous (shown as a histogram, split by instrument status).
</div>

```{r dist-binary, fig.height=3.6}
bin_df <- bind_rows(
  data.frame(Variable = "Z (draft eligible)", value = sipp$Z),
  data.frame(Variable = "D: veteran (nvstat)", value = sipp$nvstat)
) %>%
  group_by(Variable) %>%
  summarise(`= 1` = mean(value), `= 0` = 1 - mean(value), .groups = "drop") %>%
  pivot_longer(c(`= 1`, `= 0`), names_to = "level", values_to = "share")

ggplot(bin_df, aes(x = Variable, y = share, fill = level)) +
  geom_col(width = 0.55, alpha = 0.9) +
  geom_text(aes(label = sprintf("%.1f%%", 100 * share)),
            position = position_stack(vjust = 0.5), size = 3.4, colour = "white") +
  scale_fill_manual(values = c("= 1" = col_z1, "= 0" = "#bdbdbd")) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  coord_flip() +
  labs(title = "Binary variables: instrument Z and treatment D", x = NULL, y = NULL, fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", panel.grid.major.y = element_blank(),
        plot.title = element_text(face = "bold"))
```

```{r dist-y, fig.height=4.6}
sipp %>%
  mutate(Zlab = factor(ifelse(Z == 1, "Z = 1 (eligible)", "Z = 0 (ineligible)"),
                       levels = c("Z = 1 (eligible)", "Z = 0 (ineligible)"))) %>%
  ggplot(aes(x = lwage, fill = Zlab)) +
  geom_histogram(bins = 40, alpha = 0.65, colour = "white", position = "identity") +
  geom_vline(xintercept = mean(sipp$lwage), linetype = "dashed", colour = col_grey) +
  scale_fill_manual(values = c("Z = 1 (eligible)" = col_z1, "Z = 0 (ineligible)" = col_z0)) +
  labs(title = "Outcome: log wage, by instrument status",
       subtitle = "Dashed line = overall mean log wage.",
       x = "log(wage)", y = "Count", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 10))
```

```{r dist-y-summary}
data.frame(
  Group = c("All", "Z = 1 (eligible)", "Z = 0 (ineligible)"),
  N     = c(nrow(sipp), sum(sipp$Z == 1), sum(sipp$Z == 0)),
  Mean  = c(mean(sipp$lwage), mean(sipp$lwage[sipp$Z==1]), mean(sipp$lwage[sipp$Z==0])),
  SD    = c(sd(sipp$lwage),   sd(sipp$lwage[sipp$Z==1]),   sd(sipp$lwage[sipp$Z==0])),
  Min   = c(min(sipp$lwage),  min(sipp$lwage[sipp$Z==1]),  min(sipp$lwage[sipp$Z==0])),
  Max   = c(max(sipp$lwage),  max(sipp$lwage[sipp$Z==1]),  max(sipp$lwage[sipp$Z==0]))
) %>%
  mutate(across(c(Mean,SD,Min,Max), ~round(., 3))) %>%
  kbl(caption = "Log wage summary by instrument status", align = c("l","r","r","r","r","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE, font_size = 13)
```

---

# 3. Z × D Compliance Table

<div class="note-box">
**Reading it.** The 2×2 cross-tab of instrument against treatment. Under monotonicity, the first-stage gap Pr(D=1|Z=1) − Pr(D=1|Z=0) is the **complier share**, Pr(D=1|Z=0) approximates the **always-taker share** (men who served despite ineligibility), and 1 − Pr(D=1|Z=1) the **never-taker share**. A non-zero always-taker rate is the descriptive signature of *incomplete* compliance — central to this design.
</div>

```{r compliance}
compliance_block <- function(D, treat_name) {
  tab <- table(Z = sipp$Z, D = D)
  p_d_z1 <- mean(D[sipp$Z == 1])
  p_d_z0 <- mean(D[sipp$Z == 0])
  data.frame(
    Treatment       = treat_name,
    `n(Z=0,D=0)`    = tab["0","0"], `n(Z=0,D=1)` = tab["0","1"],
    `n(Z=1,D=0)`    = tab["1","0"], `n(Z=1,D=1)` = tab["1","1"],
    `Pr(D=1|Z=1)`   = round(p_d_z1, 3),
    `Pr(D=1|Z=0)`   = round(p_d_z0, 3),
    `Compliers`     = round(p_d_z1 - p_d_z0, 3),
    `Always-takers` = round(p_d_z0, 3),
    `Never-takers`  = round(1 - p_d_z1, 3),
    check.names = FALSE
  )
}

compliance_block(sipp$nvstat, "nvstat (veteran)") %>%
  kbl(caption = "Z × D compliance — counts and implied principal-strata shares (monotonicity)",
      align = c("l", rep("r", 9))) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12.5) %>%
  add_header_above(c(" " = 1, "Cell counts" = 4, "Conditional treatment rates" = 2,
                     "Implied strata" = 3)) %>%
  column_spec(9, bold = TRUE, color = "#2c7bb6") %>%
  footnote(general = "Complier share = Pr(D=1|Z=1) − Pr(D=1|Z=0). A positive always-taker rate (men serving despite ineligibility) signals incomplete compliance.")
```

---

# 4. First Stage, Reduced Form, and Wald

<div class="note-box">
**The three building blocks.** First stage = effect of Z on D; reduced form = effect of Z on Y; their ratio is the Wald / IV estimate. The HC1-robust first-stage F (squared t on a single instrument) is the weak-instrument check — compare against the rule-of-thumb 10. Three age specifications: linear, cubic, saturated.
</div>

```{r first-stage-table}
# --- HC1-robust first-stage F (coefficient on the instrument Z) ---
fs_F <- function(formula, data) {
  fit <- lm(formula, data = data)
  ct  <- coeftest(fit, vcov = vcovHC(fit, type = "HC1"))
  c(coef = unname(ct["Z","Estimate"]),
    se   = unname(ct["Z","Std. Error"]),
    F    = unname((ct["Z","Estimate"]/ct["Z","Std. Error"])^2))
}

LIN_RHS <- "age"
CUB_RHS <- "age + age2 + age3"
SAT_RHS <- "age_fac"

fs_rows <- rbind(
  data.frame(Spec="Linear age",    t(fs_F(as.formula(paste("nvstat ~ Z +", LIN_RHS)), sipp))),
  data.frame(Spec="Cubic age",     t(fs_F(as.formula(paste("nvstat ~ Z +", CUB_RHS)), sipp))),
  data.frame(Spec="Saturated age", t(fs_F(as.formula(paste("nvstat ~ Z +", SAT_RHS)), sipp)))
)
fs_rows$coef <- round(fs_rows$coef, 4)
fs_rows$se   <- round(fs_rows$se, 4)
fs_rows$F    <- round(fs_rows$F, 1)
fs_rows$Strength <- ifelse(fs_rows$F >= 10, "strong", "WEAK")

fs_rows %>%
  kbl(caption = "First-stage strength: effect of draft eligibility (Z) on veteran status",
      col.names = c("Spec","First stage δ̂","SE","F-stat","> 10?"),
      align = c("l","r","r","r","c")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  row_spec(which(fs_rows$F < 10), background = "#fdeaea", bold = TRUE) %>%
  footnote(general = "HC1-robust first-stage F on the single instrument. F < 10 (highlighted) flags a weak first stage.")
```

```{r rf-wald-table}
rf_coef <- function(rhs) {
  fit <- lm(as.formula(paste("lwage ~ Z +", rhs)), data = sipp)
  ct  <- coeftest(fit, vcov = vcovHC(fit, type = "HC1"))
  unname(ct["Z","Estimate"])
}
fs_coef <- function(rhs) unname(fs_F(as.formula(paste("nvstat ~ Z +", rhs)), sipp)["coef"])
raw_wald <- function() {
  num <- mean(sipp$lwage[sipp$Z==1]) - mean(sipp$lwage[sipp$Z==0])
  den <- mean(sipp$nvstat[sipp$Z==1]) - mean(sipp$nvstat[sipp$Z==0])
  num / den
}

wald_block <- function(spec, rhs) {
  fs <- fs_coef(rhs); rf <- rf_coef(rhs)
  data.frame(
    Spec = spec,
    `First stage δ̂` = round(fs, 4),
    `Reduced form γ̂` = round(rf, 4),
    `Wald = γ̂/δ̂ (adj.)` = round(rf / fs, 4),
    `Raw Wald (no cov.)` = round(raw_wald(), 4),
    check.names = FALSE
  )
}

rbind(
  wald_block("Linear age",    LIN_RHS),
  wald_block("Cubic age",     CUB_RHS),
  wald_block("Saturated age", SAT_RHS)
) %>%
  kbl(caption = "Reduced form, first stage, and Wald (Y = log wage)",
      align = c("l","r","r","r","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  column_spec(4, bold = TRUE) %>%
  footnote(general = "Covariate-adjusted Wald equals just-identified 2SLS. The raw Wald uses unconditional group means (no covariates) and is identical across specs.")
```

## First Stage by Age (design-specific)

<div class="note-box">
Because draft-eligibility ceilings differ by cohort, the unconditional first stage varies with age. This view (veteran rate by eligibility within each age) shows the relevance condition holding across ages and the positive ineligible-group rates that mark incomplete compliance.
</div>

```{r first-stage-age-plot, fig.height=5}
age_keep <- sipp %>% count(age_5) %>% filter(n >= 50) %>% pull(age_5)

fs_age <- sipp %>%
  filter(age_5 %in% age_keep) %>%
  mutate(Zlab = factor(ifelse(Z == 1, "Z = 1 (eligible)", "Z = 0 (ineligible)"),
                       levels = c("Z = 0 (ineligible)", "Z = 1 (eligible)"))) %>%
  group_by(age_5, Zlab) %>%
  summarise(vet_rate = mean(nvstat), .groups = "drop")

pooled_fs <- mean(sipp$nvstat[sipp$Z==1]) - mean(sipp$nvstat[sipp$Z==0])

ggplot(fs_age, aes(x = factor(age_5), y = vet_rate, fill = Zlab)) +
  geom_col(position = position_dodge(0.7), width = 0.62, alpha = 0.9) +
  scale_fill_manual(values = c("Z = 0 (ineligible)" = col_z0, "Z = 1 (eligible)" = col_z1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(title = "First stage: veteran rate by draft eligibility and age",
       subtitle = sprintf("Pooled first-stage gap = %.3f. Positive ineligible bars = incomplete compliance.", pooled_fs),
       x = "Age (age_5)", y = "Pr(veteran)", fill = NULL,
       caption = "Age cells with fewer than 50 observations excluded.") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", panel.grid.major.x = element_blank(),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 10),
        plot.caption = element_text(colour = "grey50", size = 9))
```

---

# 5. Propensity Score: Concentration & Overlap

<div class="note-box">
The propensity score is p(X) = Pr(Z = 1 | X) — the model for the *instrument*, fitted with the engine's `logit_mle`. Because draft eligibility is near-randomly assigned conditional on age, p̂(X) is expected to cluster tightly around the pooled Pr(Z=1) ≈ 0.46. This near-flat surface is the design feature that keeps weights well-behaved later.
</div>

## Summary Statistics

```{r ps-summary}
ps_summary <- function(p, name) {
  data.frame(Specification = name,
             Min = min(p), P05 = quantile(p,0.05), P25 = quantile(p,0.25),
             Mean = mean(p), P75 = quantile(p,0.75), P95 = quantile(p,0.95),
             Max = max(p), SD = sd(p),
             `near 0/1` = mean(p < 0.05 | p > 0.95),
             check.names = FALSE)
}
ps_tab <- bind_rows(ps_summary(sipp$p_lin, "Linear age"),
                    ps_summary(sipp$p_cub, "Cubic age"),
                    ps_summary(sipp$p_sat, "Saturated age"))
ps_tab[,2:10] <- lapply(ps_tab[,2:10], function(x) round(x, 4))

ps_tab %>%
  kbl(caption = "Propensity score p̂(X) = Pr(Z=1|X): summary by specification",
      align = c("l", rep("r", 9))) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12.5) %>%
  column_spec(9, bold = TRUE) %>%
  column_spec(10, bold = TRUE, color = "#d7191c") %>%
  footnote(general = "SD and the 'near 0/1' share summarise dispersion. Tight clustering around the mean reflects near-random assignment.")
```

## Distribution by Instrument Status

```{r ps-dist-plot, fig.height=7}
ps_long <- sipp %>%
  select(p_lin, p_cub, p_sat, Z) %>%
  pivot_longer(c(p_lin, p_cub, p_sat), names_to = "spec", values_to = "p_hat") %>%
  mutate(Specification = factor(spec, levels = c("p_lin", "p_cub", "p_sat"),
                                labels = c("Linear age", "Cubic age", "Saturated age")),
         Zlab = factor(ifelse(Z == 1, "Z = 1 (eligible)", "Z = 0 (ineligible)"),
                       levels = c("Z = 1 (eligible)", "Z = 0 (ineligible)")))

ggplot(ps_long, aes(x = p_hat, fill = Zlab)) +
  geom_histogram(bins = 40, alpha = 0.7, colour = "white", position = "identity") +
  geom_vline(xintercept = mean(sipp$Z), linetype = "dashed", colour = col_grey) +
  facet_wrap(~ Specification, ncol = 1, scales = "free_y") +
  scale_fill_manual(values = c("Z = 1 (eligible)" = col_z1, "Z = 0 (ineligible)" = col_z0)) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(title = "Propensity score distribution by instrument status",
       subtitle = "Overlap requires both colours present at every p̂. Tight clustering = near-random assignment.",
       x = expression(hat(p)(X) == Pr(Z == 1 ~ "|" ~ X)), y = "Count", fill = NULL,
       caption = "Dashed line = pooled Pr(Z=1).") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 9.5),
        plot.caption = element_text(colour = "grey50", size = 9))
```

---

# 6. Unadjusted Covariate Balance

<div class="note-box">
**Instrument balance.** Standardized mean differences of the age covariates across the instrument (Z = 1 vs. Z = 0), with no weighting. Because eligibility ceilings differ by cohort, age is *related* to Z by construction — so non-trivial imbalance here is exactly why the design conditions on age. Binary/continuous covariates are placed on the same standardized scale (`binary = "std"`), with KS statistics added.
</div>

```{r unadj-balance, fig.height=4.5}
covs <- data.frame(age = sipp$age, age2 = sipp$age2, age3 = sipp$age3)

# stats/binary declared in bal.tab() so cobalt computes KS upstream of love.plot()
bal <- bal.tab(covs, treat = sipp$Z,
               binary = "std", s.d.denom = "pooled",
               stats  = c("mean.diffs", "ks.statistics"))

love.plot(bal,
          stats      = c("mean.diffs", "ks.statistics"),
          binary     = "std",
          abs        = TRUE,
          thresholds = c(m = 0.1),
          var.order  = "unadjusted",
          colors     = col_z1,
          shapes     = "circle",
          title      = "Unadjusted covariate balance across the instrument Z (age polynomial)")
```

```{r unadj-balance-table}
smd_tab <- bal$Balance
smd_tab$Variable <- rownames(smd_tab)
keep_cols <- intersect(c("Variable","Type","Diff.Un"), names(smd_tab))
smd_tab[, keep_cols] %>%
  arrange(desc(abs(Diff.Un))) %>%
  mutate(Diff.Un = round(Diff.Un, 3)) %>%
  kbl(caption = "Unadjusted standardized differences across Z (sorted by |SMD|)",
      col.names = c("Variable","Type","SMD (unadjusted)"),
      align = c("l","l","r"), row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE, font_size = 12.5) %>%
  footnote(general = "SMD across the instrument. |SMD| > 0.1 (a common rule of thumb) flags covariates imbalanced with respect to Z — here, the age structure the design adjusts for.")
```

---

<div class="note-box">
**Extending this notebook.** Each section is independent and reads only from `sipp` (analysis sample) plus the engine functions. To add a descriptive, append a new `#`-level section; nothing above needs to change.
</div>
