---
title: "Angrist (1990) — Outcome Weights & Kappa Estimators"
subtitle: "Vietnam Draft Lottery: SUW Replication, DML & Learner Comparison"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body   { font-size: 15px; color: #2c2c2c; }
h1     { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2     { color: #2c7bb6; margin-top: 28px; }
h3     { color: #444; margin-top: 20px; }

.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 12,
  fig.height = 6
)
```

---

# Setup

## Packages & Functions

```{r packages-and-source}
# Core
library(haven)
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
# DML smoother + balance plots
library(OutcomeWeights)
library(cobalt)
library(viridis)
library(gridExtra)
library(ggplot2)
# DoubleML learners
library(DoubleML)
library(mlr3verse)

source("functions_all.R")   # all estimators, weight diagnostics (incl. ESS_mod),
                            # get_dml_coef / get_dml_se, make_X, make_love, etc.
cat("functions_all.R sourced.\n")
```

## Data & Design

<div class="note-box">
**Setting.** Angrist (1990) SIPP 1984 subsample (N = 3,027 white men, born 1950–53). Instrument Z = `rsncode` (draft-eligibility indicator). Treatment D = `nvstat` (Vietnam-era veteran status). Outcome: log wages in two units (`lwage = log(kwage)` dollars; `lwage_cnt = log(100·kwage)` cents) — the units only enter the SUW reproduction table. Variable definitions follow the SUW Stata replication exactly: `age = age_5`, with three covariate specifications — **Linear** (`age`), **Cubic** (`age, age², age³`), and **Saturated** (`i.age` = age dummies).
</div>

```{r data}
data_path  <- "/Users/alexung/desktop/suw_jbes_replicate"
sipp       <- read_dta(file.path(data_path, "sipp.dta"))

# Sample restriction: N = 3,027  (drop rsncode==999, missing educ / kwage)
sipp_clean <- sipp[sipp$rsncode != 999 &
                     !is.na(sipp$educ) &
                     !is.na(sipp$kwage), ]

# Age terms — EXACTLY as in suw_jbes_replicate.do: age = age_5
sipp_clean$age     <- sipp_clean$age_5
sipp_clean$age2    <- sipp_clean$age_5^2
sipp_clean$age3    <- sipp_clean$age_5^3
sipp_clean$age_fac <- factor(sipp_clean$age_5)

# Outcomes in two units (note: dollars vs. cents convention is FLIPPED vs. Card)
sipp_clean$lwage     <- log(sipp_clean$kwage)         # dollars
sipp_clean$lwage_cnt <- log(100 * sipp_clean$kwage)   # cents

# Core vectors
Y_cnt <- sipp_clean$lwage_cnt   # cents  (used only for the SUW reproduction layout)
Y_dol <- sipp_clean$lwage       # dollars (the working outcome everywhere else)
Z     <- as.integer(sipp_clean$rsncode)
D     <- as.integer(sipp_clean$nvstat)

cat(sprintf("N = %d (should be 3,027)\n", nrow(sipp_clean)))
cat(sprintf("mean(Z) = %.4f | mean(D) = %.4f\n", mean(Z), mean(D)))
```

```{r design-matrices}
# Kappa design matrices — WITH intercept (for logit / CBPS)
X_lin <- model.matrix(~ age,                data = sipp_clean)
X_cub <- model.matrix(~ age + age2 + age3,  data = sipp_clean)
X_sat <- model.matrix(~ age_fac,            data = sipp_clean)

# X_df for 2SLS (no intercept column)
X_lin_df <- data.frame(age = sipp_clean$age)
X_cub_df <- data.frame(age = sipp_clean$age, age2 = sipp_clean$age2, age3 = sipp_clean$age3)
X_sat_df <- data.frame(age_fac = sipp_clean$age_fac)

# DML design matrices — no intercept, attributes stripped for grf (make_X)
X_dml_cub      <- make_X(~ 0 + age + I(age^2) + I(age^3), data = sipp_clean)
X_dml_sat_full <- make_X(~ 0 + age_fac, data = sipp_clean)
X_dml_sat      <- X_dml_sat_full[, -1, drop = FALSE]   # drop one age dummy (reference)

cat(sprintf("Kappa: Linear %d / Cubic %d / Saturated %d covariates | DML: Cubic %dx%d / Saturated %dx%d\n",
            ncol(X_lin) - 1, ncol(X_cub) - 1, ncol(X_sat) - 1,
            nrow(X_dml_cub), ncol(X_dml_cub), nrow(X_dml_sat), ncol(X_dml_sat)))
```

---

# 1. Reproduction — SUW Table 2

<div class="note-box">
**Layout.** One treatment (veteran status), three age specifications, two outcome units (cents / dollars). Columns are spec × unit. Normalized estimators (τ_cb_u, τ_ml_u, τ_ml_a10) are translation invariant — cents and dollars match; unnormalized ones (τ_ml_a, τ_ml_t, τ_ml_a0) do not.
</div>

```{r build-columns, cache=TRUE}
# One full column = 2SLS + all six kappa estimators with analytical SEs.
run_viet_column <- function(Y, X_kappa, X_df) {
  list(
    tsls  = run_2sls(Y, D, Z, X_df, endog_name = "nvstat"),
    kappa = kappa_analytic_se_all(Y, Z, D, X_kappa)
  )
}

col_lin_cnt <- run_viet_column(Y_cnt, X_lin, X_lin_df)
col_lin_dol <- run_viet_column(Y_dol, X_lin, X_lin_df)
col_cub_cnt <- run_viet_column(Y_cnt, X_cub, X_cub_df)
col_cub_dol <- run_viet_column(Y_dol, X_cub, X_cub_df)
col_sat_cnt <- run_viet_column(Y_cnt, X_sat, X_sat_df)
col_sat_dol <- run_viet_column(Y_dol, X_sat, X_sat_df)

cat("All 6 columns computed.\n")
```

```{r table-builder, echo=FALSE}
cell_fmt <- function(est, se, digits = 3) {
  if (is.na(est)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0)
    return(sprintf("%.3f<br><span style='color:#888'>(NA)</span>", est))
  pval  <- 2 * pnorm(-abs(est / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s<br><span style='color:#888'>(%.3f)</span>", est, stars, se)
}

build_spec_table <- function(c_lin_cnt, c_lin_dol, c_cub_cnt, c_cub_dol, c_sat_cnt, c_sat_dol) {
  ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
  labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")

  rows <- list(data.frame(
    Estimator = "2SLS",
    `Linear / cents`      = cell_fmt(c_lin_cnt$tsls$coef, c_lin_cnt$tsls$se),
    `Linear / dollars`    = cell_fmt(c_lin_dol$tsls$coef, c_lin_dol$tsls$se),
    `Cubic / cents`       = cell_fmt(c_cub_cnt$tsls$coef, c_cub_cnt$tsls$se),
    `Cubic / dollars`     = cell_fmt(c_cub_dol$tsls$coef, c_cub_dol$tsls$se),
    `Saturated / cents`   = cell_fmt(c_sat_cnt$tsls$coef, c_sat_cnt$tsls$se),
    `Saturated / dollars` = cell_fmt(c_sat_dol$tsls$coef, c_sat_dol$tsls$se),
    check.names = FALSE, stringsAsFactors = FALSE
  ))

  for (i in seq_along(ests)) {
    e <- ests[i]
    rows[[length(rows) + 1]] <- data.frame(
      Estimator = labels[i],
      `Linear / cents`      = cell_fmt(c_lin_cnt$kappa$estimates[e], c_lin_cnt$kappa$se[e]),
      `Linear / dollars`    = cell_fmt(c_lin_dol$kappa$estimates[e], c_lin_dol$kappa$se[e]),
      `Cubic / cents`       = cell_fmt(c_cub_cnt$kappa$estimates[e], c_cub_cnt$kappa$se[e]),
      `Cubic / dollars`     = cell_fmt(c_cub_dol$kappa$estimates[e], c_cub_dol$kappa$se[e]),
      `Saturated / cents`   = cell_fmt(c_sat_cnt$kappa$estimates[e], c_sat_cnt$kappa$se[e]),
      `Saturated / dollars` = cell_fmt(c_sat_dol$kappa$estimates[e], c_sat_dol$kappa$se[e]),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }
  do.call(rbind, rows)
}
```

## Results

```{r table-reproduction}
build_spec_table(col_lin_cnt, col_lin_dol, col_cub_cnt, col_cub_dol, col_sat_cnt, col_sat_dol) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c","c","c"),
      caption = "SUW Table 2 — Treatment: Vietnam veteran status (nvstat). Effect of military service on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12.5) %>%
  add_header_above(c(" " = 1, "Linear age" = 2, "Cubic age" = 2, "Saturated age" = 2)) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",            1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant). Orange = unnormalized. age = age_5. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 2. Double ML — PLR-IV & Wald-AIPW

<div class="note-box">
DML uses `dml_with_smoother()` (5-fold, `tune.parameters = "all"`) on the two specifications rich enough for a machine-learning smoother: **cubic age** and **saturated age**. The linear-age spec is too low-dimensional for the ML nuisance models and is therefore not fit here (it appears in the reproduction table above via kappa / 2SLS).
</div>

```{r dml-fit-all, cache=TRUE}
# Two DML fits on Y_dol: cubic and saturated age.
set.seed(42); dml_cub <- dml_with_smoother(Y_dol, D, X_dml_cub, Z, n_cf_folds = 5, tune.parameters = "all")
set.seed(42); dml_sat <- dml_with_smoother(Y_dol, D, X_dml_sat, Z, n_cf_folds = 5, tune.parameters = "all")

res_cub <- summary(dml_cub); res_sat <- summary(dml_sat)

extract_w <- function(dml_obj) {
  om <- get_outcome_weights(dml_obj)
  check_omega_rows(om)
  list(plriv    = as.vector(om$omega["PLR-IV",    ]),
       waldaipw = as.vector(om$omega["Wald-AIPW", ]))
}
w_cub <- extract_w(dml_cub)
w_sat <- extract_w(dml_sat)

cat("Both DML fits complete.\n")
```

## Estimates {.tabset}

```{r dml-cell-display, echo=FALSE}
show_dml_cell <- function(res, wlist, Y, cap) {
  ids <- c(
    check_weight_identity(wlist$plriv,    Y, get_estimate(res, "PLR-IV")),
    check_weight_identity(wlist$waldaipw, Y, get_estimate(res, "Wald-AIPW"))
  )
  df <- as.data.frame(res[c("PLR-IV", "Wald-AIPW"), , drop = FALSE])
  df$Estimator  <- rownames(df)
  df$`ΣωY = τ̂` <- as.character(ids)
  df <- df[, c("Estimator", setdiff(names(df), c("Estimator","ΣωY = τ̂")), "ΣωY = τ̂")]
  df %>%
    kbl(caption = cap, digits = 4, row.names = FALSE,
        align = c("l", rep("r", ncol(df) - 2), "c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "ΣωY = τ̂: algebraic identity check (must be TRUE).")
}
```

### Cubic age

```{r dml-cub}
show_dml_cell(res_cub, w_cub, Y_dol, "DML — Cubic age, Y = log wages (dollars)")
```

### Saturated age

```{r dml-sat}
show_dml_cell(res_sat, w_sat, Y_dol, "DML — Saturated age, Y = log wages (dollars)")
```

## DML vs. Kappa — Point Estimates

```{r dml-summary-table}
# Kappa estimates (dollars) for both specs — reused later by the learner table.
kappa_cub <- kappa_analytic_se_all(Y_dol, Z, D, X_cub)
kappa_sat <- kappa_analytic_se_all(Y_dol, Z, D, X_sat)

cell <- function(coef, se) {
  if (is.na(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

dml_cell <- function(res, est) {
  coef_val <- get_estimate(res, est)
  se_val   <- if ("SE" %in% colnames(res)) as.numeric(res[est, "SE"])
              else if ("Std. Error" %in% colnames(res)) as.numeric(res[est, "Std. Error"])
              else NA_real_
  cell(coef_val, se_val)
}

dml_rows <- rbind(
  data.frame(Estimator = "PLR-IV (DML)",
             `Cubic age`     = dml_cell(res_cub, "PLR-IV"),
             `Saturated age` = dml_cell(res_sat, "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (DML)",
             `Cubic age`     = dml_cell(res_cub, "Wald-AIPW"),
             `Saturated age` = dml_cell(res_sat, "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE)
)

ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")
kappa_rows <- do.call(rbind, lapply(seq_along(ests), function(i) {
  e <- ests[i]
  data.frame(
    Estimator = labels[i],
    `Cubic age`     = cell(kappa_cub$estimates[e], kappa_cub$se[e]),
    `Saturated age` = cell(kappa_sat$estimates[e], kappa_sat$se[e]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}))

rbind(dml_rows, kappa_rows) %>%
  kbl(align = c("l","c","c"),
      caption = "DML vs. Kappa — Point Estimates (SE), Y = log wages (dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Double ML", 1, 2) %>%
  pack_rows("Kappa — normalized", 3, 5) %>%
  pack_rows("Kappa — unnormalized", 6, 8) %>%
  row_spec(1:2, background = "#eef6ff") %>%
  row_spec(6:8, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE). Blue = DML. Orange = unnormalized kappa. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 3. Learner Comparison — Linear / Ranger / XGBoost

```{r ext-setup}
make_dml_data <- function(X_dml) {
  df <- as.data.frame(X_dml)
  df$y <- as.numeric(Y_dol)
  df$d <- as.numeric(D)
  df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X_dml))
}

dat_cub <- make_dml_data(X_dml_cub)
dat_sat <- make_dml_data(X_dml_sat)

cat("DoubleML data objects built for both specs.\n")
```

```{r ext-fit-helpers, echo=FALSE}
ml_lm  <- function() list(g = lrn("regr.lm"),
                          m = lrn("classif.log_reg"),
                          r = lrn("classif.log_reg"))
ml_rf  <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                          m = lrn("classif.ranger", keep.inbag = TRUE),
                          r = lrn("classif.ranger", keep.inbag = TRUE))
ml_xgb <- function() list(
  g = lrn("regr.xgboost",    alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
  m = lrn("classif.xgboost", alpha = 0, subsample = 1, max_delta_step = 0),
  r = lrn("classif.xgboost", alpha = 0, subsample = 1, max_delta_step = 0))

# Fit one IIVM model. Wrapped in tryCatch: the parametric (linear+logit) learner
# can fail to produce outcome weights in the saturated spec (documented in the
# thesis); on failure we return NA gracefully instead of aborting the knit.
fit_iivm <- function(dml_data, learners, Y) {
  tryCatch({
    set.seed(42)
    obj <- DoubleMLIIVM$new(data = dml_data,
                            ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                            n_folds = 5, score = "LATE")
    obj$fit(store_models = TRUE, store_predictions = TRUE)
    om  <- get_outcome_weights(object = obj, dml_data = dml_data)
    w   <- as.vector(om$omega)
    ok  <- check_doubleml_identity(om, Y, obj$coef)
    list(obj = obj, w = w, ok = ok)
  }, error = function(e) {
    message("fit_iivm failed: ", conditionMessage(e))
    list(obj = NULL, w = rep(NA_real_, length(Y)), ok = NA)
  })
}

safe_coef <- function(f) if (is.null(f$obj)) NA_real_ else get_dml_coef(f$obj)
safe_se   <- function(f) if (is.null(f$obj)) NA_real_ else get_dml_se(f$obj)

# Compact 2-spec estimate table for one learner.
learner_estimate_table <- function(f_cub, f_sat, cap) {
  data.frame(
    Spec      = c("Cubic age", "Saturated age"),
    Estimate  = round(c(safe_coef(f_cub), safe_coef(f_sat)), 4),
    SE        = round(c(safe_se(f_cub),   safe_se(f_sat)),   4),
    `ΣωY = τ̂` = c(f_cub$ok, f_sat$ok),
    check.names = FALSE
  ) %>%
    kbl(caption = cap, align = c("l","r","r","c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "ΣωY = τ̂: algebraic identity check (must be TRUE). NA = learner could not produce outcome weights for that spec.")
}
```

## Estimates {.tabset}

### Linear + Logistic

```{r ext-fit-lm, cache=TRUE}
lm_cub <- fit_iivm(dat_cub, ml_lm(), Y_dol)
lm_sat <- fit_iivm(dat_sat, ml_lm(), Y_dol)
learner_estimate_table(lm_cub, lm_sat, "Wald-AIPW — Linear + Logistic (both specs)")
```

### Ranger

```{r ext-fit-rf, cache=TRUE}
rf_cub <- fit_iivm(dat_cub, ml_rf(), Y_dol)
rf_sat <- fit_iivm(dat_sat, ml_rf(), Y_dol)
learner_estimate_table(rf_cub, rf_sat, "Wald-AIPW — Ranger (both specs)")
```

### XGBoost

```{r ext-fit-xgb, cache=TRUE}
xgb_cub <- fit_iivm(dat_cub, ml_xgb(), Y_dol)
xgb_sat <- fit_iivm(dat_sat, ml_xgb(), Y_dol)
learner_estimate_table(xgb_cub, xgb_sat, "Wald-AIPW — XGBoost (both specs)")
```

## Comparison Table

```{r ext-comparison}
ext_cell <- function(coef, se) {
  if (is.na(coef) || !is.finite(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

learner_row <- function(label, f_cub, f_sat) {
  data.frame(
    Estimator = label,
    `Cubic age`     = ext_cell(safe_coef(f_cub), safe_se(f_cub)),
    `Saturated age` = ext_cell(safe_coef(f_sat), safe_se(f_sat)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

kappa_ref_row <- function(label, key) {
  data.frame(
    Estimator = label,
    `Cubic age`     = ext_cell(kappa_cub$estimates[key], kappa_cub$se[key]),
    `Saturated age` = ext_cell(kappa_sat$estimates[key], kappa_sat$se[key]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

rbind(
  learner_row("Wald-AIPW (linear+logit)", lm_cub,  lm_sat),
  learner_row("Wald-AIPW (ranger)",       rf_cub,  rf_sat),
  learner_row("Wald-AIPW (XGBoost)",      xgb_cub, xgb_sat),
  kappa_ref_row("τ_cb_u (kappa, ref)",  "tau_cb_u"),
  kappa_ref_row("τ_ml_u (kappa, ref)",  "tau_ml_u")
) %>%
  kbl(align = c("l","c","c"),
      caption = "Wald-AIPW Learner Comparison — DoubleML vs. Kappa (Y = log wages, dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("DoubleML learners", 1, 3) %>%
  pack_rows("Kappa reference",   4, 5) %>%
  row_spec(1:3, background = "#eef6ff") %>%
  footnote(general = "Each cell: estimate (SE). Blue = DoubleML. All score='LATE', n_folds=5. '—' = learner could not produce outcome weights for that spec. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 4. Weight Diagnostics

<div class="note-box">
**Reading the tables.** `Sum_w ≈ 0` confirms translation invariance. **`ESS_mod`** = modified effective sample size `(Σ|ωᵢ|)²/Σωᵢ²` (Chattopadhyay & Zubizarreta 2023) — the headline number, valid under negative weights and bounded in `[1, n]`. The old Kish `ESS = 1/Σωᵢ²` is shown alongside for reference but is unreliable for signed weights. `Pct_neg` and `Max_abs_w` flag weight concentration. In the near-randomly-assigned Vietnam design the diagnostics are expected to be near-uniform across estimators and learners.
</div>

```{r diag-builders, echo=FALSE}
# kappa weight bundles for both specs.
kw_cub <- kappa_weights_bundle(Z, D, X_cub)
kw_sat <- kappa_weights_bundle(Z, D, X_sat)

# weight_diag() returns 6 columns (… ESS, ESS_mod …): align has 5 "r".
DIAG_ALIGN <- c("l","r","r","r","r","r")

dml_diag_cell <- function(wlist, kw) {
  rbind(
    weight_diag(wlist$plriv,    "PLR-IV     (DML)"),
    weight_diag(wlist$waldaipw, "Wald-AIPW  (DML)"),
    weight_diag(kw$tau_cb_u,    "τ_cb_u     (kappa)"),
    weight_diag(kw$tau_ml_u,    "τ_ml_u     (kappa)"),
    weight_diag(kw$tau_ml_a10,  "τ_ml_a10   (kappa)"),
    weight_diag(kw$tau_ml_a,    "τ_ml_a     (kappa)"),
    weight_diag(kw$tau_ml_a1,   "τ_ml_t/a1  (kappa)"),
    weight_diag(kw$tau_ml_a0,   "τ_ml_a0    (kappa)")
  )
}
style_dml_diag <- function(df, cap) {
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(1:2, background = "#eef6ff") %>%
    footnote(general = "Blue = DML. Sum_w ≈ 0 ⟺ translation invariant. ESS_mod = (Σ|ω|)²/Σω² (report this); ESS = 1/Σω² (reference only).")
}

ext_diag_block <- function(kw, w_lm, w_rf, w_xgb) {
  rbind(
    weight_diag(w_lm,          "Wald-AIPW (linear+logit)"),
    weight_diag(w_rf,          "Wald-AIPW (ranger)"),
    weight_diag(w_xgb,         "Wald-AIPW (XGBoost)"),
    weight_diag(kw$tau_cb_u,   "τ_cb_u    (kappa, CBPS)"),
    weight_diag(kw$tau_ml_u,   "τ_ml_u    (kappa, MLE)"),
    weight_diag(kw$tau_ml_a10, "τ_ml_a10  (kappa, MLE)"),
    weight_diag(kw$tau_ml_a,   "τ_ml_a    (kappa, unnorm)"),
    weight_diag(kw$tau_ml_a1,  "τ_ml_t    (kappa, unnorm)"),
    weight_diag(kw$tau_ml_a0,  "τ_ml_a0   (kappa, unnorm)")
  )
}
style_ext_diag <- function(df, cap) {
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(1:3, background = "#eef6ff") %>%
    footnote(general = "Blue = DoubleML learners. ESS_mod = (Σ|ω|)²/Σω² (report this); ESS = 1/Σω² (reference only). NA rows = learner produced no weights for that spec.")
}
```

## DML vs. Kappa {.tabset}

### Cubic age

```{r dml-diag-cub}
style_dml_diag(dml_diag_cell(w_cub, kw_cub), "Weight Diagnostics — DML vs. Kappa | Cubic age")
```

### Saturated age

```{r dml-diag-sat}
style_dml_diag(dml_diag_cell(w_sat, kw_sat), "Weight Diagnostics — DML vs. Kappa | Saturated age")
```

## Learners vs. Kappa {.tabset}

### Cubic age

```{r ext-diag-cub}
style_ext_diag(ext_diag_block(kw_cub, lm_cub$w, rf_cub$w, xgb_cub$w),
               "Weight Diagnostics — Learners | Cubic age")
```

### Saturated age

```{r ext-diag-sat}
style_ext_diag(ext_diag_block(kw_sat, lm_sat$w, rf_sat$w, xgb_sat$w),
               "Weight Diagnostics — Learners | Saturated age")
```

## Weight-Statistics Bridge — RQ1 {.tabset}

<div class="note-box">
**What this is.** For every estimator, `weight_bridge()` reports the outcome weights **Overall / Treated / Control** (control sign-flipped, matching Knaus's convention), putting the two diagnostic vocabularies side by side: his six `summary()` statistics next to the two effective-sample-size measures (Kish vs. modified). The Knaus block is computed in plain R, so the kappa estimators get his statistics too.
</div>

```{r bridge-helpers, echo=FALSE}
# Named list of weight vectors per cell; drop any all-NA (failed) learner.
cell_weight_list <- function(wdml, lm_w, rf_w, xgb_w, kw) {
  lst <- list(
    "PLR-IV (DML)"     = wdml$plriv,
    "Wald-AIPW (DML)"  = wdml$waldaipw,
    "Linear+Logit"     = lm_w,
    "Ranger"           = rf_w,
    "XGBoost"          = xgb_w,
    "τ_cb_u (kappa)"   = kw$tau_cb_u,
    "τ_ml_u (kappa)"   = kw$tau_ml_u,
    "τ_ml_t (kappa)"   = kw$tau_ml_a1
  )
  Filter(function(w) !all(is.na(w)), lst)
}

bridge_cells <- list(
  "Cubic age"     = list(w = cell_weight_list(w_cub, lm_cub$w, rf_cub$w, xgb_cub$w, kw_cub)),
  "Saturated age" = list(w = cell_weight_list(w_sat, lm_sat$w, rf_sat$w, xgb_sat$w, kw_sat))
)

bridge_table <- function(wlist, cap) {
  ests <- names(wlist)
  df   <- do.call(rbind, lapply(ests, function(nm) weight_bridge(wlist[[nm]], D, nm)))
  disp <- df[, c("Scope","Min","Max","Pct_neg","Sum_top10","Sum_w","Sum_abs_w","ESS_kish","ESS_mod")]

  kk <- disp %>%
    kbl(caption = cap, row.names = FALSE,
        col.names = c("Scope","Min","Max","%Neg","ΣTop10%","Σω","Σ|ω|","ESS Kish","ESS mod"),
        align = c("l", rep("r", 8))) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 12) %>%
    add_header_above(c(" " = 1, "Knaus summary() statistics" = 6,
                       "Effective sample size" = 2)) %>%
    column_spec(9, bold = TRUE, background = "#eef6ff")
  for (i in seq_along(ests)) kk <- pack_rows(kk, ests[i], (i - 1) * 3 + 1, i * 3)
  kk %>% footnote(general = "Control rows use Knaus's sign-flipped −ω convention. ESS mod = (Σ|ω|)²/Σω² is the number to report; ESS Kish = 1/Σω² is reference only.")
}

ess_overview_df <- do.call(rbind, lapply(names(bridge_cells), function(cn) {
  cl <- bridge_cells[[cn]]
  do.call(rbind, lapply(names(cl$w), function(nm) {
    ov <- weight_bridge(cl$w[[nm]], D, nm)
    ov <- ov[ov$Scope == "Overall", ]
    data.frame(Cell = cn, Estimator = nm,
               ESS_kish = ov$ESS_kish, ESS_mod = ov$ESS_mod,
               check.names = FALSE, stringsAsFactors = FALSE)
  }))
}))
```

### ESS overview (Kish vs. modified)

```{r bridge-ess-overview}
ess_overview_df %>%
  kbl(caption = "Effective sample size, Overall scope — Kish (1/Σω²) vs. modified ((Σ|ω|)²/Σω²)",
      col.names = c("Cell","Estimator","ESS Kish","ESS mod"),
      align = c("l","l","r","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  column_spec(4, bold = TRUE, background = "#eef6ff") %>%
  pack_rows(index = table(factor(ess_overview_df$Cell, levels = unique(ess_overview_df$Cell)))) %>%
  footnote(general = "Where Kish collapses toward 1, the modified ESS recovers the genuine usable weight mass. Report the modified column.")
```

```{r bridge-ess-dumbbell, fig.height=5.5, fig.width=11}
ggplot(ess_overview_df, aes(y = Estimator)) +
  geom_segment(aes(x = ESS_kish, xend = ESS_mod, yend = Estimator),
               colour = "#bdbdbd", linewidth = 1.1) +
  geom_point(aes(x = ESS_kish, colour = "Kish"),     size = 2.6) +
  geom_point(aes(x = ESS_mod,  colour = "Modified"), size = 2.6) +
  facet_wrap(~ Cell, scales = "free_x") +
  scale_colour_manual(values = c("Kish" = "#9e9e9e", "Modified" = "#2c7bb6"), name = NULL) +
  labs(title = "Effective sample size: Kish → modified",
       x = "Effective sample size", y = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.minor = element_blank())
```

### Cubic age

```{r bridge-cub}
bridge_table(bridge_cells[["Cubic age"]]$w, "Weight-statistics bridge — Cubic age")
```

### Saturated age

```{r bridge-sat}
bridge_table(bridge_cells[["Saturated age"]]$w, "Weight-statistics bridge — Saturated age")
```

---

# 5. Love Plots

```{r love-helpers, echo=FALSE}
# Return NULL (skipped) if the weight vector is all-NA, so grid.arrange never
# receives a failed learner's weights.
maybe_love <- function(title, w, D, X) if (all(is.na(w))) NULL else make_love(title, w, D, X)
```

## DML vs. Kappa {.tabset}

### Cubic age (all 8 estimators)

```{r love-cub, fig.height=9, fig.width=13}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV (DML)",     w_cub$plriv,       D, X_dml_cub),
  maybe_love("Wald-AIPW (DML)",  w_cub$waldaipw,    D, X_dml_cub),
  maybe_love("τ_cb_u (CBPS)",    kw_cub$tau_cb_u,   D, X_dml_cub),
  maybe_love("τ_ml_u",           kw_cub$tau_ml_u,   D, X_dml_cub),
  maybe_love("τ_ml_a10",         kw_cub$tau_ml_a10, D, X_dml_cub),
  maybe_love("τ_ml_a (unnorm)",  kw_cub$tau_ml_a,   D, X_dml_cub),
  maybe_love("τ_ml_t (unnorm)",  kw_cub$tau_ml_a1,  D, X_dml_cub),
  maybe_love("τ_ml_a0 (unnorm)", kw_cub$tau_ml_a0,  D, X_dml_cub)
))
grid.arrange(grobs = grobs, nrow = 2, top = "Love Plots — Cubic age: DML vs. Kappa")
```

### Saturated age (all 8 estimators)

```{r love-sat, fig.height=9, fig.width=13}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV (DML)",     w_sat$plriv,       D, X_dml_sat),
  maybe_love("Wald-AIPW (DML)",  w_sat$waldaipw,    D, X_dml_sat),
  maybe_love("τ_cb_u (CBPS)",    kw_sat$tau_cb_u,   D, X_dml_sat),
  maybe_love("τ_ml_u",           kw_sat$tau_ml_u,   D, X_dml_sat),
  maybe_love("τ_ml_a10",         kw_sat$tau_ml_a10, D, X_dml_sat),
  maybe_love("τ_ml_a (unnorm)",  kw_sat$tau_ml_a,   D, X_dml_sat),
  maybe_love("τ_ml_t (unnorm)",  kw_sat$tau_ml_a1,  D, X_dml_sat),
  maybe_love("τ_ml_a0 (unnorm)", kw_sat$tau_ml_a0,  D, X_dml_sat)
))
grid.arrange(grobs = grobs, nrow = 2, top = "Love Plots — Saturated age: DML vs. Kappa")
```

## Learners vs. Kappa {.tabset}

### Cubic age (3 learners + 2 kappa)

```{r ext-love-cub, fig.height=5.5, fig.width=14}
grobs <- Filter(Negate(is.null), list(
  maybe_love("Linear+Logit\nCubic", lm_cub$w,        D, X_dml_cub),
  maybe_love("Ranger\nCubic",       rf_cub$w,        D, X_dml_cub),
  maybe_love("XGBoost\nCubic",      xgb_cub$w,       D, X_dml_cub),
  maybe_love("τ_cb_u\nCubic",       kw_cub$tau_cb_u, D, X_dml_cub),
  maybe_love("τ_ml_u\nCubic",       kw_cub$tau_ml_u, D, X_dml_cub)
))
grid.arrange(grobs = grobs, nrow = 1, top = "Love Plots — DoubleML learners vs. kappa | Cubic age")
```

### Saturated age (3 learners + 2 kappa)

```{r ext-love-sat, fig.height=5.5, fig.width=14}
grobs <- Filter(Negate(is.null), list(
  maybe_love("Linear+Logit\nSaturated", lm_sat$w,        D, X_dml_sat),
  maybe_love("Ranger\nSaturated",       rf_sat$w,        D, X_dml_sat),
  maybe_love("XGBoost\nSaturated",      xgb_sat$w,       D, X_dml_sat),
  maybe_love("τ_cb_u\nSaturated",       kw_sat$tau_cb_u, D, X_dml_sat),
  maybe_love("τ_ml_u\nSaturated",       kw_sat$tau_ml_u, D, X_dml_sat)
))
grid.arrange(grobs = grobs, nrow = 1, top = "Love Plots — DoubleML learners vs. kappa | Saturated age")
```
