---
title: "Card (1995) — XGBoost Nuisance Tuning for PLR-IV"
subtitle: "Nested nuisance tuning by predictive loss, tuned inside each outer cross-fitting fold"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      cache = FALSE,
                      fig.width = 11, fig.height = 5.5)
```

```{r packages, include=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
library(haven)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

source("functions_all.R")

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

SEED            <- 42
N_CF            <- 5
TUNE_EVALS      <- 15
INNER_FOLDS     <- 3
EXTRACT_WEIGHTS <- TRUE
RERUN_TUNING    <- FALSE
EXPORT_PATH     <- "card_xgb_pliv_tuning_export.rds"
```

<div class="note-box">
**Purpose.** This notebook estimates PLR-IV with untuned and
fold-specifically tuned XGBoost nuisances for the four Card design cells.
Ordinary renders read the verified export. Set `RERUN_TUNING = TRUE` only
when deliberately re-estimating the nested tuning procedure.
</div>

<div class="warn-box">
**Runtime.** A substantive rerun tunes the three regression nuisances
separately within every outer training fold and replaces
`card_xgb_pliv_tuning_export.rds`.
</div>

---

# Setup

```{r data}
card <- read_dta("card.dta")
card <- card[!is.na(card$wage) & !is.na(card$educ), ]
if (nrow(card) != 3010L) stop("Expected the Card analysis sample to contain 3,010 observations.")

card$lwage   <- log(card$wage / 100)
card$somecol <- as.integer(card$educ >= 13)
card$educ16  <- as.integer(card$educ >= 16)
card$Z_inst  <- as.integer(card$nearc4)
if (!"expersq" %in% names(card)) card$expersq <- card$exper^2

X_card <- make_X(~ 0 + exper + expersq + black + south + smsa + smsa66 +
                   reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
                 data = card)
X_kit  <- make_X(~ 0 + black + south + smsa + smsa66 + south66, data = card)

data.frame(
  Quantity = c("Analysis sample", "Card covariates",
               "Kitagawa covariates", "Treatment definitions"),
  Value = c(format(nrow(card), big.mark = ","), ncol(X_card), ncol(X_kit), 2)
) %>%
  kbl(caption = "Analysis dimensions") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE)
```

```{r learners-and-tuning-setup, include=FALSE}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X))
}

make_xgb_pliv_learners <- function() {
  base <- list(alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0)
  list(
    l = do.call(lrn, c(list("regr.xgboost"), base)),
    m = do.call(lrn, c(list("regr.xgboost"), base)),
    r = do.call(lrn, c(list("regr.xgboost"), base))
  )
}

xgb_ps <- function() {
  ps(nrounds = p_int(50, 400), max_depth = p_int(2, 5),
     eta = p_dbl(0.02, 0.20), min_child_weight = p_int(1, 8))
}

tune_param_set <- function() list(ml_l = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())

tune_settings <- function() {
  list(
    n_folds_tune = INNER_FOLDS,
    rsmp_tune    = rsmp("cv", folds = INNER_FOLDS),
    measure      = list(ml_l = "regr.rmse", ml_m = "regr.rmse", ml_r = "regr.rmse"),
    terminator   = trm("evals", n_evals = TUNE_EVALS),
    algorithm    = tnr("random_search")
  )
}

get_prediction_vector <- function(obj, learner_name) {
  pred <- tryCatch(obj$predictions[[learner_name]], error = function(e) NULL)
  if (is.null(pred)) return(numeric(0))
  as.numeric(drop(pred))
}

rmse_loss <- function(truth, pred) {
  if (length(pred) != length(truth) || length(pred) == 0) return(NA_real_)
  truth <- as.numeric(truth); pred <- as.numeric(pred)
  keep <- is.finite(truth) & is.finite(pred)
  if (!any(keep)) return(NA_real_)
  sqrt(mean((truth[keep] - pred[keep])^2))
}

calc_nuisance_losses <- function(r, cell) {
  if (is.null(r)) return(c(outcome_l_rmse = NA_real_, instrument_m_rmse = NA_real_, treatment_r_rmse = NA_real_))
  l_hat <- get_prediction_vector(r$obj, "ml_l")
  m_hat <- get_prediction_vector(r$obj, "ml_m")
  r_hat <- get_prediction_vector(r$obj, "ml_r")
  c(outcome_l_rmse    = rmse_loss(cell$Y, l_hat),
    instrument_m_rmse = rmse_loss(cell$Z, m_hat),
    treatment_r_rmse  = rmse_loss(cell$D, r_hat))
}

omega_vec <- function(om) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && "PLR-IV" %in% rownames(w)) return(as.vector(w["PLR-IV", ]))
  as.vector(w[1, ])
}

extract_xgb_w <- function(r, cell) {
  if (is.null(r) || is.null(r$obj)) {
    stop("Cannot extract weights because the fitted DoubleML object is NULL.")
  }
  dml_data <- make_dml_data(cell$Y, cell$D, cell$Z, cell$X)
  om <- get_outcome_weights(object = r$obj, dml_data = dml_data)
  omega_vec(om)
}

safe_extract_xgb_w <- function(r, cell) {
  tryCatch(extract_xgb_w(r, cell), error = function(e) NULL)
}

fit_cell <- function(dml_data, learners, label, tune = FALSE, tune_on_folds = FALSE) {
  tryCatch({
    set.seed(SEED)
    obj <- DoubleMLPLIV$new(data = dml_data,
                            ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
                            n_folds = N_CF)
    if (tune) {
      obj$tune(param_set = tune_param_set(), tune_settings = tune_settings(),
               tune_on_folds = tune_on_folds)
    }
    obj$fit(store_models = TRUE, store_predictions = TRUE)

    list(label = label, tuned = tune,
         coef = get_dml_coef(obj), se = get_dml_se(obj),
         tuning_result = tryCatch(obj$tuning_result, error = function(e) NULL),
         params        = tryCatch(obj$params,        error = function(e) NULL),
         obj = obj)
  }, error = function(e) {
    message(sprintf("Fit failed for %s: %s", label, conditionMessage(e)))
    NULL
  })
}

cells <- list(
  "Card / some college"           = list(Y = card$lwage, D = card$somecol, Z = card$Z_inst, X = X_card),
  "Card / college completion"     = list(Y = card$lwage, D = card$educ16,  Z = card$Z_inst, X = X_card),
  "Kitagawa / some college"       = list(Y = card$lwage, D = card$somecol, Z = card$Z_inst, X = X_kit),
  "Kitagawa / college completion" = list(Y = card$lwage, D = card$educ16, Z = card$Z_inst, X = X_kit)
)
```

---

# Fit: untuned versus tuned XGBoost

```{r fit-all, include=FALSE}
if (RERUN_TUNING) {
  results <- list()
  for (cell_name in names(cells)) {
    cell <- cells[[cell_name]]
    dml_data <- make_dml_data(cell$Y, cell$D, cell$Z, cell$X)

    results[[cell_name]] <- list(
      untuned = fit_cell(dml_data, make_xgb_pliv_learners(),
                         paste(cell_name, "pliv-xgb-untuned"),
                         tune = FALSE),
      tuned   = fit_cell(dml_data, make_xgb_pliv_learners(),
                         paste(cell_name, "pliv-xgb-tuned"),
                         tune = TRUE, tune_on_folds = TRUE)
    )
  }
} else {
  if (!file.exists(EXPORT_PATH)) {
    stop("Verified PLR-IV XGBoost export not found. Set RERUN_TUNING=TRUE to rebuild it.")
  }
  stored_export <- readRDS(EXPORT_PATH)
}
```

## 1. Estimate comparison

```{r comparison-table}
fmt_est <- function(r) if (is.null(r)) "\u2014" else sprintf("%.3f (%.3f)", r$coef, r$se)

if (RERUN_TUNING) {
  rows <- do.call(rbind, lapply(names(results), function(cell_name) {
    R <- results[[cell_name]]
    data.frame(
      Cell = cell_name,
      Learner = c("PLR-IV XGB untuned",
                  "PLR-IV XGB tuned (tune_on_folds=TRUE)"),
      `Estimate (SE)` = c(fmt_est(R$untuned), fmt_est(R$tuned)),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }))
} else {
  rows <- stored_export$estimate_comparison
}

rows %>%
  kbl(align = c("l","l","c"),
      caption = "Card PLR-IV: XGBoost untuned vs. fold-specifically tuned") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(rows$Learner == "PLR-IV XGB tuned (tune_on_folds=TRUE)"), background = "#eef6ff") %>%
  footnote(general = "Tuning is selected by nuisance-prediction loss within each outer training fold.")
```

## 2. Nuisance-prediction losses

```{r nuisance-loss-table}
if (RERUN_TUNING) {
  loss_rows <- do.call(rbind, lapply(names(results), function(cell_name) {
    R <- results[[cell_name]]; cell <- cells[[cell_name]]
    ul <- calc_nuisance_losses(R$untuned, cell)
    tl <- calc_nuisance_losses(R$tuned, cell)
    data.frame(
      Cell = cell_name,
      Learner = c("PLR-IV XGB untuned", "PLR-IV XGB tuned"),
      `Outcome l RMSE` = c(ul["outcome_l_rmse"], tl["outcome_l_rmse"]),
      `Instrument m RMSE` = c(ul["instrument_m_rmse"], tl["instrument_m_rmse"]),
      `Treatment r RMSE` = c(ul["treatment_r_rmse"], tl["treatment_r_rmse"]),
      check.names = FALSE, row.names = NULL
    )
  }))
} else {
  loss_rows <- stored_export$nuisance_losses
}

loss_rows %>%
  kbl(digits = 4, caption = "PLR-IV nuisance-prediction loss: untuned vs. tuned (lower is better)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(loss_rows$Learner == "PLR-IV XGB tuned"), background = "#eef6ff")
```

## 3. Tuning protocol

The tuning design, rather than a single parameter vector, is the relevant
reproducibility information because selection occurs separately within each
outer training fold.

```{r tuning-protocol}
data.frame(
  Component = c(
    "Outer cross-fitting", "Inner validation", "Search algorithm", "Budget",
    "nrounds", "max_depth", "eta", "min_child_weight",
    "Outcome nuisance", "Instrument nuisance", "Treatment nuisance"
  ),
  Specification = c(
    paste(N_CF, "folds"), paste(INNER_FOLDS, "folds"), "Random search",
    paste(TUNE_EVALS, "evaluations per nuisance and outer fold"),
    "50–400", "2–5", "0.02–0.20", "1–8",
    "RMSE", "RMSE", "RMSE"
  )
) %>%
  kbl(caption = "Nested XGBoost tuning protocol: PLR-IV") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

```{r hyperparameter-summary, include=FALSE}
extract_tuned_params <- function(r, cell_name) {
  if (is.null(r) || !r$tuned) return(data.frame())
  src <- if (!is.null(r$params)) r$params else r$tuning_result
  if (is.null(src)) return(data.frame())

  u <- unlist(src, recursive = TRUE)
  u <- u[grepl("nrounds|max_depth|eta|min_child_weight", names(u))]
  if (length(u) == 0) return(data.frame())

  hp <- data.frame(Cell = cell_name, Raw_name = names(u), Value = as.numeric(u), stringsAsFactors = FALSE)
  hp$Nuisance <- ifelse(grepl("ml_l", hp$Raw_name), "Outcome l",
                  ifelse(grepl("ml_m", hp$Raw_name), "Instrument m",
                  ifelse(grepl("ml_r", hp$Raw_name), "Treatment r", "Unknown")))
  hp$Parameter <- sub(".*(nrounds|max_depth|eta|min_child_weight).*", "\\1", hp$Raw_name)
  hp
}

if (RERUN_TUNING) {
  hp_parts <- lapply(names(results), function(cn) {
    extract_tuned_params(results[[cn]]$tuned, cn)
  })
  hp_parts <- Filter(function(x) nrow(x) > 0, hp_parts)
  hp_raw <- if (length(hp_parts)) do.call(rbind, hp_parts) else data.frame()
  if (!nrow(hp_raw)) stop("Fold-specific hyperparameters could not be extracted.")
  hp_raw <- hp_raw[hp_raw$Nuisance != "Unknown", ]
  hp_summary <- hp_raw %>%
    group_by(Cell, Nuisance, Parameter) %>%
    summarise(Median_across_folds = median(Value), N_values_seen = n(), .groups = "drop") %>%
    arrange(Cell, Nuisance, Parameter)
} else {
  hp_raw <- NULL
  hp_summary <- stored_export$hyperparameters
}
```

## 4. Outcome-weight diagnostics and export

```{r weight-helpers}
one_xgb_diag_row <- function(cell_name, learner_name, r, cell, w) {
  if (is.null(r) || is.null(r$obj)) {
    return(data.frame(
      Cell = cell_name, Learner = learner_name,
      Estimate = NA_real_, ESS_mod = NA_real_,
      `Sigma omega` = NA_real_, `Sigma abs omega` = NA_real_,
      `Treated mass` = NA_real_,
      `Control mass (-sum control weights)` = NA_real_,
      `max abs omega` = NA_real_,
      `ω′Y reproduces τ̂` = NA,
      `Weight extraction` = "fit failed",
      check.names = FALSE
    ))
  }
  if (is.null(w)) {
    return(data.frame(
      Cell = cell_name, Learner = learner_name,
      Estimate = as.numeric(r$coef), ESS_mod = NA_real_,
      `Sigma omega` = NA_real_, `Sigma abs omega` = NA_real_,
      `Treated mass` = NA_real_,
      `Control mass (-sum control weights)` = NA_real_,
      `max abs omega` = NA_real_,
      `ω′Y reproduces τ̂` = NA,
      `Weight extraction` = "failed",
      check.names = FALSE
    ))
  }

  wb <- weight_bridge(w, cell$D, learner_name)
  ov <- wb[wb$Scope == "Overall", ]
  tr <- wb[wb$Scope == "Treated", ]
  co <- wb[wb$Scope == "Control", ]
  identity_ok <- tryCatch(
    check_weight_identity(w, cell$Y, r$coef),
    error = function(e) NA
  )

  data.frame(
    Cell = cell_name, Learner = learner_name,
    Estimate = as.numeric(r$coef), ESS_mod = ov$ESS_mod,
    `Sigma omega` = ov$Sum_w, `Sigma abs omega` = ov$Sum_abs_w,
    `Treated mass` = tr$Sum_w,
    `Control mass (-sum control weights)` = co$Sum_w,
    `max abs omega` = max(abs(w)),
    `ω′Y reproduces τ̂` = identity_ok,
    `Weight extraction` = "available",
    check.names = FALSE
  )
}
```

```{r weight-table}
if (RERUN_TUNING) {
  xgb_pliv_omega_vectors <- lapply(names(cells), function(cell_name) {
    list(
      untuned = if (EXTRACT_WEIGHTS) {
        safe_extract_xgb_w(results[[cell_name]]$untuned, cells[[cell_name]])
      } else NULL,
      tuned = if (EXTRACT_WEIGHTS) {
        safe_extract_xgb_w(results[[cell_name]]$tuned, cells[[cell_name]])
      } else NULL
    )
  })
  names(xgb_pliv_omega_vectors) <- names(cells)

  xgb_pliv_weight_table <- do.call(rbind, lapply(names(cells), function(cell_name) {
    cell <- cells[[cell_name]]
    R <- results[[cell_name]]
    W <- xgb_pliv_omega_vectors[[cell_name]]
    rbind(
      one_xgb_diag_row(cell_name, "PLR-IV XGB untuned",
                       R$untuned, cell, W$untuned),
      one_xgb_diag_row(cell_name, "PLR-IV XGB tuned",
                       R$tuned, cell, W$tuned)
    )
  }))
} else {
  xgb_pliv_omega_vectors <- stored_export$omega_vectors
  xgb_pliv_weight_table <- stored_export$weight_diagnostics
}

xgb_pliv_weight_table <- xgb_pliv_weight_table[
  , setdiff(names(xgb_pliv_weight_table), "ESS"), drop = FALSE
]
if ("Identity ok" %in% names(xgb_pliv_weight_table)) {
  names(xgb_pliv_weight_table)[
    names(xgb_pliv_weight_table) == "Identity ok"
  ] <- "ω′Y reproduces τ̂"
}
if ("Weight status" %in% names(xgb_pliv_weight_table)) {
  xgb_pliv_weight_table[["Weight status"]] <- ifelse(
    xgb_pliv_weight_table[["Weight status"]] == "ok",
    "available",
    xgb_pliv_weight_table[["Weight status"]]
  )
  names(xgb_pliv_weight_table)[
    names(xgb_pliv_weight_table) == "Weight status"
  ] <- "Weight extraction"
}

xgb_pliv_weight_table %>%
  kbl(digits = 4, caption = "Card PLR-IV XGBoost outcome-weight diagnostics") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12) %>%
  row_spec(which(xgb_pliv_weight_table$Learner == "PLR-IV XGB tuned"),
           background = "#eef6ff") %>%
  footnote(general = "Weight extraction records whether a finite reconstructed weight vector is available. It is distinct from the numerical reproduction check. Modified ESS is reported for signed outcome weights; conventional Kish ESS is omitted.")
```

```{r omega-and-identities, include=FALSE}
if (RERUN_TUNING) {
  xgb_pliv_identity_checks <- lapply(names(cells), function(cell_name) {
    cell <- cells[[cell_name]]
    R <- results[[cell_name]]
    W <- xgb_pliv_omega_vectors[[cell_name]]
    list(
      untuned = tryCatch(
        if (is.null(W$untuned) || is.null(R$untuned)) NA else
          check_weight_identity(W$untuned, cell$Y, R$untuned$coef),
        error = function(e) NA
      ),
      tuned = tryCatch(
        if (is.null(W$tuned) || is.null(R$tuned)) NA else
          check_weight_identity(W$tuned, cell$Y, R$tuned$coef),
        error = function(e) NA
      )
    )
  })
  names(xgb_pliv_identity_checks) <- names(cells)
} else {
  xgb_pliv_identity_checks <- stored_export$identity_checks
}
```

```{r export, include=FALSE}
if (RERUN_TUNING) {
  xgb_pliv_export <- list(
    estimate_comparison = rows,
    fit_summaries = lapply(results, function(x) lapply(x, function(r) {
      if (is.null(r)) NULL else {
        list(coef = as.numeric(r$coef), se = as.numeric(r$se))
      }
    })),
    nuisance_losses = loss_rows,
    hyperparameters = hp_summary,
    hyperparameters_raw = hp_raw,
    weight_diagnostics = xgb_pliv_weight_table,
    omega_vectors = xgb_pliv_omega_vectors,
    identity_checks = xgb_pliv_identity_checks,
    settings = list(
      seed = SEED, n_cf = N_CF, tune_evals = TUNE_EVALS,
      inner_folds = INNER_FOLDS, tune_on_folds = TRUE,
      learner = "xgboost", dml_class = "DoubleMLPLIV",
      score = "partialling out", extract_weights = EXTRACT_WEIGHTS,
      cells = names(cells)
    ),
    sample_info = list(
      n = nrow(card), mean_Z = mean(card$Z_inst),
      somecol_mean_D = mean(card$somecol),
      educ16_mean_D = mean(card$educ16),
      cells = names(cells)
    )
  )
  saveRDS(xgb_pliv_export, EXPORT_PATH)
}
```

<div class="note-box">
The main Card analysis reads `card_xgb_pliv_tuning_export.rds`. Fold-specific
hyperparameters remain in the export for audit purposes and are not reported
as a reusable global specification.
</div>
