---
title: "Card (1995) — Replication and Outcome-Weight Diagnostics"
subtitle: "Kappa estimators, Double ML, learner sensitivity, and exact weight checks"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body   { font-size: 15px; color: #2c2c2c; }
h1     { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2     { color: #2c7bb6; margin-top: 28px; }
h3     { color: #444; margin-top: 20px; }

.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  cache   = TRUE,
  autodep = TRUE,
  fig.width  = 12,
  fig.height = 6
)
```

---

# Setup

## Packages & Functions

```{r packages-and-source, include=FALSE}
# Core
library(haven)
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(dplyr)
# DML smoother + balance plots
library(OutcomeWeights)
library(cobalt)
library(viridis)
library(gridExtra)
library(ggplot2)
# DoubleML learners
library(DoubleML)
library(mlr3verse)

source("functions_all.R")
```

## Data & Design

<div class="note-box">
**Setting.** Card (1995) NLSYM subsample (N = 3,010). Instrument Z = `nearc4` (grew up near a 4-year college). Two binary treatments: **somecol** = 1{educ > 12} and **educ16** = 1{educ ≥ 16}. Outcome: log wages, reported in two units (dollars / cents) only because SUW Table 3 is laid out that way. Two covariate specs: **Card** (exper, expersq, black, south, smsa, smsa66, reg661–668) and **Kitagawa** (black, south, smsa, smsa66, south66).
</div>

```{r data, include=FALSE}
card       <- read_dta("card.dta")

# Sample restriction: N = 3,010
card_clean <- card[!is.na(card$wage) & !is.na(card$educ), ]

# Outcomes in two units: lwage_cnt = lwage + log(100)
card_clean$lwage_cnt <- log(card_clean$wage)         # cents
card_clean$lwage     <- log(card_clean$wage / 100)   # dollars

# Binary treatments
card_clean$somecol <- as.integer(card_clean$educ > 12)
card_clean$educ16  <- as.integer(card_clean$educ >= 16)

# Instrument
card_clean$Z_inst <- as.integer(card_clean$nearc4)

# expersq if missing
if (!"expersq" %in% names(card_clean)) {
  card_clean$expersq <- card_clean$exper^2
}

```

```{r design-matrices, include=FALSE}
# Kappa design matrices — WITH intercept (for logit / CBPS)
X_card_kappa <- model.matrix(
  ~ exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card_clean
)
X_card_df <- as.data.frame(X_card_kappa[, -1, drop = FALSE])

X_kit_kappa <- model.matrix(
  ~ black + south + smsa + smsa66 + south66,
  data = card_clean
)
X_kit_df <- as.data.frame(X_kit_kappa[, -1, drop = FALSE])

# DML design matrices — no intercept, attributes stripped for grf (make_X)
X_dml_card <- make_X(
  ~ 0 + exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card_clean
)
X_dml_kit <- make_X(
  ~ 0 + black + south + smsa + smsa66 + south66,
  data = card_clean
)

# Core vectors
Y_cnt <- card_clean$lwage_cnt   # cents  (used only for the SUW Table 3 layout)
Y_dol <- card_clean$lwage       # dollars (the working outcome everywhere else)
Z     <- card_clean$Z_inst
D1    <- card_clean$somecol
D2    <- card_clean$educ16

```

# 0. Empirical Design Audit

<div class="note-box">
The compact design audit reports treatment uptake, conditional first-stage
strength, and fitted instrument-propensity support. Covariate balance is
examined later using estimator-specific outcome weights.
</div>

```{r design-audit, echo=FALSE}
first_stage_row <- function(label, D, X_df, X_kappa) {
  dat <- data.frame(D = D, Z = Z, X_df, check.names = FALSE)
  fit <- lm(D ~ ., data = dat)
  ct <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC1"))
  p <- logit_mle(Z, X_kappa)
  data.frame(
    Cell = label,
    `Adjusted first stage` = unname(ct["Z", "Estimate"]),
    `HC1 SE` = unname(ct["Z", "Std. Error"]),
    `Robust t²` = unname(ct["Z", "t value"]^2),
    `Min p̂(X)` = min(p),
    `Max p̂(X)` = max(p),
    `Boundary share` = mean(p < 0.05 | p > 0.95),
    check.names = FALSE
  )
}

sample_audit <- rbind(
  data.frame(
    Treatment = "somecol", N = length(D1), `Pr(Z=1)` = mean(Z),
    `Pr(D=1)` = mean(D1), `Pr(D=1 | Z=0)` = mean(D1[Z == 0]),
    `Pr(D=1 | Z=1)` = mean(D1[Z == 1]),
    `Raw uptake gap` = mean(D1[Z == 1]) - mean(D1[Z == 0]),
    check.names = FALSE
  ),
  data.frame(
    Treatment = "educ16", N = length(D2), `Pr(Z=1)` = mean(Z),
    `Pr(D=1)` = mean(D2), `Pr(D=1 | Z=0)` = mean(D2[Z == 0]),
    `Pr(D=1 | Z=1)` = mean(D2[Z == 1]),
    `Raw uptake gap` = mean(D2[Z == 1]) - mean(D2[Z == 0]),
    check.names = FALSE
  )
)

design_audit <- rbind(
  first_stage_row("Card / somecol", D1, X_card_df, X_card_kappa),
  first_stage_row("Card / educ16", D2, X_card_df, X_card_kappa),
  first_stage_row("Kitagawa / somecol", D1, X_kit_df, X_kit_kappa),
  first_stage_row("Kitagawa / educ16", D2, X_kit_df, X_kit_kappa)
)

sample_audit %>%
  kbl(digits = 4, caption = "Sample and unconditional treatment uptake") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13)

design_audit %>%
  kbl(digits = 4,
      caption = "Conditional first-stage strength and estimated instrument overlap") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) %>%
  footnote(general = "Robust t² is the squared HC1 t-statistic for college proximity. It is a descriptive strength diagnostic, not a formal weak-IV-robust test. Boundary share uses 0.05 and 0.95.")
```

---

# 1. Reproduction — SUW Table 3

<div class="note-box">
**Layout.** SUW Table 3 has 8 columns: 2 treatments × 2 specifications × 2 outcome units, split here into two tabs by treatment. Within each tab the four columns are spec × unit. Normalized estimators (τ_cb_u, τ_ml_u, τ_ml_a10) are translation invariant — cents and dollars columns match; unnormalized ones (τ_ml_a, τ_ml_t, τ_ml_a0) do not.
</div>

```{r build-columns, cache=TRUE}
# One full column = 2SLS + all six kappa estimators with analytical SEs.
run_card_column <- function(Y, D, X_kappa, X_df) {
  list(
    tsls  = run_2sls(Y, D, Z, X_df, endog_name = "D"),
    kappa = kappa_analytic_se_all(Y, Z, D, X_kappa)
  )
}

# Treatment 1: somecol (cols 1-4)
col1 <- run_card_column(Y_cnt, D1, X_card_kappa, X_card_df)  # cents,   Card
col2 <- run_card_column(Y_dol, D1, X_card_kappa, X_card_df)  # dollars, Card
col3 <- run_card_column(Y_cnt, D1, X_kit_kappa,  X_kit_df)   # cents,   Kitagawa
col4 <- run_card_column(Y_dol, D1, X_kit_kappa,  X_kit_df)   # dollars, Kitagawa

# Treatment 2: educ16 (cols 5-8)
col5 <- run_card_column(Y_cnt, D2, X_card_kappa, X_card_df)  # cents,   Card
col6 <- run_card_column(Y_dol, D2, X_card_kappa, X_card_df)  # dollars, Card
col7 <- run_card_column(Y_cnt, D2, X_kit_kappa,  X_kit_df)   # cents,   Kitagawa
col8 <- run_card_column(Y_dol, D2, X_kit_kappa,  X_kit_df)   # dollars, Kitagawa

```

```{r table-builder, echo=FALSE}
cell_fmt <- function(est, se, digits = 3) {
  if (is.na(est)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0)
    return(sprintf("%.3f<br><span style='color:#888'>(NA)</span>", est))
  pval  <- 2 * pnorm(-abs(est / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s<br><span style='color:#888'>(%.3f)</span>", est, stars, se)
}

build_treatment_table <- function(cA_cnt, cB_dol, cC_cnt, cD_dol) {
  ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
  labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")

  rows <- list(data.frame(
    Estimator = "2SLS",
    `Card / cents`       = cell_fmt(cA_cnt$tsls$coef, cA_cnt$tsls$se),
    `Card / dollars`     = cell_fmt(cB_dol$tsls$coef, cB_dol$tsls$se),
    `Kitagawa / cents`   = cell_fmt(cC_cnt$tsls$coef, cC_cnt$tsls$se),
    `Kitagawa / dollars` = cell_fmt(cD_dol$tsls$coef, cD_dol$tsls$se),
    check.names = FALSE, stringsAsFactors = FALSE
  ))

  for (i in seq_along(ests)) {
    e <- ests[i]
    rows[[length(rows) + 1]] <- data.frame(
      Estimator = labels[i],
      `Card / cents`       = cell_fmt(cA_cnt$kappa$estimates[e], cA_cnt$kappa$se[e]),
      `Card / dollars`     = cell_fmt(cB_dol$kappa$estimates[e], cB_dol$kappa$se[e]),
      `Kitagawa / cents`   = cell_fmt(cC_cnt$kappa$estimates[e], cC_cnt$kappa$se[e]),
      `Kitagawa / dollars` = cell_fmt(cD_dol$kappa$estimates[e], cD_dol$kappa$se[e]),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }
  do.call(rbind, rows)
}
```

## Results {.tabset}

### Treatment: Some College (educ > 12)

```{r table-somecol}
build_treatment_table(col1, col2, col3, col4) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c"),
      caption = "SUW Table 3 — Treatment: somecol (educ > 12). Effect of some college on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",           1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant). Orange = unnormalized. Card spec: exper, expersq, black, south, smsa, smsa66, reg661–668. Kitagawa spec: black, south, smsa, smsa66, south66. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

### Treatment: College Completion (educ ≥ 16)

```{r table-educ16}
build_treatment_table(col5, col6, col7, col8) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c"),
      caption = "SUW Table 3 — Treatment: educ16 (educ ≥ 16). Effect of college completion on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",           1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant). Orange = unnormalized. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 2. Double ML — PLR-IV & Wald-AIPW (honest forest, tuned vs. untuned)

<div class="note-box">
Each of the four treatment-by-covariate cells is fitted with an untuned honest
forest (`tune.parameters = "none"`) and a tuned honest forest
(`tune.parameters = "all"`), using five-fold cross-fitting. The tuned fits are
used in the downstream learner comparison. Untuned and tuned weights are both
retained for the direct forest comparison.
</div>

```{r dml-fit-all, include=FALSE}
# Four treatment-by-covariate cells.
set.seed(42); dml_card_d1_untuned <- dml_with_smoother(
  Y_dol, D1, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_card_d1_tuned <- dml_with_smoother(
  Y_dol, D1, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "all")

set.seed(42); dml_card_d2_untuned <- dml_with_smoother(
  Y_dol, D2, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_card_d2_tuned <- dml_with_smoother(
  Y_dol, D2, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "all")

set.seed(42); dml_kit_d1_untuned <- dml_with_smoother(
  Y_dol, D1, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_kit_d1_tuned <- dml_with_smoother(
  Y_dol, D1, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "all")

set.seed(42); dml_kit_d2_untuned <- dml_with_smoother(
  Y_dol, D2, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_kit_d2_tuned <- dml_with_smoother(
  Y_dol, D2, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "all")

res_card_d1_untuned <- summary(dml_card_d1_untuned)
res_card_d1_tuned   <- summary(dml_card_d1_tuned)
res_card_d2_untuned <- summary(dml_card_d2_untuned)
res_card_d2_tuned   <- summary(dml_card_d2_tuned)
res_kit_d1_untuned  <- summary(dml_kit_d1_untuned)
res_kit_d1_tuned    <- summary(dml_kit_d1_tuned)
res_kit_d2_untuned  <- summary(dml_kit_d2_untuned)
res_kit_d2_tuned    <- summary(dml_kit_d2_tuned)

# Extract PLR-IV and Wald-AIPW outcome weights for every fit.
extract_w <- function(dml_obj) {
  om <- get_outcome_weights(dml_obj)
  check_omega_rows(om)
  list(plriv    = as.vector(om$omega["PLR-IV",    ]),
       waldaipw = as.vector(om$omega["Wald-AIPW", ]))
}
w_card_d1_untuned <- extract_w(dml_card_d1_untuned)
w_card_d1_tuned   <- extract_w(dml_card_d1_tuned)
w_card_d2_untuned <- extract_w(dml_card_d2_untuned)
w_card_d2_tuned   <- extract_w(dml_card_d2_tuned)
w_kit_d1_untuned  <- extract_w(dml_kit_d1_untuned)
w_kit_d1_tuned    <- extract_w(dml_kit_d1_tuned)
w_kit_d2_untuned  <- extract_w(dml_kit_d2_untuned)
w_kit_d2_tuned    <- extract_w(dml_kit_d2_tuned)

# Headline aliases use tuned forests.
dml_card_d1 <- dml_card_d1_tuned
dml_card_d2 <- dml_card_d2_tuned
dml_kit_d1  <- dml_kit_d1_tuned
dml_kit_d2  <- dml_kit_d2_tuned
res_card_d1 <- res_card_d1_tuned
res_card_d2 <- res_card_d2_tuned
res_kit_d1  <- res_kit_d1_tuned
res_kit_d2  <- res_kit_d2_tuned
w_card_d1   <- w_card_d1_tuned
w_card_d2   <- w_card_d2_tuned
w_kit_d1    <- w_kit_d1_tuned
w_kit_d2    <- w_kit_d2_tuned
```

## Estimates {.tabset}

```{r dml-cell-display, echo=FALSE}
show_dml_cell <- function(res, wlist, Y, cap) {
  ids <- c(
    check_weight_identity(wlist$plriv,    Y, get_estimate(res, "PLR-IV")),
    check_weight_identity(wlist$waldaipw, Y, get_estimate(res, "Wald-AIPW"))
  )
  df <- as.data.frame(res[c("PLR-IV", "Wald-AIPW"), , drop = FALSE])
  df$Estimator  <- rownames(df)
  df$`ω′Y reproduces τ̂` <- as.character(ids)
  df <- df[, c("Estimator",
               setdiff(names(df), c("Estimator", "ω′Y reproduces τ̂")),
               "ω′Y reproduces τ̂")]
  df %>%
    kbl(caption = cap, digits = 4, row.names = FALSE,
        align = c("l", rep("r", ncol(df) - 2), "c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "Numerical reproduction follows Knaus's all.equal() check with R's default tolerance.")
}
```

### Card — somecol

```{r dml-card-d1-untuned}
show_dml_cell(res_card_d1_untuned, w_card_d1_untuned, Y_dol,
              "Card spec, somecol — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-card-d1-tuned}
show_dml_cell(res_card_d1_tuned, w_card_d1_tuned, Y_dol,
              "Card spec, somecol — tuned honest forest (tune.parameters = \"all\")")
```

### Card — educ16

```{r dml-card-d2-untuned}
show_dml_cell(res_card_d2_untuned, w_card_d2_untuned, Y_dol,
              "Card spec, educ16 — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-card-d2-tuned}
show_dml_cell(res_card_d2_tuned, w_card_d2_tuned, Y_dol,
              "Card spec, educ16 — tuned honest forest (tune.parameters = \"all\")")
```

### Kitagawa — somecol

```{r dml-kit-d1-untuned}
show_dml_cell(res_kit_d1_untuned, w_kit_d1_untuned, Y_dol,
              "Kitagawa spec, somecol — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-kit-d1-tuned}
show_dml_cell(res_kit_d1_tuned, w_kit_d1_tuned, Y_dol,
              "Kitagawa spec, somecol — tuned honest forest (tune.parameters = \"all\")")
```

### Kitagawa — educ16

```{r dml-kit-d2-untuned}
show_dml_cell(res_kit_d2_untuned, w_kit_d2_untuned, Y_dol,
              "Kitagawa spec, educ16 — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-kit-d2-tuned}
show_dml_cell(res_kit_d2_tuned, w_kit_d2_tuned, Y_dol,
              "Kitagawa spec, educ16 — tuned honest forest (tune.parameters = \"all\")")
```

<div class="note-box">
Within each cell, the untuned and tuned fits can be compared in their point
estimates, standard errors, and later weight geometry. Numerical reproduction
is not a tuning-quality or normalization criterion.
</div>

## DML vs. Kappa — Point Estimates

```{r dml-summary-table}
# Kappa estimates for all four cells.
kappa_card_d1 <- kappa_analytic_se_all(Y_dol, Z, D1, X_card_kappa)
kappa_card_d2 <- kappa_analytic_se_all(Y_dol, Z, D2, X_card_kappa)
kappa_kit_d1  <- kappa_analytic_se_all(Y_dol, Z, D1, X_kit_kappa)
kappa_kit_d2  <- kappa_analytic_se_all(Y_dol, Z, D2, X_kit_kappa)

cell <- function(coef, se) {
  if (is.na(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

dml_cell <- function(res, est) {
  coef_val <- get_estimate(res, est)
  se_val   <- if ("SE" %in% colnames(res)) as.numeric(res[est, "SE"])
              else if ("Std. Error" %in% colnames(res)) as.numeric(res[est, "Std. Error"])
              else NA_real_
  cell(coef_val, se_val)
}

dml_rows <- rbind(
  data.frame(Estimator = "PLR-IV (DML; tuned forest)",
             `Card / somecol`     = dml_cell(res_card_d1, "PLR-IV"),
             `Card / educ16`      = dml_cell(res_card_d2, "PLR-IV"),
             `Kitagawa / somecol` = dml_cell(res_kit_d1,  "PLR-IV"),
             `Kitagawa / educ16`  = dml_cell(res_kit_d2,  "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (DML; tuned forest)",
             `Card / somecol`     = dml_cell(res_card_d1, "Wald-AIPW"),
             `Card / educ16`      = dml_cell(res_card_d2, "Wald-AIPW"),
             `Kitagawa / somecol` = dml_cell(res_kit_d1,  "Wald-AIPW"),
             `Kitagawa / educ16`  = dml_cell(res_kit_d2,  "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE)
)

ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")
kappa_rows <- do.call(rbind, lapply(seq_along(ests), function(i) {
  e <- ests[i]
  data.frame(
    Estimator = labels[i],
    `Card / somecol`     = cell(kappa_card_d1$estimates[e], kappa_card_d1$se[e]),
    `Card / educ16`      = cell(kappa_card_d2$estimates[e], kappa_card_d2$se[e]),
    `Kitagawa / somecol` = cell(kappa_kit_d1$estimates[e],  kappa_kit_d1$se[e]),
    `Kitagawa / educ16`  = cell(kappa_kit_d2$estimates[e],  kappa_kit_d2$se[e]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}))

rbind(dml_rows, kappa_rows) %>%
  kbl(align = c("l","c","c","c","c"),
      caption = "DML vs. Kappa — Point Estimates (SE), Y = log wages (dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  add_header_above(c(" " = 1, "Card covariates" = 2, "Kitagawa covariates" = 2)) %>%
  pack_rows("Double ML", 1, 2) %>%
  pack_rows("Kappa — normalized", 3, 5) %>%
  pack_rows("Kappa — unnormalized", 6, 8) %>%
  row_spec(1:2, background = "#eef6ff") %>%
  row_spec(6:8, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE). Blue = DML. Orange = unnormalized kappa. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 3. Learner Comparison — PLR-IV and Wald-AIPW

<div class="note-box">
This section holds the five-fold DoubleML procedure and the four Card design
cells fixed while varying the nuisance learners. Linear and Ranger models are
fitted here. Untuned and fold-specifically tuned XGBoost results are imported
from the two tuning exports and are not retuned in this notebook.
</div>

```{r ext-setup, include=FALSE}
# DoubleMLData objects — one per cell. d and z must be numeric 0/1.
make_dml_data <- function(D, X_dml) {
  df <- as.data.frame(X_dml)
  df$y <- as.numeric(Y_dol)
  df$d <- as.numeric(D)
  df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X_dml))
}

dat_card_d1 <- make_dml_data(D1, X_dml_card)
dat_card_d2 <- make_dml_data(D2, X_dml_card)
dat_kit_d1  <- make_dml_data(D1, X_dml_kit)
dat_kit_d2  <- make_dml_data(D2, X_dml_kit)
```

```{r ext-fit-helpers, echo=FALSE}
ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))
ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))

omega_vec <- function(om, estimator) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && estimator %in% rownames(w)) {
    return(as.vector(w[estimator, ]))
  }
  as.vector(w[1, ])
}

fit_iivm <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLIIVM$new(data = dml_data,
                          ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5, score = "LATE")
  invisible(capture.output(obj$fit(store_models = TRUE, store_predictions = TRUE)))
  om  <- get_outcome_weights(object = obj, dml_data = dml_data)
  w   <- omega_vec(om, "Wald-AIPW")
  ok  <- check_doubleml_identity(om, Y, obj$coef)
  list(obj = obj, w = w, ok = ok)
}

# Fit one PLIV model with the same fold count and seed.
fit_pliv <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLPLIV$new(data = dml_data,
                          ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5)
  invisible(capture.output(obj$fit(store_models = TRUE, store_predictions = TRUE)))
  om <- get_outcome_weights(object = obj, dml_data = dml_data)
  w  <- omega_vec(om, "PLR-IV")
  coef <- NA_real_
  invisible(capture.output(coef <- get_dml_coef(obj)))
  ok <- check_weight_identity(w, Y, coef)
  list(obj = obj, w = w, ok = ok)
}

safe_coef <- function(f) {
  if (is.null(f$obj)) return(NA_real_)
  out <- NA_real_
  invisible(capture.output(out <- get_dml_coef(f$obj)))
  as.numeric(out)
}

safe_se <- function(f) {
  if (is.null(f$obj)) return(NA_real_)
  out <- NA_real_
  invisible(capture.output(out <- get_dml_se(f$obj)))
  as.numeric(out)
}

learner_estimate_table <- function(pliv, iivm, cap) {
  data.frame(
    Estimator = c("PLR-IV", "Wald-AIPW"),
    Estimate  = round(c(safe_coef(pliv), safe_coef(iivm)), 4),
    SE        = round(c(safe_se(pliv), safe_se(iivm)), 4),
    `ω′Y reproduces τ̂` = c(pliv$ok, iivm$ok),
    check.names = FALSE
  ) %>%
    kbl(caption = cap, align = c("l","r","r","c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "Numerical reproduction follows Knaus's all.equal() check with R's default tolerance.")
}
```

## Estimates {.tabset}

### Card — somecol

```{r ext-fit-card-d1, cache=TRUE}
lm_pliv_card_d1 <- fit_pliv(dat_card_d1, ml_lm_pliv(), Y_dol)
lm_iivm_card_d1 <- fit_iivm(dat_card_d1, ml_lm_iivm(), Y_dol)
rf_pliv_card_d1 <- fit_pliv(dat_card_d1, ml_rf_pliv(), Y_dol)
rf_iivm_card_d1 <- fit_iivm(dat_card_d1, ml_rf_iivm(), Y_dol)
learner_estimate_table(lm_pliv_card_d1, lm_iivm_card_d1,
                       "Card / somecol — Linear + Logistic")
learner_estimate_table(rf_pliv_card_d1, rf_iivm_card_d1,
                       "Card / somecol — Ranger")
```

### Card — educ16

```{r ext-fit-card-d2, cache=TRUE}
lm_pliv_card_d2 <- fit_pliv(dat_card_d2, ml_lm_pliv(), Y_dol)
lm_iivm_card_d2 <- fit_iivm(dat_card_d2, ml_lm_iivm(), Y_dol)
rf_pliv_card_d2 <- fit_pliv(dat_card_d2, ml_rf_pliv(), Y_dol)
rf_iivm_card_d2 <- fit_iivm(dat_card_d2, ml_rf_iivm(), Y_dol)
learner_estimate_table(lm_pliv_card_d2, lm_iivm_card_d2,
                       "Card / educ16 — Linear + Logistic")
learner_estimate_table(rf_pliv_card_d2, rf_iivm_card_d2,
                       "Card / educ16 — Ranger")
```

### Kitagawa — somecol

```{r ext-fit-kit-d1, cache=TRUE}
lm_pliv_kit_d1 <- fit_pliv(dat_kit_d1, ml_lm_pliv(), Y_dol)
lm_iivm_kit_d1 <- fit_iivm(dat_kit_d1, ml_lm_iivm(), Y_dol)
rf_pliv_kit_d1 <- fit_pliv(dat_kit_d1, ml_rf_pliv(), Y_dol)
rf_iivm_kit_d1 <- fit_iivm(dat_kit_d1, ml_rf_iivm(), Y_dol)
learner_estimate_table(lm_pliv_kit_d1, lm_iivm_kit_d1,
                       "Kitagawa / somecol — Linear + Logistic")
learner_estimate_table(rf_pliv_kit_d1, rf_iivm_kit_d1,
                       "Kitagawa / somecol — Ranger")
```

### Kitagawa — educ16

```{r ext-fit-kit-d2, cache=TRUE}
lm_pliv_kit_d2 <- fit_pliv(dat_kit_d2, ml_lm_pliv(), Y_dol)
lm_iivm_kit_d2 <- fit_iivm(dat_kit_d2, ml_lm_iivm(), Y_dol)
rf_pliv_kit_d2 <- fit_pliv(dat_kit_d2, ml_rf_pliv(), Y_dol)
rf_iivm_kit_d2 <- fit_iivm(dat_kit_d2, ml_rf_iivm(), Y_dol)
learner_estimate_table(lm_pliv_kit_d2, lm_iivm_kit_d2,
                       "Kitagawa / educ16 — Linear + Logistic")
learner_estimate_table(rf_pliv_kit_d2, rf_iivm_kit_d2,
                       "Kitagawa / educ16 — Ranger")
```

### XGBoost

```{r ext-xgb-import, include=FALSE}
validate_xgb_export <- function(x, expected_class) {
  required <- c("estimate_comparison", "nuisance_losses",
                "weight_diagnostics", "omega_vectors", "identity_checks",
                "settings", "sample_info")
  if (!all(required %in% names(x))) {
    stop("XGBoost export has an incomplete schema.")
  }
  if (!identical(x$settings$dml_class, expected_class)) {
    stop("XGBoost export has the wrong DoubleML class.")
  }
  if (!isTRUE(x$settings$tune_on_folds)) {
    stop("XGBoost export is not tune_on_folds=TRUE.")
  }
  required_cells <- c(
    "Card / some college", "Card / college completion",
    "Kitagawa / some college", "Kitagawa / college completion"
  )
  if (!all(required_cells %in% names(x$omega_vectors))) {
    stop("XGBoost export does not contain all four Card design cells.")
  }
  info <- x$sample_info
  if (!identical(as.integer(info$n), as.integer(length(Z))) ||
      !isTRUE(all.equal(as.numeric(info$mean_Z), mean(Z), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(info$somecol_mean_D), mean(D1), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(info$educ16_mean_D), mean(D2), tolerance = 1e-12))) {
    stop("XGBoost export metadata do not match the current Card data.")
  }
  invisible(TRUE)
}

xgb_wald_path <- "card_xgb_tuning_export.rds"
xgb_pliv_path <- "card_xgb_pliv_tuning_export.rds"
if (!file.exists(xgb_wald_path) || !file.exists(xgb_pliv_path)) {
  stop("Run both Card XGBoost tuning notebooks before card_test.Rmd.")
}
wald_xgb <- readRDS(xgb_wald_path)
pliv_xgb <- readRDS(xgb_pliv_path)
validate_xgb_export(wald_xgb, "DoubleMLIIVM")
validate_xgb_export(pliv_xgb, "DoubleMLPLIV")
```

```{r ext-xgb-export-helpers, echo=FALSE}
get_xgb_estimate_cell <- function(export, cell_name, variant) {
  tab <- export$estimate_comparison
  keep <- if (variant == "untuned") {
    grepl("untuned$", tab$Learner)
  } else {
    grepl("tune_on_folds=TRUE", tab$Learner, fixed = TRUE)
  }
  hit <- tab$Cell == cell_name & keep
  if (sum(hit) != 1L) return("—")
  as.character(tab[["Estimate (SE)"]][hit][1])
}

get_xgb_bundle <- function(export, cell_name, variant) {
  w <- tryCatch(export$omega_vectors[[cell_name]][[variant]],
                error = function(e) NULL)
  s <- tryCatch(export$fit_summaries[[cell_name]][[variant]],
                error = function(e) NULL)
  if (is.null(s)) {
    wd <- export$weight_diagnostics
    keep <- if (variant == "untuned") {
      grepl("untuned$", wd$Learner)
    } else {
      grepl(" tuned$", wd$Learner)
    }
    hit <- wd$Cell == cell_name & keep
    estimate <- if (sum(hit) == 1L) as.numeric(wd$Estimate[hit]) else NA_real_
    txt <- get_xgb_estimate_cell(export, cell_name, variant)
    nums <- regmatches(txt, gregexpr("-?[0-9]+\\.?[0-9]*", txt))[[1]]
    se <- if (length(nums) >= 2) as.numeric(nums[2]) else NA_real_
  } else {
    estimate <- as.numeric(s$coef)
    se <- as.numeric(s$se)
  }
  gap <- if (is.null(w) || !is.finite(estimate)) NA_real_ else
    sum(w * Y_dol) - estimate
  list(
    w = w, coef = estimate, se = se, gap = gap,
    ok = if (is.finite(gap)) check_weight_identity(w, Y_dol, estimate) else NA
  )
}

xgb_cells <- c(
  card_d1 = "Card / some college",
  card_d2 = "Card / college completion",
  kit_d1 = "Kitagawa / some college",
  kit_d2 = "Kitagawa / college completion"
)

xgb_card_d1_u <- get_xgb_bundle(wald_xgb, xgb_cells["card_d1"], "untuned")
xgb_card_d2_u <- get_xgb_bundle(wald_xgb, xgb_cells["card_d2"], "untuned")
xgb_kit_d1_u  <- get_xgb_bundle(wald_xgb, xgb_cells["kit_d1"], "untuned")
xgb_kit_d2_u  <- get_xgb_bundle(wald_xgb, xgb_cells["kit_d2"], "untuned")
xgb_card_d1 <- get_xgb_bundle(wald_xgb, xgb_cells["card_d1"], "tuned")
xgb_card_d2 <- get_xgb_bundle(wald_xgb, xgb_cells["card_d2"], "tuned")
xgb_kit_d1  <- get_xgb_bundle(wald_xgb, xgb_cells["kit_d1"], "tuned")
xgb_kit_d2  <- get_xgb_bundle(wald_xgb, xgb_cells["kit_d2"], "tuned")

xgb_pliv_card_d1_u <- get_xgb_bundle(pliv_xgb, xgb_cells["card_d1"], "untuned")
xgb_pliv_card_d2_u <- get_xgb_bundle(pliv_xgb, xgb_cells["card_d2"], "untuned")
xgb_pliv_kit_d1_u  <- get_xgb_bundle(pliv_xgb, xgb_cells["kit_d1"], "untuned")
xgb_pliv_kit_d2_u  <- get_xgb_bundle(pliv_xgb, xgb_cells["kit_d2"], "untuned")
xgb_pliv_card_d1 <- get_xgb_bundle(pliv_xgb, xgb_cells["card_d1"], "tuned")
xgb_pliv_card_d2 <- get_xgb_bundle(pliv_xgb, xgb_cells["card_d2"], "tuned")
xgb_pliv_kit_d1  <- get_xgb_bundle(pliv_xgb, xgb_cells["kit_d1"], "tuned")
xgb_pliv_kit_d2  <- get_xgb_bundle(pliv_xgb, xgb_cells["kit_d2"], "tuned")
```

```{r ext-xgb-comparison, echo=FALSE}
xgb_import_table <- do.call(rbind, lapply(list(
  list("PLR-IV", xgb_cells["card_d1"], pliv_xgb),
  list("PLR-IV", xgb_cells["card_d2"], pliv_xgb),
  list("PLR-IV", xgb_cells["kit_d1"], pliv_xgb),
  list("PLR-IV", xgb_cells["kit_d2"], pliv_xgb),
  list("Wald-AIPW", xgb_cells["card_d1"], wald_xgb),
  list("Wald-AIPW", xgb_cells["card_d2"], wald_xgb),
  list("Wald-AIPW", xgb_cells["kit_d1"], wald_xgb),
  list("Wald-AIPW", xgb_cells["kit_d2"], wald_xgb)
), function(x) {
  u <- get_xgb_bundle(x[[3]], x[[2]], "untuned")
  t <- get_xgb_bundle(x[[3]], x[[2]], "tuned")
  data.frame(
    Estimator = x[[1]], Cell = unname(x[[2]]),
    Untuned = get_xgb_estimate_cell(x[[3]], x[[2]], "untuned"),
    Tuned = get_xgb_estimate_cell(x[[3]], x[[2]], "tuned"),
    `ω′Y reproduces τ̂ (untuned / tuned)` = paste(
      ifelse(isTRUE(u$ok), "Yes", "No"),
      ifelse(isTRUE(t$ok), "Yes", "No"), sep = " / "
    ),
    `Absolute gap (untuned / tuned)` = sprintf(
      "%.2e / %.2e", abs(u$gap), abs(t$gap)
    ),
    check.names = FALSE
  )
}))

xgb_import_table %>%
  kbl(caption = "Imported XGBoost estimates: untuned and fold-specific tuning") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 12) %>%
  footnote(general = "Numerical reproduction follows Knaus's all.equal() check with R's default tolerance. It is separate from tuning quality and normalization.")
```

## Comparison Table

```{r ext-comparison}
ext_cell <- function(coef, se) {
  if (is.na(coef) || !is.finite(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

learner_row <- function(label, f_card_d1, f_card_d2, f_kit_d1, f_kit_d2) {
  data.frame(
    Estimator = label,
    `Card / somecol`     = ext_cell(safe_coef(f_card_d1), safe_se(f_card_d1)),
    `Card / educ16`      = ext_cell(safe_coef(f_card_d2), safe_se(f_card_d2)),
    `Kitagawa / somecol` = ext_cell(safe_coef(f_kit_d1),  safe_se(f_kit_d1)),
    `Kitagawa / educ16`  = ext_cell(safe_coef(f_kit_d2),  safe_se(f_kit_d2)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

kappa_ref_row <- function(label, key) {
  data.frame(
    Estimator = label,
    `Card / somecol`     = ext_cell(kappa_card_d1$estimates[key], kappa_card_d1$se[key]),
    `Card / educ16`      = ext_cell(kappa_card_d2$estimates[key], kappa_card_d2$se[key]),
    `Kitagawa / somecol` = ext_cell(kappa_kit_d1$estimates[key],  kappa_kit_d1$se[key]),
    `Kitagawa / educ16`  = ext_cell(kappa_kit_d2$estimates[key],  kappa_kit_d2$se[key]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

xgb_row <- function(label, a, b, c, d) {
  data.frame(
    Estimator = label,
    `Card / somecol` = ext_cell(a$coef, a$se),
    `Card / educ16` = ext_cell(b$coef, b$se),
    `Kitagawa / somecol` = ext_cell(c$coef, c$se),
    `Kitagawa / educ16` = ext_cell(d$coef, d$se),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

rbind(
  learner_row("PLR-IV (OLS)", lm_pliv_card_d1, lm_pliv_card_d2, lm_pliv_kit_d1, lm_pliv_kit_d2),
  learner_row("PLR-IV (ranger)", rf_pliv_card_d1, rf_pliv_card_d2, rf_pliv_kit_d1, rf_pliv_kit_d2),
  xgb_row("PLR-IV (XGBoost untuned)", xgb_pliv_card_d1_u, xgb_pliv_card_d2_u, xgb_pliv_kit_d1_u, xgb_pliv_kit_d2_u),
  xgb_row("PLR-IV (XGBoost tuned)", xgb_pliv_card_d1, xgb_pliv_card_d2, xgb_pliv_kit_d1, xgb_pliv_kit_d2),
  learner_row("Wald-AIPW (linear+logit)", lm_iivm_card_d1, lm_iivm_card_d2, lm_iivm_kit_d1, lm_iivm_kit_d2),
  learner_row("Wald-AIPW (ranger)", rf_iivm_card_d1, rf_iivm_card_d2, rf_iivm_kit_d1, rf_iivm_kit_d2),
  xgb_row("Wald-AIPW (XGBoost untuned)", xgb_card_d1_u, xgb_card_d2_u, xgb_kit_d1_u, xgb_kit_d2_u),
  xgb_row("Wald-AIPW (XGBoost tuned)", xgb_card_d1, xgb_card_d2, xgb_kit_d1, xgb_kit_d2),
  kappa_ref_row("τ_cb_u (kappa, ref)",  "tau_cb_u"),
  kappa_ref_row("τ_ml_u (kappa, ref)",  "tau_ml_u")
) %>%
  kbl(align = c("l","c","c","c","c"),
      caption = "Learner Comparison — DoubleML vs. Kappa (Y = log wages, dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  add_header_above(c(" " = 1, "Card covariates" = 2, "Kitagawa covariates" = 2)) %>%
  pack_rows("PLR-IV", 1, 4) %>%
  pack_rows("Wald-AIPW", 5, 8) %>%
  pack_rows("Kappa reference", 9, 10) %>%
  row_spec(1:8, background = "#eef6ff") %>%
  footnote(general = "Each cell reports estimate (SE). XGBoost rows import the stored nested-tuning fits. Their reconstructed outcome weights are assessed separately.")
```

---

# 4. Weight Diagnostics

<div class="note-box">
For a fixed weight vector, a weight sum near zero implies invariance to an
additive outcome shift. Complete-pipeline translation invariance is assessed
separately by refitting every estimator in the rerun notebook.
Treated/control mass diagnoses normalization, and arm-specific modified ESS
summarizes concentration.
Conventional Kish ESS is omitted because these are signed contrast weights.
</div>

```{r diag-builders, echo=FALSE}
# Kappa weight bundles for all four cells.
kw_card_d1 <- kappa_weights_bundle(Z, D1, X_card_kappa)
kw_card_d2 <- kappa_weights_bundle(Z, D2, X_card_kappa)
kw_kit_d1  <- kappa_weights_bundle(Z, D1, X_kit_kappa)
kw_kit_d2  <- kappa_weights_bundle(Z, D2, X_kit_kappa)

safe_outcome_diag <- function(w, D, label) {
  if (is.null(w) || !length(w) || all(is.na(w))) NULL else outcome_weight_diag(w, D, label)
}

DIAG_ALIGN <- c("l","r","c","r","c","c","c","r")

dml_diag_cell <- function(w_untuned, w_tuned, kw, D) {
  rbind(
    outcome_weight_diag(w_untuned$plriv,    D, "PLR-IV (untuned forest)"),
    outcome_weight_diag(w_untuned$waldaipw, D, "Wald-AIPW (untuned forest)"),
    outcome_weight_diag(w_tuned$plriv,      D, "PLR-IV (tuned forest)"),
    outcome_weight_diag(w_tuned$waldaipw,   D, "Wald-AIPW (tuned forest)"),
    outcome_weight_diag(kw$tau_cb_u,    D, "τ_cb_u     (kappa)"),
    outcome_weight_diag(kw$tau_ml_u,    D, "τ_ml_u     (kappa)"),
    outcome_weight_diag(kw$tau_ml_a10,  D, "τ_ml_a10   (kappa)")
  )
}
style_dml_diag <- function(df, cap) {
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(1:4, background = "#eef6ff") %>%
    footnote(general = "Blue rows are honest-forest DML. Treated/control entries use sign-oriented control weights. Arm-specific modified ESS is the headline concentration diagnostic.")
}

ext_diag_block <- function(D, X_kappa, w_pliv_lm, w_wald_lm, w_pliv_rf, w_wald_rf,
                           w_pliv_xgb, w_wald_xgb) {
  kw <- kappa_weights_bundle(Z, D, X_kappa)
  do.call(rbind, Filter(Negate(is.null), list(
    safe_outcome_diag(w_pliv_lm,     D, "PLR-IV (OLS)"),
    safe_outcome_diag(w_wald_lm,     D, "Wald-AIPW (linear+logit)"),
    safe_outcome_diag(w_pliv_rf,     D, "PLR-IV (ranger)"),
    safe_outcome_diag(w_wald_rf,     D, "Wald-AIPW (ranger)"),
    safe_outcome_diag(w_pliv_xgb,    D, "PLR-IV (XGBoost tuned)"),
    safe_outcome_diag(w_wald_xgb,    D, "Wald-AIPW (XGBoost tuned)"),
    outcome_weight_diag(kw$tau_cb_u,   D, "τ_cb_u    (kappa, CBPS)"),
    outcome_weight_diag(kw$tau_ml_u,   D, "τ_ml_u    (kappa, MLE)"),
    outcome_weight_diag(kw$tau_ml_a10, D, "τ_ml_a10  (kappa, MLE)")
  )))
}
style_ext_diag <- function(df, cap) {
  n_learn <- sum(grepl("^(PLR-IV|Wald-AIPW)", df$Estimator))
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(seq_len(n_learn), background = "#eef6ff") %>%
    footnote(general = "Blue rows are DoubleML learners. Numerical reproduction is reported separately and does not suppress reconstructed XGBoost weights.")
}
```

## DML vs. Kappa {.tabset}

### Card — somecol

```{r dml-diag-card-d1}
style_dml_diag(dml_diag_cell(w_card_d1_untuned, w_card_d1_tuned,
                            kw_card_d1, D1),
               "Weight Diagnostics — DML vs. Kappa | Card, somecol")
```

### Card — educ16

```{r dml-diag-card-d2}
style_dml_diag(dml_diag_cell(w_card_d2_untuned, w_card_d2_tuned,
                            kw_card_d2, D2),
               "Weight Diagnostics — DML vs. Kappa | Card, educ16")
```

### Kitagawa — somecol

```{r dml-diag-kit-d1}
style_dml_diag(dml_diag_cell(w_kit_d1_untuned, w_kit_d1_tuned,
                            kw_kit_d1, D1),
               "Weight Diagnostics — DML vs. Kappa | Kitagawa, somecol")
```

### Kitagawa — educ16

```{r dml-diag-kit-d2}
style_dml_diag(dml_diag_cell(w_kit_d2_untuned, w_kit_d2_tuned,
                            kw_kit_d2, D2),
               "Weight Diagnostics — DML vs. Kappa | Kitagawa, educ16")
```

## Learners vs. Kappa {.tabset}

### Card — somecol

```{r ext-diag-card-d1}
style_ext_diag(ext_diag_block(D1, X_card_kappa,
                              lm_pliv_card_d1$w, lm_iivm_card_d1$w,
                              rf_pliv_card_d1$w, rf_iivm_card_d1$w,
                              xgb_pliv_card_d1$w, xgb_card_d1$w),
               "Weight Diagnostics — Learners | Card, somecol")
```

### Card — educ16

```{r ext-diag-card-d2}
style_ext_diag(ext_diag_block(D2, X_card_kappa,
                              lm_pliv_card_d2$w, lm_iivm_card_d2$w,
                              rf_pliv_card_d2$w, rf_iivm_card_d2$w,
                              xgb_pliv_card_d2$w, xgb_card_d2$w),
               "Weight Diagnostics — Learners | Card, educ16")
```

### Kitagawa — somecol

```{r ext-diag-kit-d1}
style_ext_diag(ext_diag_block(D1, X_kit_kappa,
                              lm_pliv_kit_d1$w, lm_iivm_kit_d1$w,
                              rf_pliv_kit_d1$w, rf_iivm_kit_d1$w,
                              xgb_pliv_kit_d1$w, xgb_kit_d1$w),
               "Weight Diagnostics — Learners | Kitagawa, somecol")
```

### Kitagawa — educ16

```{r ext-diag-kit-d2}
style_ext_diag(ext_diag_block(D2, X_kit_kappa,
                              lm_pliv_kit_d2$w, lm_iivm_kit_d2$w,
                              rf_pliv_kit_d2$w, rf_iivm_kit_d2$w,
                              xgb_pliv_kit_d2$w, xgb_kit_d2$w),
               "Weight Diagnostics — Learners | Kitagawa, educ16")
```

## Detailed Outcome-Weight Statistics {.tabset}

<div class="note-box">
These tables retain the tuned honest-forest and XGBoost specifications together
with the normalized kappa estimators. They report Knaus-style overall, treated,
and sign-oriented control statistics without repeating the full learner
comparison.
</div>

```{r bridge-helpers, echo=FALSE}
cell_weight_list <- function(w_tuned, pliv_xgb_w, wald_xgb_w, kw) {
  lst <- list(
    "PLR-IV (tuned forest)" = w_tuned$plriv,
    "Wald-AIPW (tuned forest)" = w_tuned$waldaipw,
    "PLR-IV (XGBoost tuned; reconstructed)" = pliv_xgb_w,
    "Wald-AIPW (XGBoost tuned; reconstructed)" = wald_xgb_w,
    "τ_cb_u (kappa)" = kw$tau_cb_u,
    "τ_ml_u (kappa)" = kw$tau_ml_u,
    "τ_ml_a10 (kappa)" = kw$tau_ml_a10
  )
  Filter(function(w) !is.null(w) && length(w) && !all(is.na(w)), lst)
}

bridge_cells <- list(
  "Card · somecol" = list(D = D1, w = cell_weight_list(
    w_card_d1_tuned, xgb_pliv_card_d1$w, xgb_card_d1$w, kw_card_d1)),
  "Card · educ16" = list(D = D2, w = cell_weight_list(
    w_card_d2_tuned, xgb_pliv_card_d2$w, xgb_card_d2$w, kw_card_d2)),
  "Kitagawa · somecol" = list(D = D1, w = cell_weight_list(
    w_kit_d1_tuned, xgb_pliv_kit_d1$w, xgb_kit_d1$w, kw_kit_d1)),
  "Kitagawa · educ16" = list(D = D2, w = cell_weight_list(
    w_kit_d2_tuned, xgb_pliv_kit_d2$w, xgb_kit_d2$w, kw_kit_d2))
)

# Full bridge table for one cell: estimators as row-groups, scopes as rows,
# columns split into "Knaus summary()" vs. "ESS" via a spanning header.
bridge_table <- function(wlist, D, cap) {
  ests <- names(wlist)
  df   <- do.call(rbind, lapply(ests, function(nm) weight_bridge(wlist[[nm]], D, nm)))
  disp <- df[, c("Scope","Min","Max","Neg_share","Sum_top10","Sum_w",
                 "Sum_abs_w","Max_abs_w","ESS_mod")]

  kk <- disp %>%
    kbl(caption = cap, row.names = FALSE,
        col.names = c("Scope","Min","Max","Neg. share","ΣTop10%","Σω","Σ|ω|",
                      "Max |ω|","ESS mod"),
        align = c("l", rep("r", 8))) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 12) %>%
    add_header_above(c(" " = 1, "Knaus summary() statistics" = 6,
                       "Concentration" = 1,
                       "Effective sample size" = 1)) %>%
    column_spec(9, bold = TRUE, background = "#eef6ff")
  for (i in seq_along(ests)) kk <- pack_rows(kk, ests[i], (i - 1) * 3 + 1, i * 3)
  kk %>% footnote(general = "Neg. share is Knaus's observation-count statistic in [0,1]. Control rows use the sign-flipped −ω convention. Modified ESS is not repeated in the interpretation of these detailed tables.")
}
```

### Card · somecol

```{r bridge-card-d1}
bridge_table(bridge_cells[["Card · somecol"]]$w, bridge_cells[["Card · somecol"]]$D,
             "Weight-statistics bridge — Card, somecol")
```

### Card · educ16

```{r bridge-card-d2}
bridge_table(bridge_cells[["Card · educ16"]]$w, bridge_cells[["Card · educ16"]]$D,
             "Weight-statistics bridge — Card, educ16")
```

### Kitagawa · somecol

```{r bridge-kit-d1}
bridge_table(bridge_cells[["Kitagawa · somecol"]]$w, bridge_cells[["Kitagawa · somecol"]]$D,
             "Weight-statistics bridge — Kitagawa, somecol")
```

### Kitagawa · educ16

```{r bridge-kit-d2}
bridge_table(bridge_cells[["Kitagawa · educ16"]]$w, bridge_cells[["Kitagawa · educ16"]]$D,
             "Weight-statistics bridge — Kitagawa, educ16")
```

---

# 5. Love Plots

<div class="note-box">
Panels report absolute mean differences. Continuous covariates are standardized;
binary covariates remain on the raw 0/1 scale.
</div>

```{r love-helper, echo=FALSE}
maybe_love <- function(title, fit, D, X) {
  if (is.null(fit$w)) NULL else make_love(title, fit$w, D, X)
}
```

## DML vs. Kappa {.tabset}

### Card — somecol

```{r love-card-d1, fig.height=9, fig.width=13}
grid.arrange(grobs = list(
  make_love("PLR-IV\nuntuned forest", w_card_d1_untuned$plriv, D1, X_dml_card),
  make_love("Wald-AIPW\nuntuned forest", w_card_d1_untuned$waldaipw, D1, X_dml_card),
  make_love("PLR-IV\ntuned forest", w_card_d1_tuned$plriv, D1, X_dml_card),
  make_love("Wald-AIPW\ntuned forest", w_card_d1_tuned$waldaipw, D1, X_dml_card),
  make_love("τ_cb_u (CBPS)",    kw_card_d1$tau_cb_u,   D1, X_dml_card),
  make_love("τ_ml_u",           kw_card_d1$tau_ml_u,   D1, X_dml_card),
  make_love("τ_ml_a10",         kw_card_d1$tau_ml_a10, D1, X_dml_card)
), nrow = 2, top = "Love Plots — Card spec, somecol: DML vs. Kappa")
```

### Card — college completion

```{r love-card-d2, fig.height=9, fig.width=13}
grid.arrange(grobs = list(
  make_love("PLR-IV\nuntuned forest", w_card_d2_untuned$plriv, D2, X_dml_card),
  make_love("Wald-AIPW\nuntuned forest", w_card_d2_untuned$waldaipw, D2, X_dml_card),
  make_love("PLR-IV\ntuned forest", w_card_d2_tuned$plriv, D2, X_dml_card),
  make_love("Wald-AIPW\ntuned forest", w_card_d2_tuned$waldaipw, D2, X_dml_card),
  make_love("τ_cb_u (CBPS)",    kw_card_d2$tau_cb_u,   D2, X_dml_card),
  make_love("τ_ml_u",           kw_card_d2$tau_ml_u,   D2, X_dml_card),
  make_love("τ_ml_a10",         kw_card_d2$tau_ml_a10, D2, X_dml_card)
), nrow = 2, top = "Love Plots — Card spec, college completion: DML vs. Kappa")
```

### Kitagawa — somecol

```{r love-kit-d1, fig.height=9, fig.width=13}
grid.arrange(grobs = list(
  make_love("PLR-IV\nuntuned forest", w_kit_d1_untuned$plriv, D1, X_dml_kit),
  make_love("Wald-AIPW\nuntuned forest", w_kit_d1_untuned$waldaipw, D1, X_dml_kit),
  make_love("PLR-IV\ntuned forest", w_kit_d1_tuned$plriv, D1, X_dml_kit),
  make_love("Wald-AIPW\ntuned forest", w_kit_d1_tuned$waldaipw, D1, X_dml_kit),
  make_love("τ_cb_u (CBPS)",    kw_kit_d1$tau_cb_u,   D1, X_dml_kit),
  make_love("τ_ml_u",           kw_kit_d1$tau_ml_u,   D1, X_dml_kit),
  make_love("τ_ml_a10",         kw_kit_d1$tau_ml_a10, D1, X_dml_kit)
), nrow = 2, top = "Love Plots — Kitagawa spec, somecol: DML vs. Kappa")
```

### Kitagawa — college completion

```{r love-kit-d2, fig.height=9, fig.width=13}
grid.arrange(grobs = list(
  make_love("PLR-IV\nuntuned forest", w_kit_d2_untuned$plriv, D2, X_dml_kit),
  make_love("Wald-AIPW\nuntuned forest", w_kit_d2_untuned$waldaipw, D2, X_dml_kit),
  make_love("PLR-IV\ntuned forest", w_kit_d2_tuned$plriv, D2, X_dml_kit),
  make_love("Wald-AIPW\ntuned forest", w_kit_d2_tuned$waldaipw, D2, X_dml_kit),
  make_love("τ_cb_u (CBPS)",    kw_kit_d2$tau_cb_u,   D2, X_dml_kit),
  make_love("τ_ml_u",           kw_kit_d2$tau_ml_u,   D2, X_dml_kit),
  make_love("τ_ml_a10",         kw_kit_d2$tau_ml_a10, D2, X_dml_kit)
), nrow = 2, top = "Love Plots — Kitagawa spec, college completion: DML vs. Kappa")
```

## Learners vs. Kappa {.tabset}

### Card — somecol (learners + kappa)

```{r ext-love-card-d1, fig.height=5.5, fig.width=14}
grid.arrange(grobs = Filter(Negate(is.null), list(
  make_love("PLR-IV OLS\nCard, somecol", lm_pliv_card_d1$w, D1, X_dml_card),
  make_love("Wald-AIPW linear+logit\nCard, somecol", lm_iivm_card_d1$w, D1, X_dml_card),
  make_love("PLR-IV ranger\nCard, somecol", rf_pliv_card_d1$w, D1, X_dml_card),
  make_love("Wald-AIPW ranger\nCard, somecol", rf_iivm_card_d1$w, D1, X_dml_card),
  maybe_love("PLR-IV XGBoost tuned\nCard, somecol", xgb_pliv_card_d1, D1, X_dml_card),
  maybe_love("Wald-AIPW XGBoost tuned\nCard, somecol", xgb_card_d1, D1, X_dml_card),
  make_love("τ_cb_u\nCard, somecol",       kw_card_d1$tau_cb_u, D1, X_dml_card),
  make_love("τ_ml_u\nCard, somecol",       kw_card_d1$tau_ml_u, D1, X_dml_card)
)), nrow = 2, top = "Love Plots — DoubleML learners vs. kappa | Card, somecol")
```

### Card — college completion (learners + kappa)

```{r ext-love-card-d2, fig.height=7.5, fig.width=14}
grid.arrange(grobs = Filter(Negate(is.null), list(
  make_love("PLR-IV OLS\nCard, completion", lm_pliv_card_d2$w, D2, X_dml_card),
  make_love("Wald-AIPW linear+logit\nCard, completion", lm_iivm_card_d2$w, D2, X_dml_card),
  make_love("PLR-IV ranger\nCard, completion", rf_pliv_card_d2$w, D2, X_dml_card),
  make_love("Wald-AIPW ranger\nCard, completion", rf_iivm_card_d2$w, D2, X_dml_card),
  maybe_love("PLR-IV XGBoost tuned\nCard, completion", xgb_pliv_card_d2, D2, X_dml_card),
  maybe_love("Wald-AIPW XGBoost tuned\nCard, completion", xgb_card_d2, D2, X_dml_card),
  make_love("τ_cb_u\nCard, completion", kw_card_d2$tau_cb_u, D2, X_dml_card),
  make_love("τ_ml_u\nCard, completion", kw_card_d2$tau_ml_u, D2, X_dml_card)
)), nrow = 2, top = "Love Plots — DoubleML learners vs. kappa | Card, college completion")
```

### Kitagawa — somecol (learners + kappa)

```{r ext-love-kit-d1, fig.height=7.5, fig.width=14}
grid.arrange(grobs = Filter(Negate(is.null), list(
  make_love("PLR-IV OLS\nKitagawa, somecol", lm_pliv_kit_d1$w, D1, X_dml_kit),
  make_love("Wald-AIPW linear+logit\nKitagawa, somecol", lm_iivm_kit_d1$w, D1, X_dml_kit),
  make_love("PLR-IV ranger\nKitagawa, somecol", rf_pliv_kit_d1$w, D1, X_dml_kit),
  make_love("Wald-AIPW ranger\nKitagawa, somecol", rf_iivm_kit_d1$w, D1, X_dml_kit),
  maybe_love("PLR-IV XGBoost tuned\nKitagawa, somecol", xgb_pliv_kit_d1, D1, X_dml_kit),
  maybe_love("Wald-AIPW XGBoost tuned\nKitagawa, somecol", xgb_kit_d1, D1, X_dml_kit),
  make_love("τ_cb_u\nKitagawa, somecol", kw_kit_d1$tau_cb_u, D1, X_dml_kit),
  make_love("τ_ml_u\nKitagawa, somecol", kw_kit_d1$tau_ml_u, D1, X_dml_kit)
)), nrow = 2, top = "Love Plots — DoubleML learners vs. kappa | Kitagawa, somecol")
```

### Kitagawa — college completion (learners + kappa)

```{r ext-love-kit-d2, fig.height=7.5, fig.width=14}
grid.arrange(grobs = Filter(Negate(is.null), list(
  make_love("PLR-IV OLS\nKitagawa, completion", lm_pliv_kit_d2$w, D2, X_dml_kit),
  make_love("Wald-AIPW linear+logit\nKitagawa, completion", lm_iivm_kit_d2$w, D2, X_dml_kit),
  make_love("PLR-IV ranger\nKitagawa, completion", rf_pliv_kit_d2$w, D2, X_dml_kit),
  make_love("Wald-AIPW ranger\nKitagawa, completion", rf_iivm_kit_d2$w, D2, X_dml_kit),
  maybe_love("PLR-IV XGBoost tuned\nKitagawa, completion", xgb_pliv_kit_d2, D2, X_dml_kit),
  maybe_love("Wald-AIPW XGBoost tuned\nKitagawa, completion", xgb_kit_d2, D2, X_dml_kit),
  make_love("τ_cb_u\nKitagawa, completion",       kw_kit_d2$tau_cb_u, D2, X_dml_kit),
  make_love("τ_ml_u\nKitagawa, completion",       kw_kit_d2$tau_ml_u, D2, X_dml_kit)
)), nrow = 2, top = "Love Plots — DoubleML learners vs. kappa | Kitagawa, college completion")
```
