---
title: "Card (1995) — Translation Invariance: Rerun Check"
subtitle: "Kappa estimators, DML smoother (PLR-IV / Wald-AIPW), and the 3-learner Wald-AIPW comparison"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
h3   { color: #444; margin-top: 20px; }
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 12,
  fig.height = 6,
  autodep    = TRUE
)
```

---

# Purpose

This notebook checks translation invariance by the rerun method: the whole estimation
pipeline is fit once on the outcome measured in dollars and once on the same outcome
measured in cents, using the same seed, and the two resulting point estimates are
compared directly. This complements the frozen-weight check already used in the main
Card notebook, where translation invariance is inferred from whether the outcome
weights sum to zero without actually refitting anything.

Recoding wages from dollars to cents multiplies the wage by 100 before taking logs,
which is the same as adding a constant to the log wage. Since the choice between
dollars and cents carries no real economic information, a translation-invariant
estimator should return the same point estimate under both codings.

# Setup

```{r packages-and-source}
library(haven)
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(OutcomeWeights)
library(cobalt)
library(viridis)
library(gridExtra)
library(ggplot2)
library(DoubleML)
library(mlr3verse)

source("functions_all.R")
cat("functions_all.R sourced.\n")
```

```{r data}
data_path  <- "/Users/alexung/Desktop/master_thesis/code/"
card       <- read_dta(file.path(data_path, "card.dta"))

card_clean <- card[!is.na(card$wage) & !is.na(card$educ), ]

card_clean$lwage_cnt <- log(card_clean$wage)         # cents
card_clean$lwage     <- log(card_clean$wage / 100)   # dollars

card_clean$somecol <- as.integer(card_clean$educ > 12)
card_clean$educ16  <- as.integer(card_clean$educ >= 16)
card_clean$Z_inst  <- as.integer(card_clean$nearc4)

if (!"expersq" %in% names(card_clean)) {
  card_clean$expersq <- card_clean$exper^2
}

cat(sprintf("N = %d (should be 3,010)\n", nrow(card_clean)))
```

```{r design-matrices}
X_card_kappa <- model.matrix(
  ~ exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card_clean
)
X_kit_kappa <- model.matrix(
  ~ black + south + smsa + smsa66 + south66,
  data = card_clean
)

X_dml_card <- make_X(
  ~ 0 + exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card_clean
)
X_dml_kit <- make_X(
  ~ 0 + black + south + smsa + smsa66 + south66,
  data = card_clean
)

Y_dol <- card_clean$lwage        # dollars
Y_cnt <- card_clean$lwage_cnt    # cents
Z     <- card_clean$Z_inst
D1    <- card_clean$somecol
D2    <- card_clean$educ16

k <- log(100)

# the unit recoding should differ by exactly log(100); strip attributes from
# the haven-labelled columns before comparing so the check is on values only
diff_vec <- as.numeric(Y_cnt) - as.numeric(Y_dol)
max_dev  <- max(abs(diff_vec - k))
cat(sprintf("log(100) = %.12f, max deviation = %.3e\n", k, max_dev))
stopifnot(max_dev < 1e-9)

cells <- list(
  list(label = "Card / somecol",     X_kappa = X_card_kappa, X_dml = X_dml_card, D = D1),
  list(label = "Card / educ16",      X_kappa = X_card_kappa, X_dml = X_dml_card, D = D2),
  list(label = "Kitagawa / somecol", X_kappa = X_kit_kappa,  X_dml = X_dml_kit,  D = D1),
  list(label = "Kitagawa / educ16",  X_kappa = X_kit_kappa,  X_dml = X_dml_kit,  D = D2)
)
```

---

# 1. Kappa estimators

Kappa outcome weights are built only from the instrument, the treatment, and the
estimated propensity score, so they do not depend on the outcome at all. Refitting the
kappa estimators on the cents-coded outcome should therefore give exactly the same
weights as the dollar fit, and the rerun and frozen-weight checks should coincide.

```{r kappa-name-bridge}
# kappa_analytic_se_all() (functions_all.R) labels the tau_a,1 estimator "tau_ml_t",
# while kappa_weights_bundle() labels the same estimator's weight vector "tau_ml_a1".
# Bridge the two naming conventions explicitly so nothing silently mismatches.
kappa_key_map <- c(
  tau_cb_u   = "tau_cb_u",
  tau_ml_u   = "tau_ml_u",
  tau_ml_a10 = "tau_ml_a10",
  tau_ml_a   = "tau_ml_a",
  tau_ml_t   = "tau_ml_a1",
  tau_ml_a0  = "tau_ml_a0"
)
```

```{r kappa-ti-builder}
build_kappa_ti_rows <- function(cell_label, Y_dollars, Y_cents, Z, D, X_kappa, k) {
  est_d  <- kappa_analytic_se_all(Y_dollars, Z, D, X_kappa)$estimates
  est_c  <- kappa_analytic_se_all(Y_cents,   Z, D, X_kappa)$estimates
  w_list <- kappa_weights_bundle(Z, D, X_kappa)

  rows <- lapply(names(kappa_key_map), function(est_name) {
    w_name <- kappa_key_map[[est_name]]
    row <- ti_unified_row(
      name        = est_name,
      w_dollars   = w_list[[w_name]],
      tau_dollars = est_d[[est_name]],
      tau_cents   = est_c[[est_name]],
      k           = k
    )
    row$Cell <- cell_label
    row
  })
  do.call(rbind, rows)
}
```

```{r kappa-ti-run, cache=TRUE, dependson=c("data","design-matrices","kappa-name-bridge","kappa-ti-builder")}
kappa_ti_all <- do.call(rbind, lapply(cells, function(cl) {
  build_kappa_ti_rows(cl$label, Y_dol, Y_cnt, Z, cl$D, cl$X_kappa, k)
}))
```

```{r kappa-ti-table}
kappa_ti_all[, setdiff(names(kappa_ti_all), "gap_AB")] %>%
  kbl(caption = "Kappa estimators — dollars vs. cents, all 4 Card cells",
      digits = 6, row.names = FALSE,
      align = c("l", rep("r", 5), "c", "c", "l")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) %>%
  row_spec(which(!kappa_ti_all$invariant_A), background = "#fffbf0")
```

---

# 2. DML headline estimators (PLR-IV, Wald-AIPW)

The smoother-based PLR-IV and Wald-AIPW estimators rely on nuisance models that are
estimated from the data, so unlike the kappa weights, the implied outcome weights could
in principle change when the outcome is recoded. Both the dollar and cents fits below
use the same seed, the same number of cross-fitting folds, and the same tuning setting.

```{r dml-headline-builder}
build_dml_headline_ti_rows <- function(cell_label, Y_dollars, Y_cents, D, X_dml, Z, k,
                                        n_cf_folds = 5) {
  set.seed(42)
  fit_d <- dml_with_smoother(Y_dollars, D, X_dml, Z, n_cf_folds = n_cf_folds,
                              tune.parameters = "all")
  set.seed(42)
  fit_c <- dml_with_smoother(Y_cents, D, X_dml, Z, n_cf_folds = n_cf_folds,
                              tune.parameters = "all")

  res_d <- summary(fit_d)
  res_c <- summary(fit_c)
  om_d  <- get_outcome_weights(fit_d)
  check_omega_rows(om_d)

  rows <- lapply(c("PLR-IV", "Wald-AIPW"), function(est) {
    w   <- as.vector(om_d$omega[est, ])
    row <- ti_unified_row(
      name        = est,
      w_dollars   = w,
      tau_dollars = get_estimate(res_d, est),
      tau_cents   = get_estimate(res_c, est),
      k           = k
    )
    row$Cell <- cell_label
    row
  })
  do.call(rbind, rows)
}
```

```{r dml-headline-run, cache=TRUE, dependson=c("data","design-matrices","dml-headline-builder")}
dml_headline_ti_all <- do.call(rbind, lapply(cells, function(cl) {
  build_dml_headline_ti_rows(cl$label, Y_dol, Y_cnt, cl$D, cl$X_dml, Z, k)
}))
```

```{r dml-headline-table}
dml_headline_ti_all[, setdiff(names(dml_headline_ti_all), "gap_AB")] %>%
  kbl(caption = "DML smoother estimators — dollars vs. cents, all 4 Card cells",
      digits = 6, row.names = FALSE,
      align = c("l", rep("r", 5), "c", "c", "l")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) %>%
  row_spec(which(!dml_headline_ti_all$invariant_A), background = "#fffbf0")
```

---

# 3. DML Wald-AIPW across learners (lm/logit, ranger, tuned XGBoost)

For XGBoost, the default hyperparameters produced implausible point estimates in the
main notebook (large negative LATEs, far from the lm/ranger/2SLS benchmarks). The
hyperparameters used for XGBoost here were selected separately in
`card_xgboost_tuning_1.Rmd` by inner cross-validation on nuisance prediction loss, and
are plugged in below as fixed values, so the tuning search itself is not rerun here.

```{r dml-learner-setup}
make_dml_data <- function(D, X_dml, Y, Z) {
  df <- as.data.frame(X_dml)
  df$y <- as.numeric(Y)
  df$d <- as.numeric(D)
  df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                    x_cols = colnames(X_dml))
}

ml_lm <- function() list(g = lrn("regr.lm"),
                         m = lrn("classif.log_reg"),
                         r = lrn("classif.log_reg"))
ml_rf <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                         m = lrn("classif.ranger", keep.inbag = TRUE),
                         r = lrn("classif.ranger", keep.inbag = TRUE))

fit_iivm <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLIIVM$new(data = dml_data,
                          ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5, score = "LATE")
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  om  <- get_outcome_weights(object = obj, dml_data = dml_data)
  w   <- as.vector(om$omega)
  ok  <- check_doubleml_identity(om, Y, obj$coef)
  list(obj = obj, w = w, ok = ok)
}

build_dml_learner_ti_row <- function(cell_label, label, learners_fn,
                                      Y_dollars, Y_cents, D, X_dml, Z, k) {
  dat_d <- make_dml_data(D, X_dml, Y_dollars, Z)
  dat_c <- make_dml_data(D, X_dml, Y_cents,   Z)

  fit_d <- fit_iivm(dat_d, learners_fn(), Y_dollars)
  fit_c <- fit_iivm(dat_c, learners_fn(), Y_cents)

  row <- ti_unified_row(
    name        = label,
    w_dollars   = fit_d$w,
    tau_dollars = get_dml_coef(fit_d$obj),
    tau_cents   = get_dml_coef(fit_c$obj),
    k           = k
  )
  row$Cell        <- cell_label
  row$identity_ok <- fit_d$ok
  row
}

learner_specs <- list(
  list(label = "Wald-AIPW (lm + logit)", fn = ml_lm),
  list(label = "Wald-AIPW (ranger)",     fn = ml_rf)
)
```

```{r xgb-tuned-params}
# Selected by inner cross-validation on nuisance prediction loss in
# card_xgboost_tuning_1.Rmd (8-eval random search, 3-fold inner CV, seed 42).
# g0/g1 = outcome model by instrument arm, m = instrument model,
# r0/r1 = treatment model by instrument arm.
# Card/educ16 values are read off the full-precision obj$params printout; the
# other three cells use the rounded values from the hyperparameter table.
xgb_tuned_params <- list(
  "Card / educ16" = list(
    g0 = list(nrounds = 227, max_depth = 3, eta = 0.04284429, min_child_weight = 5),
    g1 = list(nrounds = 177, max_depth = 3, eta = 0.02897991, min_child_weight = 2),
    m  = list(nrounds = 74,  max_depth = 2, eta = 0.08936620, min_child_weight = 7),
    r0 = list(nrounds = 268, max_depth = 2, eta = 0.07932856, min_child_weight = 3),
    r1 = list(nrounds = 255, max_depth = 3, eta = 0.03271248, min_child_weight = 1)
  ),
  "Card / somecol" = list(
    g0 = list(nrounds = 227, max_depth = 3, eta = 0.04284429, min_child_weight = 5),
    g1 = list(nrounds = 177, max_depth = 3, eta = 0.02897991, min_child_weight = 2),
    m  = list(nrounds = 74,  max_depth = 2, eta = 0.08936620, min_child_weight = 7),
    r0 = list(nrounds = 65,  max_depth = 4, eta = 0.065,      min_child_weight = 4),
    r1 = list(nrounds = 255, max_depth = 3, eta = 0.03271248, min_child_weight = 1)
  ),
  "Kitagawa / educ16" = list(
    g0 = list(nrounds = 320, max_depth = 3, eta = 0.051, min_child_weight = 7),
    g1 = list(nrounds = 218, max_depth = 2, eta = 0.183, min_child_weight = 6),
    m  = list(nrounds = 373, max_depth = 5, eta = 0.052, min_child_weight = 6),
    r0 = list(nrounds = 282, max_depth = 3, eta = 0.133, min_child_weight = 1),
    r1 = list(nrounds = 75,  max_depth = 4, eta = 0.073, min_child_weight = 3)
  ),
  "Kitagawa / somecol" = list(
    g0 = list(nrounds = 320, max_depth = 3, eta = 0.051, min_child_weight = 7),
    g1 = list(nrounds = 218, max_depth = 2, eta = 0.183, min_child_weight = 6),
    m  = list(nrounds = 373, max_depth = 5, eta = 0.052, min_child_weight = 6),
    r0 = list(nrounds = 142, max_depth = 5, eta = 0.162, min_child_weight = 2),
    r1 = list(nrounds = 75,  max_depth = 4, eta = 0.073, min_child_weight = 3)
  )
)

make_xgb_tuned_learners <- function() {
  list(
    g = lrn("regr.xgboost",    alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
    m = lrn("classif.xgboost", predict_type = "prob"),
    r = lrn("classif.xgboost", predict_type = "prob")
  )
}

# fixes the tuned hyperparameters via set_ml_nuisance_params() instead of
# calling obj$tune(), so the tuning search is not repeated here
fit_iivm_xgb_tuned <- function(dml_data, hp, Y) {
  set.seed(42)
  learners <- make_xgb_tuned_learners()
  obj <- DoubleMLIIVM$new(data = dml_data,
                          ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5, score = "LATE")

  obj$set_ml_nuisance_params(learner = "ml_g0", treat_var = "d", params = hp$g0)
  obj$set_ml_nuisance_params(learner = "ml_g1", treat_var = "d", params = hp$g1)
  obj$set_ml_nuisance_params(learner = "ml_m",  treat_var = "d", params = hp$m)
  obj$set_ml_nuisance_params(learner = "ml_r0", treat_var = "d", params = hp$r0)
  obj$set_ml_nuisance_params(learner = "ml_r1", treat_var = "d", params = hp$r1)

  obj$fit(store_models = TRUE, store_predictions = TRUE)
  om  <- get_outcome_weights(object = obj, dml_data = dml_data)
  w   <- as.vector(om$omega)
  ok  <- check_doubleml_identity(om, Y, obj$coef)
  list(obj = obj, w = w, ok = ok)
}

build_dml_learner_ti_row_xgb_tuned <- function(cell_label, hp, Y_dollars, Y_cents,
                                                D, X_dml, Z, k) {
  dat_d <- make_dml_data(D, X_dml, Y_dollars, Z)
  dat_c <- make_dml_data(D, X_dml, Y_cents,   Z)

  fit_d <- fit_iivm_xgb_tuned(dat_d, hp, Y_dollars)
  fit_c <- fit_iivm_xgb_tuned(dat_c, hp, Y_cents)

  row <- ti_unified_row(
    name        = "Wald-AIPW (XGBoost, tuned)",
    w_dollars   = fit_d$w,
    tau_dollars = get_dml_coef(fit_d$obj),
    tau_cents   = get_dml_coef(fit_c$obj),
    k           = k
  )
  row$Cell        <- cell_label
  row$identity_ok <- fit_d$ok
  row
}
```

```{r dml-learner-run, cache=TRUE, dependson=c("data","design-matrices","dml-learner-setup","xgb-tuned-params")}
dml_learner_ti_lm_rf <- do.call(rbind, lapply(cells, function(cl) {
  do.call(rbind, lapply(learner_specs, function(ls) {
    build_dml_learner_ti_row(cl$label, ls$label, ls$fn, Y_dol, Y_cnt, cl$D, cl$X_dml, Z, k)
  }))
}))

dml_learner_ti_xgb <- do.call(rbind, lapply(cells, function(cl) {
  build_dml_learner_ti_row_xgb_tuned(cl$label, xgb_tuned_params[[cl$label]],
                                      Y_dol, Y_cnt, cl$D, cl$X_dml, Z, k)
}))

dml_learner_ti_all <- rbind(dml_learner_ti_lm_rf, dml_learner_ti_xgb)
```

```{r dml-learner-table}
drop_cols <- c("gap_AB", "identity_ok")
dml_learner_ti_all[, setdiff(names(dml_learner_ti_all), drop_cols)] %>%
  kbl(caption = "DML Wald-AIPW by learner — dollars vs. cents, all 4 Card cells",
      digits = 6, row.names = FALSE,
      align = c("l", rep("r", 5), "c", "c", "l")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) %>%
  row_spec(which(!dml_learner_ti_all$invariant_A), background = "#fffbf0")
```
