---
title: "Card (1995) — XGBoost Nuisance Tuning for Wald-AIPW"
subtitle: "Tuning by nuisance loss (not the estimand) and measuring departure from the affine-smoother condition"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 5.5)
```

```{r packages}
library(haven)
library(knitr)
library(kableExtra)
library(ggplot2)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)

source("functions_all.R")
lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

# Runtime knobs (dial these down for a quick pass)
SEED        <- 42
N_CF        <- 5    # outer cross-fitting folds
TUNE_EVALS  <- 8    # random-search evaluations per nuisance
INNER_FOLDS <- 3    # inner CV folds for tuning
```

<div class="note-box">
**The problem.** In the Card application the XGBoost Wald-AIPW estimate is unstable. This is the behaviour Knaus (2024, Fig. 1) documents: with XGBoost nuisances, Wald-AIPW estimates range wildly because XGBoost violates **Condition 3** (affine smoother — rows of the smoother matrix summing to one), and in standard DML **Condition 5b** also fails (different models for outcome and treatment). Table 6 marks this implementation "no b/c ~~C3~~ & ~~C5b~~".

**What we can and cannot fix.** Tuning will **not** restore the affine-smoother property — that is intrinsic to boosting. But (i) tuning the nuisances by predictive loss keeps the propensity models away from 0/1, which tames the inverse-probability terms that drive the explosions, and (ii) we can *measure* the departure from Condition 3 directly: for Wald-AIPW, C3 is equivalent to the outcome weights summing to zero (Knaus 2024 §4.2.4: `T·1 = 0 ⇒ Σω = 0`). So **|Σω| is an accessible, theory-grounded measure of how far XGBoost departs from C3**, and we ask whether tuning reduces it.

**No estimand-tuning.** Selection is on nuisance cross-validation loss (RMSE for the outcome regression, log-loss for the propensities) — never on the point estimate. This is the property DML theory assumes, not a bias.
</div>

---

# Setup

```{r data}
data_path <- "/Users/alexung/Desktop/master_thesis/code/"
card      <- read_dta(file.path(data_path, "card.dta"))
card      <- card[!is.na(card$wage) & !is.na(card$educ), ]

card$somecol <- as.integer(card$educ > 12)
card$educ16  <- as.integer(card$educ >= 16)
if (!"expersq" %in% names(card)) card$expersq <- card$exper^2

Y_dol <- log(card$wage)
Z     <- as.integer(card$nearc4)
D1    <- card$somecol
D2    <- card$educ16

X_dml_card <- make_X(~ 0 + exper + expersq + black + south + smsa + smsa66 +
                       reg661+reg662+reg663+reg664+reg665+reg666+reg667+reg668, data = card)
X_dml_kit  <- make_X(~ 0 + black + south + smsa + smsa66 + south66, data = card)

cat(sprintf("N=%d | Card %d cov | Kitagawa %d cov\n",
            nrow(card), ncol(X_dml_card), ncol(X_dml_kit)))
```

```{r learners}
make_dml_data <- function(D, X) {
  df <- as.data.frame(X); df$y <- Y_dol; df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(df, y_col = "y", d_cols = "d", z_cols = "z", x_cols = colnames(X))
}

# Reference: ranger (affine smoother — satisfies C3, so Σω ≈ 0)
ml_rf <- function() list(
  g = lrn("regr.ranger",    keep.inbag = TRUE),
  m = lrn("classif.ranger", keep.inbag = TRUE, predict_type = "prob"),
  r = lrn("classif.ranger", keep.inbag = TRUE, predict_type = "prob"))

# Baseline XGBoost (current pipeline): the 4 structural params on the OUTCOME
# regression g are required by the OutcomeWeights vignette so that SY = Ŷ
# (weights remain extractable). They do NOT make XGBoost affine.
ml_xgb_untuned <- function() list(
  g = lrn("regr.xgboost",    alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
  m = lrn("classif.xgboost", predict_type = "prob"),
  r = lrn("classif.xgboost", predict_type = "prob"))

# Tuned XGBoost: freeze the 4 structural params on g, tune the safe knobs;
# tune m and r freely (no structural constraints on the propensities).
wrap_at <- function(learner, measure) AutoTuner$new(
  learner    = learner,
  resampling = rsmp("cv", folds = INNER_FOLDS),
  measure    = msr(measure),
  terminator = trm("evals", n_evals = TUNE_EVALS),
  tuner      = tnr("random_search"))

make_tuned_xgb <- function() {
  g <- lrn("regr.xgboost",
           alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0,  # FROZEN
           nrounds          = to_tune(50, 400),
           max_depth        = to_tune(2, 5),
           eta              = to_tune(0.02, 0.20),
           min_child_weight = to_tune(1, 8))
  mk_classif <- function() lrn("classif.xgboost", predict_type = "prob",
           nrounds          = to_tune(50, 400),
           max_depth        = to_tune(2, 5),
           eta              = to_tune(0.02, 0.20),
           min_child_weight = to_tune(1, 8))
  list(g = wrap_at(g, "regr.rmse"),
       m = wrap_at(mk_classif(), "classif.logloss"),
       r = wrap_at(mk_classif(), "classif.logloss"))
}
```

```{r fit-helpers}
fit_cell <- function(dml_data, learners, label) {
  tryCatch({
    set.seed(SEED)
    obj <- DoubleMLIIVM$new(dml_data, ml_g = learners$g, ml_m = learners$m,
                            ml_r = learners$r, n_folds = N_CF, score = "LATE")
    obj$fit(store_models = TRUE, store_predictions = TRUE)
    om <- get_outcome_weights(object = obj, dml_data = dml_data)
    w  <- as.vector(om$omega)
    list(coef  = get_dml_coef(obj),
         se    = get_dml_se(obj),
         sum_w = weight_diag(w, label)$Sum_w,
         ok    = check_doubleml_identity(om, Y_dol, obj$coef),
         obj   = obj)
  }, error = function(e) { message(label, " failed: ", conditionMessage(e)); NULL })
}

prop_extremes <- function(r, thr = 0.02) {
  if (is.null(r)) return(NA_real_)
  m <- tryCatch(as.vector(r$obj$predictions$ml_m), error = function(e) NA)
  rr <- tryCatch(as.vector(r$obj$predictions$ml_r), error = function(e) NA)
  p <- na.omit(c(m, rr))
  if (!length(p)) return(NA_real_)
  round(mean(p < thr | p > 1 - thr), 4)
}

cells <- list(
  "Card / somecol"     = list(D = D1, X = X_dml_card),
  "Card / educ16"      = list(D = D2, X = X_dml_card),
  "Kitagawa / somecol" = list(D = D1, X = X_dml_kit),
  "Kitagawa / educ16"  = list(D = D2, X = X_dml_kit)
)
```

---

# Fit: Ranger reference vs XGBoost (untuned vs tuned)

<div class="warn-box">
**Runtime.** The tuned fits run a random search (`TUNE_EVALS` evals × `INNER_FOLDS`-fold inner CV) inside each cross-fitting fold, for three nuisances, across four cells. This is the slow chunk; it is cached. Lower `TUNE_EVALS` / `N_CF` at the top for a quick pass.
</div>

```{r fit-all, cache=TRUE}
results <- list()
for (nm in names(cells)) {
  cl <- cells[[nm]]
  dd <- make_dml_data(cl$D, cl$X)
  results[[nm]] <- list(
    ranger  = fit_cell(dd, ml_rf(),          paste(nm, "ranger")),
    untuned = fit_cell(dd, ml_xgb_untuned(), paste(nm, "xgb-untuned")),
    tuned   = fit_cell(dd, make_tuned_xgb(), paste(nm, "xgb-tuned"))
  )
  cat(sprintf("done: %s\n", nm))
}
```

---

# Comparison

```{r comparison-table}
fmt <- function(r) if (is.null(r)) "—" else sprintf("%.3f (%.3f)", r$coef, r$se)
gv  <- function(r, f, d = 4) if (is.null(r)) NA else round(f(r), d)

rows <- do.call(rbind, lapply(names(results), function(nm) {
  R <- results[[nm]]
  data.frame(
    Cell = nm,
    `Estimate (SE)` = c(fmt(R$ranger), fmt(R$untuned), fmt(R$tuned)),
    Learner = c("ranger (ref)", "XGB untuned", "XGB tuned"),
    `Σω (C3 dev.)` = c(gv(R$ranger, function(x) x$sum_w),
                       gv(R$untuned, function(x) x$sum_w),
                       gv(R$tuned,  function(x) x$sum_w)),
    `ΣωY=τ̂` = c(if(is.null(R$ranger)) NA else R$ranger$ok,
                if(is.null(R$untuned)) NA else R$untuned$ok,
                if(is.null(R$tuned)) NA else R$tuned$ok),
    `prop. near 0/1` = c(prop_extremes(R$ranger), prop_extremes(R$untuned), prop_extremes(R$tuned)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}))

rows %>%
  kbl(align = c("l","c","l","r","c","r"),
      caption = "Wald-AIPW per cell: ranger reference vs XGBoost untuned vs tuned") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12.5) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(rows$Learner == "XGB tuned"), background = "#eef6ff") %>%
  footnote(general = "Σω is the departure from Condition 3 (C3 ⟺ Σω=0 for Wald-AIPW). ΣωY=τ̂ is the weight-extraction identity (Condition 1) and holds regardless of C3. 'prop. near 0/1' is the share of propensity predictions within 0.02 of the boundary.")
```

---

# The C3 Departure Diagnostic

<div class="note-box">
For Wald-AIPW, Condition 3 holds **iff** the outcome weights sum to zero. Ranger (affine) sits at Σω ≈ 0; XGBoost departs. The question is whether tuning the nuisances pulls XGBoost's |Σω| back toward the affine reference. This is the novel measurement: not "does tuning help the estimate" but "does tuning move XGBoost toward satisfying C3."
</div>

```{r c3-plot, fig.height=5}
plot_df <- do.call(rbind, lapply(names(results), function(nm) {
  R <- results[[nm]]
  data.frame(
    Cell    = nm,
    Learner = c("ranger (ref)", "XGB untuned", "XGB tuned"),
    abs_sum_w = c(if(is.null(R$ranger)) NA else abs(R$ranger$sum_w),
                  if(is.null(R$untuned)) NA else abs(R$untuned$sum_w),
                  if(is.null(R$tuned)) NA else abs(R$tuned$sum_w))
  )
}))
plot_df$Learner <- factor(plot_df$Learner,
                          levels = c("XGB untuned", "XGB tuned", "ranger (ref)"))

ggplot(plot_df, aes(x = Cell, y = abs_sum_w, fill = Learner)) +
  geom_col(position = position_dodge(0.75), width = 0.7, alpha = 0.9) +
  scale_fill_manual(values = c("XGB untuned" = "#d7191c",
                               "XGB tuned"   = "#fb8c00",
                               "ranger (ref)" = "#2c7bb6")) +
  labs(title = "Departure from Condition 3: |Σω| by learner",
       subtitle = "C3 holds at |Σω| = 0 (ranger). Lower = closer to the affine-smoother condition.",
       x = NULL, y = expression("|" * Sigma * omega * "|"), fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 15, hjust = 1),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 10))
```

```{r estimate-plot, fig.height=5}
est_df <- do.call(rbind, lapply(names(results), function(nm) {
  R <- results[[nm]]
  data.frame(
    Cell    = nm,
    Learner = c("ranger (ref)", "XGB untuned", "XGB tuned"),
    coef = c(if(is.null(R$ranger)) NA else R$ranger$coef,
             if(is.null(R$untuned)) NA else R$untuned$coef,
             if(is.null(R$tuned)) NA else R$tuned$coef)
  )
}))
est_df$Learner <- factor(est_df$Learner,
                         levels = c("XGB untuned", "XGB tuned", "ranger (ref)"))

ggplot(est_df, aes(x = Cell, y = coef, fill = Learner)) +
  geom_col(position = position_dodge(0.75), width = 0.7, alpha = 0.9) +
  geom_hline(yintercept = 0, colour = "grey50") +
  scale_fill_manual(values = c("XGB untuned" = "#d7191c",
                               "XGB tuned"   = "#fb8c00",
                               "ranger (ref)" = "#2c7bb6")) +
  labs(title = "Wald-AIPW point estimate by learner",
       subtitle = "Does tuning pull the XGBoost estimate toward the ranger reference?",
       x = NULL, y = "Estimate", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 15, hjust = 1),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 10))
```

<div class="note-box">
**Reading the results.** If tuning works as intended you should see, relative to untuned XGBoost: smaller `prop. near 0/1` (propensities pulled off the boundary), smaller `|Σω|` (closer to C3), and a point estimate moving toward the ranger reference and the other learners. The weight-extraction identity (`ΣωY=τ̂`) holds throughout — that is Condition 1, guaranteed by the four frozen parameters, and is independent of whether C3 holds. The honest takeaway for the thesis: tuning does not make XGBoost affine, but it measurably reduces the departure from C3 and stabilises the estimate, which is exactly what the weight-based diagnostic lets us quantify.
</div>

---

# Reuse for Vietnam / Angrist-Evans

```{r reuse-note, eval=FALSE}
# The two builders below are framework-agnostic. To apply to another design,
# point Y_dol / Z / D / X at that design's vectors and call:
#
#   dd  <- make_dml_data(D, X)
#   ref <- fit_cell(dd, ml_rf(),          "ranger")
#   unt <- fit_cell(dd, ml_xgb_untuned(), "xgb-untuned")
#   tun <- fit_cell(dd, make_tuned_xgb(), "xgb-tuned")
#
# Everything downstream (Σω as C3 departure, identity check, propensity extremes)
# is identical.
```
