---
title: "Card (1995) — XGBoost Nuisance Tuning for Wald-AIPW"
subtitle: "Nested nuisance tuning by predictive loss"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 5.5)
```

```{r packages}
library(haven)
library(knitr)
library(kableExtra)
library(ggplot2)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)
library(cobalt)
library(gridExtra)
library(viridis)

source("functions_all.R")

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")


SEED = 42 # I choose this see in order to be able to reproduce the results that I had in the card notebook
N_CF = 5
TUNE_EVALS = 8
INNER_FOLDS = 3
```

<div class="note-box">
**The problem.** The main Card results show that the default XGBoost Wald-AIPW estimate can be highly unstable. This notebook asks whether this instability is partly driven by poorly chosen nuisance learners.

**What tuning can and cannot do.**  XGBoost is tuned through nuisance-prediction performance.
For that RMSE is used the outcome regression and log-loss for the binary treatment and instrument models. 

If tuning moves the Wald AIPW estimate while also improving nuisance prediction losses, it can be seen as evidence, that the default XGBoost specification was numerically fragile.
</div>

---

# Setup

```{r data}
data_path = "/Users/alexung/Desktop/master_thesis/code/"
card  = read_dta(file.path(data_path, "card.dta"))
card  = card[!is.na(card$wage) & !is.na(card$educ), ]

card$somecol = as.integer(card$educ > 12)
card$educ16  = as.integer(card$educ >= 16)
if (!"expersq" %in% names(card)) card$expersq <- card$exper^2

card$lwage_cnt = log(card$wage)        # cents
card$lwage_dol = log(card$wage / 100)  # dollars

Y_cnt = card$lwage_cnt
Y_dol = card$lwage_dol
Z     = as.integer(card$nearc4)
D1    = card$somecol
D2    = card$educ16

X_dml_card = make_X(~ 0 + exper + expersq + black + south + smsa + smsa66 +
                       reg661+reg662+reg663+reg664+reg665+reg666+reg667+reg668, data = card)
X_dml_kit  = make_X(~ 0 + black + south + smsa + smsa66 + south66, data = card)

cat(sprintf("N=%d | Card %d cov | Kitagawa %d cov\n",
            nrow(card), ncol(X_dml_card), ncol(X_dml_kit)))
```

```{r learners}

make_dml_data = function(Y, D, Z, X) {
  df = as.data.frame(X)
  df$y = as.numeric(Y)
  df$d = as.numeric(D)
  df$z = as.numeric(Z)
  
  DoubleMLData$new(
    data = df,
    y_col = "y",
    d_cols = "d",
    z_cols = "z",
    x_cols = colnames(X)
  )
}

make_xgb_learners = function() {
  list(
    g = lrn(
      "regr.xgboost",
      alpha = 0,
      subsample = 1,
      max_delta_step = 0,
      base_score = 0
    ),
    m = lrn(
      "classif.xgboost",
      predict_type = "prob"
    ),
    r = lrn(
      "classif.xgboost",
      predict_type = "prob"
    )
  )
}





ml_xgb_untuned = make_xgb_learners
ml_xgb_base = make_xgb_learners



xgb_ps = function() {
  ps(
    nrounds = p_int(50, 400),
    max_depth = p_int(2, 5),
    eta = p_dbl(0.02, 0.20),
    min_child_weight = p_int(1, 8)
  )
}



tune_param_set = function() {
  list(
    ml_g = xgb_ps(),
    ml_m = xgb_ps(),
    ml_r = xgb_ps()
  )
}




tune_settings = function() {
  list(
    n_folds_tune = INNER_FOLDS,
    rsmp_tune = rsmp("cv", folds = INNER_FOLDS),
    measure = list(
      ml_g = "regr.rmse",
      ml_m = "classif.logloss",
      ml_r = "classif.logloss"
    ),
    terminator = trm("evals", n_evals = TUNE_EVALS),
    algorithm = tnr("random_search")
  )
}


get_prediction_vector = function(obj, learner_name) {
  pred = tryCatch(
    obj$predictions[[learner_name]],
    error = function(e) NULL
  )
  
  if (is.null(pred)) {
    return(numeric(0))
  }
  
  as.numeric(pred)
}


propensity_boundary_share = function(obj, threshold = 0.02) {
  m_hat = get_prediction_vector(obj, "ml_m")
  r_hat = get_prediction_vector(obj, "ml_r")
  
  p_hat = c(m_hat, r_hat)
  p_hat = p_hat[is.finite(p_hat)]
  
  if (length(p_hat) == 0) {
    return(NA_real_)
  }
  
  mean(p_hat < threshold | p_hat > 1 - threshold)
}

fit_cell = function(dml_data, learners, label, tune = FALSE) {
  tryCatch({
    set.seed(SEED)
    
    obj = DoubleMLIIVM$new(
      data = dml_data,
      ml_g = learners$g,
      ml_m = learners$m,
      ml_r = learners$r,
      n_folds = N_CF,
      score = "LATE"
    )
    
    if (tune) {
      obj$tune(
        param_set = tune_param_set(),
        tune_settings = tune_settings()
      )
    }
    
    obj$fit(store_models = TRUE, store_predictions = TRUE)
    
    tuning_result = tryCatch(
      obj$tuning_result,
      error = function(e) NULL
    )
    
    list(
      label = label,
      tuned = tune,
      coef = get_dml_coef(obj),
      se = get_dml_se(obj),
      boundary_share = propensity_boundary_share(obj),
      tuning_result = tuning_result,
      obj = obj
    )
  }, error = function(e) {
    message(sprintf("Fit failed for %s: %s", label, conditionMessage(e)))
    NULL
  })
}

cells = list(
  "Card / somecol" = list(Y = Y_dol, D = D1, Z = Z, X = X_dml_card),
  "Card / educ16" = list(Y = Y_dol, D = D2, Z = Z, X = X_dml_card),
  "Kitagawa / somecol" = list(Y = Y_dol, D = D1, Z = Z, X = X_dml_kit),
  "Kitagawa / educ16" = list(Y = Y_dol, D = D2, Z = Z, X = X_dml_kit)
)


```

---

# Fit: XGBoost untuned vs tuned

<div class="warn-box">
**Runtime.**  The tuned fits run inner cross-validation for three nuisance functions in each of the four Card cells.
</div>

```{r fit-all, cache=TRUE}

results = list()

for (cell_name in names(cells)) {
  cell = cells[[cell_name]]
  
  dml_data = make_dml_data(
    Y = cell$Y,
    D = cell$D,
    Z = cell$Z,
    X = cell$X
  )
  
  results[[cell_name]] = list(
    untuned = fit_cell(
      dml_data = dml_data,
      learners = ml_xgb_untuned(),
      label = paste(cell_name, "xgb-untuned"),
      tune = FALSE
    ),
    tuned = fit_cell(
      dml_data = dml_data,
      learners = ml_xgb_base(),
      label = paste(cell_name, "xgb-tuned"),
      tune = TRUE
    )
  )
  
  cat(sprintf("Finished: %s\n", cell_name))
}
```


---

# Comparison

```{r comparison-table}
fmt_est = function(r) {
  if (is.null(r)) {
    return("—")
  }
  sprintf("%.3f (%.3f)", r$coef, r$se)
}

fmt_boundary = function(r, digits = 4) {
  if (is.null(r) || is.null(r$boundary_share) || !is.finite(r$boundary_share)) {
    return("—")
  }
  sprintf(paste0("%.", digits, "f"), r$boundary_share)
}

rows = do.call(
  rbind,
  lapply(names(results), function(cell_name) {
    R = results[[cell_name]]
    
    data.frame(
      Cell = cell_name,
      Learner = c("XGB untuned", "XGB tuned"),
      `Estimate (SE)` = c(
        fmt_est(R$untuned),
        fmt_est(R$tuned)
      ),
      `Propensity boundary share` = c(
        fmt_boundary(R$untuned),
        fmt_boundary(R$tuned)
      ),
      check.names = FALSE,
      stringsAsFactors = FALSE
    )
  })
)

rows %>%
  kbl(
    align = c("l", "l", "c", "r"),
    caption = "Wald-AIPW estimates: XGBoost untuned vs tuned"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    font_size = 12.5
  ) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(rows$Learner == "XGB tuned"), background = "#eef6ff") %>%
  footnote(
    general = "The propensity boundary share is the fraction of fitted treatment or instrument propensities below 0.02 or above 0.98."
  )
```

---


```{r nuisance-loss-helpers}

get_prediction_vector = function(obj, learner_name) {
  pred = tryCatch(
    obj$predictions[[learner_name]],
    error = function(e) NULL
  )
  
  if (is.null(pred)) {
    return(numeric(0))
  }
  
  as.numeric(drop(pred))
}

get_branch_prediction = function(obj, base_name, Z) {
  pred0 = get_prediction_vector(obj, paste0(base_name, "0"))
  pred1 = get_prediction_vector(obj, paste0(base_name, "1"))
  
  if (length(pred0) != length(Z) || length(pred1) != length(Z)) {
    return(numeric(0))
  }
  
  ifelse(as.numeric(Z) == 1, pred1, pred0)
}

rmse_loss = function(truth, pred) {
  if (length(pred) != length(truth) || length(pred) == 0) {
    return(NA_real_)
  }
  
  truth = as.numeric(truth)
  pred = as.numeric(pred)
  
  keep = is.finite(truth) & is.finite(pred)
  
  if (!any(keep)) {
    return(NA_real_)
  }
  
  sqrt(mean((truth[keep] - pred[keep])^2))
}

logloss_loss = function(truth, pred, eps = 1e-6) {
  if (length(pred) != length(truth) || length(pred) == 0) {
    return(NA_real_)
  }
  
  truth = as.numeric(truth)
  pred = as.numeric(pred)
  
  keep = is.finite(truth) & is.finite(pred)
  
  if (!any(keep)) {
    return(NA_real_)
  }
  
  truth = truth[keep]
  pred = pred[keep]
  pred = pmin(pmax(pred, eps), 1 - eps)
  
  -mean(truth * log(pred) + (1 - truth) * log(1 - pred))
}

calc_nuisance_losses = function(r, cell) {
  if (is.null(r)) {
    return(c(
      outcome_rmse = NA_real_,
      instrument_logloss = NA_real_,
      treatment_logloss = NA_real_
    ))
  }
  
  g_hat = get_branch_prediction(r$obj, "ml_g", cell$Z)
  m_hat = get_prediction_vector(r$obj, "ml_m")
  r_hat = get_branch_prediction(r$obj, "ml_r", cell$Z)
  
  c(
    outcome_rmse = rmse_loss(cell$Y, g_hat),
    instrument_logloss = logloss_loss(cell$Z, m_hat),
    treatment_logloss = logloss_loss(cell$D, r_hat)
  )
}
```



---

## 2. Nuisance-loss  plot




```{r nuisance-loss-plot, fig.width=9, fig.height=5}
loss_rows = do.call(
  rbind,
  lapply(names(results), function(cell_name) {
    R = results[[cell_name]]
    cell = cells[[cell_name]]
    
    untuned_losses = calc_nuisance_losses(R$untuned, cell)
    tuned_losses = calc_nuisance_losses(R$tuned, cell)
    
    data.frame(
      Cell = cell_name,
      Learner = c("XGB untuned", "XGB tuned"),
      Outcome_RMSE = c(
        untuned_losses["outcome_rmse"],
        tuned_losses["outcome_rmse"]
      ),
      Instrument_logloss = c(
        untuned_losses["instrument_logloss"],
        tuned_losses["instrument_logloss"]
      ),
      Treatment_logloss = c(
        untuned_losses["treatment_logloss"],
        tuned_losses["treatment_logloss"]
      ),
      check.names = FALSE,
      stringsAsFactors = FALSE
    )
  })
)

loss_long = rbind(
  data.frame(
    Cell = loss_rows$Cell,
    Learner = loss_rows$Learner,
    Metric = "Outcome RMSE",
    Loss = loss_rows$Outcome_RMSE
  ),
  data.frame(
    Cell = loss_rows$Cell,
    Learner = loss_rows$Learner,
    Metric = "Instrument log-loss",
    Loss = loss_rows$Instrument_logloss
  ),
  data.frame(
    Cell = loss_rows$Cell,
    Learner = loss_rows$Learner,
    Metric = "Treatment log-loss",
    Loss = loss_rows$Treatment_logloss
  )
)

loss_long = loss_long[is.finite(loss_long$Loss), ]

ggplot(
  loss_long,
  aes(x = Loss, y = Cell, color = Learner, shape = Learner)
) +
  geom_line(aes(group = Cell), color = "grey75", linewidth = 0.4) +
  geom_point(size = 2.8) +
  facet_wrap(~ Metric, scales = "free_x") +
  scale_color_manual(
    values = c(
      "XGB untuned" = "#7f7f7f",
      "XGB tuned" = "#1f77b4"
    )
  ) +
  labs(
    title = "Nuisance prediction losses before and after tuning",
    subtitle = "Lower values indicate better nuisance prediction",
    x = "Prediction loss",
    y = NULL,
    color = NULL,
    shape = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom"
  )
```




---




## 3. Hyperparameter summary

```{r hyperparameter-summary}
get_param_source = function(r) {
  if (is.null(r)) {
    return(NULL)
  }
  
  sources = list(
    tuning_result = tryCatch(r$tuning_result, error = function(e) NULL),
    obj_params = tryCatch(r$obj$params, error = function(e) NULL)
  )
  
  for (s in sources) {
    if (!is.null(s)) {
      u = unlist(s, recursive = TRUE)
      if (any(grepl("nrounds|max_depth|eta|min_child_weight", names(u)))) {
        return(s)
      }
    }
  }
  
  NULL
}

extract_tuned_params = function(r, cell_name) {
  param_source = get_param_source(r)
  
  if (is.null(param_source)) {
    return(data.frame())
  }
  
  u = unlist(param_source, recursive = TRUE)
  u = u[grepl("nrounds|max_depth|eta|min_child_weight", names(u))]
  
  if (length(u) == 0) {
    return(data.frame())
  }
  
  hp = data.frame(
    Cell = cell_name,
    Raw_name = names(u),
    Value = as.numeric(u),
    stringsAsFactors = FALSE
  )
  
  hp$Nuisance = ifelse(
    grepl("ml_g", hp$Raw_name), "Outcome g",
    ifelse(
      grepl("ml_m", hp$Raw_name), "Instrument m",
      ifelse(grepl("ml_r", hp$Raw_name), "Treatment r", "Unknown")
    )
  )
  
  hp$Branch = ifelse(
    grepl("ml_g0|ml_r0", hp$Raw_name), "Z = 0",
    ifelse(
      grepl("ml_g1|ml_r1", hp$Raw_name), "Z = 1",
      ""
    )
  )
  
  hp$Parameter = sub(".*(nrounds|max_depth|eta|min_child_weight).*", "\\1", hp$Raw_name)
  
  hp
}

hp_raw = do.call(
  rbind,
  lapply(names(results), function(cell_name) {
    extract_tuned_params(results[[cell_name]]$tuned, cell_name)
  })
)

if (nrow(hp_raw) > 0) {
  hp_raw = hp_raw[hp_raw$Nuisance != "Unknown", ]
  
  # If DoubleML does not label the two branches explicitly, give repeated
  # nuisance parameters a simple branch number.
  missing_branch = hp_raw$Branch == ""
  
  hp_raw$Branch_id = ave(
    hp_raw$Value,
    hp_raw$Cell,
    hp_raw$Nuisance,
    hp_raw$Parameter,
    FUN = seq_along
  )
  
  hp_raw$Branch[missing_branch & hp_raw$Nuisance %in% c("Outcome g", "Treatment r")] =
    paste0("Branch ", hp_raw$Branch_id[missing_branch & hp_raw$Nuisance %in% c("Outcome g", "Treatment r")])
  
  hp_table = hp_raw[, c("Cell", "Nuisance", "Branch", "Parameter", "Value")]
  hp_table = hp_table[order(hp_table$Cell, hp_table$Nuisance, hp_table$Branch, hp_table$Parameter), ]
  
  hp_table$Value = ifelse(
    hp_table$Parameter == "eta",
    sprintf("%.3f", hp_table$Value),
    sprintf("%.0f", hp_table$Value)
  )
  
  rownames(hp_table) = NULL
  
  hp_table %>%
    kbl(
      align = c("l", "l", "l", "l", "r"),
      caption = "Selected XGBoost hyperparameters after nuisance tuning"
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = TRUE,
      font_size = 12.5
    ) %>%
    collapse_rows(columns = c(1, 2, 3), valign = "middle") %>%
    footnote(
      general = "The selected values are chosen by inner cross-validation on nuisance prediction loss."
    )
} else {
  cat("No hyperparameters could be extracted automatically.")
}
```



```{r save-tuning-results}
new_path = "/Users/alexung/Desktop/master_thesis/juicy/card_tuning"
output_dir = file.path(new_path, "outputs")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

make_estimate_export = function(cell_name, learner_name, r) {
  if (is.null(r)) {
    return(data.frame(
      Cell = cell_name,
      Learner = learner_name,
      Estimate = NA_real_,
      SE = NA_real_,
      Boundary_share = NA_real_
    ))
  }
  
  data.frame(
    Cell = cell_name,
    Learner = learner_name,
    Estimate = as.numeric(r$coef),
    SE = as.numeric(r$se),
    Boundary_share = as.numeric(r$boundary_share)
  )
}

estimate_export = do.call(
  rbind,
  lapply(names(results), function(cell_name) {
    R = results[[cell_name]]
    
    rbind(
      make_estimate_export(cell_name, "XGB untuned", R$untuned),
      make_estimate_export(cell_name, "XGB tuned", R$tuned)
    )
  })
)

tuning_export = list(
  settings = list(
    seed = SEED,
    n_cf = N_CF,
    tune_evals = TUNE_EVALS,
    inner_folds = INNER_FOLDS,
    outcome = "log(wage / 100)",
    estimator = "DoubleMLIIVM Wald-AIPW",
    learner = "xgboost",
    tuning_target = "nuisance prediction loss"
  ),
  estimates = estimate_export,
  estimates_print = if (exists("rows")) rows else NULL,
  nuisance_losses = if (exists("loss_rows")) loss_rows else NULL,
  hyperparameters = if (exists("hp_table")) hp_table else NULL,
  hyperparameters_raw = if (exists("hp_raw")) hp_raw else NULL
)

#saveRDS(
#  tuning_export,
#  file = file.path(output_dir, "card_xgb_tuning_summary.rds")
#)




#cat("Saved tuning files in:", output_dir, "\n")

# In my presentation file I can then read it like : 
# xgb_tuning = readRDS(file.path(data_path, "outputs", "card_xgb_tuning_summary.rds"))
# and access:
# xgb_tuning$estimates
# xgb_tuning$hyperparameters
# xgb_tuning$settings

```

## 4. Diagnostics for untuned vs tuned
```{r xgb-tuned-weight-table, cache=TRUE}
# Compact tuned-vs-untuned outcome-weight diagnostics for XGBoost.
# This chunk does NOT use make_dml_data(), to avoid conflicts with other notebooks.

make_dml_data_xgb_diag = function(D, X) {
  df = as.data.frame(X)
  df$y = as.numeric(Y_dol)
  df$d = as.numeric(D)
  df$z = as.numeric(Z)
  
  DoubleMLData$new(
    data = df,
    y_col = "y",
    d_cols = "d",
    z_cols = "z",
    x_cols = colnames(X)
  )
}

omega_vec = function(om) {
  w = om$omega
  
  if (is.null(dim(w))) {
    return(as.vector(w))
  }
  
  if (!is.null(rownames(w)) && "Wald-AIPW" %in% rownames(w)) {
    return(as.vector(w["Wald-AIPW", ]))
  }
  
  as.vector(w[1, ])
}

extract_xgb_w = function(r, cell) {
  # If you later store w inside fit_cell(), this line reuses it immediately.
  if (!is.null(r$w)) {
    return(as.vector(r$w))
  }
  
  dml_data = make_dml_data_xgb_diag(
    D = cell$D,
    X = cell$X
  )
  
  om = get_outcome_weights(
    object = r$obj,
    dml_data = dml_data
  )
  
  omega_vec(om)
}

one_xgb_diag_row = function(cell_name, learner_name, r, cell) {
  w = extract_xgb_w(r, cell)
  
  wb = weight_bridge(w, cell$D, learner_name)
  
  ov = wb[wb$Scope == "Overall", ]
  tr = wb[wb$Scope == "Treated", ]
  co = wb[wb$Scope == "Control", ]
  
  data.frame(
    Cell = cell_name,
    Learner = learner_name,
    Estimate = as.numeric(r$coef),
    ESS = ov$ESS_kish,
    ESS_mod = ov$ESS_mod,
    `Sigma omega` = ov$Sum_w,
    `Sigma abs omega` = ov$Sum_abs_w,
    `Treated mass` = tr$Sum_w,
    `Control mass` = co$Sum_w,
    `max abs omega` = max(abs(w)),
    check.names = FALSE
  )
}

xgb_tuned_weight_table = do.call(
  rbind,
  lapply(names(cells), function(cell_name) {
    cell = cells[[cell_name]]
    R = results[[cell_name]]
    
    rbind(
      one_xgb_diag_row(cell_name, "XGB untuned", R$untuned, cell),
      one_xgb_diag_row(cell_name, "XGB tuned", R$tuned, cell)
    )
  })
)

xgb_tuned_weight_table_print = xgb_tuned_weight_table

xgb_tuned_weight_table_print$Estimate = sprintf("%.3f", xgb_tuned_weight_table_print$Estimate)
xgb_tuned_weight_table_print$ESS = sprintf("%.0f", xgb_tuned_weight_table_print$ESS)
xgb_tuned_weight_table_print$ESS_mod = sprintf("%.0f", xgb_tuned_weight_table_print$ESS_mod)
xgb_tuned_weight_table_print$`Sigma omega` = sprintf("%.4f", xgb_tuned_weight_table_print$`Sigma omega`)
xgb_tuned_weight_table_print$`Sigma abs omega` = sprintf("%.3f", xgb_tuned_weight_table_print$`Sigma abs omega`)
xgb_tuned_weight_table_print$`Treated mass` = sprintf("%.3f", xgb_tuned_weight_table_print$`Treated mass`)
xgb_tuned_weight_table_print$`Control mass` = sprintf("%.3f", xgb_tuned_weight_table_print$`Control mass`)
xgb_tuned_weight_table_print$`max abs omega` = sprintf("%.3f", xgb_tuned_weight_table_print$`max abs omega`)

xgb_tuned_weight_table_print %>%
  kbl(
    caption = "XGBoost tuning and outcome-weight diagnostics",
    align = c("l", "l", "r", "r", "r", "r", "r", "r", "r", "r")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    font_size = 12
  ) %>%
  row_spec(
    which(xgb_tuned_weight_table$Learner == "XGB tuned"),
    background = "#eef6ff"
  ) %>%
  footnote(
    general = "The control mass uses the sign-flipped convention from weight_bridge(), so treated mass and control mass should both be close to one for a normalized treated-control contrast. ESS_mod is the modified effective sample size for signed outcome weights."
  )
```



## 5 Love plots 

```{r xgb-tuned-love-card-educ16, fig.width=11, fig.height=5.5}
w_xgb_card_educ16_untuned = extract_xgb_w(
  results[["Card / educ16"]]$untuned,
  cells[["Card / educ16"]]
)

w_xgb_card_educ16_tuned = extract_xgb_w(
  results[["Card / educ16"]]$tuned,
  cells[["Card / educ16"]]
)

grid.arrange(
  make_love(
    "XGBoost untuned\nCard specification, educ16",
    w_xgb_card_educ16_untuned,
    D2,
    X_dml_card
  ),
  make_love(
    "XGBoost tuned\nCard specification, educ16",
    w_xgb_card_educ16_tuned,
    D2,
    X_dml_card
  ),
  nrow = 1,
  top = "XGBoost Wald-AIPW covariate balance before and after tuning"
)
```














