---
title: "Angrist & Evans (1998) - Stratified 3,000-Row Subsamples"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, message=FALSE, warning=FALSE}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(AER)
library(sandwich)
library(lmtest)

source("functions_all.R")

N_sub <- 3000L
final_seed <- 202L
diagnostic_seeds <- 202L:251L

paths <- c("ae98.dta", "/Users/alexung/Desktop/master_thesis/code/ae98.dta")
data_file <- paths[file.exists(paths)][1]
if (is.na(data_file)) stop("Could not find ae98.dta.")
```

# 1. Analysis Samples

The Angrist-Evans application uses two analysis samples. Labor-force
participation is defined on the full sample, while log income is defined only for
strictly positive reported income. I therefore construct one stratified 3,000-row
subsample for each outcome family.

```{r samples}
ae98_raw <- read_dta(data_file)

base_vars <- c("workedm", "morekids", "samesex", "agem", "agefstm",
               "sexk", "sex2nd", "blackm", "hispm", "othracem")

ae98_labor <- ae98_raw %>%
  filter(if_all(all_of(base_vars), ~ !is.na(.x))) %>%
  mutate(
    row_id = row_number(),
    workedm = as.integer(workedm),
    morekids = as.integer(morekids),
    samesex = as.integer(samesex),
    incomepos = as.integer(!is.na(incomem) & incomem > 0),
    race = case_when(
      blackm == 1 ~ "black",
      hispm == 1 ~ "hispanic",
      othracem == 1 ~ "other",
      TRUE ~ "white"
    ),
    age_bin = cut(agem, breaks = c(20, 25, 30, 35), include.lowest = TRUE)
  )

ae98_income <- ae98_labor %>%
  filter(incomepos == 1) %>%
  mutate(
    lincomem = log(incomem),
    inc_decile = ntile(incomem, 10)
  )

data.frame(
  Sample = c("Labor-force participation", "Log income"),
  `Parent N` = c(nrow(ae98_labor), nrow(ae98_income)),
  `Subsample N` = N_sub,
  check.names = FALSE
) %>%
  kbl(caption = "Parent samples and target subsample sizes",
      align = c("l", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

# 2. Stratified Draw

The final sample is a single proportional stratified draw with one fixed seed. No
optimization over alternative seeds is used for the final thesis sample.

```{r sampler}
largest_remainder_n <- function(n_full, n_target) {
  raw <- n_target * n_full / sum(n_full)
  n <- floor(raw)
  left <- n_target - sum(n)
  if (left > 0) {
    idx <- order(raw - n, decreasing = TRUE)[seq_len(left)]
    n[idx] <- n[idx] + 1L
  }
  pmin(n, n_full)
}

make_sampler <- function(df, strata, n_target) {
  key <- interaction(df[, strata, drop = FALSE], drop = TRUE, lex.order = TRUE)
  split_idx <- split(seq_len(nrow(df)), key)
  n_take <- largest_remainder_n(lengths(split_idx), n_target)

  function(seed) {
    set.seed(seed)
    idx <- unlist(Map(function(ii, nn) sample(ii, nn), split_idx, n_take),
                  use.names = FALSE)
    df[idx, , drop = FALSE]
  }
}

labor_sampler <- make_sampler(
  ae98_labor,
  c("samesex", "morekids", "workedm", "incomepos", "race", "age_bin"),
  N_sub
)

income_sampler <- make_sampler(
  ae98_income,
  c("samesex", "morekids", "inc_decile", "race", "age_bin"),
  N_sub
)

ae98_labor_sub <- labor_sampler(final_seed)
ae98_income_sub <- income_sampler(final_seed)

data.frame(
  Sample = c("Labor-force participation", "Log income"),
  Seed = final_seed,
  N = c(nrow(ae98_labor_sub), nrow(ae98_income_sub)),
  check.names = FALSE
) %>%
  kbl(caption = "Final honest stratified subsamples",
      align = c("l", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

# 3. Design Diagnostics

```{r design-diagnostics}
first_stage <- function(df) {
  mean(df$morekids[df$samesex == 1]) - mean(df$morekids[df$samesex == 0])
}

reduced_form <- function(df, y) {
  mean(y[df$samesex == 1]) - mean(y[df$samesex == 0])
}

fit_p <- function(df) {
  rhs <- "agem + agefstm + sexk + sex2nd + blackm + hispm + othracem"
  fit <- glm(as.formula(paste("samesex ~", rhs)), data = df, family = binomial())
  as.numeric(fitted(fit))
}

moment_labor <- function(df) {
  p_hat <- fit_p(df)
  c(
    pZ = mean(df$samesex),
    pD = mean(df$morekids),
    first_stage = first_stage(df),
    pWorked = mean(df$workedm),
    reduced_work = reduced_form(df, df$workedm),
    pIncomePos = mean(df$incomepos),
    age = mean(df$agem),
    age_sd = sd(df$agem),
    age_first = mean(df$agefstm),
    pBoy1 = mean(df$sexk),
    pBoy2 = mean(df$sex2nd),
    pBlack = mean(df$blackm),
    pHisp = mean(df$hispm),
    pOther = mean(df$othracem),
    pscore_mean = mean(p_hat),
    pscore_sd = sd(p_hat)
  )
}

moment_income <- function(df) {
  p_hat <- fit_p(df)
  qs <- quantile(df$lincomem, probs = c(.10, .25, .50, .75, .90), names = FALSE)
  c(
    pZ = mean(df$samesex),
    pD = mean(df$morekids),
    first_stage = first_stage(df),
    reduced_log_income = reduced_form(df, df$lincomem),
    log_income = mean(df$lincomem),
    log_income_sd = sd(df$lincomem),
    log_income_q10 = qs[1],
    log_income_q25 = qs[2],
    log_income_q50 = qs[3],
    log_income_q75 = qs[4],
    log_income_q90 = qs[5],
    age = mean(df$agem),
    age_sd = sd(df$agem),
    age_first = mean(df$agefstm),
    pBoy1 = mean(df$sexk),
    pBoy2 = mean(df$sex2nd),
    pBlack = mean(df$blackm),
    pHisp = mean(df$hispm),
    pOther = mean(df$othracem),
    pscore_mean = mean(p_hat),
    pscore_sd = sd(p_hat)
  )
}

diagnostic_table <- function(parent, sub, label, moment_fn) {
  target <- moment_fn(parent)
  selected <- moment_fn(sub)
  data.frame(
    Sample = label,
    Moment = names(target),
    Parent = as.numeric(target),
    Subsample = as.numeric(selected),
    Difference = as.numeric(selected - target),
    check.names = FALSE
  )
}

diag_design <- bind_rows(
  diagnostic_table(ae98_labor, ae98_labor_sub, "Labor / workedm", moment_labor),
  diagnostic_table(ae98_income, ae98_income_sub, "Income / lincomem", moment_income)
)

diag_design %>%
  mutate(across(where(is.numeric), ~ round(.x, 5))) %>%
  kbl(caption = "Design diagnostics: parent samples vs. final subsamples",
      align = c("l", "l", "r", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE)
```

# 4. Estimator Variability Across Honest Draws

The table and figure below do not choose the final sample. They repeat the same
stratified design over 50 pre-specified seeds to show the sampling variability one
should expect when reducing the sample size to 3,000 observations.

```{r estimator-functions}
RHS <- "agem + agefstm + sexk + sex2nd + blackm + hispm + othracem"

estimate_methods <- function(df, y, sample_label, seed = NA_integer_) {
  X <- model.matrix(as.formula(paste("~", RHS)), data = df)
  X_df <- df[, c("agem", "agefstm", "sexk", "sex2nd",
                 "blackm", "hispm", "othracem")]
  tsls <- run_2sls(y, df$morekids, df$samesex, X_df, "morekids")
  k <- kappa_analytic_se_all(y, df$samesex, df$morekids, X)

  data.frame(
    Sample = sample_label,
    Seed = seed,
    Method = c("2SLS", "tau_cb_u", "tau_ml_u", "tau_ml_a10",
               "tau_ml_a", "tau_ml_t", "tau_ml_a0"),
    Estimate = unname(c(tsls$coef,
                 k$estimates["tau_cb_u"], k$estimates["tau_ml_u"],
                 k$estimates["tau_ml_a10"], k$estimates["tau_ml_a"],
                 k$estimates["tau_ml_t"], k$estimates["tau_ml_a0"])),
    SE = unname(c(tsls$se,
           k$se["tau_cb_u"], k$se["tau_ml_u"], k$se["tau_ml_a10"],
           k$se["tau_ml_a"], k$se["tau_ml_t"], k$se["tau_ml_a0"])),
    check.names = FALSE
  )
}

full_estimates <- bind_rows(
  estimate_methods(ae98_labor, ae98_labor$workedm, "Labor / workedm"),
  estimate_methods(ae98_income, ae98_income$lincomem, "Income / lincomem")
) %>%
  mutate(Type = "Full sample")

final_estimates <- bind_rows(
  estimate_methods(ae98_labor_sub, ae98_labor_sub$workedm,
                   "Labor / workedm", final_seed),
  estimate_methods(ae98_income_sub, ae98_income_sub$lincomem,
                   "Income / lincomem", final_seed)
) %>%
  mutate(Type = "Final subsample")
```

```{r estimator-table}
estimator_table <- bind_rows(full_estimates, final_estimates) %>%
  mutate(
    `Estimate (SE)` = sprintf("%.3f (%.3f)", Estimate, SE)
  ) %>%
  select(Type, Sample, Seed, Method, `Estimate (SE)`)

estimator_table %>%
  kbl(caption = "Full-sample and final-subsample estimates with standard errors",
      align = c("l", "l", "r", "l", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE)
```

```{r diagnostic-draws}
diagnostic_estimates <- bind_rows(lapply(diagnostic_seeds, function(s) {
  labor_s <- labor_sampler(s)
  income_s <- income_sampler(s)
  bind_rows(
    estimate_methods(labor_s, labor_s$workedm, "Labor / workedm", s),
    estimate_methods(income_s, income_s$lincomem, "Income / lincomem", s)
  )
}))

diagnostic_summary <- diagnostic_estimates %>%
  group_by(Sample, Method) %>%
  summarise(
    Mean = mean(Estimate),
    SD_across_draws = sd(Estimate),
    Mean_SE = mean(SE),
    P05 = quantile(Estimate, .05),
    P95 = quantile(Estimate, .95),
    .groups = "drop"
  )

diagnostic_summary %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  kbl(caption = "Distribution of estimates across 50 honest stratified draws",
      align = c("l", "l", "r", "r", "r", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE)
```

```{r diagnostic-plot, fig.width=12, fig.height=7}
plot_methods <- c("2SLS", "tau_cb_u", "tau_ml_a")

plot_df <- diagnostic_estimates %>%
  filter(Method %in% plot_methods) %>%
  mutate(
    lower = Estimate - 1.96 * SE,
    upper = Estimate + 1.96 * SE,
    Method = factor(Method, levels = plot_methods)
  )

full_lines <- full_estimates %>%
  filter(Method %in% plot_methods) %>%
  mutate(Method = factor(Method, levels = plot_methods))

ggplot(plot_df, aes(x = Seed, y = Estimate)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), alpha = .25, width = .15) +
  geom_point(size = 1.7, alpha = .8) +
  geom_hline(data = full_lines, aes(yintercept = Estimate),
             colour = "#d7191c", linewidth = .7) +
  facet_grid(Sample ~ Method, scales = "free_y") +
  labs(title = "Subsample estimator variability across 50 honest stratified draws",
       subtitle = "Points are estimates; vertical intervals are +/- 1.96 SE. Red lines mark full-sample estimates.",
       x = "Seed", y = "Estimate") +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40"))
```

# 5. Save Output

```{r save}
save_obj <- list(
  method = "single honest proportional stratified draw",
  N_sub = N_sub,
  final_seed = final_seed,
  diagnostic_seeds = diagnostic_seeds,
  source_file = data_file,
  labor_parent_N = nrow(ae98_labor),
  income_parent_N = nrow(ae98_income),
  ae98_labor_sub = ae98_labor_sub,
  ae98_income_sub = ae98_income_sub,
  ae98_sub = ae98_labor_sub,
  diagnostics = list(
    design = diag_design,
    full_estimates = full_estimates,
    final_estimates = final_estimates,
    diagnostic_summary = diagnostic_summary
  )
)

saveRDS(save_obj, "ae98_sub.rds")
```
