---
title: "Angrist & Evans (1998) - Descriptive Diagnostics"
subtitle: "Selected 3,000-row diagnostic subsamples"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1 { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 34px; }
h2 { color: #2c7bb6; margin-top: 24px; }
.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
.warn-box {
  background: #fff3e0;
  border-left: 4px solid #fb8c00;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 5.2, dpi = 110)
```

```{r packages}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)

col_z1 <- "#2c7bb6"
col_z0 <- "#d7191c"
col_grey <- "#4d4d4d"
```

```{r data}
obj <- readRDS("ae98_sub.rds")
if (!all(c("ae98_labor_sub", "ae98_income_sub") %in% names(obj))) {
  stop("ae98_sub.rds must contain ae98_labor_sub and ae98_income_sub. Run ae98_make_balanced_subsample_submission.Rmd first.")
}

ae98_labor_sub <- obj$ae98_labor_sub
ae98_income_sub <- obj$ae98_income_sub

ae98_labor_sub <- ae98_labor_sub %>%
  mutate(
    workedm = as.integer(workedm),
    morekids = as.integer(morekids),
    samesex = as.integer(samesex),
    incomepos = as.integer(!is.na(incomem) & incomem > 0)
  )

ae98_income_sub <- ae98_income_sub %>%
  mutate(
    workedm = as.integer(workedm),
    morekids = as.integer(morekids),
    samesex = as.integer(samesex),
    lincomem = log(incomem)
  )
```

<div class="note-box">
Following Słoczyński, Uysal, and Wooldridge, the Angrist-Evans application uses two analysis samples. The labor-force-participation outcomes use the full AE labor sample, while log-income outcomes use the positive-income sample. The computational diagnostics therefore use two separate stratified diagnostic subsamples: one 3,000-row labor subsample and one 3,000-row income subsample.
</div>

# 1. Child Dataset: Instrument, Treatment, and Outcomes

```{r descriptive-helpers}
first_stage <- function(df) {
  mean(df$morekids[df$samesex == 1]) - mean(df$morekids[df$samesex == 0])
}

reduced_form <- function(df, y) {
  mean(y[df$samesex == 1]) - mean(y[df$samesex == 0])
}

fmt_num <- function(x) {
  ifelse(is.na(x), "--",
         ifelse(abs(x) >= 1000, format(round(x), big.mark = ",", scientific = FALSE),
                sprintf("%.4f", x)))
}
```

```{r zd-cells}
cell_table <- function(df, label) {
  df %>%
    count(samesex, morekids, name = "N") %>%
    mutate(
      Sample = label,
      Share = N / sum(N),
      Cell = paste0("Z=", samesex, ", D=", morekids)
    ) %>%
    select(Sample, Cell, N, Share)
}

zd_cells <- bind_rows(
  cell_table(ae98_labor_sub, "Labor sub"),
  cell_table(ae98_income_sub, "Income sub")
)

zd_cells %>%
  mutate(Share = sprintf("%.3f", Share)) %>%
  kbl(caption = "Instrument-treatment cells in the selected subsamples",
      align = c("l", "l", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13) %>%
  pack_rows("Labor subsample", 1, 4) %>%
  pack_rows("Income subsample", 5, 8) %>%
  footnote(general = "Z = samesex; D = morekids. These four cells determine the raw compliance structure.")
```

```{r zd-share-plot}
zd_share_plot <- bind_rows(
  ae98_labor_sub %>%
    summarise(Sample = "Labor sub",
              `Z = 1 (samesex)` = mean(samesex),
              `D = 1 (morekids)` = mean(morekids)),
  ae98_income_sub %>%
    summarise(Sample = "Income sub",
              `Z = 1 (samesex)` = mean(samesex),
              `D = 1 (morekids)` = mean(morekids))
) %>%
  pivot_longer(-Sample, names_to = "Measure", values_to = "Share")

ggplot(zd_share_plot, aes(x = Measure, y = Share, fill = Measure)) +
  geom_col(width = 0.58, alpha = 0.92) +
  geom_text(aes(label = percent(Share, accuracy = 0.1)),
            vjust = -0.35, size = 3.4) +
  facet_wrap(~ Sample) +
  scale_fill_manual(values = c("Z = 1 (samesex)" = col_z1,
                               "D = 1 (morekids)" = col_grey)) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, .62)) +
  labs(title = "Instrument and treatment shares",
       x = NULL, y = "Share", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 15, hjust = 1),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))
```

```{r iv-summary}
iv_summary <- function(df, y, label, outcome_label) {
  d_z1 <- mean(df$morekids[df$samesex == 1])
  d_z0 <- mean(df$morekids[df$samesex == 0])
  y_z1 <- mean(y[df$samesex == 1])
  y_z0 <- mean(y[df$samesex == 0])
  fs <- d_z1 - d_z0
  rf <- y_z1 - y_z0

  data.frame(
    Sample = label,
    Outcome = outcome_label,
    N = nrow(df),
    `Pr(Z = 1)` = mean(df$samesex),
    `Pr(D = 1)` = mean(df$morekids),
    `D | Z = 0` = d_z0,
    `D | Z = 1` = d_z1,
    `First stage` = fs,
    `Y | Z = 0` = y_z0,
    `Y | Z = 1` = y_z1,
    `Reduced form` = rf,
    `Wald ratio` = rf / fs,
    check.names = FALSE
  )
}

iv_summary_tab <- bind_rows(
  iv_summary(ae98_labor_sub, ae98_labor_sub$workedm,
             "Labor sub", "workedm"),
  iv_summary(ae98_income_sub, ae98_income_sub$lincomem,
             "Income sub", "log income")
)

iv_summary_tab %>%
  mutate(across(where(is.numeric), fmt_num)) %>%
  kbl(caption = "Key IV descriptives in the selected subsamples",
      align = c("l", "l", rep("r", 10))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 12.5) %>%
  footnote(general = "The first stage is E[D|Z=1] - E[D|Z=0]. The reduced form is E[Y|Z=1] - E[Y|Z=0]. The Wald ratio is the raw reduced form divided by the raw first stage.")
```

```{r first-stage-plot}
first_stage_plot <- bind_rows(
  ae98_labor_sub %>%
    group_by(samesex) %>%
    summarise(Sample = "Labor sub", Rate = mean(morekids), .groups = "drop"),
  ae98_income_sub %>%
    group_by(samesex) %>%
    summarise(Sample = "Income sub", Rate = mean(morekids), .groups = "drop")
) %>%
  mutate(`Z group` = ifelse(samesex == 1, "Z = 1 (same sex)", "Z = 0 (different sex)"))

first_stage_labels <- first_stage_plot %>%
  select(Sample, samesex, Rate) %>%
  pivot_wider(names_from = samesex, values_from = Rate, names_prefix = "z") %>%
  mutate(Gap = z1 - z0,
         Label = paste0("First stage = ", sprintf("%.3f", Gap)))

ggplot(first_stage_plot, aes(x = `Z group`, y = Rate, fill = `Z group`)) +
  geom_col(width = 0.58, alpha = 0.92) +
  geom_text(aes(label = percent(Rate, accuracy = 0.1)),
            vjust = -0.35, size = 3.4) +
  geom_text(data = first_stage_labels,
            aes(x = 1.5, y = pmax(z0, z1) + .055, label = Label),
            inherit.aes = FALSE, size = 3.3, colour = col_grey) +
  facet_wrap(~ Sample) +
  scale_fill_manual(values = c("Z = 1 (same sex)" = col_z1,
                               "Z = 0 (different sex)" = col_z0)) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, .45)) +
  labs(title = "First stage: morekids rate by sibling-sex composition",
       x = NULL, y = "Pr(morekids = 1)", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        panel.grid.major.x = element_blank(),
        axis.text.x = element_blank(),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))
```

```{r outcome-by-z}
outcome_by_z <- bind_rows(
  ae98_labor_sub %>%
    group_by(samesex) %>%
    summarise(
      Sample = "Labor sub",
      Outcome = "workedm",
      N = n(),
      Mean = mean(workedm),
      SD = sd(workedm),
      .groups = "drop"
    ),
  ae98_income_sub %>%
    group_by(samesex) %>%
    summarise(
      Sample = "Income sub",
      Outcome = "log income",
      N = n(),
      Mean = mean(lincomem),
      SD = sd(lincomem),
      .groups = "drop"
    )
) %>%
  mutate(`Z group` = ifelse(samesex == 1, "Z = 1 (same sex)", "Z = 0 (different sex)")) %>%
  select(Sample, Outcome, `Z group`, N, Mean, SD)

outcome_by_z %>%
  mutate(across(c(Mean, SD), ~ round(.x, 4))) %>%
  kbl(caption = "Outcome distribution by instrument status",
      align = c("l", "l", "l", "r", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) %>%
  footnote(general = "The mean differences across Z are the raw reduced forms summarized above.")
```

```{r worked-outcome-plot}
worked_plot <- ae98_labor_sub %>%
  group_by(samesex) %>%
  summarise(Rate = mean(workedm), .groups = "drop") %>%
  mutate(`Z group` = ifelse(samesex == 1, "Z = 1 (same sex)", "Z = 0 (different sex)"))

ggplot(worked_plot, aes(x = `Z group`, y = Rate, fill = `Z group`)) +
  geom_col(width = 0.58, alpha = 0.92) +
  geom_text(aes(label = percent(Rate, accuracy = 0.1)),
            vjust = -0.35, size = 3.4) +
  scale_fill_manual(values = c("Z = 1 (same sex)" = col_z1,
                               "Z = 0 (different sex)" = col_z0)) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, .65)) +
  labs(title = "Labor outcome: worked last year, by instrument status",
       subtitle = "The small difference is the raw reduced form for the binary outcome.",
       x = NULL, y = "Pr(worked = 1)", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 10))
```

```{r income-outcome-plot}
ae98_income_sub %>%
  mutate(`Z group` = factor(ifelse(samesex == 1, "Z = 1 (same sex)", "Z = 0 (different sex)"),
                            levels = c("Z = 1 (same sex)", "Z = 0 (different sex)"))) %>%
  ggplot(aes(x = lincomem, fill = `Z group`)) +
  geom_density(alpha = 0.45, colour = "white", adjust = 1.05) +
  geom_vline(xintercept = mean(ae98_income_sub$lincomem),
             linetype = "dashed", colour = col_grey) +
  scale_fill_manual(values = c("Z = 1 (same sex)" = col_z1,
                               "Z = 0 (different sex)" = col_z0)) +
  labs(title = "Income outcome: log income, by instrument status",
       subtitle = "Dashed line = mean log income in the selected income subsample.",
       x = "log(income)", y = "Density", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 10))
```

# 2. Overlap and Balance Checks

```{r propensity-overlap}
RHS <- "agem + agefstm + sexk + sex2nd + blackm + hispm + othracem"

fit_pscore <- function(df) {
  fit <- glm(as.formula(paste("samesex ~", RHS)), data = df, family = binomial())
  as.numeric(fitted(fit))
}

ps_summary <- function(df, label) {
  p <- fit_pscore(df)
  data.frame(
    Sample = label,
    Min = min(p),
    P05 = quantile(p, .05, names = FALSE),
    Mean = mean(p),
    P95 = quantile(p, .95, names = FALSE),
    Max = max(p),
    SD = sd(p),
    `Pr(p < .05 or p > .95)` = mean(p < .05 | p > .95),
    check.names = FALSE
  )
}

ps_tab <- bind_rows(
  ps_summary(ae98_labor_sub, "Labor sub"),
  ps_summary(ae98_income_sub, "Income sub")
)

ps_tab %>%
  mutate(across(-Sample, ~ round(.x, 4))) %>%
  kbl(caption = "Propensity-score overlap for Z = samesex",
      align = c("l", rep("r", 7))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) %>%
  footnote(general = "The instrument is close to random conditional on X, so p-hat should stay near 0.5. No mass near 0 or 1 means no obvious overlap problem for kappa weights.")
```

```{r propensity-overlap-plot}
ps_plot_df <- bind_rows(
  ae98_labor_sub %>%
    mutate(Sample = "Labor sub", p_hat = fit_pscore(ae98_labor_sub)),
  ae98_income_sub %>%
    mutate(Sample = "Income sub", p_hat = fit_pscore(ae98_income_sub))
) %>%
  mutate(`Z group` = factor(ifelse(samesex == 1, "Z = 1 (same sex)", "Z = 0 (different sex)"),
                            levels = c("Z = 1 (same sex)", "Z = 0 (different sex)")))

ggplot(ps_plot_df, aes(x = p_hat, fill = `Z group`)) +
  geom_histogram(bins = 30, alpha = 0.68, colour = "white", position = "identity") +
  geom_vline(xintercept = .5, linetype = "dashed", colour = col_grey) +
  facet_wrap(~ Sample, scales = "free_y") +
  scale_fill_manual(values = c("Z = 1 (same sex)" = col_z1,
                               "Z = 0 (different sex)" = col_z0)) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(title = "Propensity-score concentration and overlap",
       subtitle = "Both samples stay tightly centered around 0.5.",
       x = "p-hat(X) = Pr(Z = 1 | X)", y = "Count", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 10))
```

```{r covariate-balance}
covariates <- c("agem", "agefstm", "sexk", "sex2nd", "blackm", "hispm", "othracem")

smd_one <- function(x, z) {
  x1 <- x[z == 1]
  x0 <- x[z == 0]
  sp <- sqrt((var(x1) + var(x0)) / 2)
  if (!is.finite(sp) || sp == 0) return(NA_real_)
  (mean(x1) - mean(x0)) / sp
}

balance_table <- function(df, label) {
  data.frame(
    Sample = label,
    Variable = covariates,
    SMD = vapply(covariates, function(v) smd_one(df[[v]], df$samesex), numeric(1)),
    check.names = FALSE
  )
}

balance_tab <- bind_rows(
  balance_table(ae98_labor_sub, "Labor sub"),
  balance_table(ae98_income_sub, "Income sub")
) %>%
  mutate(`|SMD|` = abs(SMD)) %>%
  arrange(Sample, desc(`|SMD|`))

balance_tab %>%
  mutate(across(c(SMD, `|SMD|`), ~ round(.x, 3))) %>%
  kbl(caption = "Unadjusted covariate balance across the instrument",
      align = c("l", "l", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13) %>%
  pack_rows("Labor subsample", 1, length(covariates)) %>%
  pack_rows("Income subsample", length(covariates) + 1, nrow(balance_tab)) %>%
  footnote(general = "Small standardized mean differences are expected because sibling sex composition is close to random.")
```

```{r covariate-balance-plot}
ggplot(balance_tab, aes(x = reorder(Variable, `|SMD|`), y = `|SMD|`,
                        colour = Sample, shape = Sample)) +
  geom_hline(yintercept = 0.1, linetype = "dashed", colour = "#999999") +
  geom_point(size = 2.8, alpha = 0.92) +
  coord_flip() +
  scale_colour_manual(values = c("Labor sub" = col_z1, "Income sub" = col_z0)) +
  labs(title = "Unadjusted covariate balance across the instrument",
       subtitle = "Dashed line marks the common |SMD| = 0.10 reference threshold.",
       x = NULL, y = "|Standardized mean difference|",
       colour = NULL, shape = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(colour = "grey40", size = 10))
```

# 3. Covariate Structure Used Later

<div class="warn-box">
`samesex` is defined by whether `sexk` and `sex2nd` match. Pooled, the two child-sex indicators are nearly uncorrelated; within each instrument arm, the relationship is exact. This is the reason the later linear/logit DoubleML nuisance fits need the reduced design that drops one of them.
</div>

```{r sex-collinearity}
sex_corr <- function(df, label) {
  data.frame(
    Sample = label,
    Panel = c("Pooled", "Z = 0", "Z = 1"),
    `cor(sexk, sex2nd)` = c(
      cor(df$sexk, df$sex2nd),
      cor(df$sexk[df$samesex == 0], df$sex2nd[df$samesex == 0]),
      cor(df$sexk[df$samesex == 1], df$sex2nd[df$samesex == 1])
    ),
    check.names = FALSE
  )
}

sex_corr_tab <- bind_rows(
  sex_corr(ae98_labor_sub, "Labor sub"),
  sex_corr(ae98_income_sub, "Income sub")
)

sex_corr_tab %>%
  mutate(`cor(sexk, sex2nd)` = round(`cor(sexk, sex2nd)`, 3)) %>%
  kbl(caption = "Structural relation between first- and second-child sex indicators",
      align = c("l", "l", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13) %>%
  footnote(general = "Since Z = 1{sexk = sex2nd}, sexk and sex2nd are perfectly related within each Z arm. This is why later parametric DoubleML nuisance fits should not include both variables inside arm-specific linear/logit models.")
```

```{r sex-collinearity-plot, fig.height=4.8}
sex_corr_plot <- sex_corr_tab %>%
  mutate(Panel = paste(Sample, Panel, sep = " / "),
         Pair = "cor(sexk, sex2nd)")

ggplot(sex_corr_plot, aes(x = Pair, y = Panel, fill = `cor(sexk, sex2nd)`)) +
  geom_tile(colour = "white", linewidth = 1.2) +
  geom_text(aes(label = sprintf("%.2f", `cor(sexk, sex2nd)`)),
            size = 3.8, colour = "black") +
  scale_fill_gradient2(low = col_z0, mid = "white", high = col_z1,
                       limits = c(-1, 1), name = "Correlation") +
  labs(title = "sexk and sex2nd: nearly unrelated pooled, deterministic within Z",
       x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(face = "bold"))
```
