---
title: "Angrist & Evans (1998) — Translation Invariance: Rerun Check (Method A)"
subtitle: "Kappa estimators, DML smoother (PLR-IV / Wald-AIPW), and the learner comparison"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
h3   { color: #444; margin-top: 20px; }
table { font-size: 13px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 12, fig.height = 6, autodep = TRUE
)
```

---

# Purpose

This notebook checks translation invariance by the rerun method (Method A): the whole
estimation pipeline is fit once on the original outcome and once on the same outcome
shifted by a constant `k`, using the same seed, and the two point estimates are compared
directly. The tables report both this rerun check (Method A: `diff_A`, `invariant_A`)
and the frozen-weight algebraic prediction (Method B: `diff_B`, `invariant_B`) from
the original fit, plus `gap_AB` -- whether the two agree.

<div class="note-box">
**Two workflows only, not every SUW recoding.** Translation invariance is the claim that
`τ̂(Y+c) = τ̂(Y)` for **any** constant `c` — demonstrating it for one nonzero `c` per
outcome family is sufficient; testing every possible constant would be redundant (they
only differ in the value of `k`). This file therefore uses exactly two pure additive
shifts:

- **Labor**: `workedm` vs. `workedm + 1`, `k = 1`.
- **Income**: `lincomem` vs. `lincomem + log(100)` (i.e. the cents recoding), `k = log(100)`.

**`workedm_01 = 1 - workedm` is deliberately EXCLUDED from this file.** It is a
reflection plus a shift (`Y' = -Y+1`), not a pure translation, so it does not isolate
the same question — it belongs only in the SUW reproduction table in `child_test.Rmd`.
The mathematics of translation invariance (Σωᵢ=0 ⟺ invariant to any additive constant)
makes no distinction between a binary and a continuous outcome — `workedm+1` and
`lincomem+log(100)` exercise exactly the same mechanism, just with different `k` and a
different real-world motivation (income's shift mirrors an actual unit-of-measurement
choice; labor's is a synthetic algebraic stress test of the same property).
</div>

# Setup

```{r packages-and-source}
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(cobalt)
library(viridis)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

source("functions_all.R")
cat("functions_all.R sourced.\n")

ti_child_row <- function(name, w_original, tau_original, tau_shifted, k,
                         digits = 12) {
  row <- ti_unified_row(name = name, w_dollars = w_original,
                        tau_dollars = tau_original, tau_cents = tau_shifted,
                        k = k, digits = digits)
  names(row)[names(row) == "tau_dollars"] <- "tau_original"
  names(row)[names(row) == "tau_cents"] <- "tau_shifted"
  row$Shift_k <- round(k, digits)
  row[, c("Estimator", "Shift_k", "tau_original", "tau_shifted",
          "sum_w", "diff_B", "diff_A", "gap_AB",
          "invariant_B", "invariant_A")]
}
```

## Runtime switches

<div class="warn-box">
**XGBoost is enabled for the full child check.** Kappa, the DML headline, and the
linear/Ranger learner rows are comparatively cheap. The XGBoost Method-A rows rerun
nested tuning on the shifted outcome for Wald-AIPW and PLR-IV. The original XGBoost
estimates and weights are imported from the saved tuning exports, so the notebook does
not duplicate a completed original-outcome tuning run.
</div>

```{r switches}
RUN_KAPPA             <- TRUE
RUN_DML               <- TRUE
RUN_LM_RF             <- TRUE
RUN_XGB_METHOD_A      <- TRUE
RUN_XGB_LABOR_ONLY    <- FALSE
RUN_XGB_WALD_AIPW     <- TRUE
RUN_XGB_PLIV          <- TRUE
USE_XGB_ORIGINAL_EXPORT <- TRUE

cat(sprintf(
  "RUN_KAPPA=%s RUN_DML=%s RUN_LM_RF=%s RUN_XGB_METHOD_A=%s RUN_XGB_LABOR_ONLY=%s RUN_XGB_WALD_AIPW=%s RUN_XGB_PLIV=%s\n",
  RUN_KAPPA, RUN_DML, RUN_LM_RF, RUN_XGB_METHOD_A, RUN_XGB_LABOR_ONLY,
  RUN_XGB_WALD_AIPW, RUN_XGB_PLIV
))
```

## Data — the EXACT same subsamples as child_test.Rmd, no resampling

```{r data}
obj <- readRDS("ae98_sub.rds")
if (!all(c("ae98_labor_sub", "ae98_income_sub") %in% names(obj))) {
  stop("ae98_sub.rds must contain ae98_labor_sub and ae98_income_sub.")
}
ae98_labor_sub  <- obj$ae98_labor_sub
ae98_income_sub <- obj$ae98_income_sub

ae98_labor_sub <- ae98_labor_sub %>%
  mutate(workedm = as.integer(workedm), morekids = as.integer(morekids),
         samesex = as.integer(samesex))
ae98_income_sub <- ae98_income_sub %>%
  mutate(morekids = as.integer(morekids), samesex = as.integer(samesex),
         lincomem = log(incomem))

Z_labor <- ae98_labor_sub$samesex;  D_labor  <- ae98_labor_sub$morekids
Z_income <- ae98_income_sub$samesex; D_income <- ae98_income_sub$morekids

# The two shifts. Labor: pure +1. Income: +log(100), verified below to equal the
# cents recoding used elsewhere, so this workflow is consistent with child_test.Rmd.
k_labor  <- 1
k_income <- log(100)

Y_labor_orig   <- ae98_labor_sub$workedm
Y_labor_shift  <- ae98_labor_sub$workedm + k_labor

Y_income_orig  <- ae98_income_sub$lincomem
Y_income_shift <- ae98_income_sub$lincomem + k_income

# Sanity check: Y_income_shift should equal log(100*incomem) exactly (the cents coding).
max_dev <- max(abs(Y_income_shift - log(100 * ae98_income_sub$incomem)))
cat(sprintf("k_income = log(100) = %.12f, max deviation from cents coding = %.3e\n",
            k_income, max_dev))
stopifnot(max_dev < 1e-9)

cat(sprintf("Labor:  N=%d | Income: N=%d (same subsamples as child_test.Rmd)\n",
            nrow(ae98_labor_sub), nrow(ae98_income_sub)))
```

## Design matrices — identical construction to child_test.Rmd

```{r design-matrices}
RHS_FORMULA <- ~ agem + agefstm + sexk + sex2nd + blackm + hispm + othracem

X_labor_kappa  <- model.matrix(RHS_FORMULA, data = ae98_labor_sub)
X_income_kappa <- model.matrix(RHS_FORMULA, data = ae98_income_sub)

X_dml_labor  <- make_X(~ 0 + agem + agefstm + sexk + sex2nd + blackm + hispm + othracem,
                       data = ae98_labor_sub)
X_dml_income <- make_X(~ 0 + agem + agefstm + sexk + sex2nd + blackm + hispm + othracem,
                       data = ae98_income_sub)

# 6-covariate (sex2nd dropped) -- ONLY for linear+logit, same reasoning as child_test.Rmd
# Section 3: sexk/sex2nd are exactly collinear within a fixed Z arm, which breaks the
# arm-specific parametric (lm/logit) design-matrix inversion. Tree-based learners and
# the grf honest forest are unaffected and keep all 7 covariates.
X_dml_labor_lm  <- make_X(~ 0 + agem + agefstm + sexk + blackm + hispm + othracem,
                          data = ae98_labor_sub)
X_dml_income_lm <- make_X(~ 0 + agem + agefstm + sexk + blackm + hispm + othracem,
                          data = ae98_income_sub)

workflows <- list(
  Labor = list(label = "Labor (workedm)", D = D_labor, Z = Z_labor,
               X_kappa = X_labor_kappa, X_dml = X_dml_labor, X_dml_lm = X_dml_labor_lm,
               Y_orig = Y_labor_orig, Y_shift = Y_labor_shift, k = k_labor),
  Income = list(label = "Income (lincomem)", D = D_income, Z = Z_income,
               X_kappa = X_income_kappa, X_dml = X_dml_income, X_dml_lm = X_dml_income_lm,
               Y_orig = Y_income_orig, Y_shift = Y_income_shift, k = k_income)
)
cat("Two workflows built: Labor (k=1), Income (k=log(100)).\n")
```

---

# 1. Kappa estimators

<div class="note-box">
Kappa outcome weights are built only from the instrument, treatment, and estimated
propensity score — they never touch Y. Refitting on the shifted outcome therefore uses
IDENTICAL weights; `gap_AB` should be machine-zero for every row here. This section is
cheap regardless of `RUN_KAPPA` -- it's always safe to leave on.
</div>

```{r kappa-name-bridge}
# kappa_analytic_se_all() labels the tau_a,1 estimator "tau_ml_t"; kappa_weights_bundle()
# labels the same estimator's weight vector "tau_ml_a1". Bridge explicitly (same
# inconsistency documented in functions_all.R and handled explicitly here.
kappa_key_map <- c(
  tau_cb_u = "tau_cb_u", tau_ml_u = "tau_ml_u", tau_ml_a10 = "tau_ml_a10",
  tau_ml_a = "tau_ml_a", tau_ml_t = "tau_ml_a1", tau_ml_a0 = "tau_ml_a0"
)

build_kappa_ti_rows <- function(wf) {
  est_o  <- kappa_analytic_se_all(wf$Y_orig,  wf$Z, wf$D, wf$X_kappa)$estimates
  est_s  <- kappa_analytic_se_all(wf$Y_shift, wf$Z, wf$D, wf$X_kappa)$estimates
  w_list <- kappa_weights_bundle(wf$Z, wf$D, wf$X_kappa)

  rows <- lapply(names(kappa_key_map), function(est_name) {
    w_name <- kappa_key_map[[est_name]]
    row <- ti_child_row(name = est_name, w_original = w_list[[w_name]],
                        tau_original = est_o[[est_name]],
                        tau_shifted = est_s[[est_name]], k = wf$k)
    row$Sample <- wf$label
    row
  })
  do.call(rbind, rows)
}
```

```{r kappa-ti-run, cache=TRUE}
if (RUN_KAPPA) {
  kappa_ti_all <- do.call(rbind, lapply(workflows, build_kappa_ti_rows))
} else {
  cat("RUN_KAPPA is FALSE -- skipped.\n")
}
```

```{r kappa-ti-table}
if (RUN_KAPPA) {
  kappa_ti_all[, c("Sample", setdiff(names(kappa_ti_all), c("Sample", "gap_AB")))] %>%
    kbl(caption = "Kappa estimators -- original vs. shifted outcome, both samples",
        digits = 8, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(which(!kappa_ti_all$invariant_A), background = "#fffbf0") %>%
    footnote(general = "Orange = Method-A rerun did not reproduce invariance (expected for the unnormalized kappa rows). gap_AB should be ~0 everywhere: kappa weights never depend on Y.")
}
```

---

# 2. DML headline estimators (PLR-IV, Wald-AIPW)

<div class="note-box">
Tuned honest forest only (`tune.parameters="all"`); `child_test.Rmd` Section 2 already
covers the tuned-vs-untuned comparison as a separate question. Both the original and shifted fits use the same seed, same
`n_cf_folds`, and same tuning setting, so any difference reflects the outcome shift
itself rather than unrelated estimation randomness.
</div>

```{r dml-headline-builder}
build_dml_headline_ti_rows <- function(wf, n_cf_folds = 5) {
  set.seed(42)
  fit_o <- dml_with_smoother(wf$Y_orig, wf$D, wf$X_dml, wf$Z,
                             n_cf_folds = n_cf_folds, tune.parameters = "all")
  set.seed(42)
  fit_s <- dml_with_smoother(wf$Y_shift, wf$D, wf$X_dml, wf$Z,
                             n_cf_folds = n_cf_folds, tune.parameters = "all")

  res_o <- summary(fit_o); res_s <- summary(fit_s)
  om_o  <- get_outcome_weights(fit_o)
  check_omega_rows(om_o)

  rows <- lapply(c("PLR-IV", "Wald-AIPW"), function(est) {
    w   <- as.vector(om_o$omega[est, ])
    row <- ti_child_row(name = est, w_original = w,
                        tau_original = get_estimate(res_o, est),
                        tau_shifted = get_estimate(res_s, est), k = wf$k)
    row$Sample <- wf$label
    row
  })
  do.call(rbind, rows)
}
```

```{r dml-headline-run, cache=TRUE}
if (RUN_DML) {
  dml_headline_ti_all <- do.call(rbind, lapply(workflows, build_dml_headline_ti_rows))
} else {
  cat("RUN_DML is FALSE -- skipped.\n")
}
```

```{r dml-headline-table}
if (RUN_DML) {
  dml_headline_ti_all[, c("Sample", setdiff(names(dml_headline_ti_all), c("Sample", "gap_AB")))] %>%
    kbl(caption = "DML smoother (tuned honest forest) -- original vs. shifted, both samples",
        digits = 8, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(which(!dml_headline_ti_all$invariant_A), background = "#fffbf0")
}
```

---

# 3. Learner comparison

## 3a. Linear and Ranger learners (Wald-AIPW/IIVM and PLR-IV)

<div class="note-box">
This section mirrors the learner comparison in `child_test.Rmd`, but reruns
each estimator on the original and shifted outcomes. For Wald-AIPW/LATE we
use `DoubleMLIIVM`; for PLR-IV we use `DoubleMLPLIV`. Linear Wald-AIPW uses
linear outcome regression plus logit nuisance models, while linear PLR-IV uses
linear regression nuisance models throughout, as required by `DoubleMLPLIV`.
Ranger keeps the full 7-covariate design for both estimators.

XGBoost is handled separately below because the Method-A check requires nested
tuning on the shifted outcome.
</div>

```{r dml-learner-setup}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z", x_cols = colnames(X))
}

ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))

ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))

omega_vec <- function(om, preferred_row = NULL) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(preferred_row) && !is.null(rownames(w)) && preferred_row %in% rownames(w)) {
    return(as.vector(w[preferred_row, ]))
  }
  as.vector(w[1, ])
}

fit_iivm <- function(dml_data, learners, Y) {
  set.seed(42)
  o <- DoubleMLIIVM$new(data = dml_data, ml_g = learners$g, ml_m = learners$m,
                        ml_r = learners$r, n_folds = 5, score = "LATE")
  o$fit(store_models = TRUE, store_predictions = TRUE)
  om <- get_outcome_weights(object = o, dml_data = dml_data)
  w <- omega_vec(om, "Wald-AIPW")
  list(obj = o, w = w, ok = check_weight_identity(w, Y, get_dml_coef(o)))
}

fit_pliv <- function(dml_data, learners, Y) {
  set.seed(42)
  o <- DoubleMLPLIV$new(data = dml_data, ml_l = learners$l, ml_m = learners$m,
                        ml_r = learners$r, n_folds = 5)
  o$fit(store_models = TRUE, store_predictions = TRUE)
  om <- get_outcome_weights(object = o, dml_data = dml_data)
  w <- omega_vec(om, "PLR-IV")
  list(obj = o, w = w, ok = check_weight_identity(w, Y, get_dml_coef(o)))
}

build_dml_learner_ti_row <- function(wf, label, learners_fn, X_dml, fit_fn) {
  dat_o <- make_dml_data(wf$Y_orig,  wf$D, wf$Z, X_dml)
  dat_s <- make_dml_data(wf$Y_shift, wf$D, wf$Z, X_dml)

  fit_o <- fit_fn(dat_o, learners_fn(), wf$Y_orig)
  fit_s <- fit_fn(dat_s, learners_fn(), wf$Y_shift)

  row <- ti_child_row(name = label, w_original = fit_o$w,
                      tau_original = get_dml_coef(fit_o$obj),
                      tau_shifted = get_dml_coef(fit_s$obj), k = wf$k)
  row$Sample <- wf$label
  row$identity_ok <- fit_o$ok
  row
}
```

```{r dml-learner-run, cache=TRUE}
if (RUN_LM_RF) {
  dml_learner_ti_all <- do.call(rbind, lapply(workflows, function(wf) {
    rbind(
      build_dml_learner_ti_row(wf, "PLR-IV (linear)",          ml_lm_pliv,  wf$X_dml_lm, fit_pliv),
      build_dml_learner_ti_row(wf, "Wald-AIPW (linear+logit)", ml_lm_iivm,  wf$X_dml_lm, fit_iivm),
      build_dml_learner_ti_row(wf, "PLR-IV (ranger)",          ml_rf_pliv,  wf$X_dml,    fit_pliv),
      build_dml_learner_ti_row(wf, "Wald-AIPW (ranger)",       ml_rf_iivm,  wf$X_dml,    fit_iivm)
    )
  }))
} else {
  cat("RUN_LM_RF is FALSE -- skipped.\n")
}
```

```{r dml-learner-table}
if (RUN_LM_RF) {
  drop_cols <- c("gap_AB", "identity_ok")
  dml_learner_ti_all[, c("Sample", setdiff(names(dml_learner_ti_all), c("Sample", drop_cols)))] %>%
    kbl(caption = "DML learners -- PLR-IV and Wald-AIPW, original vs. shifted outcome",
        digits = 8, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!dml_learner_ti_all$invariant_A), background = "#fffbf0") %>%
    footnote(general = "PLR-IV rows use DoubleMLPLIV. Wald-AIPW/LATE rows use DoubleMLIIVM. Linear specifications use the 6-covariate design with sex2nd dropped; ranger uses the full 7-covariate design. Orange rows fail the Method-A rerun invariance check.")
}
```

## 3b. XGBoost Method A

<div class="warn-box">
**XGBoost rule used here.** The original-outcome tuned XGBoost estimates and weights
are imported from the saved tuning exports. For the shifted outcome, the whole nested
tuning procedure is rerun: same seed, same subsample, same cross-fitting setup, same
`tune_on_folds = TRUE`. This is the relevant Method-A check. Reusing a single frozen
hyperparameter recipe would define a different exercise because the original tuning
selected fold-specific hyperparameters.
</div>

```{r xgb-setup}
SEED <- 42; N_CF <- 5; TUNE_EVALS <- 15; INNER_FOLDS <- 3

if (!USE_XGB_ORIGINAL_EXPORT) {
  stop("This notebook expects USE_XGB_ORIGINAL_EXPORT <- TRUE for the original XGBoost side.")
}

xgb_wald_export <- readRDS("ae98_xgb_tuning_export.rds")
xgb_pliv_export <- readRDS("ae98_xgb_pliv_tuning_export.rds")

make_xgb_learners_iivm <- function() {
  list(g = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
       m = lrn("classif.xgboost", predict_type = "prob"),
       r = lrn("classif.xgboost", predict_type = "prob"))
}
make_xgb_learners_pliv <- function() {
  list(l = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
       m = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
       r = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0))
}
xgb_ps <- function() {
  ps(nrounds = p_int(50, 400), max_depth = p_int(2, 5),
     eta = p_dbl(0.02, 0.20), min_child_weight = p_int(1, 8))
}
tune_param_set_iivm <- function() list(ml_g = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())
tune_param_set_pliv <- function() list(ml_l = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())
tune_settings_iivm <- function() {
  list(n_folds_tune = INNER_FOLDS, rsmp_tune = rsmp("cv", folds = INNER_FOLDS),
       measure = list(ml_g = "regr.rmse", ml_m = "classif.logloss", ml_r = "classif.logloss"),
       terminator = trm("evals", n_evals = TUNE_EVALS), algorithm = tnr("random_search"))
}
tune_settings_pliv <- function() {
  list(n_folds_tune = INNER_FOLDS, rsmp_tune = rsmp("cv", folds = INNER_FOLDS),
       measure = list(ml_l = "regr.rmse", ml_m = "regr.rmse", ml_r = "regr.rmse"),
       terminator = trm("evals", n_evals = TUNE_EVALS), algorithm = tnr("random_search"))
}

get_xgb_original <- function(export, cell, learner) {
  row <- export$weight_diagnostics |>
    filter(.data$Cell == cell, .data$Learner == learner)
  if (nrow(row) != 1) {
    stop(sprintf("Expected one export row for %s / %s, found %d.", cell, learner, nrow(row)))
  }
  list(tau = row$Estimate[[1]],
       w = as.vector(export$omega_vectors[[cell]]$tuned),
       identity_ok = isTRUE(export$identity_checks[[cell]]$tuned))
}

fit_xgb_wald_shifted <- function(Y, D, Z, X) {
  dml_data <- make_dml_data(Y, D, Z, X)
  learners <- make_xgb_learners_iivm()
  o <- DoubleMLIIVM$new(data = dml_data, ml_g = learners$g, ml_m = learners$m,
                        ml_r = learners$r, n_folds = N_CF, score = "LATE")
  o$tune(param_set = tune_param_set_iivm(), tune_settings = tune_settings_iivm(),
         tune_on_folds = TRUE)
  o$fit(store_models = TRUE, store_predictions = TRUE)
  list(obj = o, tau = get_dml_coef(o))
}

fit_xgb_pliv_shifted <- function(Y, D, Z, X) {
  dml_data <- make_dml_data(Y, D, Z, X)
  learners <- make_xgb_learners_pliv()
  o <- DoubleMLPLIV$new(data = dml_data, ml_l = learners$l, ml_m = learners$m,
                        ml_r = learners$r, n_folds = N_CF)
  o$tune(param_set = tune_param_set_pliv(), tune_settings = tune_settings_pliv(),
         tune_on_folds = TRUE)
  o$fit(store_models = TRUE, store_predictions = TRUE)
  list(obj = o, tau = get_dml_coef(o))
}

build_xgb_method_a_rows <- function(wf) {
  rows <- list()

  if (RUN_XGB_WALD_AIPW) {
    original <- get_xgb_original(xgb_wald_export, wf$label, "XGB tuned")
    set.seed(SEED)
    shifted <- fit_xgb_wald_shifted(wf$Y_shift, wf$D, wf$Z, wf$X_dml)
    row <- ti_child_row(name = "Wald-AIPW (XGBoost tuned)",
                        w_original = original$w,
                        tau_original = original$tau,
                        tau_shifted = shifted$tau,
                        k = wf$k)
    row$Sample <- wf$label
    row$identity_ok <- original$identity_ok
    rows[["wald"]] <- row
  }

  if (RUN_XGB_PLIV) {
    original <- get_xgb_original(xgb_pliv_export, wf$label, "PLR-IV XGB tuned")
    set.seed(SEED)
    shifted <- fit_xgb_pliv_shifted(wf$Y_shift, wf$D, wf$Z, wf$X_dml)
    row <- ti_child_row(name = "PLR-IV (XGBoost tuned)",
                        w_original = original$w,
                        tau_original = original$tau,
                        tau_shifted = shifted$tau,
                        k = wf$k)
    row$Sample <- wf$label
    row$identity_ok <- original$identity_ok
    rows[["pliv"]] <- row
  }

  do.call(rbind, rows)
}

save_xgb_method_a_export <- function(xgb_ti_all) {
  saveRDS(
    list(results = xgb_ti_all,
         settings = list(seed = SEED, n_cf = N_CF, tune_evals = TUNE_EVALS,
                         inner_folds = INNER_FOLDS, tune_on_folds = TRUE,
                         original_exports = c("ae98_xgb_tuning_export.rds",
                                              "ae98_xgb_pliv_tuning_export.rds"))),
    "ae98_translation_xgb_method_a_export.rds"
  )
}
```

```{r xgb-run, cache=TRUE, cache.extra=list(RUN_XGB_METHOD_A, RUN_XGB_LABOR_ONLY, RUN_XGB_WALD_AIPW, RUN_XGB_PLIV, TUNE_EVALS, INNER_FOLDS)}
if (RUN_XGB_METHOD_A) {
  xgb_workflows <- if (RUN_XGB_LABOR_ONLY) workflows["Labor"] else workflows
  xgb_ti_all <- do.call(rbind, lapply(xgb_workflows, build_xgb_method_a_rows))
  save_xgb_method_a_export(xgb_ti_all)
} else {
  cat("XGBoost Method A skipped (RUN_XGB_METHOD_A = FALSE).\n")
}
```

```{r xgb-table}
if (RUN_XGB_METHOD_A) {
  drop_cols <- c("gap_AB", "identity_ok")
  xgb_ti_all[, c("Sample", setdiff(names(xgb_ti_all), c("Sample", drop_cols)))] %>%
    kbl(caption = "XGBoost Method A -- original export vs. shifted rerun",
        digits = 8, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!xgb_ti_all$invariant_A), background = "#fffbf0") %>%
    footnote(general = "Original tuned estimates and original outcome weights are imported from the saved XGBoost tuning exports. Shifted estimates are generated by rerunning nested tuning with the same seed, folds, tuning budget, and tune_on_folds=TRUE. Orange rows fail the Method-A rerun invariance check.")
} else {
  cat("XGBoost Method A skipped (RUN_XGB_METHOD_A = FALSE).\n")
}
```

```{r xgb-identity-table}
if (RUN_XGB_METHOD_A) {
  xgb_ti_all %>%
    select(Sample, Estimator, identity_ok, sum_w, diff_B, diff_A, gap_AB) %>%
    kbl(caption = "XGBoost weight identity and Method-A/Method-B gap",
        digits = 8, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    footnote(general = "identity_ok checks whether the imported original weights reproduce the original XGBoost point estimate exactly. For XGBoost this may fail because the estimator is not an affine smoother in the same clean sense as the kappa or grf rows; interpret the weight-based Method-B column with that caveat.")
}
```

<div class="note-box">
The XGBoost block writes `ae98_translation_xgb_method_a_export.rds` after a successful
run. That file stores only the Method-A table and settings; it does not overwrite the
original XGBoost tuning exports.
</div>

<div class="note-box">
**Reading this file.** For every row: `invariant_B` (Method B, algebraic) should match
`invariant_A` (Method A, rerun) closely for kappa and the grf honest forest (affine
smoothers). Small `gap_AB` for adaptive learners (Ranger, XGBoost) is expected and
reflects genuine re-fitting stochasticity, not a bug -- large `gap_AB` would be worth
flagging. Unnormalized kappa rows (`tau_ml_a`, `tau_ml_t`, `tau_ml_a0`) are EXPECTED to
fail `invariant_A`/`invariant_B` -- that is the whole point of including them.
</div>
