---
title: "Angrist & Evans (1998) — Main Analysis"
subtitle: "Reproduction tables, DML estimates, learner comparisons, and weight diagnostics"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body   { font-size: 15px; color: #2c2c2c; }
h1     { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2     { color: #2c7bb6; margin-top: 28px; }
h3     { color: #444; margin-top: 20px; }

.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 12,
  fig.height = 6
)
```

---

# Setup

## Packages & Functions

```{r packages-and-source}
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(cobalt)           # love.plot(), used inside make_love()
library(viridis)          # colour scale, used inside make_love()
library(OutcomeWeights)   # dml_with_smoother(), get_outcome_weights()
library(DoubleML)         # DoubleMLIIVM, DoubleMLData
library(mlr3verse)        # lrn() learners for linear/logit and ranger

lgr::get_logger("mlr3")$set_threshold("warn")

source("functions_all.R")   # run_2sls, kappa_analytic_se_all, weight_diag, etc.
cat("functions_all.R sourced.\n")
```

## Data & Design

<div class="note-box">
**Setting.** Angrist & Evans (1998) childbearing application. Instrument Z =
`samesex` (sex of first two children matches). Treatment D = `morekids` (more
than two children). Two outcome-specific analysis samples, each a 3,000-row
stratified diagnostic subsample (built in
`ae98_make_balanced_subsample_submission.Rmd`): **labor-force sample**
(`ae98_labor_sub`, outcome family = `workedm`) and **log-income sample**
(`ae98_income_sub`, outcome family = `lincomem`, restricted upstream to
positive reported income). Covariates: `agem, agefstm, sexk, sex2nd, blackm,
hispm, othracem` (7 covariates, one spec — AE98 has no Card/Kitagawa-style
second spec). The notebook first recreates the SUW Table 4 structure on the
diagnostic subsamples and then compares DML, learner-based, and kappa
weighting estimators.
</div>

```{r data}
obj <- readRDS("ae98_sub.rds")
if (!all(c("ae98_labor_sub", "ae98_income_sub") %in% names(obj))) {
  stop("ae98_sub.rds must contain ae98_labor_sub and ae98_income_sub. Run ae98_make_balanced_subsample_submission.Rmd first.")
}

ae98_labor_sub  <- obj$ae98_labor_sub
ae98_income_sub <- obj$ae98_income_sub

# --- Labor-force sample: 3 outcome recodings (SUW Table 4, cols 1-3) ---
ae98_labor_sub <- ae98_labor_sub %>%
  mutate(
    workedm    = as.integer(workedm),
    morekids   = as.integer(morekids),
    samesex    = as.integer(samesex),
    workedm_21 = workedm + 1L,        # 2 if worked, 1 otherwise
    workedm_01 = 1L - workedm         # 1 if did not work, 0 otherwise
  )

# --- Log-income sample: 4 unit recodings (SUW Table 4, cols 4-7) ---
ae98_income_sub <- ae98_income_sub %>%
  mutate(
    morekids      = as.integer(morekids),
    samesex       = as.integer(samesex),
    lincomem      = log(incomem),             # dollars   -> col (5)
    lincomem_cnt  = log(100   * incomem),     # cents     -> col (4)
    lincomem_ths  = log(incomem / 1000),       # $1000s    -> col (6)
    lincomem_hths = log(incomem / 100000)      # $100,000s -> col (7)
  )

cat(sprintf("Labor subsample:  N = %d | mean(workedm) = %.4f | mean(samesex) = %.4f | mean(morekids) = %.4f\n",
            nrow(ae98_labor_sub), mean(ae98_labor_sub$workedm),
            mean(ae98_labor_sub$samesex), mean(ae98_labor_sub$morekids)))
cat(sprintf("Income subsample: N = %d | mean(lincomem) = %.4f | mean(samesex) = %.4f | mean(morekids) = %.4f\n",
            nrow(ae98_income_sub), mean(ae98_income_sub$lincomem),
            mean(ae98_income_sub$samesex), mean(ae98_income_sub$morekids)))
```

```{r design-matrices}
RHS_FORMULA <- ~ agem + agefstm + sexk + sex2nd + blackm + hispm + othracem

# Kappa design matrices — WITH intercept (for logit / CBPS)
X_labor_kappa <- model.matrix(RHS_FORMULA, data = ae98_labor_sub)
X_labor_df    <- as.data.frame(X_labor_kappa[, -1, drop = FALSE])

X_income_kappa <- model.matrix(RHS_FORMULA, data = ae98_income_sub)
X_income_df    <- as.data.frame(X_income_kappa[, -1, drop = FALSE])

# Core vectors
Z_labor <- ae98_labor_sub$samesex
D_labor <- ae98_labor_sub$morekids

Z_income <- ae98_income_sub$samesex
D_income <- ae98_income_sub$morekids

cat(sprintf("Kappa design: labor %d covariates (+intercept) | income %d covariates (+intercept)\n",
            ncol(X_labor_kappa) - 1, ncol(X_income_kappa) - 1))

# DML design matrices — no intercept, attributes stripped for grf (make_X)
X_dml_labor  <- make_X(~ 0 + agem + agefstm + sexk + sex2nd + blackm + hispm + othracem,
                       data = ae98_labor_sub)
X_dml_income <- make_X(~ 0 + agem + agefstm + sexk + sex2nd + blackm + hispm + othracem,
                       data = ae98_income_sub)

# 6-covariate design matrices (sex2nd dropped) -- ONLY for the arm-specific
# linear+logit learner in Section 3. sexk and sex2nd are exactly collinear
# within a fixed Z arm (Z=1{sexk==sex2nd}), which breaks a design-matrix
# inversion but does not affect kappa (pooled propensity), the grf honest
# forest, Ranger, or XGBoost (all tree-based / no matrix inversion). See
# ae98_analysis_plan.txt Section 3 for the full reasoning.
X_dml_labor_lm  <- make_X(~ 0 + agem + agefstm + sexk + blackm + hispm + othracem,
                          data = ae98_labor_sub)
X_dml_income_lm <- make_X(~ 0 + agem + agefstm + sexk + blackm + hispm + othracem,
                          data = ae98_income_sub)

cat(sprintf("DML design: labor %dx%d | income %dx%d (7-covariate, used by ranger/xgboost/grf)\n",
            nrow(X_dml_labor), ncol(X_dml_labor), nrow(X_dml_income), ncol(X_dml_income)))
cat(sprintf("DML design (linear+logit only): labor %dx%d | income %dx%d (6-covariate, sex2nd dropped)\n",
            nrow(X_dml_labor_lm), ncol(X_dml_labor_lm), nrow(X_dml_income_lm), ncol(X_dml_income_lm)))
```

---

# 1. Reproduction — SUW Table 4 (split by outcome family)

<div class="note-box">
**Layout.** SUW Table 4 has 7 columns on the full samples (N = 394,840 labor /
220,502 income): 3 labor-force recodings + 4 log-income unit recodings. We
cannot reproduce those observation counts — everything below runs on the two
3,000-row diagnostic subsamples only. Reproduced here is the **structure**:
normalized estimators (τ_cb_u, τ_ml_u, τ_ml_a10) should stay translation
invariant across recodings within a panel; unnormalized ones (τ_ml_a, τ_ml_t,
τ_ml_a0) should not.
</div>

```{r build-columns, cache=TRUE}
# One column = 2SLS + all six kappa estimators with analytical SEs.
run_child_column <- function(Y, D, Z, X_kappa, X_df) {
  list(
    tsls  = run_2sls(Y, D, Z, X_df, endog_name = "D"),
    kappa = kappa_analytic_se_all(Y, Z, D, X_kappa)
  )
}

# --- Labor-force columns (cols 1-3 analogue) ---
labor_cols <- list(
  "1 if worked, 0 else"    = run_child_column(ae98_labor_sub$workedm,
                                               D_labor, Z_labor, X_labor_kappa, X_labor_df),
  "2 if worked, 1 else"    = run_child_column(ae98_labor_sub$workedm_21,
                                               D_labor, Z_labor, X_labor_kappa, X_labor_df),
  "1 if did not work, 0 else" = run_child_column(ae98_labor_sub$workedm_01,
                                               D_labor, Z_labor, X_labor_kappa, X_labor_df)
)

# --- Log-income columns (cols 4-7 analogue, same order as Table 4) ---
income_cols <- list(
  "Cents"        = run_child_column(ae98_income_sub$lincomem_cnt,
                                     D_income, Z_income, X_income_kappa, X_income_df),
  "Dollars"      = run_child_column(ae98_income_sub$lincomem,
                                     D_income, Z_income, X_income_kappa, X_income_df),
  "$1000s"       = run_child_column(ae98_income_sub$lincomem_ths,
                                     D_income, Z_income, X_income_kappa, X_income_df),
  "$100,000s"    = run_child_column(ae98_income_sub$lincomem_hths,
                                     D_income, Z_income, X_income_kappa, X_income_df)
)

cat("Labor columns (3) and income columns (4) computed.\n")
```

```{r table-builder, echo=FALSE}
cell_fmt <- function(est, se, digits = 3) {
  if (is.na(est)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0)
    return(sprintf("%.3f<br><span style='color:#888'>(NA)</span>", est))
  pval  <- 2 * pnorm(-abs(est / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s<br><span style='color:#888'>(%.3f)</span>", est, stars, se)
}

# Generic table builder — works for any number of columns (3 for labor, 4 for income).
# flip_col: name of a column whose POINT ESTIMATE (not SE) should be displayed
# with the opposite sign, matching SUW's own reporting convention for a
# reflected (not translated) outcome recoding -- see ae98_sign_convention_note.txt.
build_child_table <- function(col_list, flip_col = NULL) {
  ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
  labels <- c("\u03c4_cb_u","\u03c4_ml_u","\u03c4_ml_a10","\u03c4_ml_a","\u03c4_ml_t","\u03c4_ml_a0")

  fmt_one <- function(cn, est, se) {
    if (!is.null(flip_col) && cn == flip_col) est <- -est   # display-only sign flip; SE untouched
    cell_fmt(est, se)
  }

  row_2sls <- c(Estimator = "2SLS",
                vapply(names(col_list), function(cn)
                  fmt_one(cn, col_list[[cn]]$tsls$coef, col_list[[cn]]$tsls$se),
                  character(1)))

  kappa_rows <- lapply(seq_along(ests), function(i) {
    e <- ests[i]
    c(Estimator = labels[i],
      vapply(names(col_list), function(cn)
        fmt_one(cn, col_list[[cn]]$kappa$estimates[e], col_list[[cn]]$kappa$se[e]),
        character(1)))
  })

  df <- as.data.frame(do.call(rbind, c(list(row_2sls), kappa_rows)),
                       stringsAsFactors = FALSE)
  names(df) <- c("Estimator", names(col_list))
  rownames(df) <- NULL
  df
}
```

## Results {.tabset}

### Binary Outcomes: Work Status (labor sample, N = `r nrow(ae98_labor_sub)`)

```{r table-labor}
build_child_table(labor_cols, flip_col = "1 if did not work, 0 else") %>%
  kbl(escape = FALSE,
      align = c("l", rep("c", length(labor_cols))),
      caption = sprintf("AE98 reproduction — Labor-force participation (diagnostic subsample, N = %d; SUW Table 4 uses N = 394,840)",
                         nrow(ae98_labor_sub))) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",            1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",   2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators", 5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant across recodings). Orange = unnormalized. Covariates: agem, agefstm, sexk, sex2nd, blackm, hispm, othracem. Stars: * p<0.10, ** p<0.05, *** p<0.01. Following Slocinski, Uysal, and Wooldridge (2025), the 'did not work' column reports the additive inverse of the raw estimate (SE unchanged) so that normalized rows read identically across all three recodings, matching SUW Table 4's own display convention -- see ae98_sign_convention_note.txt for the derivation. This flip is NOT applied anywhere else (income table, Section 2 DML, Section 3 learner comparison): it is a display-only convention for this one reproduction column.")
```

### Continuous Outcome: Log Family Income (income sample, N = `r nrow(ae98_income_sub)`)

```{r table-income}
build_child_table(income_cols) %>%
  kbl(escape = FALSE,
      align = c("l", rep("c", length(income_cols))),
      caption = sprintf("AE98 reproduction — Log family income (diagnostic subsample, N = %d; SUW Table 4 uses N = 220,502)",
                         nrow(ae98_income_sub))) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",            1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",   2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators", 5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant across unit recodings). Orange = unnormalized — expect these rows to change across Cents/Dollars/$1000s/$100,000s. Covariates: agem, agefstm, sexk, sex2nd, blackm, hispm, othracem. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 2. Double ML — PLR-IV & Wald-AIPW (honest forest, tuned vs. untuned)

<div class="note-box">
**Setting.** One `dml_with_smoother()` fit per sample, on that sample's
natural outcome (`workedm` for labor, `lincomem` for income) — this is not
a "log vs. binary" choice, it is the same two-sample split carried over from
Section 1. `tune.parameters` is grf's own native tuning switch (passed
through `...` to `grf::regression_forest`/`instrumental_forest`): `"none"` is
grf's default (no internal tuning), `"all"` triggers grf's automatic,
out-of-bag-loss-based tuning of the honest forest. We run **both** for each
sample — 4 fits total — to see whether tuning changes anything before
deciding what to carry forward. `set.seed(42)` is reset immediately before
every call, since R has one global RNG stream and grf's own internal
tuning search consumes draws the untuned run never does (small residual
divergence between tuned/untuned runs is therefore expected even from the
same seed — see `ae98_analysis_plan.txt` Section 4).
</div>

```{r dml-fit-all, cache=TRUE}
# Labor sample: outcome = workedm (binary)
set.seed(42); dml_labor_untuned  <- dml_with_smoother(
  ae98_labor_sub$workedm, D_labor, X_dml_labor, Z_labor,
  n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_labor_tuned    <- dml_with_smoother(
  ae98_labor_sub$workedm, D_labor, X_dml_labor, Z_labor,
  n_cf_folds = 5, tune.parameters = "all")

# Income sample: outcome = lincomem (continuous, log dollars)
set.seed(42); dml_income_untuned <- dml_with_smoother(
  ae98_income_sub$lincomem, D_income, X_dml_income, Z_income,
  n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_income_tuned   <- dml_with_smoother(
  ae98_income_sub$lincomem, D_income, X_dml_income, Z_income,
  n_cf_folds = 5, tune.parameters = "all")

res_labor_untuned  <- summary(dml_labor_untuned)
res_labor_tuned    <- summary(dml_labor_tuned)
res_income_untuned <- summary(dml_income_untuned)
res_income_tuned   <- summary(dml_income_tuned)

# Extract PLR-IV and Wald-AIPW outcome weights for every fit.
extract_w <- function(dml_obj) {
  om <- get_outcome_weights(dml_obj)
  check_omega_rows(om)
  list(plriv    = as.vector(om$omega["PLR-IV",    ]),
       waldaipw = as.vector(om$omega["Wald-AIPW", ]))
}
w_labor_untuned  <- extract_w(dml_labor_untuned)
w_labor_tuned    <- extract_w(dml_labor_tuned)
w_income_untuned <- extract_w(dml_income_untuned)
w_income_tuned   <- extract_w(dml_income_tuned)

cat("All 4 DML fits (2 samples x untuned/tuned) complete.\n")
```

```{r dml-cell-display, echo=FALSE}
# PLR-IV + Wald-AIPW rows for one fit, with the algebraic identity check inline.
show_dml_cell <- function(res, wlist, Y, cap) {
  ids <- c(
    check_weight_identity(wlist$plriv,    Y, get_estimate(res, "PLR-IV")),
    check_weight_identity(wlist$waldaipw, Y, get_estimate(res, "Wald-AIPW"))
  )
  id_col <- "Identity: sum(w*Y)=tau"
  df <- as.data.frame(res[c("PLR-IV", "Wald-AIPW"), , drop = FALSE])
  df$Estimator <- rownames(df)
  df[[id_col]] <- as.character(ids)
  df <- df[, c("Estimator", setdiff(names(df), c("Estimator", id_col)), id_col)]
  df %>%
    kbl(caption = cap, digits = 4, row.names = FALSE,
        align = c("l", rep("r", ncol(df) - 2), "c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "Identity column: TRUE means sum(omega * Y) reproduces the estimate to machine precision.")
}
```

## Estimates {.tabset}

### Labor sample (workedm)

```{r dml-labor-untuned}
show_dml_cell(res_labor_untuned, w_labor_untuned, ae98_labor_sub$workedm,
              "Labor sample — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-labor-tuned}
show_dml_cell(res_labor_tuned, w_labor_tuned, ae98_labor_sub$workedm,
              "Labor sample — tuned honest forest (tune.parameters = \"all\")")
```

### Income sample (lincomem)

```{r dml-income-untuned}
show_dml_cell(res_income_untuned, w_income_untuned, ae98_income_sub$lincomem,
              "Income sample — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-income-tuned}
show_dml_cell(res_income_tuned, w_income_tuned, ae98_income_sub$lincomem,
              "Income sample — tuned honest forest (tune.parameters = \"all\")")
```

<div class="note-box">
**What to look at.** Compare untuned vs. tuned within each sample: does the
point estimate move, does the SE shrink or grow, does the identity check
still hold (it must — that's an algebra check on the package's own weight
extraction, not a property that tuning could break)? A stable estimate
across tuned/untuned is a reassuring sign; a large jump is worth flagging
before moving to the learner comparison in Section 3.
</div>

---

# 3. Learner Comparison — Estimates (PLR-IV and Wald-AIPW)

<div class="note-box">
**Scope of this section: estimates only.** Weight diagnostics and Love plots
are reported in Sections 4 and 5. This section pairs the existing
Wald-AIPW/LATE learner estimates with PLR-IV estimates. Linear/logit and
Ranger are fit fresh on both samples. XGBoost rows are imported from separate
tuning exports: `ae98_xgb_tuning_export.rds` for Wald-AIPW and
`ae98_xgb_pliv_tuning_export.rds` for PLR-IV. This keeps the main analysis
fast and preserves the fold-specific tuned XGBoost fits exactly as estimated.

**Covariates.** Ranger and XGBoost use the full 7-covariate
`X_dml_labor`/`X_dml_income` design. Linear/logit Wald-AIPW and linear PLR-IV
use the 6-covariate `X_dml_labor_lm`/`X_dml_income_lm` (`sex2nd` dropped),
since the arm-specific parametric nuisance model otherwise faces exact
collinearity.
</div>

```{r ext-setup}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X))
}

dat_labor_lm   <- make_dml_data(ae98_labor_sub$workedm,   D_labor,  Z_labor,  X_dml_labor_lm)
dat_labor      <- make_dml_data(ae98_labor_sub$workedm,   D_labor,  Z_labor,  X_dml_labor)
dat_income_lm  <- make_dml_data(ae98_income_sub$lincomem, D_income, Z_income, X_dml_income_lm)
dat_income     <- make_dml_data(ae98_income_sub$lincomem, D_income, Z_income, X_dml_income)

cat("DoubleML data objects built: labor (7-cov, 6-cov) and income (7-cov, 6-cov).\n")
```

```{r ext-fit-helpers, echo=FALSE}
# Learner specs -- identical to card_presentation_4_1.Rmd, so this is a like-
# for-like comparison with the Card application's own learner choices.
ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))

ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))
omega_vec <- function(om) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && "Wald-AIPW" %in% rownames(w)) return(as.vector(w["Wald-AIPW", ]))
  as.vector(w[1, ])
}

# Fit one IIVM model and return list(obj, w, ok) -- same pattern as Card.
fit_iivm <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLIIVM$new(data = dml_data,
                          ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5, score = "LATE")
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  om  <- get_outcome_weights(object = obj, dml_data = dml_data)
  w   <- omega_vec(om)
  ok  <- check_doubleml_identity(om, Y, obj$coef)
  list(obj = obj, w = w, ok = ok)
}

fit_pliv <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLPLIV$new(data = dml_data,
                          ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5)
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  om  <- get_outcome_weights(object = obj, dml_data = dml_data)
  w   <- omega_vec(om)
  ok  <- check_weight_identity(w, Y, get_dml_coef(obj))
  list(obj = obj, w = w, ok = ok)
}

learner_pair_table <- function(pliv_labor, pliv_income, iivm_labor, iivm_income, cap) {
  id_col <- "Identity: sum(w*Y)=tau"
  df <- data.frame(
    Estimator = c("PLR-IV", "Wald-AIPW / LATE"),
    `Labor estimate` = round(c(get_dml_coef(pliv_labor$obj), get_dml_coef(iivm_labor$obj)), 4),
    `Labor SE`       = round(c(get_dml_se(pliv_labor$obj),   get_dml_se(iivm_labor$obj)), 4),
    `Labor identity` = c(pliv_labor$ok, iivm_labor$ok),
    `Income estimate` = round(c(get_dml_coef(pliv_income$obj), get_dml_coef(iivm_income$obj)), 4),
    `Income SE`       = round(c(get_dml_se(pliv_income$obj),   get_dml_se(iivm_income$obj)), 4),
    `Income identity` = c(pliv_income$ok, iivm_income$ok),
    check.names = FALSE
  )

  df %>%
    kbl(caption = cap, align = c("l","r","r","c","r","r","c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    footnote(general = "PLR-IV rows use DoubleMLPLIV. Wald-AIPW/LATE rows use DoubleMLIIVM. Identity columns check algebraic sum(omega*Y) = tau_hat.")
}
```

## Estimates {.tabset}

### Linear + Logistic

```{r ext-fit-lm, cache=TRUE}
lm_labor       <- fit_iivm(dat_labor_lm,  ml_lm_iivm(), ae98_labor_sub$workedm)
lm_income      <- fit_iivm(dat_income_lm, ml_lm_iivm(), ae98_income_sub$lincomem)
pliv_lm_labor  <- fit_pliv(dat_labor_lm,  ml_lm_pliv(), ae98_labor_sub$workedm)
pliv_lm_income <- fit_pliv(dat_income_lm, ml_lm_pliv(), ae98_income_sub$lincomem)

learner_pair_table(pliv_lm_labor, pliv_lm_income, lm_labor, lm_income,
                   "Learner estimates -- Linear/Logistic Wald-AIPW and Linear PLR-IV (6 covariates)")
```

### Ranger

```{r ext-fit-rf, cache=TRUE}
rf_labor       <- fit_iivm(dat_labor,  ml_rf_iivm(), ae98_labor_sub$workedm)
rf_income      <- fit_iivm(dat_income, ml_rf_iivm(), ae98_income_sub$lincomem)
pliv_rf_labor  <- fit_pliv(dat_labor,  ml_rf_pliv(), ae98_labor_sub$workedm)
pliv_rf_income <- fit_pliv(dat_income, ml_rf_pliv(), ae98_income_sub$lincomem)

learner_pair_table(pliv_rf_labor, pliv_rf_income, rf_labor, rf_income,
                   "Learner estimates -- Ranger PLR-IV and Ranger Wald-AIPW/LATE (7 covariates)")
```

### XGBoost (imported exports)

```{r ext-xgb-import}
xgb_export_path <- "ae98_xgb_tuning_export.rds"
if (file.exists(xgb_export_path)) {
  xgb_export <- readRDS(xgb_export_path)
  xgb_export$estimate_comparison %>%
    kbl(align = c("l","l","c","r"),
        caption = "LATE / Wald-AIPW estimate -- XGBoost, untuned vs. tune_on_folds=TRUE (imported, not refit here)") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    footnote(general = "Imported from ae98_xgb_tuning_export.rds. Estimate (SE) shown without significance stars (not computed at export time), unlike the freshly-fit linear/ranger tables above. Do not refit XGBoost here -- tune_on_folds=TRUE has fold-specific hyperparameters, not one recipe to reuse.")
} else {
  cat("ae98_xgb_tuning_export.rds not found -- run child_xgb_tuning.Rmd first to generate it.\n")
}

xgb_pliv_export_path <- "ae98_xgb_pliv_tuning_export.rds"
if (file.exists(xgb_pliv_export_path)) {
  xgb_pliv_export <- readRDS(xgb_pliv_export_path)
  xgb_pliv_export$estimate_comparison %>%
    kbl(align = c("l","l","c"),
        caption = "PLR-IV estimate -- XGBoost, untuned vs. tune_on_folds=TRUE (imported, not refit here)") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    footnote(general = "Imported from ae98_xgb_pliv_tuning_export.rds. Untuned uses the fixed base settings alpha=0, subsample=1, max_delta_step=0, base_score=0. Tuned uses tune_on_folds=TRUE, so fold-specific hyperparameters are preserved by importing the saved results rather than refitting here.")
} else {
  cat("ae98_xgb_pliv_tuning_export.rds not found -- run child_xgb_pliv_tuning.Rmd first to generate it.\n")
}
```

## Comparison Table

```{r ext-comparison}
ext_cell <- function(coef, se) {
  if (is.na(coef) || !is.finite(coef)) return("\u2014")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

learner_row <- function(label, fit_labor, fit_income) {
  data.frame(Estimator = label,
             `Labor (workedm)`   = ext_cell(get_dml_coef(fit_labor$obj),  get_dml_se(fit_labor$obj)),
             `Income (lincomem)` = ext_cell(get_dml_coef(fit_income$obj), get_dml_se(fit_income$obj)),
             check.names = FALSE, stringsAsFactors = FALSE)
}

kappa_ref_row <- function(label, key) {
  data.frame(Estimator = label,
             `Labor (workedm)`   = ext_cell(labor_cols[["1 if worked, 0 else"]]$kappa$estimates[key],
                                             labor_cols[["1 if worked, 0 else"]]$kappa$se[key]),
             `Income (lincomem)` = ext_cell(income_cols[["Dollars"]]$kappa$estimates[key],
                                             income_cols[["Dollars"]]$kappa$se[key]),
             check.names = FALSE, stringsAsFactors = FALSE)
}

comparison_rows <- list(
  learner_row("PLR-IV (linear)",                  pliv_lm_labor, pliv_lm_income),
  learner_row("Wald-AIPW / LATE (linear+logit)", lm_labor, lm_income),
  learner_row("PLR-IV (ranger)",                  pliv_rf_labor, pliv_rf_income),
  learner_row("Wald-AIPW / LATE (ranger)",       rf_labor, rf_income)
)

if (exists("xgb_export")) {
  xr <- xgb_export$estimate_comparison
  get_xgb_est <- function(sample_label, learner_label) {
    row <- xr[xr$Cell == sample_label & xr$Learner == learner_label, ]
    if (nrow(row) == 0) return("\u2014")
    row[["Estimate (SE)"]][1]
  }
  comparison_rows[[length(comparison_rows) + 1]] <- data.frame(
    Estimator = "Wald-AIPW / LATE (XGBoost, untuned)",
    `Labor (workedm)`   = get_xgb_est("Labor (workedm)",   "XGB untuned"),
    `Income (lincomem)` = get_xgb_est("Income (lincomem)", "XGB untuned"),
    check.names = FALSE, stringsAsFactors = FALSE)
  comparison_rows[[length(comparison_rows) + 1]] <- data.frame(
    Estimator = "Wald-AIPW / LATE (XGBoost, tune_on_folds=TRUE)",
    `Labor (workedm)`   = get_xgb_est("Labor (workedm)",   "XGB tuned (tune_on_folds=TRUE)"),
    `Income (lincomem)` = get_xgb_est("Income (lincomem)", "XGB tuned (tune_on_folds=TRUE)"),
    check.names = FALSE, stringsAsFactors = FALSE)
}
if (exists("xgb_pliv_export")) {
  xpr <- xgb_pliv_export$estimate_comparison
  get_xgb_pliv_est <- function(sample_label, learner_label) {
    row <- xpr[xpr$Cell == sample_label & xpr$Learner == learner_label, ]
    if (nrow(row) == 0) return("\u2014")
    row[["Estimate (SE)"]][1]
  }
  comparison_rows[[length(comparison_rows) + 1]] <- data.frame(
    Estimator = "PLR-IV (XGBoost, untuned)",
    `Labor (workedm)`   = get_xgb_pliv_est("Labor (workedm)",   "PLR-IV XGB untuned"),
    `Income (lincomem)` = get_xgb_pliv_est("Income (lincomem)", "PLR-IV XGB untuned"),
    check.names = FALSE, stringsAsFactors = FALSE)
  comparison_rows[[length(comparison_rows) + 1]] <- data.frame(
    Estimator = "PLR-IV (XGBoost, tune_on_folds=TRUE)",
    `Labor (workedm)`   = get_xgb_pliv_est("Labor (workedm)",   "PLR-IV XGB tuned (tune_on_folds=TRUE)"),
    `Income (lincomem)` = get_xgb_pliv_est("Income (lincomem)", "PLR-IV XGB tuned (tune_on_folds=TRUE)"),
    check.names = FALSE, stringsAsFactors = FALSE)
}

comparison_rows[[length(comparison_rows) + 1]] <- kappa_ref_row("\u03c4_cb_u (kappa, ref)", "tau_cb_u")
comparison_rows[[length(comparison_rows) + 1]] <- kappa_ref_row("\u03c4_ml_u (kappa, ref)", "tau_ml_u")

comp_df <- do.call(rbind, comparison_rows)
n_learner_rows <- sum(grepl("^(PLR-IV|Wald-AIPW)", comp_df$Estimator))

comp_df %>%
  kbl(align = c("l","c","c"),
      caption = "Learner Comparison vs. Kappa -- Labor (workedm) and Income (lincomem)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("DoubleML learners", 1, n_learner_rows) %>%
  pack_rows("Kappa reference",   n_learner_rows + 1, nrow(comp_df)) %>%
  row_spec(1:n_learner_rows, background = "#eef6ff") %>%
  footnote(general = "Each cell: estimate (SE). Blue = DoubleML/mlr3 learners. Linear/ranger PLR-IV rows are fresh DoubleMLPLIV fits. XGBoost rows are imported from the separate Wald-AIPW and PLR-IV tuning exports and show no significance stars. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

<div class="note-box">
**Separate checks.** The 6-covariate robustness check (kappa/ranger refit
with `sex2nd` dropped) and the Method-A translation-invariance rerun are kept
outside this main analysis notebook.
</div>

---

# 4. Weight Diagnostics

<div class="note-box">
**Reading the tables.** `Sum_w ≈ 0` confirms translation invariance.
**`ESS_mod`** = modified effective sample size `(Σ|ωᵢ|)²/Σωᵢ²`
(Chattopadhyay & Zubizarreta 2023), valid under negative weights. Unlike
Card (one grf fit per cell), AE98's Section 2 fit
**both** untuned and tuned honest forests, so both appear here as separate
rows — this lets you see whether `tune.parameters="all"` changed the weight
geometry, not just the point estimate. XGBoost rows are pulled from the
imported exports in Section 3. If an export file was not run, its rows are
simply omitted rather than breaking the table.
</div>

```{r diag-builders, echo=FALSE}
# Kappa weight bundles for both samples.
kw_labor  <- kappa_weights_bundle(Z_labor,  D_labor,  X_labor_kappa)
kw_income <- kappa_weights_bundle(Z_income, D_income, X_income_kappa)

DIAG_ALIGN <- c("l","r","r","r","r","r")

# Pull imported XGBoost weight vectors + identity flags, defensively (NULL/NA if not run yet).
get_xgb_w  <- function(sample_label, tuned_label) {
  if (!exists("xgb_export")) return(NULL)
  out <- tryCatch(xgb_export$omega_vectors[[sample_label]][[tuned_label]], error = function(e) NULL)
  out
}
get_xgb_ok <- function(sample_label, tuned_label) {
  if (!exists("xgb_export")) return(NA)
  out <- tryCatch(xgb_export$identity_checks[[sample_label]][[tuned_label]], error = function(e) NA)
  if (is.null(out)) NA else out
}
get_xgb_pliv_w <- function(sample_label, tuned_label) {
  if (!exists("xgb_pliv_export")) return(NULL)
  out <- tryCatch(xgb_pliv_export$omega_vectors[[sample_label]][[tuned_label]], error = function(e) NULL)
  out
}
get_xgb_pliv_ok <- function(sample_label, tuned_label) {
  if (!exists("xgb_pliv_export")) return(NA)
  out <- tryCatch(xgb_pliv_export$identity_checks[[sample_label]][[tuned_label]], error = function(e) NA)
  if (is.null(out)) NA else out
}

w_xgb_labor_untuned   <- get_xgb_w("Labor (workedm)",   "untuned")
w_xgb_labor_tuned     <- get_xgb_w("Labor (workedm)",   "tuned")
w_xgb_income_untuned  <- get_xgb_w("Income (lincomem)", "untuned")
w_xgb_income_tuned    <- get_xgb_w("Income (lincomem)", "tuned")
w_xgb_pliv_labor_untuned   <- get_xgb_pliv_w("Labor (workedm)",   "untuned")
w_xgb_pliv_labor_tuned     <- get_xgb_pliv_w("Labor (workedm)",   "tuned")
w_xgb_pliv_income_untuned  <- get_xgb_pliv_w("Income (lincomem)", "untuned")
w_xgb_pliv_income_tuned    <- get_xgb_pliv_w("Income (lincomem)", "tuned")
ok_xgb_labor_untuned  <- get_xgb_ok("Labor (workedm)",   "untuned")
ok_xgb_labor_tuned    <- get_xgb_ok("Labor (workedm)",   "tuned")
ok_xgb_income_untuned <- get_xgb_ok("Income (lincomem)", "untuned")
ok_xgb_income_tuned   <- get_xgb_ok("Income (lincomem)", "tuned")
ok_xgb_pliv_labor_untuned  <- get_xgb_pliv_ok("Labor (workedm)",   "untuned")
ok_xgb_pliv_labor_tuned    <- get_xgb_pliv_ok("Labor (workedm)",   "tuned")
ok_xgb_pliv_income_untuned <- get_xgb_pliv_ok("Income (lincomem)", "untuned")
ok_xgb_pliv_income_tuned   <- get_xgb_pliv_ok("Income (lincomem)", "tuned")

# Visible diagnostic: tells you WHICH of the two possible causes applies for
# each cell, instead of leaving that ambiguous in the Love-plot placeholder.
cat(sprintf("xgb_export loaded: %s\n", exists("xgb_export")))
cat(sprintf("xgb_pliv_export loaded: %s\n", exists("xgb_pliv_export")))
cat("XGBoost weight/identity status per cell (w=NULL means the export is missing\n")
cat("or that cell's weight extraction failed; ok=FALSE means weights ARE present\n")
cat("but sum(omega*Y) != tau_hat -- i.e. XGBoost fails the affine-smoother identity\n")
cat("for that cell, which is a real, reportable finding, not a bug):\n")
status_tbl <- data.frame(
  Estimator = "Wald-AIPW",
  Cell   = rep(c("Labor (workedm)", "Income (lincomem)"), each = 2),
  Tuned  = rep(c("untuned", "tuned"), times = 2),
  w_is_null = c(is.null(w_xgb_labor_untuned), is.null(w_xgb_labor_tuned),
                is.null(w_xgb_income_untuned), is.null(w_xgb_income_tuned)),
  identity_ok = c(ok_xgb_labor_untuned, ok_xgb_labor_tuned,
                  ok_xgb_income_untuned, ok_xgb_income_tuned)
)
status_tbl <- rbind(
  status_tbl,
  data.frame(
    Estimator = "PLR-IV",
    Cell   = rep(c("Labor (workedm)", "Income (lincomem)"), each = 2),
    Tuned  = rep(c("untuned", "tuned"), times = 2),
    w_is_null = c(is.null(w_xgb_pliv_labor_untuned), is.null(w_xgb_pliv_labor_tuned),
                  is.null(w_xgb_pliv_income_untuned), is.null(w_xgb_pliv_income_tuned)),
    identity_ok = c(ok_xgb_pliv_labor_untuned, ok_xgb_pliv_labor_tuned,
                    ok_xgb_pliv_income_untuned, ok_xgb_pliv_income_tuned)
  )
)
status_tbl %>%
  kbl(caption = "Imported XGBoost weight availability and identity status",
      align = c("l","l","l","c","c")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE, font_size = 12) %>%
  footnote(general = "PLR-IV XGBoost estimates are imported in Section 3, but PLR-IV XGBoost weights are omitted here when the PLR-IV export was created with EXTRACT_WEIGHTS = FALSE. Wald-AIPW XGBoost weights remain imported from ae98_xgb_tuning_export.rds.")

# Skip a row cleanly if the weight vector is NULL/empty, rather than erroring.
safe_diag_row <- function(w, label) if (is.null(w) || length(w) == 0) NULL else weight_diag(w, label)

dml_diag_cell <- function(w_untuned, w_tuned, kw) {
  do.call(rbind, Filter(Negate(is.null), list(
    weight_diag(w_untuned$plriv,    "PLR-IV (untuned forest)"),
    weight_diag(w_untuned$waldaipw, "Wald-AIPW (untuned forest)"),
    weight_diag(w_tuned$plriv,      "PLR-IV (tuned forest)"),
    weight_diag(w_tuned$waldaipw,   "Wald-AIPW (tuned forest)"),
    weight_diag(kw$tau_cb_u,        "\u03c4_cb_u (kappa)"),
    weight_diag(kw$tau_ml_u,        "\u03c4_ml_u (kappa)"),
    weight_diag(kw$tau_ml_a10,      "\u03c4_ml_a10 (kappa)"),
    weight_diag(kw$tau_ml_a,        "\u03c4_ml_a (kappa)"),
    weight_diag(kw$tau_ml_a1,       "\u03c4_ml_t/a1 (kappa)"),
    weight_diag(kw$tau_ml_a0,       "\u03c4_ml_a0 (kappa)")
  )))
}
style_dml_diag <- function(df, cap) {
  n_dml <- sum(grepl("forest", df[[1]]))
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(seq_len(n_dml), background = "#eef6ff") %>%
    footnote(general = "Blue = grf honest-forest DML (untuned + tuned). Sum_w close to 0 indicates translation invariance. ESS_mod is the modified effective sample size.")
}

ext_diag_block <- function(kw, w_lm_pliv, w_lm_iivm, w_rf_pliv, w_rf_iivm,
                           w_xgb_pliv_untuned, w_xgb_pliv_tuned,
                           w_xgb_untuned, w_xgb_tuned) {
  do.call(rbind, Filter(Negate(is.null), list(
    safe_diag_row(w_lm_pliv,     "PLR-IV (linear)"),
    safe_diag_row(w_lm_iivm,     "Wald-AIPW (linear+logit)"),
    safe_diag_row(w_rf_pliv,     "PLR-IV (ranger)"),
    safe_diag_row(w_rf_iivm,     "Wald-AIPW (ranger)"),
    safe_diag_row(w_xgb_pliv_untuned, "PLR-IV (XGBoost, untuned)"),
    safe_diag_row(w_xgb_pliv_tuned,   "PLR-IV (XGBoost, tune_on_folds=TRUE)"),
    safe_diag_row(w_xgb_untuned, "Wald-AIPW (XGBoost, untuned)"),
    safe_diag_row(w_xgb_tuned,   "Wald-AIPW (XGBoost, tune_on_folds=TRUE)"),
    weight_diag(kw$tau_cb_u,     "\u03c4_cb_u (kappa, CBPS)"),
    weight_diag(kw$tau_ml_u,     "\u03c4_ml_u (kappa, MLE)"),
    weight_diag(kw$tau_ml_a10,   "\u03c4_ml_a10 (kappa, MLE)"),
    weight_diag(kw$tau_ml_a,     "\u03c4_ml_a (kappa, unnorm)"),
    weight_diag(kw$tau_ml_a1,    "\u03c4_ml_t (kappa, unnorm)"),
    weight_diag(kw$tau_ml_a0,    "\u03c4_ml_a0 (kappa, unnorm)")
  )))
}
style_ext_diag <- function(df, cap) {
  n_learn <- sum(grepl("PLR-IV|Wald-AIPW", df[[1]], ignore.case = TRUE))
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(seq_len(n_learn), background = "#eef6ff") %>%
    footnote(general = "Blue = DoubleML/mlr3 learners. XGBoost rows are imported from the separate Wald-AIPW and PLR-IV tuning exports and omitted if the corresponding export is unavailable. ESS_mod is the modified effective sample size. Check the printed XGBoost status table above before treating XGBoost weights as clean outcome-weight objects.")
}
```

## DML vs. Kappa {.tabset}

### Labor (workedm)

```{r dml-diag-labor}
style_dml_diag(dml_diag_cell(w_labor_untuned, w_labor_tuned, kw_labor),
               "Weight Diagnostics -- DML (untuned + tuned forest) vs. Kappa | Labor")
```

### Income (lincomem)

```{r dml-diag-income}
style_dml_diag(dml_diag_cell(w_income_untuned, w_income_tuned, kw_income),
               "Weight Diagnostics -- DML (untuned + tuned forest) vs. Kappa | Income")
```

## Learners vs. Kappa {.tabset}

### Labor (workedm)

```{r ext-diag-labor}
style_ext_diag(ext_diag_block(kw_labor,
                              pliv_lm_labor$w, lm_labor$w,
                              pliv_rf_labor$w, rf_labor$w,
                              w_xgb_pliv_labor_untuned, w_xgb_pliv_labor_tuned,
                              w_xgb_labor_untuned, w_xgb_labor_tuned),
               "Weight Diagnostics -- Learners | Labor")
```

### Income (lincomem)

```{r ext-diag-income}
style_ext_diag(ext_diag_block(kw_income,
                              pliv_lm_income$w, lm_income$w,
                              pliv_rf_income$w, rf_income$w,
                              w_xgb_pliv_income_untuned, w_xgb_pliv_income_tuned,
                              w_xgb_income_untuned, w_xgb_income_tuned),
               "Weight Diagnostics -- Learners | Income")
```

## Weight-Statistics Bridge {.tabset}

<div class="note-box">
**What this is.** For every estimator, `weight_bridge()` reports the outcome
weights Overall / Treated / Control (control sign-flipped, Knaus's
convention -- "Control mass" reads \u2248 1 when normalized, not \u2212 1; see
`child_xgb_tuning.Rmd`'s footnote for the same point), alongside the modified
effective sample size. The headline is **ESS mod**.
</div>

```{r bridge-helpers, echo=FALSE}
cell_weight_list <- function(w_untuned, w_tuned, lm_pliv_w, lm_iivm_w,
                             rf_pliv_w, rf_iivm_w,
                             xgb_pliv_untuned_w, xgb_pliv_tuned_w,
                             xgb_w_untuned, xgb_w_tuned, kw) {
  Filter(Negate(is.null), list(
    "PLR-IV (untuned forest)"    = w_untuned$plriv,
    "Wald-AIPW (untuned forest)" = w_untuned$waldaipw,
    "PLR-IV (tuned forest)"      = w_tuned$plriv,
    "Wald-AIPW (tuned forest)"   = w_tuned$waldaipw,
    "PLR-IV (linear)"            = lm_pliv_w,
    "Wald-AIPW (linear+logit)"   = lm_iivm_w,
    "PLR-IV (ranger)"            = rf_pliv_w,
    "Wald-AIPW (ranger)"         = rf_iivm_w,
    "PLR-IV (XGBoost untuned)"   = xgb_pliv_untuned_w,
    "PLR-IV (XGBoost tuned)"     = xgb_pliv_tuned_w,
    "Wald-AIPW (XGBoost untuned)"= xgb_w_untuned,
    "Wald-AIPW (XGBoost tuned)"  = xgb_w_tuned,
    "\u03c4_cb_u (kappa)"        = kw$tau_cb_u,
    "\u03c4_ml_u (kappa)"        = kw$tau_ml_u,
    "\u03c4_ml_t (kappa)"        = kw$tau_ml_a1
  ))
}

bridge_cells <- list(
  "Labor (workedm)"   = list(D = D_labor,  w = cell_weight_list(w_labor_untuned,  w_labor_tuned,
                                pliv_lm_labor$w, lm_labor$w,
                                pliv_rf_labor$w, rf_labor$w,
                                w_xgb_pliv_labor_untuned, w_xgb_pliv_labor_tuned,
                                w_xgb_labor_untuned,  w_xgb_labor_tuned,  kw_labor)),
  "Income (lincomem)" = list(D = D_income, w = cell_weight_list(w_income_untuned, w_income_tuned,
                                pliv_lm_income$w, lm_income$w,
                                pliv_rf_income$w, rf_income$w,
                                w_xgb_pliv_income_untuned, w_xgb_pliv_income_tuned,
                                w_xgb_income_untuned, w_xgb_income_tuned, kw_income))
)

bridge_table <- function(wlist, D, cap) {
  ests <- names(wlist)
  df   <- do.call(rbind, lapply(ests, function(nm) weight_bridge(wlist[[nm]], D, nm)))
  disp <- df[, c("Scope","Min","Max","Pct_neg","Sum_top10","Sum_w","Sum_abs_w","ESS_mod")]

  kk <- disp %>%
    kbl(caption = cap, row.names = FALSE,
        col.names = c("Scope","Min","Max","%Neg","\u03a3Top10%","\u03a3\u03c9","\u03a3|\u03c9|","ESS mod"),
        align = c("l", rep("r", 7))) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 12) %>%
    add_header_above(c(" " = 1, "Knaus summary() statistics" = 6, "Effective sample size" = 1)) %>%
    column_spec(8, bold = TRUE, background = "#eef6ff")
  for (i in seq_along(ests)) kk <- pack_rows(kk, ests[i], (i - 1) * 3 + 1, i * 3)
  kk %>% footnote(general = "Control rows use Knaus's sign-flipped convention (mass reads around 1 when normalized, not -1). ESS mod equals the squared sum of absolute weights divided by the sum of squared weights.")
}
```

### Labor (workedm)

```{r bridge-labor}
bridge_table(bridge_cells[["Labor (workedm)"]]$w, bridge_cells[["Labor (workedm)"]]$D,
             "Weight-statistics bridge -- Labor (workedm)")
```

### Income (lincomem)

```{r bridge-income}
bridge_table(bridge_cells[["Income (lincomem)"]]$w, bridge_cells[["Income (lincomem)"]]$D,
             "Weight-statistics bridge -- Income (lincomem)")
```

---

# 5. Love Plots

<div class="note-box">
The Love plots summarize covariate balance induced by each estimator's
outcome weights. XGBoost panels are shown when the imported export contains
the corresponding weight vectors.
</div>

```{r love-helpers, echo=FALSE}
maybe_love <- function(label, w, D, X, ok = TRUE) {
  if (is.null(w) || length(w) == 0) {
    reason <- "excluded: ae98_xgb_tuning_export.rds\nnot found or weight extraction failed"
    return(grid::textGrob(paste0(label, "\n", reason),
                          gp = grid::gpar(fontsize = 10, col = "#b71c1c")))
  }
  if (!is.na(ok) && isFALSE(ok)) {
    label <- paste0(label, "\n")
  }
  make_love(label, w, D, X)
}
```

## DML vs. Kappa {.tabset}

### Labor (workedm) — all 10 estimators

```{r love-labor, fig.height=20, fig.width=12}
grid.arrange(grobs = list(
  make_love("PLR-IV (untuned)",     w_labor_untuned$plriv,    D_labor, X_dml_labor),
  make_love("Wald-AIPW (untuned)",  w_labor_untuned$waldaipw, D_labor, X_dml_labor),
  make_love("PLR-IV (tuned)",       w_labor_tuned$plriv,      D_labor, X_dml_labor),
  make_love("Wald-AIPW (tuned)",    w_labor_tuned$waldaipw,   D_labor, X_dml_labor),
  make_love("\u03c4_cb_u (CBPS)",   kw_labor$tau_cb_u,        D_labor, X_dml_labor),
  make_love("\u03c4_ml_u",          kw_labor$tau_ml_u,        D_labor, X_dml_labor),
  make_love("\u03c4_ml_a10",        kw_labor$tau_ml_a10,      D_labor, X_dml_labor),
  make_love("\u03c4_ml_a (unnorm)", kw_labor$tau_ml_a,        D_labor, X_dml_labor),
  make_love("\u03c4_ml_t (unnorm)", kw_labor$tau_ml_a1,       D_labor, X_dml_labor),
  make_love("\u03c4_ml_a0 (unnorm)",kw_labor$tau_ml_a0,       D_labor, X_dml_labor)
), ncol = 2, top = "Love Plots -- Labor (workedm): DML vs. Kappa")
```

### Income (lincomem) — all 10 estimators

```{r love-income, fig.height=20, fig.width=12}
grid.arrange(grobs = list(
  make_love("PLR-IV (untuned)",     w_income_untuned$plriv,    D_income, X_dml_income),
  make_love("Wald-AIPW (untuned)",  w_income_untuned$waldaipw, D_income, X_dml_income),
  make_love("PLR-IV (tuned)",       w_income_tuned$plriv,      D_income, X_dml_income),
  make_love("Wald-AIPW (tuned)",    w_income_tuned$waldaipw,   D_income, X_dml_income),
  make_love("\u03c4_cb_u (CBPS)",   kw_income$tau_cb_u,        D_income, X_dml_income),
  make_love("\u03c4_ml_u",          kw_income$tau_ml_u,        D_income, X_dml_income),
  make_love("\u03c4_ml_a10",        kw_income$tau_ml_a10,      D_income, X_dml_income),
  make_love("\u03c4_ml_a (unnorm)", kw_income$tau_ml_a,        D_income, X_dml_income),
  make_love("\u03c4_ml_t (unnorm)", kw_income$tau_ml_a1,       D_income, X_dml_income),
  make_love("\u03c4_ml_a0 (unnorm)",kw_income$tau_ml_a0,       D_income, X_dml_income)
), ncol = 2, top = "Love Plots -- Income (lincomem): DML vs. Kappa")
```

## Learners vs. Kappa {.tabset}

### Labor (workedm)

```{r ext-love-labor, fig.height=14, fig.width=12}
grid.arrange(grobs = list(
  make_love("Linear+Logit\nLabor",          lm_labor$w,           D_labor, X_dml_labor),
  make_love("Ranger\nLabor",                rf_labor$w,           D_labor, X_dml_labor),
  maybe_love("XGBoost untuned\nLabor",      w_xgb_labor_untuned,  D_labor, X_dml_labor, ok_xgb_labor_untuned),
  maybe_love("XGBoost tuned\nLabor",        w_xgb_labor_tuned,    D_labor, X_dml_labor, ok_xgb_labor_tuned),
  make_love("\u03c4_cb_u\nLabor",           kw_labor$tau_cb_u,    D_labor, X_dml_labor),
  make_love("\u03c4_ml_u\nLabor",           kw_labor$tau_ml_u,    D_labor, X_dml_labor)
), ncol = 2, top = "Love Plots -- DoubleML learners vs. kappa | Labor (workedm)")
```

### Income (lincomem)

```{r ext-love-income, fig.height=14, fig.width=12}
grid.arrange(grobs = list(
  make_love("Linear+Logit\nIncome",         lm_income$w,           D_income, X_dml_income),
  make_love("Ranger\nIncome",               rf_income$w,           D_income, X_dml_income),
  maybe_love("XGBoost untuned\nIncome",     w_xgb_income_untuned,  D_income, X_dml_income, ok_xgb_income_untuned),
  maybe_love("XGBoost tuned\nIncome",       w_xgb_income_tuned,    D_income, X_dml_income, ok_xgb_income_tuned),
  make_love("\u03c4_cb_u\nIncome",          kw_income$tau_cb_u,    D_income, X_dml_income),
  make_love("\u03c4_ml_u\nIncome",          kw_income$tau_ml_u,    D_income, X_dml_income)
), ncol = 2, top = "Love Plots -- DoubleML learners vs. kappa | Income (lincomem)")
```

<div class="note-box">
The remaining robustness and translation-invariance rerun checks are handled
outside this main notebook.
</div>
