---
title: "AE98 (Angrist & Evans 1998) — XGBoost Nuisance Tuning for Wald-AIPW"
subtitle: "Nested nuisance tuning by predictive loss, tuned inside each outer cross-fitting fold"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 5.5)
```

```{r packages}
library(knitr)
library(kableExtra)
library(dplyr)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

source("functions_all.R")

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

SEED        = 42
N_CF        = 5
TUNE_EVALS  = 15   # started conservative given tune_on_folds=TRUE multiplies cost
                   # (retunes inside every outer fold) and we now have 2 samples,
                   # not Card's 1. Raise to 20-30 only after confirming runtime on
                   # ONE sample first (see warning below).
INNER_FOLDS = 3
```

<div class="note-box">
**What changed vs. the Card tuning script, and why.** Per the discussion with
Alexander's advisor: the Card tuning notebook tuned XGBoost hyperparameters
**once, globally**, on the whole sample (`tune_on_folds` left at its default,
`FALSE`), then reused that single recipe across all 5 outer cross-fitting
folds. This is a defensible sensitivity check but leaves a mild,
avoidable leakage: an observation that later serves as an outer-fold test
row may have influenced the hyperparameter choice used to predict it. This
notebook uses `tune_on_folds = TRUE` instead: hyperparameters are re-selected
**separately inside each outer training fold**, using only that fold's own
data, before that fold's nuisance models are fit and used to predict the
held-out rows. This removes that source of leakage; it does **not** guarantee
unbiasedness in finite samples, fix weak overlap, or fix XGBoost's
non-affine-smoother problem for the outcome-weight identity — those are
separate issues, addressed in the diagnostics below.

**Scope, per instruction:** one binary outcome (`workedm`, labor sample) and
one continuous outcome (`lincomem`, income sample) — not every recoding.
Running `tune_on_folds = TRUE` across every recoding of both outcome
families would multiply runtime far beyond what's useful here.
</div>

<div class="warn-box">
**Runtime warning.** `tune_on_folds = TRUE` reruns the inner 3-fold random
search **separately for each of the 5 outer folds**, for each of the
(likely 5) nuisance sub-models DoubleMLIIVM fits internally (`ml_g0`,
`ml_g1`, `ml_m`, `ml_r0`, `ml_r1`). That is roughly 5x the tuning cost of
the Card script's global-tuning version, per sample. **Recommendation:**
first knit with only the Labor cell active (comment out Income in the
`cells` list below), confirm the runtime is acceptable, then add Income
back in and/or raise `TUNE_EVALS`.
</div>

---

# Setup

```{r data}
obj_data <- readRDS("ae98_sub.rds")
if (!all(c("ae98_labor_sub", "ae98_income_sub") %in% names(obj_data))) {
  stop("ae98_sub.rds must contain ae98_labor_sub and ae98_income_sub.")
}

ae98_labor_sub  <- obj_data$ae98_labor_sub
ae98_income_sub <- obj_data$ae98_income_sub

ae98_labor_sub <- ae98_labor_sub %>%
  mutate(workedm = as.integer(workedm),
         morekids = as.integer(morekids),
         samesex  = as.integer(samesex))

ae98_income_sub <- ae98_income_sub %>%
  mutate(morekids  = as.integer(morekids),
         samesex   = as.integer(samesex),
         lincomem  = log(incomem))

RHS_FORMULA <- ~ 0 + agem + agefstm + sexk + sex2nd + blackm + hispm + othracem
X_dml_labor  <- make_X(RHS_FORMULA, data = ae98_labor_sub)
X_dml_income <- make_X(RHS_FORMULA, data = ae98_income_sub)

D_labor  <- ae98_labor_sub$morekids;  Z_labor  <- ae98_labor_sub$samesex
D_income <- ae98_income_sub$morekids; Z_income <- ae98_income_sub$samesex

cat(sprintf("Labor:  N=%d, %d covariates | Income: N=%d, %d covariates\n",
            nrow(ae98_labor_sub), ncol(X_dml_labor),
            nrow(ae98_income_sub), ncol(X_dml_income)))
```

```{r learners-and-tuning-setup}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                    x_cols = colnames(X))
}

make_xgb_learners <- function() {
  list(
    g = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
    m = lrn("classif.xgboost", predict_type = "prob"),
    r = lrn("classif.xgboost", predict_type = "prob")
  )
}

xgb_ps <- function() {
  ps(nrounds = p_int(50, 400), max_depth = p_int(2, 5),
     eta = p_dbl(0.02, 0.20), min_child_weight = p_int(1, 8))
}

tune_param_set <- function() list(ml_g = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())

tune_settings <- function() {
  list(
    n_folds_tune = INNER_FOLDS,
    rsmp_tune    = rsmp("cv", folds = INNER_FOLDS),
    measure      = list(ml_g = "regr.rmse", ml_m = "classif.logloss", ml_r = "classif.logloss"),
    terminator   = trm("evals", n_evals = TUNE_EVALS),
    algorithm    = tnr("random_search")
  )
}

get_prediction_vector <- function(obj, learner_name) {
  pred <- tryCatch(obj$predictions[[learner_name]], error = function(e) NULL)
  if (is.null(pred)) return(numeric(0))
  as.numeric(drop(pred))
}

get_branch_prediction <- function(obj, base_name, Z) {
  pred0 <- get_prediction_vector(obj, paste0(base_name, "0"))
  pred1 <- get_prediction_vector(obj, paste0(base_name, "1"))
  if (length(pred0) != length(Z) || length(pred1) != length(Z)) return(numeric(0))
  ifelse(as.numeric(Z) == 1, pred1, pred0)
}

# FIX vs. the Card script: DoubleMLIIVM stores the treatment nuisance by branch
# (ml_r0 / ml_r1), not as a flat "ml_r". The Card script's boundary-share
# function looked for "ml_r" directly, which silently returns numeric(0) for
# r_hat. Fixed here to use get_branch_prediction(), matching what the loss
# function already did correctly.
propensity_boundary_share <- function(obj, Z, threshold = 0.02) {
  m_hat <- get_prediction_vector(obj, "ml_m")
  r_hat <- get_branch_prediction(obj, "ml_r", Z)
  p_hat <- c(m_hat, r_hat)
  p_hat <- p_hat[is.finite(p_hat)]
  if (length(p_hat) == 0) return(NA_real_)
  mean(p_hat < threshold | p_hat > 1 - threshold)
}

rmse_loss <- function(truth, pred) {
  if (length(pred) != length(truth) || length(pred) == 0) return(NA_real_)
  truth <- as.numeric(truth); pred <- as.numeric(pred)
  keep <- is.finite(truth) & is.finite(pred)
  if (!any(keep)) return(NA_real_)
  sqrt(mean((truth[keep] - pred[keep])^2))
}

logloss_loss <- function(truth, pred, eps = 1e-6) {
  if (length(pred) != length(truth) || length(pred) == 0) return(NA_real_)
  truth <- as.numeric(truth); pred <- as.numeric(pred)
  keep <- is.finite(truth) & is.finite(pred)
  if (!any(keep)) return(NA_real_)
  truth <- truth[keep]; pred <- pmin(pmax(pred[keep], eps), 1 - eps)
  -mean(truth * log(pred) + (1 - truth) * log(1 - pred))
}

calc_nuisance_losses <- function(r, cell) {
  if (is.null(r)) return(c(outcome_rmse = NA_real_, instrument_logloss = NA_real_, treatment_logloss = NA_real_))
  g_hat <- get_branch_prediction(r$obj, "ml_g", cell$Z)
  m_hat <- get_prediction_vector(r$obj, "ml_m")
  r_hat <- get_branch_prediction(r$obj, "ml_r", cell$Z)
  c(outcome_rmse        = rmse_loss(cell$Y, g_hat),
    instrument_logloss   = logloss_loss(cell$Z, m_hat),
    treatment_logloss    = logloss_loss(cell$D, r_hat))
}

fit_cell <- function(dml_data, learners, label, Z, tune = FALSE, tune_on_folds = FALSE) {
  tryCatch({
    set.seed(SEED)
    obj <- DoubleMLIIVM$new(data = dml_data, ml_g = learners$g, ml_m = learners$m,
                             ml_r = learners$r, n_folds = N_CF, score = "LATE")
    if (tune) {
      obj$tune(param_set = tune_param_set(), tune_settings = tune_settings(),
               tune_on_folds = tune_on_folds)
    }
    obj$fit(store_models = TRUE, store_predictions = TRUE)

    list(label = label, tuned = tune,
         coef = get_dml_coef(obj), se = get_dml_se(obj),
         boundary_share = propensity_boundary_share(obj, Z),
         tuning_result = tryCatch(obj$tuning_result, error = function(e) NULL),
         params        = tryCatch(obj$params,        error = function(e) NULL),
         obj = obj)
  }, error = function(e) {
    message(sprintf("Fit failed for %s: %s", label, conditionMessage(e)))
    NULL
  })
}

# Only 2 cells: one binary outcome (labor), one continuous outcome (income) —
# not every recoding of each family.
cells <- list(
  "Labor (workedm)"   = list(Y = ae98_labor_sub$workedm,  D = D_labor,  Z = Z_labor,  X = X_dml_labor),
  "Income (lincomem)" = list(Y = ae98_income_sub$lincomem, D = D_income, Z = Z_income, X = X_dml_income)
)
# To test runtime on one cell first, uncomment:
# cells <- cells["Labor (workedm)"]
```

---

# Fit: XGBoost untuned vs. tuned (tune_on_folds = TRUE)

```{r fit-all, cache=TRUE}
results <- list()
for (cell_name in names(cells)) {
  cell <- cells[[cell_name]]
  dml_data <- make_dml_data(cell$Y, cell$D, cell$Z, cell$X)

  results[[cell_name]] <- list(
    untuned = fit_cell(dml_data, make_xgb_learners(), paste(cell_name, "xgb-untuned"),
                        Z = cell$Z, tune = FALSE),
    tuned   = fit_cell(dml_data, make_xgb_learners(), paste(cell_name, "xgb-tuned"),
                        Z = cell$Z, tune = TRUE, tune_on_folds = TRUE)
  )
  cat(sprintf("Finished: %s\n", cell_name))
}
```

## 1. Estimate comparison

```{r comparison-table}
fmt_est <- function(r) if (is.null(r)) "\u2014" else sprintf("%.3f (%.3f)", r$coef, r$se)
fmt_boundary <- function(r) if (is.null(r) || !is.finite(r$boundary_share)) "\u2014" else sprintf("%.4f", r$boundary_share)

rows <- do.call(rbind, lapply(names(results), function(cell_name) {
  R <- results[[cell_name]]
  data.frame(Cell = cell_name, Learner = c("XGB untuned", "XGB tuned (tune_on_folds=TRUE)"),
             `Estimate (SE)` = c(fmt_est(R$untuned), fmt_est(R$tuned)),
             `Propensity boundary share` = c(fmt_boundary(R$untuned), fmt_boundary(R$tuned)),
             check.names = FALSE, stringsAsFactors = FALSE)
}))

rows %>%
  kbl(align = c("l","l","c","r"), caption = "AE98 Wald-AIPW: XGBoost untuned vs. tuned inside each outer fold") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(rows$Learner == "XGB tuned (tune_on_folds=TRUE)"), background = "#eef6ff") %>%
  footnote(general = "Propensity boundary share = fraction of fitted instrument/treatment propensities below 0.02 or above 0.98. Hyperparameters for the tuned row were selected by 3-fold random-search CV on nuisance prediction loss, run separately inside each outer cross-fitting fold.")
```

## 2. Nuisance-prediction losses (untuned vs. tuned)

<div class="note-box">
This table is unaffected by the `tune_on_folds` change — `obj$predictions`
stores the out-of-fold cross-fitted predictions regardless of how the
hyperparameters were selected, so the loss numbers here are genuine
out-of-sample RMSE/log-loss either way. Presented as a table rather than a
bar chart, since with only 2 cells a plot would not add clarity over the
numbers directly.
</div>

```{r nuisance-loss-table}
loss_rows <- do.call(rbind, lapply(names(results), function(cell_name) {
  R <- results[[cell_name]]; cell <- cells[[cell_name]]
  ul <- calc_nuisance_losses(R$untuned, cell); tl <- calc_nuisance_losses(R$tuned, cell)
  data.frame(Cell = cell_name, Learner = c("XGB untuned", "XGB tuned"),
             `Outcome RMSE` = c(ul["outcome_rmse"], tl["outcome_rmse"]),
             `Instrument log-loss` = c(ul["instrument_logloss"], tl["instrument_logloss"]),
             `Treatment log-loss` = c(ul["treatment_logloss"], tl["treatment_logloss"]),
             check.names = FALSE, row.names = NULL)
}))

loss_rows %>%
  kbl(digits = 4, caption = "Nuisance-prediction loss: untuned vs. tuned (lower is better)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(loss_rows$Learner == "XGB tuned"), background = "#eef6ff") %>%
  footnote(general = "If tuned losses are lower AND the Section 1 estimate/boundary-share is more stable, that is evidence the untuned XGBoost's behavior was a bad-default artifact rather than a genuine design problem.")
```

## 3. Selected hyperparameters

<div class="warn-box">
**Please verify this section locally before trusting it.** With
`tune_on_folds = TRUE`, hyperparameters are chosen **separately for each of
the 5 outer folds**, so there is no longer one single "best" set — there
are up to 5 per nuisance sub-model. I have not been able to execute this
against your installed `DoubleML` package version, so the exact field name
and structure of `obj$params` / `obj$tuning_result` may not match what this
extraction code assumes. If this chunk errors or returns nothing useful,
run `str(results[["Labor (workedm)"]]$tuned$obj$params)` yourself first and
tell me what it looks like — I'll fix the extraction to match.
</div>

```{r hyperparameter-summary}
extract_tuned_params <- function(r, cell_name) {
  if (is.null(r) || !r$tuned) return(data.frame())
  src <- if (!is.null(r$params)) r$params else r$tuning_result
  if (is.null(src)) return(data.frame())

  u <- unlist(src, recursive = TRUE)
  u <- u[grepl("nrounds|max_depth|eta|min_child_weight", names(u))]
  if (length(u) == 0) return(data.frame())

  hp <- data.frame(Cell = cell_name, Raw_name = names(u), Value = as.numeric(u), stringsAsFactors = FALSE)
  hp$Nuisance <- ifelse(grepl("ml_g", hp$Raw_name), "Outcome g",
                  ifelse(grepl("ml_m", hp$Raw_name), "Instrument m",
                  ifelse(grepl("ml_r", hp$Raw_name), "Treatment r", "Unknown")))
  hp$Branch <- ifelse(grepl("ml_g0|ml_r0", hp$Raw_name), "Z=0",
                ifelse(grepl("ml_g1|ml_r1", hp$Raw_name), "Z=1", ""))
  hp$Parameter <- sub(".*(nrounds|max_depth|eta|min_child_weight).*", "\\1", hp$Raw_name)
  hp
}

hp_raw <- do.call(rbind, lapply(names(results), function(cn) extract_tuned_params(results[[cn]]$tuned, cn)))

if (nrow(hp_raw) > 0) {
  hp_raw <- hp_raw[hp_raw$Nuisance != "Unknown", ]
  # Since tune_on_folds=TRUE gives up to 5 values per (Cell, Nuisance, Branch,
  # Parameter) -- one per outer fold -- we summarize with the median across
  # folds, and report how many distinct fold-level values contributed, so
  # this is honestly labeled as a summary, not "the" single best set.
  hp_summary <- hp_raw %>%
    group_by(Cell, Nuisance, Branch, Parameter) %>%
    summarise(Median_across_folds = median(Value), N_folds_seen = n(), .groups = "drop") %>%
    arrange(Cell, Nuisance, Branch, Parameter)

  hp_summary %>%
    kbl(digits = 3, caption = "Selected XGBoost hyperparameters: median value across the 5 outer folds (tune_on_folds = TRUE)") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
    collapse_rows(columns = c(1,2,3), valign = "middle") %>%
    footnote(general = "Each outer fold selects its own hyperparameters via 3-fold inner CV on nuisance prediction loss; there is no single global 'best' set under tune_on_folds=TRUE. N_folds_seen should be <=5; if it is much less, some folds' parameters were not extracted -- verify obj$params structure manually.")
} else {
  cat("No hyperparameters could be extracted automatically -- run str(results[[...]]$tuned$obj$params) and share the structure so the extraction can be fixed.\n")
}
```

## 4. Diagnostics for untuned vs. tuned (outcome-weight / ESS comparison)

<div class="note-box">
This part does not depend on how tuning was done — `get_outcome_weights()`
reads off the final fitted object, which has the same structure whether
tuning was global, per-fold, or skipped entirely. Fully recuperated from the
Card tuning script's Section 4, adapted to 2 cells instead of 4, with the
DML outcome-weight identity check added explicitly (per the advisor's point
that XGBoost is not guaranteed to satisfy it).
</div>

```{r xgb-tuned-weight-table}
omega_vec <- function(om) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && "Wald-AIPW" %in% rownames(w)) return(as.vector(w["Wald-AIPW", ]))
  as.vector(w[1, ])
}

extract_xgb_w <- function(r, cell) {
  dml_data <- make_dml_data(cell$Y, cell$D, cell$Z, cell$X)
  om <- get_outcome_weights(object = r$obj, dml_data = dml_data)
  omega_vec(om)
}

one_xgb_diag_row <- function(cell_name, learner_name, r, cell) {
  w  <- extract_xgb_w(r, cell)
  wb <- weight_bridge(w, cell$D, learner_name)
  ov <- wb[wb$Scope == "Overall", ]; tr <- wb[wb$Scope == "Treated", ]; co <- wb[wb$Scope == "Control", ]

  identity_ok <- tryCatch(
    check_weight_identity(w, cell$Y, r$coef),
    error = function(e) NA
  )

  data.frame(Cell = cell_name, Learner = learner_name, Estimate = as.numeric(r$coef),
             ESS = ov$ESS_kish, ESS_mod = ov$ESS_mod,
             `Sigma omega` = ov$Sum_w, `Sigma abs omega` = ov$Sum_abs_w,
             `Treated mass` = tr$Sum_w, `Control mass (-sum control weights)` = co$Sum_w,
             `max abs omega` = max(abs(w)),
             `Identity ok` = identity_ok,
             check.names = FALSE)
}

xgb_tuned_weight_table <- do.call(rbind, lapply(names(cells), function(cell_name) {
  cell <- cells[[cell_name]]; R <- results[[cell_name]]
  rbind(one_xgb_diag_row(cell_name, "XGB untuned", R$untuned, cell),
        one_xgb_diag_row(cell_name, "XGB tuned",   R$tuned,   cell))
}))

print_tbl <- xgb_tuned_weight_table
for (col in c("Estimate","Sigma omega","Sigma abs omega","Treated mass","Control mass (-sum control weights)","max abs omega"))
  print_tbl[[col]] <- sprintf("%.4f", print_tbl[[col]])
print_tbl$ESS     <- sprintf("%.0f", print_tbl$ESS)
print_tbl$ESS_mod <- sprintf("%.0f", print_tbl$ESS_mod)

print_tbl %>%
  kbl(caption = "AE98: XGBoost outcome-weight diagnostics, untuned vs. tuned") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12) %>%
  row_spec(which(xgb_tuned_weight_table$Learner == "XGB tuned"), background = "#eef6ff") %>%
  footnote(general = "Identity ok = TRUE means sum(omega*Y) reproduces the DML coefficient to machine precision (Knaus's affine-smoother condition). If FALSE for XGBoost, treat XGBoost here as a learner-sensitivity stress test rather than a clean benchmark -- the normalized kappa estimators remain the algebraically exact reference. Control mass is reported after weight_bridge()'s sign flip (Knaus convention: it computes weight_stats_block(-w[D==0])), so full normalization corresponds to treated mass approx. 1 AND control mass approx. 1 -- the untransformed control weights themselves sum to approximately -1, not +1.")
```

<div class="note-box">
**Not built in this file, on purpose (kept out of scope per instruction):**
Love plots for XGBoost untuned vs. tuned. The mechanics are identical to the
Card tuning script's Section 5 (`make_love()` on the extracted weight
vectors) and can be added in one small chunk once the above is checked and
confirmed to run correctly.
</div>

---

## 5. Export for child_test.Rmd

<div class="warn-box">
**Do not plug the Section 3 median hyperparameters into `child_test.Rmd` as
a fixed recipe.** `tune_on_folds = TRUE` produces fold-specific
hyperparameters (up to 5 per nuisance sub-model, one per outer fold) — there
is no single global recipe to hand off. Feeding the median values into a
fresh `DoubleMLIIVM` fit in `child_test.Rmd` would estimate a *different*
object (a single-recipe-everywhere XGBoost, closer to the old Card-style
`tune_on_folds = FALSE` logic), not a reproduction of what this notebook
actually fit. Instead, this chunk exports the fitted results, weight
vectors, and diagnostic tables themselves, so `child_test.Rmd` can read and
display them without re-running (expensive) tuning and without silently
changing what was estimated.
</div>

```{r xgb-omega-vectors}
# Store the extracted outcome-weight vector for every (cell, tuned/untuned)
# combination, so downstream files can rebuild Love plots or rerun
# translation-invariance checks against the exact same fitted weights.
xgb_omega_vectors <- lapply(names(cells), function(cell_name) {
  list(
    untuned = extract_xgb_w(results[[cell_name]]$untuned, cells[[cell_name]]),
    tuned   = extract_xgb_w(results[[cell_name]]$tuned,   cells[[cell_name]])
  )
})
names(xgb_omega_vectors) <- names(cells)
```

```{r xgb-identity-checks}
# Explicit identity check per (cell, tuned/untuned), stored separately from
# the printed table so downstream files can branch on it programmatically
# (e.g. "only Love-plot XGBoost if TRUE") without re-parsing xgb_tuned_weight_table.
xgb_identity_checks <- lapply(names(cells), function(cell_name) {
  cell <- cells[[cell_name]]
  list(
    untuned = tryCatch(
      check_weight_identity(xgb_omega_vectors[[cell_name]]$untuned, cell$Y, results[[cell_name]]$untuned$coef),
      error = function(e) NA
    ),
    tuned = tryCatch(
      check_weight_identity(xgb_omega_vectors[[cell_name]]$tuned, cell$Y, results[[cell_name]]$tuned$coef),
      error = function(e) NA
    )
  )
})
names(xgb_identity_checks) <- names(cells)

cat("Identity checks (sum(omega*Y) = tau_hat):\n")
for (cn in names(xgb_identity_checks)) {
  cat(sprintf("  %-20s untuned=%s | tuned=%s\n", cn,
              xgb_identity_checks[[cn]]$untuned, xgb_identity_checks[[cn]]$tuned))
}
```

```{r xgb-export, cache=FALSE}
xgb_export <- list(
  estimate_comparison = rows,
  nuisance_losses     = loss_rows,
  hyperparameters     = if (exists("hp_summary")) hp_summary else NULL,
  weight_diagnostics  = xgb_tuned_weight_table,
  omega_vectors       = xgb_omega_vectors,
  identity_checks     = xgb_identity_checks,
  settings = list(
    seed          = SEED,
    n_cf          = N_CF,
    tune_evals    = TUNE_EVALS,
    inner_folds   = INNER_FOLDS,
    tune_on_folds = TRUE,
    learner       = "xgboost",
    dml_class     = "DoubleMLIIVM",
    score         = "LATE"
  ),
  sample_info = list(
    labor_n       = nrow(ae98_labor_sub),
    income_n      = nrow(ae98_income_sub),
    labor_mean_Z  = mean(Z_labor),
    labor_mean_D  = mean(D_labor),
    income_mean_Z = mean(Z_income),
    income_mean_D = mean(D_income)
  )
)

saveRDS(xgb_export, "ae98_xgb_tuning_export.rds")
cat("Saved: ae98_xgb_tuning_export.rds\n")
cat("Contents: estimate_comparison, nuisance_losses, hyperparameters (descriptive only, do not re-plug),\n")
cat("          weight_diagnostics, omega_vectors, identity_checks, settings, sample_info.\n")
```

<div class="note-box">
**In `child_test.Rmd` later (not done yet, per instruction):**
```r
xgb_export <- readRDS("ae98_xgb_tuning_export.rds")
xgb_export$estimate_comparison   # untuned vs. tuned point estimates/SE table
xgb_export$weight_diagnostics    # ESS / Sigma-omega / identity-check table
xgb_export$omega_vectors[["Labor (workedm)"]]$tuned   # raw weight vector, for Love plots etc.
xgb_export$identity_checks[["Income (lincomem)"]]$tuned  # TRUE/FALSE gate before trusting XGBoost weights
```
This keeps `child_test.Rmd` fast (no re-tuning on every knit) and keeps the
tuned-on-folds estimator intact (no hyperparameter re-plugging that would
silently estimate a different object).
</div>

