---
title: "Card (1995) — Outcome Weights & Kappa Estimators"
subtitle: "College Proximity Instrument: SUW Table 3 Replication, DML & Learner Comparison"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body   { font-size: 15px; color: #2c2c2c; }
h1     { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2     { color: #2c7bb6; margin-top: 28px; }
h3     { color: #444; margin-top: 20px; }

.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 12,
  fig.height = 6
)
```

---

# Setup

## Packages & Functions

```{r packages-and-source}
# Core
library(haven)
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(dplyr)
# DML smoother + balance plots
library(OutcomeWeights)
library(cobalt)
library(viridis)
library(gridExtra)
library(ggplot2)
# DoubleML learners
library(DoubleML)
library(mlr3verse)

source("functions_all.R")   # all estimators, weight diagnostics (incl. ESS_mod),
                               # get_dml_coef / get_dml_se, make_X, make_love, etc.
cat("functions_all.R sourced.\n")
```

## Data & Design

<div class="note-box">
**Setting.** Card (1995) NLSYM subsample (N = 3,010). Instrument Z = `nearc4` (grew up near a 4-year college). Two binary treatments: **somecol** = 1{educ > 12} and **educ16** = 1{educ ≥ 16}. Outcome: log wages, reported in two units (dollars / cents) only because SUW Table 3 is laid out that way. Two covariate specs: **Card** (exper, expersq, black, south, smsa, smsa66, reg661–668) and **Kitagawa** (black, south, smsa, smsa66, south66).
</div>

```{r data}
card       <- read_dta("card.dta")

# Sample restriction: N = 3,010
card_clean <- card[!is.na(card$wage) & !is.na(card$educ), ]

# Outcomes in two units: lwage_cnt = lwage + log(100)
card_clean$lwage_cnt <- log(card_clean$wage)         # cents
card_clean$lwage     <- log(card_clean$wage / 100)   # dollars

# Binary treatments
card_clean$somecol <- as.integer(card_clean$educ > 12)
card_clean$educ16  <- as.integer(card_clean$educ >= 16)

# Instrument
card_clean$Z_inst <- as.integer(card_clean$nearc4)

# expersq if missing
if (!"expersq" %in% names(card_clean)) {
  card_clean$expersq <- card_clean$exper^2
}

cat(sprintf("N = %d (should be 3,010)\n", nrow(card_clean)))
cat(sprintf("somecol mean = %.4f | educ16 mean = %.4f | nearc4 mean = %.4f\n",
            mean(card_clean$somecol), mean(card_clean$educ16),
            mean(card_clean$Z_inst)))
```

```{r design-matrices}
# Kappa design matrices — WITH intercept (for logit / CBPS)
X_card_kappa <- model.matrix(
  ~ exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card_clean
)
X_card_df <- as.data.frame(X_card_kappa[, -1, drop = FALSE])

X_kit_kappa <- model.matrix(
  ~ black + south + smsa + smsa66 + south66,
  data = card_clean
)
X_kit_df <- as.data.frame(X_kit_kappa[, -1, drop = FALSE])

# DML design matrices — no intercept, attributes stripped for grf (make_X)
X_dml_card <- make_X(
  ~ 0 + exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card_clean
)
X_dml_kit <- make_X(
  ~ 0 + black + south + smsa + smsa66 + south66,
  data = card_clean
)

# Core vectors
Y_cnt <- card_clean$lwage_cnt   # cents  (used only for the SUW Table 3 layout)
Y_dol <- card_clean$lwage       # dollars (the working outcome everywhere else)
Z     <- card_clean$Z_inst
D1    <- card_clean$somecol
D2    <- card_clean$educ16

cat(sprintf("Kappa: Card %d / Kitagawa %d covariates | DML: Card %dx%d / Kitagawa %dx%d\n",
            ncol(X_card_kappa) - 1, ncol(X_kit_kappa) - 1,
            nrow(X_dml_card), ncol(X_dml_card), nrow(X_dml_kit), ncol(X_dml_kit)))
```

## Design-Stage Instrument-Propensity Balance

<div class="note-box">
**Distinct diagnostic.** These plots assess whether the fitted instrument propensity
score balances the observed covariates between the instrument groups
(`Z = 1` versus `Z = 0`). They use positive, separately normalized instrument-IPW
weights and therefore describe conventional weighted pseudo-populations. They do not
replace the outcome-weight Love plots below, which assess the signed treated-control
comparison generating each fitted LATE estimate. Continuous covariates are standardized;
starred binary covariates are shown as raw differences in proportions.
</div>

```{r instrument-ipw-balance, fig.height=7, fig.width=14}
# Positive Hájek-normalized instrument-IPW weights. Scaling each Z arm to sum
# to one is explicit here, although cobalt's within-arm weighted means are
# invariant to a common rescaling within an arm.
instrument_ipw <- function(Z, p) {
  Z <- as.numeric(Z)
  p <- as.numeric(p)
  if (length(Z) != length(p) || any(!is.finite(p)) || any(p <= 0 | p >= 1)) {
    stop("Instrument-IPW balance requires finite propensity scores strictly inside (0, 1).")
  }
  w <- ifelse(Z == 1, 1 / p, 1 / (1 - p))
  w[Z == 1] <- w[Z == 1] / sum(w[Z == 1])
  w[Z == 0] <- w[Z == 0] / sum(w[Z == 0])
  w
}

make_instrument_ipw_love <- function(title_str, Z, X_bal, p_ml, p_cb,
                                      threshold = 0.1) {
  love.plot(
    Z ~ X_bal,
    weights = list(
      "MLE instrument IPW"  = instrument_ipw(Z, p_ml),
      "CBPS instrument IPW" = instrument_ipw(Z, p_cb)
    ),
    position   = "bottom",
    title      = title_str,
    thresholds = c(m = threshold),
    var.order  = "unadjusted",
    continuous = "std",
    binary     = "raw",
    s.d.denom  = "pooled",
    stats      = "mean.diffs",
    abs        = TRUE,
    line       = TRUE,
    stars      = "raw",
    colors     = viridis(3),
    shapes     = c("circle", "triangle", "diamond")
  )
}

p_ml_card <- logit_mle(Z, X_card_kappa)
p_cb_card <- get_cbps_p(Z, X_card_kappa)
p_ml_kit  <- logit_mle(Z, X_kit_kappa)
p_cb_kit  <- get_cbps_p(Z, X_kit_kappa)

grid.arrange(
  make_instrument_ipw_love(
    "Instrument balance — Card controls",
    Z, X_dml_card, p_ml_card, p_cb_card
  ),
  make_instrument_ipw_love(
    "Instrument balance — Kitagawa controls",
    Z, X_dml_kit, p_ml_kit, p_cb_kit
  ),
  ncol = 2,
  top = "Design-stage balance across college-proximity instrument groups"
)
```

### Positive-weight effective sample size

This table belongs to the design-stage instrument-propensity diagnostic above.
It uses conventional Kish ESS because the instrument-IPW weights are
nonnegative; it is not an ESS diagnostic for the signed outcome weights used
to represent the fitted LATE estimators.

```{r instrument-ipw-ess}
instrument_ipw_ess_rows <- function(design, Z, p_ml, p_cb) {
  rows <- lapply(list("MLE instrument IPW" = p_ml,
                      "CBPS instrument IPW" = p_cb), function(p) {
    w <- instrument_ipw(Z, p)
    do.call(rbind, lapply(c(1, 0), function(z) {
      wz <- w[Z == z]
      data.frame(
        Design = design,
        Propensity = "",
        `Z group` = paste0("Z = ", z),
        N = length(wz),
        `Kish ESS` = round(positive_weight_ess(wz), 1),
        `ESS / N` = round(positive_weight_ess(wz) / length(wz), 3),
        `Max normalized weight` = round(max(wz), 5),
        check.names = FALSE
      )
    }))
  })
  out <- do.call(rbind, rows)
  out$Propensity <- rep(names(rows), each = 2)
  rownames(out) <- NULL
  out
}

ipw_ess <- rbind(
  instrument_ipw_ess_rows("Card controls", Z, p_ml_card, p_cb_card),
  instrument_ipw_ess_rows("Kitagawa controls", Z, p_ml_kit, p_cb_kit)
)
ipw_ess_display <- ipw_ess %>% select(-Design)

ipw_ess_display %>%
  kbl(caption = "Design-stage instrument-IPW effective sample size",
      align = c("l","c","r","r","r","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12) %>%
  pack_rows(index = table(factor(ipw_ess$Design, levels = unique(ipw_ess$Design)))) %>%
  footnote(general = "Kish ESS uses (Σq)²/Σq² within each instrument group. It is appropriate here because q is nonnegative. ESS/N near one indicates limited concentration; smaller values indicate that the positive IPW pseudo-sample is supported by fewer observations.")
```

---

# 1. Reproduction — SUW Table 3

<div class="note-box">
**Layout.** SUW Table 3 has 8 columns: 2 treatments × 2 specifications × 2 outcome units, split here into two tabs by treatment. Within each tab the four columns are spec × unit. Normalized estimators (τ_cb_u, τ_ml_u, τ_ml_a10) are translation invariant — cents and dollars columns match; unnormalized ones (τ_ml_a, τ_ml_t, τ_ml_a0) do not.
</div>

```{r build-columns, cache=TRUE}
# One full column = 2SLS + all six kappa estimators with analytical SEs.
run_card_column <- function(Y, D, X_kappa, X_df) {
  list(
    tsls  = run_2sls(Y, D, Z, X_df, endog_name = "D"),
    kappa = kappa_analytic_se_all(Y, Z, D, X_kappa)
  )
}

# Treatment 1: somecol (cols 1-4)
col1 <- run_card_column(Y_cnt, D1, X_card_kappa, X_card_df)  # cents,   Card
col2 <- run_card_column(Y_dol, D1, X_card_kappa, X_card_df)  # dollars, Card
col3 <- run_card_column(Y_cnt, D1, X_kit_kappa,  X_kit_df)   # cents,   Kitagawa
col4 <- run_card_column(Y_dol, D1, X_kit_kappa,  X_kit_df)   # dollars, Kitagawa

# Treatment 2: educ16 (cols 5-8)
col5 <- run_card_column(Y_cnt, D2, X_card_kappa, X_card_df)  # cents,   Card
col6 <- run_card_column(Y_dol, D2, X_card_kappa, X_card_df)  # dollars, Card
col7 <- run_card_column(Y_cnt, D2, X_kit_kappa,  X_kit_df)   # cents,   Kitagawa
col8 <- run_card_column(Y_dol, D2, X_kit_kappa,  X_kit_df)   # dollars, Kitagawa

cat("All 8 columns computed.\n")
```

```{r table-builder, echo=FALSE}
cell_fmt <- function(est, se, digits = 3) {
  if (is.na(est)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0)
    return(sprintf("%.3f<br><span style='color:#888'>(NA)</span>", est))
  pval  <- 2 * pnorm(-abs(est / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s<br><span style='color:#888'>(%.3f)</span>", est, stars, se)
}

build_treatment_table <- function(cA_cnt, cB_dol, cC_cnt, cD_dol) {
  ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
  labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")

  rows <- list(data.frame(
    Estimator = "2SLS",
    `Card / cents`       = cell_fmt(cA_cnt$tsls$coef, cA_cnt$tsls$se),
    `Card / dollars`     = cell_fmt(cB_dol$tsls$coef, cB_dol$tsls$se),
    `Kitagawa / cents`   = cell_fmt(cC_cnt$tsls$coef, cC_cnt$tsls$se),
    `Kitagawa / dollars` = cell_fmt(cD_dol$tsls$coef, cD_dol$tsls$se),
    check.names = FALSE, stringsAsFactors = FALSE
  ))

  for (i in seq_along(ests)) {
    e <- ests[i]
    rows[[length(rows) + 1]] <- data.frame(
      Estimator = labels[i],
      `Card / cents`       = cell_fmt(cA_cnt$kappa$estimates[e], cA_cnt$kappa$se[e]),
      `Card / dollars`     = cell_fmt(cB_dol$kappa$estimates[e], cB_dol$kappa$se[e]),
      `Kitagawa / cents`   = cell_fmt(cC_cnt$kappa$estimates[e], cC_cnt$kappa$se[e]),
      `Kitagawa / dollars` = cell_fmt(cD_dol$kappa$estimates[e], cD_dol$kappa$se[e]),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }
  do.call(rbind, rows)
}
```

## Results {.tabset}

### Treatment: Some College (educ > 12)

```{r table-somecol}
build_treatment_table(col1, col2, col3, col4) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c"),
      caption = "SUW Table 3 — Treatment: somecol (educ > 12). Effect of some college on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",           1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant). Orange = unnormalized. Card spec: exper, expersq, black, south, smsa, smsa66, reg661–668. Kitagawa spec: black, south, smsa, smsa66, south66. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

### Treatment: College Completion (educ ≥ 16)

```{r table-educ16}
build_treatment_table(col5, col6, col7, col8) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c"),
      caption = "SUW Table 3 — Treatment: educ16 (educ ≥ 16). Effect of college completion on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",           1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant). Orange = unnormalized. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 2. Double ML — PLR-IV & Wald-AIPW

<div class="note-box">
**Forest specification.** For each empirical specification, the DML estimator
is fitted once with the default untuned honest forests
(`tune.parameters = "none"`) and once with all forest parameters eligible for
GRF's internal tuning (`tune.parameters = "all"`). The random seed is reset
before every fit so that the comparison isolates the tuning choice as closely
as possible. The tuned fits are retained as the headline specification and are
therefore the objects used in the subsequent learner comparison, diagnostics,
and translation-rerun notebook.
</div>

```{r dml-fit-all, cache=TRUE}
# Card/some-college specification.
set.seed(42); dml_card_d1_untuned <- dml_with_smoother(
  Y_dol, D1, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_card_d1_tuned <- dml_with_smoother(
  Y_dol, D1, X_dml_card, Z, n_cf_folds = 5, tune.parameters = "all")

# Kitagawa/college-completion specification.
set.seed(42); dml_kit_d2_untuned <- dml_with_smoother(
  Y_dol, D2, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_kit_d2_tuned <- dml_with_smoother(
  Y_dol, D2, X_dml_kit, Z, n_cf_folds = 5, tune.parameters = "all")

res_card_d1_untuned <- summary(dml_card_d1_untuned)
res_card_d1_tuned   <- summary(dml_card_d1_tuned)
res_kit_d2_untuned  <- summary(dml_kit_d2_untuned)
res_kit_d2_tuned    <- summary(dml_kit_d2_tuned)

# Extract PLR-IV and Wald-AIPW outcome weights for every fit.
extract_w <- function(dml_obj) {
  om <- get_outcome_weights(dml_obj)
  check_omega_rows(om)
  list(plriv    = as.vector(om$omega["PLR-IV",    ]),
       waldaipw = as.vector(om$omega["Wald-AIPW", ]))
}
w_card_d1_untuned <- extract_w(dml_card_d1_untuned)
w_card_d1_tuned   <- extract_w(dml_card_d1_tuned)
w_kit_d2_untuned  <- extract_w(dml_kit_d2_untuned)
w_kit_d2_tuned    <- extract_w(dml_kit_d2_tuned)

# Headline aliases: all downstream Card analyses use the tuned forest fits.
dml_card_d1 <- dml_card_d1_tuned
dml_kit_d2  <- dml_kit_d2_tuned
res_card_d1 <- res_card_d1_tuned
res_kit_d2  <- res_kit_d2_tuned
w_card_d1   <- w_card_d1_tuned
w_kit_d2    <- w_kit_d2_tuned

cat("All 4 specification-by-tuning DML fits complete.\n")
```

## Estimates {.tabset}

```{r dml-cell-display, echo=FALSE}
# PLR-IV + Wald-AIPW rows for one fit, with the algebraic identity check inline.
show_dml_cell <- function(res, wlist, Y, cap) {
  ids <- c(
    check_weight_identity(wlist$plriv,    Y, get_estimate(res, "PLR-IV")),
    check_weight_identity(wlist$waldaipw, Y, get_estimate(res, "Wald-AIPW"))
  )
  df <- as.data.frame(res[c("PLR-IV", "Wald-AIPW"), , drop = FALSE])
  df$Estimator  <- rownames(df)
  df$`ΣωY = τ̂` <- as.character(ids)
  df <- df[, c("Estimator", setdiff(names(df), c("Estimator","ΣωY = τ̂")), "ΣωY = τ̂")]
  df %>%
    kbl(caption = cap, digits = 4, row.names = FALSE,
        align = c("l", rep("r", ncol(df) - 2), "c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "ΣωY = τ̂: algebraic identity check (must be TRUE).")
}
```

### Card — somecol

```{r dml-card-d1-untuned}
show_dml_cell(res_card_d1_untuned, w_card_d1_untuned, Y_dol,
              "Card spec, somecol — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-card-d1-tuned}
show_dml_cell(res_card_d1_tuned, w_card_d1_tuned, Y_dol,
              "Card spec, somecol — tuned honest forest (tune.parameters = \"all\")")
```

### Kitagawa — educ16

```{r dml-kit-d2-untuned}
show_dml_cell(res_kit_d2_untuned, w_kit_d2_untuned, Y_dol,
              "Kitagawa spec, educ16 — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-kit-d2-tuned}
show_dml_cell(res_kit_d2_tuned, w_kit_d2_tuned, Y_dol,
              "Kitagawa spec, educ16 — tuned honest forest (tune.parameters = \"all\")")
```

## DML vs. Kappa — Point Estimates

```{r dml-summary-table}
# Kappa estimates for the two literature-aligned specifications.
kappa_card_d1 <- kappa_analytic_se_all(Y_dol, Z, D1, X_card_kappa)
kappa_kit_d2  <- kappa_analytic_se_all(Y_dol, Z, D2, X_kit_kappa)

cell <- function(coef, se) {
  if (is.na(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

dml_cell <- function(res, est) {
  coef_val <- get_estimate(res, est)
  se_val   <- if ("SE" %in% colnames(res)) as.numeric(res[est, "SE"])
              else if ("Std. Error" %in% colnames(res)) as.numeric(res[est, "Std. Error"])
              else NA_real_
  cell(coef_val, se_val)
}

dml_rows <- rbind(
  data.frame(Estimator = "PLR-IV (DML; tuned forest)",
             `somecol / Card`     = dml_cell(res_card_d1, "PLR-IV"),
             `educ16 / Kitagawa`  = dml_cell(res_kit_d2,  "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (DML; tuned forest)",
             `somecol / Card`     = dml_cell(res_card_d1, "Wald-AIPW"),
             `educ16 / Kitagawa`  = dml_cell(res_kit_d2,  "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE)
)

ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")
kappa_rows <- do.call(rbind, lapply(seq_along(ests), function(i) {
  e <- ests[i]
  data.frame(
    Estimator = labels[i],
    `somecol / Card`     = cell(kappa_card_d1$estimates[e], kappa_card_d1$se[e]),
    `educ16 / Kitagawa`  = cell(kappa_kit_d2$estimates[e],  kappa_kit_d2$se[e]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}))

rbind(dml_rows, kappa_rows) %>%
  kbl(align = c("l","c","c"),
      caption = "DML vs. Kappa — Point Estimates (SE), Y = log wages (dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Double ML", 1, 2) %>%
  pack_rows("Kappa — normalized", 3, 5) %>%
  pack_rows("Kappa — unnormalized", 6, 8) %>%
  row_spec(1:2, background = "#eef6ff") %>%
  row_spec(6:8, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE). Blue = DML. Orange = unnormalized kappa. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 3. Learner Comparison — Linear / Ranger / XGBoost

```{r ext-setup}
# DoubleMLData objects — one per cell. d and z must be numeric 0/1.
make_dml_data <- function(D, X_dml) {
  df <- as.data.frame(X_dml)
  df$y <- as.numeric(Y_dol)
  df$d <- as.numeric(D)
  df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X_dml))
}

dat_card_d1 <- make_dml_data(D1, X_dml_card)
dat_kit_d2  <- make_dml_data(D2, X_dml_kit)

cat("DoubleML data objects built for both specifications.\n")
```

```{r ext-fit-helpers, echo=FALSE}
# Three learner spec lists, reused across cells.
ml_lm  <- function() list(g = lrn("regr.lm"),
                          m = lrn("classif.log_reg"),
                          r = lrn("classif.log_reg"))
ml_rf  <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                          m = lrn("classif.ranger", keep.inbag = TRUE),
                          r = lrn("classif.ranger", keep.inbag = TRUE))
ml_xgb <- function() list(
  g = lrn("regr.xgboost",    alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
  m = lrn("classif.xgboost", predict_type = "prob"),
  r = lrn("classif.xgboost", predict_type = "prob"))

# Fit one IIVM model and return list(obj, w, ok). get_dml_se now lives in functions_all_v4.R.
fit_iivm <- function(dml_data, learners, Y) {
  set.seed(42)
  obj <- DoubleMLIIVM$new(data = dml_data,
                          ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                          n_folds = 5, score = "LATE")
  obj$fit(store_models = TRUE, store_predictions = TRUE)
  om  <- get_outcome_weights(object = obj, dml_data = dml_data)
  w   <- as.vector(om$omega)
  ok  <- check_doubleml_identity(om, Y, obj$coef)
  list(obj = obj, w = w, ok = ok)
}

# Compact two-specification estimate table for one learner.
learner_estimate_table <- function(f_card_d1, f_kit_d2, cap) {
  data.frame(
    Cell      = c("Card / some college","Kitagawa / college completion"),
    Estimate  = round(c(get_dml_coef(f_card_d1$obj), get_dml_coef(f_kit_d2$obj)), 4),
    SE        = round(c(get_dml_se(f_card_d1$obj), get_dml_se(f_kit_d2$obj)), 4),
    `ΣωY = τ̂` = c(f_card_d1$ok, f_kit_d2$ok),
    check.names = FALSE
  ) %>%
    kbl(caption = cap, align = c("l","r","r","c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "ΣωY = τ̂: algebraic identity check (must be TRUE).")
}
```

## Estimates {.tabset}

### Linear + Logistic

```{r ext-fit-lm, cache=TRUE}
lm_card_d1 <- fit_iivm(dat_card_d1, ml_lm(), Y_dol)
lm_kit_d2  <- fit_iivm(dat_kit_d2,  ml_lm(), Y_dol)

learner_estimate_table(lm_card_d1, lm_kit_d2,
                       "Wald-AIPW — Linear + Logistic")
```

### Ranger

```{r ext-fit-rf, cache=TRUE}
rf_card_d1 <- fit_iivm(dat_card_d1, ml_rf(), Y_dol)
rf_kit_d2  <- fit_iivm(dat_kit_d2,  ml_rf(), Y_dol)

learner_estimate_table(rf_card_d1, rf_kit_d2,
                       "Wald-AIPW — Ranger")
```

### XGBoost

```{r ext-xgb-import}
wald_xgb <- readRDS("card_xgb_tuning_export.rds")
pliv_xgb <- readRDS("card_xgb_pliv_tuning_export.rds")

get_xgb_row <- function(export, cell_name, tuned_label = "tuned") {
  tab <- export$estimate_comparison
  hit <- grepl(tuned_label, tab$Learner, ignore.case = TRUE) &
    !grepl("untuned", tab$Learner, ignore.case = TRUE) & tab$Cell == cell_name
  if (sum(hit) != 1L) stop(sprintf("Expected one tuned XGBoost row for '%s'.", cell_name))
  # The Wald export also contains "Propensity boundary share", whereas the
  # PLR-IV export does not. Restrict both to their common display columns before
  # base::rbind(); otherwise the two data frames have different widths.
  out <- tab[hit, c("Cell", "Learner", "Estimate (SE)"), drop = FALSE]
  out$`Weight identity` <- export$identity_checks[[cell_name]]$tuned
  out
}
get_xgb_weight <- function(export, cell_name) export$omega_vectors[[cell_name]]$tuned
get_xgb_identity <- function(export, cell_name) export$identity_checks[[cell_name]]$tuned

xgb_card_d1 <- list(
  w = get_xgb_weight(wald_xgb, "Card / some college"),
  ok = get_xgb_identity(wald_xgb, "Card / some college"))
xgb_kit_d2 <- list(
  w = get_xgb_weight(wald_xgb, "Kitagawa / college completion"),
  ok = get_xgb_identity(wald_xgb, "Kitagawa / college completion"))
xgb_pliv_card_d1 <- list(
  w = get_xgb_weight(pliv_xgb, "Card / some college"),
  ok = get_xgb_identity(pliv_xgb, "Card / some college"))
xgb_pliv_kit_d2 <- list(
  w = get_xgb_weight(pliv_xgb, "Kitagawa / college completion"),
  ok = get_xgb_identity(pliv_xgb, "Kitagawa / college completion"))

rbind(
  transform(get_xgb_row(wald_xgb, "Card / some college"), Estimator = "Wald-AIPW"),
  transform(get_xgb_row(wald_xgb, "Kitagawa / college completion"), Estimator = "Wald-AIPW"),
  transform(get_xgb_row(pliv_xgb, "Card / some college"), Estimator = "PLR-IV"),
  transform(get_xgb_row(pliv_xgb, "Kitagawa / college completion"), Estimator = "PLR-IV")
) %>%
  select(Estimator, everything()) %>%
  kbl(caption = "Imported tuned-on-folds XGBoost estimates") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE) %>%
  footnote(general = "Weight identity is TRUE only when the exported outcome weights reproduce the corresponding XGBoost coefficient exactly. FALSE does not prevent displaying the fitted estimate, but weight-based diagnostics must be interpreted as an approximate/non-affine representation.")
```

## Comparison Table

```{r ext-comparison}
ext_cell <- function(coef, se) {
  if (is.na(coef) || !is.finite(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

learner_row <- function(label, f_card_d1, f_kit_d2) {
  data.frame(
    Estimator = label,
    `somecol / Card`     = ext_cell(get_dml_coef(f_card_d1$obj), get_dml_se(f_card_d1$obj)),
    `educ16 / Kitagawa`  = ext_cell(get_dml_coef(f_kit_d2$obj),  get_dml_se(f_kit_d2$obj)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

kappa_ref_row <- function(label, key) {
  data.frame(
    Estimator = label,
    `somecol / Card`     = ext_cell(kappa_card_d1$estimates[key], kappa_card_d1$se[key]),
    `educ16 / Kitagawa`  = ext_cell(kappa_kit_d2$estimates[key],  kappa_kit_d2$se[key]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

rbind(
  learner_row("Wald-AIPW (linear+logit)", lm_card_d1, lm_kit_d2),
  learner_row("Wald-AIPW (ranger)",       rf_card_d1, rf_kit_d2),
  kappa_ref_row("τ_cb_u (kappa, ref)",  "tau_cb_u"),
  kappa_ref_row("τ_ml_u (kappa, ref)",  "tau_ml_u")
) %>%
  kbl(align = c("l","c","c"),
      caption = "Wald-AIPW Learner Comparison — DoubleML vs. Kappa (Y = log wages, dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("DoubleML learners", 1, 2) %>%
  pack_rows("Kappa reference",   3, 4) %>%
  row_spec(1:2, background = "#eef6ff") %>%
  footnote(general = "Each cell: estimate (SE). Blue = DoubleML. All score='LATE', n_folds=5. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 4. Weight Diagnostics

<div class="note-box">
**Reading the tables.** `Sum_w ≈ 0` checks translation invariance, while
`Mass T/C ≈ 1 / 1` checks the normalization of the treated and sign-oriented
control components. Following Chattopadhyay and Zubizarreta (2023, Section
5.4), the modified ESS `(Σ|ωᵢ|)²/Σωᵢ²` is reported separately for treated and
control observations; the overall value is retained only as a supplementary
concentration summary. `ESS mod/N T/C` divides each arm-specific value by that
arm's observed size, making concentration comparable across unequal groups.
`Opposite-sign mass T/C` is the percentage of absolute
weight mass carried by weights opposite to the conventional treated/control
orientation after replacing control weights by `−ω`. It detects algebraic
cancellation that ESS alone cannot reveal; it is not by itself evidence that
the IV estimator is invalid.

Conventional Kish ESS is not reported for these signed outcome weights because
it presumes a nonnegative normalized weighted sample. It appears only in the
separate positive instrument-IPW diagnostic in Section 1.
</div>

```{r diag-builders, echo=FALSE}
# Kappa weight bundles for the two specification cells.
kw_card_d1 <- kappa_weights_bundle(Z, D1, X_card_kappa)
kw_kit_d2  <- kappa_weights_bundle(Z, D2, X_kit_kappa)

# Exact coefficient decompositions enter the outcome-weight diagnostics.
# Failed XGBoost identities remain visible in the estimate/status section.
exact_xgb_w <- function(fit) {
  if (isTRUE(fit$ok) && !is.null(fit$w)) fit$w else NULL
}
safe_outcome_diag <- function(w, D, label) {
  if (is.null(w) || !length(w) || all(is.na(w))) NULL else outcome_weight_diag(w, D, label)
}

DIAG_ALIGN <- c("l","r","c","r","c","c","c","r")

dml_diag_cell <- function(wlist, kw, D) {
  rbind(
    outcome_weight_diag(wlist$plriv,    D, "PLR-IV     (DML)"),
    outcome_weight_diag(wlist$waldaipw, D, "Wald-AIPW  (DML)"),
    outcome_weight_diag(kw$tau_cb_u,    D, "τ_cb_u     (kappa)"),
    outcome_weight_diag(kw$tau_ml_u,    D, "τ_ml_u     (kappa)"),
    outcome_weight_diag(kw$tau_ml_a10,  D, "τ_ml_a10   (kappa)"),
    outcome_weight_diag(kw$tau_ml_a,    D, "τ_ml_a     (kappa)"),
    outcome_weight_diag(kw$tau_ml_t,    D, "τ_ml_t     (kappa)"),
    outcome_weight_diag(kw$tau_ml_a0,   D, "τ_ml_a0    (kappa)")
  )
}
style_dml_diag <- function(df, cap) {
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(1:2, background = "#eef6ff") %>%
    footnote(general = "Blue = DML. T/C entries report treated / sign-oriented control values. Arm-specific modified ESS is the headline concentration diagnostic; overall modified ESS is supplementary.")
}

ext_diag_block <- function(D, X_kappa, w_lm, w_rf, w_pliv_xgb, w_wald_xgb) {
  kw <- kappa_weights_bundle(Z, D, X_kappa)
  do.call(rbind, Filter(Negate(is.null), list(
    safe_outcome_diag(w_lm,          D, "Wald-AIPW (linear+logit)"),
    safe_outcome_diag(w_rf,          D, "Wald-AIPW (ranger)"),
    safe_outcome_diag(w_pliv_xgb,    D, "PLR-IV (XGBoost tuned)"),
    safe_outcome_diag(w_wald_xgb,    D, "Wald-AIPW (XGBoost tuned)"),
    outcome_weight_diag(kw$tau_cb_u,   D, "τ_cb_u    (kappa, CBPS)"),
    outcome_weight_diag(kw$tau_ml_u,   D, "τ_ml_u    (kappa, MLE)"),
    outcome_weight_diag(kw$tau_ml_a10, D, "τ_ml_a10  (kappa, MLE)"),
    outcome_weight_diag(kw$tau_ml_a,   D, "τ_ml_a    (kappa, unnorm)"),
    outcome_weight_diag(kw$tau_ml_t,   D, "τ_ml_t    (kappa, unnorm)"),
    outcome_weight_diag(kw$tau_ml_a0,  D, "τ_ml_a0   (kappa, unnorm)")
  )))
}
style_ext_diag <- function(df, cap) {
  n_learn <- sum(grepl("^(PLR-IV|Wald-AIPW)", df$Estimator))
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(seq_len(n_learn), background = "#eef6ff") %>%
    footnote(general = "Blue = DoubleML learners. T/C entries report treated / sign-oriented control values. Failed XGBoost weight identities must be read as sensitivity representations, not exact coefficient decompositions.")
}
```

## DML vs. Kappa {.tabset}

### Card — somecol

```{r dml-diag-card-d1}
style_dml_diag(dml_diag_cell(w_card_d1, kw_card_d1, D1), "Weight Diagnostics — DML vs. Kappa | Card, somecol")
```

### Kitagawa — educ16

```{r dml-diag-kit-d2}
style_dml_diag(dml_diag_cell(w_kit_d2, kw_kit_d2, D2), "Weight Diagnostics — DML vs. Kappa | Kitagawa, educ16")
```

## Learners vs. Kappa {.tabset}

### Card — somecol

```{r ext-diag-card-d1}
style_ext_diag(ext_diag_block(D1, X_card_kappa, lm_card_d1$w, rf_card_d1$w,
                              exact_xgb_w(xgb_pliv_card_d1),
                              exact_xgb_w(xgb_card_d1)),
               "Weight Diagnostics — Learners | Card, somecol")
```

### Kitagawa — educ16

```{r ext-diag-kit-d2}
style_ext_diag(ext_diag_block(D2, X_kit_kappa, lm_kit_d2$w, rf_kit_d2$w,
                              exact_xgb_w(xgb_pliv_kit_d2),
                              exact_xgb_w(xgb_kit_d2)),
               "Weight Diagnostics — Learners | Kitagawa, educ16")
```

## Weight-Statistics Bridge — RQ1 {.tabset}

<div class="note-box">
**What this is.** For every estimator, `weight_bridge()` reports the outcome
weights Overall / Treated / Control (control sign-flipped, matching Knaus's
convention). The detailed bridge retains the statistics produced by Knaus's
`summary()` method and adds modified ESS. Its `Neg. share` is the unweighted
share of negative observations on Knaus's zero-to-one numerical scale; it
should not be confused with the oriented absolute-mass diagnostic in the
headline tables. Conventional Kish ESS is omitted for signed outcome weights.
</div>

```{r bridge-helpers, echo=FALSE}
# One named list of weight vectors per cell (DML + learners + key kappa).
cell_weight_list <- function(wdml, lm_w, rf_w, pliv_xgb_w, wald_xgb_w, kw) {
  lst <- list(
    "PLR-IV (DML)"     = wdml$plriv,
    "Wald-AIPW (DML)"  = wdml$waldaipw,
    "Linear+Logit"     = lm_w,
    "Ranger"           = rf_w,
    "PLR-IV XGBoost tuned"    = pliv_xgb_w,
    "Wald-AIPW XGBoost tuned" = wald_xgb_w,
    "τ_cb_u (kappa)"   = kw$tau_cb_u,
    "τ_ml_u (kappa)"   = kw$tau_ml_u,
    "τ_ml_t (kappa)"   = kw$tau_ml_t
  )
  Filter(function(w) !is.null(w) && length(w) && !all(is.na(w)), lst)
}

bridge_cells <- list(
  "Card · somecol" = list(D = D1, w = cell_weight_list(
    w_card_d1, lm_card_d1$w, rf_card_d1$w,
    exact_xgb_w(xgb_pliv_card_d1),
    exact_xgb_w(xgb_card_d1), kw_card_d1)),
  "Kitagawa · educ16" = list(D = D2, w = cell_weight_list(
    w_kit_d2, lm_kit_d2$w, rf_kit_d2$w,
    exact_xgb_w(xgb_pliv_kit_d2),
    exact_xgb_w(xgb_kit_d2), kw_kit_d2))
)

# Full bridge table for one cell: estimators as row-groups, scopes as rows,
# columns split into "Knaus summary()" vs. "ESS" via a spanning header.
bridge_table <- function(wlist, D, cap) {
  ests <- names(wlist)
  df   <- do.call(rbind, lapply(ests, function(nm) weight_bridge(wlist[[nm]], D, nm)))
  disp <- df[, c("Scope","Min","Max","Neg_share","Sum_top10","Sum_w",
                 "Sum_abs_w","Max_abs_w","ESS_mod")]

  kk <- disp %>%
    kbl(caption = cap, row.names = FALSE,
        col.names = c("Scope","Min","Max","Neg. share","ΣTop10%","Σω","Σ|ω|",
                      "Max |ω|","ESS mod"),
        align = c("l", rep("r", 8))) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 12) %>%
    add_header_above(c(" " = 1, "Knaus summary() statistics" = 6,
                       "Concentration" = 1,
                       "Effective sample size" = 1)) %>%
    column_spec(9, bold = TRUE, background = "#eef6ff")
  for (i in seq_along(ests)) kk <- pack_rows(kk, ests[i], (i - 1) * 3 + 1, i * 3)
  kk %>% footnote(general = "Neg. share is Knaus's observation-count statistic in [0,1]. Control rows use the sign-flipped −ω convention. ESS mod is an absolute-weight concentration diagnostic; the headline tables report it separately for treated and control observations.")
}
```

### Card · somecol

```{r bridge-card-d1}
bridge_table(bridge_cells[["Card · somecol"]]$w, bridge_cells[["Card · somecol"]]$D,
             "Weight-statistics bridge — Card, somecol")
```

### Kitagawa · educ16

```{r bridge-kit-d2}
bridge_table(bridge_cells[["Kitagawa · educ16"]]$w, bridge_cells[["Kitagawa · educ16"]]$D,
             "Weight-statistics bridge — Kitagawa, educ16")
```

---

# 5. Love Plots

<div class="note-box">
**Balance scale.** Each panel reports absolute mean differences only (no KS
statistic). Continuous covariates are standardized by the pooled unadjusted
standard deviation. Starred binary covariates are left on their raw 0/1 scale.
Because the adjusted comparison uses sign-oriented outcome weights, a raw binary
difference is a signed weighted-moment difference and need not compare two
literal proportions when opposite-oriented weights remain.
</div>

```{r love-exact-helper, echo=FALSE}
maybe_love <- function(title, fit, D, X) {
  if (!isTRUE(fit$ok) || is.null(fit$w)) NULL else make_love(title, fit$w, D, X)
}
```

## DML vs. Kappa {.tabset}

### Card — somecol (all 8 estimators)

```{r love-card-d1, fig.height=9, fig.width=13}
grid.arrange(grobs = list(
  make_love("PLR-IV (DML)",     w_card_d1$plriv,       D1, X_dml_card),
  make_love("Wald-AIPW (DML)",  w_card_d1$waldaipw,    D1, X_dml_card),
  make_love("τ_cb_u (CBPS)",    kw_card_d1$tau_cb_u,   D1, X_dml_card),
  make_love("τ_ml_u",           kw_card_d1$tau_ml_u,   D1, X_dml_card),
  make_love("τ_ml_a10",         kw_card_d1$tau_ml_a10, D1, X_dml_card),
  make_love("τ_ml_a (unnorm)",  kw_card_d1$tau_ml_a,   D1, X_dml_card),
  make_love("τ_ml_t (unnorm)",  kw_card_d1$tau_ml_t,   D1, X_dml_card),
  make_love("τ_ml_a0 (unnorm)", kw_card_d1$tau_ml_a0,  D1, X_dml_card)
), nrow = 2, top = "Love Plots — Card spec, somecol: DML vs. Kappa")
```

### Kitagawa — college completion (all 8 estimators)

```{r love-kit-d2, fig.height=9, fig.width=13}
grid.arrange(grobs = list(
  make_love("PLR-IV (DML)",     w_kit_d2$plriv,       D2, X_dml_kit),
  make_love("Wald-AIPW (DML)",  w_kit_d2$waldaipw,    D2, X_dml_kit),
  make_love("τ_cb_u (CBPS)",    kw_kit_d2$tau_cb_u,   D2, X_dml_kit),
  make_love("τ_ml_u",           kw_kit_d2$tau_ml_u,   D2, X_dml_kit),
  make_love("τ_ml_a10",         kw_kit_d2$tau_ml_a10, D2, X_dml_kit),
  make_love("τ_ml_a (unnorm)",  kw_kit_d2$tau_ml_a,   D2, X_dml_kit),
  make_love("τ_ml_t (unnorm)",  kw_kit_d2$tau_ml_t,   D2, X_dml_kit),
  make_love("τ_ml_a0 (unnorm)", kw_kit_d2$tau_ml_a0,  D2, X_dml_kit)
), nrow = 2, top = "Love Plots — Kitagawa spec, college completion: DML vs. Kappa")
```

## Learners vs. Kappa {.tabset}

### Card — somecol (learners + kappa)

```{r ext-love-card-d1, fig.height=5.5, fig.width=14}
grid.arrange(grobs = Filter(Negate(is.null), list(
  make_love("Linear+Logit\nCard, somecol", lm_card_d1$w,        D1, X_dml_card),
  make_love("Ranger\nCard, somecol",       rf_card_d1$w,        D1, X_dml_card),
  maybe_love("PLR-IV XGBoost tuned\nCard, somecol", xgb_pliv_card_d1, D1, X_dml_card),
  maybe_love("Wald-AIPW XGBoost tuned\nCard, somecol", xgb_card_d1, D1, X_dml_card),
  make_love("τ_cb_u\nCard, somecol",       kw_card_d1$tau_cb_u, D1, X_dml_card),
  make_love("τ_ml_u\nCard, somecol",       kw_card_d1$tau_ml_u, D1, X_dml_card)
)), nrow = 2, top = "Love Plots — DoubleML learners vs. kappa | Card, somecol")
```

### Kitagawa — college completion (learners + kappa)

```{r ext-love-kit-d2, fig.height=5.5, fig.width=14}
grid.arrange(grobs = Filter(Negate(is.null), list(
  make_love("Linear+Logit\nKitagawa, completion", lm_kit_d2$w,        D2, X_dml_kit),
  make_love("Ranger\nKitagawa, completion",       rf_kit_d2$w,        D2, X_dml_kit),
  maybe_love("PLR-IV XGBoost tuned\nKitagawa, completion", xgb_pliv_kit_d2, D2, X_dml_kit),
  maybe_love("Wald-AIPW XGBoost tuned\nKitagawa, completion", xgb_kit_d2, D2, X_dml_kit),
  make_love("τ_cb_u\nKitagawa, completion",       kw_kit_d2$tau_cb_u, D2, X_dml_kit),
  make_love("τ_ml_u\nKitagawa, completion",       kw_kit_d2$tau_ml_u, D2, X_dml_kit)
)), nrow = 2, top = "Love Plots — DoubleML learners vs. kappa | Kitagawa, college completion")
```
