---
title: "Card (1995) — XGBoost Nuisance Tuning for PLR-IV"
subtitle: "Nested nuisance tuning by predictive loss, tuned inside each outer cross-fitting fold"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 5.5)
```

```{r packages}
library(knitr)
library(kableExtra)
library(dplyr)
library(haven)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

source("functions_all.R")

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

SEED        = 42
N_CF        = 5
TUNE_EVALS  = 15
INNER_FOLDS = 3
EXTRACT_WEIGHTS = TRUE
```

<div class="note-box">
**Purpose.** This notebook mirrors `child_xgb_tuning.Rmd`, but for
`DoubleMLPLIV` instead of `DoubleMLIIVM`. The main analysis imports the saved
RDS object from this file; it does not refit or retune XGBoost inside
`card_test.Rmd`.

**Untuned baseline.** The untuned XGBoost learners intentionally keep the
same base settings as requested:
`alpha = 0`, `subsample = 1`, `max_delta_step = 0`, `base_score = 0`.

**Tuned row.** `tune_on_folds = TRUE` retunes inside each outer cross-fitting
fold. This avoids reusing a global tuning choice for observations that later
serve as held-out fold predictions.

**Weight extraction.** `EXTRACT_WEIGHTS` is currently set to `TRUE` because
the main analysis uses the exported PLR-IV XGBoost weight vectors. This is
the slow part of the notebook. If only the estimate and tuning tables are
needed, set `EXTRACT_WEIGHTS = FALSE` before knitting.
</div>

<div class="warn-box">
**Runtime warning.** PLR-IV uses three regression nuisance learners
(`ml_l`, `ml_m`, `ml_r`) and retunes them separately inside each of the five
outer folds when `tune_on_folds = TRUE`. Keep `TUNE_EVALS = 15` unless there
is a specific reason to spend more runtime.
</div>

---

# Setup

```{r data}
card <- read_dta("card.dta")
card <- card[!is.na(card$wage) & !is.na(card$educ), ]
if (nrow(card) != 3010L) stop("Expected the Card analysis sample to contain 3,010 observations.")

card$lwage   <- log(card$wage / 100)
card$somecol <- as.integer(card$educ >= 13)
card$educ16  <- as.integer(card$educ >= 16)
card$Z_inst  <- as.integer(card$nearc4)
if (!"expersq" %in% names(card)) card$expersq <- card$exper^2

X_card <- make_X(~ 0 + exper + expersq + black + south + smsa + smsa66 +
                   reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
                 data = card)
X_kit  <- make_X(~ 0 + black + south + smsa + smsa66 + south66, data = card)

cat(sprintf("N=%d | Card/some college: %d covariates | Kitagawa/completion: %d covariates\n",
            nrow(card), ncol(X_card), ncol(X_kit)))
```

```{r learners-and-tuning-setup}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X))
}

make_xgb_pliv_learners <- function() {
  base <- list(alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0)
  list(
    l = do.call(lrn, c(list("regr.xgboost"), base)),
    m = do.call(lrn, c(list("regr.xgboost"), base)),
    r = do.call(lrn, c(list("regr.xgboost"), base))
  )
}

xgb_ps <- function() {
  ps(nrounds = p_int(50, 400), max_depth = p_int(2, 5),
     eta = p_dbl(0.02, 0.20), min_child_weight = p_int(1, 8))
}

tune_param_set <- function() list(ml_l = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())

tune_settings <- function() {
  list(
    n_folds_tune = INNER_FOLDS,
    rsmp_tune    = rsmp("cv", folds = INNER_FOLDS),
    measure      = list(ml_l = "regr.rmse", ml_m = "regr.rmse", ml_r = "regr.rmse"),
    terminator   = trm("evals", n_evals = TUNE_EVALS),
    algorithm    = tnr("random_search")
  )
}

get_prediction_vector <- function(obj, learner_name) {
  pred <- tryCatch(obj$predictions[[learner_name]], error = function(e) NULL)
  if (is.null(pred)) return(numeric(0))
  as.numeric(drop(pred))
}

rmse_loss <- function(truth, pred) {
  if (length(pred) != length(truth) || length(pred) == 0) return(NA_real_)
  truth <- as.numeric(truth); pred <- as.numeric(pred)
  keep <- is.finite(truth) & is.finite(pred)
  if (!any(keep)) return(NA_real_)
  sqrt(mean((truth[keep] - pred[keep])^2))
}

calc_nuisance_losses <- function(r, cell) {
  if (is.null(r)) return(c(outcome_l_rmse = NA_real_, instrument_m_rmse = NA_real_, treatment_r_rmse = NA_real_))
  l_hat <- get_prediction_vector(r$obj, "ml_l")
  m_hat <- get_prediction_vector(r$obj, "ml_m")
  r_hat <- get_prediction_vector(r$obj, "ml_r")
  c(outcome_l_rmse    = rmse_loss(cell$Y, l_hat),
    instrument_m_rmse = rmse_loss(cell$Z, m_hat),
    treatment_r_rmse  = rmse_loss(cell$D, r_hat))
}

omega_vec <- function(om) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && "PLR-IV" %in% rownames(w)) return(as.vector(w["PLR-IV", ]))
  as.vector(w[1, ])
}

extract_xgb_w <- function(r, cell) {
  if (is.null(r) || is.null(r$obj)) {
    stop("Cannot extract weights because the fitted DoubleML object is NULL.")
  }
  dml_data <- make_dml_data(cell$Y, cell$D, cell$Z, cell$X)
  om <- get_outcome_weights(object = r$obj, dml_data = dml_data)
  omega_vec(om)
}

safe_extract_xgb_w <- function(r, cell) {
  tryCatch(extract_xgb_w(r, cell), error = function(e) NULL)
}

fit_cell <- function(dml_data, learners, label, tune = FALSE, tune_on_folds = FALSE) {
  tryCatch({
    set.seed(SEED)
    obj <- DoubleMLPLIV$new(data = dml_data,
                            ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
                            n_folds = N_CF)
    if (tune) {
      obj$tune(param_set = tune_param_set(), tune_settings = tune_settings(),
               tune_on_folds = tune_on_folds)
    }
    obj$fit(store_models = TRUE, store_predictions = TRUE)

    list(label = label, tuned = tune,
         coef = get_dml_coef(obj), se = get_dml_se(obj),
         tuning_result = tryCatch(obj$tuning_result, error = function(e) NULL),
         params        = tryCatch(obj$params,        error = function(e) NULL),
         obj = obj)
  }, error = function(e) {
    message(sprintf("Fit failed for %s: %s", label, conditionMessage(e)))
    NULL
  })
}

cells <- list(
  "Card / some college"           = list(Y = card$lwage, D = card$somecol, Z = card$Z_inst, X = X_card),
  "Kitagawa / college completion" = list(Y = card$lwage, D = card$educ16, Z = card$Z_inst, X = X_kit)
)
```

---

# Fit: XGBoost PLR-IV untuned vs. tuned

```{r fit-all, cache=TRUE}
results <- list()
for (cell_name in names(cells)) {
  cell <- cells[[cell_name]]
  dml_data <- make_dml_data(cell$Y, cell$D, cell$Z, cell$X)

  results[[cell_name]] <- list(
    untuned = fit_cell(dml_data, make_xgb_pliv_learners(),
                       paste(cell_name, "pliv-xgb-untuned"),
                       tune = FALSE),
    tuned   = fit_cell(dml_data, make_xgb_pliv_learners(),
                       paste(cell_name, "pliv-xgb-tuned"),
                       tune = TRUE, tune_on_folds = TRUE)
  )
  cat(sprintf("Finished: %s\n", cell_name))
}
```

## 1. Estimate comparison

```{r comparison-table}
fmt_est <- function(r) if (is.null(r)) "\u2014" else sprintf("%.3f (%.3f)", r$coef, r$se)

rows <- do.call(rbind, lapply(names(results), function(cell_name) {
  R <- results[[cell_name]]
  data.frame(Cell = cell_name, Learner = c("PLR-IV XGB untuned", "PLR-IV XGB tuned (tune_on_folds=TRUE)"),
             `Estimate (SE)` = c(fmt_est(R$untuned), fmt_est(R$tuned)),
             check.names = FALSE, stringsAsFactors = FALSE)
}))

rows %>%
  kbl(align = c("l","l","c"), caption = "Card PLR-IV: XGBoost untuned vs. tuned inside each outer fold") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(rows$Learner == "PLR-IV XGB tuned (tune_on_folds=TRUE)"), background = "#eef6ff") %>%
  footnote(general = "Untuned uses the fixed base XGBoost settings listed at the top of this notebook. Tuned selects nrounds, max_depth, eta, and min_child_weight by 3-fold random-search CV inside each outer cross-fitting fold.")
```

## 2. Nuisance-prediction losses

```{r nuisance-loss-table}
loss_rows <- do.call(rbind, lapply(names(results), function(cell_name) {
  R <- results[[cell_name]]; cell <- cells[[cell_name]]
  ul <- calc_nuisance_losses(R$untuned, cell); tl <- calc_nuisance_losses(R$tuned, cell)
  data.frame(Cell = cell_name, Learner = c("PLR-IV XGB untuned", "PLR-IV XGB tuned"),
             `Outcome l RMSE` = c(ul["outcome_l_rmse"], tl["outcome_l_rmse"]),
             `Instrument m RMSE` = c(ul["instrument_m_rmse"], tl["instrument_m_rmse"]),
             `Treatment r RMSE` = c(ul["treatment_r_rmse"], tl["treatment_r_rmse"]),
             check.names = FALSE, row.names = NULL)
}))

loss_rows %>%
  kbl(digits = 4, caption = "PLR-IV nuisance-prediction loss: untuned vs. tuned (lower is better)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
  collapse_rows(columns = 1, valign = "middle") %>%
  row_spec(which(loss_rows$Learner == "PLR-IV XGB tuned"), background = "#eef6ff")
```

## 3. Selected hyperparameters

```{r hyperparameter-summary}
extract_tuned_params <- function(r, cell_name) {
  if (is.null(r) || !r$tuned) return(data.frame())
  src <- if (!is.null(r$params)) r$params else r$tuning_result
  if (is.null(src)) return(data.frame())

  u <- unlist(src, recursive = TRUE)
  u <- u[grepl("nrounds|max_depth|eta|min_child_weight", names(u))]
  if (length(u) == 0) return(data.frame())

  hp <- data.frame(Cell = cell_name, Raw_name = names(u), Value = as.numeric(u), stringsAsFactors = FALSE)
  hp$Nuisance <- ifelse(grepl("ml_l", hp$Raw_name), "Outcome l",
                  ifelse(grepl("ml_m", hp$Raw_name), "Instrument m",
                  ifelse(grepl("ml_r", hp$Raw_name), "Treatment r", "Unknown")))
  hp$Parameter <- sub(".*(nrounds|max_depth|eta|min_child_weight).*", "\\1", hp$Raw_name)
  hp
}

hp_raw <- do.call(rbind, lapply(names(results), function(cn) extract_tuned_params(results[[cn]]$tuned, cn)))

if (nrow(hp_raw) > 0) {
  hp_raw <- hp_raw[hp_raw$Nuisance != "Unknown", ]
  hp_summary <- hp_raw %>%
    group_by(Cell, Nuisance, Parameter) %>%
    summarise(Median_across_folds = median(Value), N_values_seen = n(), .groups = "drop") %>%
    arrange(Cell, Nuisance, Parameter)

  hp_summary %>%
    kbl(digits = 3, caption = "Selected PLR-IV XGBoost hyperparameters: median across fold-level choices") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12.5) %>%
    collapse_rows(columns = c(1,2), valign = "middle") %>%
    footnote(general = "With tune_on_folds=TRUE, these are descriptive summaries of fold-specific choices, not a single reusable recipe.")
} else {
  cat("No hyperparameters could be extracted automatically.\n")
}
```

## 4. Outcome-weight diagnostics

```{r xgb-pliv-weight-table}
if (isTRUE(EXTRACT_WEIGHTS)) {
  one_xgb_diag_row <- function(cell_name, learner_name, r, cell) {
    if (is.null(r) || is.null(r$obj)) {
      return(data.frame(Cell = cell_name, Learner = learner_name,
                        Estimate = NA_real_, ESS = NA_real_, ESS_mod = NA_real_,
                        `Sigma omega` = NA_real_, `Sigma abs omega` = NA_real_,
                        `Treated mass` = NA_real_,
                        `Control mass (-sum control weights)` = NA_real_,
                        `max abs omega` = NA_real_,
                        `Identity ok` = NA,
                        `Weight status` = "fit failed before weight extraction",
                        check.names = FALSE))
    }

    w <- tryCatch(extract_xgb_w(r, cell), error = function(e) {
      attr(e, "weight_error") <- TRUE
      e
    })
    if (inherits(w, "error")) {
      return(data.frame(Cell = cell_name, Learner = learner_name,
                        Estimate = as.numeric(r$coef), ESS = NA_real_, ESS_mod = NA_real_,
                        `Sigma omega` = NA_real_, `Sigma abs omega` = NA_real_,
                        `Treated mass` = NA_real_,
                        `Control mass (-sum control weights)` = NA_real_,
                        `max abs omega` = NA_real_,
                        `Identity ok` = NA,
                        `Weight status` = paste("weight extraction failed:", conditionMessage(w)),
                        check.names = FALSE))
    }

    wb <- weight_bridge(w, cell$D, learner_name)
    ov <- wb[wb$Scope == "Overall", ]; tr <- wb[wb$Scope == "Treated", ]; co <- wb[wb$Scope == "Control", ]

    identity_ok <- tryCatch(
      check_weight_identity(w, cell$Y, r$coef),
      error = function(e) NA
    )

    data.frame(Cell = cell_name, Learner = learner_name, Estimate = as.numeric(r$coef),
               ESS = ov$ESS_kish, ESS_mod = ov$ESS_mod,
               `Sigma omega` = ov$Sum_w, `Sigma abs omega` = ov$Sum_abs_w,
               `Treated mass` = tr$Sum_w, `Control mass (-sum control weights)` = co$Sum_w,
               `max abs omega` = max(abs(w)),
               `Identity ok` = identity_ok,
               `Weight status` = "ok",
               check.names = FALSE)
  }

  xgb_pliv_weight_table <- do.call(rbind, lapply(names(cells), function(cell_name) {
    cell <- cells[[cell_name]]; R <- results[[cell_name]]
    rbind(one_xgb_diag_row(cell_name, "PLR-IV XGB untuned", R$untuned, cell),
          one_xgb_diag_row(cell_name, "PLR-IV XGB tuned",   R$tuned,   cell))
  }))

  print_tbl <- xgb_pliv_weight_table
  for (col in c("Estimate","Sigma omega","Sigma abs omega","Treated mass","Control mass (-sum control weights)","max abs omega"))
    print_tbl[[col]] <- sprintf("%.4f", print_tbl[[col]])
  print_tbl$ESS     <- sprintf("%.0f", print_tbl$ESS)
  print_tbl$ESS_mod <- sprintf("%.0f", print_tbl$ESS_mod)

  print_tbl %>%
    kbl(caption = "Card: PLR-IV XGBoost outcome-weight diagnostics, untuned vs. tuned") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12) %>%
    row_spec(which(xgb_pliv_weight_table$Learner == "PLR-IV XGB tuned"), background = "#eef6ff") %>%
    footnote(general = "Identity ok = TRUE means sum(omega*Y) reproduces the PLR-IV coefficient to machine precision. If Weight status reports a failed fit, the corresponding DoubleML object was NULL before weight extraction began. If Weight status reports a weight-extraction failure, the model fit exists but get_outcome_weights() failed or was not available for that object.")
} else {
  xgb_pliv_weight_table <- data.frame()
  cat("Skipped PLR-IV XGBoost outcome-weight extraction because EXTRACT_WEIGHTS = FALSE.\n")
  cat("Reason: get_outcome_weights() is currently too slow for these PLR-IV XGBoost objects.\n")
}
```

---

## 5. Export for card_test.Rmd

```{r xgb-pliv-omega-vectors}
if (isTRUE(EXTRACT_WEIGHTS)) {
  xgb_pliv_omega_vectors <- lapply(names(cells), function(cell_name) {
    list(
      untuned = safe_extract_xgb_w(results[[cell_name]]$untuned, cells[[cell_name]]),
      tuned   = safe_extract_xgb_w(results[[cell_name]]$tuned,   cells[[cell_name]])
    )
  })
  names(xgb_pliv_omega_vectors) <- names(cells)
} else {
  xgb_pliv_omega_vectors <- list()
}
```

```{r xgb-pliv-identity-checks}
if (isTRUE(EXTRACT_WEIGHTS)) {
  xgb_pliv_identity_checks <- lapply(names(cells), function(cell_name) {
    cell <- cells[[cell_name]]
    list(
      untuned = tryCatch(
        if (is.null(xgb_pliv_omega_vectors[[cell_name]]$untuned) || is.null(results[[cell_name]]$untuned)) NA else
          check_weight_identity(xgb_pliv_omega_vectors[[cell_name]]$untuned, cell$Y, results[[cell_name]]$untuned$coef),
        error = function(e) NA
      ),
      tuned = tryCatch(
        if (is.null(xgb_pliv_omega_vectors[[cell_name]]$tuned) || is.null(results[[cell_name]]$tuned)) NA else
          check_weight_identity(xgb_pliv_omega_vectors[[cell_name]]$tuned, cell$Y, results[[cell_name]]$tuned$coef),
        error = function(e) NA
      )
    )
  })
  names(xgb_pliv_identity_checks) <- names(cells)

  cat("Identity checks (sum(omega*Y) = tau_hat):\n")
  for (cn in names(xgb_pliv_identity_checks)) {
    cat(sprintf("  %-20s untuned=%s | tuned=%s\n", cn,
                xgb_pliv_identity_checks[[cn]]$untuned,
                xgb_pliv_identity_checks[[cn]]$tuned))
  }
} else {
  xgb_pliv_identity_checks <- list()
  cat("No PLR-IV XGBoost identity checks exported because weights were not extracted.\n")
}
```

```{r xgb-pliv-export, cache=FALSE}
xgb_pliv_export <- list(
  estimate_comparison = rows,
  nuisance_losses     = loss_rows,
  hyperparameters     = if (exists("hp_summary")) hp_summary else NULL,
  weight_diagnostics  = xgb_pliv_weight_table,
  omega_vectors       = xgb_pliv_omega_vectors,
  identity_checks     = xgb_pliv_identity_checks,
  settings = list(
    seed          = SEED,
    n_cf          = N_CF,
    tune_evals    = TUNE_EVALS,
    inner_folds   = INNER_FOLDS,
    tune_on_folds = TRUE,
    learner       = "xgboost",
    dml_class     = "DoubleMLPLIV",
    score         = "partialling out",
    untuned_base  = list(alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
    extract_weights = EXTRACT_WEIGHTS
  ),
  sample_info = list(
    n = nrow(card), mean_Z = mean(card$Z_inst),
    card_somecol_mean_D = mean(card$somecol),
    kitagawa_completion_mean_D = mean(card$educ16)
  )
)

saveRDS(xgb_pliv_export, "card_xgb_pliv_tuning_export.rds")
cat("Saved: card_xgb_pliv_tuning_export.rds\n")
cat("Contents: estimate_comparison, nuisance_losses, hyperparameters, weight_diagnostics,\n")
cat("          omega_vectors, identity_checks, settings, sample_info.\n")
```
