---
title: "Card (1995) — Translation Invariance: Complete-Pipeline Rerun"
subtitle: "Kappa, honest forest, DoubleML learners, and nested XGBoost"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1   { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2   { color: #2c7bb6; margin-top: 28px; }
h3   { color: #444; margin-top: 20px; }
table { font-size: 13px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 12, fig.height = 6, cache = TRUE, autodep = TRUE,
  cache.extra = "translation-rerun-only-v3-canonical-kappa"
)
```

---

# Purpose

This notebook evaluates translation invariance by fitting each complete
estimation pipeline to log hourly wages in dollars and to
\[
Y_i^{\mathrm{shift}}=Y_i+\log(100),
\]
the cents recoding used in the SUW reproduction. The empirical check compares
the two independently fitted estimates. It does not reuse fitted nuisance
models from the original outcome.

<div class="note-box">
The full `2 treatments × 2 covariate specifications` design is evaluated.
For XGBoost, the original tuned estimate is imported from the verified tuning
export and the shifted-outcome model repeats nested tuning under the same
protocol.
</div>

# Setup

```{r packages-and-source, include=FALSE}
library(AER)
library(haven)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(cobalt)
library(viridis)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

source("functions_all.R")

RERUN_TOLERANCE <- 1e-4

ti_card_row <- function(name, tau_original, tau_shifted, k, digits = 12) {
  out <- translation_rerun_row(
    name, tau_original, tau_shifted, k,
    tolerance = RERUN_TOLERANCE, digits = digits
  )
  out$invariant <- abs(out$rerun_difference) <= RERUN_TOLERANCE
  out
}
```

## Runtime switches

```{r switches, include=FALSE}
RUN_KAPPA             <- TRUE
RUN_DML               <- TRUE
RUN_LM_RF             <- TRUE
RUN_XGB_RERUN         <- TRUE
RUN_XGB_WALD_AIPW     <- TRUE
RUN_XGB_PLIV          <- TRUE
USE_XGB_ORIGINAL_EXPORT <- TRUE
XGB_EXPORT_FILES <- c("card_xgb_tuning_export.rds", "card_xgb_pliv_tuning_export.rds")
XGB_EXPORTS_AVAILABLE <- all(file.exists(XGB_EXPORT_FILES))
```

<div class="warn-box">
**XGBoost rule.** Set `RUN_XGB_RERUN = TRUE` only for a deliberate nested
tuning rerun. Reusing the original selected hyperparameters after shifting the
outcome would not test the complete tuned procedure.
</div>

## Data — the exact main-analysis sample

```{r data-and-design}
card <- read_dta("card.dta")
card <- card[!is.na(card$wage) & !is.na(card$educ), ]
if (nrow(card) != 3010L) stop("Expected the Card analysis sample to contain 3,010 observations.")

card$somecol <- as.integer(card$educ >= 13)
card$educ16  <- as.integer(card$educ >= 16)
card$Z_inst  <- as.integer(card$nearc4)
if (!"expersq" %in% names(card)) card$expersq <- card$exper^2

k <- log(100)
Y_orig  <- log(card$wage / 100)
Y_shift <- log(card$wage)
max_dev <- max(abs((Y_shift - Y_orig) - k))
stopifnot(max_dev < 1e-9)

X_card_kappa <- model.matrix(
  ~ exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card)
X_kit_kappa <- model.matrix(~ black + south + smsa + smsa66 + south66, data = card)
X_card <- make_X(
  ~ 0 + exper + expersq + black + south + smsa + smsa66 +
    reg661 + reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668,
  data = card)
X_kit <- make_X(~ 0 + black + south + smsa + smsa66 + south66, data = card)

workflows <- list(
  "Card / some college" = list(label = "Card / some college", D = card$somecol, Z = card$Z_inst, X_kappa = X_card_kappa, X_dml = X_card, X_dml_lm = X_card, Y_orig = Y_orig, Y_shift = Y_shift, k = k),
  "Card / college completion" = list(label = "Card / college completion", D = card$educ16, Z = card$Z_inst, X_kappa = X_card_kappa, X_dml = X_card, X_dml_lm = X_card, Y_orig = Y_orig, Y_shift = Y_shift, k = k),
  "Kitagawa / some college" = list(label = "Kitagawa / some college", D = card$somecol, Z = card$Z_inst, X_kappa = X_kit_kappa, X_dml = X_kit, X_dml_lm = X_kit, Y_orig = Y_orig, Y_shift = Y_shift, k = k),
  "Kitagawa / college completion" = list(label = "Kitagawa / college completion", D = card$educ16, Z = card$Z_inst, X_kappa = X_kit_kappa, X_dml = X_kit, X_dml_lm = X_kit, Y_orig = Y_orig, Y_shift = Y_shift, k = k)
)

data.frame(
  Quantity = c(
    "Analysis sample", "Treatment definitions", "Covariate specifications",
    "Translation", "Maximum numerical deviation from exact shift"
  ),
  Value = c(
    format(nrow(card), big.mark = ","), "2", "2",
    "log wage + log(100)", format(max_dev, scientific = TRUE)
  )
) %>%
  kbl(caption = "Translation-rerun design") %>%
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE)
```

---

# 1. Kappa estimators

```{r kappa-estimator-names, include=FALSE}
kappa_estimator_names <- c(
  "tau_cb_u", "tau_ml_u", "tau_ml_a10",
  "tau_ml_a", "tau_ml_t", "tau_ml_a0"
)

build_kappa_ti_rows <- function(wf) {
  est_o <- kappa_point_estimates(wf$Y_orig,  wf$Z, wf$D, wf$X_kappa)
  est_s <- kappa_point_estimates(wf$Y_shift, wf$Z, wf$D, wf$X_kappa)
  rows <- lapply(kappa_estimator_names, function(est_name) {
    row <- ti_card_row(name = est_name, tau_original = est_o[[est_name]],
                        tau_shifted = est_s[[est_name]], k = wf$k)
    row$Specification <- wf$label
    row
  })
  do.call(rbind, rows)
}
```

```{r kappa-ti-run, include=FALSE}
if (RUN_KAPPA) {
  kappa_ti_all <- do.call(rbind, lapply(workflows, build_kappa_ti_rows))
}
```

```{r kappa-ti-table}
if (RUN_KAPPA) {
  kappa_ti_all[, c("Specification", setdiff(names(kappa_ti_all), "Specification"))] %>%
    kbl(caption = "Kappa estimators — original versus shifted outcome",
        digits = 8, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!kappa_ti_all$invariant), background = "#fffbf0") %>%
    footnote(general = paste0(
      "Orange rows have an absolute rerun difference above ",
      format(RERUN_TOLERANCE, scientific = TRUE), "."
    ))
}
```

---

# 2. Headline DML: tuned honest forest

Tuned honest forests use `tune.parameters="all"` in both reruns.

```{r dml-headline-builder, include=FALSE}
build_dml_headline_ti_rows <- function(wf, n_cf_folds = 5) {
  set.seed(42)
  fit_o <- dml_with_smoother(wf$Y_orig, wf$D, wf$X_dml, wf$Z,
                             n_cf_folds = n_cf_folds, tune.parameters = "all")
  set.seed(42)
  fit_s <- dml_with_smoother(wf$Y_shift, wf$D, wf$X_dml, wf$Z,
                             n_cf_folds = n_cf_folds, tune.parameters = "all")

  res_o <- summary(fit_o); res_s <- summary(fit_s)
  rows <- lapply(c("PLR-IV", "Wald-AIPW"), function(est) {
    row <- ti_card_row(name = est, tau_original = get_estimate(res_o, est),
                        tau_shifted = get_estimate(res_s, est), k = wf$k)
    row$Specification <- wf$label
    row
  })
  do.call(rbind, rows)
}
```

```{r dml-headline-run, include=FALSE}
if (RUN_DML) {
  dml_headline_ti_all <- do.call(rbind, lapply(workflows, build_dml_headline_ti_rows))
}
```

```{r dml-headline-table}
if (RUN_DML) {
  dml_headline_ti_all[, c("Specification", setdiff(names(dml_headline_ti_all), "Specification"))] %>%
    kbl(caption = "Tuned honest forest — original versus shifted outcome",
        digits = 8, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!dml_headline_ti_all$invariant), background = "#fffbf0")
}
```

---

# 3. OLS/logit and Ranger learners

The learner definitions match `card_test.Rmd`.

```{r dml-learner-setup, include=FALSE}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z", x_cols = colnames(X))
}

ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))

ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))

fit_iivm <- function(dml_data, learners) {
  set.seed(42)
  o <- DoubleMLIIVM$new(data = dml_data, ml_g = learners$g, ml_m = learners$m,
                        ml_r = learners$r, n_folds = 5, score = "LATE")
  o$fit(store_models = TRUE, store_predictions = TRUE)
  o
}

fit_pliv <- function(dml_data, learners) {
  set.seed(42)
  o <- DoubleMLPLIV$new(data = dml_data, ml_l = learners$l, ml_m = learners$m,
                        ml_r = learners$r, n_folds = 5)
  o$fit(store_models = TRUE, store_predictions = TRUE)
  o
}

build_dml_learner_ti_row <- function(wf, label, learners_fn, X_dml, fit_fn) {
  dat_o <- make_dml_data(wf$Y_orig,  wf$D, wf$Z, X_dml)
  dat_s <- make_dml_data(wf$Y_shift, wf$D, wf$Z, X_dml)

  fit_o <- fit_fn(dat_o, learners_fn())
  fit_s <- fit_fn(dat_s, learners_fn())

  row <- ti_card_row(name = label, tau_original = get_dml_coef(fit_o),
                      tau_shifted = get_dml_coef(fit_s), k = wf$k)
  row$Specification <- wf$label
  row
}
```

```{r dml-learner-run, include=FALSE}
if (RUN_LM_RF) {
  dml_learner_ti_all <- do.call(rbind, lapply(workflows, function(wf) {
    rbind(
      build_dml_learner_ti_row(wf, "PLR-IV (linear)",          ml_lm_pliv,  wf$X_dml_lm, fit_pliv),
      build_dml_learner_ti_row(wf, "Wald-AIPW (linear+logit)", ml_lm_iivm,  wf$X_dml_lm, fit_iivm),
      build_dml_learner_ti_row(wf, "PLR-IV (ranger)",          ml_rf_pliv,  wf$X_dml,    fit_pliv),
      build_dml_learner_ti_row(wf, "Wald-AIPW (ranger)",       ml_rf_iivm,  wf$X_dml,    fit_iivm)
    )
  }))
}
```

```{r dml-learner-table}
if (RUN_LM_RF) {
  dml_learner_ti_all[, c("Specification", setdiff(names(dml_learner_ti_all), "Specification"))] %>%
    kbl(caption = "DoubleML learners — original versus shifted outcome",
        digits = 8, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!dml_learner_ti_all$invariant), background = "#fffbf0") %>%
    footnote(general = paste0(
      "PLR-IV uses DoubleMLPLIV and Wald-AIPW uses DoubleMLIIVM. Orange rows have an absolute rerun difference above ",
      format(RERUN_TOLERANCE, scientific = TRUE), "."
    ))
}
```

# 4. Tuned XGBoost complete-pipeline rerun

Original tuned estimates are imported from the two Card tuning exports. For
the shifted outcome, tuning is repeated separately inside each outer training
fold using the same search space, folds, budget, and seed.

```{r xgb-setup, include=FALSE}
SEED <- 42; N_CF <- 5; TUNE_EVALS <- 15; INNER_FOLDS <- 3

if (!USE_XGB_ORIGINAL_EXPORT) {
  stop("This notebook expects USE_XGB_ORIGINAL_EXPORT <- TRUE for the original XGBoost side.")
}

xgb_wald_export <- if (XGB_EXPORTS_AVAILABLE) readRDS(XGB_EXPORT_FILES[[1]]) else NULL
xgb_pliv_export <- if (XGB_EXPORTS_AVAILABLE) readRDS(XGB_EXPORT_FILES[[2]]) else NULL

validate_xgb_export <- function(export, path, required_cells) {
  settings <- export$settings
  if (is.null(settings) || !isTRUE(settings$tune_on_folds)) {
    stop(sprintf("%s is not a tune_on_folds=TRUE export.", path))
  }
  expected <- c(seed = SEED, n_cf = N_CF, tune_evals = TUNE_EVALS,
                inner_folds = INNER_FOLDS)
  if (!all(names(expected) %in% names(settings))) {
    stop(sprintf("%s is missing required tuning metadata.", path))
  }
  observed <- vapply(names(expected), function(x) as.integer(settings[[x]]), integer(1))
  if (!identical(unname(observed), unname(as.integer(expected)))) {
    stop(sprintf("%s uses tuning settings that differ from this rerun.", path))
  }
  if (is.null(export$weight_diagnostics) ||
      !all(required_cells %in% export$weight_diagnostics$Cell)) {
    stop(sprintf("%s does not contain every required Card cell.", path))
  }
  invisible(TRUE)
}

if (RUN_XGB_RERUN && !XGB_EXPORTS_AVAILABLE) {
  stop("Run both Card XGBoost tuning notebooks before the translation rerun.")
}

if (XGB_EXPORTS_AVAILABLE) {
  validate_xgb_export(xgb_wald_export, XGB_EXPORT_FILES[[1]], names(workflows))
  validate_xgb_export(xgb_pliv_export, XGB_EXPORT_FILES[[2]], names(workflows))
}

make_xgb_learners_iivm <- function() {
  list(g = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
       m = lrn("classif.xgboost", predict_type = "prob"),
       r = lrn("classif.xgboost", predict_type = "prob"))
}
make_xgb_learners_pliv <- function() {
  list(l = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
       m = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0),
       r = lrn("regr.xgboost", alpha = 0, subsample = 1, max_delta_step = 0, base_score = 0))
}
xgb_ps <- function() {
  ps(nrounds = p_int(50, 400), max_depth = p_int(2, 5),
     eta = p_dbl(0.02, 0.20), min_child_weight = p_int(1, 8))
}
tune_param_set_iivm <- function() list(ml_g = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())
tune_param_set_pliv <- function() list(ml_l = xgb_ps(), ml_m = xgb_ps(), ml_r = xgb_ps())
tune_settings_iivm <- function() {
  list(n_folds_tune = INNER_FOLDS, rsmp_tune = rsmp("cv", folds = INNER_FOLDS),
       measure = list(ml_g = "regr.rmse", ml_m = "classif.logloss", ml_r = "classif.logloss"),
       terminator = trm("evals", n_evals = TUNE_EVALS), algorithm = tnr("random_search"))
}
tune_settings_pliv <- function() {
  list(n_folds_tune = INNER_FOLDS, rsmp_tune = rsmp("cv", folds = INNER_FOLDS),
       measure = list(ml_l = "regr.rmse", ml_m = "regr.rmse", ml_r = "regr.rmse"),
       terminator = trm("evals", n_evals = TUNE_EVALS), algorithm = tnr("random_search"))
}

get_xgb_original <- function(export, cell, learner) {
  row <- export$weight_diagnostics |>
    filter(.data$Cell == cell, .data$Learner == learner)
  if (nrow(row) != 1) {
    stop(sprintf("Expected one export row for %s / %s, found %d.", cell, learner, nrow(row)))
  }
  as.numeric(row$Estimate[[1]])
}

fit_xgb_wald_shifted <- function(Y, D, Z, X) {
  dml_data <- make_dml_data(Y, D, Z, X)
  learners <- make_xgb_learners_iivm()
  o <- DoubleMLIIVM$new(data = dml_data, ml_g = learners$g, ml_m = learners$m,
                        ml_r = learners$r, n_folds = N_CF, score = "LATE")
  o$tune(param_set = tune_param_set_iivm(), tune_settings = tune_settings_iivm(),
         tune_on_folds = TRUE)
  o$fit(store_models = TRUE, store_predictions = TRUE)
  list(obj = o, tau = get_dml_coef(o))
}

fit_xgb_pliv_shifted <- function(Y, D, Z, X) {
  dml_data <- make_dml_data(Y, D, Z, X)
  learners <- make_xgb_learners_pliv()
  o <- DoubleMLPLIV$new(data = dml_data, ml_l = learners$l, ml_m = learners$m,
                        ml_r = learners$r, n_folds = N_CF)
  o$tune(param_set = tune_param_set_pliv(), tune_settings = tune_settings_pliv(),
         tune_on_folds = TRUE)
  o$fit(store_models = TRUE, store_predictions = TRUE)
  list(obj = o, tau = get_dml_coef(o))
}

build_xgb_rerun_rows <- function(wf) {
  rows <- list()

  if (RUN_XGB_WALD_AIPW) {
    original <- get_xgb_original(xgb_wald_export, wf$label, "XGB tuned")
    set.seed(SEED)
    shifted <- fit_xgb_wald_shifted(wf$Y_shift, wf$D, wf$Z, wf$X_dml)
    row <- ti_card_row(name = "Wald-AIPW (XGBoost tuned)",
                        tau_original = original,
                        tau_shifted = shifted$tau,
                        k = wf$k)
    row$Specification <- wf$label
    rows[["wald"]] <- row
  }

  if (RUN_XGB_PLIV) {
    original <- get_xgb_original(xgb_pliv_export, wf$label, "PLR-IV XGB tuned")
    set.seed(SEED)
    shifted <- fit_xgb_pliv_shifted(wf$Y_shift, wf$D, wf$Z, wf$X_dml)
    row <- ti_card_row(name = "PLR-IV (XGBoost tuned)",
                        tau_original = original,
                        tau_shifted = shifted$tau,
                        k = wf$k)
    row$Specification <- wf$label
    rows[["pliv"]] <- row
  }

  do.call(rbind, rows)
}

save_xgb_rerun_export <- function(xgb_ti_all) {
  saveRDS(
    list(results = xgb_ti_all,
         settings = list(seed = SEED, n_cf = N_CF, tune_evals = TUNE_EVALS,
                         inner_folds = INNER_FOLDS, tune_on_folds = TRUE,
                         original_exports = c("card_xgb_tuning_export.rds",
                                              "card_xgb_pliv_tuning_export.rds"))),
    "card_translation_xgb_rerun_export.rds"
  )
}
```

```{r xgb-run, include=FALSE, cache.extra=list(RUN_XGB_RERUN, RUN_XGB_WALD_AIPW, RUN_XGB_PLIV, TUNE_EVALS, INNER_FOLDS)}
if (RUN_XGB_RERUN && XGB_EXPORTS_AVAILABLE) {
  xgb_ti_all <- do.call(rbind, lapply(workflows, build_xgb_rerun_rows))
  save_xgb_rerun_export(xgb_ti_all)
}
```

```{r xgb-table}
if (RUN_XGB_RERUN && XGB_EXPORTS_AVAILABLE) {
  xgb_ti_all[, c("Specification", setdiff(names(xgb_ti_all), "Specification"))] %>%
    kbl(caption = "Tuned XGBoost — original export versus shifted nested-tuning rerun",
        digits = 8, row.names = FALSE) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    collapse_rows(columns = 1, valign = "middle") %>%
    row_spec(which(!xgb_ti_all$invariant), background = "#fffbf0") %>%
    footnote(general = paste0(
      "Original estimates are imported from the verified tuning exports. ",
      "Shifted estimates repeat nested tuning. Orange rows have an absolute ",
      "rerun difference above ",
      format(RERUN_TOLERANCE, scientific = TRUE), "."
    ))
}
```

---

# 5. Save reusable rerun results

```{r save-complete-rerun, cache=FALSE, include=FALSE}
result_parts <- Filter(Negate(is.null), list(
  Kappa = if (exists("kappa_ti_all")) kappa_ti_all else NULL,
  `Tuned honest forest` = if (exists("dml_headline_ti_all")) dml_headline_ti_all else NULL,
  `Linear and Ranger` = if (exists("dml_learner_ti_all")) dml_learner_ti_all else NULL,
  `Tuned XGBoost` = if (exists("xgb_ti_all")) xgb_ti_all else NULL
))
for (nm in names(result_parts)) result_parts[[nm]]$Family <- nm
translation_results <- bind_rows(result_parts)
saveRDS(list(
  results = translation_results,
  settings = list(seed=42,n_cf=5,shift=log(100),
                  rerun_tolerance=RERUN_TOLERANCE,grf_tuning="all",
                  xgboost_tune_on_folds=TRUE,xgboost_tune_evals=TUNE_EVALS,
                  xgboost_inner_folds=INNER_FOLDS)
), "card_translation_rerun_results.rds")
```
