---
title: "Angrist (1990) — Translation Invariance: Complete-Pipeline Rerun"
subtitle: "Vietnam draft lottery: kappa, honest forest, DoubleML learners, and nested XGBoost"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size:15px; color:#2c2c2c; }
h1 { border-bottom:2px solid #2c7bb6; padding-bottom:6px; margin-top:36px; }
h2 { color:#2c7bb6; margin-top:28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size:13px; }
pre:not(.sourceCode) { display:none; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=6,
                      cache.extra="translation-rerun-only-v3-canonical-kappa")
```

# Purpose

This notebook evaluates translation invariance by fitting the complete estimation
pipeline to the original log-wage outcome and to

\[
Y_i^{\mathrm{shift}}=Y_i+\log(100),
\]

which is the cents recoding used in the SUW reproduction. The empirical check is
the direct difference between the two separately fitted point estimates. This
is the full-pipeline rerun called Method A in the thesis, but throughout this
notebook it is simply referred to as the rerun check. It extends the classical
kappa comparison to data-adaptive estimators by allowing outcome nuisances and
tuning choices to be re-estimated under the same cross-fitting protocol after
the translation. Reusing the
original fitted models or selected hyperparameters would not evaluate the
complete implemented estimator.

<div class="note-box">
**Two workflows.** The same pure additive shift is applied to the cubic-age
and saturated-age specifications. These are the two feature designs used by
the DML and learner analyses. Repeating additional constants would not test a
different transformation property. The complete estimator is rerun once for
each of these two analysis specifications.
</div>

For the table classification, estimates are described as numerically unchanged
when the absolute rerun difference does not exceed \(10^{-4}\). The exact
rerun difference remains reported in every table.

# Setup

```{r packages-and-source}
library(haven)
library(knitr)
library(kableExtra)
library(dplyr)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")
source("functions_all.R")

RERUN_TOLERANCE <- 1e-4

ti_vietnam_row <- function(name,tau_original,tau_shifted,k,digits=12) {
  translation_rerun_row(
    name,tau_original,tau_shifted,k,
    tolerance=RERUN_TOLERANCE,digits=digits
  )
}

classify_rerun_results <- function(x) {
  x$invariant <- abs(x$rerun_difference) <= RERUN_TOLERANCE
  x
}

display_rerun_results <- function(x) {
  out <- x[,c("Specification","Estimator","Shift_k","tau_original",
              "tau_shifted","rerun_difference","invariant")]
  names(out) <- c("Specification","Estimator","Shift",
                  "Original estimate","Shifted estimate","Rerun change",
                  "Numerically unchanged")
  out
}
```

## Runtime switches

```{r switches}
RUN_KAPPA <- TRUE
RUN_DML <- TRUE
RUN_LM_RF <- TRUE
RUN_XGB_RERUN <- TRUE
RUN_XGB_WALD_AIPW <- TRUE
RUN_XGB_PLIV <- TRUE
USE_XGB_ORIGINAL_EXPORT <- TRUE

SEED <- 42
N_CF <- 5
TUNE_EVALS <- 15
INNER_FOLDS <- 3
```

<div class="warn-box">
**XGBoost rule.** Original-outcome tuned estimates are
imported from the two tuning exports. For the shifted outcome, nested tuning
is rerun with the same seed, five outer folds, three inner folds, fifteen
random-search evaluations, and `tune_on_folds = TRUE`. Reusing median
hyperparameters would define a different estimator and is deliberately not
done.
</div>

## Data — the exact main-analysis sample

```{r data}
stopifnot(file.exists("sipp.dta"))
sipp <- read_dta("sipp.dta")
sipp <- sipp[sipp$rsncode!=999 & !is.na(sipp$educ) & !is.na(sipp$kwage),]
sipp$age <- sipp$age_5
sipp$age2 <- sipp$age_5^2
sipp$age3 <- sipp$age_5^3
sipp$age_fac <- factor(sipp$age_5)

Y_orig <- as.numeric(log(sipp$kwage))
k_shift <- log(100)
Y_shift <- Y_orig+k_shift
D <- as.integer(sipp$nvstat)
Z <- as.integer(sipp$rsncode)

X_kappa_cub <- model.matrix(~age+age2+age3,data=sipp)
X_kappa_sat <- model.matrix(~age_fac,data=sipp)
X_dml_cub <- make_X(~0+age+I(age^2)+I(age^3),data=sipp)
X_dml_sat_full <- make_X(~0+age_fac,data=sipp)
X_dml_sat <- X_dml_sat_full[,-1,drop=FALSE]

workflows <- list(
  "Cubic age"=list(label="Cubic age",Y_orig=Y_orig,Y_shift=Y_shift,k=k_shift,
    D=D,Z=Z,X_kappa=X_kappa_cub,X_dml=X_dml_cub),
  "Saturated age"=list(label="Saturated age",Y_orig=Y_orig,Y_shift=Y_shift,k=k_shift,
    D=D,Z=Z,X_kappa=X_kappa_sat,X_dml=X_dml_sat))

data.frame(
  Sample = "Main-analysis sample",
  N = nrow(sipp),
  Translation = "log wage + log(100)",
  Specifications = paste(names(workflows),collapse=", ")
) %>%
  kbl(caption="Rerun setup",row.names=FALSE) %>%
  kable_styling(bootstrap_options=c("striped","condensed"),full_width=FALSE)
```

# 1. Kappa estimators

<div class="note-box">
The normalized estimators should remain unchanged after the complete rerun,
whereas the unnormalized estimators are expected to move. These fits are cheap
and provide the classical benchmark for the data-adaptive estimators below.
</div>

```{r kappa-builders}
kappa_estimator_names <- c("tau_cb_u","tau_ml_u","tau_ml_a10",
  "tau_ml_a","tau_ml_t","tau_ml_a0")
build_kappa_rows <- function(wf) {
  est_o <- kappa_point_estimates(wf$Y_orig,wf$Z,wf$D,wf$X_kappa)
  est_s <- kappa_point_estimates(wf$Y_shift,wf$Z,wf$D,wf$X_kappa)
  do.call(rbind,lapply(kappa_estimator_names,function(est){
    row <- ti_vietnam_row(est,est_o[[est]],est_s[[est]],wf$k)
    row$Specification <- wf$label; row
  }))
}
```

```{r kappa-run, cache=TRUE}
if(RUN_KAPPA)kappa_ti_all<-do.call(rbind,lapply(workflows,build_kappa_rows))
```

```{r kappa-table}
if(RUN_KAPPA) {
  kappa_ti_all <- classify_rerun_results(kappa_ti_all)
  display_rerun_results(kappa_ti_all) %>%
    kbl(caption="Kappa estimators — original versus shifted log wage",digits=8,row.names=FALSE) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
    row_spec(which(!kappa_ti_all$invariant),background="#fffbf0") %>%
    footnote(general=paste0(
      "Shading marks absolute rerun differences above ",
      format(RERUN_TOLERANCE,scientific=TRUE),
      ". This occurs for the cubic unnormalized kappa variants."
    ))
}
```

# 2. Headline DML: tuned honest forest

```{r dml-headline-builder}
build_dml_headline_rows <- function(wf) {
  set.seed(SEED)
  invisible(capture.output(
    fit_o <- dml_with_smoother(wf$Y_orig,wf$D,wf$X_dml,wf$Z,
      n_cf_folds=N_CF,tune.parameters="all")
  ))
  set.seed(SEED)
  invisible(capture.output(
    fit_s <- dml_with_smoother(wf$Y_shift,wf$D,wf$X_dml,wf$Z,
      n_cf_folds=N_CF,tune.parameters="all")
  ))
  invisible(capture.output(ro <- summary(fit_o)))
  invisible(capture.output(rs <- summary(fit_s)))
  do.call(rbind,lapply(c("PLR-IV","Wald-AIPW"),function(est){
    row <- ti_vietnam_row(est,get_estimate(ro,est),
                          get_estimate(rs,est),wf$k)
    row$Specification <- wf$label; row
  }))
}
```

```{r dml-headline-run, cache=TRUE}
if(RUN_DML)dml_headline_ti_all<-do.call(rbind,lapply(workflows,build_dml_headline_rows))
```

```{r dml-headline-table}
if(RUN_DML) {
  dml_headline_ti_all <- classify_rerun_results(dml_headline_ti_all)
  display_rerun_results(dml_headline_ti_all) %>%
    kbl(caption="Tuned honest forest — original versus shifted log wage",digits=8,row.names=FALSE) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
    row_spec(which(!dml_headline_ti_all$invariant),background="#fffbf0") %>%
    footnote(general="The exact rerun differences are shown; all four are numerically unchanged at the stated reporting threshold.")
}
```

# 3. OLS/logit and Ranger learners

<div class="note-box">
“OLS/logit” describes the nuisance learner, not the linear-age feature
specification. Every row is evaluated once with cubic-age features and once
with saturated-age features. PLR-IV uses OLS for all three nuisances;
Wald-AIPW uses OLS for the outcome nuisance and logistic regression for the
binary instrument and treatment nuisances.
</div>

```{r learner-setup}
make_dml_data <- function(Y,D,Z,X){df<-as.data.frame(X);df$y<-Y;df$d<-D;df$z<-Z
  DoubleMLData$new(data=df,y_col="y",d_cols="d",z_cols="z",x_cols=colnames(X))}
ml_lm_iivm<-function()list(g=lrn("regr.lm"),m=lrn("classif.log_reg"),r=lrn("classif.log_reg"))
ml_rf_iivm<-function()list(g=lrn("regr.ranger",keep.inbag=TRUE),m=lrn("classif.ranger",keep.inbag=TRUE),r=lrn("classif.ranger",keep.inbag=TRUE))
ml_lm_pliv<-function()list(l=lrn("regr.lm"),m=lrn("regr.lm"),r=lrn("regr.lm"))
ml_rf_pliv<-function()list(l=lrn("regr.ranger",keep.inbag=TRUE),m=lrn("regr.ranger",keep.inbag=TRUE),r=lrn("regr.ranger",keep.inbag=TRUE))
fit_iivm<-function(dat,L){set.seed(SEED);o<-DoubleMLIIVM$new(data=dat,ml_g=L$g,ml_m=L$m,ml_r=L$r,n_folds=N_CF,score="LATE")
  invisible(capture.output(o$fit(store_models=TRUE,store_predictions=TRUE)));o}
fit_pliv<-function(dat,L){set.seed(SEED);o<-DoubleMLPLIV$new(data=dat,ml_l=L$l,ml_m=L$m,ml_r=L$r,n_folds=N_CF)
  invisible(capture.output(o$fit(store_models=TRUE,store_predictions=TRUE)));o}
build_learner_row<-function(wf,label,Lfun,fitfun){
  fo<-fitfun(make_dml_data(wf$Y_orig,wf$D,wf$Z,wf$X_dml),Lfun())
  fs<-fitfun(make_dml_data(wf$Y_shift,wf$D,wf$Z,wf$X_dml),Lfun())
  row<-ti_vietnam_row(label,get_dml_coef(fo),get_dml_coef(fs),wf$k)
  row$Specification<-wf$label;row}
```

```{r learner-run, cache=TRUE}
if(RUN_LM_RF)learner_ti_all<-do.call(rbind,lapply(workflows,function(wf)rbind(
  build_learner_row(wf,"PLR-IV (OLS nuisances)",ml_lm_pliv,fit_pliv),
  build_learner_row(wf,"Wald-AIPW (OLS/logit nuisances)",ml_lm_iivm,fit_iivm),
  build_learner_row(wf,"PLR-IV (Ranger)",ml_rf_pliv,fit_pliv),
  build_learner_row(wf,"Wald-AIPW (Ranger)",ml_rf_iivm,fit_iivm))))
```

```{r learner-table}
if(RUN_LM_RF) {
  learner_ti_all <- classify_rerun_results(learner_ti_all)
  display_rerun_results(learner_ti_all) %>%
    kbl(caption="DoubleML learners — original versus shifted log wage",digits=8,row.names=FALSE) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
    row_spec(which(!learner_ti_all$invariant),background="#fffbf0") %>%
    footnote(general="OLS/logit labels the nuisance learner class; cubic and saturated identify the feature specification. All reported rerun differences are within the numerical threshold.")
}
```

# 4. Tuned XGBoost complete-pipeline rerun

```{r xgb-setup}
if(RUN_XGB_RERUN){
  if(!USE_XGB_ORIGINAL_EXPORT)stop("USE_XGB_ORIGINAL_EXPORT must be TRUE.")
  req<-c("vietnam_xgb_tuning_export.rds","vietnam_xgb_pliv_tuning_export.rds")
  if(!all(file.exists(req)))stop("Run both Vietnam XGBoost tuning notebooks before the translation rerun.")
  xgb_wald_export<-readRDS(req[1]);xgb_pliv_export<-readRDS(req[2])
  for(ex in list(xgb_wald_export,xgb_pliv_export)){
    if(!isTRUE(ex$settings$tune_on_folds))stop("Export is not a tune_on_folds=TRUE fit.")
    if(!all(names(workflows)%in%ex$weight_diagnostics$Cell))stop("Export cells do not match the translation workflows.")
    if(!identical(as.integer(ex$settings$seed),as.integer(SEED))||
       !identical(as.integer(ex$settings$n_cf),as.integer(N_CF))||
       !identical(as.integer(ex$settings$tune_evals),as.integer(TUNE_EVALS))||
       !identical(as.integer(ex$settings$inner_folds),as.integer(INNER_FOLDS)))
      stop("Export tuning settings do not match the shifted-outcome rerun settings.")
  }
}

xgb_ps<-function()ps(nrounds=p_int(50,400),max_depth=p_int(2,5),eta=p_dbl(0.02,0.20),min_child_weight=p_int(1,8))
make_xgb_iivm<-function()list(g=lrn("regr.xgboost",alpha=0,subsample=1,max_delta_step=0,base_score=0),m=lrn("classif.xgboost",predict_type="prob"),r=lrn("classif.xgboost",predict_type="prob"))
make_xgb_pliv<-function(){b<-list(alpha=0,subsample=1,max_delta_step=0,base_score=0);list(l=do.call(lrn,c(list("regr.xgboost"),b)),m=do.call(lrn,c(list("regr.xgboost"),b)),r=do.call(lrn,c(list("regr.xgboost"),b)))}
tune_set_iivm<-function()list(ml_g=xgb_ps(),ml_m=xgb_ps(),ml_r=xgb_ps())
tune_set_pliv<-function()list(ml_l=xgb_ps(),ml_m=xgb_ps(),ml_r=xgb_ps())
tune_settings_iivm<-function()list(n_folds_tune=INNER_FOLDS,rsmp_tune=rsmp("cv",folds=INNER_FOLDS),measure=list(ml_g="regr.rmse",ml_m="classif.logloss",ml_r="classif.logloss"),terminator=trm("evals",n_evals=TUNE_EVALS),algorithm=tnr("random_search"))
tune_settings_pliv<-function()list(n_folds_tune=INNER_FOLDS,rsmp_tune=rsmp("cv",folds=INNER_FOLDS),measure=list(ml_l="regr.rmse",ml_m="regr.rmse",ml_r="regr.rmse"),terminator=trm("evals",n_evals=TUNE_EVALS),algorithm=tnr("random_search"))

get_original_xgb<-function(export,cell,learner_label){
  row<-export$weight_diagnostics%>%filter(.data$Cell==cell,.data$Learner==learner_label)
  if(nrow(row)!=1)stop(sprintf("Expected one diagnostic row for %s / %s.",cell,learner_label))
  tau<-as.numeric(row$Estimate[1])
  if(!is.finite(tau))stop(sprintf("Tuned estimate is unavailable for %s / %s.",cell,learner_label))
  tau}
fit_shifted_wald<-function(wf){dat<-make_dml_data(wf$Y_shift,wf$D,wf$Z,wf$X_dml);L<-make_xgb_iivm()
  o<-DoubleMLIIVM$new(data=dat,ml_g=L$g,ml_m=L$m,ml_r=L$r,n_folds=N_CF,score="LATE")
  invisible(capture.output(o$tune(param_set=tune_set_iivm(),tune_settings=tune_settings_iivm(),tune_on_folds=TRUE)))
  invisible(capture.output(o$fit(store_models=TRUE,store_predictions=TRUE)));get_dml_coef(o)}
fit_shifted_pliv<-function(wf){dat<-make_dml_data(wf$Y_shift,wf$D,wf$Z,wf$X_dml);L<-make_xgb_pliv()
  o<-DoubleMLPLIV$new(data=dat,ml_l=L$l,ml_m=L$m,ml_r=L$r,n_folds=N_CF)
  invisible(capture.output(o$tune(param_set=tune_set_pliv(),tune_settings=tune_settings_pliv(),tune_on_folds=TRUE)))
  invisible(capture.output(o$fit(store_models=TRUE,store_predictions=TRUE)));get_dml_coef(o)}
build_xgb_rows<-function(wf){out<-list()
  if(RUN_XGB_WALD_AIPW){orig<-get_original_xgb(xgb_wald_export,wf$label,"XGB tuned");set.seed(SEED);ts<-fit_shifted_wald(wf)
    r<-ti_vietnam_row("Wald-AIPW (XGBoost tuned)",orig,ts,wf$k);r$Specification<-wf$label;out$wald<-r}
  if(RUN_XGB_PLIV){orig<-get_original_xgb(xgb_pliv_export,wf$label,"PLR-IV XGB tuned");set.seed(SEED);ts<-fit_shifted_pliv(wf)
    r<-ti_vietnam_row("PLR-IV (XGBoost tuned)",orig,ts,wf$k);r$Specification<-wf$label;out$pliv<-r}
  do.call(rbind,out)}
```

```{r xgb-run, cache=TRUE, cache.extra=list(RUN_XGB_RERUN,RUN_XGB_WALD_AIPW,RUN_XGB_PLIV,TUNE_EVALS,INNER_FOLDS)}
if(RUN_XGB_RERUN){
  xgb_ti_all<-do.call(rbind,lapply(workflows,build_xgb_rows))
  saveRDS(list(results=xgb_ti_all,settings=list(seed=SEED,n_cf=N_CF,
    tune_evals=TUNE_EVALS,inner_folds=INNER_FOLDS,tune_on_folds=TRUE,
    original_exports=c("vietnam_xgb_tuning_export.rds","vietnam_xgb_pliv_tuning_export.rds"))),
    "vietnam_translation_xgb_rerun_export.rds")
}
```

```{r xgb-table}
if(RUN_XGB_RERUN) {
  xgb_ti_all <- classify_rerun_results(xgb_ti_all)
  display_rerun_results(xgb_ti_all) %>%
    kbl(caption="Tuned XGBoost — original export versus shifted nested-tuning rerun",digits=8,row.names=FALSE) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
    row_spec(which(!xgb_ti_all$invariant),background="#fffbf0") %>%
    footnote(general=paste0(
      "Original estimates are imported from the saved tuning results; shifted estimates rerun nested tuning. ",
      "Shading marks absolute rerun differences above ",
      format(RERUN_TOLERANCE,scientific=TRUE), "."
    ))
}
```

<div class="note-box">
**Interpretation.** The normalized kappa estimators and the OLS/logit, Ranger,
and honest-forest implementations are numerically unchanged after the complete
rerun. The cubic unnormalized kappa results reproduce the sensitivity reported
by SUW, while the kappa variants coincide under the saturated-age
specification. The tuned XGBoost estimates show small but visible rerun
differences. These changes are compatible with the absence of a general
normalization result for boosted-tree implementations discussed by Knaus.
Their magnitudes refer only to the present implementation and tuning protocol.
</div>

# 5. Save reusable rerun results

```{r save-complete-rerun, cache=FALSE, include=FALSE}
result_parts <- Filter(Negate(is.null), list(
  Kappa = if (exists("kappa_ti_all")) kappa_ti_all else NULL,
  `Tuned honest forest` = if (exists("dml_headline_ti_all")) dml_headline_ti_all else NULL,
  `OLS/logit and Ranger` = if (exists("learner_ti_all")) learner_ti_all else NULL,
  `Tuned XGBoost` = if (exists("xgb_ti_all")) xgb_ti_all else NULL
))
for (nm in names(result_parts)) result_parts[[nm]]$Family <- nm
translation_results <- bind_rows(result_parts)
saveRDS(list(
  results = translation_results,
  settings = list(seed=SEED,n_cf=N_CF,shift=log(100),
                  rerun_tolerance=RERUN_TOLERANCE,
                  grf_tuning="all",xgboost_tune_on_folds=TRUE,
                  xgboost_tune_evals=TUNE_EVALS,xgboost_inner_folds=INNER_FOLDS)
), "vietnam_translation_rerun_results.rds")
```
