---
title: "Angrist (1990) — Outcome Weights & Kappa Estimators"
subtitle: "Vietnam Draft Lottery: SUW Replication, DML & Learner Comparison"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body   { font-size: 15px; color: #2c2c2c; }
h1     { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2     { color: #2c7bb6; margin-top: 28px; }
h3     { color: #444; margin-top: 20px; }

.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 12,
  fig.height = 6
)
```

---

# Setup

## Packages & Functions

```{r packages-and-source}
# Core
library(haven)
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
# DML smoother + balance plots
library(OutcomeWeights)
library(cobalt)
library(viridis)
library(gridExtra)
library(ggplot2)
# DoubleML learners
library(DoubleML)
library(mlr3verse)
lgr::get_logger("mlr3")$set_threshold("warn")

source("functions_all.R")   # all estimators, weight diagnostics (incl. ESS_mod),
                            # get_dml_coef / get_dml_se, make_X, make_love, etc.
cat("functions_all.R sourced.\n")
```

## Data & Design

<div class="note-box">
**Setting.** Angrist (1990) SIPP 1984 subsample (N = 3,027 white men, born 1950–53). Instrument Z = `rsncode` (draft-eligibility indicator). Treatment D = `nvstat` (Vietnam-era veteran status). Outcome: log wages in two units (`lwage = log(kwage)` dollars; `lwage_cnt = log(100·kwage)` cents) — the units only enter the SUW reproduction table. Variable definitions follow the SUW Stata replication exactly: `age = age_5`, with three covariate specifications — **Linear** (`age`), **Cubic** (`age, age², age³`), and **Saturated** (`i.age` = age dummies).
</div>

```{r data}
sipp_path <- "sipp.dta"
stopifnot(file.exists(sipp_path))
sipp <- read_dta(sipp_path)

# Sample restriction: N = 3,027  (drop rsncode==999, missing educ / kwage)
sipp_clean <- sipp[sipp$rsncode != 999 &
                     !is.na(sipp$educ) &
                     !is.na(sipp$kwage), ]

# Age terms — EXACTLY as in suw_jbes_replicate.do: age = age_5
sipp_clean$age     <- sipp_clean$age_5
sipp_clean$age2    <- sipp_clean$age_5^2
sipp_clean$age3    <- sipp_clean$age_5^3
sipp_clean$age_fac <- factor(sipp_clean$age_5)

# Outcomes in two units (note: dollars vs. cents convention is FLIPPED vs. Card)
sipp_clean$lwage     <- log(sipp_clean$kwage)         # dollars
sipp_clean$lwage_cnt <- log(100 * sipp_clean$kwage)   # cents

# Core vectors
Y_cnt <- sipp_clean$lwage_cnt   # cents  (used only for the SUW reproduction layout)
Y_dol <- sipp_clean$lwage       # dollars (the working outcome everywhere else)
Z     <- as.integer(sipp_clean$rsncode)
D     <- as.integer(sipp_clean$nvstat)

cat(sprintf("N = %d (should be 3,027)\n", nrow(sipp_clean)))
cat(sprintf("mean(Z) = %.4f | mean(D) = %.4f\n", mean(Z), mean(D)))
```

```{r design-matrices}
# Kappa design matrices — WITH intercept (for logit / CBPS)
X_lin <- model.matrix(~ age,                data = sipp_clean)
X_cub <- model.matrix(~ age + age2 + age3,  data = sipp_clean)
X_sat <- model.matrix(~ age_fac,            data = sipp_clean)

# X_df for 2SLS (no intercept column)
X_lin_df <- data.frame(age = sipp_clean$age)
X_cub_df <- data.frame(age = sipp_clean$age, age2 = sipp_clean$age2, age3 = sipp_clean$age3)
X_sat_df <- data.frame(age_fac = sipp_clean$age_fac)

# DML design matrices — no intercept, attributes stripped for grf (make_X)
X_dml_cub      <- make_X(~ 0 + age + I(age^2) + I(age^3), data = sipp_clean)
X_dml_sat_full <- make_X(~ 0 + age_fac, data = sipp_clean)
X_dml_sat      <- X_dml_sat_full[, -1, drop = FALSE]   # drop one age dummy (reference)

cat(sprintf("Kappa: Linear %d / Cubic %d / Saturated %d covariates | DML: Cubic %dx%d / Saturated %dx%d\n",
            ncol(X_lin) - 1, ncol(X_cub) - 1, ncol(X_sat) - 1,
            nrow(X_dml_cub), ncol(X_dml_cub), nrow(X_dml_sat), ncol(X_dml_sat)))
```

---

# 1. Reproduction — SUW Table 2

<div class="note-box">
**Layout.** One treatment (veteran status), three age specifications, two outcome units (cents / dollars). Columns are spec × unit. Normalized estimators (τ_cb_u, τ_ml_u, τ_ml_a10) are translation invariant — cents and dollars match; unnormalized ones (τ_ml_a, τ_ml_t, τ_ml_a0) do not.
</div>

```{r build-columns, cache=TRUE}
# One full column = 2SLS + all six kappa estimators with analytical SEs.
run_viet_column <- function(Y, X_kappa, X_df) {
  list(
    tsls  = run_2sls(Y, D, Z, X_df, endog_name = "nvstat"),
    kappa = kappa_analytic_se_all(Y, Z, D, X_kappa)
  )
}

col_lin_cnt <- run_viet_column(Y_cnt, X_lin, X_lin_df)
col_lin_dol <- run_viet_column(Y_dol, X_lin, X_lin_df)
col_cub_cnt <- run_viet_column(Y_cnt, X_cub, X_cub_df)
col_cub_dol <- run_viet_column(Y_dol, X_cub, X_cub_df)
col_sat_cnt <- run_viet_column(Y_cnt, X_sat, X_sat_df)
col_sat_dol <- run_viet_column(Y_dol, X_sat, X_sat_df)

cat("All 6 columns computed.\n")
```

```{r table-builder, echo=FALSE}
cell_fmt <- function(est, se, digits = 3) {
  if (is.na(est)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0)
    return(sprintf("%.3f<br><span style='color:#888'>(NA)</span>", est))
  pval  <- 2 * pnorm(-abs(est / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s<br><span style='color:#888'>(%.3f)</span>", est, stars, se)
}

build_spec_table <- function(c_lin_cnt, c_lin_dol, c_cub_cnt, c_cub_dol, c_sat_cnt, c_sat_dol) {
  ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
  labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")

  rows <- list(data.frame(
    Estimator = "2SLS",
    `Linear / cents`      = cell_fmt(c_lin_cnt$tsls$coef, c_lin_cnt$tsls$se),
    `Linear / dollars`    = cell_fmt(c_lin_dol$tsls$coef, c_lin_dol$tsls$se),
    `Cubic / cents`       = cell_fmt(c_cub_cnt$tsls$coef, c_cub_cnt$tsls$se),
    `Cubic / dollars`     = cell_fmt(c_cub_dol$tsls$coef, c_cub_dol$tsls$se),
    `Saturated / cents`   = cell_fmt(c_sat_cnt$tsls$coef, c_sat_cnt$tsls$se),
    `Saturated / dollars` = cell_fmt(c_sat_dol$tsls$coef, c_sat_dol$tsls$se),
    check.names = FALSE, stringsAsFactors = FALSE
  ))

  for (i in seq_along(ests)) {
    e <- ests[i]
    rows[[length(rows) + 1]] <- data.frame(
      Estimator = labels[i],
      `Linear / cents`      = cell_fmt(c_lin_cnt$kappa$estimates[e], c_lin_cnt$kappa$se[e]),
      `Linear / dollars`    = cell_fmt(c_lin_dol$kappa$estimates[e], c_lin_dol$kappa$se[e]),
      `Cubic / cents`       = cell_fmt(c_cub_cnt$kappa$estimates[e], c_cub_cnt$kappa$se[e]),
      `Cubic / dollars`     = cell_fmt(c_cub_dol$kappa$estimates[e], c_cub_dol$kappa$se[e]),
      `Saturated / cents`   = cell_fmt(c_sat_cnt$kappa$estimates[e], c_sat_cnt$kappa$se[e]),
      `Saturated / dollars` = cell_fmt(c_sat_dol$kappa$estimates[e], c_sat_dol$kappa$se[e]),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }
  do.call(rbind, rows)
}
```

## Results

```{r table-reproduction}
build_spec_table(col_lin_cnt, col_lin_dol, col_cub_cnt, col_cub_dol, col_sat_cnt, col_sat_dol) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c","c","c"),
      caption = "SUW Table 2 — Treatment: Vietnam veteran status (nvstat). Effect of military service on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12.5) %>%
  add_header_above(c(" " = 1, "Linear age" = 2, "Cubic age" = 2, "Saturated age" = 2)) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",            1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE below). Blue = normalized (translation invariant). Orange = unnormalized. age = age_5. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 2. Double ML — PLR-IV & Wald-AIPW (honest forest, tuned vs. untuned)

<div class="note-box">
**Setting.** DML uses `dml_with_smoother()` with five-fold cross-fitting on
the **cubic-age** and **saturated-age** specifications. For each specification,
the same call jointly returns PLR-IV and Wald-AIPW. We estimate an untuned
honest forest (`tune.parameters = "none"`) and a tuned honest forest
(`tune.parameters = "all"`). The latter invokes GRF's automatic,
out-of-bag-loss-based tuning; it is distinct from the learner comparison in
Section 3. The linear-age specification remains in the kappa/2SLS reproduction
but is too low-dimensional to provide a meaningful honest-forest comparison.
`set.seed(42)` is reset before every fit so that the four calls begin from the
same random-number state, although tuning itself consumes additional random
draws.
</div>

```{r dml-fit-all, cache=TRUE}
# Cubic-age specification.
set.seed(42); dml_cub_untuned <- dml_with_smoother(
  Y_dol, D, X_dml_cub, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_cub_tuned <- dml_with_smoother(
  Y_dol, D, X_dml_cub, Z, n_cf_folds = 5, tune.parameters = "all")

# Saturated-age specification.
set.seed(42); dml_sat_untuned <- dml_with_smoother(
  Y_dol, D, X_dml_sat, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_sat_tuned <- dml_with_smoother(
  Y_dol, D, X_dml_sat, Z, n_cf_folds = 5, tune.parameters = "all")

res_cub_untuned <- summary(dml_cub_untuned)
res_cub_tuned   <- summary(dml_cub_tuned)
res_sat_untuned <- summary(dml_sat_untuned)
res_sat_tuned   <- summary(dml_sat_tuned)

extract_w <- function(dml_obj) {
  om <- get_outcome_weights(dml_obj)
  check_omega_rows(om)
  list(plriv    = as.vector(om$omega["PLR-IV",    ]),
       waldaipw = as.vector(om$omega["Wald-AIPW", ]))
}
w_cub_untuned <- extract_w(dml_cub_untuned)
w_cub_tuned   <- extract_w(dml_cub_tuned)
w_sat_untuned <- extract_w(dml_sat_untuned)
w_sat_tuned   <- extract_w(dml_sat_tuned)

cat("All four DML fits (2 age specifications x untuned/tuned) complete.\n")
```

## Estimates {.tabset}

```{r dml-cell-display, echo=FALSE}
show_dml_cell <- function(res, wlist, Y, cap) {
  ids <- c(
    check_weight_identity(wlist$plriv,    Y, get_estimate(res, "PLR-IV")),
    check_weight_identity(wlist$waldaipw, Y, get_estimate(res, "Wald-AIPW"))
  )
  df <- as.data.frame(res[c("PLR-IV", "Wald-AIPW"), , drop = FALSE])
  df$Estimator  <- rownames(df)
  df$`ΣωY = τ̂` <- as.character(ids)
  df <- df[, c("Estimator", setdiff(names(df), c("Estimator","ΣωY = τ̂")), "ΣωY = τ̂")]
  df %>%
    kbl(caption = cap, digits = 4, row.names = FALSE,
        align = c("l", rep("r", ncol(df) - 2), "c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "ΣωY = τ̂: algebraic identity check (must be TRUE).")
}
```

### Cubic age

```{r dml-cub-untuned}
show_dml_cell(res_cub_untuned, w_cub_untuned, Y_dol,
              "Cubic age — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-cub-tuned}
show_dml_cell(res_cub_tuned, w_cub_tuned, Y_dol,
              "Cubic age — tuned honest forest (tune.parameters = \"all\")")
```

### Saturated age

```{r dml-sat-untuned}
show_dml_cell(res_sat_untuned, w_sat_untuned, Y_dol,
              "Saturated age — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-sat-tuned}
show_dml_cell(res_sat_tuned, w_sat_tuned, Y_dol,
              "Saturated age — tuned honest forest (tune.parameters = \"all\")")
```

<div class="note-box">
**How to read the comparison.** Within each age specification, compare the
untuned and tuned rows for changes in the point estimate, standard error, and
later weight geometry. The identity check must remain true: it verifies the
package's algebraic outcome-weight representation and is not a tuning-quality
criterion.
</div>

## DML vs. Kappa — Point Estimates

```{r dml-summary-table}
# Kappa estimates (dollars) for both specs — reused later by the learner table.
kappa_cub <- kappa_analytic_se_all(Y_dol, Z, D, X_cub)
kappa_sat <- kappa_analytic_se_all(Y_dol, Z, D, X_sat)

cell <- function(coef, se) {
  if (is.na(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

dml_cell <- function(res, est) {
  coef_val <- get_estimate(res, est)
  se_val   <- if ("SE" %in% colnames(res)) as.numeric(res[est, "SE"])
              else if ("Std. Error" %in% colnames(res)) as.numeric(res[est, "Std. Error"])
              else NA_real_
  cell(coef_val, se_val)
}

dml_rows <- rbind(
  data.frame(Estimator = "PLR-IV (untuned forest)",
             `Cubic age`     = dml_cell(res_cub_untuned, "PLR-IV"),
             `Saturated age` = dml_cell(res_sat_untuned, "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (untuned forest)",
             `Cubic age`     = dml_cell(res_cub_untuned, "Wald-AIPW"),
             `Saturated age` = dml_cell(res_sat_untuned, "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "PLR-IV (tuned forest)",
             `Cubic age`     = dml_cell(res_cub_tuned, "PLR-IV"),
             `Saturated age` = dml_cell(res_sat_tuned, "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (tuned forest)",
             `Cubic age`     = dml_cell(res_cub_tuned, "Wald-AIPW"),
             `Saturated age` = dml_cell(res_sat_tuned, "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE)
)

ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")
kappa_rows <- do.call(rbind, lapply(seq_along(ests), function(i) {
  e <- ests[i]
  data.frame(
    Estimator = labels[i],
    `Cubic age`     = cell(kappa_cub$estimates[e], kappa_cub$se[e]),
    `Saturated age` = cell(kappa_sat$estimates[e], kappa_sat$se[e]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}))

rbind(dml_rows, kappa_rows) %>%
  kbl(align = c("l","c","c"),
      caption = "DML vs. Kappa — Point Estimates (SE), Y = log wages (dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Double ML — honest forest", 1, 4) %>%
  pack_rows("Kappa — normalized", 5, 7) %>%
  pack_rows("Kappa — unnormalized", 8, 10) %>%
  row_spec(1:4, background = "#eef6ff") %>%
  row_spec(8:10, background = "#fffbf0") %>%
  footnote(general = "Each cell: estimate (SE). Blue = DML. Orange = unnormalized kappa. Stars: * p<0.10, ** p<0.05, *** p<0.01.")
```

---

# 3. Learner Comparison — PLR-IV and Wald-AIPW

<div class="note-box">
**Purpose.** Section 2 varies GRF's internal tuning while holding the honest-
forest learner family fixed. This section instead holds the five-fold
DoubleML procedure fixed and varies the nuisance learners: parametric
linear/logistic models, Ranger, and XGBoost. Each learner family is used for
both estimators. PLR-IV is implemented with `DoubleMLPLIV`; Wald-AIPW/LATE is
implemented with `DoubleMLIIVM`. All models use the same cubic-age and
saturated-age designs defined above. Linear and Ranger are fitted in this
notebook. XGBoost is imported from `vietnam_xgb_tuning_export.rds` and
`vietnam_xgb_pliv_tuning_export.rds`, which preserve both the fixed untuned
baseline and the `tune_on_folds = TRUE` estimator without rerunning expensive
nested tuning here.

**Terminology.** “Linear learner” refers to the nuisance-model class: OLS for
PLR-IV and OLS/logit for Wald-AIPW. It does **not** refer to the linear-age
feature specification in Section 1. The learner comparison uses either the
cubic-age feature matrix or the saturated-age feature matrix throughout.
</div>

```{r ext-setup}
make_dml_data <- function(X_dml) {
  df <- as.data.frame(X_dml)
  df$y <- as.numeric(Y_dol)
  df$d <- as.numeric(D)
  df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X_dml))
}

dat_cub <- make_dml_data(X_dml_cub)
dat_sat <- make_dml_data(X_dml_sat)

cat("DoubleML data objects built for both specs.\n")
```

```{r ext-fit-helpers, echo=FALSE}
ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))
ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))
omega_vec <- function(om, estimator) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && estimator %in% rownames(w))
    return(as.vector(w[estimator, ]))
  as.vector(w[1, ])
}

# Fit one IIVM model. Wrapped in tryCatch: the parametric OLS/logit learner
# can fail to produce outcome weights in the saturated spec (documented in the
# thesis); on failure we return NA gracefully instead of aborting the knit.
fit_iivm <- function(dml_data, learners, Y) {
  tryCatch({
    set.seed(42)
    obj <- DoubleMLIIVM$new(data = dml_data,
                            ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                            n_folds = 5, score = "LATE")
    obj$fit(store_models = TRUE, store_predictions = TRUE)
    om  <- get_outcome_weights(object = obj, dml_data = dml_data)
    w   <- omega_vec(om, "Wald-AIPW")
    ok  <- check_doubleml_identity(om, Y, obj$coef)
    list(obj = obj, w = w, ok = ok)
  }, error = function(e) {
    message("fit_iivm failed: ", conditionMessage(e))
    list(obj = NULL, w = rep(NA_real_, length(Y)), ok = NA)
  })
}

# PLR-IV uses three regression nuisances: E[Y|X], E[Z|X], and E[D|X].
fit_pliv <- function(dml_data, learners, Y) {
  tryCatch({
    set.seed(42)
    obj <- DoubleMLPLIV$new(data = dml_data,
                            ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
                            n_folds = 5)
    obj$fit(store_models = TRUE, store_predictions = TRUE)
    om <- get_outcome_weights(object = obj, dml_data = dml_data)
    w  <- omega_vec(om, "PLR-IV")
    ok <- check_weight_identity(w, Y, get_dml_coef(obj))
    list(obj = obj, w = w, ok = ok)
  }, error = function(e) {
    message("fit_pliv failed: ", conditionMessage(e))
    list(obj = NULL, w = rep(NA_real_, length(Y)), ok = NA)
  })
}

safe_coef <- function(f) if (is.null(f$obj)) NA_real_ else get_dml_coef(f$obj)
safe_se   <- function(f) if (is.null(f$obj)) NA_real_ else get_dml_se(f$obj)

# Compact two-estimator, two-specification table for one learner family.
learner_pair_table <- function(pliv_cub, pliv_sat, iivm_cub, iivm_sat, cap) {
  data.frame(
    Estimator = c("PLR-IV", "Wald-AIPW / LATE"),
    `Cubic estimate` = round(c(safe_coef(pliv_cub), safe_coef(iivm_cub)), 4),
    `Cubic SE`       = round(c(safe_se(pliv_cub),   safe_se(iivm_cub)),   4),
    `Cubic identity` = c(pliv_cub$ok, iivm_cub$ok),
    `Saturated estimate` = round(c(safe_coef(pliv_sat), safe_coef(iivm_sat)), 4),
    `Saturated SE`       = round(c(safe_se(pliv_sat),   safe_se(iivm_sat)),   4),
    `Saturated identity` = c(pliv_sat$ok, iivm_sat$ok),
    check.names = FALSE
  ) %>%
    kbl(caption = cap, align = c("l","r","r","c","r","r","c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    add_header_above(c(" " = 1, "Cubic age" = 3, "Saturated age" = 3)) %>%
    footnote(general = "PLR-IV uses DoubleMLPLIV; Wald-AIPW/LATE uses DoubleMLIIVM. TRUE means the extracted outcome weights reproduce the reported coefficient. FALSE weights are approximate learner-sensitivity diagnostics and are not treated as exact representations.")
}
```

## Estimates {.tabset}

### OLS / Logistic nuisance learners

```{r ext-fit-lm, cache=TRUE}
lm_iivm_cub <- fit_iivm(dat_cub, ml_lm_iivm(), Y_dol)
lm_iivm_sat <- fit_iivm(dat_sat, ml_lm_iivm(), Y_dol)
lm_pliv_cub <- fit_pliv(dat_cub, ml_lm_pliv(), Y_dol)
lm_pliv_sat <- fit_pliv(dat_sat, ml_lm_pliv(), Y_dol)

learner_pair_table(lm_pliv_cub, lm_pliv_sat, lm_iivm_cub, lm_iivm_sat,
                   "Parametric learners — Linear PLR-IV and Linear/Logistic Wald-AIPW")
```

### Ranger

```{r ext-fit-rf, cache=TRUE}
rf_iivm_cub <- fit_iivm(dat_cub, ml_rf_iivm(), Y_dol)
rf_iivm_sat <- fit_iivm(dat_sat, ml_rf_iivm(), Y_dol)
rf_pliv_cub <- fit_pliv(dat_cub, ml_rf_pliv(), Y_dol)
rf_pliv_sat <- fit_pliv(dat_sat, ml_rf_pliv(), Y_dol)

learner_pair_table(rf_pliv_cub, rf_pliv_sat, rf_iivm_cub, rf_iivm_sat,
                   "Ranger learners — PLR-IV and Wald-AIPW/LATE")
```

### XGBoost Wald-AIPW (imported)

```{r ext-xgb-wald-import}
validate_xgb_export <- function(x,expected_class) {
  required <- c("estimate_comparison","nuisance_losses","weight_diagnostics",
                "omega_vectors","identity_checks","settings","sample_info")
  if (!all(required %in% names(x))) stop("XGBoost export has an incomplete schema.")
  if (!identical(x$settings$dml_class,expected_class)) stop("XGBoost export has the wrong DoubleML class.")
  if (!isTRUE(x$settings$tune_on_folds)) stop("XGBoost export is not tune_on_folds=TRUE.")
  if (!all(c("Cubic age","Saturated age") %in% names(x$omega_vectors)))
    stop("XGBoost export does not contain both Vietnam age specifications.")
  invisible(TRUE)
}
xgb_wald_path <- "vietnam_xgb_tuning_export.rds"
xgb_wald_available <- file.exists(xgb_wald_path)
if (xgb_wald_available) {
  xgb_wald_export <- readRDS(xgb_wald_path)
  validate_xgb_export(xgb_wald_export,"DoubleMLIIVM")
  xgb_wald_export$estimate_comparison %>%
    kbl(caption="Wald-AIPW XGBoost — untuned vs. tune_on_folds=TRUE (imported)",
        align=c("l","l","c","r")) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),
                  full_width=TRUE,font_size=13) %>%
    collapse_rows(columns=1,valign="middle") %>%
    footnote(general="Imported from vietnam_xgb_tuning_export.rds. The tuned row preserves fold-specific hyperparameters selected inside each outer training fold; no XGBoost model is refitted here.")
} else {
  cat("vietnam_xgb_tuning_export.rds not found — run vietnam_xgb_tuning.Rmd first.\n")
}
```

### XGBoost PLR-IV (imported)

```{r ext-xgb-pliv-import}
xgb_pliv_path <- "vietnam_xgb_pliv_tuning_export.rds"
xgb_pliv_available <- file.exists(xgb_pliv_path)
if (xgb_pliv_available) {
  xgb_pliv_export <- readRDS(xgb_pliv_path)
  validate_xgb_export(xgb_pliv_export,"DoubleMLPLIV")
  xgb_pliv_export$estimate_comparison %>%
    kbl(caption="PLR-IV XGBoost — untuned vs. tune_on_folds=TRUE (imported)",
        align=c("l","l","c")) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),
                  full_width=TRUE,font_size=13) %>%
    collapse_rows(columns=1,valign="middle") %>%
    footnote(general="Imported from vietnam_xgb_pliv_tuning_export.rds. Untuned and tuned estimates, weights, and identity flags refer to the exact exported fits.")
} else {
  cat("vietnam_xgb_pliv_tuning_export.rds not found — run vietnam_xgb_pliv_tuning.Rmd first.\n")
}
```

```{r ext-xgb-export-helpers, echo=FALSE}
empty_xgb_bundle <- function() list(w=NULL,ok=NA)
get_xgb_bundle <- function(export,cell,variant) {
  if (is.null(export)) return(empty_xgb_bundle())
  w <- tryCatch(export$omega_vectors[[cell]][[variant]],error=function(e)NULL)
  ok <- tryCatch(export$identity_checks[[cell]][[variant]],error=function(e)NA)
  if (is.null(ok)) ok <- NA
  list(w=w,ok=ok)
}
get_xgb_estimate_cell <- function(export,cell,variant) {
  if (is.null(export)) return("—")
  df <- export$estimate_comparison
  keep_variant <- if (variant=="untuned") grepl("untuned$",df$Learner) else grepl("tune_on_folds=TRUE",df$Learner,fixed=TRUE)
  hit <- df[df$Cell==cell & keep_variant,,drop=FALSE]
  if (!nrow(hit)) "—" else as.character(hit[["Estimate (SE)"]][1])
}

xgb_wald_obj <- if (exists("xgb_wald_export")) xgb_wald_export else NULL
xgb_pliv_obj <- if (exists("xgb_pliv_export")) xgb_pliv_export else NULL

xgb_iivm_cub_untuned <- get_xgb_bundle(xgb_wald_obj,"Cubic age","untuned")
xgb_iivm_cub_tuned   <- get_xgb_bundle(xgb_wald_obj,"Cubic age","tuned")
xgb_iivm_sat_untuned <- get_xgb_bundle(xgb_wald_obj,"Saturated age","untuned")
xgb_iivm_sat_tuned   <- get_xgb_bundle(xgb_wald_obj,"Saturated age","tuned")
xgb_pliv_cub_untuned <- get_xgb_bundle(xgb_pliv_obj,"Cubic age","untuned")
xgb_pliv_cub_tuned   <- get_xgb_bundle(xgb_pliv_obj,"Cubic age","tuned")
xgb_pliv_sat_untuned <- get_xgb_bundle(xgb_pliv_obj,"Saturated age","untuned")
xgb_pliv_sat_tuned   <- get_xgb_bundle(xgb_pliv_obj,"Saturated age","tuned")
```

## Comparison Table

```{r ext-comparison}
ext_cell <- function(coef, se) {
  if (is.na(coef) || !is.finite(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  pval  <- 2 * pnorm(-abs(coef / se))
  stars <- ifelse(pval < 0.01, "***", ifelse(pval < 0.05, "**",
                  ifelse(pval < 0.10, "*", "")))
  sprintf("%.3f%s (%.3f)", coef, stars, se)
}

learner_row <- function(label, f_cub, f_sat) {
  data.frame(
    Estimator = label,
    `Cubic age`     = ext_cell(safe_coef(f_cub), safe_se(f_cub)),
    `Saturated age` = ext_cell(safe_coef(f_sat), safe_se(f_sat)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

kappa_ref_row <- function(label, key) {
  data.frame(
    Estimator = label,
    `Cubic age`     = ext_cell(kappa_cub$estimates[key], kappa_cub$se[key]),
    `Saturated age` = ext_cell(kappa_sat$estimates[key], kappa_sat$se[key]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

comparison_rows <- list(
  learner_row("PLR-IV (OLS nuisances)",                  lm_pliv_cub,  lm_pliv_sat),
  learner_row("Wald-AIPW / LATE (OLS/logit nuisances)", lm_iivm_cub, lm_iivm_sat),
  learner_row("PLR-IV (ranger)",                  rf_pliv_cub,  rf_pliv_sat),
  learner_row("Wald-AIPW / LATE (ranger)",       rf_iivm_cub, rf_iivm_sat),
  kappa_ref_row("τ_cb_u (kappa, ref)",  "tau_cb_u"),
  kappa_ref_row("τ_ml_u (kappa, ref)",  "tau_ml_u")
)
xgb_rows <- list(
  data.frame(Estimator="PLR-IV (XGBoost, untuned)",
    `Cubic age`=get_xgb_estimate_cell(xgb_pliv_obj,"Cubic age","untuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_pliv_obj,"Saturated age","untuned"),check.names=FALSE),
  data.frame(Estimator="PLR-IV (XGBoost, tune_on_folds=TRUE)",
    `Cubic age`=get_xgb_estimate_cell(xgb_pliv_obj,"Cubic age","tuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_pliv_obj,"Saturated age","tuned"),check.names=FALSE),
  data.frame(Estimator="Wald-AIPW / LATE (XGBoost, untuned)",
    `Cubic age`=get_xgb_estimate_cell(xgb_wald_obj,"Cubic age","untuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_wald_obj,"Saturated age","untuned"),check.names=FALSE),
  data.frame(Estimator="Wald-AIPW / LATE (XGBoost, tune_on_folds=TRUE)",
    `Cubic age`=get_xgb_estimate_cell(xgb_wald_obj,"Cubic age","tuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_wald_obj,"Saturated age","tuned"),check.names=FALSE)
)
comp_df <- do.call(rbind,c(comparison_rows[1:4],xgb_rows,comparison_rows[5:6]))

comp_df %>%
  kbl(align = c("l","c","c"),
      caption = "PLR-IV and Wald-AIPW Learner Comparison — DoubleML vs. Kappa (Y = log wages, dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("DoubleML learners", 1, 8) %>%
  pack_rows("Kappa reference",   9, 10) %>%
  row_spec(1:8, background = "#eef6ff") %>%
  footnote(general = "Each cell: estimate (SE). Linear means OLS nuisances for PLR-IV and OLS/logit nuisances for Wald-AIPW; it does not mean the linear-age feature specification. XGBoost rows are imported from the separate nested-tuning exports. '—' means the required export or fit is unavailable.")
```

---

# 4. Weight Diagnostics

<div class="note-box">
**Reading the tables.** `Sum_w ≈ 0` confirms translation invariance.
**`ESS_mod`** = modified effective sample size `(Σ|ωᵢ|)²/Σωᵢ²`
(Chattopadhyay & Zubizarreta 2023), valid under negative weights and bounded
in `[1,n]`. The old Kish `ESS = 1/Σωᵢ²` is retained only as a reference.
The first pair of tabs compares untuned and tuned honest forests. The second
pair compares PLR-IV and Wald-AIPW across learner families. A learner weight
vector enters these exact diagnostics only when `ΣωᵢYᵢ = τ̂` passes; failed
XGBoost identities remain reported in Section 3 as approximate sensitivity
results but are not compared on equal footing here.
</div>

```{r diag-builders, echo=FALSE}
# kappa weight bundles for both specs.
kw_cub <- kappa_weights_bundle(Z, D, X_cub)
kw_sat <- kappa_weights_bundle(Z, D, X_sat)

# weight_diag() returns 6 columns (… ESS, ESS_mod …): align has 5 "r".
DIAG_ALIGN <- c("l","r","r","r","r","r")

xgb_status <- data.frame(
  Estimator=rep(c("PLR-IV","Wald-AIPW"),each=4),
  Specification=rep(rep(c("Cubic age","Saturated age"),each=2),2),
  Tuning=rep(c("untuned","tune_on_folds=TRUE"),4),
  `Weights available`=c(
    !is.null(xgb_pliv_cub_untuned$w),!is.null(xgb_pliv_cub_tuned$w),
    !is.null(xgb_pliv_sat_untuned$w),!is.null(xgb_pliv_sat_tuned$w),
    !is.null(xgb_iivm_cub_untuned$w),!is.null(xgb_iivm_cub_tuned$w),
    !is.null(xgb_iivm_sat_untuned$w),!is.null(xgb_iivm_sat_tuned$w)),
  `Identity passed`=c(
    xgb_pliv_cub_untuned$ok,xgb_pliv_cub_tuned$ok,
    xgb_pliv_sat_untuned$ok,xgb_pliv_sat_tuned$ok,
    xgb_iivm_cub_untuned$ok,xgb_iivm_cub_tuned$ok,
    xgb_iivm_sat_untuned$ok,xgb_iivm_sat_tuned$ok),
  check.names=FALSE)
xgb_status %>%
  kbl(caption="Imported XGBoost weight availability and identity status",
      align=c("l","l","l","c","c")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE,font_size=12) %>%
  footnote(general="Weights enter exact diagnostics, bridges, and Love plots only when available and Identity passed is TRUE. FALSE identities remain estimate-level learner-sensitivity results.")

dml_diag_cell <- function(w_untuned, w_tuned, kw) {
  rbind(
    weight_diag(w_untuned$plriv,    "PLR-IV (untuned forest)"),
    weight_diag(w_untuned$waldaipw, "Wald-AIPW (untuned forest)"),
    weight_diag(w_tuned$plriv,      "PLR-IV (tuned forest)"),
    weight_diag(w_tuned$waldaipw,   "Wald-AIPW (tuned forest)"),
    weight_diag(kw$tau_cb_u,    "τ_cb_u     (kappa)"),
    weight_diag(kw$tau_ml_u,    "τ_ml_u     (kappa)"),
    weight_diag(kw$tau_ml_a10,  "τ_ml_a10   (kappa)"),
    weight_diag(kw$tau_ml_a,    "τ_ml_a     (kappa)"),
    weight_diag(kw$tau_ml_a1,   "τ_ml_t/a1  (kappa)"),
    weight_diag(kw$tau_ml_a0,   "τ_ml_a0    (kappa)")
  )
}
style_dml_diag <- function(df, cap) {
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(1:4, background = "#eef6ff") %>%
    footnote(general = "Blue = honest-forest DML: untuned first, tuned second. Sum_w ≈ 0 means translation invariant. ESS_mod is the signed-weight effective sample size to report; ESS is reference only.")
}

exact_w <- function(fit) {
  if (isTRUE(fit$ok) && !is.null(fit$w)) fit$w else rep(NA_real_, length(Y_dol))
}

ext_diag_block <- function(kw, pliv_lm, iivm_lm, pliv_rf, iivm_rf,
                           pliv_xgb_u, pliv_xgb_t, iivm_xgb_u, iivm_xgb_t) {
  rbind(
    weight_diag(exact_w(pliv_lm),   "PLR-IV (OLS nuisances)"),
    weight_diag(exact_w(iivm_lm),   "Wald-AIPW (OLS/logit nuisances)"),
    weight_diag(exact_w(pliv_rf),   "PLR-IV (ranger)"),
    weight_diag(exact_w(iivm_rf),   "Wald-AIPW (ranger)"),
    weight_diag(exact_w(pliv_xgb_u),"PLR-IV (XGBoost, untuned)"),
    weight_diag(exact_w(pliv_xgb_t),"PLR-IV (XGBoost, tune_on_folds=TRUE)"),
    weight_diag(exact_w(iivm_xgb_u),"Wald-AIPW (XGBoost, untuned)"),
    weight_diag(exact_w(iivm_xgb_t),"Wald-AIPW (XGBoost, tune_on_folds=TRUE)"),
    weight_diag(kw$tau_cb_u,   "τ_cb_u    (kappa, CBPS)"),
    weight_diag(kw$tau_ml_u,   "τ_ml_u    (kappa, MLE)"),
    weight_diag(kw$tau_ml_a10, "τ_ml_a10  (kappa, MLE)"),
    weight_diag(kw$tau_ml_a,   "τ_ml_a    (kappa, unnorm)"),
    weight_diag(kw$tau_ml_a1,  "τ_ml_t    (kappa, unnorm)"),
    weight_diag(kw$tau_ml_a0,  "τ_ml_a0   (kappa, unnorm)")
  )
}
style_ext_diag <- function(df, cap) {
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(1:8, background = "#eef6ff") %>%
    footnote(general = "Blue = DoubleML learners, paired as PLR-IV and Wald-AIPW within each learner family. NA means the fit failed or its extracted weights did not reproduce the estimate exactly; such weights are not interpreted as exact outcome-weight representations.")
}
```

## DML vs. Kappa {.tabset}

### Cubic age

```{r dml-diag-cub}
style_dml_diag(dml_diag_cell(w_cub_untuned, w_cub_tuned, kw_cub),
               "Weight Diagnostics — Untuned/Tuned Honest Forest vs. Kappa | Cubic age")
```

### Saturated age

```{r dml-diag-sat}
style_dml_diag(dml_diag_cell(w_sat_untuned, w_sat_tuned, kw_sat),
               "Weight Diagnostics — Untuned/Tuned Honest Forest vs. Kappa | Saturated age")
```

## Learners vs. Kappa {.tabset}

### Cubic age

```{r ext-diag-cub}
style_ext_diag(ext_diag_block(kw_cub,
                              lm_pliv_cub, lm_iivm_cub,
                              rf_pliv_cub, rf_iivm_cub,
                              xgb_pliv_cub_untuned, xgb_pliv_cub_tuned,
                              xgb_iivm_cub_untuned, xgb_iivm_cub_tuned),
               "Weight Diagnostics — PLR-IV and Wald-AIPW Learners | Cubic age")
```

### Saturated age

```{r ext-diag-sat}
style_ext_diag(ext_diag_block(kw_sat,
                              lm_pliv_sat, lm_iivm_sat,
                              rf_pliv_sat, rf_iivm_sat,
                              xgb_pliv_sat_untuned, xgb_pliv_sat_tuned,
                              xgb_iivm_sat_untuned, xgb_iivm_sat_tuned),
               "Weight Diagnostics — PLR-IV and Wald-AIPW Learners | Saturated age")
```

## Weight-Statistics Bridge — RQ1 {.tabset}

<div class="note-box">
**What this is.** For every estimator, `weight_bridge()` reports the outcome weights **Overall / Treated / Control** (control sign-flipped, matching Knaus's convention), putting the two diagnostic vocabularies side by side: his six `summary()` statistics next to the two effective-sample-size measures (Kish vs. modified). The Knaus block is computed in plain R, so the kappa estimators get his statistics too.
</div>

```{r bridge-helpers, echo=FALSE}
# Named list of weight vectors per cell; drop any all-NA (failed) learner.
cell_weight_list <- function(w_untuned, w_tuned,
                             pliv_lm, iivm_lm, pliv_rf, iivm_rf,
                             pliv_xgb_u, pliv_xgb_t,
                             iivm_xgb_u, iivm_xgb_t, kw) {
  lst <- list(
    "PLR-IV (untuned forest)"    = w_untuned$plriv,
    "Wald-AIPW (untuned forest)" = w_untuned$waldaipw,
    "PLR-IV (tuned forest)"      = w_tuned$plriv,
    "Wald-AIPW (tuned forest)"   = w_tuned$waldaipw,
    "PLR-IV (OLS nuisances)"            = exact_w(pliv_lm),
    "Wald-AIPW (OLS/logit nuisances)"   = exact_w(iivm_lm),
    "PLR-IV (ranger)"            = exact_w(pliv_rf),
    "Wald-AIPW (ranger)"         = exact_w(iivm_rf),
    "PLR-IV (XGBoost, untuned)"  = exact_w(pliv_xgb_u),
    "PLR-IV (XGBoost, tuned)"    = exact_w(pliv_xgb_t),
    "Wald-AIPW (XGBoost, untuned)" = exact_w(iivm_xgb_u),
    "Wald-AIPW (XGBoost, tuned)" = exact_w(iivm_xgb_t),
    "τ_cb_u (kappa)"   = kw$tau_cb_u,
    "τ_ml_u (kappa)"   = kw$tau_ml_u,
    "τ_ml_t (kappa)"   = kw$tau_ml_a1
  )
  Filter(function(w) !all(is.na(w)), lst)
}

bridge_cells <- list(
  "Cubic age" = list(w = cell_weight_list(
    w_cub_untuned, w_cub_tuned,
    lm_pliv_cub, lm_iivm_cub, rf_pliv_cub, rf_iivm_cub,
    xgb_pliv_cub_untuned, xgb_pliv_cub_tuned,
    xgb_iivm_cub_untuned, xgb_iivm_cub_tuned, kw_cub)),
  "Saturated age" = list(w = cell_weight_list(
    w_sat_untuned, w_sat_tuned,
    lm_pliv_sat, lm_iivm_sat, rf_pliv_sat, rf_iivm_sat,
    xgb_pliv_sat_untuned, xgb_pliv_sat_tuned,
    xgb_iivm_sat_untuned, xgb_iivm_sat_tuned, kw_sat))
)

bridge_table <- function(wlist, cap) {
  ests <- names(wlist)
  df   <- do.call(rbind, lapply(ests, function(nm) weight_bridge(wlist[[nm]], D, nm)))
  disp <- df[, c("Scope","Min","Max","Pct_neg","Sum_top10","Sum_w","Sum_abs_w","ESS_kish","ESS_mod")]

  kk <- disp %>%
    kbl(caption = cap, row.names = FALSE,
        col.names = c("Scope","Min","Max","%Neg","ΣTop10%","Σω","Σ|ω|","ESS Kish","ESS mod"),
        align = c("l", rep("r", 8))) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 12) %>%
    add_header_above(c(" " = 1, "Knaus summary() statistics" = 6,
                       "Effective sample size" = 2)) %>%
    column_spec(9, bold = TRUE, background = "#eef6ff")
  for (i in seq_along(ests)) kk <- pack_rows(kk, ests[i], (i - 1) * 3 + 1, i * 3)
  kk %>% footnote(general = "Control rows use Knaus's sign-flipped −ω convention. ESS mod = (Σ|ω|)²/Σω² is the number to report; ESS Kish = 1/Σω² is reference only.")
}

ess_overview_df <- do.call(rbind, lapply(names(bridge_cells), function(cn) {
  cl <- bridge_cells[[cn]]
  do.call(rbind, lapply(names(cl$w), function(nm) {
    ov <- weight_bridge(cl$w[[nm]], D, nm)
    ov <- ov[ov$Scope == "Overall", ]
    data.frame(Cell = cn, Estimator = nm,
               ESS_kish = ov$ESS_kish, ESS_mod = ov$ESS_mod,
               check.names = FALSE, stringsAsFactors = FALSE)
  }))
}))
```

### ESS overview (Kish vs. modified)

```{r bridge-ess-overview}
ess_overview_df %>%
  kbl(caption = "Effective sample size, Overall scope — Kish (1/Σω²) vs. modified ((Σ|ω|)²/Σω²)",
      col.names = c("Cell","Estimator","ESS Kish","ESS mod"),
      align = c("l","l","r","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  column_spec(4, bold = TRUE, background = "#eef6ff") %>%
  pack_rows(index = table(factor(ess_overview_df$Cell, levels = unique(ess_overview_df$Cell)))) %>%
  footnote(general = "Where Kish collapses toward 1, the modified ESS recovers the genuine usable weight mass. Report the modified column.")
```

```{r bridge-ess-dumbbell, fig.height=5.5, fig.width=11}
ggplot(ess_overview_df, aes(y = Estimator)) +
  geom_segment(aes(x = ESS_kish, xend = ESS_mod, yend = Estimator),
               colour = "#bdbdbd", linewidth = 1.1) +
  geom_point(aes(x = ESS_kish, colour = "Kish"),     size = 2.6) +
  geom_point(aes(x = ESS_mod,  colour = "Modified"), size = 2.6) +
  facet_wrap(~ Cell, scales = "free_x") +
  scale_colour_manual(values = c("Kish" = "#9e9e9e", "Modified" = "#2c7bb6"), name = NULL) +
  labs(title = "Effective sample size: Kish → modified",
       x = "Effective sample size", y = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.minor = element_blank())
```

### Cubic age

```{r bridge-cub}
bridge_table(bridge_cells[["Cubic age"]]$w, "Weight-statistics bridge — Cubic age")
```

### Saturated age

```{r bridge-sat}
bridge_table(bridge_cells[["Saturated age"]]$w, "Weight-statistics bridge — Saturated age")
```

---

# 5. Love Plots

```{r love-helpers, echo=FALSE}
# Return NULL when a fit failed or its outcome weights are not an exact
# representation. Approximate XGBoost weights remain visible in Section 3 but
# do not enter the exact Love-plot comparison.
maybe_love <- function(title, w, D, X, ok = TRUE) {
  if (!isTRUE(ok) || all(is.na(w))) NULL else make_love(title, w, D, X)
}
```

## DML vs. Kappa {.tabset}

### Cubic age (untuned/tuned forests and kappa)

```{r love-cub, fig.height=9, fig.width=13}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nforest untuned",    w_cub_untuned$plriv,    D, X_dml_cub),
  maybe_love("Wald-AIPW\nforest untuned", w_cub_untuned$waldaipw, D, X_dml_cub),
  maybe_love("PLR-IV\nforest tuned",      w_cub_tuned$plriv,      D, X_dml_cub),
  maybe_love("Wald-AIPW\nforest tuned",   w_cub_tuned$waldaipw,   D, X_dml_cub),
  maybe_love("τ_cb_u (CBPS)",    kw_cub$tau_cb_u,   D, X_dml_cub),
  maybe_love("τ_ml_u",           kw_cub$tau_ml_u,   D, X_dml_cub),
  maybe_love("τ_ml_a10",         kw_cub$tau_ml_a10, D, X_dml_cub),
  maybe_love("τ_ml_a (unnorm)",  kw_cub$tau_ml_a,   D, X_dml_cub),
  maybe_love("τ_ml_t (unnorm)",  kw_cub$tau_ml_a1,  D, X_dml_cub),
  maybe_love("τ_ml_a0 (unnorm)", kw_cub$tau_ml_a0,  D, X_dml_cub)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — Cubic age: untuned/tuned honest forest vs. kappa")
```

### Saturated age (untuned/tuned forests and kappa)

```{r love-sat, fig.height=9, fig.width=13}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nforest untuned",    w_sat_untuned$plriv,    D, X_dml_sat),
  maybe_love("Wald-AIPW\nforest untuned", w_sat_untuned$waldaipw, D, X_dml_sat),
  maybe_love("PLR-IV\nforest tuned",      w_sat_tuned$plriv,      D, X_dml_sat),
  maybe_love("Wald-AIPW\nforest tuned",   w_sat_tuned$waldaipw,   D, X_dml_sat),
  maybe_love("τ_cb_u (CBPS)",    kw_sat$tau_cb_u,   D, X_dml_sat),
  maybe_love("τ_ml_u",           kw_sat$tau_ml_u,   D, X_dml_sat),
  maybe_love("τ_ml_a10",         kw_sat$tau_ml_a10, D, X_dml_sat),
  maybe_love("τ_ml_a (unnorm)",  kw_sat$tau_ml_a,   D, X_dml_sat),
  maybe_love("τ_ml_t (unnorm)",  kw_sat$tau_ml_a1,  D, X_dml_sat),
  maybe_love("τ_ml_a0 (unnorm)", kw_sat$tau_ml_a0,  D, X_dml_sat)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — Saturated age: untuned/tuned honest forest vs. kappa")
```

## Learners vs. Kappa {.tabset}

### Cubic age (PLR-IV and Wald-AIPW by learner)

```{r ext-love-cub, fig.height=12, fig.width=14}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nOLS nuisances", lm_pliv_cub$w, D, X_dml_cub, lm_pliv_cub$ok),
  maybe_love("Wald-AIPW\nOLS/logit nuisances", lm_iivm_cub$w, D, X_dml_cub, lm_iivm_cub$ok),
  maybe_love("PLR-IV\nranger",       rf_pliv_cub$w,   D, X_dml_cub, rf_pliv_cub$ok),
  maybe_love("Wald-AIPW\nranger",    rf_iivm_cub$w,   D, X_dml_cub, rf_iivm_cub$ok),
  maybe_love("PLR-IV\nXGBoost untuned", xgb_pliv_cub_untuned$w, D, X_dml_cub, xgb_pliv_cub_untuned$ok),
  maybe_love("PLR-IV\nXGBoost tuned", xgb_pliv_cub_tuned$w, D, X_dml_cub, xgb_pliv_cub_tuned$ok),
  maybe_love("Wald-AIPW\nXGBoost untuned", xgb_iivm_cub_untuned$w, D, X_dml_cub, xgb_iivm_cub_untuned$ok),
  maybe_love("Wald-AIPW\nXGBoost tuned", xgb_iivm_cub_tuned$w, D, X_dml_cub, xgb_iivm_cub_tuned$ok),
  maybe_love("τ_cb_u\nkappa",        kw_cub$tau_cb_u, D, X_dml_cub),
  maybe_love("τ_ml_u\nkappa",        kw_cub$tau_ml_u, D, X_dml_cub)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — PLR-IV and Wald-AIPW learners vs. kappa | Cubic age")
```

### Saturated age (PLR-IV and Wald-AIPW by learner)

```{r ext-love-sat, fig.height=12, fig.width=14}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nOLS nuisances", lm_pliv_sat$w, D, X_dml_sat, lm_pliv_sat$ok),
  maybe_love("Wald-AIPW\nOLS/logit nuisances", lm_iivm_sat$w, D, X_dml_sat, lm_iivm_sat$ok),
  maybe_love("PLR-IV\nranger",       rf_pliv_sat$w,   D, X_dml_sat, rf_pliv_sat$ok),
  maybe_love("Wald-AIPW\nranger",    rf_iivm_sat$w,   D, X_dml_sat, rf_iivm_sat$ok),
  maybe_love("PLR-IV\nXGBoost untuned", xgb_pliv_sat_untuned$w, D, X_dml_sat, xgb_pliv_sat_untuned$ok),
  maybe_love("PLR-IV\nXGBoost tuned", xgb_pliv_sat_tuned$w, D, X_dml_sat, xgb_pliv_sat_tuned$ok),
  maybe_love("Wald-AIPW\nXGBoost untuned", xgb_iivm_sat_untuned$w, D, X_dml_sat, xgb_iivm_sat_untuned$ok),
  maybe_love("Wald-AIPW\nXGBoost tuned", xgb_iivm_sat_tuned$w, D, X_dml_sat, xgb_iivm_sat_tuned$ok),
  maybe_love("τ_cb_u\nkappa",        kw_sat$tau_cb_u, D, X_dml_sat),
  maybe_love("τ_ml_u\nkappa",        kw_sat$tau_ml_u, D, X_dml_sat)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — PLR-IV and Wald-AIPW learners vs. kappa | Saturated age")
```
