---
title: "Angrist (1990) — Translation Invariance: Rerun Check (Method A)"
subtitle: "Vietnam draft lottery: kappa, honest forest, DoubleML learners, and nested XGBoost"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size:15px; color:#2c2c2c; }
h1 { border-bottom:2px solid #2c7bb6; padding-bottom:6px; margin-top:36px; }
h2 { color:#2c7bb6; margin-top:28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size:13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=6)
```

# Purpose

This notebook implements the rerun definition of translation invariance. For
every estimator, the complete pipeline is fitted to the original log-wage
outcome and to

\[
Y_i^{\mathrm{shift}}=Y_i+\log(100),
\]

which is the cents recoding used in the SUW reproduction. Method A compares
the two refitted estimates directly. Method B predicts the change from the
original outcome weights, `log(100) * sum(omega)`. `gap_AB` records whether
the rerun and frozen-weight calculations agree.

<div class="note-box">
**Two workflows.** The same pure additive shift is applied to the cubic-age
and saturated-age specifications. These are the two feature designs used by
the DML and learner analyses. Repeating additional constants would not test a
different algebraic condition: translation invariance is equivalent to zero
total outcome-weight mass for every additive constant.
</div>

# Setup

```{r packages-and-source}
library(haven)
library(knitr)
library(kableExtra)
library(dplyr)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")
source("functions_all.R")

ti_vietnam_row <- function(name,w_original,tau_original,tau_shifted,k,digits=12) {
  row <- ti_unified_row(name=name,w_dollars=w_original,
                        tau_dollars=tau_original,tau_cents=tau_shifted,
                        k=k,digits=digits)
  names(row)[names(row)=="tau_dollars"] <- "tau_original"
  names(row)[names(row)=="tau_cents"] <- "tau_shifted"
  row$Shift_k <- round(k,digits)
  row[,c("Estimator","Shift_k","tau_original","tau_shifted","sum_w",
         "diff_B","diff_A","gap_AB","invariant_B","invariant_A")]
}
```

## Runtime switches

```{r switches}
RUN_KAPPA <- TRUE
RUN_DML <- TRUE
RUN_LM_RF <- TRUE
RUN_XGB_METHOD_A <- TRUE
RUN_XGB_WALD_AIPW <- TRUE
RUN_XGB_PLIV <- TRUE
USE_XGB_ORIGINAL_EXPORT <- TRUE

SEED <- 42
N_CF <- 5
TUNE_EVALS <- 15
INNER_FOLDS <- 3
```

<div class="warn-box">
**XGBoost rule.** Original-outcome tuned estimates and outcome weights are
imported from the two tuning exports. For the shifted outcome, nested tuning
is rerun with the same seed, five outer folds, three inner folds, fifteen
random-search evaluations, and `tune_on_folds = TRUE`. Reusing median
hyperparameters would define a different estimator and is deliberately not
done.
</div>

## Data — the exact main-analysis sample

```{r data}
stopifnot(file.exists("sipp.dta"))
sipp <- read_dta("sipp.dta")
sipp <- sipp[sipp$rsncode!=999 & !is.na(sipp$educ) & !is.na(sipp$kwage),]
sipp$age <- sipp$age_5
sipp$age2 <- sipp$age_5^2
sipp$age3 <- sipp$age_5^3
sipp$age_fac <- factor(sipp$age_5)

Y_orig <- as.numeric(log(sipp$kwage))
k_shift <- log(100)
Y_shift <- Y_orig+k_shift
D <- as.integer(sipp$nvstat)
Z <- as.integer(sipp$rsncode)

X_kappa_cub <- model.matrix(~age+age2+age3,data=sipp)
X_kappa_sat <- model.matrix(~age_fac,data=sipp)
X_dml_cub <- make_X(~0+age+I(age^2)+I(age^3),data=sipp)
X_dml_sat_full <- make_X(~0+age_fac,data=sipp)
X_dml_sat <- X_dml_sat_full[,-1,drop=FALSE]

workflows <- list(
  "Cubic age"=list(label="Cubic age",Y_orig=Y_orig,Y_shift=Y_shift,k=k_shift,
    D=D,Z=Z,X_kappa=X_kappa_cub,X_dml=X_dml_cub),
  "Saturated age"=list(label="Saturated age",Y_orig=Y_orig,Y_shift=Y_shift,k=k_shift,
    D=D,Z=Z,X_kappa=X_kappa_sat,X_dml=X_dml_sat))
cat(sprintf("N=%d | shift=log(100)=%.8f | workflows=%s\n",
            nrow(sipp),k_shift,paste(names(workflows),collapse=", ")))
```

# 1. Kappa estimators

<div class="note-box">
Kappa weights depend on the instrument, treatment, covariates, and propensity
score, but not on the outcome. Consequently, Method A and Method B should
agree to numerical precision. The normalized estimators should remain
unchanged; the unnormalized estimators are expected to move.
</div>

```{r kappa-builders}
kappa_key_map <- c(tau_cb_u="tau_cb_u",tau_ml_u="tau_ml_u",
  tau_ml_a10="tau_ml_a10",tau_ml_a="tau_ml_a",tau_ml_t="tau_ml_a1",
  tau_ml_a0="tau_ml_a0")
build_kappa_rows <- function(wf) {
  est_o <- kappa_analytic_se_all(wf$Y_orig,wf$Z,wf$D,wf$X_kappa)$estimates
  est_s <- kappa_analytic_se_all(wf$Y_shift,wf$Z,wf$D,wf$X_kappa)$estimates
  W <- kappa_weights_bundle(wf$Z,wf$D,wf$X_kappa)
  do.call(rbind,lapply(names(kappa_key_map),function(est){
    row <- ti_vietnam_row(est,W[[kappa_key_map[[est]]]],est_o[[est]],est_s[[est]],wf$k)
    row$Specification <- wf$label; row
  }))
}
```

```{r kappa-run, cache=TRUE}
if(RUN_KAPPA)kappa_ti_all<-do.call(rbind,lapply(workflows,build_kappa_rows))
```

```{r kappa-table}
if(RUN_KAPPA)kappa_ti_all[,c("Specification",setdiff(names(kappa_ti_all),c("Specification","gap_AB")))]%>%
  kbl(caption="Kappa estimators — original versus shifted log wage",digits=8,row.names=FALSE)%>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE)%>%
  row_spec(which(!kappa_ti_all$invariant_A),background="#fffbf0")%>%
  footnote(general="Orange rows fail Method A, as expected for the unnormalized kappa estimators. For kappa, gap_AB should be approximately zero because the weights do not depend on Y.")
```

# 2. Headline DML: tuned honest forest

```{r dml-headline-builder}
build_dml_headline_rows <- function(wf) {
  set.seed(SEED); fit_o <- dml_with_smoother(wf$Y_orig,wf$D,wf$X_dml,wf$Z,
    n_cf_folds=N_CF,tune.parameters="all")
  set.seed(SEED); fit_s <- dml_with_smoother(wf$Y_shift,wf$D,wf$X_dml,wf$Z,
    n_cf_folds=N_CF,tune.parameters="all")
  ro <- summary(fit_o); rs <- summary(fit_s); om <- get_outcome_weights(fit_o)
  check_omega_rows(om)
  do.call(rbind,lapply(c("PLR-IV","Wald-AIPW"),function(est){
    row <- ti_vietnam_row(est,as.vector(om$omega[est,]),get_estimate(ro,est),
                          get_estimate(rs,est),wf$k)
    row$Specification <- wf$label; row
  }))
}
```

```{r dml-headline-run, cache=TRUE}
if(RUN_DML)dml_headline_ti_all<-do.call(rbind,lapply(workflows,build_dml_headline_rows))
```

```{r dml-headline-table}
if(RUN_DML)dml_headline_ti_all[,c("Specification",setdiff(names(dml_headline_ti_all),c("Specification","gap_AB")))]%>%
  kbl(caption="Tuned honest forest — original versus shifted log wage",digits=8,row.names=FALSE)%>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE)%>%
  row_spec(which(!dml_headline_ti_all$invariant_A),background="#fffbf0")
```

# 3. OLS/logit and Ranger learners

<div class="note-box">
“OLS/logit” describes the nuisance learner, not the linear-age feature
specification. Every row is evaluated once with cubic-age features and once
with saturated-age features. PLR-IV uses OLS for all three nuisances;
Wald-AIPW uses OLS for the outcome nuisance and logistic regression for the
binary instrument and treatment nuisances.
</div>

```{r learner-setup}
make_dml_data <- function(Y,D,Z,X){df<-as.data.frame(X);df$y<-Y;df$d<-D;df$z<-Z
  DoubleMLData$new(data=df,y_col="y",d_cols="d",z_cols="z",x_cols=colnames(X))}
ml_lm_iivm<-function()list(g=lrn("regr.lm"),m=lrn("classif.log_reg"),r=lrn("classif.log_reg"))
ml_rf_iivm<-function()list(g=lrn("regr.ranger",keep.inbag=TRUE),m=lrn("classif.ranger",keep.inbag=TRUE),r=lrn("classif.ranger",keep.inbag=TRUE))
ml_lm_pliv<-function()list(l=lrn("regr.lm"),m=lrn("regr.lm"),r=lrn("regr.lm"))
ml_rf_pliv<-function()list(l=lrn("regr.ranger",keep.inbag=TRUE),m=lrn("regr.ranger",keep.inbag=TRUE),r=lrn("regr.ranger",keep.inbag=TRUE))
omega_vec<-function(om,row){w<-om$omega;if(is.null(dim(w)))return(as.vector(w))
  if(!is.null(rownames(w))&&row%in%rownames(w))return(as.vector(w[row,]));as.vector(w[1,])}
fit_iivm<-function(dat,L,Y){set.seed(SEED);o<-DoubleMLIIVM$new(data=dat,ml_g=L$g,ml_m=L$m,ml_r=L$r,n_folds=N_CF,score="LATE")
  o$fit(store_models=TRUE,store_predictions=TRUE);om<-get_outcome_weights(object=o,dml_data=dat)
  w<-omega_vec(om,"Wald-AIPW");list(obj=o,w=w,ok=check_weight_identity(w,Y,get_dml_coef(o)))}
fit_pliv<-function(dat,L,Y){set.seed(SEED);o<-DoubleMLPLIV$new(data=dat,ml_l=L$l,ml_m=L$m,ml_r=L$r,n_folds=N_CF)
  o$fit(store_models=TRUE,store_predictions=TRUE);om<-get_outcome_weights(object=o,dml_data=dat)
  w<-omega_vec(om,"PLR-IV");list(obj=o,w=w,ok=check_weight_identity(w,Y,get_dml_coef(o)))}
build_learner_row<-function(wf,label,Lfun,fitfun){
  fo<-fitfun(make_dml_data(wf$Y_orig,wf$D,wf$Z,wf$X_dml),Lfun(),wf$Y_orig)
  fs<-fitfun(make_dml_data(wf$Y_shift,wf$D,wf$Z,wf$X_dml),Lfun(),wf$Y_shift)
  row<-ti_vietnam_row(label,fo$w,get_dml_coef(fo$obj),get_dml_coef(fs$obj),wf$k)
  row$Specification<-wf$label;row$identity_ok<-fo$ok;row}
```

```{r learner-run, cache=TRUE}
if(RUN_LM_RF)learner_ti_all<-do.call(rbind,lapply(workflows,function(wf)rbind(
  build_learner_row(wf,"PLR-IV (OLS nuisances)",ml_lm_pliv,fit_pliv),
  build_learner_row(wf,"Wald-AIPW (OLS/logit nuisances)",ml_lm_iivm,fit_iivm),
  build_learner_row(wf,"PLR-IV (Ranger)",ml_rf_pliv,fit_pliv),
  build_learner_row(wf,"Wald-AIPW (Ranger)",ml_rf_iivm,fit_iivm))))
```

```{r learner-table}
if(RUN_LM_RF)learner_ti_all[,c("Specification",setdiff(names(learner_ti_all),c("Specification","gap_AB","identity_ok")))]%>%
  kbl(caption="DoubleML learners — original versus shifted log wage",digits=8,row.names=FALSE)%>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE)%>%
  row_spec(which(!learner_ti_all$invariant_A),background="#fffbf0")%>%
  footnote(general="OLS/logit labels the nuisance learner class; cubic and saturated labels identify the feature specification. Orange rows fail the full rerun invariance check.")
```

# 4. Tuned XGBoost Method A

```{r xgb-setup}
if(RUN_XGB_METHOD_A){
  if(!USE_XGB_ORIGINAL_EXPORT)stop("USE_XGB_ORIGINAL_EXPORT must be TRUE.")
  req<-c("vietnam_xgb_tuning_export.rds","vietnam_xgb_pliv_tuning_export.rds")
  if(!all(file.exists(req)))stop("Run both Vietnam XGBoost tuning notebooks before Method A.")
  xgb_wald_export<-readRDS(req[1]);xgb_pliv_export<-readRDS(req[2])
  for(ex in list(xgb_wald_export,xgb_pliv_export)){
    if(!isTRUE(ex$settings$tune_on_folds))stop("Export is not a tune_on_folds=TRUE fit.")
    if(!all(names(workflows)%in%names(ex$omega_vectors)))stop("Export cells do not match Method-A workflows.")
    if(!identical(as.integer(ex$settings$seed),as.integer(SEED))||
       !identical(as.integer(ex$settings$n_cf),as.integer(N_CF))||
       !identical(as.integer(ex$settings$tune_evals),as.integer(TUNE_EVALS))||
       !identical(as.integer(ex$settings$inner_folds),as.integer(INNER_FOLDS)))
      stop("Export tuning settings do not match the Method-A rerun settings.")
  }
}

xgb_ps<-function()ps(nrounds=p_int(50,400),max_depth=p_int(2,5),eta=p_dbl(0.02,0.20),min_child_weight=p_int(1,8))
make_xgb_iivm<-function()list(g=lrn("regr.xgboost",alpha=0,subsample=1,max_delta_step=0,base_score=0),m=lrn("classif.xgboost",predict_type="prob"),r=lrn("classif.xgboost",predict_type="prob"))
make_xgb_pliv<-function(){b<-list(alpha=0,subsample=1,max_delta_step=0,base_score=0);list(l=do.call(lrn,c(list("regr.xgboost"),b)),m=do.call(lrn,c(list("regr.xgboost"),b)),r=do.call(lrn,c(list("regr.xgboost"),b)))}
tune_set_iivm<-function()list(ml_g=xgb_ps(),ml_m=xgb_ps(),ml_r=xgb_ps())
tune_set_pliv<-function()list(ml_l=xgb_ps(),ml_m=xgb_ps(),ml_r=xgb_ps())
tune_settings_iivm<-function()list(n_folds_tune=INNER_FOLDS,rsmp_tune=rsmp("cv",folds=INNER_FOLDS),measure=list(ml_g="regr.rmse",ml_m="classif.logloss",ml_r="classif.logloss"),terminator=trm("evals",n_evals=TUNE_EVALS),algorithm=tnr("random_search"))
tune_settings_pliv<-function()list(n_folds_tune=INNER_FOLDS,rsmp_tune=rsmp("cv",folds=INNER_FOLDS),measure=list(ml_l="regr.rmse",ml_m="regr.rmse",ml_r="regr.rmse"),terminator=trm("evals",n_evals=TUNE_EVALS),algorithm=tnr("random_search"))

get_original_xgb<-function(export,cell,learner_label){
  row<-export$weight_diagnostics%>%filter(.data$Cell==cell,.data$Learner==learner_label)
  if(nrow(row)!=1)stop(sprintf("Expected one diagnostic row for %s / %s.",cell,learner_label))
  w<-export$omega_vectors[[cell]]$tuned
  if(is.null(w)||!length(w))stop(sprintf("Tuned outcome weights are unavailable for %s / %s.",cell,learner_label))
  tau<-as.numeric(row$Estimate[1])
  if(!is.finite(tau))stop(sprintf("Tuned estimate is unavailable for %s / %s.",cell,learner_label))
  list(tau=tau,w=as.vector(w),identity_ok=isTRUE(export$identity_checks[[cell]]$tuned))}
fit_shifted_wald<-function(wf){dat<-make_dml_data(wf$Y_shift,wf$D,wf$Z,wf$X_dml);L<-make_xgb_iivm()
  o<-DoubleMLIIVM$new(data=dat,ml_g=L$g,ml_m=L$m,ml_r=L$r,n_folds=N_CF,score="LATE")
  o$tune(param_set=tune_set_iivm(),tune_settings=tune_settings_iivm(),tune_on_folds=TRUE)
  o$fit(store_models=TRUE,store_predictions=TRUE);get_dml_coef(o)}
fit_shifted_pliv<-function(wf){dat<-make_dml_data(wf$Y_shift,wf$D,wf$Z,wf$X_dml);L<-make_xgb_pliv()
  o<-DoubleMLPLIV$new(data=dat,ml_l=L$l,ml_m=L$m,ml_r=L$r,n_folds=N_CF)
  o$tune(param_set=tune_set_pliv(),tune_settings=tune_settings_pliv(),tune_on_folds=TRUE)
  o$fit(store_models=TRUE,store_predictions=TRUE);get_dml_coef(o)}
build_xgb_rows<-function(wf){out<-list()
  if(RUN_XGB_WALD_AIPW){orig<-get_original_xgb(xgb_wald_export,wf$label,"XGB tuned");set.seed(SEED);ts<-fit_shifted_wald(wf)
    r<-ti_vietnam_row("Wald-AIPW (XGBoost tuned)",orig$w,orig$tau,ts,wf$k);r$Specification<-wf$label;r$identity_ok<-orig$identity_ok;out$wald<-r}
  if(RUN_XGB_PLIV){orig<-get_original_xgb(xgb_pliv_export,wf$label,"PLR-IV XGB tuned");set.seed(SEED);ts<-fit_shifted_pliv(wf)
    r<-ti_vietnam_row("PLR-IV (XGBoost tuned)",orig$w,orig$tau,ts,wf$k);r$Specification<-wf$label;r$identity_ok<-orig$identity_ok;out$pliv<-r}
  do.call(rbind,out)}
```

```{r xgb-run, cache=TRUE, cache.extra=list(RUN_XGB_METHOD_A,RUN_XGB_WALD_AIPW,RUN_XGB_PLIV,TUNE_EVALS,INNER_FOLDS)}
if(RUN_XGB_METHOD_A){
  xgb_ti_all<-do.call(rbind,lapply(workflows,build_xgb_rows))
  saveRDS(list(results=xgb_ti_all,settings=list(seed=SEED,n_cf=N_CF,
    tune_evals=TUNE_EVALS,inner_folds=INNER_FOLDS,tune_on_folds=TRUE,
    original_exports=c("vietnam_xgb_tuning_export.rds","vietnam_xgb_pliv_tuning_export.rds"))),
    "vietnam_translation_xgb_method_a_export.rds")
}
```

```{r xgb-table}
if(RUN_XGB_METHOD_A)xgb_ti_all[,c("Specification",setdiff(names(xgb_ti_all),c("Specification","gap_AB","identity_ok")))]%>%
  kbl(caption="Tuned XGBoost Method A — original export versus shifted nested-tuning rerun",digits=8,row.names=FALSE)%>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE)%>%
  row_spec(which(!xgb_ti_all$invariant_A),background="#fffbf0")%>%
  footnote(general="Original estimates and weights are imported from the exact tuned exports. Shifted estimates rerun the complete nested tuning procedure. Orange indicates failure of the full rerun invariance check.")
```

```{r xgb-identity-table}
if(RUN_XGB_METHOD_A)xgb_ti_all%>%select(Specification,Estimator,identity_ok,sum_w,diff_B,diff_A,gap_AB)%>%
  kbl(caption="XGBoost outcome-weight identity and Method-A/Method-B gap",digits=8,row.names=FALSE)%>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE)%>%
  footnote(general="identity_ok refers to the imported original-outcome weight representation. When FALSE, Method B is only an approximate learner-sensitivity diagnostic; Method A remains the relevant rerun result.")
```

<div class="note-box">
**Interpretation.** Kappa and the honest forest are affine-smoother cases for
which Methods A and B should agree closely. For adaptive Ranger and XGBoost
fits, a nonzero `gap_AB` can arise because the learner is re-estimated after
the outcome shift. Large discrepancies require investigation; unnormalized
kappa failures are expected by construction.
</div>
