---
title: "Angrist (1990) — XGBoost Nuisance Tuning for PLR-IV"
subtitle: "Vietnam draft lottery: nested tuning inside each outer cross-fitting fold"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size:15px; color:#2c2c2c; }
h1 { border-bottom:2px solid #2c7bb6; padding-bottom:6px; margin-top:36px; }
h2 { color:#2c7bb6; margin-top:28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size:13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,message=FALSE,warning=FALSE,fig.width=11,fig.height=5.5)
```

```{r packages}
library(haven)
library(knitr)
library(kableExtra)
library(dplyr)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)
source("functions_all.R")
lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

SEED <- 42
N_CF <- 5
TUNE_EVALS <- 15
INNER_FOLDS <- 3
EXTRACT_WEIGHTS <- TRUE
```

<div class="note-box">
**Purpose.** This notebook mirrors `vietnam_xgb_tuning.Rmd` for PLR-IV
using `DoubleMLPLIV`. It estimates untuned and fold-specifically tuned
XGBoost nuisances for cubic and saturated age, evaluates their out-of-fold
prediction losses, extracts outcome weights, checks the outcome-weight
identity, and exports all downstream objects. The main Vietnam notebook
imports the export and never retunes XGBoost during an ordinary knit.
</div>

<div class="warn-box">
**Runtime.** PLR-IV tunes three regression nuisances (`ml_l`, `ml_m`, and
`ml_r`) separately within every outer training fold. Keep
`EXTRACT_WEIGHTS = TRUE`; otherwise the main diagnostics and Method-A check
cannot use the exact fitted vectors.
</div>

# Setup

```{r data}
stopifnot(file.exists("sipp.dta"))
sipp <- read_dta("sipp.dta")
sipp <- sipp[sipp$rsncode != 999 & !is.na(sipp$educ) & !is.na(sipp$kwage), ]
sipp$age <- sipp$age_5
sipp$age_fac <- factor(sipp$age_5)
sipp$lwage <- log(sipp$kwage)

Y <- as.numeric(sipp$lwage)
D <- as.integer(sipp$nvstat)
Z <- as.integer(sipp$rsncode)
X_cub <- make_X(~0+age+I(age^2)+I(age^3),data=sipp)
X_sat_full <- make_X(~0+age_fac,data=sipp)
X_sat <- X_sat_full[,-1,drop=FALSE]
cat(sprintf("N=%d | cubic=%d covariates | saturated=%d covariates\n",
            nrow(sipp),ncol(X_cub),ncol(X_sat)))
```

```{r learners-and-tuning}
make_dml_data <- function(Y,D,Z,X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data=df,y_col="y",d_cols="d",z_cols="z",x_cols=colnames(X))
}

make_xgb_learners <- function() {
  base <- list(alpha=0,subsample=1,max_delta_step=0,base_score=0)
  list(l=do.call(lrn,c(list("regr.xgboost"),base)),
       m=do.call(lrn,c(list("regr.xgboost"),base)),
       r=do.call(lrn,c(list("regr.xgboost"),base)))
}
xgb_ps <- function() ps(nrounds=p_int(50,400),max_depth=p_int(2,5),
                         eta=p_dbl(0.02,0.20),min_child_weight=p_int(1,8))
tune_param_set <- function() list(ml_l=xgb_ps(),ml_m=xgb_ps(),ml_r=xgb_ps())
tune_settings <- function() list(
  n_folds_tune=INNER_FOLDS,rsmp_tune=rsmp("cv",folds=INNER_FOLDS),
  measure=list(ml_l="regr.rmse",ml_m="regr.rmse",ml_r="regr.rmse"),
  terminator=trm("evals",n_evals=TUNE_EVALS),algorithm=tnr("random_search"))

get_prediction_vector <- function(obj,name) {
  p <- tryCatch(obj$predictions[[name]],error=function(e)NULL)
  if(is.null(p))return(numeric(0)); as.numeric(drop(p))
}
rmse_loss <- function(truth,pred) {
  if(length(pred)!=length(truth)||!length(pred))return(NA_real_)
  keep <- is.finite(truth)&is.finite(pred)
  if(!any(keep))return(NA_real_)
  sqrt(mean((as.numeric(truth[keep])-as.numeric(pred[keep]))^2))
}
calc_nuisance_losses <- function(r,cell) {
  if(is.null(r))return(c(outcome_l_rmse=NA,instrument_m_rmse=NA,treatment_r_rmse=NA))
  c(outcome_l_rmse=rmse_loss(cell$Y,get_prediction_vector(r$obj,"ml_l")),
    instrument_m_rmse=rmse_loss(cell$Z,get_prediction_vector(r$obj,"ml_m")),
    treatment_r_rmse=rmse_loss(cell$D,get_prediction_vector(r$obj,"ml_r")))
}

fit_cell <- function(dat,learners,label,tune=FALSE) {
  tryCatch({
    set.seed(SEED)
    obj <- DoubleMLPLIV$new(data=dat,ml_l=learners$l,ml_m=learners$m,
                            ml_r=learners$r,n_folds=N_CF)
    if(tune)obj$tune(param_set=tune_param_set(),tune_settings=tune_settings(),
                     tune_on_folds=TRUE)
    obj$fit(store_models=TRUE,store_predictions=TRUE)
    list(label=label,tuned=tune,coef=get_dml_coef(obj),se=get_dml_se(obj),
         tuning_result=tryCatch(obj$tuning_result,error=function(e)NULL),
         params=tryCatch(obj$params,error=function(e)NULL),obj=obj)
  },error=function(e){message(sprintf("Fit failed for %s: %s",label,conditionMessage(e)));NULL})
}

cells <- list("Cubic age"=list(Y=Y,D=D,Z=Z,X=X_cub),
              "Saturated age"=list(Y=Y,D=D,Z=Z,X=X_sat))
```

# Fit: untuned versus tuned XGBoost

```{r fit-all, cache=TRUE}
results <- list()
for(cn in names(cells)) {
  cell <- cells[[cn]]; dat <- make_dml_data(cell$Y,cell$D,cell$Z,cell$X)
  results[[cn]] <- list(
    untuned=fit_cell(dat,make_xgb_learners(),paste(cn,"pliv-xgb-untuned"),FALSE),
    tuned=fit_cell(dat,make_xgb_learners(),paste(cn,"pliv-xgb-tuned"),TRUE))
  cat(sprintf("Finished: %s\n",cn))
}
```

## 1. Estimate comparison

```{r comparison-table}
fmt_est <- function(r)if(is.null(r))"—"else sprintf("%.3f (%.3f)",r$coef,r$se)
rows <- do.call(rbind,lapply(names(results),function(cn){R<-results[[cn]]
  data.frame(Cell=cn,Learner=c("PLR-IV XGB untuned","PLR-IV XGB tuned (tune_on_folds=TRUE)"),
             `Estimate (SE)`=c(fmt_est(R$untuned),fmt_est(R$tuned)),check.names=FALSE)}))
rows %>% kbl(caption="Vietnam PLR-IV: XGBoost untuned vs. fold-specifically tuned",
             align=c("l","l","c")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
  collapse_rows(columns=1,valign="middle") %>%
  row_spec(which(rows$Learner=="PLR-IV XGB tuned (tune_on_folds=TRUE)"),background="#eef6ff") %>%
  footnote(general="Tuned nuisances select nrounds, max_depth, eta, and min_child_weight by three-fold random-search CV inside each outer training fold.")
```

## 2. Nuisance-prediction losses

```{r nuisance-loss-table}
loss_rows <- do.call(rbind,lapply(names(results),function(cn){R<-results[[cn]];cell<-cells[[cn]]
  u<-calc_nuisance_losses(R$untuned,cell);t<-calc_nuisance_losses(R$tuned,cell)
  data.frame(Cell=cn,Learner=c("PLR-IV XGB untuned","PLR-IV XGB tuned"),
    `Outcome l RMSE`=c(u["outcome_l_rmse"],t["outcome_l_rmse"]),
    `Instrument m RMSE`=c(u["instrument_m_rmse"],t["instrument_m_rmse"]),
    `Treatment r RMSE`=c(u["treatment_r_rmse"],t["treatment_r_rmse"]),
    check.names=FALSE,row.names=NULL)}))
loss_rows %>% kbl(digits=4,caption="PLR-IV nuisance losses: untuned vs. tuned (lower is better)") %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
  collapse_rows(columns=1,valign="middle") %>%
  row_spec(which(loss_rows$Learner=="PLR-IV XGB tuned"),background="#eef6ff")
```

## 3. Selected hyperparameters

```{r hyperparameter-summary}
extract_tuned_params <- function(r,cn){
  if(is.null(r)||!r$tuned)return(data.frame())
  src<-if(!is.null(r$params))r$params else r$tuning_result
  if(is.null(src))return(data.frame())
  u<-unlist(src,recursive=TRUE);u<-u[grepl("nrounds|max_depth|eta|min_child_weight",names(u))]
  if(!length(u))return(data.frame())
  hp<-data.frame(Cell=cn,Raw_name=names(u),Value=as.numeric(u))
  hp$Nuisance<-ifelse(grepl("ml_l",hp$Raw_name),"Outcome l",
               ifelse(grepl("ml_m",hp$Raw_name),"Instrument m",
               ifelse(grepl("ml_r",hp$Raw_name),"Treatment r","Unknown")))
  hp$Parameter<-sub(".*(nrounds|max_depth|eta|min_child_weight).*","\\1",hp$Raw_name);hp
}
hp_parts<-lapply(names(results),function(cn)extract_tuned_params(results[[cn]]$tuned,cn))
hp_parts<-Filter(function(x)nrow(x)>0,hp_parts)
hp_raw<-if(length(hp_parts))do.call(rbind,hp_parts)else data.frame()
if(nrow(hp_raw)){
  hp_summary<-hp_raw%>%filter(Nuisance!="Unknown")%>%group_by(Cell,Nuisance,Parameter)%>%
    summarise(Median_across_folds=median(Value),N_values_seen=n(),.groups="drop")
  hp_summary%>%kbl(digits=3,caption="Selected PLR-IV XGBoost hyperparameters: fold-level median")%>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE)%>%
    collapse_rows(columns=c(1,2),valign="middle")%>%
    footnote(general="These are descriptive fold-level summaries, not a reusable global recipe.")
}else cat("No hyperparameters could be extracted automatically; inspect results[[...]]$tuned$obj$params.\n")
```

## 4. Outcome-weight diagnostics and export

```{r weight-helpers}
omega_vec<-function(om){w<-om$omega;if(is.null(dim(w)))return(as.vector(w))
  if(!is.null(rownames(w))&&"PLR-IV"%in%rownames(w))return(as.vector(w["PLR-IV",]));as.vector(w[1,])}
extract_xgb_w<-function(r,cell){if(is.null(r)||is.null(r$obj))stop("fitted DoubleML object is NULL")
  omega_vec(get_outcome_weights(object=r$obj,dml_data=make_dml_data(cell$Y,cell$D,cell$Z,cell$X)))}
safe_extract_xgb_w<-function(r,cell)tryCatch(extract_xgb_w(r,cell),error=function(e)NULL)
one_diag_row<-function(cn,label,r,cell){
  w<-safe_extract_xgb_w(r,cell)
  if(is.null(r)||is.null(w))return(data.frame(Cell=cn,Learner=label,Estimate=if(is.null(r))NA else r$coef,
    ESS=NA,ESS_mod=NA,`Sigma omega`=NA,`Sigma abs omega`=NA,`Treated mass`=NA,
    `Control mass (-sum control weights)`=NA,`max abs omega`=NA,`Identity ok`=NA,
    `Weight status`=if(is.null(r))"fit failed"else"weight extraction failed",check.names=FALSE))
  wb<-weight_bridge(w,cell$D,label);ov<-wb[wb$Scope=="Overall",];tr<-wb[wb$Scope=="Treated",];co<-wb[wb$Scope=="Control",]
  data.frame(Cell=cn,Learner=label,Estimate=r$coef,ESS=ov$ESS_kish,ESS_mod=ov$ESS_mod,
    `Sigma omega`=ov$Sum_w,`Sigma abs omega`=ov$Sum_abs_w,`Treated mass`=tr$Sum_w,
    `Control mass (-sum control weights)`=co$Sum_w,`max abs omega`=max(abs(w)),
    `Identity ok`=tryCatch(check_weight_identity(w,cell$Y,r$coef),error=function(e)NA),
    `Weight status`="ok",check.names=FALSE)}
```

```{r weight-table}
xgb_pliv_weight_table<-do.call(rbind,lapply(names(cells),function(cn){R<-results[[cn]];cell<-cells[[cn]]
  rbind(one_diag_row(cn,"PLR-IV XGB untuned",R$untuned,cell),
        one_diag_row(cn,"PLR-IV XGB tuned",R$tuned,cell))}))
xgb_pliv_weight_table%>%kbl(digits=4,caption="Vietnam PLR-IV XGBoost outcome-weight diagnostics")%>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE)%>%
  row_spec(which(xgb_pliv_weight_table$Learner=="PLR-IV XGB tuned"),background="#eef6ff")%>%
  footnote(general="Identity ok verifies sum(omega*Y)=tau_hat. Failed identities remain available as approximate learner-sensitivity diagnostics but cannot enter exact weight comparisons on equal footing.")
```

```{r omega-and-identities}
xgb_pliv_omega_vectors<-lapply(names(cells),function(cn)list(
  untuned=if(EXTRACT_WEIGHTS)safe_extract_xgb_w(results[[cn]]$untuned,cells[[cn]])else NULL,
  tuned=if(EXTRACT_WEIGHTS)safe_extract_xgb_w(results[[cn]]$tuned,cells[[cn]])else NULL))
names(xgb_pliv_omega_vectors)<-names(cells)
xgb_pliv_identity_checks<-lapply(names(cells),function(cn){cell<-cells[[cn]];R<-results[[cn]];W<-xgb_pliv_omega_vectors[[cn]]
  list(
    untuned=tryCatch(if(is.null(W$untuned)||is.null(R$untuned))NA else
      check_weight_identity(W$untuned,cell$Y,R$untuned$coef),error=function(e)NA),
    tuned=tryCatch(if(is.null(W$tuned)||is.null(R$tuned))NA else
      check_weight_identity(W$tuned,cell$Y,R$tuned$coef),error=function(e)NA))})
names(xgb_pliv_identity_checks)<-names(cells)
```

```{r export, cache=FALSE}
xgb_pliv_export<-list(estimate_comparison=rows,nuisance_losses=loss_rows,
  hyperparameters=if(exists("hp_summary"))hp_summary else NULL,
  weight_diagnostics=xgb_pliv_weight_table,omega_vectors=xgb_pliv_omega_vectors,
  identity_checks=xgb_pliv_identity_checks,
  settings=list(seed=SEED,n_cf=N_CF,tune_evals=TUNE_EVALS,inner_folds=INNER_FOLDS,
    tune_on_folds=TRUE,learner="xgboost",dml_class="DoubleMLPLIV",
    score="partialling out",extract_weights=EXTRACT_WEIGHTS,cells=names(cells)),
  sample_info=list(n=nrow(sipp),mean_Z=mean(Z),mean_D=mean(D),outcome="log(kwage)",
                   specifications=names(cells)))
saveRDS(xgb_pliv_export,"vietnam_xgb_pliv_tuning_export.rds")
cat("Saved: vietnam_xgb_pliv_tuning_export.rds\n")
```

<div class="note-box">
The main analysis reads `vietnam_xgb_pliv_tuning_export.rds`. Because the
tuned model uses fold-specific choices, downstream notebooks must import the
saved estimates and weights rather than refit a single median recipe.
</div>
