---
title: "Vietnam Draft Lottery — Replication and Outcome-Weight Diagnostics"
subtitle: "Kappa estimators, Double ML, learner sensitivity, and exact weight checks"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body   { font-size: 15px; color: #2c2c2c; }
h1     { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2     { color: #2c7bb6; margin-top: 28px; }
h3     { color: #444; margin-top: 20px; }

.note-box {
  background: #eaf4ff;
  border-left: 4px solid #2c7bb6;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
.result-box {
  background: #f3f8f3;
  border-left: 4px solid #3c8d40;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
.warn-box {
  background: #fff7e6;
  border-left: 4px solid #d89000;
  padding: 10px 16px;
  border-radius: 4px;
  margin: 14px 0;
  font-size: 14px;
}
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width  = 12,
  fig.height = 6
)
options(knitr.kable.NA = "—")
```

---

# Setup

## Packages & Functions

```{r packages-and-source}
# Core
library(haven)
library(AER)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
# DML smoother + balance plots
library(OutcomeWeights)
library(cobalt)
library(viridis)
library(gridExtra)
library(ggplot2)
# DoubleML learners
library(DoubleML)
library(mlr3verse)
lgr::get_logger("mlr3")$set_threshold("warn")

source("functions_all.R")   # estimators, analytical SEs, and weight diagnostics
```

## Data & Design

<div class="note-box">
**Setting.** Angrist (1990) SIPP 1984 subsample (N = 3,027 white men, born 1950–53). Instrument Z = `rsncode` (draft-eligibility indicator). Treatment D = `nvstat` (Vietnam-era veteran status). Outcome: log wages in two units (`lwage = log(kwage)` dollars; `lwage_cnt = log(100·kwage)` cents) — the units only enter the SUW reproduction table. Variable definitions follow the SUW Stata replication exactly: `age = age_5`, with three covariate specifications — **Linear** (`age`), **Cubic** (`age, age², age³`), and **Saturated** (`i.age` = age dummies).
</div>

```{r data}
sipp_path <- "sipp.dta"
stopifnot(file.exists(sipp_path))
sipp <- read_dta(sipp_path)

# Sample restriction: N = 3,027  (drop rsncode==999, missing educ / kwage)
sipp_clean <- sipp[sipp$rsncode != 999 &
                     !is.na(sipp$educ) &
                     !is.na(sipp$kwage), ]

# Age terms — EXACTLY as in suw_jbes_replicate.do: age = age_5
sipp_clean$age     <- sipp_clean$age_5
sipp_clean$age2    <- sipp_clean$age_5^2
sipp_clean$age3    <- sipp_clean$age_5^3
sipp_clean$age_fac <- factor(sipp_clean$age_5)

# Outcomes in two units (note: dollars vs. cents convention is FLIPPED vs. Card)
sipp_clean$lwage     <- log(sipp_clean$kwage)         # dollars
sipp_clean$lwage_cnt <- log(100 * sipp_clean$kwage)   # cents

# Core vectors
Y_cnt <- sipp_clean$lwage_cnt   # cents  (used only for the SUW reproduction layout)
Y_dol <- sipp_clean$lwage       # dollars (the working outcome everywhere else)
Z     <- as.integer(sipp_clean$rsncode)
D     <- as.integer(sipp_clean$nvstat)

stopifnot(nrow(sipp_clean) == 3027L)
```

```{r design-matrices}
# Kappa design matrices — WITH intercept (for logit / CBPS)
X_lin <- model.matrix(~ age,                data = sipp_clean)
X_cub <- model.matrix(~ age + age2 + age3,  data = sipp_clean)
X_sat <- model.matrix(~ age_fac,            data = sipp_clean)

# X_df for 2SLS (no intercept column)
X_lin_df <- data.frame(age = sipp_clean$age)
X_cub_df <- data.frame(age = sipp_clean$age, age2 = sipp_clean$age2, age3 = sipp_clean$age3)
X_sat_df <- data.frame(age_fac = sipp_clean$age_fac)

# DML design matrices — no intercept, attributes stripped for grf (make_X)
X_dml_cub      <- make_X(~ 0 + age + I(age^2) + I(age^3), data = sipp_clean)
X_dml_sat_full <- make_X(~ 0 + age_fac, data = sipp_clean)
X_dml_sat      <- X_dml_sat_full[, -1, drop = FALSE]   # drop one age dummy (reference)

stopifnot(nrow(X_dml_cub) == nrow(sipp_clean),
          nrow(X_dml_sat) == nrow(sipp_clean))
```

---

# 0. Empirical Design Audit

<div class="note-box">
The lottery design is conditionally random within birth cohort. The pooled
eligibility share is therefore a sample-composition statistic, not the basis
for the assignment claim. The relevant empirical checks are whether draft
eligibility shifts veteran status after age adjustment and whether fitted
eligibility probabilities remain away from the boundaries.
</div>

```{r design-audit, echo=FALSE}
first_stage_row <- function(label, formula, X) {
  fit <- lm(formula, data = data.frame(D = D, Z = Z,
                                       age = sipp_clean$age,
                                       age2 = sipp_clean$age2,
                                       age3 = sipp_clean$age3,
                                       age_fac = sipp_clean$age_fac))
  ct <- lmtest::coeftest(fit, vcov. = sandwich::vcovHC(fit, type = "HC1"))
  p <- logit_mle(Z, X)
  data.frame(
    Specification = label,
    `Adjusted first stage` = unname(ct["Z", "Estimate"]),
    `HC1 SE` = unname(ct["Z", "Std. Error"]),
    `Robust t²` = unname(ct["Z", "t value"]^2),
    `Min p̂(X)` = min(p),
    `Max p̂(X)` = max(p),
    `Boundary share` = mean(p < 0.05 | p > 0.95),
    check.names = FALSE
  )
}

design_audit <- rbind(
  first_stage_row("Linear age", D ~ Z + age, X_lin),
  first_stage_row("Cubic age", D ~ Z + age + age2 + age3, X_cub),
  first_stage_row("Saturated age", D ~ Z + age_fac, X_sat)
)

sample_audit <- data.frame(
  N = length(Z),
  `Pr(Z=1)` = mean(Z),
  `Pr(D=1)` = mean(D),
  `Pr(D=1 | Z=0)` = mean(D[Z == 0]),
  `Pr(D=1 | Z=1)` = mean(D[Z == 1]),
  `Raw uptake gap` = mean(D[Z == 1]) - mean(D[Z == 0]),
  check.names = FALSE
)

sample_audit %>%
  kbl(digits = 4, caption = "Sample and unconditional treatment uptake",
      align = rep("r", ncol(sample_audit))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 13)

design_audit %>%
  kbl(digits = 4,
      caption = "Age-adjusted first-stage strength and estimated instrument overlap",
      align = c("l", rep("r", 6))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE, font_size = 13) %>%
  footnote(general = "Robust t² is the squared HC1 t-statistic for draft eligibility; it is a descriptive strength diagnostic, not a formal weak-IV-robust test. Boundary share uses 0.05 and 0.95.")
```

<div class="result-box">
**Design conclusion.** The raw 13.9 percentage-point uptake difference falls
to about nine percentage points after conditioning on age, but the robust
squared t-statistic remains between 26 and 30. Estimated instrument
propensities range from approximately 0.14 to 0.63, with no near-boundary
observations. The saturated-age specification most directly represents the
discrete cohort structure and is the preferred flexible specification; linear
and cubic age are sensitivity checks.
</div>

<div class="warn-box">
**Qualification.** Sample first-stage differences are not positive in every
age cell (ages 29 and 39 are negative). These noisy cell contrasts neither
establish nor refute individual-level monotonicity. The cubic raw-power basis
is also ill-conditioned; its estimates are retained because they are stable,
but the saturated specification is numerically cleaner.
</div>

---

# 1. Reproduction — SUW Table 2

<div class="note-box">
**Layout.** One treatment (veteran status), three age specifications, two outcome units (cents / dollars). Columns are spec × unit. Normalized estimators (τ_cb_u, τ_ml_u, τ_ml_a10) are translation invariant — cents and dollars match; unnormalized ones (τ_ml_a, τ_ml_t, τ_ml_a0) do not.
</div>

```{r build-columns, cache=TRUE}
# One full column = 2SLS + all six kappa estimators with analytical SEs.
run_viet_column <- function(Y, X_kappa, X_df) {
  list(
    tsls  = run_2sls(Y, D, Z, X_df, endog_name = "nvstat"),
    kappa = kappa_analytic_se_all(Y, Z, D, X_kappa)
  )
}

col_lin_cnt <- run_viet_column(Y_cnt, X_lin, X_lin_df)
col_lin_dol <- run_viet_column(Y_dol, X_lin, X_lin_df)
col_cub_cnt <- run_viet_column(Y_cnt, X_cub, X_cub_df)
col_cub_dol <- run_viet_column(Y_dol, X_cub, X_cub_df)
col_sat_cnt <- run_viet_column(Y_cnt, X_sat, X_sat_df)
col_sat_dol <- run_viet_column(Y_dol, X_sat, X_sat_df)
```

```{r table-builder, echo=FALSE}
cell_fmt <- function(est, se, digits = 3) {
  if (is.na(est)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0)
    return(sprintf("%.3f<br><span style='color:#888'>(NA)</span>", est))
  sprintf("%.3f<br><span style='color:#888'>(%.3f)</span>", est, se)
}

build_spec_table <- function(c_lin_cnt, c_lin_dol, c_cub_cnt, c_cub_dol, c_sat_cnt, c_sat_dol) {
  ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
  labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")

  rows <- list(data.frame(
    Estimator = "2SLS",
    `Linear / cents`      = cell_fmt(c_lin_cnt$tsls$coef, c_lin_cnt$tsls$se),
    `Linear / dollars`    = cell_fmt(c_lin_dol$tsls$coef, c_lin_dol$tsls$se),
    `Cubic / cents`       = cell_fmt(c_cub_cnt$tsls$coef, c_cub_cnt$tsls$se),
    `Cubic / dollars`     = cell_fmt(c_cub_dol$tsls$coef, c_cub_dol$tsls$se),
    `Saturated / cents`   = cell_fmt(c_sat_cnt$tsls$coef, c_sat_cnt$tsls$se),
    `Saturated / dollars` = cell_fmt(c_sat_dol$tsls$coef, c_sat_dol$tsls$se),
    check.names = FALSE, stringsAsFactors = FALSE
  ))

  for (i in seq_along(ests)) {
    e <- ests[i]
    rows[[length(rows) + 1]] <- data.frame(
      Estimator = labels[i],
      `Linear / cents`      = cell_fmt(c_lin_cnt$kappa$estimates[e], c_lin_cnt$kappa$se[e]),
      `Linear / dollars`    = cell_fmt(c_lin_dol$kappa$estimates[e], c_lin_dol$kappa$se[e]),
      `Cubic / cents`       = cell_fmt(c_cub_cnt$kappa$estimates[e], c_cub_cnt$kappa$se[e]),
      `Cubic / dollars`     = cell_fmt(c_cub_dol$kappa$estimates[e], c_cub_dol$kappa$se[e]),
      `Saturated / cents`   = cell_fmt(c_sat_cnt$kappa$estimates[e], c_sat_cnt$kappa$se[e]),
      `Saturated / dollars` = cell_fmt(c_sat_dol$kappa$estimates[e], c_sat_dol$kappa$se[e]),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  }
  do.call(rbind, rows)
}
```

## Results

```{r table-reproduction}
build_spec_table(col_lin_cnt, col_lin_dol, col_cub_cnt, col_cub_dol, col_sat_cnt, col_sat_dol) %>%
  kbl(escape = FALSE, align = c("l","c","c","c","c","c","c"),
      caption = "SUW Table 2 — Treatment: Vietnam veteran status (nvstat). Effect of military service on log wages.") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 12.5) %>%
  add_header_above(c(" " = 1, "Linear age" = 2, "Cubic age" = 2, "Saturated age" = 2)) %>%
  pack_rows("Panel A: 2SLS (HC1-robust)",            1, 1) %>%
  pack_rows("Panel B: Normalized kappa estimators",  2, 4) %>%
  pack_rows("Panel C: Unnormalized kappa estimators",5, 7) %>%
  row_spec(2:4, background = "#eef6ff") %>%
  row_spec(5:7, background = "#fffbf0") %>%
  footnote(general = "Each cell reports the estimate with its analytical or HC1-robust standard error below. Blue = normalized (translation invariant); orange = unnormalized. age = age_5, following the SUW replication.")
```

<div class="result-box">
**Replication conclusion.** The normalized kappa estimators and 2SLS are
unchanged when log wages are shifted from dollars to cents. The unnormalized
estimators change under linear and cubic age controls. Their equality under
saturated age is a sample age-cell calibration result, not a general
invariance property.
</div>

---

# 2. Double ML — PLR-IV & Wald-AIPW (honest forest, tuned vs. untuned)

<div class="note-box">
**Setting.** DML uses `dml_with_smoother()` with five-fold cross-fitting on
the **cubic-age** and **saturated-age** specifications. For each specification,
the same call jointly returns PLR-IV and Wald-AIPW. We estimate an untuned
honest forest (`tune.parameters = "none"`) and a tuned honest forest
(`tune.parameters = "all"`). The latter invokes GRF's automatic,
out-of-bag-loss-based tuning; it is distinct from the learner comparison in
Section 3. The linear-age specification remains in the kappa/2SLS reproduction
but is too low-dimensional to provide a meaningful honest-forest comparison.
`set.seed(42)` is reset before every fit so that the four calls begin from the
same random-number state, although tuning itself consumes additional random
draws.
</div>

```{r dml-fit-all, cache=TRUE, results='hide'}
# Cubic-age specification.
set.seed(42); dml_cub_untuned <- dml_with_smoother(
  Y_dol, D, X_dml_cub, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_cub_tuned <- dml_with_smoother(
  Y_dol, D, X_dml_cub, Z, n_cf_folds = 5, tune.parameters = "all")

# Saturated-age specification.
set.seed(42); dml_sat_untuned <- dml_with_smoother(
  Y_dol, D, X_dml_sat, Z, n_cf_folds = 5, tune.parameters = "none")
set.seed(42); dml_sat_tuned <- dml_with_smoother(
  Y_dol, D, X_dml_sat, Z, n_cf_folds = 5, tune.parameters = "all")

res_cub_untuned <- summary(dml_cub_untuned)
res_cub_tuned   <- summary(dml_cub_tuned)
res_sat_untuned <- summary(dml_sat_untuned)
res_sat_tuned   <- summary(dml_sat_tuned)

extract_w <- function(dml_obj) {
  om <- get_outcome_weights(dml_obj)
  check_omega_rows(om)
  list(plriv    = as.vector(om$omega["PLR-IV",    ]),
       waldaipw = as.vector(om$omega["Wald-AIPW", ]))
}
w_cub_untuned <- extract_w(dml_cub_untuned)
w_cub_tuned   <- extract_w(dml_cub_tuned)
w_sat_untuned <- extract_w(dml_sat_untuned)
w_sat_tuned   <- extract_w(dml_sat_tuned)
```

## Estimates {.tabset}

```{r dml-cell-display, echo=FALSE}
show_dml_cell <- function(res, wlist, Y, cap) {
  ids <- c(
    check_weight_identity(wlist$plriv,    Y, get_estimate(res, "PLR-IV")),
    check_weight_identity(wlist$waldaipw, Y, get_estimate(res, "Wald-AIPW"))
  )
  df <- as.data.frame(res[c("PLR-IV", "Wald-AIPW"), , drop = FALSE])
  df$Estimator  <- rownames(df)
  df$`ω′Y reproduces τ̂` <- as.character(ids)
  df <- df[, c("Estimator", setdiff(names(df), c("Estimator","ω′Y reproduces τ̂")),
               "ω′Y reproduces τ̂")]
  df %>%
    kbl(caption = cap, digits = 4, row.names = FALSE,
        align = c("l", rep("r", ncol(df) - 2), "c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 13) %>%
    footnote(general = "Numerical reproduction follows Knaus's all.equal() check with R's default tolerance.")
}
```

### Cubic age

```{r dml-cub-untuned}
show_dml_cell(res_cub_untuned, w_cub_untuned, Y_dol,
              "Cubic age — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-cub-tuned}
show_dml_cell(res_cub_tuned, w_cub_tuned, Y_dol,
              "Cubic age — tuned honest forest (tune.parameters = \"all\")")
```

### Saturated age

```{r dml-sat-untuned}
show_dml_cell(res_sat_untuned, w_sat_untuned, Y_dol,
              "Saturated age — untuned honest forest (tune.parameters = \"none\")")
```

```{r dml-sat-tuned}
show_dml_cell(res_sat_tuned, w_sat_tuned, Y_dol,
              "Saturated age — tuned honest forest (tune.parameters = \"all\")")
```

<div class="note-box">
**How to read the comparison.** Within each age specification, compare the
untuned and tuned rows for changes in the point estimate, standard error, and
later weight geometry. The `ω′Y reproduces τ̂` column checks the package's
algebraic outcome-weight representation; it is not a tuning-quality or
normalization criterion.
</div>

## DML vs. Kappa — Point Estimates

```{r dml-summary-table}
# Kappa estimates (dollars) for both specs — reused later by the learner table.
kappa_cub <- kappa_analytic_se_all(Y_dol, Z, D, X_cub)
kappa_sat <- kappa_analytic_se_all(Y_dol, Z, D, X_sat)

cell <- function(coef, se) {
  if (is.na(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  sprintf("%.3f (%.3f)", coef, se)
}

dml_cell <- function(res, est) {
  coef_val <- get_estimate(res, est)
  se_val   <- if ("SE" %in% colnames(res)) as.numeric(res[est, "SE"])
              else if ("Std. Error" %in% colnames(res)) as.numeric(res[est, "Std. Error"])
              else NA_real_
  cell(coef_val, se_val)
}

dml_rows <- rbind(
  data.frame(Estimator = "PLR-IV (untuned forest)",
             `Cubic age`     = dml_cell(res_cub_untuned, "PLR-IV"),
             `Saturated age` = dml_cell(res_sat_untuned, "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (untuned forest)",
             `Cubic age`     = dml_cell(res_cub_untuned, "Wald-AIPW"),
             `Saturated age` = dml_cell(res_sat_untuned, "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "PLR-IV (tuned forest)",
             `Cubic age`     = dml_cell(res_cub_tuned, "PLR-IV"),
             `Saturated age` = dml_cell(res_sat_tuned, "PLR-IV"),
             check.names = FALSE, stringsAsFactors = FALSE),
  data.frame(Estimator = "Wald-AIPW (tuned forest)",
             `Cubic age`     = dml_cell(res_cub_tuned, "Wald-AIPW"),
             `Saturated age` = dml_cell(res_sat_tuned, "Wald-AIPW"),
             check.names = FALSE, stringsAsFactors = FALSE)
)

ests   <- c("tau_cb_u","tau_ml_u","tau_ml_a10","tau_ml_a","tau_ml_t","tau_ml_a0")
labels <- c("τ_cb_u","τ_ml_u","τ_ml_a10","τ_ml_a","τ_ml_t","τ_ml_a0")
kappa_rows <- do.call(rbind, lapply(seq_along(ests), function(i) {
  e <- ests[i]
  data.frame(
    Estimator = labels[i],
    `Cubic age`     = cell(kappa_cub$estimates[e], kappa_cub$se[e]),
    `Saturated age` = cell(kappa_sat$estimates[e], kappa_sat$se[e]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}))

rbind(dml_rows, kappa_rows) %>%
  kbl(align = c("l","c","c"),
      caption = "DML vs. Kappa — Point Estimates (SE), Y = log wages (dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("Double ML — honest forest", 1, 4) %>%
  pack_rows("Kappa — normalized", 5, 7) %>%
  pack_rows("Kappa — unnormalized", 8, 10) %>%
  row_spec(1:4, background = "#eef6ff") %>%
  row_spec(8:10, background = "#fffbf0") %>%
  footnote(general = "Each cell reports estimate (SE). Blue = DML; orange = unnormalized kappa. Honest-forest tuning changes neither the substantive conclusion nor the large sampling uncertainty.")
```

<div class="result-box">
**DML conclusion.** Across cubic and saturated age controls, PLR-IV and
Wald-AIPW remain close to 2SLS and the normalized kappa estimates. Tuning the
honest forest has little effect relative to the standard errors. The evidence
therefore supports estimator stability in this low-dimensional design, not a
precisely estimated positive earnings effect.
</div>

---

# 3. Learner Comparison — PLR-IV and Wald-AIPW

<div class="note-box">
**Purpose.** Section 2 varies GRF's internal tuning while holding the honest-
forest learner family fixed. This section instead holds the five-fold
DoubleML procedure fixed and varies the nuisance learners: parametric
linear/logistic models, Ranger, and XGBoost. Each learner family is used for
both estimators. PLR-IV is implemented with `DoubleMLPLIV`; Wald-AIPW/LATE is
implemented with `DoubleMLIIVM`. All models use the same cubic-age and
saturated-age designs defined above. Linear and Ranger are fitted in this
notebook. XGBoost is imported from `vietnam_xgb_tuning_export.rds` and
`vietnam_xgb_pliv_tuning_export.rds`, which preserve both the fixed untuned
baseline and the `tune_on_folds = TRUE` estimator without rerunning expensive
nested tuning here.

**Terminology.** “Linear learner” refers to the nuisance-model class: OLS for
PLR-IV and OLS/logit for Wald-AIPW. It does **not** refer to the linear-age
feature specification in Section 1. The learner comparison uses either the
cubic-age feature matrix or the saturated-age feature matrix throughout.
</div>

```{r ext-setup}
make_dml_data <- function(X_dml) {
  df <- as.data.frame(X_dml)
  df$y <- as.numeric(Y_dol)
  df$d <- as.numeric(D)
  df$z <- as.numeric(Z)
  DoubleMLData$new(data = df, y_col = "y", d_cols = "d", z_cols = "z",
                   x_cols = colnames(X_dml))
}

dat_cub <- make_dml_data(X_dml_cub)
dat_sat <- make_dml_data(X_dml_sat)
```

```{r ext-fit-helpers, echo=FALSE}
ml_lm_iivm <- function() list(g = lrn("regr.lm"),
                              m = lrn("classif.log_reg"),
                              r = lrn("classif.log_reg"))
ml_rf_iivm <- function() list(g = lrn("regr.ranger",    keep.inbag = TRUE),
                              m = lrn("classif.ranger", keep.inbag = TRUE),
                              r = lrn("classif.ranger", keep.inbag = TRUE))
ml_lm_pliv <- function() list(l = lrn("regr.lm"),
                              m = lrn("regr.lm"),
                              r = lrn("regr.lm"))
ml_rf_pliv <- function() list(l = lrn("regr.ranger", keep.inbag = TRUE),
                              m = lrn("regr.ranger", keep.inbag = TRUE),
                              r = lrn("regr.ranger", keep.inbag = TRUE))
omega_vec <- function(om, estimator) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w)) && estimator %in% rownames(w))
    return(as.vector(w[estimator, ]))
  as.vector(w[1, ])
}

# Fit one IIVM model. Wrapped in tryCatch: the parametric OLS/logit learner
# can fail to produce outcome weights in the saturated spec (documented in the
# thesis); on failure we return NA gracefully instead of aborting the knit.
fit_iivm <- function(dml_data, learners, Y) {
  tryCatch({
    set.seed(42)
    obj <- DoubleMLIIVM$new(data = dml_data,
                            ml_g = learners$g, ml_m = learners$m, ml_r = learners$r,
                            n_folds = 5, score = "LATE")
    invisible(capture.output(
      obj$fit(store_models = TRUE, store_predictions = TRUE)
    ))
    om  <- get_outcome_weights(object = obj, dml_data = dml_data)
    w   <- omega_vec(om, "Wald-AIPW")
    ok  <- check_doubleml_identity(om, Y, obj$coef)
    list(obj = obj, w = w, ok = ok)
  }, error = function(e) {
    message("fit_iivm failed: ", conditionMessage(e))
    list(obj = NULL, w = rep(NA_real_, length(Y)), ok = NA)
  })
}

# PLR-IV uses three regression nuisances: E[Y|X], E[Z|X], and E[D|X].
fit_pliv <- function(dml_data, learners, Y) {
  tryCatch({
    set.seed(42)
    obj <- DoubleMLPLIV$new(data = dml_data,
                            ml_l = learners$l, ml_m = learners$m, ml_r = learners$r,
                            n_folds = 5)
    invisible(capture.output(
      obj$fit(store_models = TRUE, store_predictions = TRUE)
    ))
    om <- get_outcome_weights(object = obj, dml_data = dml_data)
    w  <- omega_vec(om, "PLR-IV")
    ok <- check_weight_identity(w, Y, get_dml_coef(obj))
    list(obj = obj, w = w, ok = ok)
  }, error = function(e) {
    message("fit_pliv failed: ", conditionMessage(e))
    list(obj = NULL, w = rep(NA_real_, length(Y)), ok = NA)
  })
}

safe_coef <- function(f) {
  if (is.null(f$obj)) return(NA_real_)
  out <- NA_real_
  invisible(capture.output(out <- get_dml_coef(f$obj)))
  as.numeric(out)
}
safe_se <- function(f) {
  if (is.null(f$obj)) return(NA_real_)
  out <- NA_real_
  invisible(capture.output(out <- get_dml_se(f$obj)))
  as.numeric(out)
}

# Compact two-estimator, two-specification table for one learner family.
learner_pair_table <- function(pliv_cub, pliv_sat, iivm_cub, iivm_sat, cap) {
  data.frame(
    Estimator = c("PLR-IV", "Wald-AIPW / LATE"),
    `Cubic estimate` = round(c(safe_coef(pliv_cub), safe_coef(iivm_cub)), 4),
    `Cubic SE`       = round(c(safe_se(pliv_cub),   safe_se(iivm_cub)),   4),
    `Cubic ω′Y reproduces τ̂` = c(pliv_cub$ok, iivm_cub$ok),
    `Saturated estimate` = round(c(safe_coef(pliv_sat), safe_coef(iivm_sat)), 4),
    `Saturated SE`       = round(c(safe_se(pliv_sat),   safe_se(iivm_sat)),   4),
    `Saturated ω′Y reproduces τ̂` = c(pliv_sat$ok, iivm_sat$ok),
    check.names = FALSE
  ) %>%
    kbl(caption = cap, align = c("l","r","r","c","r","r","c")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    add_header_above(c(" " = 1, "Cubic age" = 3, "Saturated age" = 3)) %>%
    footnote(general = "PLR-IV uses DoubleMLPLIV; Wald-AIPW/LATE uses DoubleMLIIVM. Numerical reproduction is checked exactly as in Knaus's notebook, using all.equal() with R's default tolerance.")
}
```

## Estimates {.tabset}

### OLS / Logistic nuisance learners

```{r ext-fit-lm, cache=TRUE, cache.extra='quiet-output-v2'}
lm_iivm_cub <- fit_iivm(dat_cub, ml_lm_iivm(), Y_dol)
lm_iivm_sat <- fit_iivm(dat_sat, ml_lm_iivm(), Y_dol)
lm_pliv_cub <- fit_pliv(dat_cub, ml_lm_pliv(), Y_dol)
lm_pliv_sat <- fit_pliv(dat_sat, ml_lm_pliv(), Y_dol)

learner_pair_table(lm_pliv_cub, lm_pliv_sat, lm_iivm_cub, lm_iivm_sat,
                   "Parametric learners — Linear PLR-IV and Linear/Logistic Wald-AIPW")
```

### Ranger

```{r ext-fit-rf, cache=TRUE, cache.extra='quiet-output-v2'}
rf_iivm_cub <- fit_iivm(dat_cub, ml_rf_iivm(), Y_dol)
rf_iivm_sat <- fit_iivm(dat_sat, ml_rf_iivm(), Y_dol)
rf_pliv_cub <- fit_pliv(dat_cub, ml_rf_pliv(), Y_dol)
rf_pliv_sat <- fit_pliv(dat_sat, ml_rf_pliv(), Y_dol)

learner_pair_table(rf_pliv_cub, rf_pliv_sat, rf_iivm_cub, rf_iivm_sat,
                   "Ranger learners — PLR-IV and Wald-AIPW/LATE")
```

### Imported XGBoost fits

```{r ext-xgb-wald-import, include=FALSE}
validate_xgb_export <- function(x,expected_class) {
  required <- c("estimate_comparison","nuisance_losses","weight_diagnostics",
                "omega_vectors","identity_checks","settings","sample_info")
  if (!all(required %in% names(x))) stop("XGBoost export has an incomplete schema.")
  if (!identical(x$settings$dml_class,expected_class)) stop("XGBoost export has the wrong DoubleML class.")
  if (!isTRUE(x$settings$tune_on_folds)) stop("XGBoost export is not tune_on_folds=TRUE.")
  if (!all(c("Cubic age","Saturated age") %in% names(x$omega_vectors)))
    stop("XGBoost export does not contain both Vietnam age specifications.")
  if (!identical(as.integer(x$sample_info$n), as.integer(length(Z))) ||
      !isTRUE(all.equal(as.numeric(x$sample_info$mean_Z), mean(Z), tolerance = 1e-12)) ||
      !isTRUE(all.equal(as.numeric(x$sample_info$mean_D), mean(D), tolerance = 1e-12)))
    stop("XGBoost export sample metadata do not match the current analysis data.")
  invisible(TRUE)
}
xgb_wald_path <- "vietnam_xgb_tuning_export.rds"
xgb_wald_available <- file.exists(xgb_wald_path)
if (!xgb_wald_available) stop("Missing vietnam_xgb_tuning_export.rds.")
xgb_wald_export <- readRDS(xgb_wald_path)
validate_xgb_export(xgb_wald_export,"DoubleMLIIVM")
```

```{r ext-xgb-pliv-import, include=FALSE}
xgb_pliv_path <- "vietnam_xgb_pliv_tuning_export.rds"
xgb_pliv_available <- file.exists(xgb_pliv_path)
if (!xgb_pliv_available) stop("Missing vietnam_xgb_pliv_tuning_export.rds.")
xgb_pliv_export <- readRDS(xgb_pliv_path)
validate_xgb_export(xgb_pliv_export,"DoubleMLPLIV")
```

```{r ext-xgb-export-helpers, echo=FALSE}
empty_xgb_bundle <- function() list(w=NULL,ok=NA,estimate=NA_real_,gap=NA_real_)
get_xgb_estimate_value <- function(export,cell,variant) {
  if (is.null(export)) return(NA_real_)
  df <- export$weight_diagnostics
  keep_variant <- if (variant=="untuned") {
    grepl("untuned$",df$Learner)
  } else {
    grepl(" tuned$",df$Learner)
  }
  hit <- df[df$Cell==cell & keep_variant,,drop=FALSE]
  if (!nrow(hit)) NA_real_ else as.numeric(hit$Estimate[1])
}
get_xgb_bundle <- function(export,cell,variant) {
  if (is.null(export)) return(empty_xgb_bundle())
  w <- tryCatch(export$omega_vectors[[cell]][[variant]],error=function(e)NULL)
  estimate <- get_xgb_estimate_value(export,cell,variant)
  gap <- if (is.null(w) || !is.finite(estimate)) NA_real_ else sum(w*Y_dol)-estimate
  ok <- if (is.finite(gap)) check_weight_identity(w,Y_dol,estimate) else NA
  list(w=w,ok=ok,estimate=estimate,gap=gap)
}
get_xgb_estimate_cell <- function(export,cell,variant) {
  if (is.null(export)) return("—")
  df <- export$estimate_comparison
  keep_variant <- if (variant=="untuned") grepl("untuned$",df$Learner) else grepl("tune_on_folds=TRUE",df$Learner,fixed=TRUE)
  hit <- df[df$Cell==cell & keep_variant,,drop=FALSE]
  if (!nrow(hit)) "—" else as.character(hit[["Estimate (SE)"]][1])
}
get_xgb_identity_cell <- function(export,cell,variant) {
  value <- tryCatch(export$identity_checks[[cell]][[variant]], error=function(e) NA)
  if (isTRUE(value)) "Yes" else if (identical(value,FALSE)) "No" else "—"
}

xgb_wald_obj <- if (exists("xgb_wald_export")) xgb_wald_export else NULL
xgb_pliv_obj <- if (exists("xgb_pliv_export")) xgb_pliv_export else NULL

xgb_iivm_cub_untuned <- get_xgb_bundle(xgb_wald_obj,"Cubic age","untuned")
xgb_iivm_cub_tuned   <- get_xgb_bundle(xgb_wald_obj,"Cubic age","tuned")
xgb_iivm_sat_untuned <- get_xgb_bundle(xgb_wald_obj,"Saturated age","untuned")
xgb_iivm_sat_tuned   <- get_xgb_bundle(xgb_wald_obj,"Saturated age","tuned")
xgb_pliv_cub_untuned <- get_xgb_bundle(xgb_pliv_obj,"Cubic age","untuned")
xgb_pliv_cub_tuned   <- get_xgb_bundle(xgb_pliv_obj,"Cubic age","tuned")
xgb_pliv_sat_untuned <- get_xgb_bundle(xgb_pliv_obj,"Saturated age","untuned")
xgb_pliv_sat_tuned   <- get_xgb_bundle(xgb_pliv_obj,"Saturated age","tuned")
```

```{r ext-xgb-comparison, echo=FALSE}
xgb_import_table <- do.call(rbind, lapply(list(
  list("PLR-IV", "Cubic age",     xgb_pliv_obj),
  list("PLR-IV", "Saturated age", xgb_pliv_obj),
  list("Wald-AIPW / LATE", "Cubic age",     xgb_wald_obj),
  list("Wald-AIPW / LATE", "Saturated age", xgb_wald_obj)
), function(x) {
  data.frame(
    Estimator=x[[1]],
    Specification=x[[2]],
    Untuned=get_xgb_estimate_cell(x[[3]],x[[2]],"untuned"),
    Tuned=get_xgb_estimate_cell(x[[3]],x[[2]],"tuned"),
    `ω′Y reproduces τ̂ (untuned / tuned)`=paste(
      get_xgb_identity_cell(x[[3]],x[[2]],"untuned"),
      get_xgb_identity_cell(x[[3]],x[[2]],"tuned"),
      sep=" / "),
    check.names=FALSE
  )
}))

xgb_import_table %>%
  kbl(caption="Imported XGBoost estimates — Untuned and fold-specific tuning",
      align=c("l","l","c","c","c")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=TRUE,font_size=13) %>%
  footnote(general="Each estimate cell reports estimate (SE). Numerical reproduction is checked as in Knaus's notebook, using all.equal() with R's default tolerance; it is separate from the normalization conditions.")
```

<div class="note-box">
**Interpretation.** Knaus's normalization results require the relevant smoother
conditions. In particular, boosted trees need not satisfy the affine-smoother
condition, so XGBoost implementations do not generally inherit the associated
normalization guarantees. Separately, none of the extracted XGBoost vectors in
the present implementation passes Knaus's default `all.equal()` criterion.
The discrepancies are nevertheless very small (at most about 1.2e-7 in
absolute value). The reconstructed weights are therefore reported together
with their numerical reproduction gaps rather than being discarded.
</div>

## Comparison Table

```{r ext-comparison}
ext_cell <- function(coef, se) {
  if (is.na(coef) || !is.finite(coef)) return("—")
  if (is.na(se) || !is.finite(se) || se <= 0) return(sprintf("%.3f (NA)", coef))
  sprintf("%.3f (%.3f)", coef, se)
}

learner_row <- function(label, f_cub, f_sat) {
  data.frame(
    Estimator = label,
    `Cubic age`     = ext_cell(safe_coef(f_cub), safe_se(f_cub)),
    `Saturated age` = ext_cell(safe_coef(f_sat), safe_se(f_sat)),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

kappa_ref_row <- function(label, key) {
  data.frame(
    Estimator = label,
    `Cubic age`     = ext_cell(kappa_cub$estimates[key], kappa_cub$se[key]),
    `Saturated age` = ext_cell(kappa_sat$estimates[key], kappa_sat$se[key]),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

comparison_rows <- list(
  learner_row("PLR-IV (OLS nuisances)",                  lm_pliv_cub,  lm_pliv_sat),
  learner_row("Wald-AIPW / LATE (OLS/logit nuisances)", lm_iivm_cub, lm_iivm_sat),
  learner_row("PLR-IV (ranger)",                  rf_pliv_cub,  rf_pliv_sat),
  learner_row("Wald-AIPW / LATE (ranger)",       rf_iivm_cub, rf_iivm_sat),
  kappa_ref_row("τ_cb_u (kappa, ref)",  "tau_cb_u"),
  kappa_ref_row("τ_ml_u (kappa, ref)",  "tau_ml_u")
)
xgb_rows <- list(
  data.frame(Estimator="PLR-IV (XGBoost, untuned)",
    `Cubic age`=get_xgb_estimate_cell(xgb_pliv_obj,"Cubic age","untuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_pliv_obj,"Saturated age","untuned"),check.names=FALSE),
  data.frame(Estimator="PLR-IV (XGBoost, tune_on_folds=TRUE)",
    `Cubic age`=get_xgb_estimate_cell(xgb_pliv_obj,"Cubic age","tuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_pliv_obj,"Saturated age","tuned"),check.names=FALSE),
  data.frame(Estimator="Wald-AIPW / LATE (XGBoost, untuned)",
    `Cubic age`=get_xgb_estimate_cell(xgb_wald_obj,"Cubic age","untuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_wald_obj,"Saturated age","untuned"),check.names=FALSE),
  data.frame(Estimator="Wald-AIPW / LATE (XGBoost, tune_on_folds=TRUE)",
    `Cubic age`=get_xgb_estimate_cell(xgb_wald_obj,"Cubic age","tuned"),
    `Saturated age`=get_xgb_estimate_cell(xgb_wald_obj,"Saturated age","tuned"),check.names=FALSE)
)
comp_df <- do.call(rbind,c(comparison_rows[1:4],xgb_rows,comparison_rows[5:6]))

comp_df %>%
  kbl(align = c("l","c","c"),
      caption = "PLR-IV and Wald-AIPW Learner Comparison — DoubleML vs. Kappa (Y = log wages, dollars)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE, font_size = 13) %>%
  pack_rows("DoubleML learners", 1, 8) %>%
  pack_rows("Kappa reference",   9, 10) %>%
  row_spec(1:8, background = "#eef6ff") %>%
  footnote(general = "Each cell reports estimate (SE). Linear denotes the nuisance learner, not the linear-age feature specification. XGBoost rows import the stored nested-tuning fits; their reconstructed outcome weights are assessed separately below.")
```

<div class="result-box">
**Learner conclusion.** Parametric, Ranger, honest-forest, and XGBoost point
estimates are similar relative to their uncertainty. This agreement indicates
limited learner sensitivity in the Vietnam design. The XGBoost smoother
reconstructions miss Knaus's default numerical-equivalence threshold by small
amounts; Section 4 reports both their weight diagnostics and the corresponding
reproduction gaps.
</div>

---

# 4. Weight Diagnostics

<div class="note-box">
**Reading the tables.** `Sum_w ≈ 0` checks translation invariance and
`Mass T/C ≈ 1 / 1` checks treated/control normalization. Following
Chattopadhyay and Zubizarreta (2023, Section 5.4), modified ESS is reported
separately for treated and control observations. The overall modified ESS is a
supplementary concentration summary. `ESS mod/N T/C` divides each arm-specific
value by that arm's observed size, making concentration comparable across
unequal groups. `Opposite-sign mass T/C` measures the
share of absolute mass opposite to the conventional treated/control orientation
and therefore reveals cancellation that ESS alone cannot show; it is not by
itself evidence of an invalid IV estimator. Conventional Kish ESS is not
reported because signed outcome weights are not a nonnegative normalized
weighted sample.
The first pair of tabs compares untuned and tuned honest forests. The second
pair compares PLR-IV and Wald-AIPW across learner families. For every available
weight vector, the tables report whether `ω′Y` reproduces `τ̂` under the same
default `all.equal()` rule used by Knaus, together with the absolute numerical
gap. This reproduction check is separate from the displayed normalization and
concentration statistics.
</div>

```{r diag-builders, echo=FALSE}
# kappa weight bundles for both specs.
kw_cub <- kappa_weights_bundle(Z, D, X_cub)
kw_sat <- kappa_weights_bundle(Z, D, X_sat)

DIAG_ALIGN <- c("l","r","c","r","c","c","c","r")

dml_diag_cell <- function(w_untuned, w_tuned, kw, D) {
  rbind(
    outcome_weight_diag(w_untuned$plriv,    D, "PLR-IV (untuned forest)"),
    outcome_weight_diag(w_untuned$waldaipw, D, "Wald-AIPW (untuned forest)"),
    outcome_weight_diag(w_tuned$plriv,      D, "PLR-IV (tuned forest)"),
    outcome_weight_diag(w_tuned$waldaipw,   D, "Wald-AIPW (tuned forest)"),
    outcome_weight_diag(kw$tau_cb_u,    D, "τ_cb_u     (kappa)"),
    outcome_weight_diag(kw$tau_ml_u,    D, "τ_ml_u     (kappa)"),
    outcome_weight_diag(kw$tau_ml_a10,  D, "τ_ml_a10   (kappa)")
  )
}
style_dml_diag <- function(df, cap) {
  df %>%
    kbl(caption = cap, align = DIAG_ALIGN) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(1:4, background = "#eef6ff") %>%
    footnote(general = "Blue = honest-forest DML. T/C entries report treated / sign-oriented control values. Arm-specific modified ESS is the headline concentration diagnostic; overall modified ESS is supplementary.")
}

exact_w <- function(fit) {
  if (isTRUE(fit$ok) && !is.null(fit$w)) fit$w else NULL
}
safe_outcome_diag <- function(w, D, label) {
  if (is.null(w) || !length(w) || all(is.na(w))) NULL else outcome_weight_diag(w, D, label)
}

identity_diag_row <- function(fit, D, Y, label) {
  if (!is.null(fit$w) && length(fit$w) && !all(is.na(fit$w))) {
    estimate <- if (!is.null(fit$estimate)) fit$estimate else safe_coef(fit)
    gap <- sum(fit$w*Y)-estimate
    out <- outcome_weight_diag(fit$w, D, label)
    out$`ω′Y reproduces τ̂` <- if (check_weight_identity(fit$w,Y,estimate)) "Yes" else "No"
    out$`|ω′Y−τ̂|` <- formatC(abs(gap),format="e",digits=2)
  } else {
    out <- data.frame(
      Estimator=label, `ω′Y reproduces τ̂`=NA_character_, `|ω′Y−τ̂|`=NA_character_,
      Sum_w=NA_real_, `Mass T/C`=NA_character_,
      `ESS mod overall`=NA_real_, `ESS mod T/C`=NA_character_,
      `ESS mod/N T/C`=NA_character_,
      `Opposite-sign mass T/C (%)`=NA_character_, `Max |w|`=NA_real_,
      check.names=FALSE
    )
  }
  out[, c("Estimator","ω′Y reproduces τ̂","|ω′Y−τ̂|",
          setdiff(names(out),c("Estimator","ω′Y reproduces τ̂","|ω′Y−τ̂|")))]
}
kappa_diag_row <- function(w, D, label) {
  out <- outcome_weight_diag(w, D, label)
  out$`ω′Y reproduces τ̂` <- "By construction"
  out$`|ω′Y−τ̂|` <- "0.00e+00"
  out[, c("Estimator","ω′Y reproduces τ̂","|ω′Y−τ̂|",
          setdiff(names(out),c("Estimator","ω′Y reproduces τ̂","|ω′Y−τ̂|")))]
}

ext_diag_block <- function(D, Y, kw, pliv_lm, iivm_lm, pliv_rf, iivm_rf,
                           pliv_xgb_t, iivm_xgb_t) {
  rbind(
    identity_diag_row(pliv_lm,   D, Y, "PLR-IV (OLS nuisances)"),
    identity_diag_row(iivm_lm,   D, Y, "Wald-AIPW (OLS/logit nuisances)"),
    identity_diag_row(pliv_rf,   D, Y, "PLR-IV (ranger)"),
    identity_diag_row(iivm_rf,   D, Y, "Wald-AIPW (ranger)"),
    identity_diag_row(pliv_xgb_t,D, Y, "PLR-IV (XGBoost, tuned)"),
    identity_diag_row(iivm_xgb_t,D, Y, "Wald-AIPW (XGBoost, tuned)"),
    kappa_diag_row(kw$tau_cb_u,   D, "τ_cb_u    (kappa, CBPS)"),
    kappa_diag_row(kw$tau_ml_u,   D, "τ_ml_u    (kappa, MLE)"),
    kappa_diag_row(kw$tau_ml_a10, D, "τ_ml_a10  (kappa, MLE)")
  )
}
style_ext_diag <- function(df, cap) {
  n_learn <- sum(grepl("^(PLR-IV|Wald-AIPW)", df$Estimator))
  df %>%
    kbl(caption = cap, align = c("l","c","r","r","c","r","c","c","c","r")) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 13) %>%
    row_spec(seq_len(n_learn), background = "#eef6ff") %>%
    footnote(general = "Blue = DoubleML learners. Numerical reproduction follows Knaus's all.equal() check with R's default tolerance. The absolute gap shows why a narrowly failed check need not be substantively important; normalization and concentration diagnostics remain distinct properties.")
}
```

## DML vs. Kappa {.tabset}

### Cubic age

```{r dml-diag-cub}
style_dml_diag(dml_diag_cell(w_cub_untuned, w_cub_tuned, kw_cub, D),
               "Weight Diagnostics — Untuned/Tuned Honest Forest vs. Kappa | Cubic age")
```

### Saturated age

```{r dml-diag-sat}
style_dml_diag(dml_diag_cell(w_sat_untuned, w_sat_tuned, kw_sat, D),
               "Weight Diagnostics — Untuned/Tuned Honest Forest vs. Kappa | Saturated age")
```

## Learners vs. Kappa {.tabset}

### Cubic age

```{r ext-diag-cub}
style_ext_diag(ext_diag_block(D, Y_dol, kw_cub,
                              lm_pliv_cub, lm_iivm_cub,
                              rf_pliv_cub, rf_iivm_cub,
                              xgb_pliv_cub_tuned, xgb_iivm_cub_tuned),
               "Weight Diagnostics — PLR-IV and Wald-AIPW Learners | Cubic age")
```

### Saturated age

```{r ext-diag-sat}
style_ext_diag(ext_diag_block(D, Y_dol, kw_sat,
                              lm_pliv_sat, lm_iivm_sat,
                              rf_pliv_sat, rf_iivm_sat,
                              xgb_pliv_sat_tuned, xgb_iivm_sat_tuned),
               "Weight Diagnostics — PLR-IV and Wald-AIPW Learners | Saturated age")
```

<div class="result-box">
**Weight conclusion.** Exact normalized representations have weight sums near
zero, as required for translation invariance. Under cubic age, the
unnormalized kappa weights retain nonzero total mass and reproduce the
cents–dollars sensitivity. Under saturated age, sample age-cell calibration
makes all kappa variants coincide. Modified ESS is reported by treatment arm;
ordinary Kish ESS is deliberately omitted because these are signed outcome
weights rather than nonnegative sampling weights.
</div>

## Detailed Outcome-Weight Statistics {.tabset}

<div class="note-box">
**What this is.** These detailed tables report Overall / Treated / Control
statistics using Knaus's sign convention. `Neg. share` matches his
observation-count summary and is distinct from the oriented opposite-sign
absolute-mass measure in the headline tables. All six kappa variants are
included here. To avoid repeating the tuning comparison, only the tuned honest-
forest and XGBoost specifications are retained. Modified ESS remains in the
preceding diagnostic tables and is not repeated.
</div>

```{r bridge-helpers, echo=FALSE}
# Named list of weight vectors per cell; drop any all-NA (failed) learner.
cell_weight_list <- function(w_tuned,
                             pliv_lm, iivm_lm, pliv_rf, iivm_rf,
                             pliv_xgb_t, iivm_xgb_t, kw) {
  lst <- list(
    "PLR-IV (tuned forest)"      = w_tuned$plriv,
    "Wald-AIPW (tuned forest)"   = w_tuned$waldaipw,
    "PLR-IV (OLS nuisances)"            = exact_w(pliv_lm),
    "Wald-AIPW (OLS/logit nuisances)"   = exact_w(iivm_lm),
    "PLR-IV (ranger)"            = exact_w(pliv_rf),
    "Wald-AIPW (ranger)"         = exact_w(iivm_rf),
    "PLR-IV (XGBoost, tuned; reconstructed)" = pliv_xgb_t$w,
    "Wald-AIPW (XGBoost, tuned; reconstructed)" = iivm_xgb_t$w,
    "τ_cb_u (kappa)"   = kw$tau_cb_u,
    "τ_ml_u (kappa)"   = kw$tau_ml_u,
    "τ_ml_a10 (kappa)" = kw$tau_ml_a10,
    "τ_ml_a (kappa, unnormalized)"  = kw$tau_ml_a,
    "τ_ml_t (kappa, untreated-unnormalized)" = kw$tau_ml_t,
    "τ_ml_a0 (kappa, treated-unnormalized)" = kw$tau_ml_a0
  )
  Filter(function(w) !is.null(w) && length(w) && !all(is.na(w)), lst)
}

bridge_cells <- list(
  "Cubic age" = list(w = cell_weight_list(
    w_cub_tuned,
    lm_pliv_cub, lm_iivm_cub, rf_pliv_cub, rf_iivm_cub,
    xgb_pliv_cub_tuned, xgb_iivm_cub_tuned, kw_cub)),
  "Saturated age" = list(w = cell_weight_list(
    w_sat_tuned,
    lm_pliv_sat, lm_iivm_sat, rf_pliv_sat, rf_iivm_sat,
    xgb_pliv_sat_tuned, xgb_iivm_sat_tuned, kw_sat))
)

bridge_table <- function(wlist, cap) {
  ests <- names(wlist)
  df   <- do.call(rbind, lapply(ests, function(nm) weight_bridge(wlist[[nm]], D, nm)))
  disp <- df[, c("Scope","Min","Max","Neg_share","Sum_top10","Sum_w",
                 "Sum_abs_w","Max_abs_w")]

  kk <- disp %>%
    kbl(caption = cap, row.names = FALSE,
        col.names = c("Scope","Min","Max","Neg. share","ΣTop10%","Σω","Σ|ω|",
                      "Max |ω|"),
        align = c("l", rep("r", 7))) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = TRUE, font_size = 12) %>%
    add_header_above(c(" " = 1, "Knaus summary() statistics" = 6,
                       "Concentration" = 1)) %>%
    column_spec(8, bold = TRUE, background = "#eef6ff")
  for (i in seq_along(ests)) kk <- pack_rows(kk, ests[i], (i - 1) * 3 + 1, i * 3)
  kk %>% footnote(general = "Neg. share is Knaus's observation-count statistic in [0,1]. Control rows use the sign-flipped −ω convention. XGBoost rows use the reconstructed tuned weight vectors; their small numerical reproduction gaps are reported in the preceding learner diagnostics. Modified ESS is not repeated here.")
}
```

### Cubic age

```{r bridge-cub}
bridge_table(bridge_cells[["Cubic age"]]$w,
             "Detailed outcome-weight statistics — Cubic age")
```

### Saturated age

```{r bridge-sat}
bridge_table(bridge_cells[["Saturated age"]]$w,
             "Detailed outcome-weight statistics — Saturated age")
```

---

# 5. Love Plots

<div class="note-box">
**Balance scale.** Each panel reports absolute mean differences only (no KS
statistic). Continuous covariates are standardized by the pooled unadjusted
standard deviation. Starred binary covariates are left on their raw 0/1 scale.
Because the adjusted comparison uses sign-oriented outcome weights, a raw binary
difference is a signed weighted-moment difference and need not compare two
literal proportions when opposite-oriented weights remain.
</div>

```{r love-helpers, echo=FALSE}
# Return NULL when a fit failed or its outcome weights are not an exact
# representation. Approximate XGBoost weights remain visible in Section 3 but
# do not enter the exact Love-plot comparison.
maybe_love <- function(title, w, D, X, ok = TRUE) {
  if (!isTRUE(ok) || all(is.na(w))) NULL else make_love(title, w, D, X)
}
```

## DML vs. Kappa {.tabset}

### Cubic age (untuned/tuned forests and kappa)

```{r love-cub, fig.height=9, fig.width=13}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nforest untuned",    w_cub_untuned$plriv,    D, X_dml_cub),
  maybe_love("Wald-AIPW\nforest untuned", w_cub_untuned$waldaipw, D, X_dml_cub),
  maybe_love("PLR-IV\nforest tuned",      w_cub_tuned$plriv,      D, X_dml_cub),
  maybe_love("Wald-AIPW\nforest tuned",   w_cub_tuned$waldaipw,   D, X_dml_cub),
  maybe_love("τ_cb_u (CBPS)",    kw_cub$tau_cb_u,   D, X_dml_cub),
  maybe_love("τ_ml_u",           kw_cub$tau_ml_u,   D, X_dml_cub),
  maybe_love("τ_ml_a10",         kw_cub$tau_ml_a10, D, X_dml_cub)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — Cubic age: untuned/tuned honest forest vs. kappa")
```

### Saturated age (untuned/tuned forests and kappa)

```{r love-sat, fig.height=9, fig.width=13}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nforest untuned",    w_sat_untuned$plriv,    D, X_dml_sat),
  maybe_love("Wald-AIPW\nforest untuned", w_sat_untuned$waldaipw, D, X_dml_sat),
  maybe_love("PLR-IV\nforest tuned",      w_sat_tuned$plriv,      D, X_dml_sat),
  maybe_love("Wald-AIPW\nforest tuned",   w_sat_tuned$waldaipw,   D, X_dml_sat),
  maybe_love("τ_cb_u (CBPS)",    kw_sat$tau_cb_u,   D, X_dml_sat),
  maybe_love("τ_ml_u",           kw_sat$tau_ml_u,   D, X_dml_sat),
  maybe_love("τ_ml_a10",         kw_sat$tau_ml_a10, D, X_dml_sat)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — Saturated age: untuned/tuned honest forest vs. kappa")
```

## Learners vs. Kappa {.tabset}

### Cubic age (PLR-IV and Wald-AIPW by learner)

```{r ext-love-cub, fig.height=12, fig.width=14}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nOLS nuisances", lm_pliv_cub$w, D, X_dml_cub, lm_pliv_cub$ok),
  maybe_love("Wald-AIPW\nOLS/logit nuisances", lm_iivm_cub$w, D, X_dml_cub, lm_iivm_cub$ok),
  maybe_love("PLR-IV\nranger",       rf_pliv_cub$w,   D, X_dml_cub, rf_pliv_cub$ok),
  maybe_love("Wald-AIPW\nranger",    rf_iivm_cub$w,   D, X_dml_cub, rf_iivm_cub$ok),
  maybe_love("PLR-IV\nXGBoost untuned", xgb_pliv_cub_untuned$w, D, X_dml_cub, xgb_pliv_cub_untuned$ok),
  maybe_love("PLR-IV\nXGBoost tuned", xgb_pliv_cub_tuned$w, D, X_dml_cub, xgb_pliv_cub_tuned$ok),
  maybe_love("Wald-AIPW\nXGBoost untuned", xgb_iivm_cub_untuned$w, D, X_dml_cub, xgb_iivm_cub_untuned$ok),
  maybe_love("Wald-AIPW\nXGBoost tuned", xgb_iivm_cub_tuned$w, D, X_dml_cub, xgb_iivm_cub_tuned$ok),
  maybe_love("τ_cb_u\nkappa",        kw_cub$tau_cb_u, D, X_dml_cub),
  maybe_love("τ_ml_u\nkappa",        kw_cub$tau_ml_u, D, X_dml_cub)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — PLR-IV and Wald-AIPW learners vs. kappa | Cubic age")
```

### Saturated age (PLR-IV and Wald-AIPW by learner)

```{r ext-love-sat, fig.height=12, fig.width=14}
grobs <- Filter(Negate(is.null), list(
  maybe_love("PLR-IV\nOLS nuisances", lm_pliv_sat$w, D, X_dml_sat, lm_pliv_sat$ok),
  maybe_love("Wald-AIPW\nOLS/logit nuisances", lm_iivm_sat$w, D, X_dml_sat, lm_iivm_sat$ok),
  maybe_love("PLR-IV\nranger",       rf_pliv_sat$w,   D, X_dml_sat, rf_pliv_sat$ok),
  maybe_love("Wald-AIPW\nranger",    rf_iivm_sat$w,   D, X_dml_sat, rf_iivm_sat$ok),
  maybe_love("PLR-IV\nXGBoost untuned", xgb_pliv_sat_untuned$w, D, X_dml_sat, xgb_pliv_sat_untuned$ok),
  maybe_love("PLR-IV\nXGBoost tuned", xgb_pliv_sat_tuned$w, D, X_dml_sat, xgb_pliv_sat_tuned$ok),
  maybe_love("Wald-AIPW\nXGBoost untuned", xgb_iivm_sat_untuned$w, D, X_dml_sat, xgb_iivm_sat_untuned$ok),
  maybe_love("Wald-AIPW\nXGBoost tuned", xgb_iivm_sat_tuned$w, D, X_dml_sat, xgb_iivm_sat_tuned$ok),
  maybe_love("τ_cb_u\nkappa",        kw_sat$tau_cb_u, D, X_dml_sat),
  maybe_love("τ_ml_u\nkappa",        kw_sat$tau_ml_u, D, X_dml_sat)
))
grid.arrange(grobs = grobs, nrow = 2,
             top = "Love Plots — PLR-IV and Wald-AIPW learners vs. kappa | Saturated age")
```

<div class="result-box">
**Balance conclusion.** The exact outcome-weight representations substantially
reduce age imbalance between the treated and control contrasts. Balance is a
diagnostic of the fitted weighted comparison; it does not test the exclusion
restriction, conditional independence, or monotonicity. XGBoost is absent
because its extracted vectors do not exactly reproduce the reported estimates.
</div>
