---
title: "Angrist (1990) — XGBoost Nuisance Tuning for Wald-AIPW"
subtitle: "Vietnam draft lottery: nested tuning inside each outer cross-fitting fold"
author: "Alexander Unger"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    code_folding: hide
    df_print: kable
---

```{css, echo=FALSE}
body { font-size: 15px; color: #2c2c2c; }
h1 { border-bottom: 2px solid #2c7bb6; padding-bottom: 6px; margin-top: 36px; }
h2 { color: #2c7bb6; margin-top: 28px; }
.note-box { background:#eaf4ff; border-left:4px solid #2c7bb6; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
.warn-box { background:#fff3e0; border-left:4px solid #fb8c00; padding:10px 16px; border-radius:4px; margin:14px 0; font-size:14px; }
table { font-size: 13px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE,
                      fig.width=11, fig.height=5.5)
```

```{r packages}
library(haven)
library(knitr)
library(kableExtra)
library(dplyr)
library(OutcomeWeights)
library(DoubleML)
library(mlr3verse)
library(mlr3tuning)

source("functions_all.R")
lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")

SEED        <- 42
N_CF        <- 5
TUNE_EVALS  <- 15
INNER_FOLDS <- 3
EXTRACT_WEIGHTS <- TRUE
RERUN_TUNING <- FALSE
EXPORT_PATH <- "vietnam_xgb_tuning_export.rds"
```

<div class="note-box">
**Purpose.** This notebook estimates Wald-AIPW/LATE with `DoubleMLIIVM`
using XGBoost nuisances for the cubic-age and saturated-age specifications.
For each specification it retains a fixed untuned baseline and a model tuned
with `tune_on_folds = TRUE`. The latter selects hyperparameters separately
inside each outer training fold, so held-out observations cannot influence
the tuning choice used to predict them. The resulting estimates, losses,
fold-level parameter summaries, outcome weights, and numerical reproduction checks are
exported for `vietnam_presentation_4.Rmd`; the main notebook does not retune.
By default, `RERUN_TUNING = FALSE` renders the report from the verified
export. It should be changed to `TRUE` only when the nested tuning procedure
itself is deliberately re-estimated.
</div>

<div class="warn-box">
**Runtime.** Wald-AIPW fits instrument, treatment, and arm-specific outcome
nuisances. With five outer folds, three inner folds, and fifteen random-search
evaluations, nested tuning is intentionally expensive. `EXTRACT_WEIGHTS` is
kept `TRUE` because the main analysis and Method-A notebook require the exact
exported vectors.
</div>

# Setup

```{r data}
stopifnot(file.exists("sipp.dta"))
sipp <- read_dta("sipp.dta")
sipp <- sipp[sipp$rsncode != 999 & !is.na(sipp$educ) & !is.na(sipp$kwage), ]

sipp$age     <- sipp$age_5
sipp$age_fac <- factor(sipp$age_5)
sipp$lwage   <- log(sipp$kwage)

Y <- as.numeric(sipp$lwage)
D <- as.integer(sipp$nvstat)
Z <- as.integer(sipp$rsncode)

X_cub      <- make_X(~ 0 + age + I(age^2) + I(age^3), data=sipp)
X_sat_full <- make_X(~ 0 + age_fac, data=sipp)
X_sat      <- X_sat_full[, -1, drop=FALSE]

data.frame(
  Quantity=c("Analysis sample","Cubic-age covariates",
             "Saturated-age covariates"),
  Value=c(format(nrow(sipp),big.mark=","),ncol(X_cub),ncol(X_sat))
) %>%
  kbl(caption="Analysis dimensions") %>%
  kable_styling(bootstrap_options=c("striped","condensed"),full_width=FALSE)
```

```{r learners-and-tuning}
make_dml_data <- function(Y, D, Z, X) {
  df <- as.data.frame(X)
  df$y <- as.numeric(Y); df$d <- as.numeric(D); df$z <- as.numeric(Z)
  DoubleMLData$new(data=df, y_col="y", d_cols="d", z_cols="z",
                   x_cols=colnames(X))
}

make_xgb_learners <- function() {
  list(
    g=lrn("regr.xgboost", alpha=0, subsample=1, max_delta_step=0, base_score=0),
    m=lrn("classif.xgboost", predict_type="prob"),
    r=lrn("classif.xgboost", predict_type="prob")
  )
}

xgb_ps <- function() {
  ps(nrounds=p_int(50,400), max_depth=p_int(2,5),
     eta=p_dbl(0.02,0.20), min_child_weight=p_int(1,8))
}

tune_param_set <- function() list(ml_g=xgb_ps(), ml_m=xgb_ps(), ml_r=xgb_ps())
tune_settings <- function() list(
  n_folds_tune=INNER_FOLDS,
  rsmp_tune=rsmp("cv", folds=INNER_FOLDS),
  measure=list(ml_g="regr.rmse", ml_m="classif.logloss", ml_r="classif.logloss"),
  terminator=trm("evals", n_evals=TUNE_EVALS),
  algorithm=tnr("random_search")
)

get_prediction_vector <- function(obj, learner_name) {
  pred <- tryCatch(obj$predictions[[learner_name]], error=function(e) NULL)
  if (is.null(pred)) return(numeric(0))
  as.numeric(drop(pred))
}

get_branch_prediction <- function(obj, base_name, Z) {
  p0 <- get_prediction_vector(obj, paste0(base_name,"0"))
  p1 <- get_prediction_vector(obj, paste0(base_name,"1"))
  if (length(p0) != length(Z) || length(p1) != length(Z)) return(numeric(0))
  ifelse(as.numeric(Z)==1, p1, p0)
}

rmse_loss <- function(truth, pred) {
  if (length(pred) != length(truth) || !length(pred)) return(NA_real_)
  keep <- is.finite(truth) & is.finite(pred)
  if (!any(keep)) return(NA_real_)
  sqrt(mean((as.numeric(truth[keep])-as.numeric(pred[keep]))^2))
}

logloss_loss <- function(truth, pred, eps=1e-6) {
  if (length(pred) != length(truth) || !length(pred)) return(NA_real_)
  keep <- is.finite(truth) & is.finite(pred)
  if (!any(keep)) return(NA_real_)
  y <- as.numeric(truth[keep]); p <- pmin(pmax(as.numeric(pred[keep]),eps),1-eps)
  -mean(y*log(p)+(1-y)*log(1-p))
}

propensity_boundary_share <- function(obj, Z, threshold=0.02) {
  m_hat <- get_prediction_vector(obj,"ml_m")
  r_hat <- get_branch_prediction(obj,"ml_r",Z)
  p_hat <- c(m_hat,r_hat); p_hat <- p_hat[is.finite(p_hat)]
  if (!length(p_hat)) return(NA_real_)
  mean(p_hat < threshold | p_hat > 1-threshold)
}

calc_nuisance_losses <- function(r, cell) {
  if (is.null(r)) return(c(outcome_rmse=NA, instrument_logloss=NA, treatment_logloss=NA))
  c(outcome_rmse=rmse_loss(cell$Y,get_branch_prediction(r$obj,"ml_g",cell$Z)),
    instrument_logloss=logloss_loss(cell$Z,get_prediction_vector(r$obj,"ml_m")),
    treatment_logloss=logloss_loss(cell$D,get_branch_prediction(r$obj,"ml_r",cell$Z)))
}

fit_cell <- function(dml_data, learners, label, Z, tune=FALSE) {
  tryCatch({
    set.seed(SEED)
    obj <- DoubleMLIIVM$new(data=dml_data, ml_g=learners$g, ml_m=learners$m,
                            ml_r=learners$r, n_folds=N_CF, score="LATE")
    if (tune) obj$tune(param_set=tune_param_set(), tune_settings=tune_settings(),
                       tune_on_folds=TRUE)
    obj$fit(store_models=TRUE, store_predictions=TRUE)
    list(label=label, tuned=tune, coef=get_dml_coef(obj), se=get_dml_se(obj),
         boundary_share=propensity_boundary_share(obj,Z),
         tuning_result=tryCatch(obj$tuning_result,error=function(e) NULL),
         params=tryCatch(obj$params,error=function(e) NULL), obj=obj)
  }, error=function(e) {
    message(sprintf("Fit failed for %s: %s",label,conditionMessage(e))); NULL
  })
}

cells <- list(
  "Cubic age"=list(Y=Y,D=D,Z=Z,X=X_cub),
  "Saturated age"=list(Y=Y,D=D,Z=Z,X=X_sat)
)
```

# Fit: untuned versus tuned XGBoost

```{r fit-all, include=FALSE}
if (RERUN_TUNING) {
  results <- list()
  for (cell_name in names(cells)) {
    cell <- cells[[cell_name]]
    dat <- make_dml_data(cell$Y,cell$D,cell$Z,cell$X)
    results[[cell_name]] <- list(
      untuned=fit_cell(dat,make_xgb_learners(),
                       paste(cell_name,"xgb-untuned"),cell$Z,FALSE),
      tuned=fit_cell(dat,make_xgb_learners(),
                     paste(cell_name,"xgb-tuned"),cell$Z,TRUE)
    )
  }
} else {
  if (!file.exists(EXPORT_PATH)) {
    stop("Verified XGBoost export not found. Set RERUN_TUNING=TRUE to rebuild it.")
  }
  stored_export <- readRDS(EXPORT_PATH)
}
```

## 1. Estimate comparison

```{r comparison-table}
fmt_est <- function(r) if (is.null(r)) "—" else sprintf("%.3f (%.3f)",r$coef,r$se)
fmt_boundary <- function(r) if (is.null(r)||!is.finite(r$boundary_share)) "—" else sprintf("%.4f",r$boundary_share)
if (RERUN_TUNING) {
  rows <- do.call(rbind,lapply(names(results),function(cn) {
    R <- results[[cn]]
    data.frame(Cell=cn,Learner=c("XGB untuned","XGB tuned (tune_on_folds=TRUE)"),
               `Estimate (SE)`=c(fmt_est(R$untuned),fmt_est(R$tuned)),
               `Propensity boundary share`=c(fmt_boundary(R$untuned),fmt_boundary(R$tuned)),
               check.names=FALSE)
  }))
} else {
  rows <- stored_export$estimate_comparison
}
rows %>% kbl(caption="Vietnam Wald-AIPW: XGBoost untuned vs. fold-specifically tuned",
             align=c("l","l","c","r")) %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
  collapse_rows(columns=1,valign="middle") %>%
  row_spec(which(rows$Learner=="XGB tuned (tune_on_folds=TRUE)"),background="#eef6ff") %>%
  footnote(general="Boundary share is the fraction of fitted instrument or treatment propensities below 0.02 or above 0.98. Tuning is selected by nuisance prediction loss within each outer training fold.")
```

## 2. Nuisance-prediction losses

```{r nuisance-loss-table}
if (RERUN_TUNING) {
  loss_rows <- do.call(rbind,lapply(names(results),function(cn) {
    R <- results[[cn]]; cell <- cells[[cn]]
    u <- calc_nuisance_losses(R$untuned,cell); t <- calc_nuisance_losses(R$tuned,cell)
    data.frame(Cell=cn,Learner=c("XGB untuned","XGB tuned"),
               `Outcome RMSE`=c(u["outcome_rmse"],t["outcome_rmse"]),
               `Instrument log-loss`=c(u["instrument_logloss"],t["instrument_logloss"]),
               `Treatment log-loss`=c(u["treatment_logloss"],t["treatment_logloss"]),
               check.names=FALSE,row.names=NULL)
  }))
} else {
  loss_rows <- stored_export$nuisance_losses
}
loss_rows %>% kbl(digits=4,caption="Wald-AIPW nuisance losses: untuned vs. tuned (lower is better)") %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
  collapse_rows(columns=1,valign="middle") %>%
  row_spec(which(loss_rows$Learner=="XGB tuned"),background="#eef6ff")
```

## 3. Tuning protocol

The tuned estimator does not have one final hyperparameter vector. The
settings are selected separately within each outer training fold, and the
corresponding fold-specific model produces the held-out predictions. The
relevant reproducibility information is therefore the prespecified search
design rather than a vector of medians assembled across different fitted
models.

```{r tuning-protocol}
data.frame(
  Component=c(
    "Outer cross-fitting","Inner validation","Search algorithm","Budget",
    "nrounds","max_depth","eta","min_child_weight",
    "Outcome nuisance","Instrument nuisance","Treatment nuisance"
  ),
  Specification=c(
    paste(N_CF,"folds"),paste(INNER_FOLDS,"folds"),"Random search",
    paste(TUNE_EVALS,"evaluations per nuisance and outer fold"),
    "50–400","2–5","0.02–0.20","1–8",
    "RMSE","Log-loss","Log-loss"
  )
) %>%
  kbl(caption="Nested XGBoost tuning protocol: Wald-AIPW") %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),
                full_width=FALSE)
```

```{r hyperparameter-summary, include=FALSE}
extract_tuned_params <- function(r,cn) {
  if (is.null(r)||!r$tuned) return(data.frame())
  src <- if (!is.null(r$params)) r$params else r$tuning_result
  if (is.null(src)) return(data.frame())
  u <- unlist(src,recursive=TRUE)
  u <- u[grepl("nrounds|max_depth|eta|min_child_weight",names(u))]
  if (!length(u)) return(data.frame())
  hp <- data.frame(Cell=cn,Raw_name=names(u),Value=as.numeric(u))
  hp$Nuisance <- ifelse(grepl("ml_g",hp$Raw_name),"Outcome g",
                 ifelse(grepl("ml_m",hp$Raw_name),"Instrument m",
                 ifelse(grepl("ml_r",hp$Raw_name),"Treatment r","Unknown")))
  hp$Branch <- ifelse(grepl("ml_g0|ml_r0",hp$Raw_name),"Z=0",
               ifelse(grepl("ml_g1|ml_r1",hp$Raw_name),"Z=1",""))
  hp$Parameter <- sub(".*(nrounds|max_depth|eta|min_child_weight).*","\\1",hp$Raw_name)
  hp
}
if (RERUN_TUNING) {
  hp_parts <- lapply(names(results),function(cn) extract_tuned_params(results[[cn]]$tuned,cn))
  hp_parts <- Filter(function(x)nrow(x)>0,hp_parts)
  hp_raw <- if (length(hp_parts)) do.call(rbind,hp_parts) else data.frame()
  if (!nrow(hp_raw)) stop("Fold-specific hyperparameters could not be extracted.")
  hp_summary <- hp_raw %>% filter(Nuisance!="Unknown") %>%
    group_by(Cell,Nuisance,Branch,Parameter) %>%
    summarise(Median_across_folds=median(Value),N_values_seen=n(),.groups="drop")
} else {
  hp_raw <- NULL
  hp_summary <- stored_export$hyperparameters
}
```

<div class="note-box">
The existing export retains the median fold summary for audit purposes.
Future substantive reruns additionally save the underlying fold-specific
values. Neither object is a reusable global XGBoost recipe.
</div>

## 4. Outcome-weight diagnostics and export

```{r weight-helpers}
omega_vec <- function(om) {
  w <- om$omega
  if (is.null(dim(w))) return(as.vector(w))
  if (!is.null(rownames(w))&&"Wald-AIPW"%in%rownames(w)) return(as.vector(w["Wald-AIPW",]))
  as.vector(w[1,])
}
extract_xgb_w <- function(r,cell) {
  if (is.null(r)||is.null(r$obj)) stop("fitted DoubleML object is NULL")
  om <- get_outcome_weights(object=r$obj,dml_data=make_dml_data(cell$Y,cell$D,cell$Z,cell$X))
  omega_vec(om)
}
safe_extract_xgb_w <- function(r,cell) tryCatch(extract_xgb_w(r,cell),error=function(e)NULL)

one_diag_row <- function(cn,label,r,cell) {
  w <- safe_extract_xgb_w(r,cell)
  if (is.null(r)||is.null(w)) return(data.frame(Cell=cn,Learner=label,Estimate=if(is.null(r))NA else r$coef,
    ESS_mod=NA,`Sigma omega`=NA,`Sigma abs omega`=NA,`Treated mass`=NA,
    `Control mass (-sum control weights)`=NA,`max abs omega`=NA,`ω′Y reproduces τ̂`=NA,
    `Weight status`=if(is.null(r))"fit failed" else "weight extraction failed",check.names=FALSE))
  wb <- weight_bridge(w,cell$D,label); ov <- wb[wb$Scope=="Overall",]; tr <- wb[wb$Scope=="Treated",]; co <- wb[wb$Scope=="Control",]
  data.frame(Cell=cn,Learner=label,Estimate=r$coef,ESS_mod=ov$ESS_mod,
    `Sigma omega`=ov$Sum_w,`Sigma abs omega`=ov$Sum_abs_w,`Treated mass`=tr$Sum_w,
    `Control mass (-sum control weights)`=co$Sum_w,`max abs omega`=ov$Max_abs_w,
    `ω′Y reproduces τ̂`=tryCatch(check_weight_identity(w,cell$Y,r$coef),error=function(e)NA),
    `Weight status`="ok",check.names=FALSE)
}
```

```{r weight-table}
if (RERUN_TUNING) {
  xgb_weight_table <- do.call(rbind,lapply(names(cells),function(cn) {
    R <- results[[cn]]; cell <- cells[[cn]]
    rbind(one_diag_row(cn,"XGB untuned",R$untuned,cell),
          one_diag_row(cn,"XGB tuned",R$tuned,cell))
  }))
} else {
  xgb_weight_table <- stored_export$weight_diagnostics
}
# Older cached exports contain conventional Kish ESS. It is not displayed
# because signed contrast weights do not define a nonnegative weighted sample.
xgb_weight_table <- xgb_weight_table[, setdiff(names(xgb_weight_table), "ESS"), drop=FALSE]
if ("Identity ok" %in% names(xgb_weight_table))
  names(xgb_weight_table)[names(xgb_weight_table)=="Identity ok"] <- "ω′Y reproduces τ̂"
xgb_weight_table %>% kbl(digits=4,caption="Vietnam Wald-AIPW XGBoost outcome-weight diagnostics") %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"),full_width=TRUE) %>%
  row_spec(which(xgb_weight_table$Learner=="XGB tuned"),background="#eef6ff") %>%
  footnote(general="Modified ESS is reported for these signed outcome weights; conventional Kish ESS is omitted. Numerical reproduction follows Knaus's all.equal() check with R's default tolerance and is separate from normalization.")
```

```{r omega-and-identities}
if (RERUN_TUNING) {
  xgb_omega_vectors <- lapply(names(cells),function(cn) list(
    untuned=if(EXTRACT_WEIGHTS)safe_extract_xgb_w(results[[cn]]$untuned,cells[[cn]])else NULL,
    tuned=if(EXTRACT_WEIGHTS)safe_extract_xgb_w(results[[cn]]$tuned,cells[[cn]])else NULL))
  names(xgb_omega_vectors) <- names(cells)
  xgb_identity_checks <- lapply(names(cells),function(cn) {
    cell <- cells[[cn]]; R <- results[[cn]]; W <- xgb_omega_vectors[[cn]]
    list(
      untuned=tryCatch(if(is.null(W$untuned)||is.null(R$untuned))NA else
        check_weight_identity(W$untuned,cell$Y,R$untuned$coef),error=function(e)NA),
      tuned=tryCatch(if(is.null(W$tuned)||is.null(R$tuned))NA else
        check_weight_identity(W$tuned,cell$Y,R$tuned$coef),error=function(e)NA))
  })
  names(xgb_identity_checks) <- names(cells)
} else {
  xgb_omega_vectors <- stored_export$omega_vectors
  xgb_identity_checks <- stored_export$identity_checks
}
```

```{r export, include=FALSE}
if (RERUN_TUNING) {
  xgb_export <- list(
    estimate_comparison=rows,nuisance_losses=loss_rows,
    hyperparameters=hp_summary,hyperparameters_raw=hp_raw,
    weight_diagnostics=xgb_weight_table,omega_vectors=xgb_omega_vectors,
    identity_checks=xgb_identity_checks,
    settings=list(seed=SEED,n_cf=N_CF,tune_evals=TUNE_EVALS,inner_folds=INNER_FOLDS,
      tune_on_folds=TRUE,learner="xgboost",dml_class="DoubleMLIIVM",score="LATE",
      extract_weights=EXTRACT_WEIGHTS,cells=names(cells)),
    sample_info=list(n=nrow(sipp),mean_Z=mean(Z),mean_D=mean(D),outcome="log(kwage)",
                     specifications=names(cells)))
  saveRDS(xgb_export,EXPORT_PATH)
}
```

<div class="note-box">
The main analysis reads `vietnam_xgb_tuning_export.rds`. With
`tune_on_folds = TRUE`, the exported estimate and weight vectors come from
fold-specific fitted models. The median summaries are descriptive and must
not be copied into a new fit.
</div>
